---
title: "Massachusetts Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Need to install mdbtools -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  snakecase, # change string case
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # text analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  batman, # rep(NA, 8) Batman!
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http query
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Massachusetts Office of Campaign and Political Finance (OCPF)][03].

> #### The Agency  
> The Office of Campaign and Political Finance is an independent state agency that administers
Massachusetts General Law Chapter 55, the campaign finance law, and Chapter 55C, the limited public
financing program for statewide candidates. Established in 1973, OCPF is the depository for
disclosure reports filed by candidates and political committees under M.G.L. Chapter 55.
> 
> Specifically, candidates who report to OCPF are those seeking statewide, legislative, county and
district office, Governor's Council candidates and two groups of municipal candidates: Candidates
for mayor, city council or alderman in the state's 14 cities with populations of 65,000 or more...
Candidates for mayor in cities with populations of less than 65,000
> 
> OCPF receives reports filed by hundreds of candidates and committees, reviews them to ensure
accurate disclosure and legal compliance, and, where appropriate, conducts legal reviews of
campaign finance activity.

> #### Our Mission  
> The fundamental purpose of the Massachusetts campaign finance law is to assist in maintaining the
integrity of the Commonwealth's electoral system. OCPF's primary mission is to ensure that accurate
and complete disclosure of campaign finance activity by those involved in the electoral process is
available in a transparent, easily accessible and timely manner and that stakeholders in the
process fully understand and comply with the statute. Stakeholders must have full confidence in the
integrity of OCPF's procedures in transmittal and disclosure of activity. OCPF is committed to
providing easily accessed resources, both in the form of disclosure and education, to all
participants seeking to influence the outcome of political campaigns. OCPF is also committed to
analyzing developments in campaign finance regulation and reform at the federal level and in other
jurisdictions, so that OCPF can suggest legislative amendments to strengthen Chapters 55 and 55C.

[03]: https://www.ocpf.us/Home/Index

## Import

We will obtain raw immutable data and import it into R as a data frame.

```{r raw_dir}
raw_dir <- here("ma", "expends", "data", "raw")
dir_create(raw_dir)
```

### Download

Data can be obtained from the OCPF in one of two ways: (1) Up to 250,000 search results can be
downloaded in text format from the [OCPF search page][04]; (2) A single [`.zip` file][05] 
containing a `.mdb` file can be downloaded from the [OCPF data page][06]. We will use the later.

> Download a zipped Microsoft Access 2000 format (.mdb) database that includes report summaries,
receipts, expenditures, in-kind contributions, liabilities, assets disposed, savings accounts,
credit card reports, reimbursement reports and subvendor reports.

[04]: https://www.ocpf.us/Reports/SearchItems
[05]: http://ocpf2.blob.core.windows.net/downloads/data/campaign-finance-reports.zip
[06]: https://www.ocpf.us/Data

```{r create_paths}
zip_url <- "http://ocpf2.blob.core.windows.net/downloads/data/campaign-finance-reports.zip"
zip_path <- url2path(zip_url, raw_dir)
```

First, check the file size before downloading.

```{r size_size}
url_file_size(zip_url, format = TRUE)
```

Then download the file to the `/data/raw` directory and unzip.

```{r download_mdb}
if (!all_files_new(raw_dir, "*.zip$")) {
  download.file(url = zip_url, destfile = zip_path)
}

if (!all_files_new(raw_dir, "*.mdb$")) {
  unzip(zipfile = zip_path, exdir = raw_dir)
}
```

### Read

To read this file, we will use `campfin::read_mdb()`, which wraps around `readr::read_csv()` and 
the `mdb-export` command from [MDB Tools][07], which must first be installed from GitHub or your
package manager.

```bash
$ sudo apt install mdbtools
```

[07]: https://github.com/brianb/mdbtools

We can use `campfin:::mdb_tables()` to find the table name we are interested in from the database.

```{r get_mdb_tables}
# get file name
mdb_file <- dir_ls(raw_dir, glob = "*.mdb$")
# list tables in file
campfin:::mdb_tables(file = mdb_file)
```

Then, use `campfin::read_mdb()` to read the table as a data frame.

```{r read_expends}
ma <- read_mdb(
  file = mdb_file,
  table = "vUPLOAD_tCURRENT_EXPENDITURES",
  na = c("", "NA", "N/A"),
  locale = locale(tz = "US/Eastern"),
  col_types = cols(
    .default = col_character(),
    Date = col_date(),
    Amount = col_double()
  )
)
```

Finally, we can standardize the data frame structure with the `janitor` package.

```{r clean_expends}
ma <- ma %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  mutate_if(is_character, str_to_upper)
```

The `report_id` variable links to the "vUPLOAD_MASTER" table of the database, which gives more 
information on the _filers_ of the reports, whose expenditures are listed in 
"vUPLOAD_tCURRENT_EXPENDITURES".

```{r read_master}
master <- read_mdb(
  file = mdb_file,
  table = "vUPLOAD_MASTER",
  na = c("", "NA", "N/A", "Unknown/ N/A"),
  col_types = cols(.default = col_character())
)

master <- master %>%
  clean_names("snake") %>% 
  filter(report_id %in% ma$report_id) %>% 
  mutate_all(str_to_upper) %>% 
  select(
    report_id,
    cpf_id,
    report_type = report_type_description,
    cand_name = full_name,
    office,
    district,
    comm_name = report_comm_name,
    comm_city = report_comm_city,
    comm_state = report_comm_state,
    comm_zip = report_comm_zip,
    category
  )
```

Then join these two tables together.

```{r join_master}
ma <- left_join(ma, master)
```

```{r rm_master, echo=FALSE, include=FALSE}
rm(master)
gc()
```

## Explore

```{r glimpse, collapse=FALSE}
head(ma)
tail(ma)
glimpse(sample_frac(ma))
```

### Missing

```{r glimpse_na, collapse=FALSE}
glimpse_fun(ma, count_na)
```

```{r flag_na}
ma <- ma %>% flag_na(date, amount, vendor, cand_name)
sum(ma$na_flag)
mean(ma$na_flag)
```

### Duplicates

```{r flag_dupes, eval=FALSE}
# repeated variable
all(ma$id == ma$line_sequence)
n_distinct(ma$id) == n_distinct(ma$line_sequence)
ma <- select(ma, -line_sequence)
ma <- ma %>% flag_dupes(-id, -line_sequence, -guid)
sum(ma$dupe_flag)
mean(ma$dupe_flag)
```

### Categorical

```{r glimpse_distinct, collapse=FALSE}
glimpse_fun(ma, n_distinct)
```

```{r report_bar, echo=FALSE}
explore_plot(
  data = ma,
  var = report_type,
  flip = TRUE,
  title = "Massachusetts Expenditures by Report Types",
  caption = "Source: MA OCPF"
)
```

```{r office_bar, echo=FALSE}
explore_plot(
  data = drop_na(ma, office),
  var = office,
  flip = TRUE,
  title = "Massachusetts Expenditures by Office Sought",
  caption = "Source: MA OCPF"
)
```

```{r category_bar, echo=FALSE}
explore_plot(
  data = drop_na(ma, category),
  var = category,
  flip = TRUE,
  title = "Massachusetts Expenditures by Category",
  caption = "Source: MA OCPF"
)
```

```{r purpose_bar, echo=FALSE, fig.height=10}
ma %>% 
  drop_na(purpose) %>% 
  unnest_tokens(word, purpose) %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  head(n = 30) %>% 
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_col(aes(fill = n)) +
  coord_flip() +
  scale_fill_continuous(guide = FALSE) +
  labs(
    title = "Massachusetts Expenditure Purpose Word",
    caption = "Source: MA OCPF",
    x = "Word in Purpose",
    y = "Frequency"
  )
```

### Continuous

#### Amounts

```{r summary_amount}
summary(ma$amount) %>% map_chr(dollar)
sum(ma$amount <= 0)
sum(ma$amount >= 1000000)
```

We can view the smallest and largest expenditures to check for range issues.

From this, we can see the minimum `amount` value  from an expenditure with a `purpose` of
`r ma$purpose[ma$amount == min(ma$amount)]`. This isn't really an expenditure in the normal
sense.

```{r glimpse_min}
glimpse(filter(ma, amount == min(amount)))
```

The maximum `amount` of `r dollar(max(ma$amount))` was made by the 
`r ma$comm_name[ma$amount == max(ma$amount)]` on `r ma$date[ma$amount == max(ma$amount)]`. However,
both the `vendor` and `purpose` values for that expenditure are missing. Searching the OCPF
database online does not return this expenditure.

```{r glimpse_max}
glimpse(filter(ma, amount == max(amount)))
```

We can use `ggplot2::geom_histogram()` to ensure a typical log-normal distribution of expenditures.

```{r amount_histogram, echo=FALSE}
ma %>%
  ggplot(aes(amount)) +
  geom_histogram() +
  geom_vline(xintercept = median(ma$amount)) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Massachusetts Expenditure Purpose Word",
    caption = "Source: MA OCPF",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can add a `year` variable from `date` using `lubridate::year()`.

```{r add_year}
ma <- mutate(ma, year = year(date))
```

There are a number of `date` values from the distant past or future.

```{r date_range, collapse=TRUE}
min(ma$date, na.rm = TRUE)
sum(ma$year < 2001, na.rm = TRUE)
max(ma$date, na.rm = TRUE)
sum(ma$date > today(), na.rm = TRUE)
count_na(ma$date)
```

We can flag these dates with a new `date_flag` variable.

```{r flag_date}
ma <- mutate(ma, date_flag = is.na(date) | date > today() | year < 2001)
sum(ma$date_flag, na.rm = TRUE)
```

Using this new flag, we can create a `date_clean` variable that's missing these erronous dates.

```{r clean_dates}
ma <- ma %>% 
  mutate(
    date_clean = as_date(ifelse(date_flag, NA, date)),
    year_clean = year(date_clean)
    )
```

The Massachusetts Governor serves four-year terms, and we can see the number of expenditures spike
every four years. 

```{r year_bar_count, echo=FALSE}
ma %>% 
  count(year_clean) %>% 
  mutate(on = is_even(year_clean)) %>%
  ggplot(aes(x = year_clean, y = n)) +
  geom_col(aes(fill = on)) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  labs(
    title = "Massachusetts Expenditure Count per Year",
    caption = "Source: MA OCPF",
    fill = "Election Year",
    x = "Year Made",
    y = "Number of Expenditures"
  ) +
  theme(legend.position = "bottom")
```

If we look at the _total_ amount spent, we can spot a fairly regular spike in the total cost of
expenditures made. One outlier seems to be 2016, when there was no Governor's race but there was
still `r dollar(sum(ma$amount[ma$year_clean == 2016], na.rm = TRUE))` spent, similar to 2018. 

```{r year_bar_sum, echo=FALSE}
ma %>% 
  mutate(on = is_even(year_clean)) %>%
  group_by(year_clean, on) %>%
  summarize(sum = sum(amount)) %>% 
  ggplot(aes(x = year_clean, y = sum)) +
  geom_col(aes(fill = on)) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Massachusetts Expenditure Amounts per Year",
    caption = "Source: MA OCPF",
    fill = "Election Year",
    x = "Year Made",
    y = "Total Amount Expended"
  ) +
  theme(legend.position = "bottom")
```

```{r month_line, echo=FALSE}
ma %>% 
  mutate(
    month = month(date_clean),
    on = is_even(year_clean)
  ) %>% 
  group_by(on, month) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(month, mean)) +
  geom_line(aes(color = on), size = 2) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  labs(
    title = "Massachusetts Expenditures per Year",
    caption = "Source: MA OCPF",
    fill = "Election Year",
    x = "Year Made",
    y = "Number of Expenditures"
  ) +
  theme(legend.position = "bottom")
```

## Wrangle

We should use the `campfin::normal_*()` functions to perform some basic, high-confidence text
normalization to improve the searchability of the database.

### Address

First, we will normalize the street address by removing punctuation and expanding abbreviations.

```{r normal_address}
ma <- ma %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      add_abbs = usps_street,
      na_rep = TRUE
    )
  )
```

We can see how this improves consistency across the `address` field.

```{r view_address_change, echo=FALSE}
drop_na(select(ma, starts_with("address")))
```

### ZIP

The `zip` address is already fairly clean, with `r percent(prop_in(ma$zip, valid_zip, na.rm = TRUE))`
of the values already in our comprehensive `valid_zip` list.

```{r count_zip_pre, collapse=TRUE}
n_distinct(ma$zip)
prop_in(ma$zip, valid_zip, na.rm = TRUE)
length(setdiff(ma$zip, valid_zip))
```

We can improve this further by lopping off the uncommon four-digit extensions and removing common
invalid codes like 00000 and 99999.

```{r normal_zip}
ma <- ma %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

This brings our valid percentage to `r percent(prop_in(ma$zip_norm, valid_zip, na.rm = TRUE))`.

```{r count_zip_post, collapse=TRUE}
n_distinct(ma$zip_norm)
prop_in(ma$zip_norm, valid_zip, na.rm = TRUE)
length(setdiff(ma$zip_norm, valid_zip))
```

### State

The `state` variable is also very clean, already at 
`r percent(prop_in(ma$state, valid_state, na.rm = TRUE))`.

```{r count_state_pre, collapse=TRUE}
n_distinct(ma$state)
prop_in(ma$state, valid_state, na.rm = TRUE)
length(setdiff(ma$state, valid_state))
setdiff(ma$state, valid_state)
```

There are still `r length(setdiff(ma$state, valid_state))` invalid values which we can remove.

```{r normal_state}
ma <- ma %>% 
  mutate(
    state_norm = normal_state(
      state = str_replace(state, "^M$", "MA"),
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r count_state_post, collapse=TRUE}
n_distinct(ma$state_norm)
prop_in(ma$state_norm, valid_state, na.rm = TRUE)
```

### City

The `city` value is the hardest to normalize. We can use a four-step system to functionally improve
the searchablity of the database.

1. **Normalize** the raw values with `campfin::normal_city()`
1. **Match** the normal values with the _expected_ value for that ZIP code
1. **Swap** the normal values with the expected value if they are _very_ similar
1. **Refine** the swapped values the [OpenRefine algorithms][08] and keep good changes

[08]: https://github.com/OpenRefine/OpenRefine/wiki/Clustering-In-Depth

The raw `city` values are relatively normal, with 
`r percent(prop_in(ma$city, valid_city, na.rm = TRUE))` already in `valid_city` (which is not
comprehensive). We will aim to get this number over 99%.

```{r count_city_pre, collapse=TRUE}
n_distinct(ma$city)
prop_in(ma$city, valid_city, na.rm = TRUE)
length(setdiff(ma$city, valid_city))
prop_na(ma$city)
```

#### Normalize

```{r normal_city}
ma <- ma %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      geo_abbs = usps_city,
      st_abbs = c("MA", "DC", "MASSACHUSETTS"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

This process brought us to `r percent(prop_in(ma$city_norm, valid_city, na.rm = TRUE))` valid.

```{r count_city_post_norm, collapse=TRUE}
n_distinct(ma$city_norm)
prop_in(ma$city_norm, valid_city, na.rm = TRUE)
length(setdiff(ma$city_norm, valid_city))
prop_na(ma$city_norm)
```

It also increased the proportion of `NA` values by 
`r percent(prop_na(ma$city_norm) - prop_na(ma$city))`. These new `NA` values were either a single
(possibly repeating) character, or contained in the `na_city` vector.

```{r new_city_na, echo=FALSE}
ma %>% 
  filter(is.na(city_norm) & !is.na(city)) %>% 
  select(zip_norm, state_norm, city, city_norm) %>% 
  distinct() %>% 
  sample_frac()
```

#### Swap

Then, we will compare these normalized `city_norm` values to the _expected_ city value for that
vendor's ZIP code. If the [levenshtein distance][09] is less than 3, we can confidently swap these
two values.

[09]: https://en.wikipedia.org/wiki/Levenshtein_distance

```{r swap_city}
ma <- ma %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = is_less_than(match_dist, 3),
      true = city_match,
      false = city_norm
    )
  )
```

This is a very fast way to increase the valid proportion to
`r percent(prop_in(ma$city_swap, valid_city, na.rm = TRUE))` and reduce the number of distinct
_invalid_ values from `r length(setdiff(ma$city_norm, valid_city))` to only
`r length(setdiff(ma$city_swap, valid_city))`

```{r count_city_post_swap, collapse=TRUE}
n_distinct(ma$city_swap)
prop_in(ma$city_swap, valid_city, na.rm = TRUE)
length(setdiff(ma$city_swap, valid_city))
```

#### Refine

Finally, we can pass these swapped `city_swap` values to the OpenRefine cluster and merge 
algorithms. These two algorithms cluster similar values and replace infrequent values with their
more common counterparts. This process can be harmful by making _incorrect_ changes. We will only
keep changes where the state, ZIP code, _and_ new city value all match a valid combination.

```{r refine_city}
good_refine <- ma %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r view_city_refines, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_raw, 
    city_refine
  ) %>% 
  sample_frac()
```

We can join these good refined values back to the original data and use them over their incorrect
`city_swap` counterparts in a new `city_refine` variable.

```{r join_refine}
ma <- ma %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

This brings us to `r percent(prop_in(ma$city_refine, valid_city, na.rm = TRUE))` valid values.

```{r count_city_post_refine, collapse=TRUE}
n_distinct(ma$city_refine)
prop_in(ma$city_refine, valid_city, na.rm = TRUE)
length(setdiff(ma$city_refine, valid_city))
```

#### Manual

We can make very few manual changes to capture the last few big invalid values. Local city
abbreviations (BOS, DORC, WORD, CAMB) often need to be changed by hand.

```{r view_final_bad}
ma %>%
  filter(city_refine %out% valid_city) %>% 
  count(state_norm, zip_norm, city_refine, sort = TRUE) %>% 
  drop_na(city_refine) %>% 
  print(n = 20)
```

```{r city_final}
ma <- ma %>% 
  mutate(
    city_manual = city_refine %>% 
      str_replace("\bBOS\b", "BOSTON") %>% 
      str_replace("^DORC$", "DORCHESTER") %>% 
      str_replace("^WORC$", "WORCHESTER") %>% 
      str_replace("^HP$", "HYDE PARK") %>% 
      str_replace("^JP$", "JAMAICA PLAIN") %>% 
      str_replace("^NY$", "NEW YORK") %>% 
      str_replace("^CRLSTRM$", "CAROL STREAM") %>% 
      str_replace("^SPFLD$", "SPRINGFIELD") %>% 
      str_replace("^SPGFLD$", "SPRINGFIELD") %>% 
      str_replace("^PLY$", "PLYMOUTH") %>% 
      str_replace("^CAMB$", "CAMBRIDGE")
  )
```

## Conclude

1. There are `r nrow(ma)` records in the database.
1. There are `sum(ma$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(ma$na_flag)` records missing either recipient or date.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 5-digit `zip_norm` variable has been created with `campfin::normal_zip(ma$zip)`.
1. The 4-digit `year` variable has been created with `lubridate::year(ma$date)`.

## Lookup

```{r lookup}
lookup_file <- here("ma", "expends", "data", "ma_city_lookup.csv")
if (file.exists(lookup_file)) {
  lookup <- read_csv(lookup_file) %>% select(1:2)
  ma <- left_join(ma, lookup, by = c("city_manual" = "city_final"))
  progress_table(
    ma$city_raw,
    ma$city_norm,
    ma$city_swap,
    ma$city_manual,
    ma$city_clean,
    compare = valid_city
  )
}
```

```{r}
progress <- progress_table(
  ma$city_raw,
  ma$city_norm,
  ma$city_swap,
  ma$city_manual,
  ma$city_clean,
  compare = valid_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Massachusetts City Normalization Progress",
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivilent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Massachusetts City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Export

```{r create_proc_dir}
proc_dir <- here("ma", "expends", "data", "processed")
dir_create(proc_dir)
```

```{r write_clean}
ma %>% 
  select(
    -city_norm,
    -city_swap,
    -city_match,
    -city_swap,
    -match_dist
  ) %>% 
  write_csv(
    path = glue("{proc_dir}/ma_expends_clean.csv"),
    na = ""
  )
```

