---
title: "Nebraska Expenditures since 2021"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state","ne", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  textreadr, # read rtf files
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("state/ne/contribs/docs/ne_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained by from the Nebraska Accountability and Disclosure Commission's
(NADC) [data download page][03]. Data prior to 2022 was available in bulk and was processed in a different workflow. (See `ne_expends_diary.Rmd`). This update deals with new data since 2021 and can be accessed on the website by clicking on `Expenditures`.

> The page provides comma separated value (CSV) downloads of contribution, loan, and expenditure data for each reporting year in a zipped file format. These files can be downloaded and imported into other applications (Microsoft Excel, Microsoft Access, etc.)

> This data is extracted from the state of Nebraska database as it existed as of  7/23/2023  2:26 AM

The [NADC home][04] page explains the role and resources provided by the 
Commission.

> This site includes the campaign statements filed with the Commission by
candidates for state and local offices, political action committees or PACs,
political parties, and ballot question committees. Campaign statements disclose
the amount and sources of contributions to the committee and the amount purposes
of expenditures by the committee. The site also includes Reports of
Contributions filed with the Commission by corporations, unions, limited
liability companies, limited partnerships, and by industry, trade or
professional associations. Reports of Contributions disclose contributions and
expenditures made to support or oppose candidates or ballot questions. Forms,
instructional materials and a list of committees may be accessed through this
site. To access the information described in this paragraph, click on the tab
above marked CAMPAIGN FINANCE.

Navigating to the [campaign finance section of the website][05], the exact
filing requirements are laid out.

> Candidates for state and local office are subject to the campaign finance
provisions of the Nebraska Political Accountability and Disclosure Act (NPADA).
Those seeking election to federal office are not subject to the NPADA. Once a
candidate raises, receives or expends $5,000 or more in a calendar year, he or
she is required to register the committee with the Commission by filing a
Statement of Organization and paying a filing fee of $100. Once the candidate
committee is formed, the committee must file periodic campaign statements.

specifically, there are **three** kinds of filings that must be made by general
candidates:

1. > Statement of Organization - This document registers the Committee with the
Commission. It discloses the name of the committee, the name of the candidate,
the office sought and the name of the treasurer.

2. > Campaign Statement - this document provides a summary of the financial
activity of the committee. It discloses the name of contributors of more than
\$250 and the amounts of the contributions. It discloses expenditures of more
than \$250.

3. > Report of Late Expenditures - this document discloses contributions of
$1,000 or more received by the committee during 14 days immediately preceding
the election. Information reported as late contributions are also reported on
subsequent campaign statements.

Aside from general candidates, the NADC also explains that political committees
must also file.

> There are **two** types of committees that are commonly referred to as
political action committees or PACs. One type of PAC is a political committee
sponsored by a corporation, union or other type of organization. Another type of
PAC is a political committee which is not sponsored by another group. PACs are
required to register with the Commission and file periodic campaign statements
disclosing receipts and expenditures.

1. > State and county political party committees with certain levels of
financial activity, register with the Commission and file periodic campaign
statements.

2. > A ballot question committee is a political committee formed for the purpose
of supporting or opposing the qualification, passage or defeat of a ballot
question. The committee must register with the Commission within 10 days after
raising, receiving or expending $5,000 or more in a calendar year. Once
registered, ballot questions committees file periodic campaign statements
disclosing receipts and expenditures.

Finally, the NADC identifies the source of the data:

> The campaign filings displayed on this website are based upon the paper
records filed with the Nebraska Accountability and Disclosure Commission. While
every effort has been made to ensure accurate data input, errors may occur.
Anyone noting an error is encouraged to contact this office.

In any discrepancy between the information displayed on this website and the
official records on file with the Commission, the official records of the
Commission take precedence.

[01]: https://github.com/irworkshop/accountability_datacleaning
[03]: https://nadc.nebraska.gov/view-campaign-filings-personal-financial-disclosures-potential-conflicts-lobbying-reports-and-more
[04]: http://www.nadc.nebraska.gov/index.html
[05]: http://www.nadc.nebraska.gov/cf/index.html
[06]: http://www.nadc.nebraska.gov/docs/B-1-2018.doc
[07]: http://www.nadc.nebraska.gov/docs/B-2-2018.doc
[08]: http://www.nadc.nebraska.gov/docs/B-4-2018.doc


```{r raw_dir}
raw_urls <- glue("https://nadc-e.nebraska.gov/PublicSite/Docs/BulkDataDownloads/{2021:2023}_ExpenditureExtract.csv.zip")
raw_dir <- dir_create(here("state","ne", "expends", "data", "raw"))
raw_zips <- path(raw_dir, basename(raw_urls))
```

```{r raw_download}
for (i in seq_along(raw_zips)) {
raw_zip <- raw_zips[i]
  if (!file_exists(raw_zip)) {
    download.file(
      url = raw_urls[i], 
      destfile = raw_zip, 
      method = "curl", 
      extra = "--cipher 'DEFAULT:!DH'"
    )
  }
  
}
```

### Extract

We can extract each of these text files into our data directory.

```{r raw_extract}
raw_all <- map(raw_zips, unzip,
  zipfile = ,
  junkpaths = TRUE,
  exdir = raw_dir
) %>% unlist()
```

```{r raw_info}
file_info(raw_all) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## Read

```{r raw_fix}
fix_csv <- path_temp(basename(raw_all))
for (i in seq_along(raw_all)) {
  read_lines(raw_all[i]) %>% 
    str_replace_all("\\s\"([A-z\\. ]+)\"[^,|\"]*", " '\\1'") %>% 
    str_replace_all("(?<!^|,)\"(?!,|$)", r"("""")") %>% 
    write_lines(fix_csv[i])
}
```

The fixed text files can be read into a single data frame.

```{r raw_read}
nee <- read_delim(
  file = fix_csv,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = TRUE,
  na = c("", " "),
  col_types = cols(
    .default = col_character(),
  )
) %>% clean_names()
```

```{r}
problems(nee)
```

### Format

Now that each of the files is read as an element of a list, we can format them
for combination and matching against the tables containing full information on
the filing committees.

#### Date

```{r form_a1}
nee  <- nee %>% mutate(across(.cols = ends_with("date"),
                             .fns = lubridate::mdy))
```

## Explore

There are `r comma(nrow(nee))` rows of `r ncol(nee)` columns. Each record
represents a single contribution from an entity to a committee.

```{r glimpse}
glimpse(nee)
tail(nee)
```

```{r}
nee <- nee %>% rename(transaction_type = expenditure_transaction_type,
                      sub_type = expenditure_sub_type,
                      payee_type = payee_or_recipient_or_in_kind_contributor_type,
                      payee_name = payee_or_recipient_or_in_kind_contributor_name)

nee <- nee %>% 
   rename_with(~str_replace(., "expenditure_", "exp_")) 
  
```


### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(nee, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("exp_date", "filer_name", 
              "exp_amount", "payee_name")
nee <- flag_na(nee, all_of(key_vars))
mean(nee$na_flag)
sum(nee$na_flag)
```

```{r na_view}
nee %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

```{r}
nee %>% 
  summarise(
    n = n(),
    no_key_comm = mean(is.na(filer_name)),
    no_key_name = mean(is.na(payee_name)),
    no_key_amt = mean(is.na(exp_amount))
  ) %>% 
  arrange(desc(n))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
nee <- flag_dupes(nee, everything())
sum(nee$dupe_flag)
mean(nee$dupe_flag)
```

### Categorical

```{r distinct_count}
col_stats(nee, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(nee, filer_type)
explore_plot(nee, transaction_type)
```

### Amounts

```{r amount_round}
nee$exp_amount <- round(as.numeric(nee$exp_amount), 2)
```

```{r amount_summary}
summary(nee$exp_amount)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(nee[c(which.max(nee$exp_amount), which.min(nee$exp_amount)), ])
```

```{r hist_amount, echo=FALSE}
nee %>%
  filter(exp_amount >= 1) %>% 
  ggplot(aes(exp_amount)) +
  geom_histogram(fill = dark2["purple"], bins = 25) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Nebraska Expenditures Amount Distribution",
    caption = "Source: NADC",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nee <- mutate(nee, year = year(exp_date))
```

```{r date_range}
min(nee$exp_date)
sum(nee$year < 1984)
max(nee$exp_date)
sum(nee$exp_date > today())
```

```{r bar_year, echo=FALSE}
nee %>% 
  count(year) %>% 
  filter(year > 2020) %>%  
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  #scale_x_continuous(breaks = 2021:2023) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nebraska Expenditures by Year",
    caption = "Source: NADC",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
   nee <- nee %>% 
   unite(
    col = addr_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address-view}
sample_n(nee, 10) %>% glimpse()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
nee <- nee %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  nee$zip,
  nee$zip_norm,
  compare = valid_zip
)
```

### State

The existing `state` column needs not be cleaned.

```{r state_check}
prop_in(nee$state, valid_state)
what_out(nee$state, valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- nee %>% 
  distinct(city, state, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NE", "DC", "NEBRASKA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
nee <- left_join(
  x = nee,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- nee %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
nee <- nee %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  nee$city,
  nee$city_norm,
  nee$city_swap,
  nee$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(nee$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.95, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Nebraska City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nebraska City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
nee <- nee %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(nee, 1000))
```

1. There are `r comma(nrow(nee))` records in the database.
1. There are `r comma(sum(nee$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nee$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-dir}
clean_dir <- dir_create(here("state","ne", "expends", "data", "clean"))
clean_csv <- path(clean_dir, glue("ne_expends_2021-20230723.csv"))
basename(clean_csv)
```

```{r clean-write}
write_csv(nee, clean_csv, na = "")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
