---
title: "Nevada Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ks", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  rvest, # read html files
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The Nevada Secretary of State (NVSOS) office requires that one register for an account to access 
"[bulk data download](https://www.nvsos.gov/sos/online-services/data-download)" service page.

The process for downloaded a report is [outlined here](https://www.nvsos.gov/SoSServices/AnonymousAccess/HelpGuides/DataDownloadUserGuide.aspx):

In brief, we will be downloading a "Full Unabridged Database Dump" of "Campaign Finance" data.

> This report will expose Contributions and Expenses report data filed within our “Aurora” Campaign
Financial Disclosure system. This would not include data filed in a Financial Disclosure report.
This bulk data report tool here should be used to pull the entire database or slightly smaller
subsets of data such as all contributions filed after 1/1/2016 by groups of type “PAC”...

The site allows users to define the format for their data download. The site generated the
following summary of our data format:

> Your report will generate 6 "~" delimited ASCII text file(s) compressed into one Zip file named
in the format "CampaignFinance.43993.<Today's Date>.zip"*. Any field capable of containing a
non-numeric character (data types char, varchar, or datetime), will be enclosed in double quotes
(") so that if the field contains your delimiter you may identify it as being contained within the
field and not an actual delimiter. Any double quotes contained within these fields will be replaced
by 2 consecutive double quotes ("") so that the end of the field's data is not erroneously
identified. Below you will find the format of each file:

The above information provides the information needed to correctly parse each file using
`vroom::vroom()`.

The report data is partitioned into multiple files, as explained on the 
[NVSOS FAQ page](https://www.nvsos.gov/SOSServices/AnonymousAccess/HelpGuides/FAQ.aspx#5):

> This is what is referred to as a normalized relational structure in the database world.  Data
items such as business entities and officers have a direct relation to one another.  There can be
any number of officers to one business entity.  Because of this many to one relationship, the
officers data is stored in a different data table (or file) than the business entities.  Then we
relate officer records to a business entity record by a common key data column, in this case the
CorporationID... By separating officers and entities into separate records we can eliminate the
redundancy and added size associated with putting the business entity data on each officer record
or eliminate the complexity of allocating an undeterminable amount of officers on the one business
entity record.  This same many-to-one relationship is true of voter history records to voter
records, UCC actions to UCC liens or Corporation Stocks to Corporations, to name a few.

The summary continues to provide individual structure summaries on each of the six files included
in the report along with an key to the file name:

```{r read_html, echo=FALSE, warning=FALSE}
here("nv", "expends", "data") %>% 
  dir_ls(glob = "*.html") %>% 
  read_html() %>% 
  html_nodes("table") %>% 
  html_table(fill = TRUE) %>% 
  map(as_tibble) %>% 
  extract(seq(3, 18, by = 3)) %>% 
  map(slice, -1) %>% 
  map(set_names, c("col", "col_type", "position")) %>%
  map(separate, 
      col, 
      sep = "\\s\\(",
      into = c("col", "key"),
      extra = "drop"
  ) %>% 
  map(mutate, key = str_remove(key, "\\)")) %>%
  map(select, -position) %>% 
  set_names(c(
    "candidates", 
    "groups",
    "reports",
    "payees",
    "contributions",
    "expenses"
  ))
```

NVSOS provides some further information on a few variables in the "Result Field" tab of the report
generator:

* `Jurisdiction` = 

> This will be name of the city or county for city/county offices currently held by the candidate
(e.g. “CITY OF YERINGTON”, “DOUGLAS COUNTY”).  This will be set to “NV SOS” for statewide offices
such as Governor, State Controller or State assemblymen.  An office assigned to a candidate could
be updated by the NV SOS Elections staff as necessary when that candidate files for a new office.

* `Contribution Type` = 

> Use this column to differentiate which one of four contribution types this contribution record
is: Monetary Contribution, In Kind Contribution, In Kind Written Commitment, or Written Commitment.

* `Last Name` = 

> When the contributor or payee is an organization as opposed to an individual, the entire
organization name will be in the Last Name field only.

* `Expense Type` = 

> Use this column to differentiate which type of expense record this is: Monetary Expense or In
Kind Expense.

* `Active` = 

> A value of F (False) indicates the group has been marked as inactive by the NV Secretary of
State's office Elections division due to submission of a "notice of inactivity" or for failure to
renew annual registration.

* `Amended` = 

> A value of T (True) indicates this contributions and expense report has been marked as an amended
report by the original filer implying this report supersedes a report for this same period, filed
earlier.   An amended report is to be full comprehensive for that report period and in essence
replaces all contributions and expenses filed in the earlier report.

* `Election Cycle` = 

> The Election Cycle is the 4 digit filing or reporting year defining a filing period grouping
together a collection of contribution and expenses reports...

* `Superseded` = 

> A report is Superseded when an amended report was filed later by the same filer for the same
reporting period.  In this case the Superseded field for the older report record will be set to T
(True)...

## Read

The following link was sent via email and downloaded to the `data/` directory:

```
https://www.nvsos.gov/yourreports/CampaignFinance.43993.102819110351.zip
```

```{r raw_dir}
raw_dir <- here("nv", "expends", "data", "raw")
dir_create(raw_dir)
```

```{r raw_read, eval=FALSE}
raw_url <- "https://www.nvsos.gov/yourreports/CampaignFinance.43993.102819110351.zip"
raw_file <- url2path(raw_url, raw_dir)
download.file(raw_url, destfile = raw_file)
```

The ZIP file contains the six individual files, as expected.

```{r unzip_list, echo=FALSE}
raw_dir %>% 
  dir_ls(glob = "*.zip") %>% 
  unzip(list = TRUE) %>% 
  clean_names()
```

The files will be unzipped into the `data/raw/` directory.

```{r unzip}
raw_dir %>% 
  dir_ls(glob = "*.zip") %>% 
  unzip(exdir = raw_dir)
```

Each file can be read using the `vroom::vroom()` function and the parameters explained by NVSOS.
Variable names will be make "clean" (lowercase snake) using the `janitor::make_clean_names()` 
function.

```{r read_candidates}
nv_candidates <- vroom(
  file = dir_ls(raw_dir, glob = "*Cnddt*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    `CandidateID` = col_character(),
    `First Name` = col_character(),
    `Last Name` = col_character(),
    `Party` = col_character(),
    `Office` = col_character(),            
    `Jurisdiction` = col_character()
  )
)

print(nv_candidates)
```

```{r read_groups}
nv_groups <- vroom(
  file = dir_ls(raw_dir, glob = "*Grp*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    `GroupID` = col_character(),
    `Group Name` = col_character(),
    `Group Type` = col_character(),
    `Contact Name` = col_character(),            
    `Active` = col_logical(),
    `City` = col_character()
  )
)

print(nv_groups)
```

```{r read_reports}
nv_reports <- vroom(
  file = dir_ls(raw_dir, glob = "*Rpr*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    `ReportID` = col_character(),
    `CandidateID` = col_character(),
    `GroupID` = col_character(),
    `Report Name` = col_character(),
    `Election Cycle` = col_number(),
    `Filing Due Date` = col_date("%m/%d/%Y"),
    `Filed Date` = col_date("%m/%d/%Y"),
    `Amended` = col_logical(),
    `Superseded` = col_logical()
  )
)

print(nv_reports)
```

```{r read_payees}
nv_payees <- vroom(
  file = dir_ls(raw_dir, glob = "*Cntrbtrs*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    `ContactID` = col_character(),
    `First Name` = col_character(),
    `Middle Name` = col_character(),
    `Last Name` = col_character()
  )
)

print(nv_payees)
```

```{r read_contribs, eval=FALSE}
nv_contributions <- vroom(
  file = dir_ls(raw_dir, glob = "*Cntrbt.*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    `ContributionID` = col_character(),
    `ReportID` = col_character(),
    `CandidateID` = col_character(),
    `GroupID` = col_character(),
    `Contribution Date` = col_date("%m/%d/%Y"),
    `Contribution Amount`	= col_number(),
    `Contribution Type` = col_character(),
    `ContributorID` = col_character()
  )
)

print(nv_contributions)
```

```{r read_expenses}
nv_expenses <- vroom(
  file = dir_ls(raw_dir, glob = "*Expn*"),
  delim = "~",
  col_names = TRUE,
  na = "",
  quote = "\"",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    `Expense Date` = col_date("%m/%d/%Y"),
    `Expense Amount`	= col_number(),
  )
)
```

## Join

We are primarily interested in the file containing data on contributions. To make the data base
more searchable on the Accountability Project database, we will be joining together the various
normalized relational tables using their respective `*_id` variables. The expenses table will not
be added.

```{r join, collapse=TRUE}
nv <- nv_expenses %>%
  # join with relational tables
  left_join(nv_reports, by = c("report_id", "candidate_id", "group_id")) %>%
  left_join(nv_candidates, by = "candidate_id") %>% 
  left_join(nv_groups, by = "group_id") %>%
  left_join(nv_payees, by = c("payee_id" = "contact_id")) %>% 
  # add origin table info to ambiguous variables
  rename(
    candidate_first = first_name.x,
    candidate_last = last_name.x,
    candidate_party = party,
    seeking_office = office,
    report_amended = amended, 
    report_superseded = superseded,
    group_contact = contact_name,
    group_active = active,
    group_city = city,
    payee_first = first_name.y,
    payee_middle = middle_name,
    payee_last = last_name.y
  )

# all rows preserved
nrow(nv) == nrow(nv_expenses)

# all cols includes
ncol(nv_expenses) %>% 
  add(ncol(nv_reports)) %>% 
  add(ncol(nv_candidates)) %>% 
  add(ncol(nv_groups)) %>% 
  add(ncol(nv_payees)) %>% 
  subtract(6) %>% # shared key cols
  equals(ncol(nv))
```

This expands our primary table from `r ncol(nv_expenses)` variables to `r ncol(nv)`
without changing the number or records included.

## Explore

Variables containing information on the contributor themselves seem to be missing. While the
`payee_id` variable from the "Expenses" database is used to identify each contributor
in the "Contributors-Payees" using the `contact_id` variable, the later table contains only names.

This information is submitted by the filer of the report and can be found using the 
[NVSOS AURORA search portal](https://www.nvsos.gov/SoSCandidateServices/AnonymousAccess/CEFDSearchUU/Search.aspx), 
so we know the data exists.

For example, in report #6991, contributor #17 (John Mueller) gave money to candidate #2360 (E
Tiras). Searching for Mr. Tiras' report from 2006-08-08 on the AURORA portal, we can see that Mr.
Mueller lives at 308 Laura Court, Incline Village, NV 89451 (see
`nv_contribs/docs/nv_example_report`).

```{r no_geo}
nv %>% 
  filter(report_id == "6991") %>% 
  select(
    report_id, 
    filed_date, 
    payee_last, 
    candidate_last
  )
```

Below is the structure of the data arranged randomly by row. There are `r nrow(nv)` rows of 
`r length(nv)` variables.

```{r glimpse_all}
glimpse(sample_frac(nv))
```

### Distinct

The variables vary in their degree of distinctiveness.

The `expense_id` is `r percent(n_distinct(nv$expense_id)/nrow(nv))` distinct and can be used to
identify a unique contribution.

```{r count_distinct}
glimpse_fun(nv, n_distinct)
```

The `*_id` variables have as many distinct values as the length of their respective tables.

```{r distinct_id, collapse=TRUE}
n_distinct(nv_payees$contact_id)/nrow(nv_payees)
n_distinct(nv_groups$group_id)/nrow(nv_groups)
```

For the least distinct variables, we can explore the most common values.

```{r tabyls_function, echo=FALSE}
print_tabyl <- function(data, ...) {
  as_tibble(arrange(tabyl(data, ...), desc(n)))
}
```

```{r explore_plots, echo=FALSE}
explore_plot(drop_na(nv, expense_type), expense_type)
explore_plot(drop_na(nv, candidate_party), candidate_party)
explore_plot(drop_na(nv, seeking_office), seeking_office)
explore_plot(drop_na(nv, jurisdiction), jurisdiction)
explore_plot(drop_na(nv, report_amended), report_amended)
explore_plot(drop_na(nv, report_superseded), report_superseded)
explore_plot(drop_na(nv, group_type), group_type)
explore_plot(drop_na(nv, group_active), group_active)
explore_plot(drop_na(nv, group_city), group_city)
```

### Ranges

For continuous variables, the ranges should be checked.

```{r ranges}
summary(nv$expense_date)
summary(nv$expense_amount)
summary(nv$filing_due_date)
summary(nv$filed_date)
```

The date variables all seem to make sense. There are no dates before 
`r min(nv$expense_date)` and none from the future (aside from the upcoming filing dates). 

The maximum contribution is for `r noquote(scales::dollar(max(nv$expense_amount)))`.

```{r max_amt}
nv %>% 
  filter(expense_amount == max(expense_amount)) %>% 
  glimpse()
```

### Plot

Visualizations can be made to better understand the distributions of continuous and distinct
variables.

```{r plot_amt_type, echo=FALSE}
nv %>%
  ggplot(aes(expense_amount)) +
  geom_histogram(aes(fill = expense_type), bins = 30) +
  scale_x_continuous(labels = scales::dollar, trans = "log10") +
  theme(legend.position = "none") +
  facet_wrap(~expense_type, ncol = 1, scales = "free_y") +
  labs(
    title = "Expense Distribution",
    subtitle = "by Expense Type",
    caption = "Source: NVSOS",
    y = "Number of Contributions",
    x = "Amount (USD)"
  )
```

```{r plot_amt_party, echo=FALSE}
top_party <- most_common(nv$candidate_party, n = 5)
nv %>%
  mutate(candidate_party = if_else(candidate_party %in% top_party, candidate_party, "Other")) %>% 
  ggplot(aes(expense_amount)) +
  geom_histogram(aes(fill = candidate_party), bins = 30) +
  scale_x_continuous(labels = dollar, trans = "log10") +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none") +
  facet_wrap(~candidate_party, ncol = 1) +
  labs(
    title = "Expense Distribution",
    subtitle = "by Political Party",
    caption = "Source: NVSOS",
    y = "Number of Contributions",
    x = "Amount (USD)"
  )
```

```{r plot_amt_group, echo=FALSE}
nv %>%
  filter(!is.na(group_type)) %>% 
  ggplot(aes(expense_amount)) +
  geom_histogram(aes(fill = group_type), bins = 30) +
  scale_x_continuous(labels = scales::dollar, trans = "log10") +
  theme(legend.position = "none") +
  facet_wrap(~group_type, scales = "free_y") +
  labs(
    title = "Expense Distribution",
    subtitle = "to groups, by type",
    caption = "Source: NVSOS",
    y = "Number of Contributions",
    x = "Amount (USD)"
  )
```

### Missing

The variables also vary in their degree of values that are `NA` (empty). 

Some variables are mutually exclusive. There cannot be, for example, both `candidate_id` and
`group_id` value for the same record, as these two variables identify different kinds of
contributions. These mutually exclusive variables cover 100% of records.

In each of their respective original tables, there are no `NA` values. When you join them together,
any contribution to a candidate will have `NA` in all variables from the "Groups" table and visa
versa.

```{r mutually_exclusive, collapse=TRUE}
# prop NA each sum to 1
mean(is.na(nv$candidate_id)) + mean(is.na(nv$group_id))
mean(is.na(nv$candidate_last)) + mean(is.na(nv$group_name))
```

It's notable that the important variables (e.g., `*_id`, `payee_last`, `contribution_*`) contain 
zero missing values.

The full count of `NA` for each variable in the data frame can be found below:

```{r count_na}
glimpse_fun(nv, count_na)
```

### Duplicates

There are no duplicate expenses in the database.

```{r get_dupes, collapse=TRUE}
nrow(get_dupes(nv_expenses))
```

## Clean

For consistency sake, we can make all character columns uppercase.

```{r str_to_upper}
nv <- mutate_if(nv, is.character, str_to_upper)
```

### Year

Since the `contribution_date` variable was parsed as an R date object through `vroom::col_date()`,
the `lubridate::year()` function makes it easy to extract the contribution year from the
contribution date.

```{r mutate_year}
nv <- mutate(nv, expense_year = year(expense_date))
```

```{r}
nv %>% 
  count(expense_year) %>% 
  mutate(even = is_even(expense_year)) %>% 
  ggplot(aes(x = expense_year, n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma)
```

### Jurisdiction

There are `r n_distinct(nv$jurisdiction)` `jurisdiction` values.  Only one appears to be invalid.

```{r clean_jurisditction}
nv$jurisdiction_clean <- na_if(nv$jurisdiction, "UNKNOWN")
```

## Conclusion

1. There are `r nrow(nv)` records in the database
1. There are no duplicate records
1. All continous ranges make sense
1. There are very few missing values where there shouldn't be 
1. consistency issues among non-individual contributor names has been fixed
1. There is no ZIP code variable
1. The `expense_year` variable has been created from the `expense_date` variable
1. Records missing any key information are flagged with the `na_flag` variable

## Write

We will write the final data set to disk to be uploaded to the Accountability Project database.
To save space, unclean columns will be removed and `NA` values will be written as empty strings.

```{r proc_dir}
proc_dir <- here("nv", "expends", "data", "processed")
dir_create(proc_dir)
```

```{r proc_write}
nv %>% 
  mutate_if(is.character, str_replace_all, "\"", "\'") %>% 
  write_csv(
    path = glue("{proc_dir}/nv_expends_clean.csv"),
    na = ""
  )
```
