---
title: "New Mexico Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nm", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("nm/contribs/docs/nm_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The data available at [the old CFIS website][old_cris] is decommissioned and no
longer updated. The most recent contributions in the data were made in October
of 2020.

> #### !!Attention Candidates, Political Committees and Lobbyists!!
>
> This system has been decommissioned and is no longer available for use by
candidates and political committees. To file a disclosure report or to register
a new political committee, please use the new CFIS at
https://login.cfis.sos.state.nm.us
>
> Lobbyists, please continue to use this system through calendar year 2020. You will use the new CFIS for 2021. More information will follow. If you have any support questions, please contact sos.elections@state.nm.us.

Data from 2020 onward can be found from the New Mexico Secretary of state at a
[new CRIS portal][new_cris].

> This page provides comma separated value (CSV) downloads of contribution,
expenditure, and loan data for each reporting year in a zipped file format.
These files can be downloaded and imported into other applications (Microsoft
Excel, Microsoft Access, etc.)
> 
> This data is extracted as it existed as of 08/17/2021 12:09 PM. 

The new CRIS page also provides a [record layout][layout] for contributions.
From this layout, we know the new data has a different structure than the old
data. We will first download the old data and format it to take an overlapping
structure to the newer data.

[old_cris]: https://www.cfis.state.nm.us/media/CFIS_Data_Download.aspx
[new_cris]: https://login.cfis.sos.state.nm.us/
[layout]: https://login.cfis.sos.state.nm.us/Template/Contributions%20and%20Loans%20File%20Layout%20Key.pdf

## Download

We can download data from both CRIS websites using a combination of GET and POST
requests made to the server. These requests define what data we are interested
in and save the file locally.

```{r raw_dir}
raw_dir <- dir_create(here("nm", "contribs", "data", "raw"))
```

### Old

The old CRIS website uses an ASP.NET server. We can visit the home page and then
use the cookies from that visit to make two subsequent requests for campaign
finance transactions made by (1) Candidates and (2) PACs.

> #### CFIS Data Download
> The files available will be downloaded in a CSV file format. To download a
file, select the type of CFIS data and corresponding filters and click Download
Data. The file will be downloaded and should appear in the lower left corner of
your browser window. If you do not see the file, please check your browser’s
pop-up blocker settings.

> ##### Transactions
> Download a listing of all contributions and expenditures for one or all filing
periods for candidates, PACs, and Lobbyists.
>
> The CFIS data available for download is updated daily at 12:00AM and 12:00PM
MST.

```{r nm_home}
nm_url <- "https://www.cfis.state.nm.us/media/CFIS_Data_Download.aspx"
set_config(config(ssl_verifypeer = 0L))
nm_home <- GET(nm_url)
nm_html <- content(nm_home, as = "parsed")
nm_cookie <- cookies(nm_home)
nm_cookie <- setNames(nm_cookie$value, nm_cookie$name)
```

```{r}
view_state <- list(
  `__VIEWSTATE` = nm_html %>% 
    html_element("#__VIEWSTATE") %>% 
    html_attr("value"),
  `__VIEWSTATEGENERATOR` = nm_html %>% 
    html_element("#__VIEWSTATEGENERATOR") %>% 
    html_attr("value"),
  `__EVENTVALIDATION` = nm_html %>% 
    html_element("#__EVENTVALIDATION") %>% 
    html_attr("value")
)
```

```{r}
nm_types <- tribble(
  ~look_for,    ~file_name,
  "Candidates", "CandidateTransactions.csv", 
# "Lobbyist",   "LobbyistTransactions.csv", 
  "PACs",       "PACTransactions.csv",
)
```

```{r echo=FALSE}
nm_types <- mutate(nm_types, file_name = path(raw_dir, file_name))
nm_types %>% 
  mutate(across(file_name, path.abbrev)) %>% 
  kable(col.names = c("\"Look For\"", "File Path"))
```

```{r}
for (look_for in nm_types$look_for) {
  message(look_for)
  csv_path <- nm_types$file_name[match(look_for, nm_types$look_for)]
  if (!file_exists(csv_path)) {
    Sys.sleep(runif(1, 1, 3))
    POST(
      url = "https://www.cfis.state.nm.us/media/CFIS_Data_Download.aspx",
      write_disk(path = csv_path, overwrite = TRUE),
      set_cookies(nm_cookie),
      progress("down"),
      body = c( # combine extracted VIEWSTATE with form options 
        view_state, # __VIEWSTATE, __VIEWSTATEGENERATOR, __EVENTVALIDATION
        list( # dropdown body arguments: ALL rows for each file type
          `ctl00$ContentPlaceHolder1$header1$ddlCSVSelect` = "Transactions",
          `ctl00$ContentPlaceHolder1$header1$ddlRegisrationYear` = "0",
          `ctl00$ContentPlaceHolder1$header1$ddlViewBy` = "Lobbyist", # keep
          `ctl00$ContentPlaceHolder1$header1$hfFilePeriodFilter` = "ALL",
          `ctl00$ContentPlaceHolder1$header1$ddlLookFor` = look_for, # change
          `ctl00$ContentPlaceHolder1$header1$ddlFilePeriodYear` = "0",
          `ctl00$ContentPlaceHolder1$header1$ddlFPCan` = "ALL",
          `ctl00$ContentPlaceHolder1$header1$Button1` = "Download+Data",
          `ctl00$ContentPlaceHolder1$header1$hfLobbyistFilingPeriod` = "ALL",
          `ctl00$ContentPlaceHolder1$header1$ddlTransRegYear` = "0",
          `ctl00$ContentPlaceHolder1$header1$ddlFPLob` = "ALL"
        )
      )
    )
  }
}
```

### New

The new CRIS website hosts separate annual files for political contributions
and expenditures. These files can be checked with a request to the page.

```{r nm_check_get}
nm_check <- GET(
  "https://login.cfis.sos.state.nm.us/api///DataDownload/GetCheckDatadownload",
  query = list(
    pageNumber = 1,
    pageSize = 10,
    sortDir = "Desc",
    sortedBy = "ElectionYear"
  )
)
```

```{r nm_check_json}
nm_check <- content(
  x = nm_check, 
  as = "parsed", 
  type = "application/json"
)
```

The data on this page reports last being updates at 
`format(as_datetime(nm_check[[length(nm_check)]]$PSTTime), "%H:%M %b %d, %Y")`.
There are `length(nm_check) - 1` files currently hosted on the new CRIS download
page. We will download the files with `TransactionType` of `CON`, indicating
they contain contributions.

```{r echo=FALSE}
post_time <- as_datetime(nm_check[[length(nm_check)]]$PSTTime)
if (is.null(nm_check[[length(nm_check)]][[1]])) {
  nm_check <- nm_check[-length(nm_check)]
}
for (i in seq_along(nm_check)) {
  nm_check[[i]][map_lgl(nm_check[[i]], is.null)] <- NA
}
nm_check <- map_df(nm_check, as_tibble)
nm_check$PSTTime <- post_time
kable(nm_check)
nm_check <- filter(nm_check, TransactionType == "CON")
```

```{r}
for (i in seq(nrow(nm_check))) {
  message(nm_check$NameOfFile[i])
  con_path <- path(raw_dir, nm_check$NameOfFile[i])
  if (!file_exists(con_path)) {
    GET(
     "https://login.cfis.sos.state.nm.us/api/DataDownload/GetCSVDownloadReport",
      query = list(
        year = nm_check$ElectionYear[i],
        transactionType = nm_check$TransactionType[i],
        reportFormat = "csv",
        fileName = nm_check$NameOfFile[i],
      )
    )
  }
}
```

## Read

We will read the old and new file formats separately. The search format of the
Accountability Project database requires that files in the same dataset have the
same structure. The older files have fewer columns containing less overall data.
We can rename the shared columns to match and remove any columns in the older
format containing data not found in the newer format.

For example, if the older data contained columns [a, b, c, d] and the newer
format contained columns [A, B, C, E, F] then the older columns would be
renamed and removed to contain only [A, B, C] to overlap with the newer data.

### Old

```{r old_read}
nm_old <- map(
  .x = nm_types$file_name,
  .f = read_delim,
  delim = ",",
  na = c("", "NA", "NULL", "N/A"),
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `IsContribution` = col_logical(),
    `IsAnonymous` = col_logical(),
    `Amount` = col_double(),
    `Date Contribution` = col_date(),
    `Date Added` = col_datetime()
  )
)
```

The older data has two very similar formats for Candidate and PAC contributions.
Once the `PAC Name` column is renamed to `Last Name`, then the two can be bound
into a single table.

```{r old_filter}
names(nm_old) <- nm_types$look_for
nm_old$PACs <- rename(nm_old$PACs, `Last Name` = `PAC Name`)
nm_old <- bind_rows(nm_old)
```

Here is a random row from the old format.

```{r echo=FALSE}
glimpse(sample_n(nm_old, 1))
```

### New

```{r}
new_csv <- path(raw_dir, nm_check$NameOfFile)
```

```{r new_read}
nm_new <- read_delim(
  file = new_csv,
  escape_backslash = FALSE,
  escape_double = TRUE,
  locale = locale(tz = "MST"),
  col_types = cols(
    .default = col_character(),
    # `Transaction Date` = col_datetime("%m/%d/%Y %H:%M:%S %p"),
    # `Filed Date` = col_datetime("%m/%d/%Y %H:%M:%S %p"),
    # `Start of Period` = col_datetime("%b  %d %Y %H:%M%p"),
    # `End of Period` = col_datetime("%b  %d %Y %H:%M%p")
  )
)
```

```{r new_parse, echo=FALSE}
# https://github.com/tidyverse/readr/issues/1276
nm_new <- nm_new %>% 
  type_convert(
    locale = locale(tz = "MST"),
    col_types = cols(
      .default = col_character(),
      `Transaction Date` = col_datetime("%m/%d/%Y %H:%M:%S %p"),
      `Filed Date` = col_datetime("%m/%d/%Y %H:%M:%S %p"),
      `Start of Period` = col_datetime("%b  %d %Y %H:%M%p"),
      `End of Period` = col_datetime("%b  %d %Y %H:%M%p")
    )
  )
```

### Combine

```{r}

```

## Explore

There are `r comma(nrow(nmc))` rows of `r ncol(nmc)` columns. Each record
represents a single Contributions...

```{r glimpse}
glimpse(nmc)
tail(nmc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(nmc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("date", "last", "amount", "committee")
nmc <- flag_na(nmc, all_of(key_vars))
sum(nmc$na_flag)
```

```{r na_view}
nmc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
nmc <- flag_dupes(nmc, -id)
sum(nmc$dupe_flag)
```

```{r dupe_view}
nmc %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r distinct_count}
col_stats(nmc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(nmc, type)
```

### Amounts

```{r amount_summary}
summary(nmc$amount)
mean(nmc$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(nmc[c(which.max(nmc$amount), which.min(nmc$amount)), ])
```

```{r hist_amount, echo=FALSE}
nmc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New Mexico Contributions Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nmc <- mutate(nmc, year = year(date))
```

```{r date_range}
min(nmc$date)
sum(nmc$year < 2000)
max(nmc$date)
sum(nmc$date > today())
```

```{r bar_year, echo=FALSE}
nmc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Mexico Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
nmc <- nmc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
nmc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nmc <- nmc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nmc$zip,
  nmc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nmc <- nmc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
nmc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  nmc$state,
  nmc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- nmc %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NM", "DC", "NEW MEXICO"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
nmc <- left_join(
  x = nmc,
  y = norm_city,
  by = c(
    "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- nmc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
nmc <- nmc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(nmc$city_raw),
  nmc$city_norm,
  nmc$city_swap,
  nmc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nmc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New Mexico City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Mexico City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
nmc <- nmc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(nmc, 50))
```

1. There are `r comma(nrow(nmc))` records in the database.
1. There are `r comma(sum(nmc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nmc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nm", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "nm_contribs_clean.csv")
write_csv(nmc, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(nmc)),
  type = md_code(map_chr(nmc, typeof)),
  def = c(
    ""
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
