---
title: "Alaska Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ak", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  httr, # http requests
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Alaska Public Offices Commission (APOC)][apoc]. 

[apoc]: https://aws.state.ak.us/ApocReports/Campaign/#

## Download

Using the [APOC income search][search], we need to search for "All Completed
Forms", "Any Names", and "Any" type. Exporting "Any" report year only returns
roughly 400,000 results despite 1,411,206 results being listed in the search.
We can export all of the results if we instead search and export individually
by report year. As of now, this needs to be done by hand.

[search]: https://aws.state.ak.us/ApocReports/CampaignDisclosure/CDIncome.aspx

```{r line_count}
line_count <- function(path) {
  as.numeric(str_extract(system(paste("wc -l", path), intern = TRUE), "^\\d+"))
}
```

```{r raw_dir}
raw_dir <- dir_create(here("state","ak", "contribs", "data", "raw"))
raw_info <- dir_info(raw_dir)
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  rowwise() %>% 
  mutate(lines = line_count(path)) %>% 
  mutate(across(path, basename))
```

```{r raw_get, eval=FALSE}
httr::GET(
  # does not work
  url = "https://aws.state.ak.us/ApocReports/CampaignDisclosure/CDIncome.aspx",
  path = raw_path,
  query = list(
    exportAll = "True",
    exportFormat = "CSV",
    isExport = "True"
  ),
)
```

## Read

The exported delimited text files have two aspects we need to adjust for; 1)
There is a column called `--------` that is empty in every file, and 2) there
is an extra comma at the end of each line. We can read this extra column at the
end as a new `null` column.

```{r raw_names}
ak_names <- make_clean_names(read_names(raw_info$path[1]))
```

All the files can be read into a single data frame using `vroom::vroom()`.

```{r raw_read}
# 1,411,206 items
akc <- vroom(
  file = raw_info$path,
  skip = 1,
  delim = ",",
  id = "file",
  num_threads = 1,
  na = c("", "NA", "N/A"),
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_names = c(ak_names, "null"),
  col_types = cols(
    .default = col_character(),
    date = col_date_mdy(),
    amount = col_number(),
    report_year = col_integer(),
    submitted = col_date_mdy(),
    null = col_logical()
  )
)
```

```{r echo=FALSE}
akc <- akc %>% 
  select(-x, -null) %>% 
  relocate(file, .after = last_col()) %>% 
  mutate(across(file, basename)) %>% 
  rename(
    id = result,
    tran_type = transaction_type,
    pay_type = payment_type,
    pay_detail = payment_detail,
    last = last_business_name,
    first = first_name,
    purpose = purpose_of_expenditure,
    filer = name
  )
```

We successfully read the same number of rows as search results.

```{r raw_rows}
nrow(akc) == 1411206
```

`r percent(mean(akc$tran_type == "Income"))` of rows have an "Income" type.

```{r raw_check}
count(akc, tran_type)
```

## Explore

There are `r comma(nrow(akc))` rows of `r ncol(akc)` columns.

```{r glimpse}
glimpse(akc)
head(akc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(akc, count_na)
```

We can flag any rows that are missing a name, date, or amount needed to identify
a transaction.

```{r na_flag}
key_vars <- c("date", "last", "amount", "filer")
akc <- flag_na(akc, all_of(key_vars))
sum(akc$na_flag)
```

All of these missing key values are the `last` name of the contributor.

```{r na_view}
akc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

```{r na_recount}
akc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

Ignoring the supposedly unique `id` variable, quite a few records are entirely
duplicated. We will not remove these records, as they may very well be valid
repetitions, but we can flag them with a new logical variable.

```{r dupe_flag}
d1 <- duplicated(akc[, -1], fromLast = TRUE)
d2 <- duplicated(akc[, -1], fromLast = FALSE)
akc <- mutate(akc, dupe_flag = (d1 | d2))
percent(mean(akc$dupe_flag), 0.1)
rm(d1, d2); flush_memory()
```

```{r dupe_view}
akc %>% 
  filter(dupe_flag) %>% 
  select(id, all_of(key_vars))
```

### Categorical

Columns also vary in their degree of distinctiveness. Some character columns 
like `first` name are obviously mostly distinct, others like `office` only have
a few unique values, which we can count.

```{r n_distinct}
col_stats(akc, n_distinct)
```

```{r bar_categorical, echo=FALSE}
explore_plot(akc, pay_type)
explore_plot(akc, report_type)
explore_plot(akc, election_type)
explore_plot(akc, office)
explore_plot(akc, filer_type)
```

### Amounts

```{r amount_summary}
noquote(map_chr(summary(akc$amount), dollar))
percent(mean(akc$amount <= 0), 0.01)
```

There are only `r comma(n_distinct(akc$amount))` values, which is an order of
magnitude less than we might expect from a distribution of values from a dataset
of this size.

In fact, more than half of all `amount` values are \$2, \$5, \$50, \$100, or 
\$500 dollars even.

```{r amount_count}
akc %>% 
  count(amount, sort = TRUE) %>% 
  add_prop(sum = TRUE)
```

```{r hist_amount, echo=FALSE}
akc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Alaska Contributions Amount Distribution",
    caption = "Source: APOC",
    x = "Amount",
    y = "Count"
  )
```

### Dates

```{r date_year}
akc <- mutate(
  .data = akc,
  date_clean = date %>% 
    # fix bad years with regex
    str_replace("^(210)(?=\\d-)", "201") %>% 
    str_replace("^(202)(?=[13-9])", "201") %>% 
    str_replace("^(29)(?=\\d-)", "20") %>% 
    str_replace("^(291)(?=\\d-)", "201") %>% 
    str_replace("^(301)(?=\\d-)", "201") %>% 
    as_date(),
  year_clean = year(date_clean)
)
```

```{r date_range}
min(akc$date_clean)
sum(akc$year_clean < 2011)
max(akc$date_clean)
sum(akc$date_clean > today())
```

```{r bar_year, echo=FALSE}
akc %>% 
  count(year_clean) %>% 
  mutate(even = is_even(year_clean)) %>% 
  ggplot(aes(x = year_clean, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2011, 2020)) +
  coord_cartesian(xlim = c(2011, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Alaska Contributions by Year",
    caption = "Source: APOC",
    fill = "Even Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
akc <- akc %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
akc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
akc <- akc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  akc$zip,
  akc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
akc <- akc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_changes}
akc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_view}
akc %>% 
  filter(state_norm %out% valid_state) %>% 
  count(state, state_norm, sort = TRUE) %>% 
  print(n = Inf)
```

```{r state_fix}
akc <- akc %>% 
  mutate(
    state_norm = state_norm %>% 
      na_in("NONE") %>% 
      str_replace("^ALBERTA$", "AB") %>% 
      str_replace("^BRITISH COLUMBIA$", "BC") %>% 
      str_replace("^NEW ZEALAND$", "NZ") %>% 
      str_replace("^NZL$", "NZ") %>% 
      str_replace("^ONTARIO$", "ON") %>% 
      str_replace("^PHILIPPINES$", "PH") %>% 
      str_replace("^PHILLIPINES$", "PH") %>% 
      str_replace("^SURRY HILLS$", "NSW") %>% 
      str_replace("^EUROPE$", "EU") %>% 
      str_remove("\\s\\w+ CANADA") %>% 
      str_squish()
  )
```

```{r state_progress}
progress_table(
  akc$state,
  akc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
akc <- akc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("AK", "DC", "ALASKA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
akc <- akc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- akc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
akc <- akc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(akc$city_raw),
  akc$city_norm,
  akc$city_swap,
  akc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(akc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Alaska City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Alaska City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
akc <- akc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(akc, 20))
```

1. There are `r comma(nrow(akc))` records in the database.
1. There are `r comma(sum(akc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(akc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("ak", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ak_contribs_clean.csv")
write_csv(akc, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=TRUE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    multipart = TRUE,
    show_progress = TRUE
  )
}
```

```{r s3_size}
r <- head_object(s3_path, "publicaccountability")
(s3_size <- as_fs_bytes(attr(r, "content-length")))
unname(s3_size == clean_size)
```
