---
title: "Indiana Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("in", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  rvest, # read html pages
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Indiana Election Division][ied].

> ##### What is the quality of the data? 	 
> The information presented in the campaign finance database is, to the best of
our ability, an accurate representation of the reports filed with the Election
Division. This information is being provided as a service to the public, has
been processed by the Election Division and should be cross-referenced with the
original report on file with the Election Division.
> 
> Some of the information in the campaign finance database was submitted in
electronic form. Most of the information was key-entered from paper reports.
Sometimes items which are inconsistent with filing requirements, such as
incorrect codes or incorrectly formatted or blank items, are present in the
results of a query. They are incorrect or missing in the database because they
were incorrect or missing on the reports submitted to the Election Division. For
some incorrect or missing data in campaign finance reports, the Election
Division has requested that the filer supply an amended report. The campaign
finance database will be updated to reflect amendments received.

> ##### What does the database contain? 	 
> By Indiana law, candidates and committees are required to disclose detailed
financial records of contributions received and expenditures made and debts owed
by or to the committee. For committees, the campaign finance database contains
all contributions, expenditures, and debts reported to the Election Division
since January 1, 1998.

[ied]: https://campaignfinance.in.gov/PublicSite/Homepage.aspx

## Import

The IED provides annuel files for both campaign contributions and expenditures.

> This page provides comma separated value (CSV) downloads of contribution and
expenditure data for each reporting year in a zipped file format. These files
can be downloaded and imported into other applications (Microsoft Excel,
Microsoft Access, etc.).
> 
> This data was extracted from the Campaign Finance database as it existed as of
2/4/2020  1:00 AM.

### Download

We can read the [IED download page][dl] to get the list of URLs to each file.

[dl]: https://campaignfinance.in.gov/PublicSite/Reporting/DataDownload.aspx

```{r raw_dir}
raw_dir <- dir_create(here("in", "contribs", "data", "raw"))
```

```{r raw_find, results='asis'}
aspx <- "https://campaignfinance.in.gov/PublicSite/Reporting/DataDownload.aspx"
raw_urls <- aspx %>% 
  read_html() %>% 
  html_node("#_ctl0_Content_dlstDownloadFiles") %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("Contribution") %>% 
  str_replace("\\\\", "/")

md_bullet(raw_urls)
```

We can download each of these files to the raw directory.

```{r raw_download}
raw_paths <- path(raw_dir, basename(raw_urls))
if (!all(this_file_new(raw_paths))) {
  download.file(raw_urls, raw_paths, method = "libcurl")
}
```

Then, we will unzip each file and delete the original.

```{r raw_unzip}
raw_paths <- as_fs_path(map_chr(raw_paths, unzip, exdir = raw_dir))
dir_ls(raw_dir, regexp = ".zip$") %>% file_delete()
```

### Read

There are two problems with each of these files:
1. When the second line of an address was entered, a `\n` newline character was
enteted between the two lines _within the same field_. The fields are surrounded
in double-quotes, but when reading the files these newlines mess things up.
2. Whenever a string itself contains `"` double-quotes, the first occurance is
registered as the end of the field itself, which begun with a `"`.

To fix these issues, we will read each file as a single character string and
use regular expressions to find and replace these errant `\n` and `"` 
characters. We will then write the edited strings to a new file.

```{r raw_fix}
fix_dir <- dir_create(here("in", "contribs", "data", "fix"))
if (!any(file_exists(dir_ls(fix_dir)))) {
  for (file in raw_paths) {
    read_file(file) %>% 
      # find newlines not at end of line
      str_replace_all("(?<!\"(\r|1|0)\")\n(?!\"\\d{1,10}\")", " ") %>%
      # find quotes not at end of field
      str_replace_all("(?<!(\n|^|,))\"(?!(,(?=\"))|$|\r)", "\'") %>% 
      str_trim(side = "both") %>% 
      # save to disk
      write_file(path = path(fix_dir, basename(file)))
    gc()
  }
}
fix_paths <- dir_ls(fix_dir)
```

These fixed files can be read into a single data frame with `purrr::map_df()`
and `readr::read_delim()`.

```{r raw_read}
inc <- map_df(
  fix_paths,
  read_delim,
  delim = ",",
  quote = "\"",
  na = c("", "n/a", "NA", "N/A"),
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    FileNumber = col_integer(),
    Amount = col_double(),
    ContributionDate = col_datetime(),
    Amended = col_logical()
  )
)
```

```{r raw_rename}
inc <- inc %>% 
  clean_names("snake") %>% 
  rename(
    file = file_number,
    type = contributor_type,
    date = contribution_date,
    method = type
  )
```

## Explore

```{r glimpse}
head(inc)
tail(inc)
glimpse(sample_n(inc, 20))
```

### Missing

```{r count_na}
col_stats(inc, count_na)
```

```{r flag_na}
inc <- inc %>% flag_na(committee, name, amount, date)
mean(inc$na_flag)
```

### Duplicates

```{r flag_dupes}
inc <- flag_dupes(inc, everything())
mean(inc$dupe_flag)
```

### Categorical

```{r n_distinct}
col_stats(inc, n_distinct)
```

```{r comm_type_bar, echo=FALSE}
explore_plot(
  data = inc,
  var = committee_type,
  title = "Indiana Contribution Recipient Types"
)
```

```{r cont_type_bar, echo=FALSE}
explore_plot(
  data = inc,
  var = type,
  title = "Indiana Contribution Contributor Types"
)
```

```{r method_bar, echo=FALSE}
explore_plot(
  data = filter(inc, !is.na(method)),
  var = method,
  title = "Indiana Contribution Contribution Method"
)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(inc$amount)
mean(inc$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
inc %>%
  filter(amount > 1, amount < 1e6) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contribution Distribution",
    caption = "Source: Indiana Election Division",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_comm_violin, echo=FALSE}
inc %>%
  filter(amount > 1, amount < 1e6) %>% 
  ggplot(aes(x = committee_type, y = amount)) +
  geom_violin(aes(fill = committee_type)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contribution Distribution",
    caption = "Source: Indiana Election Division",
    x = "Amount",
    y = "Count"
  )
```

