---
title: "Indiana Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("in", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  rvest, # read html pages
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Indiana Election Division][ied].

> ##### What is the quality of the data? 	 
> The information presented in the campaign finance database is, to the best of
our ability, an accurate representation of the reports filed with the Election
Division. This information is being provided as a service to the public, has
been processed by the Election Division and should be cross-referenced with the
original report on file with the Election Division.
> 
> Some of the information in the campaign finance database was submitted in
electronic form. Most of the information was key-entered from paper reports.
Sometimes items which are inconsistent with filing requirements, such as
incorrect codes or incorrectly formatted or blank items, are present in the
results of a query. They are incorrect or missing in the database because they
were incorrect or missing on the reports submitted to the Election Division. For
some incorrect or missing data in campaign finance reports, the Election
Division has requested that the filer supply an amended report. The campaign
finance database will be updated to reflect amendments received.

> ##### What does the database contain? 	 
> By Indiana law, candidates and committees are required to disclose detailed
financial records of contributions received and expenditures made and debts owed
by or to the committee. For committees, the campaign finance database contains
all contributions, expenditures, and debts reported to the Election Division
since January 1, 1998.

[ied]: https://campaignfinance.in.gov/PublicSite/Homepage.aspx

## Import

The IED provides annuel files for both campaign contributions and expenditures.

> This page provides comma separated value (CSV) downloads of contribution and
expenditure data for each reporting year in a zipped file format. These files
can be downloaded and imported into other applications (Microsoft Excel,
Microsoft Access, etc.).
> 
> This data was extracted from the Campaign Finance database as it existed as of
2/4/2020  1:00 AM.

### Download

We can read the [IED download page][dl] to get the list of URLs to each file.

[dl]: https://campaignfinance.in.gov/PublicSite/Reporting/DataDownload.aspx

```{r raw_dir}
raw_dir <- dir_create(here("in", "contribs", "data", "raw"))
```

```{r raw_find, results='asis'}
aspx <- "https://campaignfinance.in.gov/PublicSite/Reporting/DataDownload.aspx"
raw_urls <- aspx %>% 
  read_html() %>% 
  html_node("#_ctl0_Content_dlstDownloadFiles") %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("Contribution") %>% 
  str_replace("\\\\", "/")

md_bullet(raw_urls)
```

We can download each of these files to the raw directory.

```{r raw_download}
raw_paths <- path(raw_dir, basename(raw_urls))
if (!all(this_file_new(raw_paths))) {
  download.file(raw_urls, raw_paths, method = "libcurl")
}
```

Then, we will unzip each file and delete the original.

```{r raw_unzip}
if (length(dir_ls(raw_dir, regexp = ".csv$")) == 0) {
  raw_paths <- as_fs_path(map_chr(raw_paths, unzip, exdir = raw_dir))
} else {
  raw_paths <- dir_ls(raw_dir, regexp = ".csv$")
}
```

### Read

There are two problems with each of these files:
1. When the second line of an address was entered, a `\n` newline character was
enteted between the two lines _within the same field_. The fields are surrounded
in double-quotes, but when reading the files these newlines mess things up.
2. Whenever a string itself contains `"` double-quotes, the first occurance is
registered as the end of the field itself, which begun with a `"`.

To fix these issues, we will read each file as a single character string and
use regular expressions to find and replace these errant `\n` and `"` 
characters. We will then write the edited strings to a new file.

```{r raw_fix}
fix_dir <- dir_create(here("in", "contribs", "data", "fix"))
if (!any(file_exists(dir_ls(fix_dir)))) {
  for (file in raw_paths) {
    read_file(file) %>% 
      # find newlines not at end of line
      str_replace_all("(?<!\"(\r|1|0)\")\n(?!\"\\d{1,10}\")", " ") %>%
      # find quotes not at end of field
      str_replace_all("(?<!(\n|^|,))\"(?!(,(?=\"))|$|\r)", "\'") %>% 
      str_trim(side = "both") %>% 
      # save to disk
      write_file(path = path(fix_dir, basename(file)))
    gc()
  }
}
fix_paths <- dir_ls(fix_dir)
```

These fixed files can be read into a single data frame with `purrr::map_df()`
and `readr::read_delim()`.

```{r raw_read}
inc <- map_df(
  fix_paths,
  read_delim,
  delim = ",",
  quote = "\"",
  na = c("", "n/a", "NA", "N/A"),
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    FileNumber = col_integer(),
    Amount = col_double(),
    ContributionDate = col_datetime(),
    Amended = col_logical()
  )
)
```

```{r raw_rename}
inc <- inc %>% 
  clean_names("snake") %>% 
  rename(
    file = file_number,
    type = contributor_type,
    date = contribution_date,
    method = type
  )
```

## Explore

```{r glimpse}
head(inc)
tail(inc)
glimpse(sample_n(inc, 20))
```

### Missing

```{r count_na}
col_stats(inc, count_na)
```

```{r flag_na}
inc <- inc %>% flag_na(committee, name, amount, date)
mean(inc$na_flag)
```

### Duplicates

```{r flag_dupes}
inc <- flag_dupes(inc, everything())
mean(inc$dupe_flag)
```

### Categorical

```{r n_distinct}
col_stats(inc, n_distinct)
```

```{r comm_type_bar, echo=FALSE}
explore_plot(
  data = inc,
  var = committee_type,
  title = "Indiana Contribution Recipient Types"
)
```

```{r cont_type_bar, echo=FALSE}
explore_plot(
  data = inc,
  var = type,
  title = "Indiana Contribution Contributor Types"
)
```

```{r method_bar, echo=FALSE}
explore_plot(
  data = filter(inc, !is.na(method)),
  var = method,
  title = "Indiana Contribution Contribution Method"
)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(inc$amount)
mean(inc$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
inc %>%
  filter(amount > 1, amount < 1e6) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contribution Distribution",
    caption = "Source: Indiana Election Division",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_comm_violin, echo=FALSE}
inc %>%
  filter(amount > 1, amount < 1e6) %>% 
  ggplot(aes(x = committee_type, y = amount)) +
  geom_violin(aes(fill = committee_type)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contribution Distribution",
    caption = "Source: Indiana Election Division",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
inc <- mutate(
  .data = inc,
  date = as_date(date),
  year = year(date),
  year = case_when(
    year < 1998 ~ 1998,
    year > 2020 ~ 2020,
    year %>% between(1998, 2020) ~ year
  )
)
```

```{r year_bar}
inc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1998, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Indiana Contributions by Year",
    caption = "Source: Indiana Election Division",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
inc <- inc %>%
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

We can see how this process improved consistency.

```{r address_view}
inc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
inc <- inc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  inc$zip,
  inc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
inc <- inc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_view}
inc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

We can further improve these values by checking the state abbreviation against
the _expected_ abbreviation for that record's `zip_norm` variable. If the 
invalid abbreviation is only 1 letter off the expected value, we can confidently
repair these typos.

```{r state_swap}
inc <- inc %>% 
  left_join(
    y = select(zipcodes, -city), 
    by = c("zip_norm" = "zip"),
    suffix = c("_raw", "_match")
  ) %>% 
  mutate(
    match_dist = str_dist(state_raw, state_match),
    state_norm = if_else(
      condition = !is.na(state_match) & match_dist == 1,
      true = state_match,
      false = state_norm
    )
  ) %>% 
  rename(state = state_raw)

inc %>% 
  filter(match_dist == 1) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_select, echo=FALSE}
inc <- inc %>% 
  select(
    -state_match,
    -match_dist,
  )
```

```{r state_progress}
progress_table(
  inc$state,
  inc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats. The `campfin::normal_city()` 
function is a good start, again converting case, removing punctuation, but
_expanding_ USPS abbreviations. We can also remove `invalid_city` values.

```{r city_invalid, echo=FALSE}
invalid_city <- c(
  invalid_city,
  "CITY OMITTED ON REPORT",
  "OMITTED ON REPORT",
  "OMITTED FROM REPORT"
)
```

```{r city_norm}
inc <- inc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("IN", "DC", "INDIANA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

Again, we can further improve normalization by comparing our normalized value
agaist the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
inc <- inc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(city_match) & match_abb | match_dist <= 2,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_other}
many_city <- c(valid_city, extra_city)
inc %>% 
  count(city_swap, state_norm, sort = TRUE) %>% 
  filter(!is.na(city_swap), city_swap %out% many_city)
```

```{r city_replace}
inc <- inc %>% 
  mutate(
    city_swap = city_swap %>% 
      str_replace("^INDY$", "INDIANAPOLIS") %>% 
      str_replace("^INDPLS$", "INDIANAPOLIS") %>% 
      str_replace("^ABBOTT PARKS$", "ABBOTT PARK") %>% 
      str_replace("^OVERLAND PARKS$", "OVERLAND PARK") %>% 
      str_remove("\\sIN$")
  )
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(inc$city_raw),
  inc$city_norm,
  inc$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Indiana City Normalization Progress",
    subtitle = "Raw at 21.5% before converted to uppercase",
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Indiana City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(inc, 20))
```

1. There are `r comma(nrow(inc))` records in the database.
1. There are `r comma(sum(inc$dupe_flag))` duplicate records in the database 
(`r percent(mean(inc$dupe_flag), accuracy = 0.01)`).
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(inc$na_flag))` records missing a contributor or 
recipient name, date, or amount (`r percent(mean(inc$na_flag), accuracy = 0.01)`).
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("in", "contribs", "data", "clean"))
```

```{r clean_write}
inc <- inc %>% 
  select(
    -city_norm,
    city_norm = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean"))

write_csv(
  x = inc,
  path = path(clean_dir, "in_contribs_clean.csv"),
  na = ""
)
```
