---
title: "Data Diary"
subtitle: "Nevada Contributions"
author: "Kiernan Nicholls"
date: "`r format(Sys.time())`"
output:
  html_document: 
    df_print: tibble
    fig_caption: yes
    highlight: tango
    keep_md: yes
    max.print: 32
    toc: yes
    toc_float: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Objectives

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called ZIP5
1. Create a YEAR field from the transaction date
1. For campaign donation data, make sure there is both a donor AND recipient

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, warning=FALSE, error=FALSE}
# install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # date strings
  magrittr, # pipe opperators
  janitor, # data cleaning
  zipcode, # clean and data
  refinr, # cluster and merge
  rvest, # scrape web pages
  knitr, # knit documents
  here, # navigate local storage
  fs # search local storage 
)
```

```{r fix_fun, echo=FALSE}
here <- here::here
"%out%" <- Negate("%in%")
print_all <- function(df) df %>% print(n = nrow(.)) 
```

## Data

If the raw data has not been downloaded, it can be retrieved from the 
[Oklahoma Ethics Commision's website](https://www.ok.gov/ethics/public/login.php) as a ZIP archive.

> Everyone has access to the public disclosure system, the only secured access point is the
downloadable raw data option. This option provides an entire database dump in comma separated value
(.csv) or tab delimited (.txt) formats. This secure area is intended for use by the media and
import into an existing database.

```{r any_old_files, echo=FALSE}
any_old_files <- function(path, glob) {
  path %>%
    dir_ls(type = "file", glob = glob) %>% 
    file_info() %>% 
    pull(modification_time) %>% 
    floor_date("day") %>% 
    equals(today()) %>% 
    not() %>% 
    any()
}
```

```{r download_zip}
if (here("ok_contribs", "data", "raw") %>% any_old_files("*.zip")) {
  download.file(
    url = "https://www.ok.gov/ethics/public/dfile.php?action=csv",
    destfile = here("ok_contribs", "data", "raw", "ethicscsvfile.zip")
  )
}
```

There are 48 individual CSV files contained within the ZIP archive. Many of these files are not
relevant to this project, but all will be unzipped into the `data/raw` directory.

```{r unzip_list, echo=FALSE}
unzip(
  zipfile = here("ok_contribs", "data", "raw", "ethicscsvfile.zip"),
  list = TRUE
)
```

If these files have not yet been unzipped, they will be now.

```{r unzip}
if (here("ok_contribs", "data", "raw") %>% any_old_files("*.csv")) {
  unzip(
    zipfile = here("ok_contribs", "data", "raw", "ethicscsvfile.zip"),
    exdir = here("ok_contribs", "data", "raw"),
    overwrite = TRUE
  )
}
```

## Read

The data of interest is spread across a number of different files than can be joined along their
respective `*_id` variables. The `transaction.csv` contains the list of contributions and expenses,
and data on those transactions is spread across other tables. 

The relationship between these files is described in the `data/relations_db.xls` Excel file.
Descriptions for each of these files is provided int the `data/descriptions.doc` Word file.

In general, there are three _types_ of files that need to be read and joined together

1. All transactions
    * `transaction.csv`
1. Contributor information
    * `contributor.csv`
      * `cont_type.csv`
    * `individual_cont.csv`
    * `business_cont.csv`
    * `committee_cont.csv`
    * `vendor_cont.csv`
1. Recipient information
    * `so1.csv`
    * `so2.csv`
    * `party.csv`
    * `district.csv`
    * `office.csv`
    * `affiliation.csv`
    * `report.csv`
    * `lump_fund.csv`
    * `surplus.csv`
    * `refund.csv`

They will each be read as data frames using `readr::read_csv()`. All files contain an erroneous
trailing column in the header resulting in an empty column that will be removed. All variable names
will be made "clean" (lowercase and snake_case) using `janitor::clean_names()`.

### Transactions

> Holds all the contribution and expenditure transactions. Has the transaction date, amount, the contributor id and report number (report_num) that it ties back to in the report table.

```{r read_transaction}
transactions <- read_csv(
  file = "ok_contribs/data/raw/transaction.csv",
  col_types = cols(
    TRANS_INDEX = col_character(),
    TRANSACTION_DATE = col_date("%d-%b-%y"),
    CONTRIBUTOR_ID = col_character(),
    TRANS_AMOUNT = col_double(),
    REP_NUM = col_character()
  )
)

transactions <- transactions %>% 
  select(-X6) %>% 
  clean_names() %>% 
  rename(
    cont_id = contributor_id,
    trans_date = transaction_date
    )

print(transactions)
```

### Contributors

> Holds address, phone and type of any contributor using [`contributor_id`] as its identifier in
other tables.

```{r read_contributor}
contributors <- read_csv(
  file = "ok_contribs/data/raw/contributor.csv",
  col_types = cols(.default = col_character())
)

contributors <- contributors %>%
  select(-X9, -PHONE, -EXT) %>% 
  clean_names() %>% 
  rename(
    cont_id = contributor_id,
    cont_type = type,
    cont_street = street,
    cont_city = city,
    cont_state = state,
    cont_zip = zip
  )

print(contributors)
```

> Holds the different contributor types available (Individual, Business, Committee, Vendor)

```{r read_cont_type}
cont_types <- read_csv(
  file = "ok_contribs/data/raw/cont_type.csv",
  col_types = cols(.default = col_character())
)

cont_types <- cont_types %>%
  select(-X3) %>% 
  clean_names()
```

#### Individual Contributors

> Holds information relating to any individual contributor. Name, employer and occupation. Contributor id is the key that goes back to the contributor table and into either the transaction table for a transaction list or contributor aggregate table for tie ins to the `ethics_num` (committee) with aggregate totals.

```{r read_individual_cont}
individual_conts <- read_csv(
  file = "ok_contribs/data/raw/individual_cont.csv",
  col_types = cols(.default = col_character())
)

individual_conts <- individual_conts %>% 
  select(-X7) %>% 
  clean_names() %>% 
  rename(
    cont_id = contributor_id,
    cont_employer = employer,
    cont_occupation = occupation
  )
```

#### Business Contributors

> Holds the business name (`cont_name`) and business activity of a business contributor.
Contributor id is the key that goes back to the contributor table and into either the transaction
table for a transaction list or contributor aggregate table for tie ins to the `ethics_num`
(committee) with aggregate totals.

```{r read_business_cont}
business_conts <- read_csv(
  file = "ok_contribs/data/raw/business_cont.csv",
  col_types = cols(.default = col_character())
)

business_conts <- business_conts %>% 
  select(-X4) %>% 
  clean_names() %>% 
  rename(
    cont_bname = cont_name,
    cont_activity = business_activity
    )
```

#### Committee Contributors

> Holds the principal interest, contributor committee name and contributor FEC number and
committees ethics number for any committee contributors (`contributor_id`). Contributor id is the
key that goes back to the contributor table and into either the transaction table for a transaction
list or contributor aggregate table for tie ins to the ethics_num (committee) with aggregate
totals.

```{r read_committee_cont}
committee_conts <- read_csv(
  file = "ok_contribs/data/raw/committee_cont.csv",
  col_types = cols(.default = col_character())
)

committee_conts <- committee_conts %>% 
  select(-X6) %>% 
  clean_names() %>% 
  rename(
    cont_interest = principal_interest,
    cont_id = id,
    ethics_id = ethics_num,
    cont_cname = committee_name
  )
```

#### Vendor Contributors

> Holds the Vendor Contributor name for any expenditure transaction

```{r read_vendor_cont}
vendor_conts <- read_csv(
  file = "ok_contribs/data/raw/vendor_cont.csv",
  col_types = cols(.default = col_character())
)

vendor_conts <- vendor_conts %>% 
  select(-X3) %>% 
  clean_names() %>% 
  rename(cont_vname = cont_name)
```

### Recipients

The information on the recipients of each transaction are held in other databases.

#### Statement of Organization

The "SO-1" form applies to committees formed to support a political candidate.

```{r read_so1}
so1 <- read_csv(
  file = "ok_contribs/data/raw/so1.csv",
  col_types = cols(
    .default = col_character(),
    STRICKEN_WITHDRAWN = col_logical(),
    ORGANIZATION_DATE = col_date("%m/%d/%Y")
    )
  )

so1 <- so1 %>% 
  select(-X27) %>% 
  clean_names() %>% 
  rename(ethics_id = ethics_num)

# fix logical parse
so1$special_election <- 
  so1$special_election %>% 
  str_replace("yes", "1") %>% 
  str_replace("no", "0") %>% 
  parse_logical()
```

The "SO-2" form applies to committees formed to support non-candidate issues.

```{r read_so2}
so2 <- read_csv(
  file = "ok_contribs/data/raw/so2.csv",
  col_types = cols(
    .default = col_character(),
    ORGANIZATION_DATE = col_date("%m/%d/%Y")
  )
)

so2 <- so2 %>% 
  select(-X15) %>% 
  clean_names() %>% 
  rename(ethics_id = ethics_num)

# fix logical parse
so2$stmnt_of_intent <- 
  so2$stmnt_of_intent %>%
  recode("y" = "1", "n" = "0") %>% 
  parse_logical()
```

#### Affiliation

> Contains the banking information for the SO1 or SO2 that is filed. The rep_num refers to the
appropriate SO1 or SO2 in the Report table. The name, address, street, city, state and zip for the
bank are contained in this table.

```{r read_affiliation}
affiliations <- read_csv(
  file = "ok_contribs/data/raw/affiliation.csv",
  col_types = cols(.default = col_character())
)

affiliations <- affiliations %>% 
  select(-X8) %>% 
  clean_names()
```

#### Ballot Measures

> Holds the ballot year for the appropriate SO1 or SO2 that is filed. The rep_num refers to the
appropriate SO1 or SO2 in the Report table.

```{r read_ballot_measure}
ballot_measures <- read_csv(
  file = "ok_contribs/data/raw/ballot_measure.csv",
  col_types = cols(.default = col_character())
)

ballot_measures <- ballot_measures %>% 
  select(-X3) %>% 
  clean_names()
```

#### Depository

> Holds the full name and address for the depository information for the SO1 or SO2 from the
report_num field that relates to the report table.

```{r read_depository}
depositories <- read_csv(
  file = "ok_contribs/data/raw/depository.csv",
  col_types = cols(.default = col_character())
)

depositories <- depositories %>% 
  select(-X8) %>% 
  clean_names()
```

#### Parties

> Has the different party affiliation types

```{r read_party}
parties <- read_csv(
  file = "ok_contribs/data/raw/party.csv",
  col_types = cols(
    VIEWABLE = col_logical(),
    PARTY_ID = col_character(),
    PARTY_DESC = col_character()
  )
)

parties <- parties %>% 
  select(-X4) %>% 
  clean_names()
```

#### Offices

> Description of office types (mainly for elections)

```{r read_office}
offices <- read_csv(
  file = "ok_contribs/data/raw/office.csv",
  col_types = cols(.default = col_character())
)

offices <- offices %>% 
  select(-X3) %>% 
  clean_names() %>% 
  rename(office_id = id)
```

#### Districts

> List of the districts for elections

```{r read_district}
districts <- read_csv(
  file = "ok_contribs/data/raw/district.csv",
  col_types = cols(.default = col_character())
)

districts <- districts %>% 
  select(-X3) %>% 
  clean_names()
```

#### Candidates

> Holds the candidate name and birthdate tied to the specific ethics_num (committee)

```{r read_candidate}
candidates <- read_csv(
  file = "ok_contribs/data/raw/candidate.csv",
  col_types = cols(.default = col_character())
)

candidates <- candidates %>% 
  select(-X6, -BIRTHDATE) %>% 
  clean_names() %>% 
  rename(ethics_id = ethics_num)
```

#### Lump Funds

> Holds lump fund information for the respective report_num

```{r read_lump_funds}
lump_funds <- read_csv(
  file = "ok_contribs/data/raw/lump_fund.csv",
  col_types = cols(
    .default = col_character(),
    LUMP_AMOUNT = col_double(),
    LUMP_DATE = col_date("%d-%b-%y")
    )
)

lump_funds <- lump_funds %>% 
  select(-X8) %>% 
  clean_names()
```

### Report

> Holds all the `report_num` for all filed reports in the system from the SO1, SO2s to all the C1R,
C3R, C4R, and C5R reports. C6R reports are stored in the c6r_report table. Contains the date the
report was submitted, the `ethics_num` (committee) that it ties to, period id, the report type,
signature field, admin entered (means the report was filed by administrator), the amended reference
(if null, is the latest report, if not then that report was amended to the `report_num` that is
displayed in that field.), the final flag determines if that was the final report they will be
filing and `supp_year` is just a field on the form to show the year.

```{r read_report}
reports <- read_csv(
  file = "ok_contribs/data/raw/report.csv",
  col_types = cols(
    .default = col_character(),
    SUBMITTED_DATE = col_date("%d-%b-%y"),
    FINAL = col_logical()
  )
)

reports <- reports %>% 
  select(-X11) %>% 
  clean_names() %>% 
  rename(ethics_id = ethics_num)
```

> Description of each type of report available  (SO1, SO2, C1R, C3R, C4R, C5R, C6R)

```{r read_rep_type}
rep_types <- read_csv(
  file = "ok_contribs/data/raw/report_type.csv",
  col_types = cols(.default = col_character())
)

rep_types <- rep_types %>%
  select(-X3) %>% 
  clean_names()
```

## Join

Our primary interest is when a transaction was made, for how much, from whom, and to whom. The
transaction database contains the when and how much, but uses keys to identify the who.

The contributor of a transaction (giving money) is identified by the `cont_id` variable.

The recipient of a transaction (getting money) are the ones filing the report on which each
transaction appears, identifying by the `rep_num` variable. In the database of reports, the filer
of each report is identified with their `ethics_id`.

By joining each transaction with the filer of the respective report, we can identify the filer.

```{r join_recs}
ok <- transactions %>% 
  left_join(
    reports %>% select(rep_num, ethics_id), 
    by = "rep_num"
  )

print(ok)
```

To improve the searchability of this database of transactions, we will add the name and location
of each contributor and recipient.

### Contributors

First, we will join the `contributors` table, which contains geographic data on each contributor
(city, state, zip), which the full tables of each contributor type.

There are four types of contributors, each identified with different `cont_*name` variables:

1. Individuals with `cont_fname` (first), `cont_mname` (middle), and `cont_lname` (last)
    * With `cont_employer` and `cont_occupation`
1. Businesses with a `cont_bname`
    * With `cont_activity`
1. Committees with a `cont_cname`
    * with `cont_interest` and `ethics_id`
1. Vendors with a `cont_vname`
    * With OK Ethics Commission `ethics_id`
    
It's important to note that the transactions database contains both contributions _and_
expenditures reported by the filer. For expenditures, the "contributor" is actually the vendor
recipient of the money

```{r join_conts, collapse=TRUE}
nrow(contributors)
nrow(individual_conts)
nrow(business_conts)
nrow(committee_conts)
nrow(vendor_conts)

contributors2 <- contributors %>% 
  left_join(individual_conts, by = "cont_id") %>% 
  left_join(business_conts, by = "cont_id") %>% 
  left_join(committee_conts, by = "cont_id") %>% 
  left_join(vendor_conts, by = "cont_id") %>% 
  left_join(cont_types, by = "cont_type")

nrow(contributors2)
```

There appears to be more total contributors than listed in the contributors database. There are
`r nrow(individual_conts) - nrow(contributors)` more _individual_ contributors alone than there are
total records in the contributors database.

```{r cont_miss, collapse=TRUE}
nrow(individual_conts) - nrow(contributors)
mean(individual_conts$cont_id %in% contributors$cont_id)
```

### Recipients

When a committee is formed to receive contributions, the file a "Statement of Organization" report.
Committees formed to receive funds on behalf of a candidate file an "SO-1" form, and non-candidate
organizations file an "SO-2" form.

These forms contain a lot of information, but we will extract only the geographic information of
each, so that we can better search the contributions and expenditures in the transactions database.

First, we will create a new table of candidate committee information from the SO-1 database.

```{r can_recs}
candidate_rec <- so1 %>%
  left_join(candidates, by = "ethics_id") %>% 
  left_join(parties, by = c("party_num" = "party_id")) %>% 
  left_join(offices, by = c("office_num" = "office_id")) %>% 
  rename(
    rec_street   = street,
    rec_city     = city,
    rec_state    = state, 
    rec_zip      = zip,
    rec_cname    = comname,
    rec_party    = party_desc,
    rec_office   = office_desc
  ) %>% 
  select(ethics_id, starts_with("rec_")) %>%
  # multiple entries per ethics id
  # make all upper
  mutate_if(is_character, str_to_upper) %>% 
  # take only the first
  group_by(ethics_id) %>% 
  slice(1) %>% 
  ungroup() %>% 
  distinct()

print(candidate_rec)
```

The same can be done with non-candidate committee recipients from SO-2 filings.

```{r com_recs}
committee_rec <- so2 %>% 
  rename(
    rec_cname    = comname,
    rec_street   = street,
    rec_city     = city,
    rec_state    = state,
    rec_zip      = zip
  ) %>% 
  select(ethics_id, starts_with("rec_")) %>%
  mutate_if(is_character, str_to_upper) %>% 
  group_by(ethics_id) %>% 
  slice(1) %>% 
  ungroup() %>% 
  distinct()

print(committee_rec)
```

Combine the two types of recipients into a single table that can be joined to the transactions
database along the `ethics_id` of each transaction's report filer.

```{r bind_recs, collapse=TRUE}
recipients <- bind_rows(candidate_rec, committee_rec)
dim(recipients)
n_distinct(recipients$ethics_id) == nrow(recipients)
```

There are `r nrow(recipients)` unique committees that have filed SO-1 or S0-2 reports, each
identified by their unique `ethics_id` variable.

### Total Join

With our new tables of unique contributors and unique recipients, we can better identify the
parties to each transaction. We will join all three tables by their respective `*_id` variables.

```{r}
ok <- ok %>% 
  left_join(contributors2 %>% select(-ethics_id), by = "cont_id") %>% 
  left_join(recipients, by = "ethics_id")
```

## Explore

There are `r nrow(ok)` records of `r length(ok)` variables in the full database.

```{r dims}
nrow(ok)
length(ok)
names(ok)
sample_frac(ok)
glimpse(sample_frac(ok))
```

### Distinct

The variables range in their degree of distinctness.

The `trans_index` is `r noquote(scales::percent(n_distinct(ok$trans_index)/nrow(ok)))` distinct and
can be used to identify a unique transaction.

```{r}
ok %>% 
  map(n_distinct) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_distinct") %>% 
  mutate(prop_distinct = round(n_distinct / nrow(ok), 4)) %>%
  print(n = length(ok))
```

The `*_id` variables have as many distinct values as the length of their respective tables.

```{r distinct_id, collapse=TRUE}
n_distinct(individual_conts$cont_id)/nrow(individual_conts)
n_distinct(business_conts$cont_id)/nrow(business_conts)
```

```{r tabyls_function, echo=FALSE}
print_tabyl <- function(data, ...) {
  as_tibble(arrange(tabyl(data, ...), desc(n)))
}
```

```{r tabyls}
print_tabyl(ok, cont_desc)
print_tabyl(ok, cont_city)
print_tabyl(ok, cont_state)
print_tabyl(ok, cont_occupation)
print_tabyl(ok, cont_interest)
print_tabyl(ok, rec_city)
print_tabyl(ok, rec_state)
print_tabyl(ok, rec_party)
print_tabyl(ok, rec_office)
```

### Missing

The variables also vary in their degree of values that are `NA` (missing). 

Some variables are mutually exclusive. There cannot be, for example, both `cont_lname` and
`cont_bname` value for the same record, as these two variables identify different kinds of
transactions.

It's notable that the important variables (e.g., `*_id`, `trans_*`) contain 
mostly no missing values. There are `r sum(is.na(ok$trans_amount))` records without a 
`trans_amount`. They will be flaged with a new `na_flag` variable.

```{r missin_important, collapse=TRUE}
sum(is.na(ok$cont_id))
sum(is.na(ok$ethics_id))
sum(is.na(ok$trans_date))
sum(is.na(ok$trans_amount))
ok <- ok %>% 
  mutate(na_flag = is.na(trans_amount))
```

The full count of `NA` for each variable in the data frame can be found below:

```{r count_na}
ok %>% 
  map(function(var) sum(is.na(var))) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_na") %>% 
  mutate(prop_na = n_na / nrow(ok)) %>% 
  print(n = length(ok))
```

### Ranges

The range of continous variables will need to be checked for data integrity. There are only two
quasi-continuous variables, the `trans_amount` and `trans_date`

#### Transaction Amounts

The range for `trans_amount` seems reasonable enough.

```{r amount_range}
summary(ok$trans_amount)
```

There are only `r sum(ok$trans_amount > 500000, na.rm = T)` transactions greater than \$500,000. 

```{r plot_amt_nonlog, collapse=TRUE}
sum(ok$trans_amount > 500000, na.rm = TRUE)
ggplot(ok, aes(trans_amount)) + 
  geom_histogram() + 
  scale_y_log10() +
  scale_x_continuous(labels = scales::dollar) +
  geom_hline(yintercept = 10)
```

```{r glimpse_min_max}
glimpse(ok %>% filter(trans_amount == min(trans_amount, na.rm = T)))
glimpse(ok %>% filter(trans_amount == max(trans_amount, na.rm = T)))
```

```{r reports_min_max}
reports %>%
  filter(rep_num == ok$rep_num[which(ok$trans_amount == min(ok$trans_amount, na.rm = TRUE))])
reports %>%
  filter(rep_num == ok$rep_num[which(ok$trans_amount == max(ok$trans_amount, na.rm = TRUE))])
```

### Transaction Dates

There are 

```{r date_range}
summary(ok$trans_date)
```

### Plots

```{r plot_amt_hist}
ok %>%
  filter(rec_party %in% c("REPUBLICAN", "DEMOCRAT", "NON-PARTISAN", "INDEPENDENT")) %>% 
  ggplot(aes(trans_amount)) +
  geom_histogram(bins = 30, aes(fill = rec_party)) +
  scale_x_continuous(labels = scales::dollar, trans = "log10") +
  scale_y_log10() +
  scale_fill_manual(values = c("blue", "forestgreen", "purple", "red")) +
  facet_wrap(~rec_party) +
  theme(legend.position = "none") +
  labs(
    title = "Transaction Distribution",
    caption = "Source: OKEC",
    y = "Number of Contributions",
    x = "Amount (USD)"
  )
```

```{r plot_amt_hist}
ok %>%
  ggplot(aes(trans_amount)) +
  geom_histogram(bins = 30, aes(fill = cont_desc)) +
  scale_x_continuous(labels = scales::dollar, trans = "log10") +
  scale_y_log10() +
  facet_wrap(~cont_desc) +
  theme(legend.position = "none") +
  labs(
    title = "Transaction Distribution",
    subtitle = "by Contribution Type",
    caption = "Source: OKEC",
    y = "Number of Contributions",
    x = "Amount (USD)"
  )
```

```{r plot_amt_year}
ok %>% 
  group_by(year = year(trans_date)) %>% 
  summarise(sum = sum(trans_amount, na.rm = TRUE)) %>% 
  ggplot(aes(year, sum)) +
  geom_col() +
  labs(
    title = "Transaction Sum by Year",
    caption = "Source: OKEC",
    y = "Total Transaction Amount",
    x = "Transaction Date"
  )
```


## Clean

We can now create new variables on our comprehensive table of transactions. The objective of these
cleaning steps will be to create `*_clean` variables with a reduced number of distinct values by
fixing spelling mistakes. We will also create new variables to better identify each transaction in
the Accountability Database.

### Year

```{r mutate_year}
ok <- ok %>% 
  mutate(trans_year = year(trans_date))
```

### ZIPs

```{r clean_zipcodes}
n_distinct(ok$cont_zip)

ok <- ok %>% 
  mutate(cont_zip_clean = clean.zipcodes(cont_zip) %>% 
           na_if("00000") %>% 
           na_if("11111") %>% 
           na_if("99999") %>% 
           str_sub(1, 5)
  )

# reduced by half
n_distinct(ok$cont_zip_clean)
```

```{r clean_zipcodes}
n_distinct(ok$rec_zip)

ok <- ok %>% 
  mutate(rec_zip_clean = clean.zipcodes(rec_zip) %>% 
           na_if("00000") %>% 
           na_if("11111") %>% 
           na_if("99999") %>% 
           str_sub(1, 5)
  )

# reduced by half
n_distinct(ok$rec_zip_clean)
```

### States

```{r zip_data}
data("zipcode")
zipcode <-
  as_tibble(zipcode) %>% 
  mutate_if(is.character, str_to_upper) %>% 
  select(city, state, zip)

valid_abb <- unique(zipcode$state)
setdiff(valid_abb, state.abb)
```

```{r count_bad_states}
n_distinct(ok$cont_state)
setdiff(ok$cont_state, valid_abb)
ok %>% 
  filter(cont_state %out% valid_abb) %>% 
  print_tabyl(cont_state)
```

```{r upper_states}
ok <- ok %>% mutate(cont_state_clean = str_to_upper(cont_state))
```

```{r view_bad_states}
ok %>% 
  filter(cont_state_clean %out% valid_abb) %>% 
  filter(!is.na(cont_state_clean)) %>% 
  select(
    trans_index,
    cont_street,
    cont_city,
    cont_state_clean,
    cont_zip_clean
  ) %>% 
  left_join(
    zipcode %>% select(city, state, zip), 
    by = c("cont_zip_clean" = "zip")
  )
```

```{r fix_bad_states, collapse=TRUE}
ok$cont_state_clean <- 
  ok$cont_state_clean %>% 
  str_replace("^K$",  "OK") %>% 
  str_replace("^OL$", "OK") %>% 
  str_replace("^FO$", "OK") %>% 
  # address matches tulsa
  # zip matches new york
  str_replace("^74$", "OK") %>% 
  na_if("NG")

# all fixed
setdiff(ok$cont_state_clean, valid_abb)
```

```{r count_bad_states, collapse=TRUE}
n_distinct(ok$rec_state)
setdiff(ok$rec_state, valid_abb)
```

### Cities

```{r valid_city}
valid_city <- unique(zipcode$city)
length(valid_city)

valid_ok_city <- unique(zipcode$city[zipcode$state == "OK"])
length(valid_ok_city)
```

```{r count_bad_city, collapse=TRUE}
n_distinct(ok$cont_city) # 7424
length(setdiff(ok$cont_city, valid_city))
sum(ok$cont_city %out% valid_city)
mean(ok$cont_city %out% valid_city)
```

```{r}
ok %>% 
  filter(cont_city_clean %out% valid_city) %>% 
  count(cont_city_clean) %>% 
  arrange(desc(n))
```

```{r}
ok$cont_city_clean <- 
  str_to_upper(ok$cont_city) %>%
  str_remove_all("[:punct:]") %>%
  str_remove_all("\\d+") %>% 
  str_remove_all("(^|\\b)OK(\\b|$)") %>% 
  str_remove_all("(^|\\b)DC(\\b|$)") %>%
  str_replace_all("\\s+", "\\s") %>% 
  str_trim() %>% 
  # make common NA
  na_if("N/A") %>% 
  na_if("NA") %>% 
  na_if("N A") %>% 
  na_if("N.A") %>% 
  na_if("NONE") %>% 
  na_if("NONR") %>% 
  na_if("NON") %>% 
  na_if("NONE GIVEN") %>% 
  na_if("NOT GIVEN") %>%
  na_if("NOT GIVE") %>% 
  na_if("NOT REQUIRED") %>%
  na_if("INFORMATION REQUESTED") %>% 
  na_if("REQUESTED INFORMATION") %>% 
  na_if("INFO REQUESTED") %>% 
  na_if("REQUESTED INFO") %>% 
  na_if("UNKOWN") %>%
  na_if("UNKNOWN") %>% 
  na_if("NOTAPPLICABLE") %>% 
  na_if("NOT APPLICABLE") %>% 
  na_if("VARIOUS") %>% 
  na_if("UNDER $1000") %>% 
  na_if("ANYWHERE") %>% 
  na_if("X") %>% 
  na_if("XX") %>% 
  na_if("XXX") %>% 
  na_if("XXXX") %>% 
  na_if("XXXXX") %>% 
  na_if("XXXXXXX") %>% 
  na_if("XXXXXXXX") %>% 
  na_if("-") %>% 
  na_if("")

ok$cont_city_clean <- 
  ok$cont_city_clean %>%
  # very frequent city abbs
  str_replace_all("^OKC$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLA CITY$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLAHOMA$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OK CITY$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLAHOM CITY$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLAHMA CITY$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLAHOMA CITH$", "OKLAHOMA CITY") %>% 
  str_replace_all("^OKLAHOMA CIT$", "OKLAHOMA CITY") %>% 
  # location type abbs
  str_replace_all("^FT\\s", "FORT") %>% 
  str_replace_all("^ST\\s", "SAINT") %>% 
  str_replace_all("^MT\\s", "MOUNT")

n_distinct(ok$cont_city)
n_distinct(ok$cont_city_clean)
```

```{r}
ok_city_fix <- ok %>%
  filter(!is.na(cont_city)) %>% 
  mutate(cont_city_fix = cont_city_clean %>%
           key_collision_merge(dict = valid_ok_city) %>% 
           n_gram_merge()
  ) %>% 
  mutate(fixed = (cont_city_fix != cont_city_clean))

sum(x$fixed, na.rm = TRUE)
tabyl(x$fixed)

ok_city_fix %>% 
  filter(fixed) %>%
  count(cont_state, cont_city_clean, cont_city_fix) %>% 
  arrange(desc(n))
```

```{r}
ok_city_fix <- ok_city_fix %>%
  filter(fixed) %>% 
  select(
    cont_city,
    cont_city_clean,
    cont_city_fix,
    cont_state_clean,
    cont_zip_clean
  ) %>% 
  distinct()

n_distinct(x$cont_city)
n_distinct(x$cont_city_clean)
n_distinct(x$cont_city_fix)
```

```{r}
good_fix <- ok_city_fix %>% 
  inner_join(
    zipcode,
    by = c(
      "cont_state_clean" = "state",
      "cont_zip_clean" = "zip",
      "cont_city_fix" = "city"
    )
  )

nrow(good_fix)

bad_fix <- ok_city_fix %>% 
  anti_join(
    zipcode,
    by = c(
      "cont_state_clean" = "state",
      "cont_zip_clean" = "zip",
      "cont_city_fix" = "city"
    )
  )

nrow(bad_fix)
```

