---
title: "United States Health Center Funding Diary"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "covid", "hrsa_cares", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

> The Health Resources and Services Administration [HRSA] is an agency of the U.S.
Department of Health and Human Services located in North Bethesda, Maryland. It
is the primary federal agency for improving access to health care services for
people who are uninsured, isolated or medically vulnerable.

Per the [HRSA supplemental funding FAQ page][faq]: 

> On Wednesday, April 8, HRSA announced the release of more than $1.3 billion in
funding provided by the Coronavirus Aid, Relief, and Economic Security (CARES)
Act (PDF - 696 KB) (activity code H8D). For additional information, see the
CARES technical assistance webpage.

[faq]: https://bphc.hrsa.gov/program-opportunities/coronavirus-disease-2019/faq

## Download

The HRSA website lists all the health center grantees by state on individual 
web pages. We can loop through each state and scrape the HTML table and save it
as a local text file.

```{r raw_dir}
raw_dir <- dir_create(here("us", "covid", "hrsa_cares", "data", "raw"))
```

```{r raw_download}
x <- "https://bphc.hrsa.gov/emergency-response/coronavirus-cares-FY2020-awards/"
for (s in valid_abb) {
  st_url <- str_c(x, str_to_lower(s))
  st_path <- path(raw_dir, path_ext_set(s, "csv"))
  if (file_exists(st_path)) {
    next()
  } else {
    st_get <- GET(st_url)
    if (status_code(st_get) == 200) {
      content(x = st_get) %>% 
        html_node(css = "table") %>% 
        html_table(header = TRUE) %>% 
        write_csv(path = st_path)
    }
  }
}
```

```{r raw_paths}
raw_paths <- dir_ls(raw_dir)
```

## Read

The `r length(raw_paths)` text files can be read into a single data frame.

```{r raw_read}
hrsa <- vroom(
  file = raw_paths,
  delim = ",",
  .name_repair = make_clean_names,
  col_types = cols(
    `HEALTH CENTER GRANTEE` = col_character(),
    `CITY` = col_character(),
    `STATE` = col_character(),
    `FUNDING AMOUNT` = col_number()
  )
)
```

## Explore

```{r glimpse}
glimpse(hrsa)
tail(hrsa)
```

### Missing

There are no missing values.

```{r na_count}
col_stats(hrsa, count_na)
```

### Duplicates

There are no duplicate records.

```{r dupe_count}
sum(duplicated(hrsa))
```

### Geographical

```{r state_count}
count(hrsa, state, sort = TRUE)
```

```{r city_normal}
hrsa <- mutate(hrsa, across(city, normal_city, abbs = usps_city))
```

```{r city_valid}
many_city <- c(valid_city, extra_city)
percent(prop_in(hrsa$city, many_city), 0.01)
```

```{r}
hrsa %>% 
  filter(city %out% many_city) %>% 
  count(city, state, sort = TRUE)
```

### Amounts

```{r ammount_summary}
noquote(map_chr(summary(hrsa$funding_amount), dollar))
```

```{r hist_amount, echo=FALSE}
hrsa %>%
  ggplot(aes(funding_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "United States Health Center Funding Amount Distribution",
    x = "Amount",
    y = "Count"
  )
```

### Add

We can add funding agency, date, and year manually.

```{r date_year}
hrsa <- mutate(
  .data = hrsa,
  .before = 1,
  date = mdy("04082020"), 
  year = year(date),
  agency = "Health Resources and Services Administration",
  govt = "US"
)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(hrsa, 50))
```

1. There are `r comma(nrow(hrsa))` records in the database.
1. There are `r comma(sum(hrsa$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(hrsa$na_flag))` records missing key variables.
1. The 4-digit `year` variable has been created manually.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "covid", "hrsa_cares", "data", "clean"))
clean_path <- path(clean_dir, "us_hrsa_cares.csv")
write_csv(hrsa, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(hrsa)),
  type = md_code(map_chr(hrsa, typeof)),
  def = c(
    "Date funding released",
    "Year funding released (2020)",
    "Distributing agency name",
    "Agency government abbreviation",
    "Health center grantee name",
    "Grantee city name",
    "Grantee state abbreviation",
    "CARES Act funding amount"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
