---
title: "Indiana Lobbying Expenditure Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("co", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://data.colorado.gov/Lobbyist/Directory-of-Lobbyist-Clients-in-Colorado/35k5-cv8s

## Data

Lobbyist data is obtained from the [Colorado Open Data Portal][03]. The data is as current as March 11, 2020.

> About:   
Lobbyist name and address and the names and addresses of their associated clients provided by the Colorado Department of State (CDOS).

```{r raw_dir}
raw_dir <- dir_create(here("in", "lobby", "data", "raw", "exp"))
```


```{r download, eval=FALSE}
landing_url <- 'https://www.in.gov/ilrc/2335.htm'

urls <- GET(landing_url) %>% content() %>% html_nodes("a") %>% html_attr("href")

comp_urls <- paste0('https://www.in.gov',urls[str_detect(urls, 'Compensated Lobbyist Total')] %>% html_attr('href'))

wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose",
      "--content-disposition",
      url,
      paste("-P", raw_dir)
    )
  )
}

if (!all_files_new(raw_dir)) {
  map(comp_urls, wget, raw_dir)
}

```

These files come in various formats, and we will need to parse them differently according to their file extensions.
```{r read csv}
inle_csv <- read_csv(dir_ls(raw_dir) %>% str_subset('csv'), 
                 col_types = cols(.default = col_character())) %>% 
  clean_names()
in_excel <- map_dfr(dir_ls(raw_dir) %>% 
                               str_subset('xl'), read_excel) %>% 
                      clean_names()
```

We'll do some basic cleaning by turning all text columns to uppercase and transform column names to snake case.
```{r clean}
colr <- colr %>% 
  mutate_if(is.character, str_to_upper) %>% 
  clean_names()
```

```{r examine}
head(colr)
tail(colr)
glimpse(sample_n(colr, 20))
```

According to the [data portal](https://data.colorado.gov/Lobbyist/Directory-of-Lobbyist-Clients-in-Colorado/35k5-cv8s): 

> All of the lobbyists and all of the clients for Colorado, except for state lobbyists, also known as official State Liaisons, who are responsible for lobbying on behalf of a state agency and therefore have no "client". The primaryLobbyistID field is the number assigned when the lobbyist registers with the state, and the first four numbers are the year of original registration. The annual lobbyist registration identification number is updated when they register in a current fiscal year, and the first four numbers reflect status of the lobbyist when registering in a current fiscal year (for example, if they were registered in 2015, but did not update their registration in 2016, then the registration id would remain 2015 until they register in a new fiscal year). The data is updated as necessary to the main lobbyist online system via manual entry, however the updates to the Colorado Information Marketplace is only done weekly, every Monday night.

### Year
We can add a year based on the first four digit of `annual_lobbyist_registration_id`.
```{r add year}
colr <- colr %>% 
  mutate(year = str_sub(annual_lobbyist_registration_id, start = 1L, end = 4L),
         first_year = str_sub(primary_lobbyist_id, start = 1L, end = 4L))
```

### Missing

```{r na_count}
col_stats(colr, count_na)
```

```{r na_flag}
colr <- colr %>% flag_na(lobbyist_last_name, client_name)
sum(colr$na_flag)
```

### Duplicates
We can see there's no duplicate entry.
```{r dupe_flag}
colr <- flag_dupes(colr, dplyr::everything())
sum(colr$dupe_flag)
```
### Categorical

#### Year
We can see that the data is pretty up to date. However, there're some fileds indicating the annual registration was recorded in `2055`, which is presumably a human error. 
```{r check year}
colr <- colr %>% 
  mutate(year = as.numeric(year))
tabyl(colr$year)
```
### Wrangle
To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
colr <- colr %>% 
    # combine street addr
  unite(
    col = lobbyist_address,
    starts_with("lobbyist_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
    unite(
    col = client_address,
    starts_with("client_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate_at(.vars = vars(ends_with('address')), .funs = list(norm = ~ normal_address(.,abbs = usps_street,
      na_rep = TRUE))) %>% 
  select(-ends_with("address"))
```

```{r address_view}
colr %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10) %>% 
  glimpse()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
colr <- colr %>% 
    mutate_at(.vars = vars(ends_with('zip')), 
              .funs = list(norm = ~ normal_zip(.))) %>% 
  rename(lobbyist_zip5 = lobbyist_zip_norm,
         client_zip5 = client_zip_norm)
```

```{r zip_progress}
progress_table(
  colr$lobbyist_zip,
  colr$lobbyist_zip5,
  colr$client_zip,
  colr$client_zip5,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
prop_in(colr$lobbyist_state, valid_state, na.rm = T)
prop_in(colr$client_state, valid_state, na.rm = T)
colr <- colr %>% 
   mutate(client_state_norm = normal_state(client_state,abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state))
```

```{r state_view}
colr %>% 
  filter(client_state != client_state_norm) %>% 
  count(client_state, sort = TRUE)
```

```{r state_progress}
progress_table(
  colr$client_state,
  colr$client_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good colrart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
colr <- colr %>% 
      mutate_at(.vars = vars(ends_with('city')), .funs = list(norm = ~ normal_city(.,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE)))

prop_in(colr$lobbyist_city_norm, valid_city, na.rm = T)
prop_in(colr$client_city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
colr <- colr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lobbyist_state" = "state",
      "lobbyist_zip5" = "zip"
    )
  ) %>% 
  rename(lobbyist_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lobbyist_city_norm, lobbyist_city_match),
    match_dist = str_dist(lobbyist_city_norm, lobbyist_city_match),
    lobbyist_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = lobbyist_city_match,
      false = lobbyist_city_norm
    )
  ) %>% 
  select(
    -lobbyist_city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap clients}
colr <- colr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "client_state_norm" = "state",
      "client_zip5" = "zip"
    )
  ) %>% 
  rename(client_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(client_city_norm, client_city_match),
    match_dist = str_dist(client_city_norm, client_city_match),
    client_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = client_city_match,
      false = client_city_norm
    )
  ) %>% 
  select(
    -client_city_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  colr$lobbyist_city,
  colr$client_city,
  colr$lobbyist_city_norm,
  colr$client_city_norm,
  colr$lobbyist_city_swap,
colr$client_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "NY City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(colr, 20))
```

1. There are `r nrow(colr)` records in the database.
1. There are `r sum(colr$dupe_flag)` duplicate records in the database.
1. The range and distribution of `year` seems mostly reasonable except for a few entries.
1. There are `r sum(colr$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("co", "lobby", "data", "exp","clean"))
```

```{r write_clean}
write_csv(
  x = inle %>% rename(lobbyist_city_clean = lobbyist_city_swap) %>% rename(client_city_clean = client_city_swap),
  path = path(clean_dir, "co_lob_summary_clean.csv"),
  na = ""
)
```

