---
title: "District Of Columbia Contributions"
author: "Kiernan Nicholls & Aarushi Sahejpal"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("dc", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  aws.s3, # upload to aws s3
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
setwd("/Volumes/TAP/accountability_datacleaning/state")
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data comes courtesy of the Washington, [DC Office of Campaign Finance (OCF)][03].

As the [OCF website][04] explains: 

> The Office of Campaign Finance (OCF) provides easy access to all contributions
and expenditures reported from 2003, through the current reporting period.
Because the system is updated on a daily basis, you may be able to retrieve data
received by OCF after the latest reporting period. This data is as reported, but
may not be complete.

[03]: https://ocf.dc.gov/
[04]: https://ocf.dc.gov/service/view-contributions-expenditures

The data is found on the dc.gov [OpenData website][05]. The file abstract reads:

> The Office of Campaign Finance (OCF) is pleased to publicly share election
campaign contribution data. The Campaign Finance Office is charged with
administering and enforcing the District of Columbia laws pertaining to campaign
finance operations, lobbying activities, conflict of interest matters, the
ethical conduct of public officials, and constituent service and statehood fund
programs. OCF provides easy access to all contributions and expenditures
reported from 2003, through the current reporting period. Because the system is
updated on a daily basis, you may be able to retrieve data received by OCF after
the latest reporting period. This data is as reported, but may not be complete.
Visit the http://ocf.dc.gov for more information.

[05]: https://opendata.dc.gov/datasets/DCGIS::campaign-financial-contributions/explore
[06]: http://geospatial.dcgis.dc.gov/ocf/

## Download

Contributions should be downloaded by hand to the raw directory.

```{r raw_dir}
raw_dir <- dir_create(here("dc", "contribs", "data", "raw"))
```

Contributions are retrieved by filter type drop down menu.

```{r raw_options, results='asis'}
read_html("https://efiling.ocf.dc.gov/ContributionExpenditure") %>% 
  html_nodes("#FilerTypeId> option") %>%
  html_text() %>% 
  extract(-1) %>% 
  str_trim() %>% 
  md_bullet()
```

## Read

```{r raw_paths, results='asis'}
raw_paths <- dir_ls(raw_dir)
md_bullet(md_code(path.abbrev(raw_paths)))
```

These files have a troublesome encoding. We can read and re-write them.

```{r}
library(readr)

dcc <- read_csv("/Volumes/TAP/accountability_datacleaning/dc/contribs/data/raw/Campaign_Financial_Contributions.csv")
```


----
### exclude
{r raw_convert, eval=FALSE}
for (p in raw_paths) {
  write_csv(
    path = p,
    x = read.csv(
      file = p, skip = 1,
      fileEncoding = "UTF-16LE", 
    )
  )
}


The new files can be read into a single data frame.

dcc <- map_df(
  .x = raw_paths,
  .f = read_csv,
  skip = 1,
  col_names = read_names(raw_paths[2]),
  col_types = cols(
    .default = col_character(),
    Receipt.Date = col_date_usa(),
    Amount = col_number()
  )
)

----



```{r raw_rename}
dcc <- dcc %>% 
  clean_names("snake") 

```

```{r}
dcc <- dcc %>% 
  rename(
    committee = committeename,
    contributor = contributorname,
    address = address,
    type = contributortype,
    method = contributiontype,
    employer = employer,
    date = dateofreceipt)
```


## Explore

There are `r comma(nrow(dcc))` rows of `r ncol(dcc)` columns.

```{r glimpse}
glimpse(dcc)
tail(dcc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(dcc, count_na)
```

We can then flag any records missing a key variable like a name or number.

```{r na_flag}
key_vars <- c("date", "contributor", "amount", "committee")
dcc <- flag_na(dcc, all_of(key_vars))
sum(dcc$na_flag)
```

```{r na_view}
dcc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```


### Categorical

```{r distinct_count}
col_stats(dcc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(dcc, type)
explore_plot(dcc, method)
```

### Amounts

```{r ammount_summary}
summary(dcc$amount)
mean(dcc$amount <= 0)
```

```{r hist_amount, echo=FALSE}
dcc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "District Of Columbia Contracts Amount Distribution",
    caption = "Source: DC OCF",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
dcc <- mutate(dcc, year = year(date))
dcc <- dcc %>%
  mutate(date = as.Date(date, format = "%Y/%m/%d %H:%M:%S+00")) %>%
  mutate(date = format(date, "%Y-%m-%d"))
```

```{r date_range}
min(dcc$date)
sum(dcc$year < 2000)
max(dcc$date)
sum(dcc$date > today())
```

```{r bar_year, echo=FALSE}
dcc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2023, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "District Of Columbia Expenditures by Year",
    caption = "Source: DC OCF",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
dcc <- dcc %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
dcc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```


### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_glimpse}
glimpse(sample_n(dcc, 50))
```

1. There are `r comma(nrow(dcc))` records in the database.
1. There are `r comma(sum(dcc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dcc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("dc", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "dc_contribs_20230517.csv")
write_csv(dcc, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
s3_head <- head_object(s3_path, "publicaccountability")
(s3_size <- as_fs_bytes(attr(s3_head, "content-length")))
unname(s3_size == clean_size)
```


