---
title: "Colorado Lobbying Expenditure Data Diary"
author: "Yanqi Xu"  
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("co", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.


## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://data.colorado.gov/Lobbyist/Directory-of-Lobbyists-in-Colorado/bqa5-gr84


## Data

Lobbyist data is obtained from the [Colorado Open Data Portal][03]. The data is as current as March 11, 2020.

> About:   
Information for each registered lobbyist, including contact details, and their associated income and expenses as summarized by month and associated report date for the State of Colorado dating back to 1995 provided by the Colorado Department of State (CDOS).


## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("co", "lobby", "data", "raw", "exp"))
```
### Download from web
```{r download, eval=FALSE}
summary_url <- 'https://data.colorado.gov/api/views/bqa5-gr84/rows.csv?accessType=DOWNLOAD'
exp_url <- 'https://data.colorado.gov/api/views/eqsm-7ah7/rows.csv?accessType=DOWNLOAD'
  
urls <- c(summary_url, exp_url)
wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose",
      "--content-disposition",
      url,
      paste("-P", raw_dir)
    )
  )
}

if (!all_files_new(raw_dir)) {
  map(urls, wget, raw_dir)
}
```

### Read
> Description:  
Expenses and income summarized by month associated with lobbying activities. In a reporting month, income is reported based on the date that it is actually received. For example, in the event that a payment is made cumulatively over a three month period, they must report the full three month payment when the payment was received. If a lobbyist has multiple payments received by various clients, each dollar amount is reported separately by each client. If a lobbyist receives multiple payments on multiple bills from a single client, the monthly income report will be the total of what the client is paying the lobbyist for all of lobbying activities on the multiple bills. The statute specifies that lobbyists report the total amount they receive from the client, so in some cases the monthly dollar amount reported by a lobbyist is the total amount received may reflect a payment paid for a number of bills and activities. The payments are not itemized by bills or activities. In this dataset income is summarized by month.

> Each row represents: 	Lobbyist name, id, address, and income and expense by month

According to Colorado data portal, this dataset can be joined to: 
> Bill Information and Position with Income of Lobbyist in Colorado, Characterization of Lobbyist Clients in Colorado, Directory of Lobbyist Clients in Colorado, Expenses for Lobbyists in Colorado, Subcontractors for Lobbyists in Colorado based on "primaryLobbyistId"

```{r raw_read}
cole <- read_csv(dir_ls(raw_dir, regexp = "Directory.+"), 
                 col_types = cols(.default = col_character())) %>% 
  clean_names()

co_exp <- read_csv(dir_ls(raw_dir, regexp = "Expenses.+"), 
                 col_types = cols(.default = col_character())) %>% 
  clean_names()
```

## Explore

```{r glimpse}
head(cole)
tail(cole)

head(co_exp)
tail(co_exp)
```


### Missing
All records seem to be pretty complete. 
```{r na_count}
col_stats(cole, count_na)
col_stats(co_exp, count_na)
```

We'll flag records without lobbyists' addresses.

```{r flag na}
cole <- cole %>% 
  flag_na(lobbyist_address1)

co_exp <- co_exp %>% 
  flag_na(lobbyist_address1, expenditure_amount)
```


### Duplicates
There isn't any duplicate column. 
```{r dupe_flag}
cole <- flag_dupes(cole, dplyr::everything())
sum(cole$dupe_flag)

co_exp <- flag_dupes(co_exp, dplyr::everything())
sum(co_exp$dupe_flag)
```

### Categorical

#### Dates
Since the dates are all read as characters, we will convert them back in to date objects. We can add a year variable to the dataframe based on the registration date.

After examining the results, we can clearly see that there're some human errors when entering the date. 

```{r date_range}
cole <- cole%>% 
   mutate_at(.vars = vars(ends_with('date')), as.Date, format = "%m/%d/%Y %H:%M:%S %p") %>% 
   mutate(year = year(report_due_date))

co_exp <- co_exp%>% 
   mutate_at(.vars = vars(ends_with('date')), as.Date, format = "%m/%d/%Y %H:%M:%S %p") %>% 
   mutate(year = year(report_due_date))

min(co_exp$expenditure_receipt_date, na.rm = T)
max(co_exp$expenditure_receipt_date, na.rm = T)
```

#### Year
We can see the distribution of the `year` variable as such.
```{r year distribute}
tabyl(co_exp$fiscal_year)
```


```{r year count, echo=FALSE}
cole %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year < 2021 & year > 1994) %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Colorado Lobbyist Expenses by year",
    caption = "Source: Colorado Secretary of State via data.colorado.gov",
    fill = "Election year",
    x = "year",
    y = "Count"
  )
```


### Continuous
We can examine the amounts in both the summary and the expenditure database
```{r cont}
cole <- cole %>% 
  mutate_at(.vars = vars(starts_with('total_monthly')), .funs = as.numeric)

co_exp <- co_exp %>% 
  mutate(expenditure_amount = as.numeric(expenditure_amount))

summary(cole$total_monthly_income)
summary(cole$total_monthly_expenses)

summary(co_exp$expenditure_amount)
```


### Wrangle
To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
cole <- cole %>% 
    # combine street addr
  unite(
    col = lobbyist_address,
    starts_with("lobbyist_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate(lobbyist_address_norm = normal_address(lobbyist_address,abbs = usps_street,
      na_rep = TRUE)) %>% 
  select(-ends_with("address"))

co_exp <- co_exp %>% 
    # combine street addr
  unite(
    col = lobbyist_address,
    starts_with("lobbyist_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate(lobbyist_address_norm = normal_address(lobbyist_address,abbs = usps_street,
      na_rep = TRUE)) %>% 
  select(-ends_with("address"))
```

```{r address_view}
cole %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10) %>% 
  glimpse()

co_exp %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10) %>% 
  glimpse()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
cole <- cole %>% 
    mutate(lobbyist_zip_norm = normal_zip(lobbyist_zip, na_rep = T)) %>% 
  rename(lobbyist_zip5 = lobbyist_zip_norm)

co_exp <- co_exp %>% 
    mutate(lobbyist_zip_norm = normal_zip(lobbyist_zip, na_rep = T)) %>% 
  rename(lobbyist_zip5 = lobbyist_zip_norm)
```

```{r zip_progress}
progress_table(
  cole$lobbyist_zip,
  cole$lobbyist_zip5,
  compare = valid_zip
)

progress_table(
  co_exp$lobbyist_zip,
  co_exp$lobbyist_zip5,
  compare = valid_zip
)
```

### State

By examining the percentage of lobbyist_state that are considered valid, we can see that the variable in both datasets doesn't need to be normalized.

```{r state_norm}
prop_in(cole$lobbyist_state, valid_state, na.rm = T)
prop_in(co_exp$lobbyist_state, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good coleart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
cole <- cole %>% 
      mutate(lobbyist_city_norm = normal_city(lobbyist_city,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE))

prop_in(cole$lobbyist_city_norm, valid_city, na.rm = T)

co_exp <- co_exp %>% 
      mutate(lobbyist_city_norm = normal_city(lobbyist_city,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE))

prop_in(co_exp$lobbyist_city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
cole <- cole %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lobbyist_state" = "state",
      "lobbyist_zip5" = "zip"
    )
  ) %>% 
  rename(lobbyist_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lobbyist_city_norm, lobbyist_city_match),
    match_dist = str_dist(lobbyist_city_norm, lobbyist_city_match),
    lobbyist_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = lobbyist_city_match,
      false = lobbyist_city_norm
    )
  ) %>% 
  select(
    -lobbyist_city_match,
    -match_dist,
    -match_abb
  )

co_exp <- co_exp %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lobbyist_state" = "state",
      "lobbyist_zip5" = "zip"
    )
  ) %>% 
  rename(lobbyist_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lobbyist_city_norm, lobbyist_city_match),
    match_dist = str_dist(lobbyist_city_norm, lobbyist_city_match),
    lobbyist_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = lobbyist_city_match,
      false = lobbyist_city_norm
    )
  ) %>% 
  select(
    -lobbyist_city_match,
    -match_dist,
    -match_abb
  )
```


After the two normalization steps, the percentage of valid cities is close to 100% for both datasets. 

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  cole$lobbyist_city,
  cole$lobbyist_city_norm,
  cole$lobbyist_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))

progress_exp <- progress_table(
  co_exp$lobbyist_city,
  co_exp$lobbyist_city_norm,
  co_exp$lobbyist_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```


```{r progress_print, echo=FALSE}
kable(progress, digits = 3, caption = 'CO Lobbyists Summary Data Normalization Progress')
kable(progress_exp, digits = 3, caption = 'CO Lobbyists Expenditure Data Normalization Progress')
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Colorado Lobbyists Summary Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )


progress_exp %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Colorado Lobbyists' Expenditures Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.


## Conclude

```{r clean_glimpse}
glimpse(sample_n(cole, 20))
glimpse(sample_n(co_exp, 20))
```

1. There are `r nrow(cole)` records in the summary database and `r nrow(co_exp)` in the expenditure database.
1. There're `r sum(cole$dupe_flag)` duplicate records in the summary database and `r sum(co_exp$dupe_flag)` in the expenditure database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(co_exp$na_flag)` records missing either address or expenditure amount.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("co", "lobby", "data", "exp","clean"))
```

```{r write_clean}
write_csv(
  x = cole %>% rename( lobbyist_city_clean = lobbyist_city_swap),
  path = path(clean_dir, "co_lob_summary_clean.csv"),
  na = ""
)


write_csv(
  x = co_exp %>% rename(lobbyist_city_clean = lobbyist_city_swap),
  path = path(clean_dir, "co_lob_exp_clean.csv"),
  na = ""
)
```
