---
title: "Rhode Island Lobbying Registration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ri", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit zip Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.


## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The [Rhode Island Secretary of State's office](https://www.sos.ri.gov/divisions/Open-Government/Transparency/lobbying) makes the data available in web pages, but does not provide bulk download options. The workshop filed a open records request and obtained lobbying registration and expenditure data. The record layouts for both datasets are slightly different. The request was filled on Feb. 27, 2020.

## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("ri", "lobby", "data", "raw", "reg"))
```

### Read
We can see that the date is formatted in Excel and we need to use `excel_numeric_to_date`
```{r raw_read}
rilr_new <- read_xlsx(
  path = dir_ls(raw_dir) %>% str_subset("19")
) %>% clean_names()

rilr_old <- read_xlsx(
  path = dir_ls(raw_dir) %>% str_subset("16")
) %>% clean_names()

```

## Explore

```{r glimpse}
head(rilr_new)
tail(rilr_new)
glimpse(sample_n(rilr_new, 20))

head(rilr_old)
tail(rilr_old)
glimpse(sample_n(rilr_old, 20))
```

### Missing
The important fields are almost complete. 
```{r na_count}
col_stats(rilr_new, count_na)
col_stats(rilr_old, count_na)
```


### Duplicates
There're some duplicate records in the new dataset, and we will flag them with `dupe_fla`. Note that if two entries are identical, both entries are flagged.
```{r dupe_flag}
rilr_new <- flag_dupes(rilr_new, dplyr::everything())
sum(rilr_new$dupe_flag)

rilr_old <- flag_dupes(rilr_old, dplyr::everything())
```

### Categorical

```{r n_dirilr_newinct}
col_stats(rilr_new, n_distinct)
col_stats(rilr_old, n_distinct)
```

Visualize number of lobbyists registered each year. 
First, we will create a 

```{r create year}
rilr_old <- rilr_old %>% 
  mutate(year = year(relationship_date))
rilr_old$year %>% tabyl()

rilr_new$year %>% tabyl()
```
We can see that most of the data is from 2019.


#### Dates

Data from 2016 onward doesn't have dates.

```{r date_range}
min(rilr_old$relationship_date)
max(rilr_old$relationship_date)

min(rilr_old$)
max(rilr_old$)
```

```{r year_bar}
rilr_new %>% 
  count(year) %>% 
  rbind(rilr_old %>% count(year)) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Rhode Island Lobbyists by Year",
    caption = "Source: Rhode Island Secretary of State: Lobbyists",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
zip codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.

We will separate the compensation into a numeric field and a text field. 


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
rilr_new <- rilr_new %>% 
  unite(starts_with("address"),col = "address_full", remove = F) %>% 
  # normalize combined addr
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    ))
```
The old file didn't have any address fields.
```{r address_view}
rilr_new %>% 
  select(contains("ADDR")) %>% 
  distinct() %>% 
  sample_n(10)
```

### zip

For zip codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the zip+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
rilr_new <- rilr_new %>% 
  mutate(
    zip5 = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  rilr_new$zip,
  rilr_new$zip5,
  compare = valid_zip
)
```

### state

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.
The state field is almost clean
```{r state_norm}
prop_in(rilr_new$state, valid_state)
```

```{r firm_state_view}
rilr_new %>% 
  filter(FIRM_STATE != FIRM_STATE_norm) %>% 
  count(FIRM_STATE, sort = TRUE)
```

```{r state_progress}
progress_table(
  rilr_new$STATE,
  rilr_new$STATE_norm,
  compare = valid_state
)
```

### city

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good rilr_newart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
rilr_new <- rilr_new %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and zip code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
rilr_new <- rilr_new %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist) == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r firm_city_swap}
rilr_new <- rilr_new %>% 
  left_join(
    y = zipcodes,
    by = c(
      "FIRM_STATE_norm" = "state",
      "FIRM_zip5" = "zip"
    )
  ) %>% 
  rename(FIRM_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(FIRM_city_norm, FIRM_city_match),
    match_dist = str_dist(FIRM_city_norm, FIRM_city_match),
    FIRM_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = FIRM_city_match,
      false =FIRM_city_norm
    )
  ) %>% 
  select(
    -FIRM_city_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  rilr_new$city,
  rilr_new$FIRM_city,
  rilr_new$city_norm,
  rilr_new$FIRM_city_norm,
  rilr_new$city_swap,
rilr_new$FIRM_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Rhode Island city Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "RI city Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(rilr_new, 20))
```

1. There are `r nrow(rilr_new)` records in the database.
1. There are `r sum(rilr_new$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(rilr_new$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `YEAR` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("ri", "lobby", "data", "reg","clean"))
```

```{r write_clean}
write_csv(
  x = rilr_new %>% rename(city_clean = city_swap) %>% rename( FIRM_city_clean = FIRM_city_swap),
  path = path(clean_dir, "rilr_new_clean.csv"),
  na = ""
)
```
