---
title: "Georgia Lobbyying Registration"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ga", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  readxl, #read excel files
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Lobbying data is obtained from the [Georgia Government Transparency and Campaign Finance Commission][ethics].
We use the Lobbyist Roster under the **Search Lobbyist** feature to return a list of lobbyists. There're two options for the data output, organized by group or lobbyist. Many records have matches in the other data set, so we'll need to join the two datasets for the addresses of both lobbyists and groups they're associated with. We will use the Lobbyist Group Roster so that the data is organized by clients in relation to the lobbyists they employed.

Note that the data is separated by year, and we will use `httr` to obtain each year's data via a `POST` request, whose parameters can be found from network activities.

## Download
We can set up the raw directory.
```{r}
raw_dir <- dir_create(here("ga", "lobby", "data", "raw","reg"))
raw_lob_paths <- path(raw_dir, glue("ga_lob_{2004:2020}.xls"))
raw_grp_paths <- path(raw_dir, glue("ga_grp_{2004:2020}.xls"))
```
The curl command is preprocessed to include `listGroup={c(LGR,LR)}` and `listYear={2004:2020}`
```{r download, eval=FALSE}
ga_lob_curls <- glue(read_file(path(raw_dir,"ga_lob_curl.sh")))
ga_grp_curls <- glue(read_file(path(raw_dir,"ga_grp_curl.sh")))

ga_lob_download <- function(ga_curl,curl_type) { 
  #listGroup=LGR for lobbyist roster, and listGroup=LR for group roster
  #curl_type <- ifelse(test = str_detect(ga_curl,"listGroup=LGR"),yes = "lob", no = "grp")
  this_year <- ga_curl %>% str_extract("(?<=listYear=)\\d{4}")
  raw_path <- path(raw_dir, glue("ga_{curl_type}_{this_year}.xls"))
  if (!this_file_new(raw_path)) {
  write_lines(
  # x = system(paste("bash", path(raw_dir,"ga_curl.sh")), intern = TRUE),
  x = system(ga_curl, intern = TRUE),
  path = raw_path
)
  }
}

map2(ga_lob_curls,"lob",ga_lob_download)
map2(ga_grp_curls,"grp",ga_lob_download)

### the 2001-2003 xls file can be accessed via a GET request
download.file("https://media.ethics.ga.gov/search/Lobbyist/Exp/Ros_2003_2001.xls", path(raw_dir,"ga_lob_0103.xls"))
```


We can view the file details here.
```{r}
dir_info(raw_dir)
```

We can see that for year 2004 and 2005 the files are actually empty, so we'll disregard these two when reading.

### Read
The 2001—2003 data has rows to escape. So we'll read the two datasets separately. The file's also not in a standard Excel format, so we will use `rvest` to scrape the html content. We'll also add the year to each dataframe. 
```{r raw_read}
ga_lob_read <- function(ga_path){
  year_from_file <- str_extract(ga_path,"20\\d{2}")
  ga_html <- ga_path %>% read_html()
  ga_node <- ga_html %>% html_node("table")
  ga_table <- ga_node %>% html_table()
  names(ga_table) <- ga_table[1,]
  ga_table <- ga_table[2:nrow(ga_table),]
  ga_table <- ga_table %>% 
    mutate(Year = year_from_file)
  return(ga_table)
}

ga_lob <- map_dfr(raw_lob_paths[3:length(raw_lob_paths)], ga_lob_read)
ga_grp <- map_dfr(raw_grp_paths[3:length(raw_grp_paths)], ga_lob_read)
```
According to the website, the `PaymentExceeds` column is a logical vector indicating whether payments have exceeded $10,000.

## Explore
```{r glimpse}
glimpse(galr)
tail(galr)
```

### Missing

```{r na_count}
col_stats(galr, count_na)
```

```{r na_flag}
galr <- galr %>% flag_na(filer, filing_date, total_expenditures)
sum(galr$na_flag)
```

```{r na_view}
galr %>% 
  filter(na_flag) %>% 
  select(filer, filing_date, total_expenditures)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
galr <- flag_dupes(galr, everything())
```

### Categorical

```{r distinct_count}
col_stats(galr, n_distinct)
```

### Dates

We can examine the validity of `date_clean`
```{r date_range}
min(galr$date_clean)
max(galr$date_clean)
sum(galr$date_clean > today())
```

```{r bar_year, echo=FALSE}
galr <- galr %>% mutate(year = as.integer(year))

galr %>%
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 2)) +
  coord_cartesian(xlim = c(1989, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Jersey Lobbyists Expenditures by Year",
    caption = "Source: NJ ELEC",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

The address field only exists in certain datasets of certain years, but we can separate them by the comma delimiter.

```{r address_sep}
galr <- galr %>% 
  separate(
    col = address,
    into = c("addr_sep", "city_sep", "state_zip"),
    sep = "([:blank:]+)?,",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  )

st_regex <- valid_state %>% paste0(collapse = "|")

galr <- galr %>% 
  mutate(state_sep = 
    str_extract(state_zip,st_regex),
    zip_sep = str_remove(state_zip, state_sep) %>% str_remove(",") %>% str_trim()
  )
```

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
galr <- mutate(
   .data = galr,
   addr_norm = normal_address(
     address = addr_sep,
     abbs = usps_street,
     na = invalid_city
   )
 )
```

```{r address_view}
galr %>% 
  select(addr_sep, addr_norm) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
galr <- mutate(
  .data = galr,
  zip_norm = normal_zip(
    zip = zip_sep,
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  galr$zip_sep,
  galr$zip_norm,
  compare = valid_zip
)
```

### State
The two-letter state abbreviations are all valid and don't need to be normalized. 

```{r state_norm}
prop_in(galr$state_sep, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
galr <- galr %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(galr$city_sep),
  galr$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
galr <- galr %>% 
  rename_all(~str_replace(., "_norm", "_clean")
             ) %>%
  rename(state_clean = state_sep) %>% 
  select(-state_zip)
```

```{r clean_glimpse}
glimpse(sample_n(galr, 20))
```

1. There are `r comma(nrow(galr))` records in the database.
1. There are `r comma(sum(galr$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(galr$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nj", "lobby", "data", "clean"))
clean_path <- path(clean_dir, "nj_lobby_exp_clean.csv")
write_csv(galr, clean_path, na = "")
file_size(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
