---
title: "Georgia Lobbyying Expenditures"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ga", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  readxl, #read excel files
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects
[ethics]:https://media.ethics.ga.gov/search/Lobbyist/Lobbyist_ByExpenditures.aspx
## Data

Lobbying data is obtained from the [Georgia Government Transparency and Campaign Finance Commission][ethics].
We use the Lobbyist Roster under the **Search Lobbyist** feature to return a list of lobbyists. There're two options for the data output, organized by group or lobbyist. We will use the Lobbyist Group Roster so that the data is organized by clients in relation to the lobbyists they employed.

Note that the data is separated by year, and we will use `httr` to obtain each year's data via a `POST` request, whose parameters can be found from network activities.

## Download
We can set up the raw directory.
```{r}
raw_dir <- dir_create(here("ga", "lobby", "data", "raw","exp"))
```
We can download the data files by year from 2016 to 2020. The 2020 data file is downloaded on September 29. 



Then We can view the file details here.
```{r}
dir_info(raw_dir)
```


### Read
The file's also not in a standard Excel format, so we will use `rvest` to scrape the html content.
```{r raw_read}

ga_lob_read <- function(ga_path){
  year_from_file <- str_extract(ga_path,"20\\d{2}")
  #lob_type <- str_extract(ga_path,"(?<=_)\\w{3}(?=_\\d+)")
  lob_file <- path(raw_dir, glue("ga_lob_exp_{year_from_file}.csv"))
if (file_exists(lob_file)) {
  message("File for year {year_from_file} already converted, skipping")
} else {
  message(glue("Start converting file for year {year_from_file}"))
  ga_html <- ga_path %>% read_html()
  ga_node <- ga_html %>% html_node("table")
  ga_table <- ga_node %>% html_table()
  names(ga_table) <- ga_table[1,]
  ga_table <- ga_table[2:nrow(ga_table),]
    write_csv(
    x = ga_table,
    path = lob_file
  )
    message(glue("Conversion complete for year {year_from_file}"))
}
}

ga_lob <- map_dfr(dir_ls(raw_dir), ga_lob_read)
```


Now we can read the converted csv files.
```{r}
gale <- dir_ls(raw_dir,regexp = ".*exp_.*.csv") %>% 
  map_dfr(read_csv,col_types = cols(.default = col_character()))

gale <- gale %>% 
  mutate(Amount = as.numeric(Amount),
         Date = as.Date(Date, format = "%m/%d/%Y"))
```



### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
gale <- flag_dupes(gale, everything())
sum(gale$dupe_flag)
```

## Explore
```{r glimpse}
glimpse(gale)
tail(gale)
```


### Missing

```{r na_count}
col_stats(gale, count_na)
```


### Categorical

```{r distinct_count}
col_stats(gale, n_distinct)
```

### Dates

We can examine the validity of `date_clean`. It looks pretty clean.
```{r date_range}
min(gale$Date)
max(gale$Date)
sum(gale$Date > today())
```

```{r bar_year, echo=FALSE}
gale <- gale %>% mutate(ReportYear = as.integer(ReportYear))

gale %>%
  count(ReportYear) %>% 
  mutate(even = is_even(ReportYear)) %>% 
  ggplot(aes(x = ReportYear, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2006, 2020, by = 2)) +
  coord_cartesian(xlim = c(2006, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Georgia Lobbyists Expenditures by Year",
    caption = "Source: Georgia Government Transparency and Campaign Finance Commission",
    x = "Year",
    y = "Count"
  )
```

### Continuous
```{r hist}
brewer_dark2 <- RColorBrewer::brewer.pal(n = 8, name = "Dark2")
gale %>% 
  ggplot(aes(Amount)) +
  geom_histogram(fill = brewer_dark2[1]) +
  geom_vline(xintercept = median(gale$Amount, na.rm = TRUE)) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "GA Lobbying Expenditures Amount Distribution",
    caption = "Source: Georgia Government Transparency and Campaign Finance Commission",
    x = "Amount",
    y = "Count"
  )
```



## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
gale <- gale %>% 
  rename_all(~str_replace(., "_norm", "_clean")
             )
```

```{r clean_glimpse}
glimpse(sample_n(gale, 20))
```

```{r total count}
nrow(gale)
```


1. There are `r comma(nrow(gale))` records in the database.
1. There are `r comma(sum(gale$dupe_flag))` duplicate records in the database.
1. The range and distribution of `date` seem reasonable.
1. There are few records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ga", "lobby", "data", "clean","exp"))
clean_path <- path(clean_dir, "ga_lobby_exp_clean.csv")
write_csv(gale, clean_path, na = "")
file_size(clean_path)
```

## Upload

Using the `aws.s3` package, we can upload the file to the IRW server.

```{r upload,eval=FALSE}
s3_path <- path("csv", basename(clean_path))
put_object(
  file = clean_path,
  object = s3_path, 
  bucket = "publicaccountability",
  acl = "public-read",
  multipart = TRUE,
  show_progress = TRUE
)
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```

