---
title: "Nevada Lobbying Registration Data Diary"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project


The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.


Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:


1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved


## Objectives


This document describes the process used to complete the following objectives:


1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction


## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  rvest, # read html tables
  httr, # interact with http requests
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  vroom, #read deliminated files
  readxl, #read excel files
  pdftools #wrangle pdf texts
)
```

```{r fix_fun, echo=FALSE, collapse = TRUE}
# fix conflict
here <- here::here
```
This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.


The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.


## Download
Set the download directory first.
```{r create raw_dir}
# create a directory for the raw data
doc_dir <- here("sd", "lobby", "docs")

raw_dir <- here("sd", "lobby", "data", "raw","reg")

dir_create(c(raw_dir, doc_dir))
```

## Scrape
We will use `rvest` to scrape the website. The [South Dakota SOS site](https://sosenterprise.sd.gov/BusinessServices/Lobbyist/LobbyistSearch.aspx) makes available all registry from year 2016 to 2019. 

The underlying URLs for each year is generated by inspecting the network traffic tool. 
We can see that while the website does provide a table, it does not include lobbyist Employment Date, or client registration and authorization date. We'll scrape such information from its child pages. 
```{r download}
sd_lob_url <- glue("https://sosenterprise.sd.gov/BusinessServices/Lobbyist/LobbyistSearch.aspx?__LASTFOCUS=&__EVENTTARGET=ctl00%24MainContent%24SearchButton&__EVENTARGUMENT=&__VIEWSTATE=QMUH%2F71bweLgBn%2BznQBR6rfE%2BXa%2F%2BStHPt1eZzPiGNdd6v4mDKH0DVVlNYfiG4oSt3BAIRCmyWlwoe6RrxZVMJOEYmejeDHzHkGw7j5PJerXmVLKXLbSWWipmGrNtmVvwUY0E3nmtXzmdF6ipkbBNVyVXEM2OPP8b4QTbOykcVPzkfhMxpyVqkuZCloJ7ek4xutnPgnMc5GuueAmZLHSo8UYe4xEagZxXfGvtOj6FOjORYsEE1fIVeLgbzrgqr8Sj0J8CX295pjHuDZclsdA7P%2FoB4VHrbq8HVNfL3bllAR%2Fkdf6Xx0z839EAoKOGfKyjzVCKFeV3Jc%2FX7DeEP1zc9LLNBuqkl1s4HsqwOA%2FLGGayjS8qwJcqOn4xt%2FyRnbm%2F4YWEFAVbeF2ANDzqKGl52VAja9dm8OFBL4R9cBTPrERqIAfQ4V5N92MpToQmEpCNmVk38titrJaxR%2BEZWNZN4qHxd3qU5qe8gj41oXIwfGEf6Hw6mLtVTQ31MxygpIwtjbGS08ojx9aEXenEYJ3Q7nciaoPzh%2B7KexvC5urZxZJryUV4oQbrYjYOSsZegGNo9AvPnDWb8kTQ5ZtkgpXj2owXRqCmy9PYjbRavJ78CoLZzev0GsF9Jrbs1J%2FyA5cPZX0uDt0ltopT207FyPoAqP5yqj1GGtOjrN6tM49muI6S9DMpEgqpgupGGvFj5W10BQBjcPf4dq3HHXXmc4%2FJuNaqr9FM0Dcglt%2FRrfOuapK9KYB09LEV40%2BqDwuA39U2zaFr%2FZxJZRgSIDHibV%2BJjlId56Br%2BFzu%2FqpnaNx7LBMU%2FhaGSeE%2BuKE%2BqRl6c7dhp6FeLakZf6UA1BRvkEMNeQIqqKyLyEws9QkW1cpCXvkxp0X9KwcEw2yjnJMnPwQVhOUa1XXjtqs%2FDHxl99%2BuX0UDUvSZx5CX8OreUgXxrDemvBLmDypFNj%2FRfzw7m2DTiApftHk83R4ncpL7OmAcvg%2FDvEhm%2Bbd9P4iKwLbGXiRLfbW&__VIEWSTATEGENERATOR=C5C1E4B2&__VIEWSTATEENCRYPTED=&__EVENTVALIDATION=lrPcPU9%2BUYIbTw9luMruEs9a93D61wh3%2FmT2MhlqL2WKQW1dDE7utJX6N1hlDGvdBrTXBlPwTaB0OG4Opc0aOwvcsVNsDyI1rQrkLDaeoYZ5O%2FB66lal9ya6149GgVoi9qjcY366%2BFySL10nVzziMsu3p2C05MVFpF5OQWz%2Bd95YACoUm%2B9m72Z4Lskw0XweIf%2BAqEszex3mllm066myEMFE6%2FRVMcmdjGSa2SXtk7dzrMxj7TT3BDKI4n5%2BAlyWDY0h80FPow61mkHk5d5P96Z82tgBO4fRej4x4w%2BnDoNCse%2BwcqEgTWO09PmP%2BT5klZZ9Ytfyy3p01KrujXgdyMg%2BfpKcgiEAlOIYSsdKVaFVy9fmkguKt9PCZgXA25MliNdtZjObTitOyJlxmgTwlmRnBNJlCAsyYvvFJfnyhYPON2DC&ctl00%24MainContent%24searchBy=chkSearchByPrivate&ctl00%24MainContent%24txtFirstName=&ctl00%24MainContent%24txtMiddleName=&ctl00%24MainContent%24txtLastName=&ctl00%24MainContent%24txtEmployerName=&ctl00%24MainContent%24txtSubject=&ctl00%24MainContent%24slctYears={2016:2020}")


parseweb <- function(lob_url){
  page <- GET(lob_url)
  
  table_content <- page %>% content() %>% html_nodes('table')

parsetb <- table_content[[1]] %>% html_table(header = T, fill = F) %>% distinct() %>% as_tibble()

  string <- 
  str_match_all(as.character(table_content[[1]]),'a href=(.+) class=')[[1]][,2] %>% 
    str_extract_all("L.+\\d") %>% unlist()

parsetb <- parsetb %>% mutate(ext_url = string %>% str_c("https://sosenterprise.sd.gov/BusinessServices/Lobbyist/",.))
  
return(parsetb)
}

sd_reg <- sd_lob_url %>% map_dfr(parseweb)
# function fetchspec goes into each link to the lobbyist ID and grab information.
fetchspec <- function(ext_url) {
  x <- GET(ext_url) 
  employment_date <- x %>% content() %>% html_node("#ctl00_MainContent_txtEmploymentDate") %>% html_text()
  employer_registration_date <- x %>% content() %>% html_node("#ctl00_MainContent_txtRegistrationDate") %>% html_text()
  authorization_date <- x %>% content() %>% html_node("#ctl00_MainContent_txtAuthorizationDate") %>% html_text()
  status <- x %>% content() %>% html_node("#ctl00_MainContent_txtAuthorizationDate") %>% html_text()
  occupation <- x %>% content() %>% html_node("#ctl00_MainContent_txtOccupation") %>% html_text()
  legislation_subject <- x %>% content() %>% html_node("#ctl00_MainContent_txtSubject") %>% html_text()
  agent_name <- x %>% content() %>% html_node("#ctl00_MainContent_txtAgentName") %>% html_text()
  tib <- tibble(ext_url,employment_date, occupation, status,employer_registration_date, authorization_date, legislation_subject)
  return(tib)
}


```

Now we have a structured dataframe to work with. 
We can now save it as raw

```{r save raw}
raw_dir <- here("sd", "lobby", "data", "raw","reg")
dir_create(raw_dir)
sd_reg %>% 
  select(-lob_url) %>% 
  write_csv(
    path = glue("{raw_dir}/sd_lobby_reg.csv"),
    na = ""
  )

sd_reg %>% 
  write_csv()
```


## 
