---
title: "FAA Pilot Registration"
author:
  - "Victor Brew"
  - "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120, knitr.kable.NA = "")
  set.seed(5)
}
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  pdftools, # read documentation
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

```{r where_here}
# where does this document knit?
here::i_am("national/faa_pilots/docs/faa_pilots_basic.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The database of Federal Aviation Administration (FAA) pilot registrations can
be obtained from [the FAA website][faa].

[faa]: https://www.faa.gov/licenses_certificates/airmen_certification/releasable_airmen_download/

> ### Airmen Certification Database
> * Airmen Certification Branch is not the authoritative source for medical
data.
> * The expiration date provided in the downloadable file is for informational
purposes only.
> * Any questions regarding medical information should be directed to Aerospace
Medical Certification Division.
> 
> We update these files monthly. The records in each database file are stored in
either fixed length ASCII text format (TXT) or comma-delimited text format (CSV)
which is already separated into airmen basic records and certificate records.
Both formats can be manipulated by common database applications such as MS
Access.
> 
> This information does not include airmen certificate number data, nor does it
include the records of those airmen who do not want their addresses released.
You can also elect to Change the Releasability Status of your Address if you do
not want it listed in the database.

We will be downloading the data in fixed length ASCII text format and can use 
the [provided documentation][doc] to learn more about the data and how to read
it.

[doc]: https://www.faa.gov/licenses_certificates/airmen_certification/media/Help.pdf

```{r download_docs}
pdf_tmp <- file_temp(ext = "pdf")
download.file(
  "https://www.faa.gov/licenses_certificates/airmen_certification/media/HelpComm.pdf",
  destfile = pdf_tmp
)
```

### Overview

> On April 5, 2000, the Wendell H. Ford Aviation Investment and Reform Act for
the 21st Century became Public Law 106-181. Section 715 of that law requires the
Federal Aviation Administration to release names, addresses, and ratings
information for all airmen after the 120th day following the date of enactment.
> 
> The law also requires that the airmen be given an opportunity to elect that
their address information be withheld from release under this law. Accordingly,
the FAA sent letters to all active airmen informing them of the provisions of
the law, and giving them the option to withhold their address information. The
FAA will be continuing this procedure for airmen who become active. Responses
from the letters have been processed
> 
> This file contains the names, addresses, and certificate information of those
airmen who did not respond to indicate that they wished to withhold their
address information. It is the intent of the Airmen Certification Branch to
produce this file, in its entirety, on a monthly basis. The file may be
downloaded from the Civil Aviation Registry web site at
<http://registry.faa.gov>.

### Format

> #### PILOT INFORMATION FILES LAYOUT – COMMA DELIMITED FORMAT
> Files: `PILOT_BASIC.csv` and `PILOT_CERT.csv`
> 
> These files contain Basic and Certificate data for Pilot related Certificate
Types. They have all fields separated by commas for ACCESS databases. Each file
has a heading record containing the data field literals. The Unique ID is the
key to each record and is the link between the 2 files. Each Basic file record
has 1 or more active Certificate file records. Certificate Types in this file
include: P, F, A, U, E, H, and X. The non-Pilot files may contain additional
certificate records for the Airman.

> **Notes:** 
> * File is in Unique ID order.  
> * Only active Certificates are processed; At least 1 Certificate must be
active for Airman to be in file.

##### Pilot Basic record (Record Type `00`)

A table of column formats can be read from the PDF.

```{r read_docs}
# read table and split by newline
pdf_txt <- str_split(pdf_text(pdf_tmp), pattern = "\n")
# read text tables as data frame
basic_cols <- read_table(
  file = pdf_txt[[3]][16:30],
  col_names = str_trim(str_split(pdf_txt[[3]][12], "\\s{2,}")[[1]])
)
```

```{r echo=FALSE}
kable(basic_cols)
```

## Download

```{r raw_dir}
raw_url <- "https://registry.faa.gov/database/CS052021.zip"
raw_dir <- dir_create(here("national", "faa_pilots", "data", "raw"))
raw_zip <- path(raw_dir, basename(raw_url))
```

```{r raw_download}
if (!file_exists(raw_zip)) {
  download.file(raw_url, raw_zip)
}
```

```{r zip_content, echo=FALSE}
zip_list <- 
  unzip(raw_zip, list = TRUE) %>% 
  as_tibble() %>% 
  mutate(across(Length, fs_bytes))
kable(zip_list)
```

```{r zip_unzip}
raw_csv <- path(raw_dir, zip_list$Name)
if (!all(file_exists(raw_csv))) {
  raw_csv <- unzip(raw_zip, exdir = raw_dir)
}
```

## Read

```{r raw_read}
faa <- map_df(
  .x = str_subset(raw_csv, "_BASIC.csv"),
  .f = read_delim,
  delim = ",",
  trim_ws = TRUE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `MED CLASS` = col_integer(),
    `MED DATE` = col_date("%m%Y"),
    `MED EXP DATE` = col_date("%m%Y"),
    `BASIC MED COURSE DATE` = col_date("%Y%m%d"),
    `BASIC MED CMEC DATE` = col_date("%Y%m%d"),
    `X16` = col_skip() # empty trailing comma
  )
)
```

## Match

We will be adding the 2021 FAA registration data to an existing database. To do
so, we will have to match our column names and position to that data.

```{r old_match}
faa <- faa %>% 
  select(-`BASIC MED COURSE DATE`, -`BASIC MED CMEC DATE`) %>% 
  rename_with(~str_remove_all(., "\\s"))
```

## Explore

There are `r comma(nrow(faa))` rows of `r ncol(faa)` columns. Each record
represents a single pilot registered with the FAA.

```{r glimpse}
glimpse(faa)
tail(faa)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(faa, count_na)
```

Of those pilots missing a first or last name, most appear to be from cultures
where not everybody has a name that matches the Western surname tradition.

```{r na_flag}
faa %>% 
  filter(is.na(FIRSTNAME) | is.na(LASTNAME))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
faa <- flag_dupes(faa, -UNIQUEID)
sum(faa$dupe_flag)
```

The few existing duplicates do have their own `UNIQUEID` but duplicate
everything else, including the address and medical evaluation dates.

```{r dupe_view}
faa %>% 
  filter(dupe_flag) %>% 
  arrange(FIRSTNAME, LASTNAME)
```

The existing data does not have a variable indication duplicates, so we will
have to remove this new flag.

```{r dupe_unflag}
faa <- select(faa, -dupe_flag)
```

### Categorical

```{r distinct_count}
col_stats(faa, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(faa, REGION)
explore_plot(faa, STATE)
```

### Dates

```{r date_range}
min(faa$MEDDATE)
max(faa$MEDDATE)
sum(faa$MEDDATE > today())
```

```{r bar_year, echo=FALSE}
faa %>% 
  count(year = year(MEDDATE)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col() + 
  scale_y_continuous(labels = comma) +
  coord_cartesian(xlim = c(2000, 2021)) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "FAA Pilots by Medical Evaluation Date",
    x = "Year Evaluated",
    y = "Count"
  )
```

## Wrangle

We can add the 5-digit ZIP code in the proper place to match existing format.

```{r zip_add}
faa <- mutate(faa, ZIP5 = str_sub(ZIPCODE, end = 5), .after = ZIPCODE)
```

## Conclude

Any missing checks were avoided due to the existing data format.

1. There are `r comma(nrow(faa))` records in the database.
1. There a handful of un-flagged duplicate records.
1. Date range is mostly reasonable, but will be formatted as `mm/dd/yyyy`.
1. There are `r comma(sum(faa$na_flag))` records missing key variables.
1. Consistency in geographic data has not been improved.
1. The 4-digit `year` has not been created.

We must also format our date columns as character strings in mm/dd/yyyy format.

```{r date_mdy}
faa <- faa %>% 
  mutate(across(where(is.Date), format, "%m/%d/%Y"))
```

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(path(dirname(raw_dir), "clean"))
clean_path <- path(clean_dir, glue("airmen_{year(today())}.csv"))
write_csv(faa, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
