---
title: "Arizona Expenditures"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("az", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  mdbr, # read mdb files
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("az/expends/docs/az_expends_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

Expenditures data for the state of Arizona was retrieved under open records law 
Arizona Rev. Stat. Ann. Secs. 39-121 to 39-122. After a $25 fee was paid to the
state by the Investigative Reporting Workshop we were given access to the
state's file sharing website from which a Microsoft Access Database was
downloaded.

```{r raw-dir}
raw_dir <- dir_create(here("az", "expends", "data", "raw"))
mdb_file <- path(raw_dir, "CFS_Export.mdb")
```

```{r raw-download}
file_size(mdb_file)
```

## Read

The Microsoft Access database provided by the state cannot easily be read into
R like one might with a text file. We can use the [mdbr] package to access the
open source [MDB Tools][mdbtools] command line utility. This package can list
the tables in a database and read them as data frames.

[mdbtools]: https://github.com/brianb/mdbtools
[mdbr]: https://github.com/kiernann/mdbr

To use the mdbr package, MDB Tools must first be installed. The `mdbtools`
utility can be installed from the command line on Debian systems:

```bash
# https://github.com/brianb/mdbtools
sudo apt install mdbtools
```

We can use the `mdb_tableS()` function to list all the tables in the database.

```{r mdb-tables}
az_tables <- mdb_tables(mdb_file)
```

|Table                  |Description                                     |Rows |
|:----------------------|:-----------------------------------------------|:----|
|`BallotMeasures`       |Information on ballot measures voted on         |50   |
|`Categories`           |Category names with codes                       |71   |
|`Counties`             |County names with codes                         |15   |
|`Cycles`               |Election cycle start and dates with codes       |15   |
|`EntityTypes`          |Filing entity type names with codes             |43   |
|`IncomeExpenseNeutral` |Filing type with codes                          |3    |
|`Names`                |The names and addresses of recipients      |1,197,512 |
|`Offices`              |Office names sought with codes                  |70   |
|`Parties`              |Political party names with codes                |6    |
|`ReportNames`          |Filing report type, usual a time of year        |237  |
|`Reports`              |All of the reports filed by committees       |143,497 |
|`ReportTypes`          |Report types with codes                         |8    |
|`TransactionTypes`     |Transaction type name and `IncomeExpenseNeutral` |159 |
|`Committees`           |Names, IDs, and addresses of filing committees |4,556 |
|`Transactions`         |All expenditures with amounts and dates    |6,241,937 |

These tables in this file can be exported to separate comma-delimited text 
files that can then be read into a single list.

```{r raw-csv}
raw_csv <- path(raw_dir, path_ext_set(make_clean_names(az_tables), "csv"))
```

```{r export-mdb}
if (!all(file_exists(raw_csv))) {
  for (i in seq_along(az_tables)) {
    message(az_tables[i])
    export_mdb(
      file = mdb_file,
      table = az_tables[i],
      path = raw_csv[i]
    )
  }
}
```

These CSV files can be read using the schema from the Access database.

```{r mdb-read}
az <- map(
  .x = seq_along(raw_csv),
  .f = function(i) {
    read_csv(
      file = raw_csv[i],
      col_types = mdb_schema(
        file = mdb_file,
        table = az_tables[i]
      )
    )
  }
)
```

```{r raw-name}
names(az) <- az_tables
```

```{r as-date}
az$Transactions$TransactionDate <- as_date(az$Transactions$TransactionDate)
```

## Join

The relational database contains `r length(az)` tables. The `Transactions` table
contains all the campaign finance transactions with details, such as the party
names or transaction types, only identified with ID variables. We need to use
the smaller tables to add information about each expenditure.

### Expenses

The `Transactions` table contains `r nrow(az$TransactionTypes)` different types
of transactions. Each transaction type can be related to income, an expenses, or
neutral.

```{r tran-type}
expense_types <- left_join(
  az$TransactionTypes,
  az$IncomeExpenseNeutral,
  by = "IncomeExpenseNeutralID"
)
```

We can then identify only those transaction types which relate to expenses.

```{r tran-type2}
expense_types <- expense_types %>% 
  filter(IncomeExpenseNeutral == "Expense") %>% 
  select(-starts_with("IncomeExpenseNeutral"))
```

The table of transactions can then be filtered to only those transaction types.

```{r tran-filter}
aze <- az$Transactions %>% 
  inner_join(expense_types, by = "TransactionTypeID") %>% 
  relocate(TransactionTypeName, .after = TransactionTypeID) %>% 
  select(-TransactionTypeID, TransactionType = TransactionTypeName)
```

### Committee

The spending party to each expenditure transaction is only identified by the
`CommitteeID`. The committees themselves are identified in the separate
`Committees` table, which in turn only identifies the committees by name using
the `NameID` column and `Names` table. We can join all these together and select
the subset of columns needed to identify the parties to the transaction.

```{r comm-names}
comm_names <- az$Names %>% 
  select(NameID, CommitteeName = LastName, EntityTypeID) %>% 
  left_join(az$EntityTypes, by = "EntityTypeID") %>% 
  select(-EntityTypeID) %>% 
  rename(CommitteeType = EntityTypeName)
```

```{r comm-make}
az_comm <- az$Committees %>% 
  select(1:2, starts_with("Physical")) %>% 
  rename_with(~str_replace(., "Physical", "Committee")) %>% 
  left_join(comm_names, by = "NameID") %>% 
  relocate(CommitteeName, .after = NameID) %>% 
  select(-NameID)
```

Before we join the committee names and addresses to the list of transactions, we
can first normalize the geographic data independently.

```{r comm-addr}
az_comm <- az_comm %>% 
  mutate(
    across(
      contains("Address"),
      list(xnorm = normal_address),
      abbs = usps_street
    )
  ) %>% 
  unite(
    CommitteeAddressNorm, sep = " ",
    matches("CommitteeAddress\\d_xnorm")
  )
```

```{r comm-city}
az_comm <- az_comm %>% 
  mutate(
    CommitteeCityNorm = normal_city(
      city = CommitteeCity,
      abbs = usps_city,
      states = c("AZ", "DC"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r comm-state}
az_comm <- az_comm %>% 
  mutate(
    CommitteeStateNorm = normal_state(
      state = CommitteeState,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r comm-zip}
az_comm <- az_comm %>% 
  mutate(
    CommitteeZipNorm = normal_zip(
      zip = CommitteeZipCode,
      na_rep = TRUE
    )
  )
```

```{r comm-check}
prop_in(az_comm$CommitteeCityNorm, valid_city)
prop_in(az_comm$CommitteeState, valid_state)
prop_in(az_comm$CommitteeZipNorm, valid_zip)
```

```{r comm-join}
aze <- aze %>% 
  left_join(az_comm, by = "CommitteeID") %>% 
  relocate(all_of(names(az_comm)), .after = CommitteeID)
```

### Vendor

```{r vend-name}
az_vend <- az$Names %>% 
  left_join(az$EntityTypes, by = "EntityTypeID") %>% 
  select(
    NameID, LastName, FirstName, MiddleName, Suffix, EntityTypeName,
    Address1, Address2, City, State, ZipCode
  ) %>% 
  rename(VendorType = EntityTypeName)
```

We will perform the same kind of geographic normalization on our vendor data.

```{r vend-addr}
az_vend <- az_vend %>% 
  mutate(
    across(
      contains("Address"),
      list(xnorm = normal_address),
      abbs = usps_street
    )
  ) %>% 
  unite(
    AddressNorm, sep = " ",
    matches("^Address\\d_xnorm")
  )
```

```{r vend-state}
az_vend <- az_vend %>% 
  mutate(
    StateNorm = normal_state(
      state = State,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r vend-zip}
az_vend <- az_vend %>% 
  mutate(
    ZipNorm = normal_zip(
      zip = ZipCode,
      na_rep = TRUE
    )
  )
```

```{r vend-city}
az_city <- az_vend %>% 
  distinct(City, StateNorm, ZipNorm) %>% 
  mutate(
    CityNorm = normal_city(
      city = City,
      abbs = usps_city,
      states = c("AZ", "DC"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r city-swap}
az_city <- az_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "StateNorm" = "state",
      "ZipNorm" = "zip"
    )
  ) %>% 
  rename(CityMatch = city) %>% 
  mutate(
    match_abb = is_abbrev(CityNorm, CityMatch),
    match_dist = str_dist(CityNorm, CityMatch),
    CitySwap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = CityMatch,
      false = CityNorm
    )
  ) %>% 
  select(
    -CityMatch,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
az_vend <- left_join(az_vend, az_city)
```

```{r vend-check}
prop_in(az_vend$CitySwap, valid_city)
prop_in(az_vend$State, valid_state)
prop_in(az_vend$ZipNorm, valid_zip)
```

```{r vend-join}
aze <- aze %>% 
  left_join(az_vend, by = "NameID") %>% 
  relocate(all_of(names(az_vend)), .after = Amount) %>% 
  relocate(ends_with("_norm"), .after = last_col())
```

### Finalize

```{r cat-join}
aze <- aze %>% 
  left_join(az$Categories, by = "CategoryID") %>% 
  relocate(CategoryName, .after = CategoryID) %>% 
  select(-CategoryID, Category = CategoryName)
```

```{r clean-names}
aze <- clean_names(aze, case = "snake")
```

## Explore

There are `r comma(nrow(aze))` rows of `r ncol(aze)` columns. Each record
represents a single expenditure from a committee to a vendor or another party.

```{r glimpse}
glimpse(aze)
tail(aze)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(aze, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("transaction_date", "committee_name", "amount", "last_name")
aze <- flag_na(aze, all_of(key_vars))
sum(aze$na_flag)
```

```{r na-view}
aze %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
aze <- flag_dupes(aze, -transaction_id)
sum(aze$dupe_flag)
```

```{r dupe-view}
aze %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(transaction_date, committee_name)
```

### Categorical

```{r distinct-count}
col_stats(aze, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(aze, transaction_type) + scale_x_wrap()
explore_plot(aze, category) + scale_x_wrap()
```

### Amounts

```{r amount-round}
# fix floating point precision
aze$amount <- round(aze$amount, digits = 2)
```

All expenditures in the transaction table have a negative amounts (to 
distinguish the flow of money from contributions). Since we are only dealing
with expenditures in this dataset, and we will be clarifying the giving and
receiving parties differently than contributions, we can safely convert these
values to positive amounts.

```{r amount-positive}
aze$amount <- aze$amount * -1
```

```{r amount-summary}
summary(aze$amount)
mean(aze$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(aze[c(which.max(aze$amount), which.min(aze$amount)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
aze %>%
  ggplot(aes(x = abs(amount))) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Arizona Expenditures Amount Distribution",
    caption = "Source: AZ CFS",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
aze <- mutate(aze, transaction_year = year(transaction_date))
```

```{r date-range}
min(aze$transaction_date)
sum(aze$transaction_year < 2000)
max(aze$transaction_date)
sum(aze$transaction_date > today())
```

It's common to see an increase in the number of contributins in elections years.

```{r bar-year, echo=FALSE}
aze %>% 
  count(transaction_year) %>% 
  mutate(even = is_even(transaction_year)) %>% 
  ggplot(aes(x = transaction_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1994, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona Expenditures by Year",
    caption = "Source: AZ CFS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

### Wrangle

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- aze %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
aze <- aze %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(aze$city),
  aze$city_norm,
  aze$city_swap,
  aze$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(aze$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
aze <- aze %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  relocate(ends_with("_clean"), .after = last_col()) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(aze, 1000))
```

1. There are `r comma(nrow(aze))` records in the database.
1. There are `r comma(sum(aze$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(aze$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- min(aze$transaction_date[aze$transaction_year >= 1996]) %>% 
  str_remove_all("-")
max_dt <- str_remove_all(max(aze$transaction_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("az", "expends", "data", "clean"))
clean_csv <- path(clean_dir, glue("az_expends_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(aze, clean_csv, na = "")
write_rds(aze, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
