---
title: "Arkansas Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ar", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Arkansas Secretary of State's [Financial Disclosure portal][03].

>  This page provides comma separated value (CSV) downloads of contribution,
expenditure, and loan data for each reporting year in a zipped file format.
These files can be downloaded and imported into other applications (Microsoft
Excel, Microsoft Access, etc.)
> 
> This data is extracted from the Arkansas Campaign Finance database as it
existed as of 06/24/2020 12:02 PM.

The AR SOS also provides a [data layout key][04]:

|Field |Field Name           |Description                                     |
|:-----|:--------------------|:-----------------------------------------------|
|A     |`ORG ID`             |Unique ID of the paying candidate or committee. |
|B     |`EXPENDITURE AMOUNT` |Expenditure Amount.                             |
|C     |`EXPENDITURE  DATE`  |Expenditure Date.                               |
|D     |`LAST NAME`          |Payee Last or Full Name.                        |
|E     |`FIRST NAME`         |Payee First Name.                               |
|F     |`MIDDLE NAME`        |Payee Middle Initial or Name if provided.       |
|G     |`SUFFIX`             |Payee Name Suffix.                              |
|H     |`ADDRESS 1`          |Payee Street, PO Box, etc.                      |
|I     |`ADDRESS 2`          |Payee Suite/Apartment number.                   |
|J     |`CITY`               |Payee City                                      |
|K     |`STATE`              |Payee State                                     |
|L     |`ZIP`                |Payee Zip Code                                  |
|M     |`EXPLANATION`        |Explanation provided for the expenditure.       |
|N     |`EXPENDITURE ID`     |Unique Expenditure internal ID.                 |
|O     |`FILED DATE`         |Expenditure Filed Date                          |
|P     |`PURPOSE`            |Purpose of the Expenditure.                     |
|Q     |`EXPENDITURE TYPE`   |Indicates Type of Expenditure.                  |
|R     |`COMMITTEE TYPE`     |Indicates Type of Committee                     |
|S     |`COMMITTEE NAME`     |Name of the paying committee.                   |
|T     |`CANDIDATE NAME`     |Name of the paying candidate.                   |
|U     |`AMENDED`            |Y/N if an amendment was filed.                  |

[03]: https://financial-disclosures.sos.arkansas.gov/index.html#/dataDownload
[04]: https://financial-disclosures.sos.arkansas.gov//CFISAR_Service/Template/KeyDownloads/Expenditures,%20Debts,%20and%20Payments%20to%20Workers%20File%20Layout%20Key.pdf

## Download

To download the expenditure files, we can make a series of direct `httr::GET()`
requests to the AR SOS server, downloaded the CSV files locally.

```{r raw_dir}
raw_dir <- dir_create(here("ar", "contribs", "data", "raw"))
raw_url <- str_c(
  base = "https://financial-disclosures.sos.arkansas.gov",
  path = "/CFISAR_Service/api/DataDownload/GetCSVDownloadReport"
)
```

```{r raw_download}
ys <- seq(2017, year(today()))
fix_dir <- dir_create(path(dirname(raw_dir), "fix"))
for (y in ys) {
  # download raw file
  raw_file <- glue("CON_{y}.csv")
  raw_path <- path(raw_dir, raw_file)
  GET(
    url = raw_url,
    write_disk(raw_path, overwrite = TRUE),
    query = list(
      year = "2017",
      transactionType = "CON",
      reportFormat = "csv",
      fileName = raw_file
    )
  )
  # fix downloaded file
  read_file(raw_path) %>% 
    str_replace_all("(?<!,|^)\"(?!,|$)", "'") %>% 
    str_remove_all(",\\s(?=\r\n)") %>% 
    write_file(path(fix_dir, basename(raw_path)))
  message(paste("finished", y))
}
```

```{r fix_paths}
fix_paths <- dir_ls(fix_dir)
```

## Read

```{r raw_read}
arc <- map_df(
  .x = fix_paths,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Receipt Amount` = col_double(),
    `Receipt Date` = col_date("%m/%d/%Y %H:%M:%S %p"),
    `Filed Date` = col_date("%m/%d/%Y %H:%M:%S %p")
  )
)
```

```{r raw_rename, echo=FALSE}
old_names <- names(arc)
arc <- arc %>% 
  clean_names("snake") %>% 
  rename_all(str_remove, "_name") %>% 
  rename_all(str_remove, "receipt_") %>% 
  rename_all(str_replace, "committee_", "com_") %>% 
  rename_all(str_replace, "candidate_", "cand_") %>% 
  mutate(across(amended, equals, "Y")) %>% 
  remove_empty("cols") %>% 
  select(-occupation_comment)
```

We can count the distinct values of a discrete variable to check file reading.

```{r raw_check}
count(arc, type)
```

## Explore

```{r glimpse}
glimpse(arc)
tail(arc)
```

### Missing

Variables differ in the degree of values they are missing.

```{r na_count}
col_stats(arc, count_na)
```

With `campfin::flag_na()`, we can flag any record missing a key variable.

```{r na_flag}
arc <- arc %>% 
  flag_na(date, last, amount, candidate)
```

```{r na_mean}
percent(mean(arc$na_flag), 0.1)
```

These records are missing either the `last` or `candidate` names.

```{r na_view}
arc %>% 
  filter(na_flag) %>% 
  select(date, last, amount, candidate) %>%
  col_stats(count_na)
```

Records missing a contributor `last` name are non-itemized.

```{r na_last_source}
arc %>% 
  filter(is.na(last)) %>% 
  count(source_type, sort = TRUE) %>% 
  add_prop()
```

Records missing the receiving `candidate` name are given to PACs.

```{r na_cand_type}
arc %>% 
  filter(is.na(candidate)) %>% 
  count(com_type, sort = TRUE) %>% 
  add_prop()
```

### Duplicates

If we ignore the supposedly unique `id` variable, there are a number of 
duplicate records.

```{r dupe_flag}
arc <- flag_dupes(arc, -id)
percent(mean(arc$dupe_flag), 0.1)
```

```{r dupe_view}
arc %>% 
  filter(dupe_flag) %>% 
  select(date, last, address1, amount, candidate) %>% 
  drop_na()
```

Even more of these duplicate records are missing a `last` name.

```{r dupe_na}
percent(mean(is.na(arc$last[arc$dupe_flag])), 0.1)
```

### Categorical

```{r distinct_count}
col_stats(arc, n_distinct)
```

```{r distinct_plots}
explore_plot(arc, type)
explore_plot(arc, com_type)
```

### Amounts

```{r ammount_summary}
summary(arc$amount)
mean(arc$amount <= 0)
```

The `amount` values for records missing a contributor name are much smaller on
average, however their doesn't appear to be a firm upper limit. Could still
possibly be related to itemization.

```{r hist_amount, echo=FALSE}
arc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  facet_wrap(~na_flag, scales = "free", ncol = 1) +
  labs(
    title = "Arkansas Contributions Amount Distribution",
    caption = "Source: AK SOS",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_cents}
arc %>% 
  count(cents = amount %% 1, sort = TRUE) %>% 
  add_prop()
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
arc <- mutate(arc, year = year(date))
```

```{r date_range}
min(arc$date)
sum(arc$year < 2017)
max(arc$date)
sum(arc$date > today())
```

```{r bar_year, echo=FALSE}
arc %>% 
  filter(year >= 2017) %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = ys) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arkansas Contributions by Year",
    caption = "Source: AK SOS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
arc <- arc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
arc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
arc <- mutate(
  .data = arc,
  zip_norm = normal_zip(
    zip = zip,
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  arc$zip,
  arc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
arc <- mutate(
  .data = arc,
  state_norm = normal_state(
    state = state,
    abbreviate = TRUE,
    na_rep = TRUE,
    valid = valid_state
  )
)
```

```{r state_view}
arc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  arc$state,
  arc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
arc <- mutate(
  .data = arc,
  city_norm = normal_city(
    city = city, 
    abbs = usps_city,
    states = c("AR", "DC", "ARKANSAS"),
    na = invalid_city,
    na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
arc <- arc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- arc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
arc <- arc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r}
arc %>% 
  filter(city_refine %out% valid_city) %>% 
  count(city_refine, state_norm, sort = TRUE)
```

```{r}
extra_city <- c(extra_city, "ROCHESTER HILLS", "HOLIDAY ISLAND", "GREERS FERRY")
```

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(arc$city_raw),
  arc$city_norm,
  arc$city_swap,
  arc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(arc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Arkansas City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arkansas City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
arc <- arc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_frac(arc))
```

1. There are `r comma(nrow(arc))` records in the database.
1. There are `r comma(sum(arc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(arc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ar", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ar_contribs_clean.csv")
write_csv(arc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=TRUE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE
  )
}
```

```{r s3_size}
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(arc)),
  old = c(md_code(old_names), rep("", ncol(arc) - length(old_names))),
  type = md_code(map_chr(arc, rlang::as_label)),
  def = c(
    "Unique ID of the paying candidate or committee",
    "Contribution Amount",
    "Contribution Date",
    "Payee Last or Full Name",
    "Payee First Name",
    "Payee Middle Initial or Name if provided",
    "Payee Name Suffix",
    "Payee Street, PO Box, etc",
    "Payee Suite/Apartment number",
    "Payee City",
    "Payee State",
    "Payee Zip Code",
    "Description of Contribution",
    "Unique Contribution internal ID",
    "Contribution Filed Date",
    "Cobtribution source",
    "Indicates Type of Contribution",
    "Indicates Type of Committee",
    "Name of the paying candidate",
    "Y/N if an amendment was filed",
    "Contributor employer name",
    "Contributor occupation",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original", "Type", "Definition")
))
```
