---
title: "New Jersey Lobbyying Expenditures"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nj", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  readxl, #read excel files
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Lobbying data is obtained from the [Election Law Enforcement Commission][elec].

> The Election Law Enforcement Commission (ELEC) is dedicated to administering
“The New Jersey Campaign Contributions and Expenditures Reporting Act‚” “The
Gubernatorial‚ Legislative Disclosure Statement Act‚” “Legislative and
Governmental Process Activities Disclosure Act‚” and parts of several other
laws.

According to ELEC, the overall expenditures associated with lobbying were reported by year, and can be found 
in the annual reports.

ELEC [provides a portal][portal] for each year's summary data. We can download ["All 2019 Summary Data Download Excel Spreadsheet"][2009] and for [previous years on the archive page][prev].


[elec]: https://www.elec.state.nj.us/
[portal]: https://www.elec.state.nj.us/publicinformation/gaa_annual.htm
[2009]: https://www.elec.state.nj.us/pdffiles/Lobby2019/Lobby2019.xlsx
[prev]: https://www.elec.state.nj.us/publicinformation/lobby_statistics_archive.htm

### Download
We can set up the raw directory.
```{r}
raw_dir <- dir_create(here("nj", "lobby", "data", "raw","exp"))
```


```{r raw_dir, eval = FALSE}
exp_xlsx_urls <- glue("https://www.elec.state.nj.us/pdffiles/Lobby{c(1997:1998,2014:2019)}/Lobby{c(1997:1998,2014:2019)}.xlsx")
exp_xls_urls <- glue("https://www.elec.state.nj.us/pdffiles/Lobby{c(1999:2013)}/Lobby{str_sub(c(1999:2013),start = 3L, end = 4L)}.xls")

exp_urls <- c(exp_xls_urls, exp_xlsx_urls)
wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose",
      "--content-disposition",
      url,
      paste("-P", raw_dir)
    )
  )
}

if (!all_files_new(raw_dir)) {
  map(exp_urls, wget, raw_dir)
}
```
We can view the file details here.
```{r}
raw_info <- as_tibble(dir_info(raw_dir))
```

### Read
There are a lot of individual sheets in each spreadsheet for each year. These tables are generated from different schedules of disclosure forms filed by lobbyists.
According to the ELEC forms [factsheet](https://www.elec.state.nj.us/pdffiles/Lobbying/quickfacts.pdf)


> Form L1-L
Annual Report for use by a Represented Entity. A Represented Entity is any person, partnership, committee,
association, trade association, corporation, labor union, or any other organization that employs, retains,
designates, engages, or otherwise uses the services of a Governmental Affairs Agent to influence legislation,
regulations, governmental processes, or to communicate with the general public.
Form L1-A
Annual Report for use by a lobbying firm, a law firm, a public relations firm, or other business that employs
or engages a Governmental Affairs Agent(s). A Governmental Affairs Agent is the individual(s) who
communicates with, or provides a benefit to, the State officials covered by the Act.
Form L1-G
Annual Report for use by a person whose only lobbying activity is communication with the general public,
referred to as “grassroots lobbying.” Note that “person” includes an individual, partnership, committee,
association, corporation, and any other organization or group of persons.
Form L-2
For use by a Represented Entity designating a Governmental Affairs Agent to file an Annual Report on its
behalf. Also for use by a person who engages in communication with the general public who designates a
Governmental Affairs Agent to file an Annual Report on its behalf. Note that “person” includes an individual,
partnership, committee, association, corporation, and any other organization or group of persons. The
compensation paid to the Governmental Affairs Agent or Governmental Affairs Agent Firm must be reported.
Form L-3
For use by an out-of-state person or entity for the purpose of consenting to service of process.


Since the data structure is consistent for each year and we will combine all the records into a single table, we will create a function to work with a table first and map that function to each year's spreadsheets.

#### Expenditures
We'll deal with "Summary Expend by Category" tables first, which captures expenditures of different categories. 

> SCHEDULE B - SALARY & COMPENSATION
PURPOSE: To report the salary and compensation paid by the Represented Entity to its Governmental Affairs Agent(s). Include the
reimbursement of an Agent's expenses in amounts reported.
For the Governmental Affairs Agents who are employees of the Represented Entity named on page 1, question 1, please
report the salary and other compensation paid. NOTE: Only the pro rata share of each employee's salary and
compensation need be included if the employee spends only a portion of his/her time lobbying.

> SCHEDULE C - SUPPORT PERSONNEL
PURPOSE: To report the costs of support personnel who, over the course of the reporting year, individually spend 450 or more
hours supporting the activities of the Represented Entity or Governmental Affairs Agent(s).
After determining to which person(s) this applies, report the pro rata share of those costs which are attributable to
supporting the activities of the Represented Entity or Governmental Affairs Agent(s) in influencing legislation, regulations,
governmental processes, or communicating with the general public.

> SCHEDULES D-1 & D-2 - ASSESSMENTS (A), MEMBERSHIP FEES (M), OR DUES (D)
Schedule D-1 - Specific Intent
PURPOSE: To report the amount of assessments, membership fees, or dues paid by the Represented Entity. If the assessments, membership
fees, or dues were paid by the Represented Entity with the specific intent to influence legislation, regulations, governmental
processes, or to communicate with the general public, please provide the information below:
PART I – For assessments, membership fees, or dues exceeding $100 for the calendar year:
PART II – For assessments, membership fees, or dues $100 or less for the calendar year:

> Schedule D-2 - Major Purpose
PURPOSE: To report the pro rata amount of assessments, membership fees, or dues paid by the Represented Entity. If the assessments,
membership fees, or dues were paid by the Represented Entity to an entity whose major purpose is to influence legislation,
regulations, governmental processes, or to communicate with the general public, and, was not reported on Schedule D-1, ''Specific
Intent,'' please provide the information below:
PART I – For assessments, membership fees, or dues exceeding $100 for the calendar year:
PART II – For assessments, membership fees, or dues $100 or less for the calendar year:

> SCHEDULE E - COMMUNICATION EXPENSES
PURPOSE: To report the costs of the preparation and distribution of materials related to influencing legislation, regulations,
governmental processes, and conducting communications with the general public.

> SCHEDULE F - TRAVEL/LODGING
NAME OF GOVERNMENTAL AFFAIRS AGENT AMOUNT
PURPOSE: To report the travel and lodging costs of the Governmental Affairs Agents who are employees of the Represented Entity named on page 1, question 1, related to influencing legislation, regulations, governmental processes, or communicating with the general public.

> SCHEDULE G-1 ITEMIZATION OF BENEFITS WHICH EXCEEDED $25 PER DAY OR $200 PER CALENDAR
YEAR TO STATE OFFICIALS AND THEIR IMMEDIATE FAMILY MEMBERS
PURPOSE: To report detailed information concerning benefits passed to State officials covered by the Act, as well as the immediate
family members of these officials. If the value of a benefit exceeded $25 per day or $200 per calendar year, report below.
(Select one description item for each entry from the drop down list. When selecting "O - Other", enter a description in the space provided.

Schedule G-1 is represented in the data as "benefits passing".
The summary expenditure table is, according to the spreadsheet,an
> ALPHABETICAL LISTING OF REPRESENTED ENTITIES, GOVERNMENTAL AFFAIRS AGENTS AND PERSONS COMMUNICATING WITH THE GENERAL PUBLIC  -                   SUMMARY OF EXPENDITURES BY CATEGORY

According to NJ ELEC, the `total_expenditures` field is the sum of `in-house salaries`, `support_personnel`, `assessments_membership_fees_and_dues`, `communication_expenses`, `travel_and_lodging` as well as `benefits_passing`. The `total_receipts` captures the amount paid to lobbying firms for their lobbying efforts, and the compensation to each individual lobbyist also likely come from such receipts. 

In the past, the L form likely corresponds to the L1-L forms, which are forms filed by lobbying entities(clients), while the A form likely corresponds to the L1-A forms, which are filed by lobbying firms. 

Note that due to the particular structure of the original data, it's not possible to determine one-to-one relationship between the client and lobbying firms. We'll just arrange the data by filer, which can contain both clients(usually with a 0 or NA `total_receipts` amount. 

```{r raw_read}
read_exp <- function(short_path){
  path <- path(raw_dir,short_path)
  year_on_file <- str_extract(short_path,"(?<=by)\\d{2,}(?=.xls)")
  year <- if_else(condition = nchar(year_on_file) == 2,
                  true = case_when(
                    year_on_file == "99" ~ "1999",
                    TRUE ~ paste0("20",year_on_file)
                  ),
                  false = year_on_file)
# spreadsheet lob00's data structure is slightly different from others
  target_sheet <- if_else(condition = year %in% c("2000","1997","1999"), true = 1L, false = 4L)
    # we use the "Summary Expend by Category" tab
  if (year %in% c("2016","2018","2019")) {
    
  df <- read_excel(path, sheet = target_sheet,col_types = "text",skip = 2) %>% clean_names()
  } else if(year == "1998"){
    df <- read_excel(path, sheet = target_sheet,col_types = "text",skip = 7,col_names = c("form","date","lobbyist_or_legislative_agent`","in_house_salaries","support_personnel","assessments_membership_fees_dues","communication_expenses","travel_and_lodging","benefit_passing","total_expenditures","reimbursed","compensation_paid_to_outside_agents")) %>% clean_names()
  } else{
    df <- read_excel(path, sheet = target_sheet,col_types = "text") %>% clean_names()
  }
  #the last row is the total value, and we will remove that.
  #df <-  df[1:nrow(df)-1,] 
  df <- df %>% mutate(year = year)
  if (year == "1998"){
    df <- df %>% rename(reimbursed_benefits = reimbursed)
  } else if (year %in% as.character(c(2001:2010))){
    df <- df %>% rename(assessments_membership_fees_dues = assessment_membership_fees_dues,
                        travel_and_lodging = travel_lodging)
  } else if (year %in% as.character(c(2011:2019))) {
    df <- df %>% rename(assessments_membership_fees_dues = assesments_membership_fees_dues,
                        support_personnel = support_personel)
  }
if (year %in% as.character(c(1997:2005))) {
  df <- df %>% rename(
                      filer = lobbyist_or_legislative_agent)
}
  if (year %in% as.character(c(2001:2005))) {
  df <- df %>% rename(
                      compensation_paid_to_outside_agents = out_of_house_salaries)
  }

  
  if (year %in% c("2000","1999")) {
  df <- df %>% rename(
    reimbursed_benefits = reimbursed,
    in_house_salaries = in_house,
    compensation_paid_to_outside_agents = out_of_house,
    support_personnel = support,
    communication_expense = communi,
    travel_and_lodging = travel,
    benefit_passing = benefit,
    total_expenditures = e_total,
    total_receipts = r_total,
    assessments_membership_fees_dues = assessment
  )  
    #select(-c(r_total, e_total,address))
  } else if(year == "1997") {
     df <- df %>% rename(
    in_house_salaries = in_house,
    total_receipts = receipts,
    compensation_paid_to_outside_agents= out_of_house,
    support_personnel = support,
    travel_and_lodging = travel_lodging,
    communication_expenses = communications,
    total_expenditures = expenditures,
    assessments_membership_fees_dues = assessment
  )
  }
      if (year %in% as.character(c(1999:2010))) {
  df <- df %>% rename(communication_expenses = communication_expense
                      )
}
  return(df)
}

njle <- list.files(raw_dir)%>% map_dfr(read_exp) %>% rename(filing_date = date)
```
NJ ELEC also provides a [guide](https://www.elec.state.nj.us/download/lobby/Annual_Lobbying.pdf) to each form.

We'll need to convert the years in excel numeric format to dates
```{r}
  njle <- njle %>% mutate(
                      date_clean = if_else(nchar(filing_date)==5,
                                     true = excel_numeric_to_date(as.numeric(filing_date)),
                                     false = as.Date(filing_date, format = "%m/%d/%y")))
```
We also need to remove the rows of total values.
```{r}
njle <- njle %>% filter(!is.na(form))
```


## Explore
```{r glimpse}
glimpse(njle)
tail(njle)
```

### Missing

```{r na_count}
col_stats(njle, count_na)
```

```{r na_flag}
njl <- njle %>% flag_na(filer, filing_date, total_expenditures)
sum(njle$na_flag)
```

```{r na_view}
njl %>% 
  filter(na_flag) %>% 
  select(filer, filing_date, total_expenditures)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
njl <- flag_dupes(njle, everything())
```

### Categorical

```{r distinct_count}
col_stats(njle, n_distinct)
```

### Dates

We can examine the validity of `date_clean`
```{r date_range}
min(njle$date_clean)
max(njle$date_clean)
sum(njle$date_clean > today())
```

```{r bar_year, echo=FALSE}
njle <- njle %>% mutate(year = as.integer(year))

njle %>%
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 2)) +
  coord_cartesian(xlim = c(1989, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Jersey Lobbyists Expenditures by Year",
    caption = "Source: NJ ELEC",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

The address field only exists in certain datasets of certain years, but we can separate them by the comma delimiter.

```{r address_sep}
njle <- njle %>% 
  separate(
    col = address,
    into = c("addr_sep", "city_sep", "state_zip"),
    sep = "([:blank:]+)?,",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  )

st_regex <- valid_state %>% paste0(collapse = "|")

njle <- njle %>% 
  mutate(state_sep = 
    str_extract(state_zip,st_regex),
    zip_sep = str_remove(state_zip, state_sep) %>% str_remove(",") %>% str_trim()
  )
```

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
njle <- mutate(
   .data = njle,
   addr_norm = normal_address(
     address = addr_sep,
     abbs = usps_street,
     na = invalid_city
   )
 )
```

```{r address_view}
njle %>% 
  select(addr_sep, addr_norm) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
njle <- mutate(
  .data = njle,
  zip_norm = normal_zip(
    zip = zip_sep,
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  njle$zip_sep,
  njle$zip_norm,
  compare = valid_zip
)
```

### State
The two-letter state abbreviations are all valid and don't need to be normalized. 

```{r state_norm}
prop_in(njle$state_sep, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
njle <- njle %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(njle$city_sep),
  njle$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
njle <- njle %>% 
  rename_all(~str_replace(., "_norm", "_clean")
             ) %>%
  rename(state_clean = state_sep) %>% 
  select(-state_zip)
```

```{r clean_glimpse}
glimpse(sample_n(njle, 20))
```

1. There are `r comma(nrow(njle))` records in the database.
1. There are `r comma(sum(njle$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(njle$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nj", "lobby", "data", "clean"))
clean_path <- path(clean_dir, "nj_lobby_exp_clean.csv")
write_csv(njle, clean_path, na = "")
file_size(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
