---
title: "Louisiana Expenditures"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state","la", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Expenditures records is from the [Louisiana Ethics Administration's][lea]:

>  The mission of the Ethics Administration Program is to administer and to
enforce Louisiana’s conflicts of interest legislation, campaign finance
registration and reporting requirements and lobbyist registration and disclosure
laws to achieve compliance by governmental officials, public employees,
candidates, and lobbyists and to provide public access to disclosed information.

We can search expenditure records from the LEA's expenditures search portal:

> * [Campaign Finance Expenditures][portal]  
>     Choose this option to sort and view campaign expense records.  
>     Expenditures may appear multiple times in the search.  

[lea]: http://ethics.la.gov/AboutTheBoard.aspx
[portal]: http://www.ethics.la.gov/CampaignFinanceSearch/SearchEfilingExpenditures.aspx

## Download

We can search for expenditures between two dates, however the number of results
that can be returned at a time is 100,000.

>  Due to the high volume of contribution receipts, these search results are
limited to the top 100,000 of 1,473,715 records that match your search criteria
and sorting selection.

To circumvent this cap, we perform multiple searches between the start and end
of the years between 2000 and 2022. The [cURL] commands to download these chunks
are stored in the `raw_curl.sh` text file. We can run these commands one by one
and save the returned files locally.

We will start our next update from Jan 1, 2023. 

[cURL]: https://en.wikipedia.org/wiki/CURL

```{r raw_dir}
raw_dir <- dir_create(here("state","la", "expends", "data", "raw"))
#raw_path <- path(raw_dir, "la_exp_raw.csv")
#raw_curl <- read_lines(here("state","la", "expends", "raw_curl.sh"))
```

```{r raw_curl, eval=FALSE}
for (i in seq_along(raw_curl)) {
  out_path <- path(raw_dir, glue("SearchResults-{【  seq(2000, 2022)[i]}.csv"))
  write_lines(system(raw_curl[i], intern = TRUE), out_path)
  flush_memory(); Sys.sleep(5)
}
```

```{r raw_paths}
raw_info <- dir_info(raw_dir)
nrow(raw_info)
sum(raw_info$size)
as_tibble(raw_info) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## Read

All of these yearly files can be read into a single data frame with `vroom()`.

```{r raw_read}
lae <- vroom(
  file = raw_info$path,
  escape_backslash = FALSE,
  escape_double = FALSE,
  trim_ws = TRUE,
  id = "file",
  num_threads = 1,
  #.name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    ExpenditureDate = col_date_mdy(),
    ExpenditureAmt = col_number()
  )
) %>% clean_names()
```

```{r raw_parse}
old_names <- names(lae)
lae <- lae %>% 
  mutate(across(file, path.abbrev)) %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  rename_all(str_remove, "_name$") %>% 
  rename_all(str_remove, "^filer_") %>% 
  rename_all(str_remove, "^report_") %>% 
  rename_all(str_remove, "^recipient_") %>% 
  rename_all(str_remove, "^expenditure_") %>% 
  rename_all(str_remove, "^candidate_") %>% 
  rename(amount = amt)
```

## Explore

```{r glimpse}
glimpse(lae)
tail(lae)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(lae, count_na)
```

We can flag any record missing a key variable like a name or date.

```{r na_flag}
lae <- lae %>% flag_na(date, last, amount, recipient)
sum(lae$na_flag)
```

All such records are missing a beneficiary.

```{r na_view}
lae %>% 
  filter(na_flag) %>% 
  select(date, last, amount, recipient)
```

```{r na_recount}
lae %>% 
  filter(na_flag) %>% 
  select(date, last, amount, recipient) %>% 
  col_stats(count_na)
```

### Duplicates

We can flag any record that's duplicated at least once across all variables.

```{r dupe_flag}
d1 <- duplicated(lae, fromLast = FALSE)
d2 <- duplicated(lae, fromLast = TRUE)
lae <- mutate(lae, dupe_flag = d1 | d2)
rm(d1, d2); flush_memory()
```

Over `r percent(mean(lae$dupe_flag))` of records are such duplicates

```{r dupe_count}
percent(mean(lae$dupe_flag), 0.01)
```

```{r dupe_view}
lae %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, recipient, number)
```

### Categorical

```{r distinct_count}
col_stats(lae, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(lae, code)
explore_plot(lae, type)
explore_plot(lae, schedule)
explore_plot(lae, schedule)
```

### Amounts

```{r ammount_summary}
summary(lae$amount)
mean(lae$amount <= 0)
```

```{r hist_amount, echo=FALSE}
lae %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Louisiana Expenditures Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
lae <- mutate(lae, year = year(date))
```

```{r date_range}
min(lae$date)
sum(lae$year < 2000)
max(lae$date)
sum(lae$date > today())
```

```{r bar_year, echo=FALSE}
lae %>% 
  count(year) %>% 
  mutate(even = !is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Louisiana Expenditures by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
lae <- lae %>% 
  unite(
    col = addr_full,
    starts_with("addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address_view}
lae %>% 
  select(contains("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
lae <- lae %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  lae$zip,
  lae$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
lae <- lae %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_view}
lae %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  lae$state,
  lae$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
lae <- lae %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("LA", "DC", "LOUISIANA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
lae <- lae %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- lae %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
lae <- lae %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(lae$city_raw),
  lae$city_norm,
  lae$city_swap,
  lae$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(lae$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Louisiana City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Louisiana City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
lae <- lae %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_frac(lae))
```

1. There are `r comma(nrow(lae))` records in the database.
1. There are `r comma(sum(lae$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(lae$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("la", "expends", "data", "clean"))
clean_path <- path(clean_dir, "la_expends_clean.csv")
write_csv(lae, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(lae)),
  old = c(md_code(old_names), rep("", ncol(lae) - length(old_names))),
  type = md_code(map_chr(lae, typeof)),
  def = c(
    "Source file path",
    "Spending candidate last name",
    "Spending candidate first name",
    "Expenditure code",
    "Expenditure type",
    "Expenditure number",
    "Schedule reported on",
    "Recipient vendor name",
    "Recipient street address",
    "Recipient secondary address",
    "Recipient city name",
    "Recipient state abbreviation",
    "Recipient ZIP+4 code",
    "Expenditure description",
    "Expenditure other beneficiary name",
    "Date contribution was made",
    "Contribution amount or correction",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original name", "Type", "Definition")
))
```
