---
title: "Massachusetts Contributions"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","ma", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load_gh("kiernann/mdbr")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  aws.s3, # aws cloud storage
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Massachusetts [Office of Campaign and Political 
Finance (OCPF)][https://www.ocpf.us/]. 

> The Office of Campaign and Political Finance is an independent state agency
that administers Massachusetts General Law [Chapter 55][55] the campaign finance
law, and [Chapter 55C][55c], the limited public financing program for statewide
candidates. Established in 1973, OCPF is the depository for disclosure reports
filed by candidates and political committees under M.G.L. Chapter 55.

[55]: https://www.ocpf.us/Legal/CampaignFinanceLaw
[55C]: https://www.ocpf.us/Legal/PublicFinancingLaw
[dl]: https://www.ocpf.us/Data

## Download

We can download zip files by year. These zip files contain multiple data files.

[mdb]: http://ocpf2.blob.core.windows.net/downloads/data/campaign-finance-reports.zip

We can download this archive and extract the file to the `ma/contribs/data/raw`
directory.

```{r raw_dir}
#raw_dir <- dir_create(here("state","ma", "contribs", "data", "raw"))
raw_dir <- here("state","ma", "contribs", "data", "raw")
zip_name <- glue("ocpf-{2020:2023}-reports.zip")
zip_url <- str_c("http://ocpf2.blob.core.windows.net/downloads/data2/", zip_name)
zip_paths <- path(raw_dir, zip_name) 
```

```{r raw_download, eval=FALSE}
for (i in seq_along(zip_paths)) {
 if (!this_file_new(zip_paths[i])) {
  download.file(zip_url[i], zip_paths[i])
  unzip(zip_paths[i], exdir = str_remove(zip_paths[i],".zip"))
} 
}
```

## Read

We can see that `read_me` file provided in the zip details the file layouts for these tables. For contributions, everthing is contained in the `report-items.txt` file.

```{r}
data_meta <- read_lines(dir_ls(raw_dir, regexp = "readme.txt",recurse = T)[1])

data_dict <- data_meta[175:246]

data_dict <- data_dict %>% I() %>%  read_tsv(col_names = c("variable","notes"))

data_dict <- data_dict %>% 
  mutate(notes = coalesce(notes, X3)) %>% select(-X3)
```


```{r raw_read}
raw_paths <- dir_ls(raw_dir, regexp = "report-items.txt",recurse = T)

mac <- read_tsv(
  file = raw_paths,
  na = c("", "NA", "N/A"),
  col_types = cols(
    .default = col_character(),
    Amount = col_double()
  )
)

mac <- mac %>% select(-ncol(mac))
```

```{r raw_rename, echo=FALSE}
mac <- mac %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  #rename_all(~str_replace(., "contributor", "cont")) %>% 
  rename_all(~str_remove(., "ocpf")) %>% 
  rename(
    first = first_name,
    last = name
  ) 
```

```{r echo=FALSE}
flush_memory()
```

According to the data dictionary, this file contains both receipts and expenditures. We need to narrow down to records related to receipts only using the `record_type_id` indicator.

```{r}
mac$record_type_id <- mac$record_type_id %>% as.double()

mac <- mac %>% filter(record_type_id %in% c(201:204,211,220,303,401:403))
```

### Contribution type
The type of contribution (individual, committee) is recorded in the `record_type_id`` field and each number represents a contribution category described in the data dictionary. We will join that with the original data

```{r}
cont_tb <- data_meta[90:156] %>% I() %>%  read_tsv(col_names = c("record_type_id","cont_type"))
```


```{r}
mac <- mac %>% left_join(cont_tb)
```

### Previous
Since there are some entries that were already included in the last update, we'll get rid of these rows by matching the `item_id`.
```{r}
mac_prev <- read_csv(here("state","ma", "contribs", "data","previous","ma_contribs_clean.csv"))
```

We'll then standardize these column names so that they're consistent. 
```{r}
mac <- mac %>% 
  rename(id = item_id,
          rpt_id = report_id,
          address = street_address,
          cpf_id = related_cpf_id,
          officer = principal_officer,
          tender = tender_type_id
          ) %>%
  mutate(date = mdy(date),
         tender = tender %>% str_replace("1", "Check") %>% 
str_replace("2", "Cash or Money Order") %>% 
str_replace("3", "Credit Card") %>% 
str_replace("4", "Transfer") %>% 
str_replace("5", "Other") %>% 
str_replace("6", "Cash") %>% 
str_replace("7", "Money Order"))

mac$id = as.double(mac$id)
```

```{r}
mac <- mac %>% filter(id %out% mac_prev$id)
```

```{r}
mac$rpt_id <- as.numeric(mac$rpt_id)
```


## Filer

This file doesn't contain the information about filers and recipients of contributions. We need to extract that information from the data file `reports.csv`.

```{r}
mac_f <- read_tsv(dir_ls(raw_dir, regexp = "reports.txt",recurse = T)) %>% clean_names()
```

We only need the committee name in this case, and will match the filer data with the contribution data base on the `line` (or `report_id` in the filer data base) field.

```{r}
mac_f <- mac_f %>% 
  select(rpt_id = report_id, recip_any = ocpf_full_name,rpt_type = report_type_description)
```

```{r}
mac <- mac %>% left_join(mac_f)
```

## Explore

There are `r comma(nrow(mac))` rows of `r ncol(mac)` columns.

```{r glimpse}
glimpse(mac)
tail(mac)
```

## Missing

We should flag any records missing one of the key variables needed to properly
identify a unique contribution.

```{r na_count}
col_stats(mac, count_na)
```

We can first `dplyr::coalesce()` the contributor and recipient variables to
only flag records missing _any_ kind of name.

```{r na_flag}
mac <- mac %>% 
  unite(
    col = contrib_any,
    first, last, 
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  relocate(contrib_any, .after = last_col()) %>% 
  mutate(across(contrib_any, na_if, "")) %>% 
  flag_na(contrib_any, recip_any, date, amount)
```

The only variable missing from theses columns is the coalesced contributor name.

```{r na_view}
mac %>% 
  filter(na_flag) %>% 
  select(contrib_any, recip_any, date, amount) %>% 
  col_stats(count_na)
```

For all records with a `cont_type` of "OTHER", there is no given contributor
name. We can remove these flags.

```{r na_remove}
prop_na(mac$contrib_any[which(mac$cont_type != "OTHER")])
prop_na(mac$contrib_any[which(mac$cont_type == "OTHER")])
mac$na_flag[which(mac$cont_type == "OTHER")] <- FALSE
# very few remain
percent(mean(mac$na_flag), accuracy = 0.01)
```

## Duplicates

We can create a file containing every duplicate record in the data.

```{r dupe_write}
dupe_file <- path(dirname(raw_dir), "dupes.tsv")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  mac <- mutate(mac, group = str_sub(date, end = 7))
  ma_id <- split(mac$id, mac$group)
  mas <- mac %>%
    select(-id) %>% 
    group_split(group, .keep = FALSE)
  pb <- txtProgressBar(max = length(mas), style = 3)
  for (i in seq_along(mas)) {
    d1 <- duplicated(mas[[i]], fromLast = FALSE)
    d2 <- duplicated(mas[[i]], fromLast = TRUE)
    dupes <- data.frame(id = ma_id[[i]], dupe_flag = d1 | d2)
    dupes <- filter(dupes, dupe_flag)
    vroom_write(
      x = dupes,
      path = dupe_file,
      append = TRUE,
      progress = FALSE,
    )
    rm(d1, d2, dupes)
    mas[[i]] <- NA
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
}
```

```{r dupe_read}
dupes <- vroom(
  file = dupe_file,
  col_names = c("id", "dupe_flag"),
  col_types = cols(
    id = col_number(),
    dupe_flag = col_logical()
  )
)
```

This file can then be joined against the contributions using the transaction ID.

```{r dupe_flag}
mac <- left_join(mac, dupes, by = "id")
mac <- mutate(mac, dupe_flag = !is.na(dupe_flag))
percent(mean(mac$dupe_flag), 0.1)
```

```{r dupe_view}
mac %>% 
  filter(dupe_flag) %>% 
  select(contrib_any, recip_any, date, amount) %>% 
  arrange(date, contrib_any)
```

## Categorical

```{r n_distinct}
col_stats(mac, n_distinct)
```

```{r bar_type, echo=FALSE}
explore_plot(
  data = mac,
  var = description
)
```

```{r bar_method, echo=FALSE}
explore_plot(
  data = mac,
  var = tender
)
```

```{r bar_report, echo=FALSE}
explore_plot(
  data = mac,
  var = rpt_type
) + theme(axis.text.x = element_text(angle = 15, vjust = 0.5))
```

```{r bar_category, echo=FALSE}
explore_plot(
  data = mac,
  var = cont_type
)
```

## Amounts

```{r ammount_summary}
summary(mac$amount)
mean(mac$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
mac %>%
  filter(amount >= 1, !is.na(amount)) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Massachusetts Contributions Amount Distribution",
    subtitle = "from 2020 to 2023",
    caption = "Source: MA OCPF",
    x = "Amount",
    y = "Count"
  )
```

## Dates

The actual year a contribution was made sometimes differs from the year in which
it was reported. We will
create a new `year` variable from `date` using `lubridate::year()`. This will
more accurately identify the contribution.

```{r date_year}
mac <- mutate(mac, year = year(date))

max_year <- max(mac$year, na.rm = T)
min_year <- min(mac$year, na.rm = T)
```

```{r}
mac %>%
  count(year) %>%
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) +
  coord_cartesian(xlim = c(2020, max_year)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = seq(2020, max_year, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Massachusetts Contributions by Year",
    caption = "Source: MA OCPF",
    fill = "Election Year",
    x = "Amount",
    y = "Count"
  )
```

```{r date_range}
min(mac$date)
max(mac$date)
sum(mac$date > today())
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to
facilitate this process.

```{r pre_count}
comma(nrow(mac))
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official
USPS suffixes.

This can be done by creating a separate table of unique normalized addresses.

```{r addr_norm}
addr_norm <- mac %>%
  count(address, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r addr_th, echo=FALSE}
addr_norm <- addr_norm %>% 
  mutate(address_norm = str_remove(address_norm, "(?<=\\d)\\s(?=TH\\b)"))
```

```{r addr_view}
print(addr_norm)
```

Then joining that table on to the original contributions.

```{r addr_join}
mac <- left_join(mac, addr_norm, by = "address")
rm(addr_norm)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
zip_norm <- mac %>%
  count(zip, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_join}
mac <- left_join(mac, zip_norm, by = "zip")
rm(zip_norm)
```

```{r zip_progress}
progress_table(
  mac$zip,
  mac$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the
`campfin::normal_state()` function.

```{r state_norm}
state_norm <- mac %>%
  count(state, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_join}
mac <- left_join(mac, state_norm, by = "state")
rm(state_norm)
```

```{r state_view}
mac %>%
  filter(state != state_norm) %>%
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  mac$state,
  mac$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_south}
usps_city <- usps_city %>% 
  add_row(abb = "SO", full = "SOUTH")
```

```{r city_norm}
city_norm <- mac %>%
  count(city, state_norm, zip_norm, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    city_norm = normal_city(
      city = city,
      abbs = usps_city,
      states = c("MA", "DC", "MASSACHUSETTS"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
city_norm <- city_norm %>%
  rename(city_raw = city) %>%
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>%
  rename(city_match = city) %>%
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>%
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  rename(city = city_raw)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r refine_city}
good_refine <- city_norm %>%
  mutate(
    city_refine = city_swap %>%
      key_collision_merge() %>%
      n_gram_merge(numgram = 1)
  ) %>%
  filter(city_refine != city_swap) %>%
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  select(
    city_swap, 
    city_refine
  ) 
```

```{r view_city_refines, echo=FALSE}
good_refine %>% 
  count(city_swap, city_refine, sort = TRUE)
```

Then we can join the refined values back to the database.

```{r join_refine}
good_refine <- good_refine %>% unique()

city_norm <- city_norm %>%
  left_join(good_refine, by = "city_swap") %>%
  mutate(city_refine = coalesce(city_refine, city_swap))
```

```{r city_review}
city_norm %>% 
  filter(city_norm != city_refine) %>% 
  sample_n(10)
```

```{r city_rejoin}
mac <- left_join(
  x = mac,
  y = distinct(city_norm),
  by = c("city", "state_norm", "zip_norm")
)
```

```{r post_count}
comma(nrow(mac))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(mac$city),
  mac$city_norm,
  mac$city_swap,
  mac$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```
  
```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
raw_in <- percent(prop_in(mac$city, valid_city))
progress %>%
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.90, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Massachusetts City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>%
  select(
    stage,
    all = n_distinct,
    bad = n_diff
  ) %>%
  mutate(good = all - bad) %>%
  pivot_longer(c("good", "bad")) %>%
  mutate(name = name == "good") %>%
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Massachusetts City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Percent Valid",
    fill = "Valid"
  )
```
## Conclude

```{r city_remove}
mac <- mac %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  relocate(state_clean, zip_clean, .after = city_clean)
```

```{r clean_glimpse}
glimpse(sample_n(mac, 100))
```

1. There are `r comma(nrow(mac))` records in the database.
1. There are `r comma(sum(mac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mac$na_flag))` records missing a recipient or date.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("state","ma", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ma_contribs_clean_2020-2023.csv")
write_csv(mac, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_md, echo=FALSE}
data_dict_clean <- 
  tibble(
      var = md_code(names(mac)),
  type = md_code(map_chr(mac, typeof)),
  def = c(data_dict$notes[49:69],
          "Contributor type (individual or not)",
    "Coalesced recipient name",
    "Report on which contribution reported",
              "Full combined contributor name",
    "Flag indicating missing values",
    "Flag indicating duplicate record",
    "Calendar year contribution made",
    "Normalized full street address",
    "Normalized city name",
    "Normalized 2-letter state",
    "Normalized 5-digit ZIP code"))

(dict_md <- kable(
  x = data_dict_clean,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")))
```
