---
title: "Maryland Expenditures"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("md", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is taken from the [Maryland Campaign Reporting Information System][cris].

As explained by this [CRIS help page][help]:

> ## General Information on Expenditures and Outstanding Obligations
>
> An ***expenditure*** is defined as a gift, transfer, disbursement, or promise
of money or valuable thing by or on behalf of a political committee to promote
or assist in promoting the success or defeat of a candidate, political party, or
question at  an election.
> 
> Expenditures must be election related; that is, they must enhance the
candidates election chances, such that they would not have been incurred if
there had been no candidacy. Furthermore, expenditures, including loans, may not
be for the personal use of the candidate or any other individual.
> 
> An outstanding obligation is any unpaid debt that the committee has incurred
at the end of a reporting period.

[cris]: https://campaignfinance.maryland.gov/Home/Logout
[help]: https://campaignfinance.maryland.gov/home/viewpage?title=View%20Expenditures%20/%20Outstanding%20Obligations&link=Public/ViewExpenses

## Download
Our last update included data up until June 30, 2020. In this update, we will include everything up to June 30, 2023. The next update should start with July 1, 2023. 
```{r raw_dir}
raw_dir <- dir_create(here("state","md", "expends", "data", "raw"))
raw_path <- path(raw_dir, "ExpenseInfo.xls")
raw_gone <- !file_exists(raw_path)
```

To download a copy of the search results locally, we can first `POST()` our
form information, leaving everything blank but a start date. This is the same
as filling out the form manually on the website and clicking "Search".

```{r raw_post, eval=raw_gone}
md_post <- POST(
  url = "https://campaignfinance.maryland.gov/Public/ViewExpenses",
  body = list(
    txtPayeeLastName = "",
    txtPayeeFirstName = "",
    ddlPayeeType = "",
    MemberId = "",
    txtRegistrant = "",
    CommitteeType = "",
    txtStreet = "",
    txtTown = "",
    ddlState = "",
    txtZipCode = "",
    txtZipExt = "",
    ddlCountyofResidences = "",
    ddlExpenCategory = "",
    ddlExpensePurpose = "",
    FilingYear = "",
    FilingPeriodName = "",
    ddlFundType = "",
    dtStartDate = "07/01/2020",
    dtStartDate = "06/30/2023",
    #dtEndDate = format(today(), "%m/%d/%Y"),
    txtAmountfrom = "",
    txtAmountto = "",
    Submit = "Search"
  )
)
```

From this `POST()`, we can extract the cookies needed to then submit the
corresponding `GET()` request, essentially telling the server to then click the
"Export" button on our previously searched results.

```{r raw_cookies, eval=raw_gone}
md_cookie <- cookies(md_post)$value
names(md_cookie) <- cookies(md_post)$name
```

While we _could_ export to a CSV file, the formatting they use has no
quotation escapes which makes it difficult to read the entire file properly.
While more troublesome than CSV, we can export as a Microsoft Excel file and
then _convert_ that file to CSV.

```{r raw_get, eval=raw_gone}
md_get <- GET(
  url = "https://campaignfinance.maryland.gov/Public/ExportExpensestoExcel",
  set_cookies(md_cookie),
  write_disk(raw_path, overwrite = TRUE),
  query = list(
    page = "1",
    orderBy = "~",
    filter = "~",
    `Grid-size` = "15",
    theme = "vista"
  )
)
```

This `GET()` created a local file.

```{r raw_info}
file_info(raw_path) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```


## Read

```{r}
raw_files <- dir_ls(raw_dir, glob = "*.csv")
```

```{r raw_read}
mde <- map_dfr(
  raw_files,
  vroom,
  delim = ",",
  quote = "\"",
  escape_backslash = FALSE,
  escape_double = FALSE,
  guess_max = 0,
  num_threads = 1,
  col_types = cols(
    .default = col_character(),
    `Expenditure Date` = col_date("%m/%d/%Y"),
    `Amount($)` = col_double()
  )
)
```

```{r raw_parse, echo=FALSE}
old_names <- names(mde)
mde <- mde %>% 
  clean_names("snake") %>%
  mutate(across(where(is.character), str_squish)) %>% 
  rename_all(str_remove, "expense_") %>% 
  rename_all(str_remove, "expenditure_") %>% 
  rename_all(str_remove, "_name") %>% 
  rename(type = payee_type) %>% 
  mutate(payee = coalesce(payee, vendor))
```

The read data frame should have the same number of rows as results returned from
the CRIS web portal, which we'll have to check manually. We can also count the
distinct values of a discrete variable like `method`.

```{r raw_check}
nrow(mde) # 668245 from search
count(mde, method, sort = TRUE)
```

## Explore

```{r glimpse}
glimpse(mde)
tail(mde)
```

### Missing

```{r na_count}
col_stats(mde, count_na)
```

```{r na_flag}
mde <- mutate(mde, payee = coalesce(payee, vendor))
mde <- mde %>% flag_na(date, payee, amount, committee)
percent(mean(mde$na_flag), 0.01)
```

```{r na_view}
mde %>% 
  filter(na_flag) %>% 
  select(date, payee, amount, committee)
```

```{r na_recheck}
mde %>% 
  filter(na_flag) %>% 
  select(date, payee, amount, committee) %>% 
  col_stats(count_na)
```

### Duplicates

```{r dupe_flag}
d1 <- duplicated(mde, fromLast = FALSE)
d2 <- duplicated(mde, fromLast = TRUE)
mde <- mutate(mde, dupe_flag = d1 | d2)
percent(mean(mde$dupe_flag), 0.01)
rm(d1, d2); flush_memory()
```

```{r dupe_view}
mde %>% 
  filter(dupe_flag) %>% 
  select(date, payee, amount, committee)
```

A significant amount of these duplicate values have an `amount` of zero.

```{r dupe_amounts}
mean(mde$amount == 0, na.rm = TRUE)
mean(mde$amount[mde$dupe_flag] == 0, na.rm = TRUE)
```

### Categorical

```{r distinct_count}
col_stats(mde, n_distinct)
```

```{r distinct_plots}
explore_plot(mde, type)
explore_plot(mde, category) + scale_x_truncate()
explore_plot(mde, purpose) + scale_x_truncate()
explore_plot(mde, method)
explore_plot(mde, fundtype)
```

### Amounts

```{r ammount_summary}
summary(mde$amount)
mean(mde$amount <= 0)
```

```{r hist_amount, echo=FALSE}
mde %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Maryland Expenditures Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
mde <- mutate(mde, year = year(date))
```

```{r date_range}
min(mde$date)
sum(mde$year < 2000)
max(mde$date)
sum(mde$date > today())
```

```{r bar_year, echo=FALSE}
mde %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2020:2022) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maryland Expenditures by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

The `address` variable contains all geographic information in a single string.

```{r results='asis'}
md_bullet(head(mde$address))
```

By using `tidyr::separate()` and `tidyr::unite()`, we can split this single
string into it's component pieced in new variables ending in `_sep`.

```{r}
mde <- mde %>% 
  separate(
    col = address,
    into = c(glue("addr_sep{1:10}"), "city_sep", "state_zip"),
    sep = ",\\s",
    remove = FALSE,
    fill = "left",
    extra = "merge"
  ) %>% 
  unite(
    starts_with("addr_sep"),
    col = "addr_sep",
    sep = " ",
    na.rm = TRUE,
    remove = TRUE
  ) %>% 
  mutate(across(where(is.character), na_if, "")) %>%
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s(?=\\d|-|x)",
    remove = TRUE,
    fill = "right"
  )
```



### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
mde <- mde %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_sep,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mde <- mde %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip_sep,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mde$zip_sep,
  mde$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
mde <- mde %>% 
  mutate(
    state_norm = normal_state(
      state = state_sep,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_view}
mde %>% 
  filter(state_sep != state_norm) %>% 
  count(state_sep, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  mde$state_sep,
  mde$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mde <- mde %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep, 
      abbs = usps_city,
      states = c("MD", "DC", "MARYLAND"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
mde <- mde %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- mde %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_refine_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
mde <- mde %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_count}
many_city <- c(valid_city, extra_city)
mde %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, sort = TRUE)
```

```{r city_manual}
mde <- mde %>% 
  mutate(
    city_refine = city_refine %>% 
      str_replace("^DC$", "WASHINGTON") %>% 
      str_replace("^BALTO$", "BALTIMORE") %>% 
      str_replace("^BALTIMORE CITY$", "BALTIMORE") %>% 
      str_replace("^BALTIMORE CO$", "BALTIMORE") %>% 
      str_replace("^BALTIMORE CO$", "BALTIMORE")
  )
```

```{r city_others}
many_city <- c(many_city, "COLMAR MANOR", "ANNE ARUNDEL", "PRINCE GEORGES")
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(mde$city_sep),
  mde$city_norm,
  mde$city_swap,
  mde$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(mde$city_sep, many_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Maryland City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maryland City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
mde <- mde %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine,
    -ends_with("_sep")
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(mde, 20))
```

1. There are `r comma(nrow(mde))` records in the database.
1. There are `r comma(sum(mde$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mde$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("state","md", "expends", "data", "clean"))
clean_path <- path(clean_dir, "md_expends_clean_202007-202306.csv")
write_csv(mde, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(mde)),
  old = c(md_code(old_names), rep("", ncol(mde) - length(old_names))),
  type = md_code(map_chr(mde, typeof)),
  def = c(
    "Date expenditure was made",
    "Payee name",
    "Payee full address",
    "Payee type",
    "Expenditure amount or correction",
    "Spending committee name",
    "Expenditure category",
    "Expenditure method",
    "Expenditure purpose",
    "Expenditure helping other committee",
    "Payee vendor name (unused)",
    "Funds source type (Electoral, Administrative)",
    "Freeform comment text",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original name", "Type", "Definition")
))
```
