---
title: "Mississippi Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ms", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

## Import

### Download

```{r raw_dir}
raw_dir <- dir_create(here("ms", "contracts", "data", "raw"))
raw_json <- path(raw_dir, "ms_contracts.json")
```

```{r raw_post}
ms_post <- POST(
  url = "https://www.ms.gov/dfa/contract_bid_search/Contract/ContractData",
  set_cookies(BIGipServerpl_msi_prod_https = "rd1o00000000000000000000ffff0a0df71fo443"),
  write_disk(raw_json, overwrite = TRUE),
  query = list(
    sEcho = 2,
    iColumns = 8,
    sColumns = ",,,,,,,",
    iDisplayStart = 0,
    iDisplayLength = 9999,
    mDataProp_0 = "ContractNumber",
    bSortable_0 = TRUE,
    mDataProp_1 = "ObjectID",
    bSortable_1 = TRUE,
    mDataProp_2 = "PrimaryVendor",
    bSortable_2 = TRUE,
    mDataProp_3 = "Attachments",
    bSortable_3 = TRUE,
    mDataProp_4 = "StartDate",
    bSortable_4 = TRUE,
    mDataProp_5 = "EndDate",
    bSortable_5 = TRUE,
    mDataProp_6 = "PCardEnabled",
    bSortable_6 = TRUE,
    mDataProp_7 = 7,
    bSortable_7 = FALSE,
    iSortCol_0 = 0,
    sSortDir_0 = "asc",
    iSortingCols = 1
  )
)
```

### Read

This the raw JSON content from this `POST()` can be converted to a list.

```{r raw_content}
ms_list <- content(ms_post)
```

For each element of the list, we can convert it to a single tibble row.

```{r raw_unlist}
list2tibble <- function(list_element) {
  list_element %>% 
    extract(-11) %>% 
    unlist(recursive = FALSE) %>% 
    compact() %>% 
    as_tibble() %>% 
    na_if("")
}
```

Calling this function on every element of the list and binding them together
produces a single data frame.

```{r raw_bind}
msc <- map_df(ms_list$aaData, list2tibble)
rm(ms_list)
```

Then we can parse the data frame columns into R objects.

```{r raw_parse}
mili_date <- function(x) {
  as_datetime(as.numeric(str_extract(x, "\\d+"))/1000)
}
msc <- msc %>% 
  mutate(across(ends_with("Date"), mili_date))
```

## Explore

```{r glimpse}
glimpse(msc)
tail(msc)
```
