---
title: "Florida Voters"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 95)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("fl", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  tabulizer, # scraping pdfs
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs, # local storage 
  tabulizer # read pdf columns
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

[Voter data][voter] for the state of Florida can be obtained from the state
[Division of Elections][doe]. The file was provided to the Investigative
Reporting Workshop by mail on a compact disc and uploaded to the Workshop's AWS
server.

> Voter registration and voting history information is public record under
Florida law. As a public service, the Division of Elections provides a monthly
disk on registered voters. The disk contains information extracted from the
Florida Voter Registration System and data reported by the 67 county Supervisors
of Elections. The data includes voter registration and voting information on all
voters registered up to the last day of the month prior to the date the disk
generated.Further information on the file specifications can be found in the
[Voter Extract File Layout - PDF][layout].

> In order to receive a copy of the monthly disk, email or mail your request to
the Division of Elections (see Contact Information below) before 4 pm on the
respective dates below. Please include in your request, a phone number, an email
address, and a mailing address to receive the disk by mail.The disks will be
mailed out or available for pick-up from 9:00 A.M. until 4:00 P.M. on that day.
A request received after the date and time specified will be filled with next
month’s extract disk. There is no fee associated with this complimentary disk.

> A standing request may be made to receive all monthly extract disks issued
subsequently through the end of the calendar year.All standing requests expire
at the end of the calendar year in which made and must be renewed annually.

[voter]: https://dos.myflorida.com/elections/data-statistics/voter-registration-statistics/voter-extract-disk-request/
[doe]: https://dos.myflorida.com/elections/
[layout]: https://dos.myflorida.com/media/696057/voter-extract-file-layout.pdf

## Download

The raw data from the disk can be downloaded from the Workshop's AWS server. It was last updated on Oct. 10, 2023. 


```{r aws_bucket,eval=F, echo=F}
aws_info <- get_bucket_df(
  bucket = "publicaccountability", 
  prefix = "FOR_REVIEW/fl_voters"
)
```

```{r aws_info, echo=FALSE, eval=F}
aws_info <- aws_info %>% 
  as_tibble() %>% 
  transmute(
    path = as_fs_path(Key), 
    size = as_fs_bytes(Size), 
    modification_time = parse_datetime(LastModified)
  )
print(aws_info)
```

```{r raw_dir}
raw_dir <- dir_create(here("state","fl", "voters", "data", "raw"))
```

We will save each object to a local directory.

```{r aws_get,eval=FALSE, echo=FALSE}
for (key in aws_info$path[aws_info$size > 0]) {
  p <- path(raw_dir, basename(key))
  if (!file_exists(p)) {
    save_object(
      object = key,
      bucket = "publicaccountability",
      file = p
    )
  }
}
```

```{r raw_info}
raw_info <- dir_info(raw_dir)
raw_pdf <- raw_info$path[1]
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## About

Information about the data is found in the Voter Extract Disk File Layout file:

### Content

> Content: The extract files are created monthly in tab-delimited ASCII text
file format and contain:
>
> 1. Official voter registration information. Data is extracted from the Florida
Voter Registration System and includes information on voters who are officially
registered as of the end of the prior month. Public information is included
except in those cases in which an exemption applies, including if a registered
voter requested exemption from public disclosure pursuant to Section 119.071,
Fla. Stat. (Section 97.0585, Fla. Stat.). All information on registered voters
who are participants in the Attorney General’s Address Confidentiality Program
pursuant to 741.401 – 741.465, Fla. Stat., is excluded.
> 
> 2. Unofficial voting history information. Data is extracted from independently
reported voting history from the 67 county supervisors of elections. The data is
captured at a fixed point in time and may not always correspond to information
in a county’s voter registration file because the voter may have moved to
another county after voting. You must look to the voter’s registration
information in that county-specific voter registration file. Disclaimer: The
Division of Elections cannot warrant the content or accuracy of the unofficial
voting history extract. The most accurate and current official voting history
resides with the supervisors of elections. Any discrepancies in this extract
must be resolved in favor of the official voting records in the county where the
voter actually voted.

### Format

> The disk contains two zipped (compressed) folders. One folder consists of 67
county voter registration files; the other consists of 67 county voting history
files. The files are in tab-delimited ASCII text. To unzip folder and access
file(s), open the zipped folder and drag the file(s) to a new location. File(s)
can then be opened using a wide variety of text, spreadsheet, or database
applications, including MS Access, MS Excel, Open Office, MySQL, and Oracle. The
naming convention for these files is as follows

|File Type          |File Type  |Individual File Name              |
|:------------------|:--------  |:---------------------------------|
|Voter Registration |Compressed |`Voter_Registration_YYYYMMDD.zip` |
|Voter Registration |Individual |`CountyCode_YYYYMMDD.txt`         |
|Voting History     |Compressed |`Voter _History_YYYYMMDD.zip`     |
|Voting History     |Individual |`CountyCode_H_YYYYMMDD.txt`       |

### Layout

The registration table has 38 rows:

```{r raw_cols, echo=FALSE}
raw_cols <- suppressWarnings(extract_tables(
  file = raw_pdf,
  pages = 2,
  output = "data.frame"
))
raw_cols <- raw_cols[[1]][[2]][4:42]
raw_cols <- raw_cols[-8]
raw_cols[7] <- "Requested Exemption"
```

| Field|Name                         | Length|Format             |Protected |
|-----:|:----------------------------|------:|:------------------|:---------|
|     1|`County Code`                |      3|See Code Table     |FALSE     |
|     2|`Voter ID`                   |     10|                   |FALSE     |
|     3|`Name Last`                  |     30|                   |TRUE      |
|     4|`Name Suffix`                |      5|                   |TRUE      |
|     5|`Name First`                 |     30|                   |TRUE      |
|     6|`Name Middle`                |     30|                   |TRUE      |
|     7|`Requested exemption`        |      1|Yes or No          |FALSE     |
|     8|`Residence Address Line 1`   |     50|                   |TRUE      |
|     9|`Residence Address Line 2`   |     40|                   |TRUE      |
|    10|`Residence City (USPS)`      |     40|                   |TRUE      |
|    11|`Residence State`            |      2|State Abbreviation |TRUE      |
|    12|`Residence Zipcode`          |     10|                   |TRUE      |
|    13|`Mailing Address Line 1`     |     40|                   |TRUE      |
|    14|`Mailing Address Line 2`     |     40|                   |TRUE      |
|    15|`Mailing Address Line 3`     |     40|                   |TRUE      |
|    16|`Mailing City`               |     40|                   |TRUE      |
|    17|`Mailing State`              |      2|State Abbreviation |TRUE      |
|    18|`Mailing Zipcode`            |     12|                   |TRUE      |
|    19|`Mailing Country`            |     40|                   |TRUE      |
|    20|`Gender`                     |      1|“F” “M” or “U”     |FALSE     |
|    21|`Race`                       |      1|See Code Table     |FALSE     |
|    22|`Birth Date`                 |     10|MM/DD/YYYY         |TRUE      |
|    23|`Registration Date`          |     10|MM/DD/YYYY         |FALSE     |
|    24|`Party Affiliation`          |      3|See Code Table     |FALSE     |
|    25|`Precinct`                   |      6|                   |TRUE      |
|    26|`Precinct Group`             |      3|                   |TRUE      |
|    27|`Precinct Split`             |      6|                   |TRUE      |
|    28|`Precinct Suffix`            |      3|                   |TRUE      |
|    29|`Voter Status`               |      3|Active or Inactive |FALSE     |
|    30|`Congressional District`     |      3|                   |TRUE      |
|    31|`House District`             |      3|                   |TRUE      |
|    32|`Senate District`            |      3|                   |TRUE      |
|    33|`County Commission District` |      3|                   |TRUE      |
|    34|`School Board District`      |      2|                   |TRUE      |
|    35|`Daytime Area Code`          |      3|                   |TRUE      |
|    36|`Daytime Phone Number`       |      7|                   |TRUE      |
|    37|`Daytime Phone Extension`    |      4|                   |TRUE      |

The voter history file has 5 rows:

| Field|Name            | Length|Format                  |
|-----:|:---------------|------:|:-----------------------|
|     1|`County Code`   |      3|See Code Table          |
|     2|`Voter ID`      |     10|                        |
|     3|`Election Date` |     10|MM/DD/YYYY              |
|     4|`Election Type` |      3|PPP, PRI, RUN, GEN, OTH |
|     5|`History Code`  |      1|See Code Tables         |

### Codes

The information PDF contains code tables used to identify the values for any
variable relying on codes instead of full values.

There are 6 codes in the `History Code` column:

|History Code |History Description             |
|:------------|:-------------------------------|
|A            |Voted by Mail                   |
|B            |Vote-by-Mail Ballot Not Counted |
|E            |Voted Early                     |
|N            |Did Not Vote (not required)     |
|P            |Provisional Ballot Not Counted  |
|Y            |Voted at Polls                  |

There are 7 codes in the `Race Code` column:

|Race Code |Race Description                  |
|:---------|:---------------------------------|
|         1|American Indian or Alaskan Native |
|         2|Asian Or Pacific Islander         |
|         3|Black, Not Hispanic               |
|         4|Hispanic                          |
|         5|White, Not Hispanic               |
|         6|Other                             |
|         7|Multi-racial                      |
|         9|Unknown                           |


There are 10 codes in the `Party Affiliation` column:

|Party Codes |Party Description                            |
|:-----------|:--------------------------------------------|
|CPF         |Constitution Party of Florida                |
|DEM         |Florida Democratic Party                     |
|ECO         |Ecology Party of Florida                     |
|GRE         |Green Party of Florida                       |
|IND         |Independent Party of Florida                 |
|LPF         |Libertarian Party of Florida                 |
|NPA         |No Party Affiliation                         |
|PSL         |Party for Socialism and Liberation - Florida |
|REF         |Reform Party of Florida                      |
|REP         |Republican Party of Florida                  |

There are 67 codes in the `County Name` column:

|County Code |County Name  |
|:-----------|:------------|
|ALA         |Alachua      |
|BAK         |Baker        |
|BAY         |Bay          |
|BRA         |Bradford     |
|BRE         |Brevard      |
|BRO         |Broward      |
|CAL         |Calhoun      |
|CHA         |Charlotte    |
|CIT         |Citrus       |
|CLA         |Clay         |
|CLL         |Collier      |
|CLM         |Columbia     |
|DAD         |Miami-Dade   |
|DES         |Desoto       |
|DIX         |Dixie        |
|DUV         |Duval        |
|ESC         |Escambia     |
|FLA         |Flagler      |
|FRA         |Franklin     |
|GAD         |Gadsden      |
|GIL         |Gilchrist    |
|GLA         |Glades       |
|GUL         |Gulf         |
|HAM         |Hamilton     |
|HAR         |Hardee       |
|HEN         |Hendry       |
|HER         |Hernando     |
|HIG         |Highlands    |
|HIL         |Hillsborough |
|HOL         |Holmes       |
|IND         |Indian River |
|JAC         |Jackson      |
|JEF         |Jefferson    |
|LAF         |Lafayette    |
|LAK         |Lake         |
|LEE         |Lee          |
|LEO         |Leon         |
|LEV         |Levy         |
|LIB         |Liberty      |
|MAD         |Madison      |
|MAN         |Manatee      |
|MRN         |Marion       |
|MRT         |Martin       |
|MON         |Monroe       |
|NAS         |Nassau       |
|OKA         |Okaloosa     |
|OKE         |Okeechobee   |
|ORA         |Orange       |
|OSC         |Osceola      |
|PAL         |Palm Beach   |
|PAS         |Pasco        |
|PIN         |Pinellas     |
|POL         |Polk         |
|PUT         |Putnam       |
|SAN         |Santa Rosa   |
|SAR         |Sarasota     |
|SEM         |Seminole     |
|STJ         |St. Johns    |
|STL         |St. Lucie    |
|SUM         |Sumter       |
|SUW         |Suwannee     |
|TAY         |Taylor       |
|UNI         |Union        |
|VOL         |Volusia      |
|WAK         |Wakulla      |
|WAL         |Walton       |
|WAS         |Washington   |

## Read

```{r raw_unzip}
det_dir <- dir_create(path(raw_dir, "details"))
if (length(dir_ls(det_dir)) != 67) {
   unzip(
    zipfile = raw_info$path[2],
    exdir = det_dir,
    junkpaths = TRUE
  )
} else {
  dir_ls(det_dir)
} -> raw_paths
```

```{r raw_read}
flv <- vroom(
  file = raw_paths,
  delim = "\t",
  quote = "",
  guess_max = 0,
  num_threads = 1,
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "*"),
  col_names = make_clean_names(raw_cols),
  col_types = cols(
    .default = col_character(),
    birth_date = col_date("%m/%d/%Y"),
    registration_date = col_date("%m/%d/%Y")
  )
)
```

```{r raw_rename, echo=FALSE}
flv <- rename(
  .data = flv,
  county      = county_code,
  req_exempt  = requested_exemption,
  addr1       = residence_address_line_1,
  addr2       = residence_address_line_2,
  city        = residence_city_usps,
  state       = residence_state,
  zip         = residence_zipcode,
  ml_addr1    = mailing_address_line_1,
  ml_addr2    = mailing_address_line_2,
  ml_addr3    = mailing_address_line_3,
  ml_city     = mailing_city,
  ml_state    = mailing_state,
  ml_zip      = mailing_zipcode,
  ml_county   = mailing_country,
  reg_date    = registration_date,
  party       = party_affiliation,
  prct_name   = precinct,
  prct_group  = precinct_group,
  prct_split  = precinct_split,
  prct_suffix = precinct_suffix,
  status      = voter_status,
  congress    = congressional_district,
  house       = house_district,
  senate      = senate_district,
  commish      = county_commission_district,
  school      = school_board_district,
  phone_area  = daytime_area_code,
  phone_num   = daytime_phone_number,
  phone_ext   = daytime_phone_extension,
  email       = email_address
)
```


To check the file has been correctly read, we can count the distinct values in
a known discrete variable like `party` affiliation. IF the file was ready
incorrectly due to a delimiter or quote, we would see erroneous values.

```{r raw_check}
count(flv, party, sort = TRUE)
```

### Old

```{r old_download}
prev_dir <- dir_create(raw_dir %>% str_replace("raw","previous"))
old_path <- path(prev_dir, "fl_voters_clean.csv")
```

```{r eval=FALSE}
if (!file_exists(old_path)) {
  save_object(
    file = old_path,
    object = "csv/fl_voters_clean.csv",
    bucket = getOption("aws.bucket"),
    show_progress = TRUE
  )
}
```


```{r old_read,eval=F}
flo <- vroom(
  file = old_path,
  na = c("", "*", "NA"),
  col_types = cols(
    .default = col_character(),
    dob = col_date("%m/%d/%Y"),
    reg_date = col_date("%m/%d/%Y")
  )
)
```

```{r old_reverse,eval=F}
flo <- flo %>% 
  # remove the added columns
  select(-c(setdiff(names(flo), names(flv)))) %>% 
  # keep only voters not in new data
  filter(voter_id %out% flv$voter_id)
  # spit the combined phone
  # separate(
  #   col = PHONE,
  #   into = c("area", "num"),
  #   sep = "(?<=.{3})(?=.{7})",
  #   extra = "merge",
  #   fill = "left"
  # ) %>% 
  # match names to new data
  #set_names(names(flv))
```

```{r old_inactive,eval=F}
add_prop(count(flo, status))
add_prop(count(flv, status))
flo <- mutate(flo, status = "INA")
#flo %>% write_csv(path(prev_dir, "flo_processed.csv", na=""))
flo$birth_date <- flo$birth_date %>% ymd()
```

```{r,echo=F,eval=FALSE}
flo <-read_csv(path(prev_dir, "flo_processed.csv", na=""),
               col_types = cols(
     .default = col_character(),
     birth_date = col_date("%m/%d/%Y"),
    reg_date = col_date("%m/%d/%Y")
   ))
flv$req_exempt <- as.logical(flv$req_exempt)
```

```{r old_bind,eval=FALSE}
flv <- bind_rows(flv, flo, .id = "old_flag")
rm(flo)
flv <- mutate(flv, old_flag = (old_flag == 2), req_exempt = (req_exempt == "Y"))
flv <- relocate(flv, old_flag, .after = email)
comma(nrow(flv))
```

```{r, echo=FALSE}
rm(flv)
flv <- read_csv(path(prev_dir, "flv_processed.csv"),
                col_types = cols(
                     .default = col_character(),
     birth_date = col_date("%m/%d/%Y"),
    reg_date = col_date("%m/%d/%Y")))
```



## Explore

There are `r scales::comma(nrow(flv))` rows of `r ncol(flv)` columns.

```{r glimpse}
glimpse(flv)
tail(flv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(flv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("name_last", "birth_date", "reg_date")
flv <- flag_na(flv, all_of(key_vars))
percent(mean(flv$na_flag), 0.1)
```

All of these missing rows have no `last_name` or `birth_date`.

```{r na_view}
flv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

```{r na_recount}
flv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("state","fl", "voters", "dupes_2023.csv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  fls <- flv %>% 
    select(-voter_id) %>% 
    group_split(county)
  split_id <- split(flv$voter_id, flv$county)
  pb <- txtProgressBar(max = length(fls), style = 3)
  for (i in seq_along(fls)) {
    d1 <- duplicated(fls[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(fls[[i]], fromLast = TRUE)
      dupes <- tibble(voter_id = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(fls)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("voter_id", "dupe_flag"),
  col_types = cols(
    voter_id = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(flv)
flv <- left_join(flv, dupes, by = "voter_id")
flv <- mutate(flv, dupe_flag = !is.na(dupe_flag))
sum(flv$dupe_flag)
```

We can see that, despite unique IDs, there are duplicate voters. All of these
apparent duplicates are missing many of the key values needed to identify a
record.

```{r dupe_view}
flv %>% 
  filter(dupe_flag) %>% 
  select(voter_id, all_of(key_vars), gender, party) %>% 
  arrange(reg_date)
```

```{r dupe_recount}
flv %>% 
  filter(dupe_flag) %>% 
  col_stats(count_na)
```

### Categorical

```{r distinct_count}
col_stats(flv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
fls <- sample_frac(flv, 0.01)
race_codes <- tribble(
  ~race,       ~race_name,
  "1", "Native", "2", "AAPI",
  "3", "Black",  "4", "Hispanic" ,
  "5", "White",  "6", "Other",
  "7", "Multi",  "9", "Unknown"
)
explore_plot(fls, county)
explore_plot(fls, gender)
explore_plot(left_join(fls, race_codes), race_name)
explore_plot(fls, party)
explore_plot(fls, status)
fls %>% 
  filter(!is.na(congress)) %>% 
  mutate(across(congress, as.numeric)) %>% 
  ggplot(aes(x = congress, fill = congress)) +
  geom_bar(fill = dark2["purple"]) +
  scale_x_continuous(breaks = 1:27)
rm(fls)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
flv <- mutate(flv, reg_year = year(reg_date), birth_year = year(birth_date))
```

```{r date_range}
min(flv$reg_date)
max(flv$reg_date)
```

```{r bar_year, echo=FALSE}
flv %>% 
  count(reg_year) %>% 
  mutate(
    type = ifelse(
      test = reg_year %% 4 == 0,
      yes = "pres.",
      no = ifelse(
        test = reg_year %% 2 == 0,
        yes = "mid.",
        no = "off"
      )
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = type)) + 
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1960, 2020, by = 10)) +
  coord_cartesian(xlim = c(1960, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida Voters by Registration Year",
    caption = "Source: FL DOE",
    fill = "Election Year",
    x = "Year Registered",
    y = "Count"
  )
```

```{r bar_birth, echo=FALSE}
flv %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(binwidth = 2, fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1910, 2010, by = 10)) +
  coord_cartesian(xlim = c(1910, 2010)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida Voters by Birth Year",
    caption = "Source: FL DOE",
    x = "Year Born",
    y = "Count"
  )
```

```{r bar_age, echo=FALSE}
flv %>% 
  mutate(
    age = floor(as.numeric(as.Date("2020-11-03") - birth_date)/365)
  ) %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 1, fill = dark2["purple"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(18, 108, by = 5)) +
  coord_cartesian(xlim = c(18, 108)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida Voters Age on Election Day 2020",
    caption = "Source: FL DOE",
    x = "Age",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

```{r rows_pre}
comma(nrow(flv))
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
addr_file <- path(dirname(raw_dir), "addr.tsv")
if (!file_exists(addr_file)) {
  addr_norm <- flv %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  unite(
    col = addr_full,
    everything(),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
  vroom_write(addr_norm, addr_file)
} else {
  addr_norm <- vroom(addr_file)
}
```

```{r addr_join}
flv <- left_join(flv, addr_norm)
rm(addr_norm)
```

```{r address_view}
flv %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
flv <- flv %>%
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  flv$zip,
  flv$zip_norm,
  compare = valid_zip
)
```

### State

There are _no_ state values in the voter database. We can assume all the voters
are from the state of Florida, but we will double check this against the
normalized ZIP code.

```{r state_check}
fl_zips <- zipcodes$zip[zipcodes$state == "FL"]
out_zips <- na.omit(setdiff(flv$zip_norm, fl_zips))
prop_in(flv$zip_norm, fl_zips)
zipcodes$state[which(zipcodes$zip %in% out_zips)]
```

```{r state_add}
flv <- mutate(
  .data = flv, 
  state_norm = if_else(is.na(city) & is.na(zip), NA_character_, "FL")
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm,eval=FALSE}
norm_city <- flv %>% 
  count(city, state_norm, zip_norm, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    city_norm = normal_city(
      city = str_replace(city, "^PT ST", "PORT ST"), 
      abbs = usps_city,
      states = c("FL", "DC", "FLORIDA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap,eval=FALSE}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  rename(city = city_raw)
```

```{r,echo=FALSE}
#norm_city %>% write_csv(path(prev_dir, "norm_city.csv"),na="")
norm_city <- read_csv(path(prev_dir, "norm_city.csv"),col_types = cols(.default = col_character()))
flv <- read_csv(here("state","fl", "voters", "data", "clean", "fl_voters_clean_20231010.csv"),col_types = cols(.default = col_character()))
```


```{r city_join2,eval=FALSE}
flv <- left_join(flv, norm_city)
```

```{r}
many_city <- c(valid_city, extra_city)
prop_in(flv$city_swap, many_city)
```


#### Progress

```{r city_out}
flv %>% 
  filter(city_swap %out% many_city) %>% 
  count(city, city_swap, state_norm, sort = TRUE) %>% 
  add_prop()
```

```{r city_beach}
beach_city <- c(
  "HALLANDALE BEACH",
  "DANIA BEACH",
  "LAKE WORTH BEACH",
  "PONTE VEDRA",
  "SUNNY ISLES BEACH",
  "HIALEAH GARDENS"
)
```

```{r many_beach}
many_city <- c(many_city, beach_city)
```

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(flv$city),
  flv$city_norm,
  flv$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
  mutate(across(prop_in, percent, 0.1)) %>% 
  mutate(across(n_out, comma)) %>% 
  mutate(across(prop_na, percent, 0.01)) %>% 
  kable()
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(flv$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
flv <- flv %>% 
  select(
    -city_norm,
    city_clean = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  relocate(state_clean, zip_clean, .after = city_clean)
```

```{r clean_glimpse}
glimpse(sample_n(flv, 50))
```

1. There are `r comma(nrow(flv))` records in the database.
1. There are `r comma(sum(flv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(flv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `reg_year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r echo=FALSE}
flush_memory()
```

```{r clean_dir,eval=FALSE}
clean_dir <- dir_create(here("state","fl", "voters", "data", "clean"))
clean_path <- path(clean_dir, "fl_voters_clean.csv")
write_csv(flv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
