---
title: "West Virginia Voter"
author: "Victor"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("wv", "voter", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data
Voter data for the state of West Virginia was obtained via a zip file download from the state's data portal
from the [Secretary of State] but then extracted and made availableon our IRW database  (https://publicaccountability.s3.amazonaws.com/public/PPP/InvestigatingReporting121620_wvvoters.zip). 
## Extract

```{r raw_dir}
raw_dir <- dir_create(here("wv", "voter", "data", "raw"))
raw_zip <- path(raw_dir,"InvestigatingReporting121620_wvvoters.zip")
```

The ZIP archive itself contains a number of subsequent ZIP archives.

```{r zip_list}
(zip_files <- raw_zip %>% 
  unzip(list = TRUE) %>% 
  as_tibble(.name_repair = tolower) %>% 
  mutate(across(length, as_fs_bytes)) %>% 
  mutate(across(name, as_fs_path)))
```

We can extract only the file containing the latest voter data. #There might not be a flag in here for overwrite 

```{r zip_extract}
raw_path <- unzip(
  zipfile = raw_zip, 
  files = str_subset(zip_files$name,"^State"), 
  exdir = raw_dir,
  junkpaths = TRUE,
  overwrite = TRUE
)
```

```{r}
length(raw_path) #Run this before code to see length before changes 
```

```{r}
#
file_info(raw_path) %>% 
  group_by(size) %>% 
  slice(1)
```

```{r}
file_info(raw_path) %>% # make table of file into
  group_by(size) %>% # group them by file size
  slice(1) %>% # keep only ONE of the same size
  pull(path) -> raw_path # pull the path column as vector
```

```{r}
length(raw_path) #Run this after code to see if changes worked. 
```

## Read
The .txt file containing Voter data is not a fixed-width file (FWF) with each column
found at unique line position. 

We can use this information in `readr::read_fwf()` to parse the text file as a 
data frame for exploration.

```{r raw_read}
wv <- read_delim(
  file = raw_path,
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  quote = "",
  col_types = cols(
    .default = col_character(),
    `DATE OF BIRTH` = col_date("%m/%d/%Y %H:%M:%S"),
    `REGISTRATION DATE` = col_date("%m/%d/%Y %H:%M:%S")
  )
)
```


```{r}
#wv <- wv %>%
  #mutate(
  #across(`REGISTRATION DATE`,
         #mdy)
#)

```

SnakeCase: Cleaning column names 
```{r}
#Making column names consistent
wv <- janitor::clean_names(wv)
```

## Explore

```{r glimpse}
glimpse(wv)
tail(wv)
```

### Missing

Most columns are missing some amount of values; we can flag any records missing
one of the key variables needed to identify a transaction.

```{r na_count}
#Go across every column, count the number of NA rows per column and then calculate the % of NA 
col_stats(wv, count_na)
```

```{r na_flag}
wv <- wv %>% flag_na(county_name, id_voter, last_name, first_name, date_of_birth, registration_date) #Creates a new column with trues and falses and puts a true on any row that's missing a variable ex. .5% of rows is missing one of those values. .5 is normal though - we only check for columns that are relevant to figuring out the information we need. 
mean(wv$na_flag) %>% #change this to sum to count the rows vs find %
  percent(0.01)
```


### Duplicates

There are only a handful of entirely duplicated records.

```{r dupe_flag}
wv <- flag_dupes(wv, -id_voter, .check = F) #create another logical column that puts true's next to each row with duplicates 
sum(wv$dupe_flag)
```

```{r dupe_view}
dups <- wv %>% 
  filter(dupe_flag) %>% 
  select(county_name, id_voter, last_name, first_name, date_of_birth, registration_date) %>% 
  arrange(county_name, last_name, registration_date) 
```
### Categorical

```{r distinct_count}
col_stats(wv, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(wv, county_name) + scale_x_truncate()
explore_plot(wv, poll_name) + scale_x_truncate()
```

### Dates

We can add the calendar year from `registration_date` with `lubridate::year()`

```{r date_year}
wv <- mutate(wv, year = year(registration_date), birth_year = year(date_of_birth))
```

There are 0 entries in Registration dates since the count was the exact same as the number of rows. Maybe remove that column?
```{r date_range_registration}
count_na(wv$registration_date)
min(wv$registration_date)
mean(wv$registration_date == "1900-01-01")
wv$registration_date[wv$registration_date == "1900-01-01"] <- NA
max(wv$registration_date)
```

```{r date_range_birth}
min(wv$date_of_birth)
sum(wv$date_of_birth == "1900-01-01")
wv$date_of_birth[wv$date_of_birth == "1900-01-01"] <- NA
wv$date_of_birth[wv$date_of_birth == "1900-01-01"] <- NA
max(wv$date_of_birth, na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
wv %>% 
  count(birth_year) %>% 
  ggplot(aes(x = birth_year, y = n)) +
  geom_col() + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2020, by = 10)) +
  coord_cartesian(xlim = c(1900, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "West Virginia Voter by Birth Year",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
wv <- wv %>% 
  unite(
    col = address_full,
    house_no, street, street2, unit, 
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% select(-address_full)
  
```

```{r address_view}
wv %>% 
  select(house_no, street, street2, unit, address_norm) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
wv <- wv %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  wv$zip,
  wv$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
wv <- wv %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
wv %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  wv$state,
  wv$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
wv <- wv %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("WV", "DC", "WEST VIRGINIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
wv <- wv %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- wv %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
wv <- wv %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(wv$city_raw),
  wv$city_norm,
  wv$city_swap,
  wv$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(wv$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "West Virginia City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "West Virginia City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
wv <- wv %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(wv, 50))
```

1. There are `r comma(nrow(wv))` records in the database.
1. There are `r comma(sum(wv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("wv", "voter", "data", "clean"))
clean_path <- path(clean_dir, "wv_voter_clean.csv")
write_csv(wv, clean_path, na = "")
file_size(clean_path)
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE
  )
}
```

```{r s3_size}
#as_fs_bytes(object_size(s3_path, "publicaccountability"))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wv)),
  type = md_code(map_chr(wv, typeof)),
  def = c(
    "County that the vote was registered in", 
    "Unique voter ID used to identify a voter",
    "The voters full name last name",
    "The voters full name first name",
    "The voters full name middle name",
    "The voters full name suffix",
    "The voters DOB",
    "The sex of the voter",
    "The house number of the voters mailing address",
    "The Street Address of the voters mailing address",
    "The 2nd part of an adress (if there is a PO box or apt number) of the voters mailing address",
    "The unit number (if living in an apartment) of the voters mailing address",
    "Normalized city name",
    "Normalized 2-digit state abbreviation",
    "5-digit ZIP code",
    "(Mail delivered to somewhere else they don't live at)The Street Address of the voters mailing address",
    "(Mail delivered to somewhere else they don't live at)The 2nd part of an adress (if there is a PO box or apt number) of the voters mailing address",
    "(Mail delivered to somewhere else they don't live at)The unit number (if living in an apartment) of the voters mailing address",
    "(Mail delivered to somewhere else they don't live at) Normalized city name",
    "(Mail delivered to somewhere else they don't live at) Normalized 2-digit state abbreviation",
    "(Mail delivered to somewhere else they don't live at) 5-digit ZIP code",
    "The date that the voter registered to vote",
    "The voter status when voting",
    "The congressional district of this voter",
    "The senatorial district of this voter",
    "The delegate district of this voter",
    "Legal district of magistrate court",
    "Voting precinct number",
    "Polling place within voting precinct ",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Normalized year",
    "Normalized birth year",
    "Normalized combined street address",
    "The cleaned normalized zip file from the original data",
    "normalized clean 2 letter state variable",
    "normalized clean city variable"
))
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
