---
title: "District Of Columbia Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("dc", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

A list of contracts by fiscal year can be obtained from the District of Columbia
[Contracts and Procurement Transparency Portal][cptp]. A search can be run for
results by fiscal years 2016 through 2021. This search can then be exported as
a single CSV file which we can read into R. The same thing can be done on the 
[purchase orders portal][pop].

[cptp]: https://contracts.ocp.dc.gov/contracts/search
[pop]: https://contracts.ocp.dc.gov/purchase/search

## Read

```{r raw_dir}
raw_dir <- dir_create(here("dc", "contracts", "data", "raw"))
```

First, we will read the contract awards file.

```{r raw_read_con}
dcc <- vroom(
  file = path(raw_dir, "ContractAwards.csv"),
  delim = ",",
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    `Start Date` = col_date_usa(),
    `End Date` = col_date_usa(),
    `Award Date` = col_date_usa(),
    `Amount` = col_number()
  )
)
```

Then the we will read the purchase orders file.

```{r raw_read_po}
dcp <- vroom(
  file = path(raw_dir, "PurchaseOrders.csv"),
  delim = ",",
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    `Total Amount` = col_number(),
    `Order Date` = col_date_usa(),
  )
)
```

After making some column names match, the two data frames can be bound together
into a single data frame. For contracts, the `award_date` will be used as a 
single date, paired with the `order_date` for purchase orders. The purchase
order `total_amount` will be paired with the contract's single `amount`.

```{r raw_rename_con}
dcc <- rename(
  .data = dcc,
  id = contract_number,
  vendor = vendor_supplier,
  date = award_date
)
```

```{r raw_rename_po}
dcp <- rename(
  .data = dcp,
  amount = total_amount,
  id = po_number,
  vendor = vendor_supplier,
  date = order_date
)
```

After the two data frames are bound together, columns not found in both files
(e.g., a contract's `start_date`) will be filed with `NA` for records from the 
other data type.

```{r raw_bind}
dcc <- bind_rows(contract = dcc, purchase = dcp, .id = "type")
```

## Explore

```{r glimpse}
glimpse(dcc)
tail(dcc)
```

### Missing

There are a handful of records missing a name or date needed to properly
identify a transaction, mostly the agency name.

```{r na_count}
col_stats(dcc, count_na)
```

These records can be flagged with `campfin::flag_na()`.

```{r na_flag}
dcc <- dcc %>% flag_na(date, vendor, amount, agency)
percent(mean(dcc$na_flag), 0.01)
```

```{r na_view}
dcc %>% 
  filter(na_flag) %>% 
  select(date, vendor, amount, agency, type)
```

### Duplicates

Ignoring the `id` variable, there are a handful of completely duplicated 
records. These can be flagged with `campfin::flag_dupes()`.

```{r dupe_flag, warning=TRUE}
dcc <- flag_dupes(dcc, -id)
percent(mean(dcc$dupe_flag), 0.01)
```

```{r dupe_view}
dcc %>% 
  filter(dupe_flag) %>% 
  select(date, vendor, amount, agency, type) %>% 
  arrange(date)
```

### Categorical

```{r n_distinct}
col_stats(dcc, n_distinct)
```

```{r count_distinct}
add_prop(count(dcc, agency, sort = TRUE))
add_prop(count(dcc, option_period, sort = TRUE))
```

### Continuous

#### Amounts

```{r ammount_summary}
noquote(map_chr(summary(dcc$amount), dollar))
sum(dcc$amount <= 0)
```

```{r amount_minmax}
glimpse(dcc[c(which.min(dcc$amount), which.max(dcc$amount)), ])
```

```{r hist_amount, echo=FALSE}
dcc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:10)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "District Of Columbia Contracts Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: Contracts and Procurement Transparency Portal",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
dcc <- mutate(dcc, year = year(date))
```

Aside from a few of contracts awarded much earlier, the date columns are clean.

```{r date_range}
count_na(dcc$date)
min(dcc$date, na.rm = TRUE)
sum(dcc$year < 2012, na.rm = TRUE)
max(dcc$date, na.rm = TRUE)
sum(dcc$date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
dcc %>% 
  count(year, type) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = type)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2008, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "District Of Columbia Contracts & Purchases by Year",
    caption = "Source: Contracts and Procurement Transparency Portal",
    fill = "Transaction Type",
    x = "Year Made",
    y = "Count"
  )
```

## Conclude

1. There are `r comma(nrow(dcc))` records in the database.
1. There are `r comma(sum(dcc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `award_date` seem reasonable.
1. There are `r comma(sum(dcc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `award_year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("dc", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "dc_contracts_clean.csv")
write_csv(dcc, clean_path, na = "")
file_size(clean_path)
```

The encoding of the exported file should be UTF-8 or ASCII.

```{r clean_encode}
enc <- system2("file", args = paste("-i", clean_path), stdout = TRUE)
str_replace_all(enc, clean_path, basename)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
both <- rep(TRUE, ncol(dcc))
both[c(5:7, 12)] <- FALSE
both[13:15] <- NA
dict_raw <- tibble(
  var = md_code(names(dcc)),
  type = md_code(map_chr(dcc, typeof)),
  both = both,
  def = c(
    "Transaction type (contract or purchae)",
    "Unique contract number",
    "Contract title",
    "Awarding agency name",
    "Option period awarded",
    "Contract start date",
    "Contract end date",
    "Contract awarded date, purchase made date",
    "National Institute of Governmental Purchasing code",
    "Recipient vendor name",
    "Contract amount awarded, total purchase amount",
    "Purchase order fiscal year",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year contract awarded"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Found in both", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# District Of Columbia Contracts Data Dictionary\n", dict_md),
  path = here("dc", "contracts", "dc_contracts_dict.md"),
)
```

