---
title: "State Data"
author: "First Last"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

### About

### Variables

## Import

### Download

```{r raw_dir}
raw_dir <- here("ca", "lobbying", "data", "raw")
dir_create(raw_dir)
```

```{r zip_file}
zip_url <- "https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip"
zip_file <- str_c(raw_dir, basename(zip_url), sep = "/")
```

```{r download_raw}
if (!this_file_new(zip_file)) {
  download.file(
    url = zip_url,
    destfile = zip_file
  )
}
```

```{r unzip_raw}
cal_dir <- dir_ls(raw_dir, recurse = TRUE, type = "directory", glob = "*DATA$")
if (is_empty(cal_dir)) {
  unzip(zip_file, exdir = raw_dir)
  cal_dir <- dir_ls(raw_dir, recurse = TRUE, type = "directory", glob = "*DATA$")
}
```

### Read

```{r warn_vroom}
if (packageVersion("vroom") < "1.0.2.9000") {
  warning("vroom version greater than 1.0.2.9000 is needed")
}
```

```{r read_conts}
cal_conts <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_CONTRIBUTIONS\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      FILER_ID = col_character(),
      FILING_PERIOD_START_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
      FILING_PERIOD_END_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
      CONTRIBUTION_DT = col_character(),
      RECIPIENT_NAME = col_character(),
      RECIPIENT_ID = col_character(),
      AMOUNT = col_double()
    )
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

```{r read_emp_lob}
cal <- cal_emp_lob <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMP_LOBBYIST\\d_CD.TSV"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      SESSION_ID = col_double()
    )
  ) %>% 
  clean_names("snake")
```

```{r read_emp_total}
cal_emp_total <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      CURRENT_QTR_AMT = col_double(),
      SESSION_TOTAL_AMT = col_double(),
      SESSION_YR_1 = col_double(),
      SESSION_YR_2 = col_double(),
      YR_1_YTD_AMT = col_double(),
      YR_2_YTD_AMT = col_double(),
      QTR_1 = col_double(),
      QTR_2 = col_double(),
      QTR_3 = col_double(),
      QTR_4 = col_double(),
      QTR_5 = col_double(),
      QTR_6 = col_double(),
      QTR_7 = col_double(),
      QTR_8 = col_double()
    )
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

```{r read_emp_firms}
cal_emp_firms <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER_FIRMS\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      SESSION_ID = col_double()
    )
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

```{r read_emp_hist}
# empty file
cal_lob_hist <- 
dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER_HISTORY_CD.TSV$"
  ) %>% 
  file_size()
```

```{r read_firms}
cal_firm_totals <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_FIRM\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      CURRENT_QTR_AMT = col_double(),
      SESSION_TOTAL_AMT = col_double(),
      SESSION_YR_1 = col_double(),
      SESSION_YR_2 = col_double(),
      YR_1_YTD_AMT = col_double(),
      YR_2_YTD_AMT = col_double(),
      QTR_1 = col_double(),
      QTR_2 = col_double(),
      QTR_3 = col_double(),
      QTR_4 = col_double(),
      QTR_5 = col_double(),
      QTR_6 = col_double(),
      QTR_7 = col_double(),
      QTR_8 = col_double()
    )
  ) %>% 
  clean_names("snake")
```

## Explore

```{r}
head(cal)
tail(cal)
glimpse(sample_frac(cal))
```

