---
title: "Indiana Lobbying Expenditure Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","in", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel
  httr, # GET request
  rvest, # scrape web page
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://www.in.gov/ilrc/2335.htm

## Data

Lobbyist data is obtained from the [Indiana Lobby Registration Commission][03]. The data was updated through 2022.

```{r raw_dir}
raw_dir <- dir_create(here("state","in", "lobby", "data", "raw", "exp"))
```

```{r download, eval=FALSE}
landing_url <- 'https://www.in.gov/ilrc/2335.htm'

urls <- GET(landing_url) %>% content() %>% html_nodes("a")

comp_urls <- paste0('https://www.in.gov',urls[str_detect(html_text(urls), 'Compensated Lobbyist Total')] %>% html_attr('href'))

wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose",
      "--content-disposition",
      url,
      paste("-P", dir)
    )
  )
}

if (!this_file_new(raw_dir)) {
  map(comp_urls, wget, raw_dir)
}
```

### Read

Lobbyist data is obtained from the [Indiana Lobby Registration Commission][03]. At the time of data aquisition, exployer lobbyist lists are available in xls format from 2006 to 2022.

These files come in various formats, and we will need to parse them differently according to their file extensions.
The excel files have certain rows that we need to skip. 
For the terminated column, TA=Term. April, TO=Term. October

```{r read csv}
inle_csv <- read_csv(dir_ls(raw_dir) %>% str_subset('csv'), 
                 col_types = cols(.default = col_character())) %>% 
  clean_names() 

in_0809_fs <- dir_ls(raw_dir) %>% str_subset('2008|2009')

in_1112_fs <- dir_ls(raw_dir) %>% str_subset('2011|2012')

in_1322_fs <- dir_ls(raw_dir) %>% str_subset('2013|2014|2015|2016|2017|2019|2020|2021|2022')

# Get the vector of column names by reading in the 2019 file and accessing its column headers
names_inle <- names(inle_csv)


#for (in_file in dir_ls(raw_dir) %>% str_subset('xl')) {
  
read_inxl <- function(in_file) {
df <- read_excel(in_file,range = cell_cols(1:29), col_types = "text")
                 #, col_types = c(rep("text",4), rep("numeric",25)))
# find the index of the first row whose first column (year) is not blank
 start_index <- which(!is.na(df[,1]))[1]
    df <- df[start_index:nrow(df),]
    names(df) <- names_inle
        # we also need to fill down the  lobbyist column and year column that are often shared by multiple clients
    df <- df %>% 
      mutate(year_clean = str_extract(in_file, "\\d{4}"),
             lobbyist_clean = lobbyist) %>% 
      fill(lobbyist_clean)
    return(df)
}


read_1112_inxl <- function(in_file) {
  # for 2011 and 2012
df <- read_excel(in_file, col_types = "text", range = cell_cols(1:27))
# find the index of the first row that has zero blanks
start_index <- which(rowSums(is.na(df)) == 0)[1]
    df <- df[start_index:nrow(df),]
    df <- df %>% 
      add_column(year = str_extract(in_file, "\\d{4}"),.before = 1) %>% 
      add_column(terminated = NA, .after = 3) 
    names(df) <- names_inle
    
    df <- df %>% 
      mutate(year_clean = str_extract(in_file, "\\d{4}"),
             lobbyist_clean = lobbyist) %>% 
      fill(lobbyist_clean)
    
    return(df)
}

inle_csv <- inle_csv %>% 
      mutate(year_clean = year,
             lobbyist_clean = lobbyist) %>% 
      fill(year_clean, lobbyist_clean)

in_1322 <- map_dfr(in_1322_fs, read_inxl) %>% 
  bind_rows(inle_csv)

in_1112 <- map_dfr(in_1112_fs, read_1112_inxl)

inle <- read_inxl(dir_ls(raw_dir) %>% str_subset("2010")) %>% 
  bind_rows(in_1112) %>% 
  bind_rows(in_1322)
```



Next, we can see that the file structures for 2008 and 2009 are different from the one later on. The main difference is that: 
1. The grand_totals for each year is in its own row.
2. The `first_period_*` and `second_period_*` columns use the same column for each category, but there's a `PD` column for period. 


We will transform the data accordingly.

```{r transform 0809}
read_0809_inxl <- function(in_file){
  # Step 1: move the orphan columns to the right
  df <- read_excel(in_file, col_types = "text")
  # fix year-end
  df <- df %>% 
    # create a new column total_gross filled with all net figure columns from the left
    mutate(grand_totals = `Net Figure`) %>% 
    rename_all(.funs = str_to_lower) 
    
  #identify the index of all columns that only have three valid fields (the Year-End rows), the year-end field, the actual net figure field, and the grand_total field that we just added
  grand_total_index <- which(rowSums(!is.na(df))== 3)
  # replace all other rows' grand_total as NA
  other_index <- setdiff(1:nrow(df), grand_total_index) 
  df$grand_totals[other_index] <- NA_integer_
  # Fill the NA_interger_ with actual total_gross from the bottom
  df <- df %>% 
    fill(grand_totals, .direction = "up")
  # After this step problem 1 is fixed.
  
  # Step 2. Pivot_wider based on pd
  # We'll remove the orphan rows.
  df <- df[-grand_total_index,]
  
    # create a new column that fills down column names
  df <- df %>% 
    add_column(year = NA_character_,.before = 1) %>% 
    mutate(year_clean = str_extract(in_file,"\\d{4}"))
  
  x <- df %>% filter(pd == 1)
  x <- x %>% select(-pd)
  names(x) <- c("year","lobbyist", "client", names_inle %>% str_subset("first"),"grand_totals", "year_clean")
  
  y <- df %>% filter(pd == 2)
  y <- y %>% select(-pd)
  names(y) <- c("year","lobbyist", "client", names_inle[17:length(names_inle)], "year_clean")
  
  y <- unique(y)
  x <- unique(x)
  combined <- x %>% full_join(y, by = c("year","client","lobbyist", "grand_totals","year_clean"),multiple="all")
  
  return(combined)
}

in_0809 <- map_dfr(in_0809_fs, read_0809_inxl) %>% 
  add_column(terminated = NA_character_, .after = 3)

inle <- inle %>% 
  bind_rows(in_0809)
```


We'll do some basic cleaning by turning all text columns to uppercase. We also need to get rid of the commas in supposedly numeric columns, which we deliberately read as plain text at first.
```{r clean}
inle <-inle %>% 
  mutate_if(is.character, str_to_upper) %>% 
  clean_names() %>% 
  mutate_at(.vars = vars(-c("year", "lobbyist", "terminated", "client", "lobbyist_clean")), 
              .funs=funs(as.numeric(str_remove_all(.,"\\$|,|`"))))
```

```{r examine}
head(inle)
tail(inle)
glimpse(sample_n(inle, 20))
```

### Missing

```{r na_count}
col_stats(inle, count_na)
```

```{r na_flag}
inle <- inle %>% flag_na(client, lobbyist_clean)
sum(inle$na_flag)
```

### Duplicates
We can see there's no duplicate entry.
```{r dupe_flag}
inle <- flag_dupes(inle, dplyr::everything())
sum(inle$dupe_flag)
```
### Categorical
All the numeric columns are read as plain text. We will clean these columns with `stringr::str_remove()`.
```{r conv num}
inle %>% 
  ggplot(aes(grand_totals)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:5)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contributions Amount Distribution",
    subtitle = "from 2008 to 2022",
    caption = "Source: Indiana Lobby Registration Commission",
    x = "Amount",
    y = "Count"
  )
```


## Conclude
```{r}
inle <- inle %>% mutate(lobbyist_clean = coalesce(lobbyist_clean,lobbyist))
```


```{r clean_glimpse}
glimpse(sample_n(inle, 20))
min(inle$year_clean)
max(inle$year_clean)
```

1. There are `r nrow(inle)` records in the database.
1. There are `r sum(inle$dupe_flag)` duplicate records in the database.
1. The range and distribution of `year` seems mostly reasonable except for a few entries.
1. There are `r sum(inle$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("state","in", "lobby", "data", "exp","clean"))
```

```{r write_clean}
write_csv(
  x = inle,
  path = path(clean_dir, "in_lob_exp_clean.csv"),
  na = ""
)
```

We did a little spot check for the data and it checks out.
