---
title: "District Of Columbia Licenses"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "dc", "licenses", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  jsonlite, # read json files
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/dc/licenses/docs/dc_licenses_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

We can search for professional licenses from the Department of Consumer and
Regulatory Affairs (DCRA), which provided a search portal.

## Read

First, we will submit an empty request for all licenses. Then we can save the
results to a local file.

```{r raw-dir}
raw_dir <- dir_create(here("state", "dc", "licenses", "data", "raw"))
raw_csv <- path(raw_dir, "dc_licenses.csv")
```

```{r raw-download}
if (!file_exists(raw_csv)) {
  api_url <- "https://govservices.dcra.dc.gov/oplalicensesearch/Home/GetLicenseSearchDetailsByFilter"
  # find how many records total
  find_n <- POST(api_url, body = list(pageSize = 1))
  get_all <- POST(
    url = api_url,
    encode = "form",
    body = list(
      licenseType = "",
      licenseNumber = "",
      licenseeName = "",
      licenseStatus = "",
      searchBeginDate = "",
      searchEndDate = "",
      sortName = "LicenseNumber",
      pageIndex = 1,
      # request all of the rows on one page
      pageSize = content(find_n)$recordCount
    )
  )
  # pull the json results as a csv and parse 
  all_got <- content(get_all, simplifyDataFrame = TRUE)
  dcl <- as_tibble(all_got$licenseSearchDetailsList)
  dcl <- dcl %>% 
    clean_names(case = "snake") %>% 
    mutate(
      across(where(is.character), ~na_if(str_squish(.), "")),
      across(ends_with("_date"), parse_date, "%m/%d/%Y")
    )
  # save for future use
  write_csv(dcl, raw_csv, na = "")
  rm(find_n, get_all, all_got)
} else {
  # read if already exists
  dcl <- read_delim(
    file = raw_csv,
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    name_repair = make_clean_names,
    # locale = locale(date_format = "%m/%d/%Y"),
    col_types = cols(
      .default = col_character(),
      effective_date = col_date(),
      license_expiration_date = col_date(),
      initial_issue_date = col_date()
    )
  )
}
```

## Explore

There are `r comma(nrow(dcl))` rows of `r ncol(dcl)` columns. Each record
represents a single occupational & professional License

```{r glimpse}
glimpse(dcl)
tail(dcl)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(dcl, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("initial_issue_date", "licensee_name", "license_type")
dcl <- flag_na(dcl, all_of(key_vars))
sum(dcl$na_flag)
```

All `r sum(dcl$na_flag)` rows missing a `licensee_name` are cosmetology 
apprentices that were originally issued on January 1, 2001. Suspicious.

```{r na-view}
dcl %>% 
  filter(na_flag) %>% 
  count(initial_issue_date, licensee_name, license_type)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
dcl <- flag_dupes(dcl, -license_number)
sum(dcl$dupe_flag)
```

```{r dupe-view}
dcl %>% 
  filter(dupe_flag) %>% 
  count(initial_issue_date, licensee_name, license_type, sort = TRUE)
```

### Categorical

```{r distinct-count}
col_stats(dcl, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(dcl, license_type)
explore_plot(dcl, license_status)
```

### Dates

We can add the calendar year from date with `lubridate::year()`

```{r}
dcl <- dcl %>% 
  mutate(
    date_fix = initial_issue_date %>% 
      na_if("1900-01-01") %>% 
      na_if("1901-01-01"),
    year_fix = year(date_fix)
  )
```

```{r date-range}
min(dcl$date_fix, na.rm = TRUE)
sum(dcl$year_fix < 2000, na.rm = TRUE)
max(dcl$date_fix, na.rm = TRUE)
sum(dcl$date_fix > today(), na.rm = TRUE)
```

```{r bar-year, echo=FALSE}
dcl %>% 
  count(year_fix) %>% 
  filter(year_fix > 1950) %>% 
  ggplot(aes(x = year_fix, y = n)) +
  geom_col(aes(fill = n), color = "black") + 
  scale_fill_viridis_c(option = "C", guide = "none") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2020, by = 10)) +
  theme(legend.position = "bottom") +
  labs(
    title = "District Of Columbia Licenses Issued by Year",
    caption = "Source: DCRA",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(dcl, 1000))
```

1. There are `r comma(nrow(dcl))` records in the database.
1. There are `r comma(sum(dcl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dcl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-dir}
clean_dir <- dir_create(here("state", "dc", "licenses", "data", "clean"))
clean_csv <- path(clean_dir, glue("dc_licenses_19500101-20220531.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(dcl, clean_csv, na = "")
write_rds(dcl, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
