---
title: "Georgia Lobbyying Registration"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ga", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  readxl, #read excel files
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects
[ethics]:https://media.ethics.ga.gov/search/Campaign/Campaign_ByContributions.aspx
## Data

Lobbying data is obtained from the [Georgia Government Transparency and Campaign Finance Commission][ethics].
We use the Lobbyist Roster under the **Search Lobbyist** feature to return a list of lobbyists. There're two options for the data output, organized by group or lobbyist. We will use the Lobbyist Group Roster so that the data is organized by clients in relation to the lobbyists they employed.

Note that the data is separated by year, and we will use `httr` to obtain each year's data via a `POST` request, whose parameters can be found from network activities.

## Download
We can set up the raw directory.
```{r}
raw_dir <- dir_create(here("ga", "lobby", "data", "raw","reg"))
raw_lob_paths <- path(raw_dir, glue("ga_lob_{2004:2020}.xls"))
```
The curl command is preprocessed to include `listGroup={c(LGR,LR)}` and `listYear={2004:2020}`
```{r download, eval=FALSE}
ga_lob_curls <- glue(read_file(path(raw_dir,"ga_lob_curl.sh")))

ga_lob_download <- function(ga_curl,curl_type) { 
  #listGroup=LGR for lobbyist roster, and listGroup=LR for group roster
  #curl_type <- ifelse(test = str_detect(ga_curl,"listGroup=LGR"),yes = "lob", no = "grp")
  this_year <- ga_curl %>% str_extract("(?<=listYear=)\\d{4}")
  raw_path <- path(raw_dir, glue("ga_{curl_type}_{this_year}.xls"))
  if (!this_file_new(raw_path)) {
  write_lines(
  # x = system(paste("bash", path(raw_dir,"ga_curl.sh")), intern = TRUE),
  x = system(ga_curl, intern = TRUE),
  path = raw_path
)
  }
}

map2(ga_lob_curls,"lob",ga_lob_download)

### the 2001-2003 xls file can be accessed via a GET request
download.file("https://media.ethics.ga.gov/search/Lobbyist/Exp/Ros_2003_2001.xls", path(raw_dir,"ga_lob_0103.xls"))
```


We can view the file details here.
```{r}
dir_info(raw_dir)
```

We can see that for year 2004 and 2005 the files are actually empty, so we'll disregard these two when reading.

### Read
The 2001—2003 data has rows to escape. So we'll read the two datasets separately. The file's also not in a standard Excel format, so we will use `rvest` to scrape the html content. We'll also add the year to each dataframe. 
```{r raw_read}
ga_lob_read <- function(ga_path){
  year_from_file <- str_extract(ga_path,"20\\d{2}")
  lob_type <- str_extract(ga_path,"(?<=_)\\w{3}(?=_\\d+)")
  lob_file <- path(raw_dir, glue("ga_{lob_type}_{year_from_file}.csv"))
if (file_exists(lob_file)) {
  message("File for year {year_from_file} already converted, skipping")
} else {
  message(glue("Start converting file for year {year_from_file}"))
  ga_html <- ga_path %>% read_html()
  ga_node <- ga_html %>% html_node("table")
  ga_table <- ga_node %>% html_table()
  names(ga_table) <- ga_table[1,]
  ga_table <- ga_table[2:nrow(ga_table),]
  ga_table <- ga_table %>% 
    mutate(Year = year_from_file)
    write_csv(
    x = ga_table,
    path = lob_file
  )
    message(glue("Conversion completed for year {year_from_file}"))
}
}

ga_lob <- map_dfr(raw_lob_paths[3:length(raw_lob_paths)], ga_lob_read)
```



```{r}
galr <- dir_ls(raw_dir,regexp = ".*lob_.*.csv") %>% 
  map_dfr(read_csv,col_types = cols(.default = col_character()))

galr_early <- dir_ls(raw_dir, regexp = ".+0103.xls") %>% read_xls(skip = 1,col_types = "text")
```


According to the website, the `PaymentExceeds` column is a logical vector indicating whether payments have exceeded $10,000.

We will also need to rename some columns in the 01-03 files to keep it consistent with the later files. 
```{r}
galr_early <- galr_early %>% 
  mutate(Year = str_sub(`Docket Year`, start = 1L, end = 4L),
         Registered = excel_numeric_to_date(as.numeric(Registered))) %>% 
  rename(Association=`Association Name`,
         DateRegistered = Registered,
         FirstName = First,
         MiddleName = Middle,
         DocketYear = `Docket Year`,
         Address1 = `Address 1`,
         LastName = Last)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
galr <- flag_dupes(galr, everything())
galr_early <- flag_dupes(galr_early, everything())
sum(galr_early$dupe_flag)
```

## Explore
```{r glimpse}
glimpse(galr)
tail(galr)
glimpse(galr_early)
tail(galr_early)
```


### Missing

```{r na_count}
col_stats(galr, count_na)
col_stats(galr_early, count_na)
```

```{r na_flag}
galr <- galr %>% flag_na(FilerID, FirstName,LastName, DateRegistered,Association)
galr_early <- galr_early %>% flag_na(FirstName,LastName, DateRegistered,Association)
sum(galr$na_flag)
sum(galr_early$na_flag)
```

```{r na_view}
galr %>% 
  filter(na_flag) %>% 
  select(FilerID, FirstName,LastName, DateRegistered,Association)

galr_early %>% 
  filter(na_flag) %>% 
  select(FirstName,LastName, DateRegistered,Association)
```


### Categorical

```{r distinct_count}
col_stats(galr, n_distinct)
```

### Dates

We can examine the validity of `date_clean`. It looks pretty clean.
```{r date_range}
min(galr$DateRegistered)
max(galr$DateRegistered)
sum(galr$DateRegistered > today())
```

```{r date_range early}
min(galr_early$DateRegistered)
max(galr_early$DateRegistered)
sum(galr_early$DateRegistered > today())
```

```{r bar_year, echo=FALSE}
galr <- galr %>% mutate(Year = as.integer(Year))
galr_early <- galr_early %>% mutate(Year = as.integer(Year))

galr %>%
  count(Year) %>% 
  bind_rows(galr_early %>% count(Year)) %>% 
  mutate(even = is_even(Year)) %>% 
  ggplot(aes(x = Year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 2)) +
  coord_cartesian(xlim = c(1989, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Georgia Lobbyists Registration by Year",
    caption = "Source: Georgia Government Transparency and Campaign Finance Commission",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.
To normalize the lobbyist phone number variable, will will combine the number and extension with
`tidyr::unite()` and pass the united string to `campfin::normal_phone()`.

```{r phone_normal}
galr <- galr %>% 
  unite(
    Phone, PhoneExt,
    col = "Phone_norm",
    sep = "x",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    Phone_norm = normal_phone(
      number = "Phone_norm",
      na_bad = FALSE,
      rm_ext = FALSE
    )
  )

galr_early <- galr_early %>% 
  mutate(
    Phone_norm = normal_phone(
      number = "Phone",
      na_bad = FALSE,
      rm_ext = FALSE
    )
  )
```

```{r phone_view, echo=FALSE}
galr %>% 
  select(Phone, PhoneExt, Phone_norm) %>% 
  distinct() %>% 
  sample_frac()

galr_early %>% 
  select(Phone, Phone_norm) %>% 
  distinct() %>% 
  sample_frac()
```


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
galr <- galr %>% 
  # combine street addr
  unite(
    col = address_full,
    starts_with("Address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
   address_norm = normal_address(
     address = address_full,
     abbs = usps_street,
     na = invalid_city
   )
 ) 

galr_early <- galr_early %>% 
  # combine street addr
  unite(
    col = address_full,
    starts_with("Address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
   address_norm = normal_address(
     address = address_full,
     abbs = usps_street,
     na = invalid_city
   )
 )
```

```{r address_view}
galr %>% 
  select(address_full, address_norm) %>% 
  distinct() %>% 
  sample_n(10)

galr_early %>% 
  select(address_full, address_norm) %>% 
  distinct() %>% 
  sample_n(10)

galr <- galr %>% 
  select(-address_full)

galr_early <- galr_early %>% 
  select(-address_full)

```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
galr <- mutate(
  .data = galr,
  zip_norm = normal_zip(
    zip = Zip,
    na_rep = TRUE
  )
)

galr_early <- mutate(
  .data = galr_early,
  zip_norm = normal_zip(
    zip = Zip,
    na_rep = TRUE
  )
)

```

```{r zip_progress}
progress_table(
  galr$Zip,
  galr_early$Zip,
  galr$zip_norm,
  galr_early$zip_norm,
  compare = valid_zip
)
```

### State
The two-letter state abbreviations are almost valid and don't need to be normalized. 

```{r state_norm}
prop_in(galr$State, valid_state, na.rm = T)
prop_in(galr_early$State, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
galr <- galr %>% 
  mutate(
    city_norm = normal_city(
      city = City, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )

galr_early <- galr_early %>% 
  mutate(
    city_norm = normal_city(
      city = City, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )
many_city <- c(valid_city, extra_city)
prop_in(galr_early$city_norm,many_city)
```

#### Progress

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(galr$City),
  galr$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
galr <- galr %>% 
  rename_all(~str_replace(., "_norm", "_clean")
             )
galr_early <- galr_early %>% 
  rename_all(~str_replace(., "_norm", "_clean")
             )
```

```{r clean_glimpse}
glimpse(sample_n(galr, 20))
glimpse(sample_n(galr_early, 20))
```

```{r total count}
nrow(galr)
nrow(galr_early)
```


1. There are `r comma(nrow(galr)+nrow(galr_early))` records in the database.
1. There are `r comma(sum(galr$dupe_flag))` duplicate records in the database.
1. The range and distribution of `date` seem reasonable.
1. There are `r comma(sum(galr$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ga", "lobby", "data", "clean","reg"))
clean_path <- path(clean_dir, "ga_lobby_reg_clean_2006_2020.csv")
early_path <- path(clean_dir, "ga_lobby_reg_clean_2001_2003.csv")
write_csv(galr, clean_path, na = "")
write_csv(galr_early, early_path, na = "")
file_size(clean_path)
file_size(early_path)
```

## Upload

Using the `aws.s3` package, we can upload the file to the IRW server.

```{r upload,eval=FALSE}
s3_path <- path("csv", basename(clean_path))
s3_path_early <- path("csv", basename(early_path))
put_object(
  file = clean_path,
  object = s3_path, 
  bucket = "publicaccountability",
  acl = "public-read",
  multipart = TRUE,
  show_progress = TRUE
)
put_object(
  file = early_path,
  object = s3_path_early, 
  bucket = "publicaccountability",
  acl = "public-read",
  multipart = TRUE,
  show_progress = TRUE
)
as_fs_bytes(object_size(s3_path, "publicaccountability"))
as_fs_bytes(object_size(s3_path_early, "publicaccountability"))
```


## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(galr)),
  type = md_code(map_chr(galr, typeof)),
  def = c(
    "ID of the filer (lobbyist)",
    "Lobbyist last name",
    "Lobbyist name suffix",
    "Lobbyist first name",
    "Lobbyist middle name",
    "Lobbyist street address line 1",
    "Lobbyist street address line 2",
    "Lobbyist City",
    "Lobbyis State",
    "Lobbyist ZIP code",
    "Normalized Lobbyist phone",
    "Lobbyist phone",
    "Lobbyist phone extension",
    "Secondary lobbyist phone",
    "Lobbyist email",
    "Organization to which lobbyists were associated",
    "Payment exceeds $10,000",
    "Date registered",
    "Date terminated",
    "Level of lobbying activity",
    "Year of data publication",
    "Flag for missing date, organization, or, filerID or name",
    "Flag for completely duplicated record",
    "Normalized lobbyist street address",
    "Normalized 5-digit ZIP code",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Georgia Lobbying Registration Data Dictionary\n", dict_md),
  path = here("ga", "lobby", "ga_contribs_dict.md"),
)
```
