---
title: "Michigan Lobbyists"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mi", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][01] GitHub repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Registration

Data is obtained from the [Michigan Secretary of State][sos]. The data is
provided by NICUSA, Inc., which provides information technology services for the
SOS. The data accessed on April 9, 2023 has a total of 8,261 records. This update includes all the rows in the last dataset. The next update will likely replace the current dataset. 

> Use this page to search for information concerning individuals and
organizations registered under the Michigan Lobby Registration Act, as well as
expenditures disclosed by these registrants on required financial reports. The
record for each registrant will also include a listing of any reported employees
compensated by each registrant for lobbying, as well as employers who report the
registrant as an employee compensated for lobbying on behalf of the employer.
>
> You may download the entire list of Michigan registrants by clicking on the
Spreadsheet Format box and following the instructions provided.

The website certificates are out of date, so we can only obtain the file by
using `curl` with the `--insecure` option.

[sos]: https://www.michigan.gov/sos/

```{r raw_dir}
raw_dir <- here("state","mi", "lobby", "data", "raw")
dir_create(raw_dir)
```

```{r raw_curl}
# changes from time to time
#lob_url <- "https://miboecfr.nictusa.com/cfr/dumpdata/aaaMxayCb/mi_lobby.sh"
lob_url <- "https://miboecfr.nictusa.com/cfr/dumpdata/aaaeNaqWQ/mi_lobby.sh"
lob_path <- url2path(lob_url, raw_dir)
if (!file_exists(lob_path)) {
  download.file(
    url = lob_url,
    destfile = lob_path,
    method = "curl",
    extra = "--insecure"
  )
}
```

### Vars

```{r vars, echo=FALSE}
var_names <- c(
    "id", "type", "last", "first", "mi", "sig", "addr", 
    "city", "state", "zip", "phone", "reg", "term"
  )
vars <- tribble(
  ~Name, ~Description,
  "Bureau Lobby ID#", "Unique Bureau ID# of this Lobbyist or Agent",
  "Lobby Type", "Type of Lobby (A = Agent, L = Lobbyist)",
  "Lobby Last Name or Organization", "Last or Full Name of the Individual or Lobby Organization",
  "Lobby Individual First Name", "First Name of the Individual Lobbyist or Agent",
  "Lobby Individual Middle Name", "Middle Name of the Individual Lobbyist or Agent",
  "Lobby Signatory", "Official Signatory or Contact Person for this Lobby",
  "Lobby Mailing Address", "Mailing Street Address of this Lobby",
  "Lobby Mailing City", "Mailing City of this Lobby",
  "Lobby Mailing State", "Mailing State of this Lobby",
  "Lobby Mailing Zipcode", "Mailing Zipcode of this Lobby",
  "Lobby Phone#", "Phone Number of this Lobby",
  "Lobby Active Date", "Date this Lobby became an Active Lobbyist or Agent",
  "Lobby Termination Date", "Date this Lobby Terminated all Lobbying activity" 
)
vars %>% 
  mutate(Variable = md_code(var_names)) %>% 
  select(Variable, Description) %>% 
  kable(format = "markdown")
```

### Import

As described on the [data website][raw]:

> #### Other Notes:...
> The file is TAB delimited and NO quotes surround string text.
>
> The first record DOES contain the field names.
>
> The second record is a 'dummy' record used primarily to clue database programs
like Access in as to how to import the data, as well as some other useful
information. You may want to delete this record AND the record(s) at the end of
the file containing counts once you have gotten any use from them.
>
> When saving the mi_lobby.sh file, you may want to rename it with an extension
of .txt, so that certain database programs will import it correctly. The Bureau
of Elections makes every effort to provide accurate information to the public.
However, any data taken from the database should be verified against the actual
report filed by the lobby. The information provided here is deemed reliable but
not guaranteed.

We can use this information to define the parameters of `readr::read_delim()`.

[raw]: https://miboecfr.nicusa.com/cgi-bin/cfr/lobby_srch_res.cgi

```{r raw_read}
milr <- read_delim(
  file = lob_path,
  delim = "\t",
  skip = 2,
  col_names = var_names,
  col_types = cols(
    .default = col_character(),
    type = col_factor(),
    reg = col_date_mdy(),
    term = col_date_mdy()
  )
)
```

### Explore

```{r raw_glimpse}
head(milr)
tail(milr)
glimpse(sample_frac(milr))
```

As we can see from `tail()`, the last two rows still need to be removed.

The `id` variable is unique to each lobbyist, so we can use it to remove the
summary rows at the bottom of the file.

```{r n_distinct}
col_stats(milr, n_distinct)
```

```{r raw_filter}
milr <- filter(milr, !is.na(id))
```

Now, there are no rows missing the key information needed to identify lobbyists.

```{r count_na}
col_stats(milr, count_na)
```

There are no duplicate rows in the database.

```{r duplicated}
sum(duplicated(milr))
```

The database contains both outside lobbyist and lobbying agents.

```{r plot_agent, echo=FALSE}
explore_plot(
  data = milr,
  var = type
) + labs(title = "Michigan Lobbyist Types",
  subtitle = "A = Agent, L = Lobbyist")
```

`r percent(prop_na(milr$term))` of lobbyists in the database have a termination 
date, meaning only `r percent(1-prop_na(milr$term))` of the records identify 
active lobbyists.

```{r term_na}
prop_na(milr$term)
```

We can add the registration year using `lubridate::year()` on the date column.

```{r year_add}
milr <- mutate(milr, year = year(reg))
```

```{r year_plot, echo=FALSE}
milr %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  labs(
    title = "Michigan Lobbyists Registered per Year",
    x = "Year Registered",
    y = "Lobbyists"
  )
```

### Wrangle

To improve the searchability and consistency of the database, we can perform
some very basic and confident text normalization.

#### Phone

We can convert the phone numbers into a standard charatcer (i.e., non-numeric) format.

```{r phone_norm}
milr <- mutate(milr, phone_norm = normal_phone(phone))
```

```{r phone_view, echo=FALSE}
milr %>% 
  select(starts_with("phone")) %>% 
  distinct() %>% 
  sample_frac()
```

#### Address

We can use `campfin::normal_address()` to improve the consistency in the `addr` variable.

```{r addr_norm}
milr <- mutate(milr, addr_norm = normal_address(addr, abbs = usps_street))
```

```{r addr_view, echo=FALSE}
milr %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_frac()
```

#### ZIP

```{r zip_norm}
milr <- mutate(milr, zip_norm = normal_zip(zip, na_rep = TRUE))
```

```{r zip_view, echo=FALSE}
milr %>% 
  select(starts_with("zip")) %>% 
  distinct() %>% 
  sample_frac()
```

```{r zip_progress}
progress_table(
  milr$zip,
  milr$zip_norm,
  compare = valid_zip
)
```

#### State

The `state` variable does not need to be cleaned.

```{r state_prop}
prop_in(milr$state, valid_state)
```

#### City

```{r city_norm}
milr <- mutate(
  .data = milr, 
  city_norm = normal_city(
    city = city, 
    abbs = usps_city, 
    na = invalid_city
  )
)
```

```{r city_swap}
milr <- milr %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(city_match) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_abb,
    -match_dist
  )
```

```{r city_filter}
out <- milr %>% 
  filter(city_swap %out% valid_city) %>% 
  count(city_swap, state, sort = TRUE) %>% 
  drop_na()
```

```{r check_combine}
many_city <- c(valid_city, extra_city)
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  milr$city_raw,
  milr$city_norm,
  milr$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

### Export

```{r clean_write}
clean_dir <- dir_create(here("state","mi", "lobby", "data", "clean"))
clean_path <- path(clean_dir, "mi_lobby_reg.csv")
write_csv(milr %>% rename(city_clean = city_swap) %>% select(-city_norm), path = clean_path, na = "")
```

```{r clean_check}
file_size(clean_path)
guess_encoding(clean_path)
```

## Contributions

Lobbying expenditure data be queried as [itemized lobby expenditure analysis](https://miboecfr.nictusa.com/cgi-bin/cfr/lobby_exp_anls.cgi). 

```{r exp_down}
exp_url <- "https://miboecfr.nictusa.com/cfr/dumpall/miloball.sh"
exp_path <- url2path(exp_url, raw_dir)
if (!file_exists(exp_path)) {
  download.file(
    url = exp_url,
    destfile = exp_path,
    method = "curl",
    extra = "--insecure"
  )
}
```

```{r echo=FALSE}
exp_names <- c(
  "rpt_year",
  "rpt_type",
  "lob_last",
  "lob_first",
  "lob_mi",
  "lob_type",
  "lob_id",
  "exp_type",
  "po_title",
  "po_last",
  "po_first",
  "po_mi",
  "lob_why",
  "exp_date",
  "exp_amt",
  "ytd_fb",
  "doc_id"
)
```

```{r exp_read}
mile <- read_delim(
  file = exp_path,
  delim = "\t",
  skip = 2,
  col_names = exp_names,
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    rpt_year = col_integer(),
    exp_date = col_date_usa(),
    exp_amt = col_double(),
    ytd_fb = col_double()
  )
) %>% 
  filter(!is.na(rpt_year))
```

```{r exp_glimpse}
head(mile)
tail(mile)
glimpse(sample_frac(mile))
```

```{r}
col_stats(mile, count_na)
```

```{r}
col_stats(mile, n_distinct)
```

```{r lob_exp_amt, echo=FALSE}
mile %>%
  ggplot(aes(exp_amt)) +
  geom_histogram(aes(fill = lob_type)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Michigan Lobbyist Expense Amounts",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~lob_type, ncol = 1)
```

```{r lob_exp_year, echo=FALSE}
mile %>% 
  count(rpt_year) %>% 
  mutate(odd = !is_even(rpt_year)) %>% 
  ggplot(aes(rpt_year, n)) +
  geom_col(aes(fill = odd)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(
    title = "Michigan Lobbyist Expense Counts per Year",
    x = "Report Year",
    y = "Count"
  )
```


```{r year and state}
mile <- mile %>% rename(year = rpt_year) %>% mutate(state = "MI")
```

```{r}
write_csv(
  x = mile,
  path = path(clean_dir, "mi_lobby_exp.csv"),
  na = ""
)
```

