---
title: "Chicago Contracts"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "il", "chicago", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  jsonlite, # read json files
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  readxl, # read excel files
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/il/chicago/contracts/docs/chicago_contracts_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

> Contracts and modifications awarded by the City of Chicago since 1993. This
data is currently maintained in the City’s Financial Management and Purchasing
System (FMPS), which is used throughout the City for contract management and
payment.
>
> Legacy System Records: Purchase Order/Contract Numbers that begin with alpha
characters identify records imported from legacy systems. Records with a null
value in the Contract Type field were imported from legacy systems.
>
> "Comptroller-Other" Contract Type: Some records where the Contract Type is
"COMPTROLLER-OTHER" are ordinance-based agreements and may have start dates
earlier than 1993.
>
> Depends Upon Requirements Contracts: If the contract Award Amount is $0, the
contract is not cancelled, and the contract is a blanket contract, then the
contract award total Depends Upon Requirements. A Depends Upon Requirements
contract is an indefinite quantities contract in which the City places orders as
needed and the vendor is not guaranteed any particular contract award amount.
>
> Blanket vs. Standard Contracts: Only blanket contracts (contracts for repeated
purchases) have FMPS end dates. Standard contracts (for example, construction
contracts) terminate upon completion and acceptance of all deliverables. These
dates are tracked outside of FMPS.
>
> Negative Modifications: Some contracts are modified to delete scope and money
from a contract. These reductions are indicated by negative numbers in the Award
Amount field of this dataset.
>
> Data Owner: Procurement Services.  
> Time Period: 1993 to present.  
> Frequency: Data is updated daily.  

## Download

```{r raw-dir}
raw_url <- "https://data.cityofchicago.org/api/views/rsxa-ify5/rows.csv"
raw_dir <- dir_create(here("state", "il", "contracts", "data", "raw"))
raw_csv <- path(raw_dir, basename(raw_url))
```

```{r raw-download}
if (!file_exists(raw_csv)) {
  download.file(raw_url, raw_csv)
}
```

## Read

```{r raw-read}
chic <- read_delim(
  file = raw_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  locale = locale(date_format = "%m/%d/%Y"),
  col_types = cols(
    .default = col_character(),
    `Start Date` = col_date(),
    `End Date` = col_date(),
    `Approval Date` = col_date(),
    `Award Amount` = col_double()
  )
)
```

```{r clean-names}
chic <- clean_names(chic, case = "snake")
```

## Explore

There are `r comma(nrow(chic))` rows of `r ncol(chic)` columns. Each record
represents a single contract between the city of Chicago and an outside vendor
for goods and services.

```{r glimpse}
glimpse(chic)
tail(chic)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(chic, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("approval_date", "department", "award_amount", "vendor_name")
chic <- flag_na(chic, all_of(key_vars))
sum(chic$na_flag)
```

```{r na-view}
chic %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
chic <- flag_dupes(chic, everything())
sum(chic$dupe_flag)
```

```{r dupe-view}
chic %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(approval_date)
```

### Categorical

```{r distinct-count}
col_stats(chic, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(chic, contract_type) + scale_x_truncate()
explore_plot(chic, department) + scale_x_truncate()
explore_plot(chic, procurement_type) + scale_x_truncate()
```

### Amounts

```{r amount-round}
# fix floating point precision
chic$award_amount <- round(chic$award_amount, digits = 2)
```

```{r amount-summary}
summary(chic$award_amount)
mean(chic$award_amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(chic[c(which.max(chic$award_amount), which.min(chic$award_amount)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
chic %>%
  ggplot(aes(award_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Chicago Contracts Amount Distribution",
    caption = "Source: Chicago Data Portal",
    x = "Amount",
    y = "Count"
  )
```

```{r dept-sum, echo=FALSE}
chic %>%
  group_by(department) %>% 
  summarise(award_amount = sum(award_amount) / 1e9) %>% 
  arrange(desc(award_amount)) %>% 
  head() %>% 
  ggplot(aes(reorder(department, -award_amount), award_amount)) +
  geom_col(aes(fill = award_amount), color = "black") +
  scale_y_continuous(labels = comma) +
  scale_x_wrap() +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Chicago Contracts Amount by Department",
    caption = "Source: Chicago Data Portal",
    x = "Department",
    y = "Amount Sum ($Bil.)"
  )
```

```{r vend-sum, echo=FALSE}
chic %>%
  group_by(vendor_name) %>% 
  summarise(award_amount = sum(award_amount) / 1e9) %>% 
  arrange(desc(award_amount)) %>% 
  head() %>% 
  ggplot(aes(reorder(vendor_name, -award_amount), award_amount)) +
  geom_col(aes(fill = award_amount), color = "black") +
  scale_y_continuous(labels = comma) +
  scale_x_wrap() +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Chicago Contracts Amount by Vendor",
    caption = "Source: Chicago Data Portal",
    x = "Vendor",
    y = "Amount Sum ($Bil.)"
  )
```

### Dates

We can add the calendar year from `approval_date` with `lubridate::year()`.

```{r date-year}
chic <- mutate(chic, approval_year = year(approval_date))
```

```{r date-range}
min(chic$approval_date, na.rm = TRUE)
sum(chic$approval_year < 2000, na.rm = TRUE)
max(chic$approval_date, na.rm = TRUE)
sum(chic$approval_date > today(), na.rm = TRUE)
```

```{r bar-year, echo=FALSE}
chic %>% 
  count(approval_year) %>% 
  mutate(even = is_even(approval_year)) %>% 
  ggplot(aes(x = approval_year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1990, 2022, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Chicago Contracts by Year",
    caption = "Source: Chicago Data Portal",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

The `address_*`, `city`, and `state` variables are all already fairly normalized
and most of the "bad" addresses are foreign, so they shouldn't be changed.

We also needed to add fields for the city and state of the department, which
will be Chicago and Illinois in every instance.

```{r add-geo}
chic <- chic %>% 
  mutate(
    dept_city = "CHICAGO",
    dept_state = "IL"
  )
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
chic <- chic %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  chic$zip,
  chic$zip_norm,
  compare = valid_zip
)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(chic, 1000))
```

1. There are `r comma(nrow(chic))` records in the database.
1. There are `r comma(sum(chic$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(chic$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(chic$approval_date, na.rm = TRUE), "-")
max_dt <- str_remove_all(max(chic$approval_date, na.rm = TRUE), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("state", "il", "chicago", "contracts", "data", "clean"))
clean_csv <- path(clean_dir, glue("il-chicago_contracts_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(chic, clean_csv, na = "")
write_rds(chic, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
