---
title: "Texas Lobbying"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("tx", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  readxl, # read excel files
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The data can be obtained from the [Texas Ethics Commission][tec]. The download page lists data files from 2001 and on. Since the older files are in a format relatively hard to wrangle. We here include only lobbying records from 2016 on. The 2023 file was updated on 2023-08-04.

[tec]:https://www.ethics.state.tx.us/data/search/lobby/

## Import

### Download

```{r create_raw_dir}
raw_dir <- dir_create(here("state","tx", "lobby", "data", "raw"))
```

```{r download_raw}
txl_urls <- c(
  "https://www.ethics.state.tx.us/data/search/lobby/2016/2016LobbyistGroupByLobbyist.nopag.xlsx",
  "https://www.ethics.state.tx.us/data/search/lobby/2017/2017LobbyistGroupByLobbyist.xlsx",
  glue("https://www.ethics.state.tx.us/data/search/lobby/{2018:2023}/{2018:2023}LobbyGroupByLobbyist.xlsx")
)

if (!all_files_new(raw_dir)) {
  for (xlsx_url in txl_urls) {
    download.file(
      url = xlsx_url,
      destfile = path(raw_dir, basename(xlsx_url))
    )
  }
}
```

### Read

```{r read_raw_xlsx}
txl <- map_df(
  .x = dir_ls(raw_dir), 
  .f = read_excel,
  col_types = "text"
)

txl <- txl %>% 
  clean_names("snake") %>% 
  rename(
    lob_id = filer_id,
    lob_name = filer_name,
    lob_biz = business,
    lob_addr1 = addr_1_4,
    lob_addr2 = addr_2_5,
    lob_city = city_6,
    lob_state = state_7,
    lob_zip = zip_8,
    pri_name = client_name,
    pri_addr1 = addr_1_10,
    pri_addr2 = addr_2_11,
    pri_city = city_12,
    pri_state = state_13,
    pri_zip = zip_14
  )
```

```{r raw_parse}
txl <- mutate_at(txl, vars(begin, stop), ~parse_date(., "%m/%d/%Y"))
txl <- mutate_at(txl, vars(lob_name), ~str_remove(., "\\s\\(.*\\)$"))
```

## Explore

```{r glimpse}
head(txl)
tail(txl)
glimpse(sample_n(txl, 20))
```

### Missing

```{r na_count}
col_stats(txl, count_na)
```

```{r na_flag}
txl <- txl %>% flag_na(lob_name, pri_name)
percent(mean(txl$na_flag), 0.1)
```

```{r na_view}
txl %>% 
  filter(na_flag) %>% 
  select(lob_name, pri_name) %>% 
  sample_frac()
```

### Duplicates

Most of the duplicate rows come from a repeated single variable and missing
identifying variables (like the date). We will flag them nonetheless.

```{r dupe_flag}
txl <- flag_dupes(txl, everything(), .check = TRUE)
percent(mean(txl$dupe_flag), 0.1)
```

```{r dupe_view}
txl %>% 
  filter(dupe_flag) %>% 
  select(lob_name, pri_name, begin, stop, na_flag) %>% 
  arrange(lob_name)
```

### Categorical

```{r n_distinct}
col_stats(txl, n_distinct)
```

```{r bar_distinct, fig.keep=FALSE, echo=FALSE}
explore_plot(txl, reporting_interval)
explore_plot(txl, method)
explore_plot(txl, amount)
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

We will also add a single year variable.

```{r year_add}
txl <- mutate(txl, year = year(begin))
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r lob_address_norm}
txl <- txl %>% 
  unite(
    col = lob_addr,
    starts_with("lob_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    lob_addr_norm = normal_address(
      address = lob_addr,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-lob_addr)
```

```{r pri_address_norm}
txl <- txl %>% 
  unite(
    col = pri_addr,
    starts_with("pri_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    pri_addr_norm = normal_address(
      address = pri_addr,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-pri_addr)
```

```{r address_view}
txl %>% 
  select(contains("lob_addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
txl <- mutate_at(
  .tbl = txl,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_progress}
progress_table(
  txl$lob_zip,
  txl$lob_zip_norm,  
  txl$pri_zip,
  txl$pri_zip_norm,
  compare = valid_zip
)
```

### State

```{r}
prop_in(txl$lob_state, valid_state)
prop_in(txl$pri_state, valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
txl <- mutate_at(
  .tbl = txl,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  states = c("TX", "DC", "TEXAS"),
  na = invalid_city,
  na_rep = TRUE
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r lob_city_swap}
txl <- txl %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lob_state" = "state",
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lob_city_norm, city_match),
    match_dist = str_dist(lob_city_norm, city_match),
    lob_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r pri_city_swap}
txl <- txl %>% 
  left_join(
    y = zipcodes,
    by = c(
      "pri_state" = "state",
      "pri_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(pri_city_norm, city_match),
    match_dist = str_dist(pri_city_norm, city_match),
    pri_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = pri_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress_table(
  str_to_upper(txl$lob_city),
  txl$lob_city_norm,
  txl$lob_city_swap,
  str_to_upper(txl$pri_city),
  txl$pri_city_norm,
  txl$pri_city_swap,
  compare = many_city
)
```

## Conclude

```{r clean_select}
txl <- txl %>% 
  select(
    -lob_city_norm,
    -pri_city_norm,
  ) %>% 
  rename_all(~str_replace(., "_(norm|swap)", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(txl, 20))
```

1. There are `r comma(nrow(txl))` records in the database.
1. There are `r comma(sum(txl$dupe_flag))` duplicate records in the database.
1. There are `r comma(sum(txl$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("state","tx", "lobby", "data", "clean"))
```

```{r clean_write}
write_csv(
  x = txl,
  path = path(clean_dir, "tx_lobby_reg_2016-2023.csv"),
  na = ""
)
```

