---
title: "Fatal Shootings and Military Equipment"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "leso", "docs"))
```

## Packages

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  readxl, # read excel files
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  usmap, # plot us maps
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

## Police Violence

We can download a database of fatal police shootings from the Washington Post's
[Fatal Force project][fatal].

>  After Michael Brown, an unarmed black man, was killed in 2014 by police in
Ferguson, Mo., a Post investigation found that the FBI undercounted fatal police
shootings by more than half. This is because reporting by police departments is
voluntary and many departments fail to do so.
> 
> The Post’s data relies primarily on news accounts, social media postings and
police reports. Analysis of more than five years of data reveals that the number
and circumstances of fatal shootings and the overall demographics of the victims
have remained relatively constant.

[db]: https://github.com/washingtonpost/data-police-shootings
[fatal]: https://www.washingtonpost.com/graphics/investigations/police-shootings-database/

```{r github_url, echo=FALSE}
github_url <- function(user, repo, file, branch = "master") {
  str_c(
    url = "https://raw.githubusercontent.com",
    user = user, 
    repo = repo,
    branch = branch,
    file = file,
    sep = "/"
  )
}
```

```{r fatal_url}
fatal_url <- github_url(
  user = "washingtonpost", 
  repo = "data-police-shootings",
  file = "fatal-police-shootings-data.csv"
)
```

The Fatal Force database can be read directly from GitHub with `vroom::vroom()`.

```{r fatal_read}
fatal <- vroom(
  file = fatal_url,
  col_types = cols(
    .default = col_character(),
    id = col_integer(),
    date = col_date(),
    age = col_double(),
    signs_of_mental_illness = col_logical(),
    body_camera = col_logical()
  )
)
```

```{r fatal_glimpse}
glimpse(fatal)
```

## Equipment Transfers

The [1033 program][1033] allows for the Department of Defense (DOD) to transfer
excess military equipment to civilian law enforcement agencies.

[1033]: https://en.wikipedia.org/wiki/1033_program

> The program has been criticized over the years by local media, by the Office
of the Inspector General, U.S. Department of Defense in 2003, and by the GAO
which found waste, fraud and abuse. It was not until media coverage of police
during August 2014 Ferguson unrest that the program drew nationwide public
attention; the Ferguson Police Department had equipment obtained through the
1033 program.

An Excel spreadsheet of equipment transfers can be downloaded from the Defense
Logistic Agency's [electronic reading room][err].

[err]: https://www.dla.mil/DispositionServices/FOIA/EFOIALibrary/

```{r raw_dir}
mil_url <- str_c(
  url = "https://www.dla.mil/Portals/104/Documents/DispositionServices/LESO/",
  file = "DISP_AllStatesAndTerritories_06302018.xlsx"
)
mil_path <- file_temp(ext = "xlsx")
download.file(mil_url, mil_path)
```

The Excel spreadsheet lists transfers to police departments to each state in
separate states. We can combine `purrr::map_df()` and `readxl::read_excel()` to
read all the sheets into a single data frame of transfers.

```{r raw_read}
mil_equip <- mil_path %>%
  readxl::excel_sheets() %>%
  purrr::set_names() %>%
  purrr::map_df(
    .f = readxl::read_excel,
    .name_repair = make_clean_names,
    path = mil_path
  )
```

## Compare

We will compare states on their number of fatal police shootings from 2015 to
2020 and the monetary value of military equipment transfers from 1990 to 2018;
our assumption being that value is a useful proxy for the lethality of the
equipment. Our null hypothesis holds that there will be no relationship between the value of transfers per capita and the number of fatal shootings per capita.

To normalize our shooting and transfer values between states, we will use a
database of 2020 state populations from my 'usa' package.

```{r pops}
pops <- usa::states %>% 
  left_join(usa::facts) %>% 
  select(state = abb, population)
```

We will first calculate the total equipment transfer values per capita for each
state. We have not removed any equipment from the 1033 database, which includes
ammunition, small arms and, armored vehicles but also clothing and supplies.

```{r weapons}
mil_equip <- mil_equip %>% 
  group_by(state) %>% 
  summarise(value = sum(acquisition_value), items = n()) %>% 
  left_join(pops, by = "state") %>% 
  mutate(val_per = value/population) %>% 
  select(-population)
```

We can then combine our normalized equipment values to fatal shootings per
capita. 

```{r compare}
compare <- fatal %>% 
  count(state, name = "deaths") %>% 
  left_join(pops) %>% 
  mutate(deaths_100k = deaths/(population/1e5)) %>% 
  select(-population) %>% 
  left_join(mil_equip, by = "state")
```

Using linear regression, we can determine there is in fact a relationship
between the military equipment received by the police departments in a state and
that state's number of fatal police shootings per capita.

```{r lm}
lm_mil <- lm(deaths_100k ~ val_per, data = compare)
summary(lm_mil)
```

With a p-value of `r round(summary(lm_mil)$coefficients[,4][2], 4)`, we are able
to reject our null hypothesis at a 95% confidence interval.

```{r states_plot}
compare %>% 
  ggplot(aes(x = val_per, y = deaths_100k)) +
  # geom_smooth(method = "lm") +
  geom_label(aes(label = state)) +
  scale_x_continuous(labels = scales::dollar) +
  labs(
    title = "Militarization of Police via the 1033 Program",
    x = "Military Equipment Value per capita (1990-2020)",
    y = "Fatal Police Shootings per capita centum (2015-2018)"
  )
```

