---
title: "District Of Columbia Voters"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("dc", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  readxl, # read excel files
  refinr, # cluster & merge
  scales, # format strings
  digest, # hash strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::dr_here(show_reason = FALSE)
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The DC voter registration database can be requested from the Board of Elections
by filling out the [Data Request Form][drf] PDF and submitting it alongside a
small fee, depending on the requested format.

Data was requested by the Investigative Reporting Workshop and received on
September 28, 2020. 

[drf]: https://www.dcboe.org/dcboe/media/PDFFiles/Data_Request_Form.pdf

Two files were provided to the IRW:

1. `D.C. Registered Voters (All).xlsx`
2. `Read Me.txt`

The README file contains a number of tables to describe the data.

```{r readme}
raw_dir <- here("dc", "voters", "data", "raw")
readme <- read_lines(file = path(raw_dir, "Read Me.txt"))
```

#### Columns

```{r raw_cols}
raw_cols <- read_tsv(
  file = readme[3:22],
  col_names = str_split(readme[1], "\t", simplify = TRUE)
)
kable(raw_cols)
```

#### Status

|Code | Meaning         |
|:----|:----------------|
| `A` | ACTIVE          |
| `X` | ACTIVE (ID REQ) | 
| `F` | FEDERAL ONLY    |

#### History

```{r hist_codes}
hist_codes <- tribble(
  ~code, ~action,
  NA,  "NO VOTING HISTORY",
  "V", "POLL",
  "A", "ABSENTEE",
  "N", "NOT ELIGIBLE TO VOTE",
  "E", "ELIGIBLE BUT DID NOT VOTE",
  "Y", "EARLY VOTER"
)
```

```{r echo=FALSE}
kable(hist_codes)
```

## Read

```{r raw_path}
raw_path <- dir_ls(raw_dir, regexp = "xlsx$")
```

This text file can then be easily read as a data frame.

```{r raw_read}
dcv <- read_excel(raw_path, col_types = "text")
```

We can ensure the data was properly read by checking the unique values of
the city column, which should obviously all be Washington.

```{r raw_check}
n_distinct(dcv$RES_CITY)
unique(dcv$RES_CITY)
```

```{r raw_parse}
dcv <- mutate(dcv, across(REGISTERED, ~excel_numeric_to_date(as.numeric(.))))
```

The last 50 columns of the file are the voter's behavior in past elections.
This data will be moved to a new object and converted to a more format that
is more easily analyzed.

```{r vote_write}
hist_file <- path(raw_dir, "dc_vote_hist.tsv.xz")
if (!file_exists(hist_file)) {
  dcv %>% 
    select(FIRSTNAME, LASTNAME, REGISTERED, matches("^\\d")) %>%
    pivot_longer(
      cols = !1:3,
      names_to = "date_type",
      values_to = "vote_code",
      values_drop_na = TRUE
    ) %>% 
    separate(
      col = date_type,
      into = c("elect_date", "elect_type"),
      sep = "-"
    ) %>% 
    mutate(across(elect_date, mdy)) %>% 
    clean_names("snake") %>% 
    write_tsv(xzfile(hist_file))
  rm(vote_hist)
  flush_memory(1)
}
```

```{r raw_rename, echo=FALSE}
dcv <- dcv %>% 
  select(-matches("^\\d")) %>% 
  clean_names("snake") %>% 
  rename_with(~str_remove(., "res_")) %>% 
  rename(
    reg_date = registered,
    last = lastname,
    first = firstname
  )
```

## Explore

There are `r scales::comma(nrow(dcv))` rows of `r ncol(dcv)` columns.

```{r glimpse}
glimpse(dcv)
tail(dcv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(dcv, count_na)
```

No columns are missing the registration date or last name needed to identify a
voter.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
d1 <- duplicated(dcv, fromLast = FALSE)
d2 <- duplicated(dcv, fromLast = TRUE)
dcv <- mutate(dcv, dupe_flag = d1 | d2)
sum(dcv$dupe_flag)
```

```{r dupe_rm, echo=FALSE}
rm(d1, d2)
flush_memory(1)
```

```{r dupe_view}
dcv %>% 
  filter(dupe_flag) %>% 
  select(reg_date, last, party, smd) %>% 
  arrange(reg_date)
```

### Categorical

```{r distinct_count}
col_stats(dcv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=2.5}
explore_plot(dcv, party)
dcv %>% 
  count(ward) %>% 
  ggplot(aes(ward, n)) +
  geom_col(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma)
```

### Dates

We can add the registration year from `reg_date` with `lubridate::year()`.

```{r date_year}
dcv <- mutate(dcv, reg_year = year(reg_date))
```

```{r date_range}
min(dcv$reg_date)
max(dcv$reg_date)
sum(dcv$reg_date > today())
```

```{r bar_year, echo=FALSE}
dcv %>% 
  count(reg_year) %>% 
  mutate(
    elect_type = case_when(
      reg_year %% 4 == 0 ~ "pres.",
      reg_year %% 2 == 0 ~ "mid.",
      TRUE ~ "none"
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = elect_type)) + 
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1968, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  coord_cartesian(xlim = c(1968, 2020)) +
  labs(
    title = "District Of Columbia Voters by Registration Year",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

The `address` variable is already sufficiently normalized and abbreviated,
we can simply map together all the individual columns.

```{r addr_view}
dcv %>% 
  select(house, frac, street, apt) %>% 
  sample_n(20)
```

The city, state, and ZIP code variables are easy because it's a single city.

```{r geo_check}
unique(dcv$city)
unique(dcv$state)
prop_in(dcv$zip, valid_zip)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(dcv, 50))
```

1. There are `r comma(nrow(dcv))` records in the database.
1. There are `r comma(sum(dcv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dcv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("dc", "voters", "data", "clean"))
clean_path <- path(clean_dir, "dc_voters_clean.csv")
write_csv(dcv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(dcv)),
  type = md_code(map_chr(dcv, typeof)),
  def = c(
    "Date registered",
    "Voter last name",
    "Voter first name",
    "Voter middle name",
    "Voter name suffix",
    "Voter status (Active, Fed, ID Required)",
    "Political party",
    "House number",
    "House fraction",
    "Apartment number",
    "Street name",
    "City name (Washington)",
    "State (DC)",
    "ZIP code",
    "ZIP+4 code",
    "Precinct number",
    "Ward number (1-8)",
    "Advisory Neighborhood Commission code",
    "Single Member District code",
    "Flag indicating duplicate record",
    "Calendar year registered"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
