---
title: "Nebraska Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ne", "lobbying", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  pdftools, # read pdf file text
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

A list of registered lobbyists can be obtained from the [Nebraska Legislature's website][nel].

[nel]: https://nebraskalegislature.gov/reports/lobby.php

> The following reports identify lobbyists registered in Nebraska with the Office of the Clerk of
the Legislature.
> 
> ##### Lists of Registered Lobbyists
> * [Lobby Registration Report by Principal][1]
> * [Lobby Registration Report by Lobbyist][2]
> * [Lobbyist/Principal Expenditures Report][3]
> * [Lobbyist/Principal Statement of Activity][4]
> * [Counts of Lobbyists/Principals][5]

[1]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/principallist.pdf
[2]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/lobbyistlist.pdf
[3]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/expense.pdf
[4]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/activity_final_by_bill.pdf
[5]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/counts.pdf

Each of these data files comes in PDF format only.

## Import

We will have to use the `pdftools` package to read these files and extract the text content. We
are first interested in the list of registered lobbyists by principal clients.

```{r pdf_file}
pdf_file <- "https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/principallist.pdf"
```

The `pdftools::pdf_text()` file can read each page of the file into a list of character strings,
with each page being read as a _single_ character string list element.

```{r pdf_text}
nelr <- pdf_text(pdf_file)
cat(nelr[[1]])
```

We can define a function that uses the `stringr` and `tibble` packages to takes these page strings
and split them up and wrangle them into a data frame. The file is structured so that lobbyists are
listed below each of their principals, indented by 3 spaces. We can use that structure to identify
which rows are lobbyists, then split the string into three columns.

```{r pdf_table}
pdf_table <- function(page) {
  # split the page into lines
  x <- page %>%
    str_remove_all("\\sWD\\s") %>%
    str_split("\n") %>%
    `[[`(1)
  # define the rows to remove
  heading <- 1:str_which(x, "^Principal and Lobbyist\\s")
  footer <- str_which(x, "Lobbyist Registration, Room 2014 State Capitol"):length(x)
  x <- x[-c(heading, footer)]
  # enframe the rows and separate
  x <- x %>%
    enframe(name = NULL, value = "line") %>%
    mutate(
      indent = str_detect(line, "^\\s{3}\\w"),
      line = str_trim(line)
    ) %>%
    separate(
      col = line,
      into = c("name", "address"),
      sep = "\\s{2,}",
      extra = "merge"
    ) %>% 
    separate(
      col = address,
      into = c("address", "phone"),
      sep = "\\s+(?=\\()"
    )
}
```

```{r pdf_parse}
nelr <- nelr %>% 
  map_dfr(pdf_table) %>%
  filter(phone %>% str_detect("^\\(")) %>%
  mutate(
    pri_name = ifelse(!indent, name, NA),
    pri_geo = ifelse(!indent, address, NA),
    pri_phone = ifelse(!indent, phone, NA)
  ) %>%
  fill(starts_with("pri")) %>%
  filter(indent) %>%
  select(-indent) %>% 
  rename(
    lob_name = name,
    lob_geo = address,
    lob_phone = phone
  )
```

```{r echo=FALSE}
print(nelr)
```

## Wrangle

This new `lob_address` column can now be split into it's components with `tidyr::separate()`.

```{r address_split_lob}
nelr <- nelr %>%
  separate(
    col = lob_geo,
    into = c(glue("lob_street{1:10}"), "lob_city", "lob_state_zip"),
    sep = ",\\s",
    fill = "left",
    remove = FALSE
  ) %>%
  unite(
    starts_with("lob_street"),
    col = "lob_address",
    sep = " ",
    na.rm = TRUE
  ) %>%
  separate(
    col = lob_state_zip,
    sep = "\\s(?=\\d)",
    into = c("lob_state", "lob_zip")
  )
```

The same process needs to be done for the `pri_address`.

```{r address_split_pri}
nelr <- nelr %>%
  separate(
    col = pri_geo,
    into = c(glue("pri_street{1:10}"), "pri_city", "pri_state_zip"),
    sep = ",\\s",
    fill = "left",
    remove = FALSE
  ) %>%
  unite(
    starts_with("pri_street"),
    col = "pri_address",
    sep = " ",
    na.rm = TRUE
  ) %>%
  separate(
    col = pri_state_zip,
    sep = "\\s(?=\\d)",
    into = c("pri_state", "pri_zip")
  )
```

### Phone

```{r phone_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("phone")),
    .fun = list(norm = normal_phone)
  )
```

### Address

```{r address_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("address")),
    .fun = list(norm = normal_address),
    abbs = usps_street,
    na_rep = TRUE
  )
```

```{r address_view, echo=FALSE}
nelr %>% 
  select(
    pri_address,
    pri_address_norm
  ) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

```{r zip_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("zip")),
    .fun = list(norm = normal_zip),
    na_rep = TRUE
  )
```

```{r zip_progress, echo=FALSE}
progress_table(
  nelr$lob_zip,
  nelr$lob_zip_norm,
  nelr$pri_zip,
  nelr$pri_zip_norm,
  compare = valid_zip
)
```

### State

The `*_state` components do not need to be wrangled.

```{r state_progress}
prop_in(nelr$lob_state, valid_state)
prop_in(nelr$pri_state, valid_state)
```

### City

```{r city_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("city")),
    .fun = list(norm = normal_city),
    na_rep = TRUE
  )
```

```{r city_swap_lob}
nelr <- nelr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lob_state" = "state",
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lob_city_norm, city_match),
    match_dist = str_dist(lob_city_norm, city_match),
    lob_city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap_pri}
nelr <- nelr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "pri_state" = "state",
      "pri_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(pri_city_norm, city_match),
    match_dist = str_dist(pri_city_norm, city_match),
    pri_city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = pri_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_progress_lob, echo=FALSE}
progress_table(
  nelr$lob_city,
  nelr$lob_city_norm,
  nelr$lob_city_swap,
  compare = c(valid_city, extra_city)
)
```

```{r city_progress_pri, echo=FALSE}
progress_table(
  nelr$pri_city,
  nelr$pri_city_norm,
  nelr$pri_city_swap,
  compare = c(valid_city, extra_city)
)
```

## Export

```{r proc_dir}
proc_dir <- here("ne", "lobbying", "data", "processed")
dir_create(proc_dir)
```

```{r write_clean}
nelr %>% 
  select(
    lob_name,
    lob_geo,
    lob_addr = lob_address_norm,
    lob_city = lob_city_swap,
    lob_state,
    lob_zip = lob_zip_norm,
    pri_name,
    pri_geo,
    pri_addr = pri_address_norm,
    pri_city = pri_city_swap,
    pri_state,
    pri_zip = pri_zip_norm,
  ) %>% 
  write_csv(
    path = glue("{proc_dir}/ne_lobbyist_clean.csv"),
    na = ""
  )
```

