---
title: "Maryland Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  # it's nice to un-collapse df print
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("kiernann/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Maryland Campaign Reporting Information System][03] (CRIS).

> ## General Information on Expenditures and Outstanding Obligations
>
> An ***expenditure*** is defined as a gift, transfer, disbursement, or promise of money or
valuable thing by or on behalf of a political committee to promote or assist in promoting the
success or defeat of a candidate, political party, or question at  an election.
> 
> Expenditures must be election related; that is, they must enhance the candidates election
chances, such that they would not have been incurred if there had been no candidacy. Furthermore,
expenditures, including loans, may not be for the personal use of the candidate or any other
individual.
> 
> An outstanding obligation is any unpaid debt that the committee has incurred at the end of a
reporting period.

[03]: https://campaignfinance.maryland.gov/Home/Logout

## Import

### Download

```{r create_raw_dir}
raw_dir <- here("md", "expends", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw}

```

### Read

```{r read_raw}
md <- read_delim(
  file = dir_ls(raw_dir),
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character(),
    `Expenditure Date` = col_date_usa(),
    `Amount($)` = col_double()
  )
)
```

```{r read_problems}
problems(md) %>% 
  select(-row, -file) %>% 
  distinct()
```

```{r}
md <- 
  read_lines(file = dir_ls(raw_dir)) %>% 
  str_remove_all("\"") %>% 
  str_c(collapse = "\n") %>% 
  vroom(
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character()
  )
)
```

```{r}
md <- md[which(str_count(md, ",") != 12)] %>% str_split(",")
```

