---
title: "New York Attorneys"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "ny", "attorneys", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  jsonlite, # read json files
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/ny/attorneys/docs/ny_attorneys_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

We can obtain the New York attorneys data from the [state open data portal][na],
which gets it's data from the NYS Unified Court System and managed by the NYS 
Office of Court Administration.

> The data included here is the information in the NYS Attorney Registration
Database that is deemed public information pursuant to 22 NYCRR 118.

> Data derived from most recent biennial registrations filed by NYS Attorneys

> All registered attorneys admitted in NYS, whether they are resident or
non-resident, active or retired, or practicing law in NY or anywhere else.

[na]: https://data.ny.gov/Transparency/NYS-Attorney-Registrations/eqw2-r5nb

The dataset was created on April 12, 2013 and is updated daily.

```{r echo=FALSE}
about <- fromJSON("https://data.ny.gov/api/views/eqw2-r5nb.json")
about$columns |> 
  select(
    position,
    name,
    fieldName,
    dataTypeName
  ) |> 
  mutate(across(fieldName, md_code)) %>% 
  kable()
```

## Download

```{r raw-dir}
raw_url <- "https://data.ny.gov/api/views/eqw2-r5nb/rows.tsv"
raw_dir <- dir_create(here("state", "ny", "licenses", "data", "raw"))
raw_csv <- path(raw_dir, basename(raw_url))
```

```{r raw-download}
if (!file_exists(raw_csv)) {
  GET(
    url = raw_url,
    body = list(accessType = "DOWNLOAD"),
    write_disk(raw_csv),
    progress("down")
  )
}
```

## Read
```{r raw-read}
nya <- read_delim(
  file = raw_csv,
  delim = "\t",
  skip = 1, # use fieldName header
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = about$columns$fieldName,
  col_types = cols(
    .default = col_character(),
    year_admitted = col_integer(),
    next_registration = col_date("%b %Y")
  )
)
```

## Explore

There are `r comma(nrow(nya))` rows of `r ncol(nya)` columns. Each record
represents a single attorney registered in the state of New York.

```{r glimpse}
glimpse(nya)
tail(nya)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(nya, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("year_admitted", "last_name", "judicial_department_of_admission")
nya <- flag_na(nya, all_of(key_vars))
sum(nya$na_flag)
```

```{r na-view}
nya %>% 
  filter(na_flag) %>% 
  select(year_admitted, ends_with("name"))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
nya <- flag_dupes(nya, -registration_number)
sum(nya$dupe_flag)
```

There are no duplicates in the data, even when ignoring `registration_number`.

### Categorical

```{r distinct-count}
col_stats(nya, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(nya, county) + scale_x_wrap()
explore_plot(nya, judicial_department_of_admission)
explore_plot(nya, status) + scale_x_wrap()
```

### Dates

There is already a `year_admitted` value, but no date exists for when they 
Attorney was admitted. There is a month of when they must next register.

```{r bar-year, echo=FALSE}
nya %>% 
  count(year_admitted) %>% 
  ggplot(aes(x = year_admitted, y = n)) +
  geom_col(aes(fill = n), color = "black") + 
  scale_fill_viridis_c(option = "C", end = 0.90, guide = "none") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2022, by = 10)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Attorneys by Year",
    caption = "Source: NYS Unified Court System",
    fill = "Election Year",
    x = "Year Admitted",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

It should be noted that a not insignificant number of Attorneys are not from the
United States.

```{r}
nya$country <- str_to_upper(nya$country)
not_usa <- nya$country != "UNITED STATES OF AMERICA" & !is.na(nya$country)
mean(not_usa)
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
addr_norm <- nya %>% 
  distinct(street_1, street_2) %>% 
  mutate(
    street_norm_1 = normal_address(
      address = street_1,
      abbs = usps_street,
      na_rep = TRUE
    ),
    street_norm_2 = normal_address(
      address = street_2,
      abbs = usps_street,
      na_rep = TRUE,
      abb_end = FALSE
    )
  ) %>% 
  unite(
    col = street_norm,
    starts_with("street_norm"),
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  na_if("")
```

```{r address-view}
sample_n(addr_norm, 10)
```

```{r address-join}
nya <- left_join(nya, addr_norm, by = c("street_1", "street_2"))
```

### ZIP

The `zip` column already contains the trimmed 5-digit ZIP code and an additional
but separate `zip_plus_four` column.

### State

The `state` column already contains only clean abbreviations.

```{r zip-show}
head(na.omit(nya$zip))
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

```{r state-check}
prop_in(nya$state, valid_state)
```

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- nya %>% 
  distinct(city, state, zip) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NY", "DC", "NEW YORK"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c("state", "zip")
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
nya <- left_join(
  x = nya,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state", 
    "zip"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- nya %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state",
      "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
nya <- nya %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r}
many_city <- c(valid_city, extra_city)
```

Most of the remaining "invalid" cities are overseas.

```{r}
out_city <- nya %>% 
  count(city_refine, country, sort = TRUE) %>% 
  filter(city_refine %out% many_city, !is.na(city_refine)) %>% 
  add_prop(sum = TRUE)
out_city
```

```{r}
head(out_city$city_refine, 10)
```

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(nya$city[!not_usa]),
  nya$city_norm[!not_usa],
  nya$city_swap[!not_usa],
  nya$city_refine[!not_usa],
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(nya$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    subtitle = glue("IGNORING NON-USA CITIES"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York City Normalization Progress",
    subtitle = "IGNORING NON-USA CITIES",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
nya <- nya %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) 
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(nya, 1000))
```

1. There are `r comma(nrow(nya))` records in the database.
1. There are `r comma(sum(nya$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nya$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-dir}
clean_dir <- dir_create(here("state", "ny", "attorneys", "data", "clean"))
clean_csv <- path(clean_dir, glue("ny_attorneys_20220524.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(nya, clean_csv, na = "")
write_rds(nya, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
