---
title: "Rhode Island Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & database
  knitr, # knit documents
  here, # relative storage
  fs # search storage 
)
```

```{r fix_fun, echo=FALSE}
# fix conflict
here <- here::here
# custom utility functions
"%out%" <- Negate("%in%")
print_all <- function(dc) dc %>% print(n = nrow(.)) 
# source functions
source(here("R", "code", "normalize_geo.R"))
source(here("R", "code", "all_files_new.R"))
source(here("R", "code", "glimpse_fun.R"))
# load data
geo <- read_csv(here("R", "data", "geo.csv"))
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Describe _where_ the data is coming from. [Link to the data download][03] page if possible.

Describe the data set that is going to be cleaned. A file name, age, and unit of observation.

[03]: https://example.com "source"

### About

> If the publisher provides any information on the file, you can directly quote that here.

### Variables

`variable_name`:

> Directly quote the definition given for variables of interest.

## Import

### Download

### Read

```{r read_raw}
ri <- read_csv(
  file = "ri/expends/data/raw/1c4eb100-d3cc-4130-9374-c02782ebf646.csv",
  col_types = cols(
    .default = col_character(),
    exp_date = col_date("%m/%d/%Y"),
    amount = col_double(),
    OSAP =  col_double(),
    ZeroedByCF7 = col_logical()
  )
)

ri <- ri %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper)
```

## Explore

There are `r nrow(ri)` records of `r length(ri)` variables in the full database.

```{r glimpse}
glimpse(sample_frac(ri))
```

### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
ri %>% glimpse_fun(n_distinct)
```

```{r plot_bar, echo=FALSE}
ggplot(data = ri) + 
  geom_bar(mapping = aes(exp_desc)) +
  coord_flip()
```

```{r plot_bar, echo=FALSE}
ggplot(data = ri) + 
  geom_bar(mapping = aes(receipt_desc)) +
  coord_flip()
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
ri %>% glimpse_fun(count_na)
```

We will flag those records missing a `FullName` value.

```{r na_flag, collapse=TRUE}
ri <- ri %>% mutate(na_flag = is.na(full_name))
mean(ri$na_flag)
```

### Duplicates

We can use `janitor::get_dupes()` to create a table of records duplicated more than once.

```{r get_dupes, collapse=TRUE}
ri_dupes <- distinct(get_dupes(ri))
nrow(ri_dupes)
sum(ri_dupes$dupe_count)
```

We can then join this table back with the full dataset, flagging any duplicated rows with
`dupe_flag()`.

```{r join_dupes}
ri <- ri %>%
  left_join(ri_dupes) %>% 
  mutate(dupe_flag = !is.na(dupe_count))
```

### Ranges

#### amounts

There are `r sum(ri$amount < 0)` records with `amount` values less than zero. 

```{r range_amount, collapse=TRUE}
summary(ri$amount)
sum(ri$amount < 0)
```

```{r amount_hist}
ri %>% 
  ggplot(aes(amount)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10", labels = scales::dollar)
```

```{r amount_box_type, fig.height=10}
ri %>% 
  ggplot(aes(exp_desc, amount)) +
  geom_boxplot(varwidth = TRUE, outlier.alpha = 0.1) +
  scale_y_continuous(trans = "log10", labels = scales::dollar) +
  coord_flip()
```

### Dates

There are `r sum(ri$exp_date > today())` records with an `exp_date` past `r today()`.

```{r range_date, collapse=TRUE}
summary(ri$exp_date)
sum(ri$exp_date > today())
```

```{r year_bar, echo=FALSE}
ri %>% 
  ggplot() +
  geom_bar(aes(year(exp_date)))
```

```{r month_line_desc, echo=FALSE}
ri %>% 
  group_by(exp_desc, month = month(exp_date)) %>% 
  summarize(median_amount = median(amount)) %>% 
  ggplot(aes(month, median_amount)) +
  geom_line(aes(color = exp_desc), size = 2) +
  scale_y_continuous(labels = scales::dollar)
```

## Wrangle

### Year

Add a `year` variable from `date` after `col_date()` using `lubridate::year()`.

```{r add_year}
ri <- ri %>% mutate(exp_year = year(exp_date))
```

### Address

Before wrangling the geographic data, we need to separate the `city_st_zip` variable into it's 
respective parts.

```{r sep_geo}
ri <- ri %>% 
  separate(
    col = city_st_zip,
    into = c("city_sep", "state_zip"),
    sep = ",\\s",
    remove = FALSE
  ) %>% 
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s",
    remove = TRUE
  )
```

The `address` variable should be minimally cleaned by removing punctuation and fixing white-space.

```{r norm_address}
ri <- ri %>% mutate(address_clean = normalize_address(Address))
ri %>% 
  filter(address_clean != address) %>% 
  select(address, address_clean) %>% 
  head(10)
```

```{r norm_zip, collapse=TRUE}
ri <- ri %>% mutate(zip_clean = normalize_zip(zip_sep, na_rep = TRUE))
mean(is.na(ri$zip_clean))
mean(na.omit(ri$zip_clean) %in% geo$zip)
```

### City

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Prep

```{r prep_city}
df <- df %>% mutate(city_prep = normalize_city(city))
```

#### Match

```{r match_dist}
df <- df %>%
  left_join() %>%
  rename() %>%
  mutate(match_dist = stringdist())
```

#### Swap

```{r}
df <- df %>% 
  mutate(
    city_swap = if_else(
      condition = match_dist == 1, 
      true = city_match, 
      false = city_prep)
  )
```

#### Refine

```{r valid_city}
valid_city <- unique(zipcode$city)
```

```{r view_refine}
df_refined <- df %>%
  filter(var == "DF") %>% 
  filter(match_dist != 1) %>% 
  mutate(
    city_refine = var %>% 
      key_collision_merge(dict = valid_city) %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_swal != city_refine)
```

#### Merge

```{r join_refine}
df <- df %>% 
  left_join(refine_table, by = "id") %>% 
  mutate(city_clean = coalesce(city_refine, city_swap))
```

Each step of the cleaning process reduces the number of distinct city values.

## Conclude

1. There are `r nrow(df)` records in the database
1. There are `r sum(df$dupe_flag)` records with duplicate filer, recipient, date, _and_ amount
(flagged with `dupe_flag`)
1. The ranges for dates and amounts are reasonable
1. Consistency has been improved with `stringr` package and custom `normalize_*()` functions.
1. The five-digit `zip_clean` variable has been created with `zipcode::clean.zipcode()`
1. The `year` variable has been created with `lubridate::year()`
1. There are `r sum(is.na(df$name))` records with missing `name` values and `r sum(is.na(df$date))`
records with missing `date` values (both flagged with the `na_flag`)

## Export

```{r write_clean}

```

