---
title: "Missouri Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  # it's nice to un-collapse df print
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("kiernann/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # text analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  rvest, # scrape html pages
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [New York Board of Elections][03] (SBOE). 

> The State Board of Elections was established in the Executive Department June 1, 1974 as a
bipartisan agency vested with the responsibility for administration and enforcement of all laws
relating to elections in New York State. The Board is also responsible for regulating disclosure
and limitations of a Fair Campaign Code intended to govern campaign practices. In conducting these
wide-ranging responsibilities, the Board offers assistance to local election boards and
investigates complaints of possible statutory violations. In addition to the regulatory and
enforcement responsibilities the board is charged with the preservation of citizen confidence in
the democratic process and enhancement in voter participation in elections.

The NYSBOE database can be obstained from their Campaign Finance [discolure reports page][04]. On
that page, they elaborate on the availability and accuracy of the website.

> ### Data Availability  
> This database contains all financial disclosure reports filed with NYSBOE from July of 1999 to
the present. Financial disclosure reports filed prior to the 1999 July Periodic report are either
on file with the New York State Archives or in storage with the New York State Board of Elections.
For further information or to obtain copies of these archived or stored filings, please call
1-800-458-3453. Each page costs 25¢ plus postage and copy orders must be prepaid.
> 
> Electronically filed disclosure reports are generally available in the database on the day they
are received. A small number of candidates and committees are either statutorily exempt or have
applied for and obtained exemptions from electronic filing. These filers will continue filing on
paper and their disclosure reports will become available as they are manually entered into the
database by NYSBOE staff.

> ### Data Accuracy  
> The majority of financial disclosure reports filed at NYSBOE are entered into the database
directly from e-mail, diskette, CD or DVD filings submitted by committee treasurers or candidates.
The information contained in paper filings will be entered into the database exactly as it appears
on the forms. Because database searches retrieve information exactly the way it is reported and
then entered into the database, search results may be inaccurate and/or incomplete. This will
occur, for example, if filers do not adhere to the required format, do not use the proper codes,
misspell words or leave items blank. Although NYSBOE carefully reviews disclosure reports and
requires treasurers to submit amended reports as needed, there will necessarily be delays before
the review process is completed and the information in the database is corrected.

The page also describes the format of their campaign finance database.

> ### Database Files in ASCII Delimited Format
> **Updated data files are uploaded during active filing periods after 4:00 P.M. daily until the
filing is complete.**
>
> **Note:** To match the filing data files to Filer Names by filer ID you will need to Download the
Filer data file. Commcand.zip is a zipped file containing the data file (commcand.asc) in ASCII
delimited and two text files. (filerec.txt contains the data file layout - codes.txt explains the
codes used in the data file).
>
> **All downloadable files are zipped files containing a data file in ASCII delimited format and
two text files. (efsrecb.txt contains the data file layout - efssched.txt explains the different
schedules as they apply to the database).**
>
> [Download Data file containing ALL filings][05]. **Note:** This file is a large file (238,994
KB) that contains over 6 million records. Do not attempt to download this file unless you have a
database to download the file to.

[03]: https://www.elections.ny.gov/INDEX.html
[04]: https://www.elections.ny.gov/CFViewReports.html
[05]: https://www.elections.ny.gov/NYSBOE/download/ZipDataFiles/ALL_REPORTS.zip

## Import

We can use the link above to download a copy of the NYSBOE database to the `/data/raw` directory.

```{r raw_dir}
raw_dir <- here("ny", "expends", "data", "raw")
dir_create(raw_dir)
```

### Download

```{r create_zip_paths}
zip_url <- "https://www.elections.ny.gov/NYSBOE/download/ZipDataFiles/ALL_REPORTS.zip"
zip_path <- glue("{raw_dir}/{basename(zip_url)}")
```

Like they suggest, we will double check the size of the file before downloading.

```{r check_file_size}
url_file_size(zip_url, format = TRUE)
```

If the `ALL_REPORTS.zip` file hasn't yet been downloaded, do so now with `download.file()`.

```{r download_zip}
if (!this_file_new(zip_path)) {
  download.file(
    url = zip_url,
    destfile = zip_path
  )
}
```

### Unzip

If the `ALL_REPORTS.zip` file hasn't yet been unziped, we can do so now with `unzip()`. First, we
will list the files in the ZIP archive.

```{r}
zip_content <- as_tibble(
  x = unzip(zip_path, list = TRUE),
  .name_repair = make_clean_names
)
print(zip_content)

out_file <- glue("{raw_dir}/{zip_content$name[3]}")
```

```{r}
if (!this_file_new(out_file)) {
  unzip(
    zipfile = zip_path,
    exdir = raw_dir,
    overwrite = TRUE
  )
}
```

### Read

```{r out_layout, echo=FALSE}
layout_file <- glue("{raw_dir}/{zip_content$name[1]}")
out_layout <- 
  read_lines(file = layout_file) %>% 
  extract(11:42) %>%
  enframe(NULL) %>% 
  filter(str_detect(value, "^$", negate = TRUE)) %>% 
  map_df(str_trim) %>% 
  separate(
    col = value,
    into = c("FIELD", "LOCATION", "TYPE", "FORMAT", "IMPORT"),
    sep = "   +"
  ) %>% 
  na_if("REQUIRED") %>% 
  select(-IMPORT)

print(out_layout, n = 30)
```

```{r out_format, echo=FALSE}
out_format <- 
  read_lines(file = layout_file) %>% 
  extract(44:50) %>%
  enframe(NULL) %>% 
  separate(
    col = value,
    into = c("aspect", "value"),
    sep = ":"
  ) %>% 
  map_df(str_trim) %>% 
  mutate(aspect = snakecase::to_snake_case(aspect))

print(out_format)
```

From the `EFSSCHED.TXT` file, we know Schedule L records contain the expenditures we are interested
in.

```{r out_schedules, echo=FALSE}
sched_file <- glue("{raw_dir}/{zip_content$name[2]}")
out_scheds <- 
  read_lines(file = sched_file) %>% 
  extract(60:78) %>% 
  enframe(NULL) %>% 
  separate(
    col = value, 
    into = c("sched", "desc"), 
    sep = "\\s-\\s"
  ) %>% 
  map_df(str_trim)

print(out_scheds)
```

```{r}
out_cols <- str_remove(read_lines(file = sched_file, skip_empty_rows = TRUE)[4:34][-2], "\n")
out_cols_header <- unlist(str_split(out_cols[1], "\\s+"))
x <- str_c(out_cols[str_which(out_cols, "X")], collapse = "\n")
z <- read_fwf(file = x, col_positions = fwf_empty(x, col_names = out_cols_header))
names <- str_remove(na.omit(z$FIELD[z$J == "X"]), "_\\d+")
```

```
FIELD NAMES     A  B  C D  E  F  G  H  I  J K  L  M  N  O  P Q
--------------------------------------------------------------
DATE1           X  X  X X  X  X  X  X  X  X X  X  X  X     X X
DATE2                                       X  X  X
CONTRIB_CODE    X       X                               X  X
CONTRIB_TYPE            X
CORP            X  X  X X  X  X  X  X  X  X X  X  X  X  X  X X
FIRST_NAME      X       X                                  X
MID_INIT        X       X                                  X
LAST_NAME       X       X                                  X
ADDR            X  X  X X  X  X  X  X  X  X X  X  X  X     X X
CITY            X  X  X X  X  X  X  X  X  X X  X  X  X     X X
STATE           X  X  X X  X  X  X  X  X  X X  X  X  X     X X
ZIP             X  X  X X  X  X  X  X  X  X X  X  X  X     X X
CHECK_NO        X  X  X       X  X  X     X       X        X X
CHECK_DATE                                X
AMOUNT          X  X  X X  X  X  X  X  X  X X  X  X  X     X X
AMOUNT2                                     X        X  X
DESCRIPTION             X
OTHER_RECPT                X
PURPOSE_CODE1                 X                      X  X
PURPOSE_CODE2                                                X
EXPLANATION                   X                      X  X    X
XFER_TYPE                        X  X
CHKBOX                                 X                      
```

```{r}
ny <- read_delim(
  file = out_file, 
  delim = ",", 
  col_names = FALSE,
  col_types = cols(
    .default = col_character()
  ),
  escape_double = FALSE,
  escape_backslash = FALSE
)

ny <- filter(ny, X2 == "J")
```
