---
title: "Data Diary"
subtitle: "Nevada Contributions"
author: "Kiernan Nicholls"
date: "`r format(Sys.time())`"
output:
  html_document: 
    df_print: tibble
    fig_caption: yes
    highlight: tango
    keep_md: yes
    max.print: 32
    toc: yes
    toc_float: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Objectives

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called ZIP5
1. Create a YEAR field from the transaction date
1. For campaign donation data, make sure there is both a donor AND recipient

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, warning=FALSE, error=FALSE}
# install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # date strings
  magrittr, # pipe opperators
  janitor, # data cleaning
  zipcode, # clean and compare
  refinr, # cluster and merge
  vroom, # read files fast
  rvest, # scrape web pages
  knitr, # knit documents
  httr, # acess web API
  here, # navigate local storage
  fs # search local storage 
)
```

## Data

The Nevada Secretary of State (NVSOS) office requires that one register for an account to access 
"[bulk data download](https://www.nvsos.gov/sos/online-services/data-download)" service page.

> Welcome to the Nevada Secretary of State online unified login system.  Here you may access the following systems all with one login account:
>
> * Document Preparation Service *New
> * Commercial Registered Agent Tools
> * Bulk Data Download
> * NV SOS Trust Account
> * Nevada Ministers Database
> * Notary Class / Notary Address Change

The process for downloaded a report is [outlined here](https://www.nvsos.gov/SoSServices/AnonymousAccess/HelpGuides/DataDownloadUserGuide.aspx):

In brief, we will be downloading a "Full Unabridged Database Dump" of "Campaign Finance" data,
which will be delivered via email. We have chosen the file to be in the CSV format with column
names in the first row. When selected a report with this criteria, the NVSOS website provides
the following message: 

> This report will expose Contributions and Expenses report data filed within our “Aurora” Campaign
Financial Disclosure system.  This would not include data filed in a Financial Disclosure report.
This bulk data report tool here should be used to pull the entire database or slightly smaller
subsets of data such as all contributions filed after 1/1/2016 by groups of type “PAC”.  For more
detailed searches such as looking up all contributions filed by candidate “John Doe” or looking up
all expenses filed between $100,000 and $300,000 use the free public search here.

The website also provides a brief message regarding the age of the data:

> The data being reported off of is no more than 24 hours old. This data is copied very late each night from live data to minimize the large burden of bulk reporting on the production system.

This report was generated on `r today()`.

The report data is partitioned into multiple files, as explained on the 
[NVSOS FAQ page](https://www.nvsos.gov/SOSServices/AnonymousAccess/HelpGuides/FAQ.aspx#5):

> This is what is referred to as a normalized relational structure in the database world.  Data
items such as business entities and officers have a direct relation to one another.  There can be
any number of officers to one business entity.  Because of this many to one relationship, the
officers data is stored in a different data table (or file) than the business entities.  Then we
relate officer records to a business entity record by a common key data column, in this case the
CorporationID... By separating officers and entities into separate records we can eliminate the
redundancy and added size associated with putting the business entity data on each officer record
or eliminate the complexity of allocating an undeterminable amount of officers on the one business
entity record.  This same many-to-one relationship is true of voter history records to voter
records, UCC actions to UCC liens or Corporation Stocks to Corporations, to name a few.

Specifically, a campaign finance report is partitioned into six individual reports:

1. Candidates
1. Contributions
1. Contributors-Payees
1. Expenses
1. Groups
1. Reports

NVSOS provides information on some variables in the "Result Field" tab of the report generator:

In the "Candidates" report:

* `Jurisdiction` = 

> This will be name of the city or county for city/county offices currently held by the candidate
(e.g. “CITY OF YERINGTON”, “DOUGLAS COUNTY”).  This will be set to “NV SOS” for statewide offices
such as Governor, State Controller or State assemblymen.  An office assigned to a candidate could
be updated by the NV SOS Elections staff as necessary when that candidate files for a new office.

In the "Contributions" report:

* `Contribution Type` = 

> Use this column to differentiate which one of four contribution types this contribution record
is: Monetary Contribution, In Kind Contribution, In Kind Written Commitment, or Written Commitment.

In the "Contributors-Payees" report:

* `Last Name` = 

> When the contributor or payee is an organization as opposed to an individual, the entire
organization name will be in the Last Name field only.

In the "Expenses" report:

* `Expense Type` = 

> Use this column to differentiate which type of expense record this is: Monetary Expense or In
Kind Expense.

In the "Groups" report:

* `Active`

> A value of F (False) indicates the group has been marked as inactive by the NV Secretary of
State's office Elections division due to submission of a "notice of inactivity" or for failure to
renew annual registration.

In the "Reports" report:

* `Amended` = 

> A value of T (True) indicates this contributions and expense report has been marked as an amended
report by the original filer implying this report supersedes a report for this same period, filed
earlier.   An amended report is to be full comprehensive for that report period and in essence
replaces all contributions and expenses filed in the earlier report.

* `Election Cycle` = 

> The Election Cycle is the 4 digit filing or reporting year defining a filing period grouping
together a collection of contribution and expenses reports...

* `Superseded` = 

> A report is Superseded when an amended report was filed later by the same filer for the same
reporting period.  In this case the Superseded field for the older report record will be set to T
(True)...

## Read

The following link was sent via email and downloaded to the `data/` directory:

```
https://www.nvsos.gov/yourreports/CampaignFinance.43993.060419121813.zip
```

The ZIP file contains six individual files:

```{r unzip_list, echo=FALSE}
here("nv_contribs", "data") %>% 
  dir_ls(, glob = "*.zip") %>% 
  unzip(list = TRUE) %>% 
  clean_names()
```

The files will be unzipped into the `data/` directory.

```{r unzip}
dir_create(here("nv_contribs", "data"))
here("nv_contribs", "data") %>% 
  dir_ls(, glob = "*.zip") %>% 
  unzip(
    exdir = here("nv_contribs", "data"),
    overwrite = TRUE
  )
```

Each file can be read using the `vroom::vroom()` function:

```{r read_candidates}
nv_candidates <- 
  here("nv_contribs", "data") %>% 
  dir_ls(glob = "*.csv") %>% 
  extract(1) %>% 
  vroom(col_types = cols(.default = "c")) %>% 
  clean_names()

print(nv_candidates)
```

```{r read_contribs}
nv_contribs <- 
  here("nv_contribs", "data") %>% 
  dir_ls(glob = "*.csv") %>% 
  extract(2) %>% 
  vroom(
    .name_repair = make_clean_names,
    col_types = cols(
      .default = col_character(),
      `Contribution Date` = col_date("%m/%d/%Y"),
      `Contribution Amount` = col_number()
    )
  )

print(nv_contribs)
```

```{r read_groups}
nv_groups <- 
 here("nv_contribs", "data") %>% 
  dir_ls(glob = "*.csv") %>% 
  extract(5) %>% 
  vroom(
    .name_repair = make_clean_names,
    col_types = cols(
      .default = col_character(),
      Active = col_logical()
    )
  )

print(nv_groups)
```

```{r read_reports}
nv_reports <- 
 here("nv_contribs", "data") %>% 
  dir_ls(glob = "*.csv") %>% 
  extract(6) %>% 
  vroom(
    .name_repair = make_clean_names,
    col_types = cols(
      .default = col_character(),
      `Filing Due Date` = col_date("%m/%d/%Y"),
      `Filed Date` = col_date("%m/%d/%Y"),
      Amended = col_logical(),
      Superseded = col_logical()
    )
  )

print(nv_reports)
```

We are primarily interested in the file containing data on contributions. To make the data base
more searchable on the Accountability Project database, we will be joining together the various
normalized relational tables using their respective `*_id` variables.

```{r join, collapse=TRUE}
nv <- nv_contribs %>%
  # join with relational tables
  left_join(nv_candidates, by = "candidate_id") %>% 
  left_join(nv_reports, by = c("report_id", "candidate_id", "group_id")) %>%
  left_join(nv_groups, by = "group_id") %>% 
  # add origin table info to ambiguous variables
  rename(
    candidate_first = first_name,
    candidate_last = last_name,
    candidate_party = party,
    seeking_office = first_name,
    report_amended = amended, 
    report_superseded = superseded,
    group_contact = contact_name,
    group_active = active,
    group_city = city
  )

nrow(nv) == nrow(nv_contribs)
```

This expands our primary table from `r length(nv_contribs)` variables to `r length(nv)`
without changing the number or records included.

## Explore

Below is the structure of the data arranged randomly by row. There are `r nrow(nv)` rows of 
`r length(nv)` variables.

```{r glimpse_all}
glimpse(sample_frac(nv))
```

### Distinct

The variables vary in their degree of distinctiveness.

```{r count_distinct}
nv %>% 
  map(n_distinct) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_distinct") %>% 
  mutate(prop_distinct = round(n_distinct / nrow(nv), 4)) %>%
  print(n = length(nv))
```

For the least distinct variables, we can explore the most common values.

```{r tabyls_function, echo=FALSE}
print_tabyl <- function(data, ...) {
  as_tibble(arrange(tabyl(data, ...), desc(n)))
}
```

```{r tabyls}
print_tabyl(nv, contribution_type)
print_tabyl(nv, candidate_party)
print_tabyl(nv, seeking_office)
print_tabyl(nv, jurisdiction)
print_tabyl(nv, report_name)
print_tabyl(nv, election_cycle)
print_tabyl(nv, report_amended)
print_tabyl(nv, report_superseded)
print_tabyl(nv, group_type)
print_tabyl(nv, group_active)
print_tabyl(nv, group_city)
```

### Ranges

For continuous variables, the ranges should be checked.

```{r ranges}
summary(nv$contribution_date)
summary(nv$contribution_amount)
summary(nv$filing_due_date)
summary(nv$filed_date)
```

The date variables all seem to make sense. There are no dates before 
`r min(nv$contribution_date)` and none from the future (aside from the upcoming filing dates). The
maximum contribution is for `r noquote(scales::dollar(max(nv$contribution_amount)))`.

```{r max_amt}
nv %>% 
  filter(contribution_amount == max(contribution_amount)) %>% 
  glimpse()
```

### Plot

```{r plot_amt_party}
nv %>% 
  filter(candidate_party == "Democratic Party" |
           candidate_party == "Republican Party" |
             candidate_party == "Nonpartisan" |
                candidate_party == "Unspecified") %>% 
  ggplot(aes(contribution_amount)) +
  geom_histogram(aes(fill = candidate_party)) +
  scale_x_log10() +
  scale_y_log10() +
  scale_fill_manual(values = c("blue", "purple", "red", "black")) +
  theme(legend.position = "none") +
  facet_wrap(~candidate_party) +
  labs(
    title = "Contribution Distribution",
    subtitle = "by political Party",
    y = "Number of Contributions",
    x = "Amount ($USD)"
  )
```

### Missing

### Duplicates

## Clean

## Conclusion

## Write

```{r write_csv, eval=FALSE}
write_csv(
  x = nv,
  path = here("nv-contribs", "data", "nv_contribs_clean.csv"),
  na = ""
)
```

