---
title: "Missouri Voters"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("mo", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Missouri Centralized Voter Registration (MCVR) data was obtained as an open
record pursuant to Section 115.157, RSMo. from the Elections Division, Office of
Secretary of State.

The MCVR data was provided as a ZIP archive through a file sharing site.

```{r raw_dir}
raw_dir <- dir_create(here("mo", "voters", "data", "raw"))
raw_zip <- path(raw_dir, "Files.zip")
raw_out <- as_fs_path(unzip(raw_zip, exdir = raw_dir))
```

The archive contains a README file to explain the data:

```{r readme_about, echo=FALSE, results='asis'}
readme <- read_lines(raw_out[3])
md_quote(readme[1:5])
```

The README also contains a disclaimer:

```{r readme_disclaim, echo=FALSE, results='asis'}
md_quote(readme[18:21])
```

And a record layout describing the columns of the file:

```{r readme_layout}
readme[31:83] %>% 
  str_replace("(?<=\\d)  ", "\t") %>% 
  str_replace("\t{2,}", "\t") %>% 
  read_tsv(col_names = c("Position", "Name", "Description")) %>% 
  mutate(across(Name, md_code)) %>% 
  select(-Position) %>% 
  kable()
```

## Read

We can read the tab-delimited file as a dataframe.

```{r raw_read}
mov <- read_delim(
  file = raw_out[1],
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "NA", "XXXXX"),
  col_types = cols(
    .default = col_character(),
    Birthdate = col_date_usa(),
    `Registration Date` = col_date_usa()
  )
)
```

There are 20 columns at the end of the dataframe containing all of the past
elections in which each person has voted. We are going to keep the most recent
election and then save all the columns as a separate data frame. This data frame
will be kept in a _long_ format, with a row for every election.

```{r voter_split, eval=FALSE}
hist_file <- path(dirname(raw_dir), "vote_history.csv")
if (file_exists(hist_file)) {
  vote_hist <- vroom(
    file = hist_file,
    col_types = cols(
      voter_id = col_character(),
      order = col_integer(),
      date = col_date(),
      election = col_character()
    )
  )
} else {
  vote_hist <- select(mov, `Voter ID`, starts_with("Voter History"))
  vote_hist <- pivot_longer(
    data = vote_hist,
    cols = starts_with("Voter History"),
    names_to = "order",
    values_to = "election"
  )
  vote_hist <- vote_hist %>% 
    clean_names("snake") %>% 
    filter(!is.na(election))
  vote_hist <- separate(
    data = vote_hist,
    col = election,
    sep = "(?<=\\d)\\s",
    into = c("date", "election")
  )
  vote_hist <- mutate(
    .data = vote_hist,
    order = as.integer(str_extract(order, "\\d+")),
    date = parse_date(date, "%m/%d/%Y")
  )
  write_csv(
    x = vote_hist,
    path = hist_file
  )
}
```

We can then remove the election columns.

```{r voter_remove}
mov <- rename(mov, `Last Election` = `Voter History 1`)
mov <- select(mov, -starts_with("Voter History"))
```

```{r clean_names}
mov <- clean_names(mov, "snake")
```

```{r raw_rename}
mov <- mov %>% 
  select(-starts_with("mailing_")) %>% 
  rename_all(str_remove, "residential_") %>% 
  rename_all(str_remove, "_new") %>% 
  rename(
    zip = zip_code,
    birth_date = birthdate,
    reg_date = registration_date
  )
```

## Old

In 2018, the Workshop received a similar file. We are going to keep any
registered voters _not_ found in the current MCVR file.

```{r old_read}
moo <- s3read_using(
  FUN = readr::read_delim,
  filename = "MO_VOTERS.csv",
  object = "csv/MO_VOTERS.csv",
  bucket = "publicaccountability",
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "NA", "XXXXX"),
  col_types = cols(
    .default = col_character(),
    Birthdate = col_date_usa(),
    RegistrationDate = col_date_usa()
  )
)
```

```{r old_rename}
moo <- moo %>% 
  clean_names("snake") %>% 
  select(-year) %>% 
  rename(
    zip = zip5,
    birth_date = birthdate,
    reg_date = registration_date,
    state_senate = statesenate
  )
```

```{r old_names}
prop_in(names(moo), names(mov))
```

Most of the voters in the _old_ data are still found in the new data.

```{r old_overlap}
prop_in(moo$voter_id, mov$voter_id)
prop_in(mov$voter_id, moo$voter_id)
```

Using the unique `voter_id` we will remove any voter found in the newer data.

```{r old_filter}
nrow(moo)
moo <- filter(moo, voter_id %out% mov$voter_id)
nrow(moo)
```

The unique old data can then be joined to the most recent voter registrations.

```{r old_bind}
mov <- bind_rows(mov, moo, .id = "source")
mov <- relocate(mov, source, .after = last_col())
add_prop(count(mov, source))
```

## Explore

There are `r scales::comma(nrow(mov))` rows of `r ncol(mov)` columns.

```{r glimpse}
glimpse(mov)
tail(mov)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(mov, count_na)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_write}
dupe_file <- here("mo", "voters", "dupes.csv")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  write_lines("voter_id,dupe_flag", dupe_file)
  mos <- mov %>% 
    select(-voter_id) %>% 
    group_split(county)
  split_id <- split(mov$voter_id, mov$county)
  pb <- txtProgressBar(max = length(mos), style = 3)
  for (i in seq_along(mos)) {
    write_csv(
      path = dupe_file,
      append = TRUE,
      col_names = FALSE,
      x = tibble(
        voter_id = split_id[[i]],
        dupe_flag = or(
          e1 = duplicated(mos[[i]], fromLast = FALSE),
          e2 = duplicated(mos[[i]], fromLast = TRUE)
        )
      )
    )
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(mos)
}
```

```{r dupe_read}
dupes <- read_csv(
  file = dupe_file,
  col_types = cols(
    voter_id = col_character(),
    dupe_flag = col_logical()
  )
)
```

```{r dupe_join}
nrow(mov)
mov <- left_join(mov, dupes)
nrow(mov)
mov <- mutate(mov, dupe_flag = !is.na(dupe_flag))
sum(mov$dupe_flag)
```

We can see that, despite unique IDs, there are duplicate voters.

```{r dupe_view}
mov %>% 
  filter(dupe_flag) %>% 
  select(voter_id, first_name, last_name, birth_date, zip)
```

### Categorical

```{r distinct_count}
col_stats(mov, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(mov, county)
mov$voter_status[mov$voter_status == "SE-N 02"] <- "Active"
explore_plot(mov, voter_status)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
mov <- mutate(mov, reg_year = year(reg_date))
```

```{r date_range}
min(mov$reg_date)
mean(mov$reg_year < 2000)
max(mov$reg_date)
sum(mov$reg_date > today())
```

```{r bar_year, echo=FALSE}
mov %>% 
  count(reg_year) %>% 
  mutate(even = is_even(reg_year)) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1960, 2020, by = 20)) +
  coord_cartesian(xlim = c(1960, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Missouri Voters by Registration Year",
    caption = "Source: MDVR",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

We can create a single, unified normalized address field.

```{r addr_unite}
mov <- mov %>% 
  unite(
    col = address_norm,
    house_number:non_standard_address,
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  relocate(address_norm, .after = last_col())
```

```{r addr_view}
mov %>% 
  select(address_norm, house_number:non_standard_address) %>% 
  sample_n(20) %>% 
  remove_empty("cols")
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mov <- mov %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mov$zip,
  mov$zip_norm,
  compare = valid_zip
)
```

### State

As we would expect, all the Missouri voters live in Missouri.

```{r state_count}
count(mov, state, sort = TRUE)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mov <- mov %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("MO", "DC", "MISSOURI"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(mov$city),
  mov$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
mov <- rename_all(mov, ~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(mov, 50))
```

1. There are `r comma(nrow(mov))` records in the database.
1. There are `r comma(sum(mov$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mov$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("mo", "voters", "data", "clean"))
clean_path <- path(clean_dir, "mo_voters.csv")
write_csv(mov, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
