---
title: "West Virginia Lobbying Registration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wv", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.


## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs, # search storage 
  gluedown # generate dictionary
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[guide]: http://ethics.dls.West Virginia.gov/LDS%20Guide.pdf

## Data
Lobbyist data is obtained via a public record request from the [West Virginia State Ethics Commission][03]. The data was obtained on April 9, 2020.

It is a list of all current lobbyists for the 2019-2020 session.

According to the [West Virginia State Ethics Commission](https://ethics.wv.gov/lobbyist/Pages/Registration.aspx),

>   Lobbyists may register at any time.  Registration must be completed before lobbying begins.  Lobbyist registration cycles consist of two-year periods, and all lobbyist registrations expire at the end of each even-numbered year.

## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("state","wv", "lobby", "data", "raw","reg"))
```

### Read
```{r raw_read}
wvlr <- dir_ls(raw_dir) %>% read_xlsx() %>% 
  clean_names() %>% 
  mutate_if(is.character, str_to_upper)
```


## Explore

```{r glimpse}
head(wvlr)
tail(wvlr)
glimpse(sample_n(wvlr, 20))
```


### Missing
There're almost no missing fields, except for one record missing "represents".
```{r na_count}
col_stats(wvlr, count_na)
```

### Duplicates
There's no duplicate columns. 
```{r dupe_flag}
wvlr <- flag_dupes(wvlr, dplyr::everything())
```

## Wrangle
To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.


### Year,
We'll add a common `year` field which is the mos current year, 2020.
```{r}
wvlr <-  wvlr %>% mutate(year = "2020")
```


### Phone
We can normalize the phone numbers.

```{r normal_phone}
wvlr <- wvlr %>% 
      mutate(phone_primary_norm = normal_phone(phone_primary))
```

### Address

For the street `addresss` variable, the `campfin::normal_address()`
function will force consistence case, remove punctuation, and
abbreviation official USPS suffixes.
```{r norm address}
wvlr <- wvlr %>% 
    # combine street addr
  unite(
    col =  address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
    mutate(address_norm = normal_address(address_full,abbs = usps_street,
      na_rep = TRUE)) %>% 
  select(-ends_with("full"))
```

### Separate
We will first separate city, state and zip with regular expressions. Afterwards, we will also normalize them using `campfin::normal_*()` functions.
```{r sep csz city}
wvlr <- wvlr %>% 
  separate(col = city_state_zip, into = c("city", "state_zip"), sep = ", ", remove = FALSE) %>% 
  separate(col = state_zip, into = c("state", "zip"), sep = " ", remove = TRUE)
```

### ZIP
For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
prop_in(wvlr$zip, valid_zip, na.rm = TRUE)
wvlr <- wvlr %>% 
    mutate(zip5 =normal_zip(zip))
prop_in(wvlr$zip5, valid_zip, na.rm = TRUE)
```


### State
The state column is clean and doesn't need to be normalized.
```{r state norm}
prop_in(wvlr$state, valid_state, na.rm = TRUE)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
prop_in(wvlr$city, valid_city, na.rm = T)

wvlr <- wvlr %>% 
  mutate(city_norm = normal_city(city,abbs = usps_city,
                                 states = usps_state,
                                 na = invalid_city,
                                 na_rep = TRUE))
prop_in(wvlr$city_norm, valid_city, na.rm = T)
```


After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  wvlr$city,
  wvlr$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 5)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "West Virginia City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equiwvlrnt.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "West Virginia City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
```


## Conclude

```{r clean_glimpse}
glimpse(sample_n(wvlr, 20))
```

1. There are `r nrow(wvlr)` records in the database.
1. There's no duplicate record in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(wvlr$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("wv", "lobby", "data", "processed","exp"))
```
   
```{r write_clean}
wvlr <- wvlr %>% 
  rename(city_clean = city_norm)

write_csv(
  x = wvlr,
  path = path(clean_dir, "wv_lob_reg_clean.csv"),
  na = ""
)
```

## Dictionary

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wvlr)),
  # raw = c(md_code(raw_names), rep("", 7)),
  type = md_code(map_chr(wvlr, rlang::as_label)),
  def = c(rep("Original Field", 3),
          glue("City field extracted from", md_code("city_state_zip")),
          glue("State field extracted from", md_code("city_state_zip")),
          glue("Zip field extracted from", md_code("city_state_zip")),
          rep("Original Field",4),
          "The latest year in the current lobbying cycle",
          glue("Normalized primary phone numbers from", md_code("phone_primary_norm")),
          "Normalized combined street address",
          "Normalized 5-digit ZIP code",
          "Normalized 2-digit state abbreviation",
          "Normalized city name")
)
```
    
 
```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# West Virginia Lobbying Registration Dictionary\n", dict_md),
  path = here("wv", "lobby", "wi_contribs_dict.md"),
)
```
