---
title: "Idaho Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("id", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `zip`
1. Create a `year` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  snakecase, # column naming
  lubridate, # datetime strings
  magrittr, # pipe opperators
  gluedown, # printing markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  rvest, # read html pages
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][repo] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[repo]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

This data is obtained from the Michigan [Board of Elections (BOE)][boe] 
[Campaign Finance Reporting (CFR)][cfr] system. The data is provided as 
[annual ZIP archive files][data] for the years 1998 through 2020. These files
are updated nightly.

[boe]: https://www.michigan.gov/sos/0,4670,7-127-1633---,00.html
[cfr]: https://www.michigan.gov/sos/0,4670,7-127-1633_8723---,00.html
[data]: https://miboecfr.nictusa.com/cfr/dumpall/cfrdetail/

## Import

To import the data for processing, we will have to download each archive file
and read them together into a single data frame object.

### Download

We will scrape the download page for every archive link, then downloading each
to a local directory.

```{r download_raw}
raw_dir <- dir_create(here("id", "contribs", "data", "raw"))
raw_base <- "https://sos.idaho.gov/elect/finance/downloads.html"
raw_page <- read_html(raw_base)
raw_urls <- raw_page %>% 
  html_node("table") %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("con") %>%
  str_subset("^2") %>% 
  str_c(dirname(raw_base), ., sep = "/")
raw_paths <- path(raw_dir, basename(raw_urls))
if (!all(this_file_new(raw_paths))) {
  for(i in seq_along(raw_urls)) {
    download.file(raw_urls[i], raw_paths[i])
    Sys.sleep(runif(1))
  }
}
```

### Read

Each file has a different column order and names. We will first use 
`purrr::map()` to use `readxl::read_excel()` and create a list of data frames.

```{r read_excel}
idc <- map(
  .x = raw_paths,
  .f = read_excel,
  col_types = "text"
)
```

Since the variety of column names is so great, we will have to chain together
a number of `stringr::str_replace()` functions to create some consistency.

```{r rename}
consistent_names <- function(nm) {
  nm %>% 
    to_snake_case() %>% 
    str_replace("contributor_", "contr_") %>% 
    str_replace("contrib_", "contr_") %>% 
    str_replace("contr_name", "contr_last") %>% 
    str_replace("^first_name$", "cand_first") %>% 
    str_replace("^last_name$", "cand_last") %>% 
    str_replace("^middle_name$", "cand_mi") %>% 
    str_replace("suf$", "suffix") %>% 
    str_replace("_mid$", "_mi") %>% 
    str_replace("middle", "mi") %>% 
    str_replace("_cp$", "_type") %>% 
    str_remove("_name$") %>% 
    str_replace("zipcode", "zip") %>% 
    str_replace("_st$", "_state") %>% 
    str_replace("mailing", "address") %>% 
    str_replace("line_1", "address_1") %>% 
    str_replace("line_2", "address_2") %>% 
    str_remove("^contr_") %>% 
    str_remove("^contributing_") %>% 
    str_remove("^contribution_") %>% 
    str_replace("^address$", "address_1")
}

new_names <- idc %>% 
  map(names) %>% 
  map(consistent_names)

for (i in seq_along(idc)) {
  names(idc[[i]]) <- new_names[[i]]
}
```

Now that each individual data frame has similar column names, we can use
`dplyr::bind_rows()` to bind all `r length(idc)` data frames together.

```{r bind_rows}
idc <- bind_rows(idc, .id = "file")
```

Then, we can apply `readr::parse_guess()` to every column that were previously
all read as character vectors.

```{r parse_guess}
idc <- mutate_all(idc, parse_guess)
```

We also need to reorder and recode these variables to be consistent across each
year.

```{r reorder, echo=FALSE}
idc <- idc %>% 
  mutate(
    last = coalesce(last, committee_company)
  ) %>% 
  select(
    file,
    party,
    cand_first,
    cand_mi,
    cand_last,
    cand_suffix,
    committee,
    office,
    district,
    type,
    amount,
    date,
    last,
    first,
    mi,
    suffix,
    address_1,
    address_2,
    city,
    state,
    zip,
    country,
    election,
    election_type,
  )
```

```{r recode}
idc <- idc %>% 
  mutate_if(is_character, str_to_upper) %>% 
  mutate(
    office = office %>% 
      str_replace("REPRESENTATIVE", "REP."),
    cand_suffix = cand_suffix %>% 
      str_remove_all("[:punct:]"),
    party = party %>% 
      str_replace("OTHER", "OTH") %>% 
      str_replace("REPUBLICAN", "REP") %>% 
      str_replace("DEMOCRATIC", "DEM") %>% 
      str_replace("INDEPENDENT", "IND") %>% 
      str_replace("LIBERTARIAN", "LIB") %>% 
      str_replace("CONSTITUTION", "CON") %>% 
      str_replace("NON-PARTISAN", "NON"),
    type = type %>% 
      str_replace("COMPANY", "C") %>% 
      str_replace("PERSON", "P") %>% 
      str_replace("LOAN", "L") %>% 
      str_replace("IN KIND", "I"),
    election = election %>% 
      str_replace("GENERAL", "G") %>% 
      str_replace("PRIMARY", "P")
  )

count_na(idc$date) # 1276
slash_dates <- str_which(idc$date, "\\d+/\\d+/\\d{4}")
idc$date[slash_dates] <- as.character(mdy(idc$date[slash_dates]))
excel_dates <- str_which(idc$date, "[:punct:]", negate = TRUE)
idc$date[excel_dates] %>% 
  as.numeric() %>% 
  excel_numeric_to_date() %>% 
  as.character() -> idc$date[excel_dates]

idc$date <- as_date(idc$date)
count_na(idc$date) # 1276
```

## Explore

```{r glimpse}
head(idc)
tail(idc)
glimpse(sample_frac(idc))
```

We should first identify which columns are missing the kinds of key information
we need to properly identify all parties to a contribution. We can do this
with `campfin::flag_na()` after creating a new

```{r count_na}
col_stats(idc, count_na)
```

```{r flag_na}
idc <- idc %>% 
  # combine cand and comm names in new col
  mutate(recip = coalesce(cand_last, committee)) %>% 
  flag_na(last, recip, date, amount)

sum(idc$na_flag)
mean(idc$na_flag)
```

Records that are entirely duplicated at least once across all columns should
also be identified with `campfin::flag_dupes()`. The first occurrence of the
record is not flagged, but all subsequent duplicates are. Not all these records
are true duplicates, since it is technically possible to make the same
contribution to the same person on the same day for the same amount.

```{r flag_dupes}
idc <- flag_dupes(idc, everything(), .check = TRUE)
sum(idc$dupe_flag)
mean(idc$dupe_flag)
idc %>% 
  filter(dupe_flag) %>% 
  select(recip, last, date, amount)
```

### Categorical

```{r n_distinct}
col_stats(idc, n_distinct)
```

```{r bar_office, echo=FALSE}
explore_plot(
  data = filter(idc, !is.na(office)),
  var = office,
  title = "Idaho Contributions by Office",
  caption = "Source: Idaho Elections Office"
) + theme(axis.text.x = element_text(angle = 10, vjust = 0.7))
```

```{r bar_party, echo=FALSE}
explore_plot(
  data = filter(idc, !is.na(party)),
  var = party,
  title = "Idaho Contributions by Office",
  caption = "Source: Idaho Elections Office"
)
```

```{r type_party, echo=FALSE}
explore_plot(
  data = filter(idc, !is.na(type)),
  var = type,
  title = "Idaho Contributions by Office",
  caption = "Source: Idaho Elections Office"
)
```

### Amounts

```{r summary_amount}
summary(idc$amount)
sum(idc$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
idc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Idaho Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: Idaho Elections Office",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_violin, echo=FALSE}
idc %>%
  filter(
    amount > 1,
    amount < 1e6,
    party %in% most_common(idc$party, 4)
  ) %>% 
  ggplot(aes(x = party, y = amount)) +
  geom_violin(aes(fill = party)) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Idaho Contributions Amount Distribution",
    subtitle = "from 2000 to 2019, by Party",
    caption = "Source: Idaho Elections Office",
    x = "Amount",
    y = "Count"
  )
```

### Dates

```{r year_add}
idc <- mutate(idc, year = year(date))
```

```{r date_range}
min(idc$date, na.rm = TRUE)
max(idc$date, na.rm = TRUE)
idc <- mutate(idc, date_flag = date > today() | year < 1999 | is.na(date))
count_na(idc$date) # 1276
sum(idc$date_flag) # 1339 = 63
mean(idc$date_flag)
```

```{r date_replace}
x <- idc$date[idc$date_flag & !is.na(idc$date)]
x <- str_replace(x, "^202(?=[^2])", "201")
x <- str_replace(x, "^19([^9])", "199")
x <- str_replace(x, "^2([^2])", "20")
x <- str_replace(x, "2061", "2016")
x[which(x > today() | year(x) < 1999)] <- NA
idc$date[idc$date_flag & !is.na(idc$date)] <- as_date(x)
idc <- mutate(
  .data = idc,
  date_flag = date > today() | year < 1999 | is.na(date),
  year = year(date)
)
count_na(idc$date) # 1307
sum(idc$date_flag) # 1307
```

For some reason there no records from 2005 to 2010.

```{r year_bar, echo=FALSE}
idc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) +
  scale_y_continuous(label = comma) +
  labs(
    title = "Idaho Contributions by Year",
    caption = "Source: Idaho Elections Office",
    x = "Year Made",
    y = "Count"
  )
idc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) +
  scale_y_continuous(label = comma, trans = "log10") +
  labs(
    title = "Idaho Contributions by Year",
    caption = "Source: Idaho Elections Office",
    x = "Year Made",
    y = "Count"
  )
```

We know these files were
read by comparing the `file` variable to the urls downloaded. For the records
loaded from the file downloaded from 2008, the year of the `date` variable seems
to primarily be from 2003 and 2004.

```{r file_year}
idc %>% 
  filter(file %in% str_which(raw_urls, "2008")) %>% 
  count(year = year(date))
```



## Wrangle

### Address

```{r address_norm}
idc <- idc %>% 
  # combine street addr
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
idc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

```{r zip_norm}
idc <- idc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  idc$zip,
  idc$zip_norm,
  compare = valid_zip
)
```

### State

```{r state_norm}
idc <- idc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_progress}
progress_table(
  idc$state,
  idc$state_norm,
  compare = valid_state
)
```

### City

```{r city_norm}
idc <- idc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("ID", "DC", "IDAHO"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r city_swap}
idc <- idc %>%
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -match_abb,
    -match_dist,
    -city_match
  )
```

```{r city_progress}
many_city <- c(valid_city, extra_city)
progress_table(
  idc$city_raw,
  idc$city_norm,
  idc$city_swap,
  compare = many_city
)
```

```{r view_city}
idc %>% 
  filter(city_swap %out% many_city) %>% 
  count(city_swap, sort = TRUE)
```

## Conclude

1. There are `r comma(nrow(idc))` records in the database.
1. There are `r comma(sum(idc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(idc$na_flag))` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 5-digit `zip_norm` variable has been created with `campfin::normal_zip(df$zip)`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r create_proc_dir}
proc_dir <- dir_create(here("id", "contribs", "data", "processed"))
```

```{r write_clean}
idc <- idc %>% 
  select(
    -city_norm,
    city_norm = city_swap
  ) %>%
  rename_at(
    .vars = vars(ends_with("norm")),
    .funs = ~str_replace(., "_(.*)", "_clean")
  )

write_csv(
  x = idc,
  path = path(proc_dir, "id_contribs_clean.csv"),
  na = ""
)
```

