---
title: "Nursing Home Compare Data Diary -- Ownership"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
options(knitr.kable.NA = '')
set.seed(5)
```

```{r create_docs_dir, echo=FALSE}
docs <- fs::dir_create(here::here("nursing_home", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://www.medicare.gov/nursinghomecompare/Data/About.html

## Data

The master flat file is obtained from the [Medicare.gov][03]. The data is as current as March 16, 2020. 

```{r raw_dir}
raw_dir <- dir_create(here("nursing_home","data", "raw"))
```

```{r download, eval=FALSE}
landing_url <- 'https://data.medicare.gov/views/bg9k-emty/files/6ef81c64-2ebd-4ece-93af-8e462427939a?content_type=&filename=NursingHomeCompare_Revised_Flatfiles.zip'

zip_path <- path(raw_dir,str_extract(landing_url,"(?<=filename\\=).+"))

if (!all_files_new(raw_dir)) {
   download.file(url = landing_url, destfile = zip_path)
}

emgc_prp <- 'https://data.medicare.gov/api/views/9ezk-fzua/rows.csv?accessType=DOWNLOAD&sorting=true'

wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose",
      "--content-disposition",
      url
    )
  )
}

wget(emgc_prp, raw_dir)
```

```{r handle zip}
zip_contents <- 
  unzip(zip_path, list = TRUE) %>% 
  as_tibble() %>% 
  clean_names()

zip_data <- str_subset(zip_contents$name, "csv")

zip_others <- setdiff(zip_contents$name, zip_data)

if (not(all(file_exists(path(raw_dir, zip_data))))) {
  unzip(
    zipfile = zip_path,
    files = zip_data,
    exdir = raw_dir
  )
}

if (not(all(file_exists(path(docs, zip_others))))) {
  unzip(
    zipfile = zip_path,
    files = zip_others,
    exdir = docs
  )
}
```


The abbreviation information is stored in `NHC_DataDictionary_SNFQRP.pdf` and the record layouts are contained in `DataMedicareGov_MetadataAllTabs_v23.xlsx`. 

Secondly, we'll read the health deficiency data that contains the infection-control related citations. We can also view the record layout from the `HealthDeficiencies` sheet `DataMedicareGov_MetadataAllTabs_v23.xlsx`.

```{r read owner}
owner <- read_csv(dir_ls(raw_dir,regexp = "Owner.+")) %>% clean_names()
```                                                                                                              


```{r read owner dict}
owner_dict <- read_xlsx(str_subset(path(docs, zip_others),"DataMedicareGov"),
                        sheet = "Ownership")

kable(owner_dict)
```

### Missing

```{r na_count}
col_stats(owner, count_na)
```
We'll flag the records without any owner name (`r sum(owner$na_flag)` entries).
```{r na_flag}
owner <- owner %>% flag_na(owner_name)
```

### Duplicates
We can see there's no duplicate entry.
```{r dupe_flag}
owner <- flag_dupes(owner, dplyr::everything())
```

## Wrangle
To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.
### Date
From the record layout, we can see that we can clean the `association_date`with either an `NA` or an acutal date without the word "wince". 
```{r clean date}
owner <- owner %>% 
  mutate(association_date_clean = na_if(association_date, "No Date Provided") %>% 
           str_remove("since\\s") %>% 
           as.Date("%m/%d/%Y"))
```
### Percentage
We can also clean the `owner_percentage` field.

```{r clean perc}
owner <- owner %>% 
  mutate(owner_percentage_clean = owner_percentage %>% 
           na_if("NOT APPLICABLE") %>% 
           na_if("NO PERCENTAGE PROVIDED") %>% 
           str_remove("%") %>% 
           as.numeric())

```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
owner <- owner %>% 
    mutate(address_norm = normal_address(address,abbs = usps_street,
      na_rep = TRUE))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
prop_in(owner$zip, valid_zip, na.rm = T)

owner <- owner %>% 
    mutate(zip5 = normal_zip(zip, na_rep = T))

prop_in(owner$zip5, valid_zip, na.rm = T)
```

### State

The two digit state abbreviations are all valid.

```{r state_norm}
prop_in(owner$state, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good ownerart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
owner <- owner %>% 
      mutate(city_norm = normal_city(city,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE))

prop_in(owner$city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
owner <- owner %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```


After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  owner$city_raw,
  owner$city_norm,
  owner$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Nursing Home Compare Ownership City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nursing Home Compare Ownership City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Explore
### Categorical

#### Year
We can see that the data is pretty up to date. However, there're some fileds indicating the annual registration was recorded in `2055`, which is presumably a human error. 
```{r check year}
owner <- owner %>% 
  mutate(year = year(association_date_clean))
tabyl(owner$year)
```


## Conclude

```{r clean_glimpse}
glimpse(sample_n(owner, 20))
```

1. There are `r nrow(owner)` records in the database.
1. There are `r sum(owner$dupe_flag)` duplicate records in the database.
1. The range and distribution of `year` seems mostly reasonable except for a few entries.
1. There are `r sum(owner$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(path(raw_dir, "processed"))
```

```{r write_clean}
write_csv(
  x = owner %>% 
    mutate_if(is.character, str_to_upper) %>% 
    select(-city_norm) %>% 
    rename(city_clean = city_swap),
  path = path(clean_dir, "nursing_owner_clean.csv"),
  na = ""
)
```



