---
title: "New York Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ny", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [New York State Board of Elections][sboe] (SBOE). 

> The State Board of Elections was established in the Executive Department June
1, 1974 as a bipartisan agency vested with the responsibility for administration
and enforcement of all laws relating to elections in New York State. The Board
is also responsible for regulating disclosure and limitations of a Fair Campaign
Code intended to govern campaign practices. In conducting these wide-ranging
responsibilities, the Board offers assistance to local election boards and
investigates complaints of possible statutory violations. In addition to the
regulatory and enforcement responsibilities the board is charged with the
preservation of citizen confidence in the democratic process and enhancement in
voter participation in elections.

The NYSBOE database can be obtained from their Campaign Finance 
[disclosure reports page][cfv]. On that page, they elaborate on the availability
and accuracy of the website.

> ### Data Availability  
> This database contains all financial disclosure reports filed with NYSBOE from
July of 1999 to the present. Financial disclosure reports filed prior to the
1999 July Periodic report are either on file with the New York State Archives or
in storage with the New York State Board of Elections. For further information
or to obtain copies of these archived or stored filings, please call
1-800-458-3453. Each page costs 25¢ plus postage and copy orders must be
prepaid.
> 
> Electronically filed disclosure reports are generally available in the
database on the day they are received. A small number of candidates and
committees are either statutorily exempt or have applied for and obtained
exemptions from electronic filing. These filers will continue filing on paper
and their disclosure reports will become available as they are manually entered
into the database by NYSBOE staff.

> ### Data Accuracy  
> The majority of financial disclosure reports filed at NYSBOE are entered into
the database directly from e-mail, diskette, CD or DVD filings submitted by
committee treasurers or candidates. The information contained in paper filings
will be entered into the database exactly as it appears on the forms. Because
database searches retrieve information exactly the way it is reported and then
entered into the database, search results may be inaccurate and/or incomplete.
This will occur, for example, if filers do not adhere to the required format, do
not use the proper codes, misspell words or leave items blank. Although NYSBOE
carefully reviews disclosure reports and requires treasurers to submit amended
reports as needed, there will necessarily be delays before the review process is
completed and the information in the database is corrected.

The page also describes the format of their campaign finance database.

> ### Database Files in ASCII Delimited Format
> **Updated data files are uploaded during active filing periods after 4:00 P.M.
daily until the filing is complete.**
>
> **Note:** To match the filing data files to Filer Names by filer ID you will
need to [Download the Filer data file][fil]. Commcand.zip is a zipped file
containing the data file (commcand.asc) in ASCII delimited and two text files.
(filerec.txt contains the data file layout - codes.txt explains the codes used
in the data file).
>
> **All downloadable files are zipped files containing a data file in ASCII
delimited format and two text files. (efsrecb.txt contains the data file layout
- efssched.txt explains the different schedules as they apply to the
database).**
>
> [Download Data file containing ALL filings][all]. **Note:** This file is a
large file (238,994 KB) that contains over 6 million records. Do not attempt to
download this file unless you have a database to download the file to.

[sboe]: https://www.elections.ny.gov/INDEX.html
[cfv]: https://www.elections.ny.gov/CFViewReports.html
[fil]: https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/commcand.zip
[all]: https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/ALL_REPORTS.zip

## Download

We can use the link above to download a copy of the NYSBOE database to the
`/data/raw` directory.

```{r raw_dir}
raw_dir <- dir_create(here("ny", "expends", "data", "raw"))
```

First, we will download the `ALL_REPORTS.zip` file.

```{r create_zip_paths}
sboe_url <- "https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/"
zip_url <- str_c(sboe_url, "ALL_REPORTS.zip")
zip_path <- path(raw_dir, basename(zip_url))
```

Like they suggest, we will double check the size of the file before downloading.

```{r check_file_size}
url_file_size(zip_url)
```

```{r download_zip}
if (!file_exists(zip_path)) {
  download.file(
    url = zip_url,
    destfile = zip_path
  )
}
```

## Extract

If the `ALL_REPORTS.zip` file hasn't yet been unziped, we can do so now with
`unzip()`. First, we will list the files in the ZIP archive.

```{r zip_list}
(zip_content <- zip_path %>% 
  unzip(list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>% 
  mutate(across(length, as_fs_bytes)))
```

```{r zip_extract}
if (length(dir_ls(raw_dir)) == 0) {
  unzip(
    zipfile = zip_path,
    exdir = raw_dir,
    overwrite = TRUE
  )
}
```

## About

We can use the `EFSRECB.TXT` file to understand the layout and format of the
raw data file. This will help us read it properly.

```{r layout_file}
layout_file <- path(raw_dir, "EFSRECB.TXT")
efsrecb <- read_lines(
  file = layout_file, 
  skip_empty_rows = TRUE
)
```

First, we see the table describing the columns with their format and type.

```{r layout_read, echo=FALSE}
out_layout <- read_table(efsrecb[6:37], na = "")
out_layout %>% 
  select(-last_col()) %>% 
  mutate(across(FIELD, str_remove, "_\\d+")) %>% 
  mutate(across(everything(), str_replace_na)) %>% 
  mutate(across(everything(), str_replace, "NA", "")) %>% 
  kable()
```

```{r out_format, echo=FALSE}
out_format <- enframe(efsrecb[39:44]) %>% 
  separate(
    col = value,
    into = c("aspect", "value"),
    sep = ":"
  ) %>% 
  map_df(str_trim) %>% 
  mutate(across(aspect, make_clean_names))

out_format <- as.list(out_format$value) %>% 
  set_names(out_format$aspect) %>% 
  map(parse_guess)

noquote(out_format)
```

From the `EFSSCHED.TXT` file, we know Schedule L records contain the
expenditures we are interested in.

```{r sched_file}
sched_file <- path(raw_dir, "EFSSCHED.TXT")
efssched <- read_lines(sched_file, skip_empty_rows = TRUE)
```

```{r sched_trim, echo=FALSE}
efssched <- str_remove(efssched, "^(\n|\t+)")
```

```{r sched_data, echo=FALSE}
out_scheds <- efssched[35:53] %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value, 
    into = c("sched", "desc"), 
    sep = "\\s-\\s"
  ) %>% 
  map_df(str_trim)

kable(out_scheds, col.names = c("Schedule", "Description"))
```

```{r out_columns, echo=FALSE}
x <- str_pad(
  string = efssched[4:34], 
  width = nchar(efssched[4]), 
  pad = " ",
  side = "right"
)
out_cols <- read_table(x[str_detect(x, "X|Q")], na = "NULL")
out_cols <- clean_names(na_if(out_cols, ""))
```

```
FIELD NAMES     A  B  C D  E  F  G  H  I  J K  L  M  N  O  P Q
--------------------------------------------------------------
DATE1           X  X  X X  X  X  X  X  X  X X  X  X  X     X X
DATE2                                       X  X  X
CONTRIB_CODE    X       X                               X  X
CONTRIB_TYPE            X
CORP            X  X  X X  X  X  X  X  X  X X  X  X  X  X  X X
FIRST_NAME      X       X                                  X
MID_INIT        X       X                                  X
LAST_NAME       X       X                                  X
ADDR            X  X  X X  X  X  X  X  X  X X  X  X  X     X X
CITY            X  X  X X  X  X  X  X  X  X X  X  X  X     X X
STATE           X  X  X X  X  X  X  X  X  X X  X  X  X     X X
ZIP             X  X  X X  X  X  X  X  X  X X  X  X  X     X X
CHECK_NO        X  X  X       X  X  X     X       X        X X
CHECK_DATE                                X
AMOUNT          X  X  X X  X  X  X  X  X  X X  X  X  X     X X
AMOUNT2                                     X        X  X
DESCRIPTION             X
OTHER_RECPT                X
PURPOSE_CODE1                 X                      X  X
PURPOSE_CODE2                                                X
EXPLANATION                   X                      X  X    X
XFER_TYPE                        X  X
CHKBOX                                 X                      
```

## Fix

```{r fix_replace}
out_file <- path(raw_dir, "ALL_REPORTS.out")
fix_file <- path(raw_dir, "FIX_REPORTS.out")
if (!file_exists(fix_file)) {
  read_lines(out_file) %>% 
    str_replace_all('"""(\\w+),"""', "\"\\1\"") %>%
    str_replace_all("(?<!,)\"\"(?!,)", "'") %>%
    str_replace_all("(?<=,)\"\"(?!,|$)", "\"") %>% 
    str_replace_all("(?<!,|^)\"\"(?=,)", "\"") %>% 
    str_replace_all("(?=!\",)\"\"\"(?=,\")", "\"\"") %>% 
    str_replace_all("(?=!\",)\"(?=,\")", "\"\"") %>% 
    str_replace_all("(?<!^|,)\"(?!,(?=\")|$)", "'") %>%
    write_lines(fix_file)
  Sys.sleep(10); flush_memory()
}
```

## Read

First, we'll read the `ALL_REPORTS.out` text file containing, well, all reports.

```{r ny_names, echo=FALSE}
ny_names <- c(
  "filer_id",
  "rpt_id",
  "xfer_code",
  "eyear",
  "t3_trid",
  "date",
  "date2",
  "contrib_code",
  "type_code",
  "corp",
  "first",
  "middle",
  "last",
  "addr",
  "city",
  "state",
  "zip",
  "check_no",
  "check_date",
  "amount",
  "amount2",
  "descr",
  "recpt",
  "purpose1",
  "purpose2",
  "explain",
  "xfer_type",
  "checkbox",
  "crerec_uid",
  "crerec_date"
)
```

```{r read_out}
nye <- read_delim( # 12,037,086
  file = fix_file, 
  delim = ",", 
  col_names = ny_names,
  escape_double = TRUE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character(),
    eyear = col_integer(),
    date = col_date("%m/%d/%Y"),
    date2 = col_date("%m/%d/%Y"),
    amount = col_double(),
    amount2 = col_double(),
    crerec_date = col_datetime("%m/%d/%Y %H:%M:%S")
  )
)
```

```{r read_problems}
# two problems left
# first encoding
# second one last quote
select(problems(nye), -file)
```

The New York campaign finance database is a single data frame with 
`r comma(nrow(nye))` rows and `r ncol(nye)` columns. This data frame contains 
data for _all_ campaign finance. We will filter out any transaction that is not
an expenditure.

```{r filter_rows}
# filter only expenditures
expend_sched <- out_scheds$sched[str_which(out_scheds$desc, "Expenditure")]
all_rows <- nrow(nye)
nye <- filter(nye, xfer_code %in% expend_sched)
percent(nrow(nye)/all_rows, 0.1)
flush_memory()
```

We can remove any columns not relevant to schedules F and L via `EFSSCHED.TXT`.

```{r filter_cols}
(sched_cols <- which(out_cols$f == "X" | out_cols$l == "X"))
nye <- select(nye, c(1:5, sched_cols + 5))
```

```{r raw_head}
head(nye)
```

## Join

Now we must use the separate `commcand.zip` file to add information on the
_filers_ of the expenditure reports. This file can similarly be downloaded from
the SBOE, extracted, and read as was done by with the reports file.

```{r com_download}
com_url <- str_c(sboe_url, "commcand.zip")
com_zip <- path(raw_dir, basename(com_url))
if (!file_exists(com_zip)) {
  download.file(com_url, com_zip)
}
```

```{r com_list, echo=FALSE}
(com_content <- com_zip %>% 
  unzip(list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>% 
  mutate(across(length, as_fs_bytes)))
```

```{r com_extract}
com_files <- unzip(
  zipfile = com_zip,
  exdir = raw_dir,
  overwrite = TRUE
)
```

First, we can read `FILEREC.TXT` which lists filers by their `filer_id`.

```{r com_fil}
fil_path <- path(raw_dir, "FILEREC.TXT")
fil_layout <- read_lines(fil_path)
fil_layout <- clean_names(read_table(fil_layout[8:22]))
fil_layout$field <- make_clean_names(fil_layout$field)
kable(fil_layout)
```

```{r com_read}
(commcand <- raw_dir %>% 
  path("COMMCAND.txt") %>% 
  read_lines() %>% 
  # fix quote issues
  str_replace_all("(?<!,|^)\"(?!,|$)", "'") %>% 
  read_delim(
    delim = ",",
    col_names = fil_layout$field,
    col_types = cols(
      .default = col_character(),
      district = col_integer(),
      office = col_integer()
    )
  ))
```

Then, we can read the various other files used to define the codes in the
`COMMCAND.txt` file.

```{r com_codes}
code_file <- path(raw_dir, "CODES.TXT")
codes <- str_remove(read_lines(code_file), "^\t|\\s+")
```

```{r com_office}
off_codes <- read_tsv(
  file = str_replace(codes[9:83], "\\s", "\t"), 
  col_names = c("office_code", "office"),
  col_types = cols(
    office_code = col_integer()
  )
)
```

```{r com_district}
dis_codes <- codes[88:90] %>% 
  str_replace("\\s+(?=\\d)", "\t") %>% 
  read_tsv(col_names = c("district_type", "range")) %>% 
  # convert to long format
  separate(range, c("a", "b"), convert = TRUE) %>% 
  rowwise() %>% 
  mutate(district = list(seq(a, b))) %>% 
  unnest(district) %>% 
  select(district, district_type) %>% 
  mutate(across(district_type, str_remove, "\\s.*"))
```

```{r com_types}
com_codes <- read_tsv(
  file = codes[96:116], 
  col_names = c("committee_code", "com_type")
)
```

```{r com_report}
rpt_codes <- str_replace(codes[121:136], "\\s{4}", "\t")
rpt_codes <- read_tsv(rpt_codes, col_names = c("code", "report"))
```

Now these codes can be joined to the filers data frame and the column names
can be clarified for joining to the reports data.

```{r com_join}
commcand <- commcand %>% 
  left_join(off_codes, by = c("office" = "office_code")) %>% 
  select(-office, office = office.y) %>% 
  relocate(office, .before = district) %>% 
  rename(committee_code = committee_type) %>% 
  left_join(com_codes) %>% 
  relocate(com_type, .after = committee_code) %>% 
  select(-committee_code) %>% 
  rename_with(~str_c("fil_", .), 10:13) %>% 
  rename_with(~str_replace(., "filer_", "fil_")) %>% 
  rename_with(~str_remove(., "_name"), 8:9) %>% 
  mutate(active = status == "ACTIVE", .keep = "unused", .after = fil_type) %>%
  rename(fil_addr = fil_address) %>% 
  filter(fil_id %in% nye$filer_id)
```

Now that the `COMMCAND.txt` file is complete, we can join it onto the reports
data frame, clarifying which columns belong to filers and which to vendors.

```{r raw_join}
nye <- nye %>% 
  rename(ven_name = corp, fil_id = filer_id) %>% 
  rename_with(~str_c("ven_", .), 9:12) %>% 
  left_join(commcand, by = "fil_id")
```

## Explore

```{r head_tail}
glimpse(nye)
tail(nye)
```

### Missing

```{r na_glimpse}
col_stats(nye, count_na)
```

```{r na_flag}
key_vars <- c("ven_name", "fil_name", "amount", "date")
nye <- nye %>% flag_na(all_of(key_vars))
percent(mean(nye$na_flag), 0.01)
```

```{r na_view}
nye %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

### Duplicate

```{r dupe_flag}
d1 <- duplicated(nye, fromLast = FALSE)
d2 <- duplicated(nye, fromLast = TRUE)
nye <- mutate(nye, dupe_flag = d1 | d2)
rm(d1, d2); flush_memory()
mean(nye$dupe_flag)
```

```{r dupe_view}
nye %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r dist_count}
col_stats(nye, n_distinct)
```

```{r bar_report, echo=FALSE}
explore_plot(nye, rpt_id)
explore_plot(nye, active)
explore_plot(nye, com_type) + scale_x_truncate(15)
explore_plot(nye, office)
```

### Amount

```{r amount_summary}
summary(nye$amount)
```

`r percent(mean(nye$amount == 0, na.rm = TRUE))` of `amount` values are zero.

```{r zero_amount}
mean(nye$amount <= 0, na.rm = TRUE)
```

The largest amount of `r dollar(max(nye$amount, na.rm = T))` has an 
`explanation` of "Transfer."

```{r glimpse_max}
glimpse(nye[which.max(nye$amount), ])
```

```{r amount_hist, echo=FALSE}
nye %>% 
  filter(amount >= 1) %>% 
  ggplot(aes(x = amount)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(8, "Dark2")[3]) +
  geom_vline(xintercept = median(nye$amount, na.rm = TRUE)) +
  geom_vline(xintercept = mean(nye$amount, na.rm = TRUE), linetype = 2) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "New York Expenditure Amount Distribution",
    subtitle = "Solid for mean, dashed for median",
    x = "Amount",
    y = "Count",
    caption = "Source: New York State Board of Elections"
  )
```

#### Date

We can add the calendar date of the expenditure with `lubridate::year()`.

```{r year_add}
nye <- mutate(nye, year = year(date))
```

```{r date_bad}
count_na(nye$year)
sum(nye$year < 1999, na.rm = TRUE)
sum(nye$year > 2020, na.rm = TRUE)
```

```{r date_flag}
nye <- nye %>% 
  mutate(
    date_flag = year < 1999 | year > 2020,
    date_clean = as_date(ifelse(date_flag, NA, date)),
    year_clean = year(date_clean)
  )
```

```{r date_count}
percent(mean(nye$date_flag, na.rm = TRUE), 0.01)
```

```{r bar_year, echo=FALSE}
nye %>% 
  count(year_clean) %>% 
  mutate( even = !is_even(year_clean)) %>% 
  ggplot(aes(x = year_clean, y = n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 1999:2020) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Expenditures Made by Year",
    x = "Year",
    y = "Count",
    fill = "Election Year",
    caption = "Source: New York State Board of Elections"
  )
```

```{r year_bar_median, echo=FALSE}
nye %>% 
  group_by(year_clean) %>% 
  summarise(median = median(amount, na.rm = TRUE)) %>% 
  mutate(even = is_even(year_clean)) %>% 
  ggplot(aes(x = year_clean, y = median)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_x_continuous(breaks = 1999:2020) +
  scale_y_continuous(labels = dollar) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Expenditure Medians",
    x = "Year",
    y = "Median Amount",
    fill = "Election Year",
    caption = "Source: New York State Board of Elections"
  )
```

```{r year_bar_sum, echo=FALSE}
nye %>% 
  group_by(year_clean) %>% 
  summarise(sum = sum(amount, na.rm = TRUE)) %>% 
  mutate(even = is_even(year_clean)) %>% 
  ggplot(aes(x = year_clean, y = sum)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_y_continuous(labels = function(x) dollar(x/1e6)) +
  scale_x_continuous(breaks = 1999:2020) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Expenditures Totals",
    x = "Year",
    y = "Total Amount (Million)",
    fill = "Election Year",
    caption = "Source: New York State Board of Elections"
  )
```

```{r month_line_amount}
nye %>% 
  filter(!date_flag) %>% 
  mutate(
    month = month(date_clean),
    even = is_even(year_clean)
  ) %>% 
  group_by(even, month) %>% 
  summarise(sum = sum(amount, na.rm = TRUE)) %>% 
  ggplot(aes(x = month, y = sum)) +
  geom_line(aes(color = even), size = 2) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_y_continuous(labels = function(x) dollar(x/1e6)) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Expenditures Mean by Month",
    x = "Year",
    y = "Total Amount (Million)",
    color = "Election Year",
    caption = "Source: New York State Board of Elections"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

Each normalization will be applied to both `filer_*` and `vendor_*` variables.

### Address

First, we will normalize the street address by removing punctuation and
abbreviating common USPS words.

```{r address_normal}
nye <- mutate(
  .data = nye,
  across(
    .cols = ends_with("_addr"),
    .fns = list(norm = normal_address),
    abbs = usps_street,
    na_rep = TRUE
  )
)
```

We can see how this improves consistency across the address fields.

```{r address_view, echo=FALSE}
nye %>% 
  select(contains("fil_addr")) %>% 
  drop_na() %>% 
  sample_n(10)
```

### ZIP

We can `*_zip` by lopping off the uncommon four-digit extensions and removing
common invalid codes like 00000 and 99999.

```{r zip_normal}
nye <- mutate(
  .data = nye,
  across(
    .cols = ends_with("zip"),
    .fns = list(norm = normal_zip),
    na_rep = TRUE
  )
)
```

This brings our valid percentage to 
`r percent(prop_in(nye$fil_zip_norm, valid_zip, na.rm = TRUE), 0.01)`.

```{r count_zip_post, collapse=TRUE}
progress_table(
  nye$ven_zip,
  nye$ven_zip_norm,
  compare = valid_zip
)
```

### State

```{r state_normal}
nye <- mutate(
  .data = nye,
  across(
    .cols = ends_with("state"),
    .fns = list(norm = normal_state),
    abbreviate = TRUE,
    na_rep = TRUE,
    valid = NULL
  )
)
```

We can also manually fix typos for NY.

```{r fix_ny_states}
state_ny_fix <- function(string) {
  string %>% 
    str_replace("^N$",  "NY") %>% 
    str_replace("^MY$", "NY") %>% 
    str_replace("^NT$", "NY") %>% 
    str_replace("^NU$", "NY") %>% 
    str_replace("^BY$", "NY")
}

nye <- mutate(nye, across(ends_with("state_norm"), state_ny_fix))
```

```{r state_progress}
progress_table(
  nye$ven_state,
  nye$ven_state_norm,
  compare = valid_state
)
```

```{r}
nye %>% 
  filter(ven_state != ven_state_norm) %>% 
  count(ven_state, ven_state_norm, sort = TRUE)
```

### City

The `city` value is the hardest to normalize. We can use a four-step system to
functionally improve the searchability of the database.

1. **Normalize** raw values with `campfin::normal_city()`
1. **Match** normal values with the _expected_ value for that ZIP code
1. **Swap** normal values with the expected value if they are _very_ similar
1. **Refine** swapped values the [OpenRefine][or] and keep good changes

[or]: https://github.com/OpenRefine/OpenRefine/wiki/Clustering-In-Depth

The raw `filer_city` values are fairly normal, with
`r percent(prop_in(nye$filer_city, valid_city, na.rm = TRUE))` already in 
`valid_city`. We will aim to get this number over 99% using the above steps.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r fix_ny_city}
city_ny_fix <- function(string) {
  string %>% 
    str_replace("^nyc$",  "NEW YORK") %>% 
    str_replace("^NYC$", "NEW YORK") %>%  
    str_replace("^S\\.I\\.$", "SI")
}
nye <- mutate(nye, across(ends_with("city"), list(norm = city_ny_fix)))
```

```{r city_normal}
nye <- mutate(
  .data = nye,
  across(
    .cols = ends_with("city_norm"),
    .fns = normal_city,
    abbs = usps_city,
    states = c("NY", "DC", "NEW YORK"),
    na = invalid_city,
    na_rep = TRUE
  )
)
```

```{r new_city_na, echo=FALSE}
nye %>% 
  filter(is.na(ven_city_norm) & !is.na(ven_city)) %>% 
  count(ven_city, ven_city_norm, sort = TRUE)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

[09]: https://en.wikipedia.org/wiki/Levenshtein_distance

```{r ven_city_swap}
nye <- nye %>% 
  left_join(
    y = zipcodes,
    by = c(
      "ven_state_norm" = "state",
      "ven_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(ven_city_norm, city_match),
    match_dist = str_dist(ven_city_norm, city_match),
    ven_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = ven_city_norm
    )
  )
```

This swap caught many small differences between city names.

```{r city_swap_dist}
nye %>% 
  filter(match_dist == 1) %>% 
  count(ven_city_norm, ven_city_swap, sort = TRUE)
```

We can also see how many abbreviations were found and replaced.

```{r city_swap_abb}
nye %>% 
  filter(match_abb) %>% 
  count(ven_city_norm, ven_city_swap, sort = TRUE)
```

But others were missed if the city match by `zip_norm` was different.

```{r city_swap_non_abb}
nye %>% 
  filter(ven_city_norm == "LIC") %>% 
  count(ven_city_norm, city_match, ven_city_swap, sort = TRUE)
```

We can use the _successful_ abbreviation swaps as a database to catch others.

```{r city_abbs}
ny_city_abbs <- nye %>% 
  filter(match_abb) %>%
  count(ven_city_norm, ven_city_swap, sort = TRUE) %>% 
  select(ven_city_norm, ven_city_abb = ven_city_swap) %>% 
  head(20)
```

```{r city_abb_look}
nye <- nye %>% 
  left_join(ny_city_abbs, by = "ven_city_norm") %>% 
  mutate(ven_city_swap = coalesce(ven_city_abb, ven_city_swap))
```

```{r city_remove}
nye <- nye %>% 
  select(
    -city_match,
    -match_abb,
    -match_dist,
    -ven_city_abb
  )
```

```{r fil_city_swap}
nye <- nye %>% 
  left_join(
    y = zipcodes,
    by = c(
      "fil_state_norm" = "state",
      "fil_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(fil_city_norm, city_match),
    match_dist = str_dist(fil_city_norm, city_match),
    fil_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = fil_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_abb,
    -match_dist
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

```{r refine_city}
good_refine <- nye %>% 
  mutate(
    ven_city_refine = ven_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(ven_city_refine != ven_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "ven_city_refine" = "city",
      "ven_state_norm" = "state",
      "ven_zip" = "zip"
    )
  )
```

```{r view_city_refines, echo=FALSE}
good_refine %>%
  count(
    ven_state_norm, 
    ven_zip_norm, 
    ven_city_norm, 
    ven_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
nye <- nye %>% 
  left_join(good_refine) %>% 
  mutate(ven_city_refine = coalesce(ven_city_refine, ven_city_swap))
```

#### Progress

```{r city_remain, include=FALSE}
many_city <- c(valid_city, extra_city)
nye %>% 
  filter(ven_city_refine %out% many_city) %>% 
  count(ven_city_refine, ven_state_norm, ven_zip_norm, sort = TRUE) %>% 
  drop_na()
```

```{r city_extra}
extra_city <- extra_city %>% 
  c(
    "WINNIPEG", 
    "LAKE SUCCESS", 
    "NORTH VALLEY STREAM",
    "NORTH BELLMORE",
    "SOLVAY",
    "PELHAM MANOR",
    "GREEN ISLAND"
  )
```

```{r echo=FALSE}
nye$ven_city_refine <- nye$ven_city_refine %>% 
  str_replace("\\sCIT$", "CITY") %>% 
  str_replace("^GARDEN CITY PAR$", "GARDEN CITY PARK")
```

```{r ven_city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(nye$ven_city),
  nye$ven_city_norm,
  nye$ven_city_swap,
  nye$ven_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

```{r fil_city_progress}
progress_table(
  str_to_upper(nye$fil_city),
  nye$fil_city_norm,
  nye$fil_city_swap,
  compare = many_city
)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nye$ven_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
nye <- nye %>% 
  select(
    -ven_city_norm,
    -ven_city_swap,
    ven_city_clean = ven_city_refine,
    -fil_city_norm,
    fil_city_norm = fil_city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(nye, 50))
```

1. There are `r comma(nrow(nye))` records in the database.
1. There are `r comma(sum(nye$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nye$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.
The data frame will be split into two files, the original file types for loans
over and under $150,000.

```{r clean_paths}
clean_dir <- dir_create(here("ny", "expends", "data", "clean"))
clean_path <- path(clean_dir, "ny_expends_clean.csv")
write_csv(nye, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(nye)),
  type = md_code(map_chr(nye, typeof)),
  def = c(
    "Unique filer ID",
    "Report type code",
    "Transfer type code",
    "Election year made",
    "Semi-unqiue T3 ID",
    "Date expenditure made",
    "Date expenditure refunded",
    "Payee vendor name",
    "Payee vendor address",
    "Payee vendor city",
    "Payee vendor state",
    "Payee vendor ZIP code",
    "Check number",
    "Expenditure amount",
    "Expenditure primary purpose",
    "Expenditure explaination",
    "Filer committee name",
    "Filer committee type",
    "Filer active (TRUE/FALSE)",
    "Committee type",
    "Office candidate seeking",
    "District candidate seeking",
    "Committee treasurer first name",
    "Committee treasurer last name",
    "Filer committee address",
    "Filer committee city",
    "Filer committee state",
    "Filer committee ZIP code",
    "Flag indicating missing variable",
    "Flag indicating duplicate record",
    "Calendar year expenditure made",
    "Flag indicating invalid date",
    "Date with invalid values removed",
    "Year with invalid years removed",
    "Normalized vendor address",
    "Normalized filer address",
    "Normalized vendor ZIP code",
    "Normalized filer ZIP code",
    "Normalized vendor state",
    "Normalized filer state",
    "Normalized vendor city",
    "Normalized filer city"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
