---
title: "Nebraska Lobbyists"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","ne", "lobbying", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  pdftools, # read pdf file text
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][01] GitHub repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects 

## Data

The data is obtained via a public records request fulfilled on July 27, 2023. A zip archive was provided by the legislative clerk's office. The data includes records from 2000 to July 21, 2023. 

> The following reports identify lobbyists registered in Nebraska with the
Office of the Clerk of the Legislature.
> 
> ##### Lists of Registered Lobbyists
> * [Lobby Registration Report by Principal][1]
> * [Lobby Registration Report by Lobbyist][2]
> * [Lobbyist/Principal Expenditures Report][3]
> * [Lobbyist/Principal Statement of Activity][4]
> * [Counts of Lobbyists/Principals][5]

[1]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/principallist.pdf
[2]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/lobbyistlist.pdf
[3]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/expense.pdf
[4]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/activity_final_by_bill.pdf
[5]: https://nebraskalegislature.gov/FloorDocs/Current/PDF/Lobby/counts.pdf


## Import

```{r}
raw_dir <- dir_create(here::here("state","ne", "lobbying", "data","raw"))

nelr <- read_csv(dir_ls(raw_dir, glob = "*.csv"))

nelr <- nelr %>% clean_names()
```

```{r}
nelr <- nelr %>% mutate(across(.cols = ends_with("_date"), mdy))
```


## Explore

```{r glimpse}
glimpse(nelr)
tail(nelr)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(nelr, count_na)
```


### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
nelr <- flag_dupes(nelr, everything())
sum(nelr$dupe_flag)
mean(nelr$dupe_flag)
```

## Wrangle
First, we will rename some columns. 
```{r}
nelr <- nelr %>% rename_with(~str_replace(., "principal_", "pri_")) %>% 
  rename_with(~str_replace(., "lobbyist_", "lob_"))
```
### Phone

```{r phone_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("phone")),
    .fun = list(norm = normal_phone)
  )
```

### Address

```{r address_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("address")),
    .fun = list(norm = normal_address),
    abbs = usps_street,
    na_rep = TRUE
  )
```

```{r address_view, echo=FALSE}
nelr %>% 
  select(
    pri_address,
    pri_address_norm,
    lob_address,
    lob_address_norm
  ) %>% 
  distinct() %>% 
  sample_frac()
```



### ZIP

```{r zip_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("zip")),
    .fun = list(norm = normal_zip),
    na_rep = TRUE
  )
```

```{r zip_progress, echo=FALSE}
progress_table(
  nelr$lob_zip,
  nelr$lob_zip_norm,
  nelr$pri_zip,
  nelr$pri_zip_norm,
  compare = valid_zip
)
```

### State

The `*_state` components do not need to be wrangled.

```{r state_progress}
prop_in(nelr$lob_state, valid_state)
prop_in(nelr$pri_state, valid_state)
```

### City

```{r city_normal}
nelr <- nelr %>% 
  mutate_at(
    .vars = vars(ends_with("city")),
    .fun = list(norm = normal_city),
    na_rep = TRUE
  )
```

```{r city_swap_lob}
nelr <- nelr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lob_state" = "state",
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lob_city_norm, city_match),
    match_dist = str_dist(lob_city_norm, city_match),
    lob_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap_pri}
nelr <- nelr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "pri_state" = "state",
      "pri_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(pri_city_norm, city_match),
    match_dist = str_dist(pri_city_norm, city_match),
    pri_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = pri_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_progress_lob, echo=FALSE}
progress_table(
  nelr$lob_city,
  nelr$lob_city_norm,
  nelr$lob_city_swap,
  compare = c(valid_city, extra_city)
)
```

```{r city_progress_pri, echo=FALSE}
progress_table(
  nelr$pri_city,
  nelr$pri_city_norm,
  nelr$pri_city_swap,
  compare = c(valid_city, extra_city)
)
```

## Export

```{r clean_order}
nelr <- nelr %>% 
  select(-c(
    lob_city_clean = lob_city_swap,
    pri_city_clean = pri_city_swap
  )) %>% 
  mutate(
    year = year(registration_date),
    .before = lob_name
  ) %>% 
  rename_with(~str_replace(., "_norm", "_clean"))
```

```{r clean_write}
clean_dir <- dir_create(here("state","ne", "lobbying", "data", "clean"))
clean_path <- path(clean_dir, "ne_lobby_reg.csv")
write_csv(nelr, clean_path, na = "")
```

```{r clean_check}
nrow(nelr)
file_size(clean_path)
guess_encoding(clean_path)
```
