---
title: "Oregon Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("or", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data for the state of Oregon can be obtained from the state's
[Open Data][od] portal. Each record represents a single contract between the
state and an outside vendor providing goods or services.

[od]: https://data.oregon.gov

> This composite dataset is sorted by agency, and contains all versions of
contracts and amendments issued in ORPIN from July 1, 2012 to June 30, 2019. The
contract start date may be different than the contract issued date, or date the
data entry was completed, and the contract was entered into ORPIN system. For
more information: https://www.oregon.gov/transparency/Pages/index.aspx

The [Oregon Procurement Information Network (ORPIN)][orpin] also provides a
[data dictionary][dict] as a PDF.

[orpin]: https://orpin.oregon.gov/open.dll/welcome
[dict]: https://www.oregon.gov/transparency/Documents/2019_ORPIN_Contracts_Data_Dictionary.pdf

|Variable            |Description                                              |
|:-------------------|:--------------------------------------------------------|
|Agency Number       |Unique identifier assigned to the agency                 |
|Agency Name         |Name of the agency for whom the contract was issued      |
|Award Number        |Contract number, revisions displays as decimal points.   |
|Award Title         |Title of the contract                                    |
|Award Type          |Indicates which contract type was used                   |
|Contractor          |Company name of the contractor                           |
|Street Address      |Street address of the contractor                         |
|City                |City of the contractor                                   |
|State               |State of the contractor                                  |
|Zip                 |Zip code of the contractor                               |
|Original Start Date |Date the contract first became active                    |
|Amendment Date      |Date when an amendment or revision was issued            |
|Expiration Date     |Date when the contract expired or will expire            |
|Award Value         |Estimated contract value when first awarded              |
|Amendment Value     |Dollar amount change for amendment version               |
|Total Award Value   |Total Estimated Contract Value at the time of issuing    |

## Read

The raw data can be read directly from the portal using `vroom::vroom()`.

```{r raw_read}
orc <- vroom(
  file = "https://data.oregon.gov/api/views/6e9e-sfc4/rows.csv",
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    `Original Start Date` = col_date_usa(),
    `Amendment Date` = col_date_usa(),
    `Expiration Date` = col_date_usa(),
    `Original Award Value` = col_double(),
    `Amendment Value` = col_double(),
    `Total Award Value` = col_double()
  )
)
```

```{r raw_rename, echo=FALSE}
orc <- rename(
  .data = orc,
  agency_id = agency_number,
  agency = agency_name,
  doc_num = document_number,
  amend_num = amendment_number,
  vendor = contractor,
  address1 = street_address,
  address2 = address_line_2,
  zip = zip_code,
  start_date = original_start_date,
  amend_date = amendment_date,
  end_date = expiration_date,
  amount = original_award_value,
  amend_amount = amendment_value,
  total_amount = total_award_value
)
```

## Explore

```{r glimpse}
glimpse(orc)
tail(orc)
```

### Missing

There are very few contracts missing values. None need to be flagged.

```{r na_count}
col_stats(orc, count_na)
```

### Duplicates

Ignoreing the semi-unique `doc_num` variable, there are a number of duplicated
records that need to be flagged.

```{r dupe_flag}
orc <- flag_dupes(orc, -doc_num)
sum(orc$dupe_flag)
```

```{r dupe_view}
orc %>% 
  filter(dupe_flag) %>% 
  select(start_date, agency, amount, vendor)
```

### Categorical

```{r distinct_count}
col_stats(orc, n_distinct)
```

```{r distinct_plots}
explore_plot(orc, award_type)
explore_plot(orc, agency) + scale_x_truncate()
```

### Amounts

```{r ammount_summary}
noquote(map_chr(summary(orc$amount), dollar))
percent(mean(orc$amount <= 0), 0.01)
```

```{r hist_amount, echo=FALSE}
orc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Oregon Contracts Amount Distribution",
    caption = "Source: ORPIN",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
orc <- mutate(orc, start_year = year(start_date))
```

There are a handful of very old or future contracts.

```{r date_range}
min(orc$start_date)
sum(orc$start_year < 2004)
max(orc$start_date)
sum(orc$start_date > today())
```

```{r bar_year, echo=FALSE}
orc %>% 
  count(start_year) %>% 
  mutate(even = is_even(start_year)) %>% 
  ggplot(aes(x = start_year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  coord_cartesian(xlim = c(2004, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oregon Contracts by Year",
    caption = "Source: ORPIN",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
orc <- orc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
orc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
orc <- orc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  orc$zip,
  orc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
orc <- orc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
orc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  orc$state,
  orc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
orc <- orc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("OR", "DC", "OREGON"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
orc <- orc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Progress

```{r city_count}
many_city <- c(valid_city, extra_city)
orc %>% 
  filter(city_swap %out% many_city) %>% 
  count(city_swap, state_norm, sort = TRUE)
```

```{r city_progress, echo=FALSE}
many_city <- c(many_city, "MILWAUKIE")
progress <- progress_table(
  str_to_upper(orc$city_raw),
  orc$city_norm,
  orc$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(orc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Oregon City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oregon City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
orc <- orc %>% 
  select(
    -city_norm,
    city_clean = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(orc, 20))
```

1. There are `r comma(nrow(orc))` records in the database.
1. There are `r comma(sum(orc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(orc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("or", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "or_contracts_clean.csv")
write_csv(orc, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, path.abbrev))
```

## Upload

Using the duckr R package, we can wrap around the [duck] commnand line tool to
upload the file to the IRW S3 server.

[duck]: https://duck.sh

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
