---
title: "Wyoming Campaign Expenditures Data Diary"
subtitle: "Wyoming Expenditures July 31"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("kiernann/campfin")
pacman::p_load(
  campfin,
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & database
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  rvest # scrape html
)
```

```{r fix_fun, echo=FALSE}
# fix conflict
here <- here::here
# custom utility functions
"%out%" <- Negate("%in%")
print_all <- function(df) df %>% print(n = nrow(.)) 
# load data
data("zipcode")
zipcode <- geo
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

The data comes from the Wyoming Secretary of State. [Link to the data download][03].

This txt file contains 11 years worth of data (2008-2018) in 8 columns, Filer Type,	Filer Name,	Payee,	Purpose,	Date,	City,State & ZIP and Filing Status	Amount as of this writing. 

[03]: https://www.wycampaignfinance.gov/WYCFWebApplication/GSF_SystemConfiguration/SearchExpenditures.aspx "source"

### About

> Wyoming's Campaign Finance Information System (WYCFIS) exists to provide a mechanism for online filing of campaign finance information and to provide full disclosure to the public. This website contains detailed financial records and related information that candidates, committees, organizations and parties are required by law to disclose.
Wyoming requires all statewide candidates, candidate committees, political action committees, organizations and political parties to file electronically online using this system. 


## Import

### Download

Download raw, **immutable** data file. Go to https://www.wycampaignfinance.gov/WYCFWebApplication/GSF_SystemConfiguration/SearchExpenditures.aspx, leave the fields blank, and click the "All" tab and hit "Search". After the table is populated, click "Export"

```{r raw_dir}
# create a directory for the raw data
raw_dir <- here("wy", "expends", "data", "raw")
dir_create(raw_dir)
```


### Read

```{r read_many, echo=FALSE}
wy <- 
  dir_ls(path = raw_dir) %>% 
  map(
    read_delim,
    delim = ",",
    escape_double = FALSE,
    escape_backslash = FALSE,
    col_types = cols(
      .default = col_character(),
      Date = col_date("%m/%d/%Y"),
      Amount = col_double()
    )  
    ) %>% 
  bind_rows() %>% 
  distinct() %>% 
  select(-starts_with("X")) %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper) %>% 
  map_if(is_character, str_replace_all, "\\\"", "\'") %>% 
  as_tibble()
```

## Explore

There are `nrow(wy)` records of `length(wy)` variables in the full database.

```{r glimpse}
head(wy)
tail(wy)
glimpse(wy)
```

### Distinct

The variables range in their degree of distinctness.


```{r n_distinct}
wy %>% glimpse_fun(n_distinct)
```

We can explore the distribution of the least distinct values with `ggplot2::geom_bar()`.

```{r plot_bar, echo=FALSE}
ggplot(data = wy) +
  geom_bar(aes(filer_type)) 
```

Or, filter the data and explore the most frequent discrete data.

```{r plot_bar2, echo=FALSE}
wy %>% 
  group_by(payee)  %>% 
  summarize(total_spent = sum(amount)) %>% 
  arrange(desc(total_spent)) %>% 
  head(10) %>% 
  ggplot(aes(x=payee, y=total_spent)) + 
  geom_col() +
    labs(title = "Wyoming Campaign Expenditures",
       caption = "Source: Wyoming Secretary of State") +
  scale_y_continuous(labels = scales::dollar) +
  coord_flip() +
  theme_minimal()
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
wy %>% glimpse_fun(count_na)
```

We will flag any records with missing values in the key variables used to identify an expenditure.
There are `r sum(wy$na_flag)` columns in city_state_zip that are NAs
```{r na_flag}
wy <- wy %>% mutate(na_flag = is.na(city_state_zip))
```

### Duplicates
There are no duplicates
```{r get_dupes, collapse=TRUE}
wy_dupes <- get_dupes(wy)
```

### Ranges

#### Amounts

```{r}
summary(wy$amount)
```

See how the campaign expenditures were distributed

```{r}
wy %>% 
  ggplot(aes(x = amount)) + 
  geom_histogram() +
  scale_x_continuous(
    trans = "log10", labels = dollar)
```

Distribution of expenses by filer
```{r box_plot_by_type, echo=FALSE}

wy %>% 
  ggplot(
    mapping = aes(
      x = filer_type, 
      y = amount
    )
  ) +
  geom_boxplot(
    mapping  = aes(fill = filer_type), 
    varwidth = TRUE,
    outlier.alpha = 0.01
  ) +
  scale_fill_brewer(
    type    = "qual",
    palette = "Set1",
    guide   = FALSE
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar
  ) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
  labs(
    title = "Wyoming Expenditure Amount Ranges",
    x     = "Expenditure Type",
    y     = "Amount",
    caption = "Source: Wyoming Secretary of State"
  )
```

### Dates
The dates seem to be reasonable, with records dating back to `r summary(wy$date)[1]` till `r summary(wy$date)[-1]` 
```{r}
summary(wy$date)
```

### Year

Add a `year` variable from `date` after `col_date()` using `lubridate::year()`.

```{r add_year}
wy <- wy %>% mutate(year = year(date), on_year = is_even(year))
```

```{r year_count_bar, echo=FALSE}

wy %>% 
  count(on_year, year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill=on_year)) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  labs(
    title = "Wyoming Expenditure Counts per Year",
    caption = "Source: Wyoming Secretary of State",
    x = "Year",
    y = "Count"
  )
  
```

```{r amount_year_bar, echo=FALSE}
wy %>% 
  group_by(year, on_year) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_col(aes(fill = on_year)) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  scale_y_continuous(labels = scales::dollar) +
  labs(
    title = "Wyoming Expenditure Mean Amount per Year",
    caption = "Source: Wyoming Secretary of State",
    x = "Year",
    y = "Amount"
  ) 
```

```{r amount_month_line}
wy %>% 
  mutate(month = month(date)) %>% 
  group_by(on_year, month) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(month, mean)) +
  geom_line(aes(color = on_year), size = 2) +
  scale_y_continuous(labels = dollar) +
  scale_x_continuous(labels = month.abb, breaks = 1:12) +
  scale_color_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wyoming Expenditure Amount by Month",
    caption = "Source: Wyoming Secretary of State",
    color = "Election Year",
    x = "Month",
    y = "Amount"
  )
```
## Wrangle
### Indexing
```{r}
wy <- tibble::rowid_to_column(wy, "id")
```

The lengths of city_state_zip column differ, and regular expressions can be used to separate the components.

The original data the city, state, and ZIP all in one column. The following code seperates them. 

### Zipcode
First, we'll extract any numbers whose lengths range from 1 to 5 and normalize them under "zip_clean". 

```{r}
wy <- wy %>% 
  mutate(
    zip_clean = city_state_zip %>% 
      str_extract("\\d{2,5}") %>% 
      normal_zip(na_rep = TRUE))
sample(wy$zip_clean, 10)
```


### State

In this regex, state is considered to consist of two upper-case letters following a space, or two upper-case letters with a trailing space at the end.

```{r separate state}
wy <- wy %>% 
  mutate( state_clean =
            trimws(str_extract(wy$city_state_zip, "\\s([A-Z]{2})\\s|^([A-Z]{2})\\s$")))
count_na(wy$state_clean)
```

```{r normal_state, collapse=TRUE}
wy <- wy %>% mutate(state_clean = normal_state(state_clean))
```

### City
First, we can get a list of incorporated cities and towns in Wyoming. The Wyoming State Archives provided the list in a web table. We use the `rvest` package to scrape the names of Wyoming cities and towns. http://wyoarchives.state.wy.us/index.php/incorporated-cities.

```{r}
wyoming_cities_page <- read_html("http://wyoarchives.state.wy.us/index.php/incorporated-cities")

wy_city <- wyoming_cities_page %>%  html_nodes("tr") %>% 
  html_text()

wy_city <- str_match(wy_city[2:100],"(^\\D{2,})\\r")[,2]
wy_city <- toupper(wy_city[!is.na(wy_city)])
```
```{r valid_city}
valid_city <- unique(c(wy_city,zipcode$city))
```

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Prep
Find the cities before a comma first, if not, find the non-numeric string. 
``` {r two-step regex match}
wy <- wy %>% 
  mutate(
    city_raw = str_match(wy$city_state_zip,"(^\\D{3,}),")[,2]) 

wy <- wy %>% mutate(city_raw=ifelse(is.na(city_raw)==TRUE, 
               str_extract(city_state_zip, "[A-Z]{4,}"), paste(city_raw)))

wy$city_raw <- wy$city_raw %>% 
  str_replace("^ROCK$", "ROCK SPRING") 
```

```{r city_raw summary, collapse= TRUE}
count_na(wy$city_raw)
n_distinct(wy$city_raw)
prop_in(wy$city_raw, valid_city, na.rm = TRUE)
sum(unique(wy$city_raw) %out% valid_city)
```

`r sum(!is.na(wy$city_raw))` cities were found.
```{r prep_city}
wy <- wy %>% mutate(city_prep = normal_city(city_raw))
```

#### Match

```{r match_dist}
wy <- wy %>%
  left_join(
    y = zipcode,
    by = c(
      "zip_clean" = "zip",
      "state_clean" = "state"
    )
  ) %>%
  rename(city_match = city) 

```

#### Swap

To replace city names with expected city names from zipcode when the two variables are no more than two characters different
```{r }
wy <- wy %>% 
  mutate(
    match_dist = stringdist(city_prep, city_match),
    city_swap = if_else(condition = is.na(city_match) == FALSE,
                        if_else(
      condition = match_dist <= 2,
      true = city_match,
      false = city_prep
    ),
      false = city_prep
  ))

wy$city_swap <- wy$city_swap %>% 
  str_replace("^CAS$", "CASPER") %>% 
  str_replace("^CA$", "CASPER") %>% 
  str_replace("^RS$","ROCK SPRINGS") %>% 
  str_replace("^AF$", "AFTON") %>% 
  str_replace("^M$", "MOUNTAIN VIEW") %>% 
  str_replace("^GR$", "GREEN RIVER") %>% 
  na_if("WY") %>% 
  str_replace(" WYOMING","") %>% 
  str_replace("^SLC$", "SALT LAKE CITY") %>% 
  str_replace("^COD$", "CODY") 

  
summary(wy$match_dist)
sum(wy$match_dist == 1, na.rm = TRUE)
n_distinct(wy$city_swap)
```
This ZIP match swapping made `r sum(wy$city_swap != wy$city_prep, na.rm = TRUE)` changes.

#### Refine

Instead of using the OpenRefine algorithms' `refinr::key_collision_merge()` and `refinr::n_gram_merge()` functions, we use `adist` and `agrep` to fuzzy match the swapped city data with valid city names.
```{r view_refine}
wy_cities <- tibble(city = wy_city, state = rep("WY",length(wy_city)))
cities <- unique(rbind(wy_cities, unique(select(zipcode, -zip))))
# Get a table of cities that are not in the valid_city vector
wy_out <- wy %>% 
  count(state_clean, city_swap, sort = TRUE) %>% 
  filter(city_swap %out% valid_city) %>% 
  drop_na()

# Fuzzy-matching city names with the names out of such list
prep_refine <- wy_out %>% 
  # Join on all names in the relevant state
  left_join(cities, by=c(state_clean = "state")) %>%
  # Calculate the distances, per original city name.
  group_by(city) %>%                
  mutate(dist = diag(adist(city_swap, city, partial=TRUE))) %>%
  # Append the agrepl result with the Levenshtein edit distance
  rowwise() %>%
  mutate(string_agrep = agrepl(city_swap, city, max.distance = 0.3)) %>%
  ungroup() %>%  
  # Only retain minimum distances
  group_by(city_swap) %>%   
  filter(dist == min(dist))

# Refine the entries where city_swap is six letter apart from a single match in cities (It's a rather safe switch, after examining the prep_refine table). Overcorrection can be manually changed later.
to_refine <- prep_refine %>% filter(n()==1) %>% filter(city_swap %in% prep_refine$city_swap[prep_refine$dist<6])
```

#### Merge 

```{r join_refine}
wy_refined <- wy %>% left_join(to_refine, by = "city_swap") %>% 
  select(-n, -state_clean.y, -dist, - string_agrep) %>% 
   mutate(city_refine = if_else(
    condition = is.na(city) == TRUE,
    true = city_swap,
    false = city
  )) %>% select(-city)

```

Manually change the city_refine fields due to overcorrection/undercorrection.

```{r}
wy_refined$city_refine <- wy_refined$city_refine %>% 
  str_replace("^RIO VISTA$", "LAGO VISTA") %>% 
  str_replace("^OGEN$", "OGDEN") %>%
  str_replace("^ANNIPOLIS$", "ANNAPOLIS") %>% 
  str_replace("^LAR$", "LARAMIE") %>%
  str_replace("^LARA$", "LARAMIE") %>%
  str_replace("^CHE$", "CHEYENNE") %>%
  str_replace("^COLO SPGS$", "COLORADO SPRINGS") %>%
  str_replace("^WASHNGTON$", "WASHINGTON") %>% 
  str_replace("^WASHINGTON DC$", "WASHINGTON") %>% 
  str_replace("^ST.\\s", "SAINT " ) %>% 
  str_replace("^PINE$", "PINEDALE")
```

This process reduces the number of distinct city value by `r n_distinct(wy_refined$city_raw) - n_distinct(wy_refined$city_refine)`
```{r city_progress1, collapse=TRUE}
n_distinct(wy_refined$city_raw)
n_distinct(wy_refined$city_prep)
n_distinct(wy_refined$city_swap)
n_distinct(wy_refined$city_refine)
```

Each process also increases the percent of valid city names.

```{r city_progress2, collapse=TRUE}
prop_in(wy_refined$city_raw, valid_city, na.rm = TRUE)
prop_in(wy_refined$city_prep, valid_city, na.rm = TRUE)
prop_in(wy_refined$city_swap, valid_city, na.rm = TRUE)
prop_in(wy_refined$city_refine, valid_city, na.rm = TRUE)
```
Each step of the cleaning process reduces the number of distinct city values.
There are `r sum(!is.na(wy_refined$city_raw))` entries of cities identified in the original data matching the regex with `r n_distinct(wy_refined$city_raw)` distinct values, after the swap and refine processes, there are `r sum(!is.na(wy_refined$city_refine))` entries with `r n_distinct(wy_refined$city_refine)` distinct values. 

## Conclude

1. There are `r nrow(wy)` records in the database
1. There are `r sum(wy$dupe_flag)` records with duplicate filer, recipient, date, _and_ amount
(flagged with `dupe_flag`)
1. The ranges for dates and amounts are reasonable
1. Consistency has been improved with `stringr` package and custom `normal_*()` functions.
1. The five-digit `zip_clean` variable has been created with `zipcode::clean.zipcode()`
1. The `year` variable has been created with `lubridate::year()`
1. There are `r sum(is.na(wy$filer_name))` records with missing `name` values and `r sum(is.na(wy$date))`
records with missing `date` values (both flagged with the `na_flag`)

## Export

```{r write_clean}
clean_dir <- here("wy", "expends", "data", "processed")
dir_create(clean_dir)
wy_refined %>% 
  rename(city_clean = city_refine) %>% 
  select(
    -city_prep,
    -on_year,
    -city_match,
    -match_dist,
    -city_swap,
  ) %>% 
  write_csv(
    path = glue("{clean_dir}/wy_expends_clean.csv"),
    na = ""
  )
```

