---
title: "Iowa Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ia", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  jsonlite, # import json api
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  aws.s3, # aws cloud storage
  scales, # format strings
  rvest, # read html pages
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Iowa Ethics and Campaign Disclosure Board][iae].

> In order to accomplish its Mission, the Board will enforce the provisions of
the "Campaign Disclosure Act" in Iowa Code chapter 68A, the "Government Ethics
and Lobbying Act" in Iowa Code chapter 68B, the reporting of gifts and bequests
received by agencies under Iowa Code section 8.7, and the Board's administrative
rules in Chapter 351 of the Iowa Administrative Code.

[iae]: https://ethics.iowa.gov/

The Board provides the file through the [state open data portal][odp] under the
title "Iowa Campaign Contributions Received." The data can be accessed as a
tabular CSV file or through a number of direct APIs.

The database was created June 18, 2015 and last updated December 10, 2019.

[odp]: https://data.iowa.gov/

> This dataset contains information on contributions and in kind donations made
by organizations and individuals to state-wide, legislative or local candidate
committees, state PACs, county central committees, state parties, and state and
local ballot issue committees in Iowa. Data is available beginning in 2003 for
all reports filed electronically, and some paper filed reports.

> Data is provided through reports submitted by candidate committees, state
political committees, federal/out-of-state political committees, county central
committees, ballot issue committees and organizations making contributions or
independent expenditures. Quality of the data provided in the dataset is
dependent upon the accuracy of the data reported electronically.

The Board also provides a disclaimer on the completness of the database:

> Data on paper filed reports is not available except for the following:
contributions to individual candidates between 2003 and 2006 from political and
party committees; contributions to individual candidates 2007 on; contributions
to party committees between 2003 and 2007 from political and candidate
committees; contributions from State Political Committees to candidates between
2003 and 2004; contributions from Federal/Out-of-State Political Committees over
$50 from 2005 on; and contributions from county central committees from 2008 on.

The database license is as follows:

> Pursuant to Iowa Code section 68B.32A(7), the information obtained from
statements or reports filed with the board under Iowa Code chapter 68A, Iowa
Code chapter 68B, Iowa Code section 8.7, or rules adopted by the board shall not
be copied or otherwise used for any commercial purpose. For purposes of this
rule, "commercial purposes" shall include solicitations by a business or
charitable organization. Information used in newspapers, magazines, books, or
other similar communications, so long as the principal purpose of such
communications is for providing information to the public and not for other
commercial purpose, and for soliciting political campaign contributions is
permissable.

## Read

These fixed files can be read into a single data frame with `purrr::map_df()`
and `readr::read_delim()`.

```{r raw_dir}
raw_dir <- dir_create(here("ia", "contribs", "data", "raw"))
raw_url <- "https://data.iowa.gov/api/views/smfg-ds7h/rows.csv"
raw_path <- path(raw_dir, basename(raw_url))
if (!this_file_new(raw_path)) {
  download.file(raw_url, raw_path)
}
```

```{r raw_read}
iac <- vroom(
  file = raw_path,
  na = c("", "N/A", "NA", "n/a", "na"),
  col_types = cols(
    .default = col_character(),
    `Date` = col_date_usa(),
    `Contribution Amount` = col_double()
  )
)
```

```{r raw_rename, echo=FALSE}
iac <- iac %>% 
  clean_names("snake") %>%
  rename(
    code = committee_code,
    committee = committee_name,
    type = transaction_type,
    cont_comm_cd = contributing_committee_code,
    cont_org = contributing_organization,
    first = first_name,
    mi = middle_initial,
    last = last_name,
    addr1 = address_line_1,
    addr2 = address_line_2,
    zip = zip_code,
    amount = contribution_amount,
    tx = transaction_id,
  ) %>% 
  mutate(last = coalesce(last, cont_org)) %>% 
  select(-cont_org, -cont_comm_cd)
```

We can ensure this file was read correctly by counting distinct values of a 
known discrete variable.

```{r raw_check}
n_distinct(iac$type) == 2
```

## Explore

There are `r comma(nrow(iac))` rows of `r ncol(iac)` columns.

```{r glimpse}
glimpse(iac)
tail(iac)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(iac, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("date", "last", "amount", "committee")
iac <- flag_na(iac, all_of(key_vars))
sum(iac$na_flag)
```

All of the flagged rows are only missing a contributor `last` name.

```{r na_view}
iac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_n(10)
```

```{r na_recount}
iac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

We can create a file containing every duplicate record in the data.

```{r dupe_write}
dupe_file <- path(dirname(raw_dir), "dupes.csv")
if (!file_exists(dupe_file)) {
  write_lines("tx,dupe_flag", dupe_file)
  iac <- mutate(iac, group = str_sub(date, end = 7))
  ia_tx <- split(iac$tx, iac$group)
  ias <- iac %>%
    select(-tx) %>% 
    group_split(group, .keep = FALSE)
  pb <- txtProgressBar(max = length(ias), style = 3)
  for (i in seq_along(ias)) {
    write_csv(
      path = dupe_file,
      append = TRUE,
      x = tibble(
        tx = ia_tx[[i]],
        dupe_flag = or(
          duplicated(ias[[i]], fromLast = FALSE),
          duplicated(ias[[i]], fromLast = TRUE)
        )
      )
    )
    setTxtProgressBar(pb, i)
    ias[i] <- NA
    flush_memory(1)
  }
}
```

```{r dupe_read}
dupes <- read_csv(
  file = dupe_file,
  col_types = cols(
    tx = col_character(),
    dupe_flag = col_logical()
  )
)
```

This file can then be joined against the contributions using the transaction ID.

```{r dupe_flag}
iac <- left_join(iac, dupes)
iac <- mutate(iac, dupe_flag = !is.na(dupe_flag))
percent(mean(iac$dupe_flag), 0.1)
```

```{r dupe_view}
iac %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(date, last)
```
  
## Categorical

```{r n_distinct}
col_stats(iac, n_distinct)
```

```{r rec_type_bar, echo=FALSE}
add_prop(count(iac, type))
```

## Amounts

```{r ammount_summary}
summary(iac$amount)
mean(iac$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
iac %>%
  filter(!is.na(amount), amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Iowa Contribution Distribution",
    caption = "Source: Iowa Election Division",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_comm_violin, echo=FALSE}
iac %>%
  filter(!is.na(amount), amount >= 1) %>% 
  ggplot(aes(x = type, y = amount)) +
  geom_violin(aes(fill = type), adjust = 3) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Iowa Contribution Distribution",
    subtitle = "by type, direct or in-kind",
    caption = "Source: Iowa Election Division",
    x = "Amount",
    y = "Count"
  )
```

## Dates

```{r date_year}
iac <- mutate(iac, year = year(date))
```

```{r year_bar}
iac %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1998, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Iowa Contributions by Year",
    caption = "Source: Iowa Election Division",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
addr_norm <- iac %>%
  select(starts_with("addr")) %>% 
  distinct() %>% 
  unite(
    everything(),
    col = addr_full,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address_join}
iac <- left_join(iac, addr_norm)
rm(addr_norm)
```

We can see how this process improved consistency.

```{r address_view}
iac %>% 
  sample_n(10) %>% 
  select(starts_with("addr"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
iac <- iac %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  iac$zip, 
  iac$zip_norm, 
  compare = valid_zip
)
```

### State

Very little needs to be done to clean the `state` variable.

```{r state_repair}
x <- iac$state
length(x)
prop_in(x, valid_state)
count_out(x, valid_state)
st_zip <- iac$zip %in% zipcodes$zip[zipcodes$state == "IA"]
st_out <- x %out% valid_state
st_rx <- str_detect(x, "^[Ii]|[Aa]$")
st_na <- !is.na(x)
# has ia zip, ia regex, not valid, not na
x[st_zip & st_rx & st_out & st_na] <- "IA"
length(x)
iac <- mutate(iac, state_norm = x)
```

```{r state_progress}
progress_table(
  iac$state, 
  iac$state_norm, 
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats. The `campfin::normal_city()` 
function is a good start, again converting case, removing punctuation, but
_expanding_ USPS abbreviations. We can also remove `invalid_city` values.

```{r city_norm}
ia_city <- iac %>% 
  count(city, state_norm, zip_norm, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("IA", "DC", "IOWA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

Again, we can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ia_city <- ia_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(city_match) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_other}
many_city <- c(valid_city, extra_city)
ia_city %>% 
  count(city_swap, state_norm, sort = TRUE) %>% 
  filter(!is.na(city_swap), city_swap %out% many_city)
```

```{r city_replace}
ia_city <- ia_city %>% 
  mutate(
    city_swap = city_swap %>% 
      str_replace("^OVERLAND PARKS$", "OVERLAND PARK") %>% 
      str_replace("^NEW YORK CITY$", "NEW YORK") %>% 
      str_replace("^NYC$", "NEW YORK") %>% 
      str_replace("^WASHINGTON D C$", "WASHINGTON") %>% 
      str_replace("\\sPK$", "PARK") %>% 
      str_remove("\\sD\\sC$") %>% 
      str_remove("\\sIN$") %>% 
      na_if("UNITEMIZED") %>% 
      na_if("IA")
  )
```

```{r city_join}
ia_city <- rename(ia_city, city = city_raw)
iac <- left_join(iac, ia_city, by = c("city", "state_norm", "zip_norm"))
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(iac$city),
  iac$city_norm,
  iac$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Iowa City Normalization Progress",
    subtitle = "Raw at 21.5% before converted to uppercase",
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Iowa City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
iac <- iac %>% 
  select(
    -city_norm,
    city_clean = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(iac, 50))
```

1. There are `r comma(nrow(iac))` records in the database.
1. There are `r comma(sum(iac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(iac$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ia", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ia_contribs_clean.csv")
write_csv(iac, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=TRUE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(iac)),
  type = md_code(map_chr(iac, typeof)),
  def = c(
    "Unique transaction hash",
    "Date contribution was made",
    "Recipient committee code",
    "Recipient committee name",
    "Type of contribution (direct, in-kind)",
    "Contributor first name",
    "Contributor middle initial",
    "Contributor last name or organization",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor state abbreviation",
    "Contributor city name",
    "Contributor ZIP+4 code",
    "Amount or correction",
    "Flag for missing value",
    "Flag for duplicate row",
    "Calendar year contribution made",
    "Normalized street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-letter state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
