---
title: "Florida Expenditures"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("fl", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("fl/expends/docs/fl_expends_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Florida Division of Elections.

As the [agency home page][fdoe] explains:

> By Florida law, campaigns, committees, and electioneering communications
organizations are required to disclose detailed financial records of campaign
contributions and expenditures. Chapter 106, Florida Statutes, regulates
campaign financing for all candidates, including judicial candidates, political
committees, electioneering communication organizations, affiliated party
committees, and political parties. It does not regulate campaign financing for
candidates for federal office.

[fdoe]: https://dos.myflorida.com/elections/candidates-committees/campaign-finance/

### About

A more detailed description of available data can be found on the 
[Campaign Finance page][cf]:

[cf]: https://dos.myflorida.com/elections/candidates-committees/campaign-finance/campaign-finance-database/

> #### Quality of Data
> The information presented in the campaign finance database is an accurate
representation of the reports filed with the Florida Division of Elections. > >
Some of the information in the campaign finance database was submitted in
electronic form, and some of the information was key-entered from paper reports.
Sometimes items which are not consistent with filing requirements, such as
incorrect codes or incorrectly formatted or blank items, are present in the
results of a query. They are incorrect in the database because they were
incorrect on reports submitted to the division.

> #### What does the Database Contain?
> By law candidates and committees are required to disclose detailed financial
records of contributions received and expenditures made. For committees, the
campaign finance database contains all contributions and expenditures reported
to the Florida Division of Elections since January 1, 1996. For candidates, the
campaign finance database contains all contributions and expenditures reported
to the Division since the candidacy was announced, beginning with the 1996
election.

> #### Whose Records are Included?
> Included are campaign finance reports which have been filed by candidates for
any multi-county office, with the exception of U.S. Senator and U.S.
Representative, and by organizations that receive contributions or make
expenditures of more than $500 in a calendar year to support or oppose any
multi-county candidate, issue, or party. To obtain reports from local county or
municipal candidates and committees, contact county or city filing offices.

> #### When are the Records Available?
> Campaign finance reports are posted to the database as they are received from
the candidates and committees. Our data is as current as possible, consistent
with the reporting requirements of Florida law.

## Download

We will use the [Expenditure Records][exp] query form to download three separate
files covering all campaign expenditures. [The home page][cf] lists
instructions on how to download the desired files:

> #### How to Use the Campaign Finance Database
> 1. Specify a subset of the [Expenditure]...
> 2. Select an election year entry from the list box.
> 3. Select a candidate/committee option...
> 4. Select contribution criteria (for Detail report only)...
> 5. Select how you would like the records sorted. 
> 6. Select the format in which you would like the data returned. 
> 7. Limit the number of records to return....
>     * Choosing "Return Query Results in a Tab Delimited Text File" will return
        the data in a file of tab-separated columns suitable for importing into 
        almost any spreadsheet or database. This option allows you to download 
        the data for further analysis offline.
> 8. Click on the Submit Query button.

To get all files covering all expenditures:

1. Select "All" from the **Election Year** drop down menu
1. In the **From Date Range** text box, enter "01/01/2008"
1. Delete "500" from the **Limit Records** text box
1. Select the "Return Results in a Tab Delimited Text File" **Retrieval Format**
1. Save to the `/fl/expends/data/raw` directory

[exp]: https://dos.elections.myflorida.com/campaign-finance/expenditures/

```{r raw_tsv}
raw_dir <- dir_create(here("fl", "expends", "data", "raw"))
raw_tsv <- path(raw_dir, "Expend.txt")
```   

```{r raw_post}
if (!file_exists(raw_tsv) && FALSE) {
  POST(
    write_disk(raw_tsv),
    user_agent("https://publicaccountability.org/"),
    url = "https://dos.elections.myflorida.com/cgi-bin/expend.exe",
    # must get cookies from browser :(
    set_cookies(),
    encode = "form",
    body = list(
      election = "All",
      search_on = "1",
      CanFName = "",
      CanLName = "",
      CanNameSrch = "2",
      office = "All",
      cdistrict = "",
      cgroup = "",
      party = "All",
      ComName = "",
      ComNameSrch = "2",
      committee = "All",
      cfname = "",
      clname = "",
      namesearch = "2",
      ccity = "",
      cstate = "",
      czipcode = "",
      cpurpose = "",
      cdollar_minimum = "",
      cdollar_maximum = "",
      rowlimit = "",
      csort1 = "DAT",
      csort2 = "CAN",
      cdatefrom = "01/01/1900",
      cdateto = "",
      queryformat = "2",
      Submit = "Submit"
    )
  )
}
```

## Fix

The tab-separated file has no quotes surrounding problematic fields. That
results in a handful with new-line character breaks (e.g., typed in an 
address). To fix this problem, we have to read each line separately and then
manually combine any of those split lines.

```{r fix_read}
fll <- read_lines(raw_tsv)
length(fll)
fl_head <- fll[1] # save header
fll <- fll[-1] # remove header
```

We can identify the problematic lines by looking for any line without a `Date`
character pattern following the first tab-delimiter.

```{r fix_which}
bad_nl <- str_which(
  string = fll,
  # find rows without date in 2nd col
  pattern = "^(.*)\t(\\d{2}/\\d{2}/\\d{4})",
  negate = TRUE
)
```

```{r fix_show}
length(bad_nl)
fll[bad_nl]
```

We can then combine these identified lines with the previous, making a new valid
line. When the broken line is removed, the file can be read.

```{r fix_merge}
fl_combine <- paste0(fll[bad_nl - 1], fll[bad_nl])
read_tsv(file = c(fl_head, fl_combine)) # test lines
fll[bad_nl - 1] <- fl_combine # replace broken lines
fll <- fll[-bad_nl] # remove extra already merged
```

```{r}
fll <- iconv(fll, to = "UTF-8", sub = "")
```

```{r fix_write}
tmp_tsv <- file_temp(ext = "tsv")
write_lines(x = c(fl_head, fll), file = tmp_tsv)
```

```{r fix_rm, echo=FALSE}
rm(fll, fl_head, fl_combine, bad_nl)
gc(reset = TRUE, full = TRUE)
```

## Read

```{r raw_read}
fle <- read_delim(
  file = tmp_tsv,
  delim = "\t",
  quote = "",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    Date = col_date_mdy(),
    Amount = col_double()
  )
)
```

```{r raw_names}
fle <- fle %>% 
  rename(`Comm Name` = `Candidate/Committee`) %>% 
  clean_names(case = "snake")
```

## Explore

There are `r comma(nrow(fle))` rows of `r ncol(fle)` columns. Each record
represents a single expenditure made by a political campaign to an outside
vendor for goods or services.

```{r glimpse}
glimpse(fle)
tail(fle)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(fle, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("date", "payee_name", "amount", "comm_name")
fle <- flag_na(fle, all_of(key_vars))
sum(fle$na_flag)
```

```{r na_view}
fle %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
fle <- flag_dupes(fle, everything())
mean(fle$dupe_flag)
sum(fle$dupe_flag)
```

```{r dupe_view}
fle %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(payee_name, amount)
```

### Categorical

```{r distinct_count}
col_stats(fle, n_distinct)
```

```{r type_upper}
fle <- mutate(fle, across(type, str_to_upper))
```

```{r distinct_plots, echo=FALSE}
explore_plot(fle, type)
```

### Amounts

```{r amount_summary}
summary(fle$amount)
mean(fle$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(fle[c(which.max(fle$amount), which.min(fle$amount)), ])
```

```{r hist_amount, echo=FALSE}
fle %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Florida Expenditures Amount Distribution",
    caption = "Source: FL DOE",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
fle <- mutate(fle, year = year(date))
```

```{r date_range}
min(fle$date)
sum(fle$year < 2000)
max(fle$date)
sum(fle$date > today())
```

```{r bar_year, echo=FALSE}
fle %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1995, 2021, by = 2)) +
  coord_cartesian(xlim = c(1995, 2021)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida Expenditures by Year",
    caption = "Source: FL DOE",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Separate

```{r}
fle <- extract(
  data = fle,
  col = city_state_zip,
  into = c("city", "state", "zip"),
  regex = "^(.*),\\s+(.*)\\s+(.*)$",
  remove = FALSE
)
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

The `address` also seems to be separated by white space not tab-characters. We
can create a new variable with the secondary address.

```{r}
fle %>% 
  select(address) %>% 
  head(10)
```

```{r address_split}
fl_addr <- fle %>% 
  distinct(address) %>% 
  separate(
    col = address,
    into = c("address1", "address2"),
    sep = "\\s{3,}",
    remove = FALSE,
    extra = "merge",
    fill = "right"
  )
```

```{r address_norm}
fl_addr <- mutate(
  fl_addr,
  across(
    .cols = c(address1, address2),
    .fns = ~normal_address(
      address = .,
      abbs = usps_street,
      na_rep = TRUE,
      na = invalid_city
    )
  )
)
```

```{r address_join}
fle <- left_join(fle, fl_addr, by = "address")
fle <- fle %>% 
  relocate(address1, address2, .after = address) %>% 
  select(-address)
```

```{r address_view}
fle %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

```{r echo=FALSE}
rm(fl_addr)
flush_memory()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
fle <- fle %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  fle$zip,
  fle$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
fle <- fle %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
fle %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  fle$state,
  fle$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- fle %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("FL", "DC", "FLORIDA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
fle <- left_join(
  x = fle,
  y = norm_city,
  by = c(
    "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- fle %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
fle <- fle %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(fle$city_raw),
  fle$city_norm,
  fle$city_swap,
  fle$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(fle$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
fle <- fle %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(fle, 50))
```

1. There are `r comma(nrow(fle))` records in the database.
1. There are `r comma(sum(fle$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(fle$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("fl", "expends", "data", "clean"))
clean_path <- path(clean_dir, "fl_expends_clean.csv")
write_csv(fle, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(fle)),
  type = md_code(map_chr(fle, typeof)),
  def = c(
    ""
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
