---
title: "Louisiana Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("la", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data can be obtained from the [Louisiana Checkbook website][chk]. We
can download "Annual Report Source Data" in Microsoft Excel format. This data
was last updated June 1, 2020.

[chk]: https://checkbook.la.gov/contracts/index.cfm

## Download

Data is available from 2016 through 2019.

```{r raw_download}
raw_dir <- dir_create(here("la", "contracts", "data", "raw"))
raw_source <- "https://checkbook.la.gov/Reports/AnnualSource/"
raw_names <- glue("FY{16:19}AnnualReportSourceDataAct589.xlsx")
raw_urls <- str_c(raw_source, raw_names)
raw_paths <- path(raw_dir, raw_names)
if (!all(file_exists(raw_paths))) {
  download.file(raw_names, raw_paths)
}
```

## Read

The four Excel spreadsheets can be read into a single data frame using
`purrr::map_df()` and `readxl::read_excel()`.

```{r raw_read}
lac <- map_df(
  .x = raw_paths,
  .f = read_excel,
  .id = "source",
  skip = 3,
  col_types = "text"
)
```

After reading every column as text, we can clean the variable names and parse
the numeric columns accordingly. We can also change the `source` variable from
combining files into the corresponding fiscal year.

```{r raw_parse}
lac <- lac %>% 
  clean_names("snake") %>% 
  filter(vendor_name != "Sum:") %>% 
  mutate(across(total_amount, parse_double)) %>% 
  mutate(across(total_count, parse_integer)) %>% 
  mutate(
    source = basename(raw_paths[as.integer(source)]),
    year = source %>% 
      str_extract("(?<=FY)\\d{2}") %>% 
      str_c("20", .) %>% 
      as.integer()
  ) %>% 
  mutate(
    description = coalesce(
      description, published_description, published_text
    )
  ) %>% 
  select(-published_description, -published_text)
```

Records belonging to contracts have a `contract_no` and single purchase orders
have a `po_number` variable and the inverse is missing for each. We will create
a new variable indicating the record _type_, combine these two numbers as a new
single `id` variable, and then remove the two original number columns.

```{r raw_ids}
lac <- lac %>% 
  mutate(
    .before = contract_no,
    id = coalesce(po_number, contract_no)
  ) %>% 
  mutate(
    .after = id,
    .keep = "unused",
    type = case_when(
      is.na(contract_no) ~ "purchase",
      is.na(po_number) ~ "contract"
    )
  )
```

## Explore

```{r glimpse}
glimpse(lac)
tail(lac)
```

### Missing

The only variables missing any values are the `description` variable and
similar free-form text cells at the end of each spreadsheets. They do not need
to be flagged.

```{r na_count}
col_stats(lac, count_na)
```

### Duplicates

If we ignore the supposedly unique `id` variable, there are a few records that
are entirely duplicated at least once across every variable. Without an exact
data column, these could very well be contracts or purchase orders made for the
same amount in the same year.

```{r dupe_flag}
lac <- flag_dupes(lac, -id)
sum(lac$dupe_flag)
```

```{r dupe_view}
lac %>% 
  filter(dupe_flag) %>% 
  select(id, year, vendor_name, department_name, total_amount)
```

### Categorical

```{r distinct_count}
col_stats(lac, n_distinct)
```

```{r distinct_plots}
explore_plot(lac, type)
explore_plot(lac, service_type_name)
explore_plot(lac, department_name) + scale_x_truncate()
```

### Amounts

A significant portion of the `total_amount` variable are less than or equal to
zero.

```{r ammount_summary}
summary(lac$total_amount)
percent(mean(lac$total_amount < 0), 0.1)
percent(mean(lac$total_amount == 0), 0.1)
```

```{r hist_amount, echo=FALSE}
lac %>%
  ggplot(aes(total_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Louisiana Contracts Amount Distribution",
    caption = "Source: LA Checkbook",
    x = "Amount",
    y = "Count"
  )
```

It's not clear what the `total_count` variable represents. Perhaps the total
number of each contract ordered.

```{r amount_count_mean}
lac %>% 
  group_by(total_count) %>% 
  summarise(mean_amount = mean(total_amount))
```

### Dates

We already added the fiscal year based on the source file. Without a start date,
we can't add the calendar year.

```{r bar_year, echo=FALSE}
lac %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Louisiana Contracts by Year",
    caption = "Source: LA Checkbook",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

There are no geographic variables that require normalization. At the very least,
we know that each `department_name` value has an associated state in Louisiana.

```{r state_add}
lac <- mutate(lac, state = "LA", .after = department_name)
```

## Conclude

1. There are `r comma(nrow(lac))` records in the database.
1. There are `r comma(sum(lac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(lac$na_flag))` records missing key variables.
1. There are no geographic variables to normalize, `state` was added manually.
1. The 4-digit fiscal `year` was determined from source file.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("la", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "la_contracts_clean.csv")
write_csv(lac, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(lac)),
  type = md_code(map_chr(lac, typeof)),
  def = c(
    "Source Excel file name",
    "Unique contract or purchase ID",
    "Contract or single purchase order",
    "Service type code",
    "Service type full name",
    "Spending department code",
    "Spending department name",
    "Spending department state abbreviation",
    "Common vendor ID",
    "Unique vendor ID",
    "Vendor full name",
    "Total contract or purchase amount",
    "Total number of contracts ordered",
    "Free-form text description of contracts",
    "Fiscal year ordered from source file",
    "Flag indicating duplicate record"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Louisiana Contracts Data Dictionary\n", dict_md),
  path = here("la", "contracts", "la_contracts_dict.md"),
)
```
