---
title: "Kentucky Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("df", "data", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  textreadr, # read text files
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

The data is obtained from the [Kentucky Legislative Ethics Commission (KLEC)][03]:

[03]: https://klec.ky.gov/Pages/default.aspx

> KLEC has jurisdiction over
> 
> * Legislative agents (lobbyists),
> * Employers (individuals or entities who engage legislative agents), and
> * Members of the General Assembly.
> 
> The legislative ethics law covers four broad subject matters
> 
> * Registration of legislative agents and employers;
> * Statements by legislative agents and employers of:
>     * lobbying expenditures and expenses, and financial transactions;
> * Conduct of members of the General Assembly; and
> * Financial disclosure statements of the General Assembly, legislative candidates, and
> * key legislative staff.

The KLEC provides [a rich text file (RTF)][04] containing a list of legislative agents (lobbyists).

[04]: https://klec.ky.gov/Reports/Reports/Agents.rtf

## Import

The text file is a table of lobbyists with their clients indented below them.  With the `textreadr`
package, we can read the RTF file into a character vector of the lines on the document.

```{r raw_read}
lob_url <- "https://klec.ky.gov/Reports/Reports/Agents.rtf"
kylr <- read_rtf(file = lob_url)
```

```{r raw_head, echo=FALSE}
head(kylr)
```

First, we need to remove the header and footer from each page and keep only those lines which
contain the lobbyist name table information.

```{r raw_trim}
kylr <- str_subset(kylr, "Kentucky Registered Legislative Agents", negate = TRUE)
kylr <- str_subset(kylr, "November 22, 2019", negate = TRUE)
kylr <- str_subset(kylr, "Legislative Agents/Employer\tPhone\tContact\tAddress", negate = TRUE)
kylr <- str_subset(kylr, "11/22/2019\t\\d{2}", negate = TRUE)
kylr <- str_replace_all(kylr, "\"", "\'")
```

## Wrangle

We need to `tibble::enframe()` the character vector and turn it into a single column data frame.
From there, we can `tidyr::separate()` the column into it's four component elements.

```{r raw_sep}
kylr <- kylr %>%
  enframe(name = "line", value = "text") %>%
  separate(
    col = text,
    into = c("name", "phone", "contact", "address"),
    sep = "\t"
  )
```

```{r echo=FALSE}
head(kylr)
```

Then, we have to use the indentation of the text file to identify which rows belong to lobbyist
information and which belong to their principal clients.

```{r raw_indent}
indent <- which(is.na(kylr$address))
kylr <- mutate(kylr, address = coalesce(address, contact))
kylr$contact[indent] <- NA
```

Using this identation, we can shift the lobbyist names over into a _new_ column and `dplyr::fill()`
that name and address down _alongside_ each of their clients below. Then this new data frame
is re-arranged into a table with each record identifying a lobbyist and a single client. In this
sense, the lobbyist names are now repeated for each client.

```{r raw_shift}
kylr <- kylr %>%
  mutate(
    lob_name = if_else(
      condition = is.na(contact),
      true = name,
      false = NA_character_
    ),
    lob_phone = if_else(
      condition = is.na(contact),
      true = phone,
      false = NA_character_
    ),
    lob_address = if_else(
      condition = is.na(contact),
      true = address,
      false = NA_character_
    )
  ) %>%
  fill(starts_with("lob")) %>%
  mutate_if(is_character, str_trim) %>%
  mutate_all(str_to_upper) %>% 
  filter(!is.na(contact)) %>%
  rename(
    pri_name = name,
    pri_phone = phone,
    pri_contact = contact,
    pri_address = address
  ) %>%
  select(
    starts_with("lob"),
    starts_with("pri")
  )
```

Now, we need to `tidyr::separate()` the two new `*_address` columns into the other components.
First, we will split the lobbyist's address into the street, city, state, and ZIP code. We will

```{r raw_sep_lob}
kylr <- kylr %>%
  separate(
    col = lob_address,
    into = c(glue("lob_addr{1:10}"), "lob_extra"),
    sep = ",\\s",
    fill = "left",
    extra = "merge"
  ) %>%
  na_if("") %>%
  unite(
    starts_with("lob_addr"),
    col = "lob_addr",
    na.rm = TRUE,
    sep = " ",
  ) %>% 
  separate(
    col = lob_extra,
    into = c("lob_extra", "lob_zip"),
    sep = "\\s(?=\\d)",
    fill = "left",
    extra = "merge"
  ) %>% 
  separate(
    col = lob_extra,
    into = c("lob_city", "lob_state"),
    sep = "\\s(?=[^ ]*$)",
    fill = "left",
    extra = "merge"
  )
```

```{r raw_sep_pri}
kylr <- kylr %>%
  separate(
    col = pri_address,
    into = c(glue("pri_addr{1:10}"), "pri_extra"),
    sep = ",\\s",
    fill = "left",
    extra = "merge"
  ) %>%
  na_if("") %>%
  unite(
    starts_with("pri_addr"),
    col = "pri_addr",
    na.rm = TRUE,
    sep = " ",
  ) %>% 
  separate(
    col = pri_extra,
    into = c("pri_extra", "pri_zip"),
    sep = "\\s(?=\\d)",
    fill = "left",
    extra = "merge"
  ) %>% 
  separate(
    col = pri_extra,
    into = c("pri_city", "pri_state"),
    sep = "\\s(?=[^ ]*$)",
    fill = "left",
    extra = "merge"
  )
```

```{r raw_sep_name}
kylr <- kylr %>% 
  separate(
    col = lob_name,
    into = c("lob_last", "lob_first"),
    sep = ",\\s",
    extra = "merge" 
  )
```

```{r echo=FALSE}
head(kylr)
```

## Normal

Now that the text file has been wrangled into a database format, we can proceed to manipulate the
_content_ of the file to improve the searchability of the database.

### Phone

```{r phone_norm}
kylr <- mutate_at(
  .tbl  = kylr,
  .vars = vars(ends_with("phone")),
  .funs = list(norm = normal_phone)
)
```

```{r phone_view, echo=FALSE}
kylr %>% 
  select(starts_with("lob_phone")) %>%
  distinct() %>% 
  sample_frac()
```

### Address

```{r addr_norm}
kylr <- mutate_at(
  .tbl  = kylr,
  .vars = vars(ends_with("addr")),
  .funs = list(norm = normal_address),
  abbs  = usps_street
)
```

```{r addr_view, echo=FALSE}
kylr %>% 
  select(starts_with("lob_addr")) %>%
  distinct() %>% 
  sample_frac()
```

### ZIP

```{r zip_norm}
kylr <- mutate_at(
  .tbl  = kylr,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip)
)
```

```{r zip_progress, echo=FALSE}
progress_table(
  kylr$lob_zip,
  kylr$lob_zip_norm,
  kylr$pri_zip,
  kylr$pri_zip_norm,
  compare = valid_zip
)
```

### City

```{r city_norm}
kylr <- mutate_at(
  .tbl  = kylr,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs  = usps_city
)
```

```{r city_swap_lob}
kylr <- kylr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lob_state" = "state",
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lob_city_norm, city_match),
    match_dist = str_dist(lob_city_norm, city_match),
    lob_city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap_pri}
kylr <- kylr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "pri_state" = "state",
      "pri_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(pri_city_norm, city_match),
    match_dist = str_dist(pri_city_norm, city_match),
    pri_city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = pri_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_progress, echo=FALSE}
progress_table(
  kylr$lob_city,
  kylr$lob_city_norm,
  kylr$lob_city_swap,
  compare = c(valid_city, extra_city)
)
progress_table(
  kylr$pri_city,
  kylr$pri_city_norm,
  kylr$pri_city_swap,
  compare = c(valid_city, extra_city)
)
```

## Export

```{r proc_glimpse}
glimpse(kylr)
```

```{r proc_dir}
proc_dir <- here("ky", "lobbying", "data", "processed")
dir_create(proc_dir)
```

```{r proc_write}
kylr %>% 
  select(-ends_with("city_norm")) %>% 
  write_csv(
    path = glue("{proc_dir}/ky_lobby_reg.csv"),
    na = ""
  )
```
