---
title: "California Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("kiernann/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http query
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the California Secretary of State's [Cal-Access database][03].

> Cal-Access provides financial information supplied by state candidates, donors, lobbyists, and
others. Simply start by selecting either Campaign Finance Activity, Lobbying Activity, or
Cal-Access Resources and we will guide you to the information.

The [Political Reform Division][04] (PRD) collects the data regarding expenditures made.

> The Political Reform Division administers provisions of California's Political Reform Act, including the law's most fundamental purpose of ensuring that "receipts and expenditures in election campaigns should be fully and truthfully disclosed in order that the voters may be fully informed and the activities of lobbyists should be regulated and their finances disclosed..."

> * Receive campaign disclosure statements (itemizing contributions received and expenditures made)
filed by individuals and committees raising or spending campaign funds to support or oppose state
candidates or ballot measures. (Local campaign committees file their itemized disclosure statements
with local filing officers).

### About

All California campaign finance data can be downloaded in a single file. As described on PRD
website:

> ### Raw Data for Campaign Finance and Lobbying Activity  
> In addition to presenting California campaign finance and lobbying activity on the user-friendly
[CAL-ACCESS website][05], the Secretary of State provides the raw data to allow people with 
technical expertise to create their own databases.
> 
> The raw data is presented in tab-delimited text files from corresponding tables in the CAL-ACCESS
database. Users can uncompress and extract the data with standard software such as PKZIP, WinZip,
or MacZip. The Secretary of State does not provide technical support to users who want to download
and analyze CAL-ACCESS data in this advanced way. However, the Secretary of State offers guides to
the CAL-ACCESS data structure and fields.

> ### Download Files  
* [Guides to CAL-ACCESS data structure and fields (ZIP)][06]
* [CAL-ACCESS raw data (ZIP)][07]
> 
> All CAL-ACCESS users should keep in mind that campaign finance and lobbying activity information
changes often. The raw data extracts are updated once a day. Campaign finance and lobbying activity
filings can also be obtained in hard copy by contacting the Secretary of State's Political Reform
Division.

[03]: https://www.sos.ca.gov/
[04]: https://www.sos.ca.gov/campaign-lobbying/about-political-reform-division/
[05]: http://cal-access.sos.ca.gov/
[06]: https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip
[07]: https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip

We will download the file to the `/raw` directory.

```{r create_raw_dir}
raw_dir <- here("ca", "expends", "data", "raw")
dir_create(raw_dir)
```

### Variables

Using the data key provided by Cal-Access, we can find the expenditure file and understand it's 
contents.

```{r check_key}
key_url <- "https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip"
key_file <- str_c(raw_dir, basename(key_url), sep = "/")
url_file_size(key_url, format = TRUE)
```

If they ZIP file containing the documentation files has not yet been downloaded, we can do so now.

```{r download_key}
if (!this_file_new(key_file)) {
  download.file(
    url = key_url,
    destfile = key_file
  )
}
```

Before we unzip the file, we can view it's contents.

```{r list_key}
key_content <- as_tibble(
  .name_repair = make_clean_names,
  x = unzip(
    zipfile = key_file,
    list = TRUE
  )
)
```


```{r unzip_key}
key_exists <- dir_exists(glue("{raw_dir}/CalAccess-Documentation"))
if (!key_exists) {
  unzip(
    zipfile = key_file,
    exdir = raw_dir
  )
}
```

## Import

From the documentation, we know the `EXPN` table is the one containing the expenditures we are
interested in.

### Download

If the CAL-ACCESS raw data hasn't yet been downloaded, we can do so.

```{r check_zip}
zip_url <- "https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip"
zip_file <- str_c(raw_dir, basename(zip_url), sep = "/")
url_file_size(zip_url, format = TRUE)
```

```{r download_zip}
if (!this_file_new(zip_file)) {
  download.file(
    url = zip_url,
    destfile = zip_file
  )
}
```

We can use `unzip(list = TRUE)` to view the contents of the ZIP file.

```{r zip_contents}
zip_contents <- as_tibble(
  .name_repair = make_clean_names,
  x = unzip(
    zipfile = zip_file,
    list = TRUE
  )
)
```

```{r zip_contents_expends}
expn_file <- zip_contents %>% 
  filter(name %>% str_detect("EXPN")) %>% 
  pull(name)
```

```{r unzip_zip}
zip_exists <- dir_exists(glue("{raw_dir}/CalAccess"))
if (!zip_exists) {
  unzip(
    zipfile = zip_file,
    exdir = raw_dir,
    files = expn_file
  )
}
```

## Join

### Read

```{r read_expn}
expn_file <- str_c(raw_dir, expn_file, sep = "/")
ca <- read_tsv(
  file = expn_file,
  col_types = cols(
    .default = col_character(),
    EXPN_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    AMOUNT = col_double(),
    CUM_YTD = col_double(),
    CUM_OTH = col_double()
  )
)
```

## Explore

```{r glimpse}
head(ca)
tail(ca)
glimpse(sample_frac(ca))
```

