---
title: "South Carolina Expenditures"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document:
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options:
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("sc", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  jsonlite, # read json data
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("sc/expends/docs/sc_expends_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

South Carolina expenditure data can be obtained from the
[State Ethics Commission][sec], which operates a [search portal][sp].

[sec]: https://ethics.sc.gov/
[sp]: https://ethicsfiling.sc.gov/public/campaign-reports/Expenditures

## Download

We can use the **Advance Search** functions of the portal to request all
Expenditures made between two dates. We will request all Expenditures since
the year 2000 and save the results to a local JSON file.

```{r raw-dir}
raw_dir <- dir_create(here("sc", "expends", "data", "raw"))
raw_json <- path(raw_dir, "Expenditure-Search-Results.xlsx")
```

```{r raw-download}
if (!file_exists(raw_json)) {
  a <- POST(
    url = "https://ethicsfiling.sc.gov/api/Candidate/Expenditure/Public/Get/All/Advanced/Campaign/Expenditures",
    encode = "json",
    write_disk(path = raw_json),
    progress(type = "down"),
    body = list(
      candidate = "",
      expDesc = "",
      fromamount = 0,
      fromDate = "2000-01-01T05:00:00.000Z",
      toDate = Sys.Date(), # thru today
      office = "",
      toamount = 0,
      vendorCity = "",
      vendorName = "",
      vendorState = "",
      vendorZip = ""
    )
  )
}
```

## Read

The JSON file can be read as a flat table with the `fromJSON()` function.

```{r raw-read}
sce <- as_tibble(fromJSON(raw_json))
sce <- clean_names(sce, case = "snake")
```

The columns must be parsed after the fact.

```{r parse-read}
sce <- sce %>%
  mutate(
    across(ends_with("date"), as_date),
    across(where(is_character), str_trim),
    across(where(is_character), na_if, "")
  )
```

## Explore

There are `r comma(nrow(sce))` rows of `r ncol(sce)` columns. Each record
represents a single contribution made from an individual to a campaign.

```{r glimpse}
glimpse(sce)
tail(sce)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(sce, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("exp_date", "vendor_name", "amount", "candidate_name")
```

Only the `exp_desc` column is missing data.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
sce <- flag_dupes(sce, -exp_id)
sum(sce$dupe_flag)
mean(sce$dupe_flag)
```

```{r dupe-view}
sce %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(exp_date)
```

### Categorical

```{r distinct-count}
col_stats(sce, n_distinct)
```

### Amounts

```{r amount-round}
# fix floating point precision
sce$amount <- round(sce$amount, digits = 2)
```

```{r amount-summary}
summary(sce$amount)
mean(sce$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(sce[c(which.max(sce$amount), which.min(sce$amount)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
sce %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "South Carolina Expenditures Amount Distribution",
    caption = "Source: SC SEC",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
sce <- mutate(sce, exp_year = year(exp_date))
```

```{r date-range}
min(sce$exp_date)
sum(sce$exp_year < 2000)
max(sce$exp_date)
sum(sce$exp_date > today())
```

It's common to see an increase in the number of expenditures in elections years.

```{r bar-year, echo=FALSE}
sce %>%
  count(exp_year) %>%
  mutate(even = is_even(exp_year)) %>%
  ggplot(aes(x = exp_year, y = n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "South Carolina Expenditures by Year",
    caption = "Source: SC SEC",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to
facilitate this process.

### Separate

The addresses in this data are contained in a single string with a comma
separating the city name from the state abbreviation and then a space before
the ZIP code. However, there is no way to easily distinguish the street address
from the city name (given the variety in city name length, etc).

This issue does not matter too much given the way the Accountability Project
combined street addresses and city names in our variable mapping. However, we
can try and go row by row and check in fact try to match _every_ known city in 
each state to the address and see if they match. If we find a match, we will 
remove the city name from the street address and place it in its own column.

```{r address-split}
sce$address <- str_replace_all(sce$address, "\n", " ")
sce <- extract(
  data = sce,
  col = address,
  into = c("address_sep", "state_sep", "zip_sep"),
  regex = "^(.*),(.*) (.*)$",
  remove = FALSE
)
```

```{r addr-csv}
addr_csv <- here("sc", "expends", "data", "addr.csv")
```

```{r echo=FALSE}
zipcodes <- zipcodes %>% 
  add_row(city = "MT PLEASANT", state = "NC")
```

```{r addr-look}
if (!file_exists(addr_csv)) {
  x <- sce %>% 
    distinct(address, address_sep, state_sep) %>% 
    mutate(
      address_sep = address_sep %>% 
        str_remove_all("[:punct:]") %>% 
        str_to_upper() %>% 
        str_squish(),
      state_sep = normal_state(state_sep, na_rep = TRUE),
      city_sep = NA_character_
    )
  try_city <- rep(list(NA), length(valid_abb))
  names(try_city) <- valid_abb
  for (st in valid_abb) {
    message("\n", st)
    st_city <- zipcodes$city[zipcodes$state == st]
    y <- filter(x, state_sep == st)
    if (nrow(y) > 0) {
      pb <- txtProgressBar(max = nrow(y), style = 3)
      for (i in seq(nrow(y))) {
        find <- NA
        j <- 1
        while(is.na(find) & j < length(st_city)) {
          find <- str_extract(y$address_sep[i], paste0(st_city[j], "$"))
          j <- j + 1
        }
        if (!is.na(find)) {
          y$city_sep[i] <- find
          y$address_sep[i] <- str_squish(str_remove(y$address_sep[i], find))
        }
        setTxtProgressBar(pb, i)
      }
    }
    try_city[[which(st == valid_abb)]] <- y
  }
  try_city <- bind_rows(try_city)
  write_csv(try_city, addr_csv)
} else {
  try_city <- read_csv(addr_csv)
}
```

```{r city-rejoin}
sce <- sce %>% 
  select(-address_sep, -state_sep) %>% 
  left_join(try_city) %>% 
  relocate(address_sep, city_sep, state_sep, zip_sep, .after = last_col())
```

About `r percent(prop_na(sce$city_sep))` of rows contained an address where the
city name could not be matched against a list of cities in the state. This may
be due to a discrepancy in the state part of the address, a simple misspelling,
or a smaller city name not contained in our list.

```{r city-missing}
prop_na(sce$city_sep)
```

### Address

Since we can't confidently say all address are void of the city name, it's not
easy to perform our usual address normalization. We have already removed all
punctuation and forced a consistent case when we were matching city names.


### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
sce <- sce %>%
  mutate(
    zip_norm = normal_zip(
      zip = zip_sep,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  sce$zip_sep,
  sce$zip_norm,
  compare = valid_zip
)
```

```{r zip-view}
sce %>%
  filter(zip_sep != zip_norm | !is.na(zip_sep) & is.na(zip_norm)) %>%
  count(zip_sep, zip_norm, sort = TRUE)
```

```{r zip-remove}
sce <- select(sce, -zip_sep)
```

### State

We have already normalized state values when looking for cities in the address.

```{r state-remove}
prop_in(sce$state_sep, valid_state)
sce <- rename(sce, state_norm = state_sep)
```

### City

Because we have only extracted city names from addresses that _already_ matched
our list, we can't improve on those city names in any way. Any unknown or 
misspelled city named in the address that might have been fixed has been left as
a part of that string.

```{r city.rename}
sce <- rename(sce, city_norm = city_sep)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(sce, 1000))
```

1. There are `r comma(nrow(sce))` records in the database.
1. There are `r comma(sum(sce$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(sce$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(sce$exp_date[sce$exp_year > 2000]), "-")
max_dt <- str_remove_all(max(sce$exp_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("sc", "expends", "data", "clean"))
clean_csv <- path(clean_dir, glue("sc_expends_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(sce, clean_csv, na = "")
write_rds(sce, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key,
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
