---
title: "Ohio Voters"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 95)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("oh", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  textreadr, # read docx files
  docxtractr, # read docx tables
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("oh/voters/docs/oh_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

A database of voters registered in the state of Ohio can be retrieved from the
Secretary of State's [File Transfer Page][ftp].

> Welcome to the Ohio Secretary of State's Voter Files Download Page. At this
page you can download files of all registered voters by county, state senate
district, state representative district, congressional district, or the entire
state of Ohio. The statewide file is divided into four .zip files SWVF_1_22
(Adams-Erie), SWVF_23_44 (Fairfield-Lawrence), SWVF_45_66 (Licking-Pike) and
SWVF_67_88 (Portage-Wyandot). The voter files available on this page are updated
every week.

> The data is in a "comma delimited" format that can be imported into Microsoft
Excel or opened in Access as well as many other spreadsheet or database
programs. Many of the available files contain a significant quantity of data
records. A spreadsheet program, such as MS Excel, may not allow all of the data
in a file to be loaded because of a limit on the number of available rows. For
this reason, it is advised that a database application be utilized to load and
work with the data available at this site. For technical issues, please contact
us at (614) 466-8467 or voterdownload@OhioSoS.gov.

[ftp]: https://www6.ohiosos.gov/ords/f?p=VOTERFTP:STWD:::#stwdVtrFiles

We can also download a Microsoft Word document [record layout file][lay].

[lay]: https://www6.ohiosos.gov/ords/f?p=111:2::FILE_LAYOUT:NO:RP,2::

```{r layout_download}
download.file(
  url = "https://www6.ohiosos.gov/ords/f?p=111:2::FILE_LAYOUT:NO:RP,2::",
  destfile = layout_file <- file_temp(ext = "docx")
)
```

### About

```{r layout_read}
readme <- read_docx(layout_file)
```

> This document contains a description of the Voter Registration Data Files that
are available on the Ohio Secretary of State’s Election Division’s FTP site.
These files are downloaded from the current Centralized Ohio Voter Registration
Database. The database is a record collection of registered voters in the state
of Ohio, as submitted by each county Board of Elections. These records are
submitted and maintained in accordance with the Ohio Revised Code.

> Current files include voting histories for elections from year 2000 as
provided by the counties. November 20 17 General Election history is complete
as all the counties have finished the loading process.

```{r layout}
layout_file %>% 
  read_docx() %>% 
  docx_extract_tbl() %>% 
  kable()
```

## Download

We are going to download a record of voters statewide. This can be done by
scraping the download page itself for the links of the listed files and then
download each individually.

```{r raw_get}
nc_dl <- GET(
  url = "https://www6.ohiosos.gov/ords/f",
  query = list(
    p = "VOTERFTP:STWD:::#stwdVtrFiles"
  )
)
```

The files are listed in an HTML table at the bottom of the page.

```{r raw_files_table}
nc_table <- html_node(content(nc_dl), "table")
```

```{r raw_files_href}
raw_urls <- nc_table %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_c(dirname(nc_dl$url), ., sep = "/")
```

```{r raw_files_find}
nc_files <- html_table(nc_table) %>% 
  mutate(across(2, dmy)) %>% 
  mutate(across(3, ~as_fs_bytes(str_remove_all(., "[^\\w]"))))
```

```{r  raw_files_print}
nc_files %>%
  mutate(across(1, md_code)) %>% 
  mutate(across(1, md_link, raw_urls)) %>% 
  kable()
```

As the page notes, these files are individually compressed text files.

> Note: These statewide voter files are available as compressed (gzip) files.
Please use a compression tool to uncompress these files after downloading.

```{r raw_dir}
raw_dir <- dir_create(here("oh", "voters", "data", "raw"))
raw_paths <- dir_ls(raw_dir, glob = "*.gz")
```

```{r raw_head}
if (length(raw_paths) < nrow(nc_files)) {
  raw_paths <- character(nrow(nc_files))
  for (i in seq_along(raw_urls)) {
    raw_head <- HEAD(raw_urls[i])
    raw_disp <- headers(raw_head)[["content-disposition"]]
    raw_name <- str_extract(raw_disp, "(?<=filename\\=\").*(?=\";)")
    raw_paths[i] <- path(raw_dir, raw_name)
    download.file(raw_urls[i], raw_paths[i])
  }
}
```

## Read

The Workshop has an older version of this data on the TAP already. We will 
download all the new files, then our current older version, keep only those
voters _not_ found in the new version (using their unique ID), and then 
combining the two versions. This will give us the most comprehensive historical
record of registered voters.

### New

```{r raw_read}
ohv <- vroom(
  file = raw_paths,
  delim = ",",
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    DATE_OF_BIRTH = col_date(),
    REGISTRATION_DATE = col_date()
  )
)
```

```{r}
comma(nrow(ohv))
```

```{r}
vote_hist <- select(ohv, SOS_VOTERID, matches("\\d{2}/\\d{2}/\\d{4}"))
ohv <- select(ohv, -matches("\\d{2}/\\d{2}/\\d{4}"))
flush_memory(1)
```

```{r}
hist_file <- here("oh", "voters", "data", "vote_hist.tsv")
if (!file_exists(hist_file)) {
  file_create(dupe_file)
  n <- 100000 # rows per chunk
  nr <- nrow(vote_hist)
  # split file into chunks
  split_hist <- split(
    x = vote_hist, 
    f = rep(x = seq(ceiling(nr/n)), each = n, length.out = nr)
  )
  row_done <- 0
  pb <- txtProgressBar(max = nrow(ohv), style = 3)
  for (i in seq_along(split_hist)) {
    row_new <- nrow(split_hist[[i]])
    split_long <- pivot_longer(
      data = split_hist[[i]],
      cols = -SOS_VOTERID,
      names_sep = "-",
      names_to = c("type", "date"),
      names_transform = list(date = mdy),
      values_to = "vote",
      values_drop_na = TRUE
    )
    write_tsv(
      x = split_long, 
      file = hist_file, 
      append = TRUE
    )
    rm(split_long)
    split_hist[[i]] <- NA
    Sys.sleep(2)
    flush_memory(1)
    row_done <- row_done + row_new
    setTxtProgressBar(pb, value = row_done)
  }
}
```

```{r echo=FALSE}
rm(vote_hist)
flush_memory(1)
```

### Old

```{r old_save}
old_path <- here("oh", "voters", "data", "oh_voters_old.csv")
if (!file_exists(old_path)) {
  save_object(
    object = "csv/oh_voters.csv", 
    bucket = "publicaccountability", 
    file = old_path,
    show_progress = TRUE
  )
}
```

```{r old_read}
oho <- vroom(
  file = old_path,
  quote = "",
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character(),
    DATE_OF_BIRTH = col_date(),
    REGISTRATION_DATE = col_date()
  )
)
```

```{r old_filter}
comma(nrow(oho))
oho <- filter(oho, SOS_VOTERID %out% ohv$SOS_VOTERID)
comma(nrow(oho))
```

```{r echo=FALSE}
flush_memory(2)
```

```{r}
setdiff(names(oho), names(ohv))
oho <- select(oho, -all_of(setdiff(names(oho), names(ohv))))
all(names(oho) %in% names(ohv))
```

### Bind

```{r old_bind}
ohv <- bind_rows(new = ohv, old = oho, .id = "source_file")
comma(nrow(ohv))
rm(oho)
```

```{r echo=FALSE}
flush_memory(1)
```

```{r raw_rename, echo=FALSE}
ohv <- ohv %>% 
  relocate(source_file, .after = last_col()) %>% 
  clean_names("snake") %>% 
  rename(vote_city = city) %>% 
  rename_with(~str_remove(., "^residential_")) %>% 
  rename_with(~str_replace(., "^mailing_", "mail_")) %>% 
  rename(
    birth_date = date_of_birth,
    reg_date = registration_date,
    party = party_affiliation,
    addr1 = address1,
    addr2 = secondary_addr,
    zip4 = zip_plus4,
    postal = postalcode,
    mail_addr1 = mail_address1,
    mail_addr2 = mail_secondary_address,
    mail_zip4 = mail_zip_plus4,
    mail_postal = mail_postal_code
  )
```

## Explore

There are `r scales::comma(nrow(ohv))` rows of `r ncol(ohv)` columns.

```{r glimpse}
glimpse(ohv)
tail(ohv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(ohv, count_na)
```

We can flag any record missing a key variable needed to identify a voter.

```{r na_flag}
key_vars <- c("sos_voterid", "last_name", "birth_date", "reg_date")
ohv <- flag_na(ohv, all_of(key_vars))
sum(ohv$na_flag)
```

```{r na_view}
ohv %>% 
  filter(na_flag) %>% 
  select(sos_voterid, first_name, middle_name, last_name, birth_date)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("oh", "voters", "dupes.tsv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  ohs <- ohv %>% 
    select(-sos_voterid, -county_id) %>% 
    group_split(county_number)
  split_id <- split(ohv$sos_voterid, ohv$county_number)
  pb <- txtProgressBar(max = length(ohs), style = 3)
  for (i in seq_along(ohs)) {
    d1 <- duplicated(ohs[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(ohs[[i]], fromLast = TRUE)
      dupes <- tibble(sos_voterid = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(ohs)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("sos_voterid", "dupe_flag"),
  col_types = cols(
    sos_voterid = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(ohv)
ohv <- left_join(ohv, dupes, by = "sos_voterid")
ohv <- mutate(ohv, dupe_flag = !is.na(dupe_flag))
sum(ohv$dupe_flag)
```

We can see that, despite unique IDs, there are duplicate voters.

```{r dupe_view}
ohv %>% 
  filter(dupe_flag) %>% 
  select(sos_voterid, all_of(key_vars)) %>% 
  arrange(last_name)
```

### Categorical

```{r distinct_count}
col_stats(ohv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(ohv, voter_status)
ohv %>% 
  count(party) %>%
  mutate(p = n/sum(n)) %>% 
  filter(p > 0.1) %>% 
  ggplot(aes(party, p)) +
  geom_col(aes(fill = party)) +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(
    values = c("#3333FF", "#E81B23"),
    na.value = "#666666"
  )
ohv %>% 
  count(county_number) %>% 
  mutate(across(1, parse_integer)) %>% 
  group_by(county_number) %>% 
  summarise(n = sum(n)) %>% 
  ggplot(aes(county_number, n)) +
  geom_col()
```

### Dates

`r percent(mean(ohv$reg_date == "1900-01-01"), 0.1)` of voters are registered
on January 1st, 1900. This must be the default when missing a date. We will
remove these values.

```{r reg_date_range}
min(ohv$reg_date)
mean(ohv$reg_date == "1900-01-01")
ohv$reg_date[ohv$reg_date == "1900-01-01"] <- NA
max(ohv$reg_date, na.rm = TRUE)
```

The birth dates are similar.

```{r birth_date_range}
min(ohv$birth_date)
sum(ohv$birth_date == "1800-01-01")
sum(ohv$birth_date == "1900-01-01")
ohv$birth_date[ohv$birth_date == "1800-01-01"] <- NA
ohv$birth_date[ohv$birth_date == "1900-01-01"] <- NA
max(ohv$birth_date, na.rm = TRUE)
```

We can add the calendar year from `reg_date` with `lubridate::year()`.

```{r date_year}
ohv <- mutate(ohv, reg_year = year(reg_date), birth_year = year(birth_date))
```

```{r bar_reg_year, echo=FALSE}
ohv %>% 
  count(reg_year) %>% 
  filter(!is.na(reg_year)) %>% 
  mutate(
    type = if_else(
      condition = reg_year %% 4 == 0,
      true = "pres.",
      false = if_else(
        condition = reg_year %% 2 == 0,
        true = "mid.",
        false = "none"
      )
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = type)) + 
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1956, 2020, by = 4)) +
  coord_cartesian(xlim = c(1956, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Ohio Voters Registered by Year",
    caption = "Source: OH SOS",
    fill = "Election Registered",
    x = "Year Made",
    y = "Count"
  )
```

```{r bar_birth_year, echo=FALSE}
ohv %>% 
  filter(!is.na(birth_year)) %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(fill = dark2["purple"], binwidth = 1) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2004, by = 4)) +
  coord_cartesian(xlim = c(1900, 2004)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Ohio Voters by Birth Year",
    caption = "Source: OH SOS",
    x = "Year Born",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

The residential `addr1` variable is already normalized how we want.

```{r addr_view}
ohv %>% 
  select(addr1) %>% 
  sample_n(20)
```

As we would expect, the `state` and `zip` variables are also clean.

```{r geo_check}
prop_in(ohv$zip, valid_zip)
prop_in(ohv$state, valid_state)
```

The cities are mostly valid, but the database uses abbreviations like "HTS"
rather than "HEIGHTS", as we would in our database. We can expand these
abbreviations with `campfin::normal_city()`.

```{r city_clean}
many_city <- c(valid_city, extra_city)
prop_in(ohv$city, many_city)
norm_city <- ohv %>% 
  count(city, sort = TRUE) %>% 
  select(-n) %>% 
  filter(city %out% many_city) %>% 
  mutate(across(city, list(clean = normal_city), abbs = usps_city))
norm_city
prop_in(norm_city$city_clean, many_city)
```

This brings our known value above 99%.

```{r city_join}
ohv <- left_join(ohv, norm_city)
ohv <- mutate(ohv, city_clean = coalesce(city_clean, city))
prop_in(ohv$city_clean, many_city)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(ohv, 50))
```

1. There are `r comma(nrow(ohv))` records in the database.
1. There are `r comma(sum(ohv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ohv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("oh", "voters", "data", "clean"))
clean_path <- path(clean_dir, "oh_voters_clean.csv")
write_csv(ohv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
