---
title: "CMS AAP Program COVID-19 Payments"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "contract", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tabulizer, # read pdf tables
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("us/covid/cms_aapp/docs/cms_aapp_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Centers for Medicare & Medicaid Services (CMS) Accelerated and Advance Payment
(AAP) Program [Fact Sheet][fact]:

> On March 28 2020, CMS expanded the existing Accelerated and Advance Payments
Program to a broader group of Medicare Part A providers and Part B suppliers. An
accelerated or advance payment is a payment intended to provide necessary funds
when there is a disruption in claims submission and/or claims processing. CMS
can also offer these payments in circumstances such as national emergencies, or
natural disasters in order to accelerate cash flow to the impacted health care
providers and suppliers.
>
> The subsequent passage of the Coronavirus Aid, Relief, and Economic Security
(CARES) Act (P.L. 116-136) on March 27, 2020, amended the existing Accelerated
Payments Program to provideadditional benefits and flexibilit ies, including
extended repayment timeframes, to the subset ofproviders specifically referenced
in the CARES Act, including inpatient hospitals, children’s hospitals,certain
cancer hospitals, and critical access hospitals.
>
> The Continuing Appropriations Act, 2021 and Other Extensions Act (P.L.
116-159), enacted on October 1, 2020,amended the repayment terms for
allproviders and suppliers who requested and received accelerated and advance
payment(s) during the COVID-19 Public Health Emergency (PHE). Details on
repayment terms are provided below.
>
> Although we announced the pause of the Accelerated Payments Program and the
suspension of the Advance Payments Program on April 26, 2020, CMS continued to
accept applications from providers as they related to the COVID-19 public health
emergency (PHE). Please note that, as of October 8, 2020, CMS will no longer
accept applications for accelerated or advance payments as they relate to the
COVID-19 PHE, although CMS will continue to monitor the ongoing impacts of
COVID-19 on the Medicare provider and supplier community.

[fact]: https://www.cms.gov/files/document/accelerated-and-advanced-payments-fact-sheet.pdf

## Download

The PDF containing the AAP Program payments can be downloaded from the CMS.

```{r raw_dir}
raw_url <- "https://www.cms.gov/files/document/covid-medicare-accelerated-and-advance-payments-program-covid-19-public-health-emergency-payment.pdf"
raw_dir <- dir_create(here("us", "covid", "cms_aapp", "data", "raw"))
raw_path <- path(raw_dir, basename(raw_url))
```

```{r raw_download}
if (!file_exists(raw_path)) {
  download.file(raw_url, raw_path)
}
```

## Read

The tables from that PDF are extracted using the free [Tabula] tool.

[Tabula]: https://tabula.technology/

```{r raw_path}
raw_zip <- path(raw_dir, "tabula-extract.zip")
raw_csv <- unzip(raw_zip, exdir = raw_dir)
```

The extracted CSV files can be read into a single data frame.

```{r raw_read}
aapp <- map_df(
  .x = raw_csv,
  .f = read_csv,
  col_types = "ccn",
  col_names = c(
    "National Provider Identifier",
    "Provider/Supplier Name",
    "Payment Amount "
  )
)
```

```{r raw_trim}
aapp[nrow(aapp), ]
aapp <- aapp[-nrow(aapp), ]
names(aapp) <- c("npi", "name", "amount")
```

```{r raw_newline}
aapp$name <- str_replace_all(aapp$name, "\r", " ")
```

## Join

Providers/suppliers are identified by their National Provider Identifier (NPI)
number. We can download a separate table of hospitals, etc. by their NPI.

> In September 2007, CMS began disclosing NPPES health care provider data that
are disclosable under the Freedom of Information Act (FOIA) to the public.  In
accordance with the e-FOIA Amendments, CMS has disclosed these data via the
Internet in two forms:
> 
> 1. [NPI Registry][reg]: The NPI Registry is a query-only database which is 
     updated daily to enable users to query the NPPES (e.g., search by NPI, 
     provider name, etc.) and retrieve the FOIA-disclosable data from the search 
     results. There is no charge to view the data.
> 2. [NPI Downloadable File][df]: Full Replacement Monthly NPI File, Weekly 
     Incremental NPI File, and Full Replacement NPI Deactivation File. There is 
     no charge to download the data.

[reg]: https://npiregistry.cms.hhs.gov/
[df]: http://download.cms.gov/nppes/NPI_Files.html

We want the [full download file][df].

> The zipped NPPES Downloadable file will include the following 3 reference 
files: 
> 
> 1. **Other Name Reference File** - this file contains additional Other Names 
     associated with Type 2 NPIs.
> 2. **Practice Location Reference File** - this file contains all of the 
     non-primary Practice Locations associated with Type 1 and Type 2 NPIs.
> 3. **Endpoint Reference File** - this file contains all Endpoints associated 
     with Type 1 and Type 2 NPIs.
> 
> Detailed information about the new files will be included in the Data
Dissemination File - Readme file.

> ##### Full Replacement Monthly NPI File
> * [NPPES Data Dissemination (February 09, 2021)][dd] - ZIP format (778 MB)
>
> Beginning with the May 2008 file, the size of the `.csv` Header File increased
because we added Authorized Official data elements to it that were inadvertently
not previously included. As a result of the larger size, the `.csv` Header File
can no longer be opened in Excel. Even so, the `.csv` Header File is small
enough to be opened by a simple text editor like Notepad. As always, the `.csv`
data file must be opened with a text editor capable of opening large files. Two
examples are TextPad and UltraEdit.
> 
> The data dissemination file exceeds 4 GB. Older file compression utilities may
not be able to extract this file properly. Please check your file compression
utility for any known problems and the latest version in order to avoid any
problems extracting this file. Utilities, such as WinZip and 7-zip, have been
successful in extracting the file.

[dd]: https://download.cms.gov/nppes/NPPES_Data_Dissemination_February_2021.zip

```{r}
npi_url <- "https://download.cms.gov/nppes/NPPES_Data_Dissemination_February_2021.zip"
npi_zip <- path(raw_dir, basename(npi_url))
if (!file_exists(npi_zip)) {
  download.file(npi_url, npi_zip)
}
```

## Explore

There are `r comma(nrow(aapp))` rows of `r ncol(aapp)` columns. Each record
represents a single payment made to a provider/supplier under the AAP program.

```{r glimpse}
glimpse(aapp)
tail(aapp)
```

### Missing

No records are missing any values.

```{r na_count}
col_stats(aapp, count_na)
```

### Duplicates

There are no entirely duplicate rows, but there are a few duplicate IDs. These
are multiple payments to the same provider/supplier. Without a date value, it's
not clear what's the cause.

```{r dupe_flag}
aapp <- aapp %>% 
  group_by(npi) %>% 
  mutate(dupe_flag = n() > 1) %>% 
  ungroup()
```

```{r dupe_view}
filter(aapp, dupe_flag)
```

### Amounts

```{r amount_summary}
summary(aapp$amount)
mean(aapp$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(aapp[c(which.max(aapp$amount), which.min(aapp$amount)), ])
```

```{r hist_amount, echo=FALSE}
aapp %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Medicare Accelerated and Advance Payments Program",
    subtitle = "COVID-19 Public Health Emergency Payment Data",
    caption = "Source: Centers for Medicare & Medicaid Services",
    x = "Amount",
    y = "Count"
  )
```

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "covid", "cms_aapp", "data", "clean"))
clean_path <- path(clean_dir, "us_cms_covid_aapp.csv")
write_csv(aapp, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
