---
title: "New Jersey Lobbying Registration"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("nj", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  httr, # interact with http responses
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://www.elec.state.nj.us/forcandidates/gaa_pub_info.htm
[04]: https://www3-elec.mwg.state.nj.us/ELEC_AGAA/entitysearch.aspx

## Data

Lobbyist data is obtained from the [New Jersey Election Law Enforcement Commission][03].

#### Download
> "View Quarterly Reports and Quarterly Summary" -> "Entity Search" -> "Search" -> "Download Records"

The [Entity Search page][04], we can see a table is generated and can be retrieved:

The data is downloaded on June 12, 2020.


First, we must download a reporting linking lobbyists to their principals.This information is included in the Index of Registered Entities.

Then we can join this table with another table `Agent Name` which contains lobbyists address information, as well as the `Lobbying Entity/Business` table which lists all geographical data of lobbying firms.

```{r raw_dir}
raw_dir <- dir_create(here("nj", "lobby", "data", "raw", "reg"))
```


```{r eval=FALSE}
njlr <- read_csv(dir_ls(raw_dir, regexp = "Index.+"))
nj_lob <- read_csv(dir_ls(raw_dir, regexp = "Agent.csv"))
```

## Import

We can separate the firm address into address, city, state and ZIP.

```{r read_lobs}
njlr <- njlr %>%
  mutate(firm_address = str_extract(Firm_Address, "^[^,]+(?=\\s*,)"),
         city_state_zip = str_remove(Firm_Address, "^[^,]+(\\s*,)"))

njlr <- njlr %>% 
  separate(col = city_state_zip, into = c("city","state_zip"),sep = ",", remove = F) %>% 
  separate(col = state_zip, into = c("state","zip"),sep = "\\s", remove = F)
```

```{r read_pris}
pris <- 
  read_delim(
    file = in_raw("sc_pris.csv"),
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    na = c("", " ")
  ) %>% 
  remove_empty("cols") %>% 
  clean_names("snake")

names(pris) <- names(pris) %>% 
  str_remove("_(.*)") %>% 
  str_remove("code$") %>%
  str_replace("^lpname$", "name") %>% 
  str_c("pri", ., sep = "_")
```

```{r read_xref}
xref <- read_delim(
    file = in_raw("sc_xref.csv"),
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    col_types = cols(.default = "c"),
    na = c("", " ")
  ) %>% 
  remove_empty("cols") %>% 
  clean_names("snake") %>% 
  rename(
    lob_last = lastname,
    lob_first = firstname,
    lob_address = address,
    lob_city = city,
    lob_state = state,
    lob_zip = zipcode,
    lob_phone = phone,
    pri_name = principal_name,
    lob_mi = middle_init,
    lob_suffix = suffix
    )
```

```{r join_xref}
sclr <- xref %>% 
  left_join(pris, by = "pri_name") %>% 
  select(
    starts_with("lob"), 
    starts_with("pri")
  )
```

```{r}
prop_in(
  x = str_normal(paste(lobs$lob_first, lobs$lob_last)),
  y = str_normal(paste(xref$lob_first, xref$lob_last)),
)
```

