---
title: "New Jersey Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nj", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Lobbying data is obtained from the [Election Law Enforcement Commission][elec].

> The Election Law Enforcement Commission (ELEC) is dedicated to administering
“The New Jersey Campaign Contributions and Expenditures Reporting Act‚” “The
Gubernatorial‚ Legislative Disclosure Statement Act‚” “Legislative and
Governmental Process Activities Disclosure Act‚” and parts of several other
laws.

ELEC [provides a portal][portal] where users can search for information on
lobbyist as well as their firms and clients. From the portal, we can run empty
searches on five different categories to return all results and export them:

1. `Agent.csv` for lobbyist names and addresses.
2. `BillSearch.csv` for lobbying activity on certain bills.
3. `Client.csv` for client names.
4. `Firm.csv` for firm names and addresses.
5. `IndexClient.csv` for agent, firm, client relationships.

[elec]: https://www.elec.state.nj.us/
[portal]: https://www3-elec.mwg.state.nj.us/ELEC_AGAA/EntitySearch.aspx

## Read

The results data was manually exported to the `data/raw/` directory.

```{r raw_dir}
raw_dir <- dir_create(here("nj", "lobby", "data", "raw"))
raw_info <- as_tibble(dir_info(raw_dir))
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

First, we will read the `IndexClient.csv` file containing the relationships
between lobbying agents, their firms, and the client entities they represent.

```{r raw_read}
njl <- vroom(
  file = path(raw_dir, "IndexClient.csv"),
  delim = ",",
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    Active_Date = col_date_usa()
  )
)
```

```{r raw_select, echo=FALSE}
njl <- select(njl, -phone_number2)
```

We can confirm the file was read properly by counting the distinct values of a
discrete variable.

```{r raw_check}
count(njl, agent_status)
```

The only information we need to add from one of the other files is the agent's
termination date. This can be found in the `Agent.csv` file.

```{r raw_agent}
agent <- vroom(
  file = path(raw_dir, "Agent.csv"),
  delim = ",",
  col_select = c(agent_name, firm_name, active_date, agent_term_date),
  col_types = cols(
    .default = col_character(),
    active_date = col_date_usa(),
    agent_term_date = col_date_usa()
  )
)
```

```{r agent_rename, echo=FALSE}
agent <- distinct(agent)
agent <- rename(agent, term_date = agent_term_date)
```

This termination date can then be joined to the index data frame using the
agent's full name.

```{r raw_join}
mean(njl$agent_name %in% agent$agent_name)
njl <- left_join(njl, agent)
```

## Explore

```{r glimpse}
glimpse(njl)
tail(njl)
```

### Missing

```{r na_count}
col_stats(njl, count_na)
```

```{r na_flag}
njl <- njl %>% flag_na(active_date, agent_name, firm_name, client_name)
sum(njl$na_flag)
```

```{r na_view}
njl %>% 
  filter(na_flag) %>% 
  select(active_date, agent_name, firm_name, client_name)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
njl <- flag_dupes(njl, everything())
```

### Categorical

```{r distinct_count}
col_stats(njl, n_distinct)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
njl <- mutate(njl, year = year(active_date))
```

```{r date_range}
min(njl$active_date)
max(njl$active_date)
sum(njl$active_date > today())
```

```{r bar_year, echo=FALSE}
njl %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 2)) +
  coord_cartesian(xlim = c(1989, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Jersey Lobbyists Registered by Year",
    caption = "Source: NJ ELEC",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

First, we must `tidyr::separate()` the single `firm_address` variable into it's
component pieces.

```{r address_sep}
njl <- njl %>% 
  separate(
    col = firm_address,
    into = c("addr_sep", "city_sep", "state_zip"),
    sep = "([:blank:]+)?,",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) %>% 
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s+(?=\\d)",
    remove = TRUE,
    extra = "merge"
  )
```

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
njl <- mutate(
   .data = njl,
   addr_norm = normal_address(
     address = addr_sep,
     abbs = usps_street,
     na = invalid_city
   )
 )
```

```{r address_view}
njl %>% 
  select(addr_sep, addr_norm) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
njl <- mutate(
  .data = njl,
  zip_norm = normal_zip(
    zip = zip_sep,
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  njl$zip_sep,
  njl$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
njl <- njl %>% 
  mutate(
    state_norm = normal_state(
      state = state_sep,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
njl %>% 
  filter(state_sep != state_norm) %>% 
  count(state_sep, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  njl$state_sep,
  njl$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
njl <- njl %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep, 
      abbs = usps_city,
      states = c("NJ", "DC", "NEW JERSEY"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(njl$city_sep),
  njl$city_norm,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
njl <- njl %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  select(-ends_with("_sep"))
```

```{r clean_glimpse}
glimpse(sample_n(njl, 20))
```

1. There are `r comma(nrow(njl))` records in the database.
1. There are `r comma(sum(njl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(njl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nj", "lobby", "data", "clean"))
clean_path <- path(clean_dir, "nj_lobby_clean.csv")
write_csv(njl, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(njl)),
  type = md_code(map_chr(njl, typeof)),
  def = c(
    "Lobbyist full name",
    "Unique firm ID",
    "Firm name",
    "A = active, T = terminated",
    "Date lobbyist became active",
    "Lobbyist badge number",
    "Full firm address",
    "Firm telephone number",
    "Client entity name",
    "Firm full name (w/ address)",
    "Date lobbyist was terminated",
    "Flag indicating missing variable",
    "Calendar year lobbyist active",
    "Normalized firm business address",
    "Normalized firm business city",
    "Normalized firm business state",
    "Normalized firm business ZIP code"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
