---
title: "Montana Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mt", "lobbying", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Montana Commissioner of Political Practices][cpp].

> The Commissioner of Political Practices also makes information about lobbying money in Montana
publicly available. Any member of the public can see who is registered as a lobbyist, who is a
principal (a person or entity that pays lobbyists to lobby on their behalf), and how money is spent
to influence legislation and legislators.
> 
> See a list of registered principals and lobbyists for the 2019 legislative session [here][xlsx]
as an excel doc or [here][pdf] as a PDF.

[cpp]: https://politicalpractices.mt.gov/lobbying
[xlsx]: https://politicalpractices.mt.gov/Portals/144/2019%20Communication%20assets/PDFs%20for%20website/2019%20list%20of%20principals%20and%20lobbyists%20for%20website.xlsx?ver=2019-04-05-161928-000
[pdf]: https://politicalpractices.mt.gov/Portals/144/2019%20Communication%20assets/PDFs%20for%20website/2019%20list%20of%20principals%20and%20lobbyists%20for%20website.pdf?ver=2019-04-17-105845-647

### Download

We can download the excel version of the file.

```{r create_raw_dir}
raw_dir <- here("mt", "lobbying", "reg", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw}
mtlr_base <- "https://politicalpractices.mt.gov/Portals/144"
mtlr_folders <- str_c("2019 Communication assets", "PDFs for website", sep = "/")
mtlr_file <- "2019 list of principals and lobbyists for website.xlsx"
mtlr_url <- str_c(mtlr_base, mtlr_folders, mtlr_file, sep = "/")
mtlr_path <- url2path(mtlr_url, raw_dir)
download.file(url = mtlr_url, destfile = mtlr_path)
```

### Read

Using the `readxl::read_excel()` function, we can import this file as a dataframe.

```{r read_raw}
mtlr <- read_excel(mtlr_path, skip = 2)
```

## Pivot

However, the file is in _wide_ format, with rows representing clients and multiple columns
containing the lobbyists for that client. Using `tidyr::pivot_longer()`, we can pivot this wide
format to a long format, with one row per client/lobbyist relationship.

```{r pivot_longer}
mtlr <- mtlr %>% 
  pivot_longer(
    cols = starts_with("Lobbyist"),
    values_to = "Lobbyist Name",
    names_to = c("Type", "Number"),
    names_sep = " ",
    names_ptypes = list(
      "Type" = character(), 
      "Number" = numeric()
    )
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  drop_na(lobbyist_name)
```

```{r rename}
names(mtlr)[2:3] %>% 
  str_extract( "(.*)_report") %>% 
  str_remove_all("\\d") %>% 
  str_remove_all("_._") -> names(mtlr)[2:3]
```

## Explore

```{r glimpse}
head(mtlr)
tail(mtlr)
glimpse(sample_frac(mtlr))
```

