---
title: "Pennsylvania Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("pa", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Pennsylvania Department of State](https://www.dos.pa.gov). From their
[search page](https://www.palobbyingservices.pa.gov/Public/wfSearch.aspx), we can "Export All" from
the advance search.

This provides an `Export.zip` archive file.

```{r raw_dir}
raw_dir <- here("pa", "lobby", "data", "raw")
dir_create(raw_dir)
raw_zip <- dir_ls(raw_dir, glob = "*.zip")
```

This file contains a single CSV.

```{r zip_list}
unzip(raw_zip, list = TRUE)
```

## Import

We can directly pass this ZIP file to `readr::read_csv()`.

```{r zip_read}
export <- raw_zip %>% 
  read_lines() %>% 
  str_replace("\"I\"", "'I'") %>% 
  read_delim(
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    col_types = cols(
      .default = col_character(),
      IsRelatedToPrincipal = col_logical(),
      IsRelatedToFirm = col_logical(),
      IsRelatedToLobbyist = col_logical(),
      IsRelatedToPAC = col_logical(),
      IsOfficerCandidatePAC = col_logical(),
      IsLicensedAttorney = col_logical(),
      IsLicensedInPA = col_logical(),
      DuePayingMembers = col_integer(),
      CalendarYear = col_integer(),
      IsLobbyingFirm = col_logical(),
      IsLawFirm = col_logical(),
      IsOtherFirm = col_logical(),
      LastUpdatedDate = col_date_usa(),
      SubmittedDate = col_date_usa(),
      RenewalDate = col_date_usa(),
      LobbyingCommencedDate = col_date_usa(),
      ExpirationDate = col_date_usa(),
      TerminatedDate = col_date_usa(),
      AffiliatedStartDate = col_date_usa(),
      AffiliatedEndDate = col_date_usa(),
      IsNeverAffiliated = col_logical(),
      IsRepresentThirdParty = col_logical(),
      AffirmedDate = col_date_usa()
    )
  )
```

```{r trim, echo=FALSE}
export <- mutate_if(export, is_character, str_trim)
```

This data frame contains `r comma(nrow(export))` rows, with a separate record for each relationship
between a lobbyist, lobbying firm, or principal client. That means a single relationship between a
lobbyist and a principal is listed twice, once with the lobbyist as the registrant and again with
the principal. We will restructure and filter this data frame so that only lobbyists are listed as
registrants, with a separate record for each of their principals.

```{r raw_rows}
palr <- filter(export, RegistrationType == "Lobbyist") %>% clean_names("snake")
```

We can also remove all the columns associated with the affirmation of the
registration. As per the [disclosure instructions][05]:

> **Affirmations and Acknowledgments:** The registration must include a typed or printed name of an
individual, which, subject to 18 Pa.C.S. § 4904 (unsworn falsifications to authorities), states
that the information provided is true and correct to the best of the filer’s knowledge, information
and belief. The registration statement also contains an acknowledgment that the registrant or their
designee has received, read and understands the Act.

[03]: https://www.dos.pa.gov/OtherServices/LobbyingDisclosure/Documents/LDR/InstructionsLobbyistonlinefiling061311.pdf

```{r raw_cols}
palr <- select(palr, -starts_with("affirm"), -is_represent_third_party)
```

To make this data set a little easier to understand, we will prefix each registrant column, now
filtered to include only lobbyists, with `lob_`.

```{r raw_lob_rename}
names(palr)[1:15] <- names(palr)[1:15] %>% 
  str_remove("^registration_") %>% 
  str_c("lob", ., sep = "_")
```

Then, we can abbreviate the other columns... This just makes things easier to type.

```{r raw_rename}
names(palr) <- names(palr) %>% 
  str_replace("(?<=(_|^))affili.*?(?=($|_))", "pri") %>% 
  str_replace("(?<=(_|^))lobby.*?(?=($|_))", "lob") %>% 
  str_replace("(?<=(_|^))num.*?(?=($|_))", "id") %>% 
  str_replace("(?<=(_|^))regist.*?(?=($|_))", "reg") %>% 
  str_replace("business", "biz") %>% 
  str_replace("candidate", "cand") %>% 
  str_replace("calendar", "cal") %>% 
  str_replace("description", "desc") %>% 
  str_replace("expiration", "expire") %>% 
  str_replace("principal", "pri") %>% 
  str_replace("terminated", "term") %>% 
  str_replace("related", "rel") %>% 
  str_replace("submitted", "sub") %>% 
  str_replace("postal_code", "zip") %>% 
  str_remove("_name$")
```

## Explore

```{r}
head(palr)
tail(palr)
glimpse(sample_frac(palr))
```

Having filtered `r percent(1-(nrow(palr)/nrow(export)))` of the rows in the database, we can now
remove some of the columns which now contain no information.

```{r count_na}
col_stats(palr, count_na)
```

```{r remove_empty}
palr <- remove_empty(palr, which = "cols")
```

We can also check how many distinct values are present in each column.

```{r n_distinct}
col_stats(palr, n_distinct)
```

To better understand the type of records in the database, we can create a new variable indicating
the year of the relationship registration.

```{r date_view}
select(palr, ends_with("date"))
```

```{r date_year}
palr <- mutate(palr, sub_year = year(sub_date))
```

```{r}
palr %>% 
  count(sub_year) %>% 
  ggplot(aes(x = sub_year, y = n)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Pennsylvanian Lobbyist Registrations per Year",
    subtitle = "Carryover date to 2007?",
    x = "Year submitted",
    y = "Disclosures Submitted"
  )
```

## Wrangle

To improve the uniformity and searchability of our database, we will perform some very consistant
and confident text normalization.

### Phone

First, will create a single new telephone number variable for each lobbyist and principal by
using `tidyr::unite()` to combine the number and extension and then pass that variable to
`campfin::normal_phone()` to convert to a new format.

```{r phone_norm}
palr <- palr %>% 
  unite(
    starts_with("lob_phone"),
    col = lob_phone_full,
    sep = "x",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  unite(
    starts_with("pri_phone"),
    col = pri_phone_full,
    sep = "x",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    lob_phone_norm = normal_phone(lob_phone_full),
    pri_phone_norm = normal_phone(pri_phone_full)
  ) %>% 
  select(-ends_with("phone_full")) %>% 
  mutate_at(
    .vars = vars(ends_with("fax")),
    .funs = list(norm = normal_phone)
  )
```

```{r phone_view, echo=FALSE}
palr %>% 
  select(contains("phone")) %>% 
  distinct() %>% 
  sample_frac()
```

### Address

We will perform a similar change to the street address, `tidyr:unite()`ing the two line variables
and passing them to `campfin:normal_address()` to expand abbreviations and improve consistency.

```{r address_norm}
palr <- palr %>% 
  unite(
    starts_with("lob_line"),
    col = lob_addr,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  unite(
    starts_with("pri_line"),
    col = pri_addr,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate_at(
    .vars = vars(ends_with("addr")),
    .funs = list(norm = normal_address),
    abb = usps_street
  ) %>% 
  select(-ends_with("addr"))
```

```{r address_view, echo=FALSE}
palr %>% 
  select(lob_line1, lob_line2, lob_addr_norm) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

We will use `campfin:normal_zip()` to coerce 5-digit ZIP codes.

```{r zip_norm}
palr <- mutate_at(
  .tbl = palr,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

This improves the proportion of ZIP codes that are valid.

```{r zip_progress, echo=FALSE}
progress_table(
  palr$lob_zip,
  palr$lob_zip_norm,
  palr$pri_zip,
  palr$pri_zip_norm,
  compare = valid_zip
)
```

### State

The 2-character state codes for both lobbyist and principal need no further cleaning.

```{r state}
prop_in(palr$lob_state, valid_state)
prop_in(palr$pri_state, valid_state)
```

### City

The city strings are the most difficult to clean, given the variety in city names and the wide
range of quasi-valid formats in which they might be writte. We can use `campfin::normal_city()`
to normalize the strings.

```{r city_normal, echo=FALSE}
palr <- mutate_at(
  .tbl = palr,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  na = invalid_city
)
```

Then, we compare these normalized strings to the expected city name for that record's state and ZIP
code. If the normalized string is extremelly similar to the expected string, we can safely replace
what is written with what we would expect.

```{r}
palr <- palr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lob_state" = "state",
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(lob_city_match = city) %>% 
  mutate(
    lob_match_abb = is_abbrev(lob_city_norm, lob_city_match),
    lob_match_dist = str_dist(lob_city_norm, lob_city_match),
    lob_city_swap = if_else(
      condition = !is.na(lob_city_match) & (lob_match_abb | lob_match_dist == 1),
      true = lob_city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -lob_city_match,
    -lob_match_abb,
    -lob_match_dist
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "pri_state" = "state",
      "pri_zip_norm" = "zip"
    )
  ) %>% 
  rename(pri_city_match = city) %>% 
  mutate(
    pri_match_abb = is_abbrev(pri_city_norm, pri_city_match),
    pri_match_dist = str_dist(pri_city_norm, pri_city_match),
    pri_city_swap = if_else(
      condition = !is.na(pri_city_match) & (pri_match_abb | pri_match_dist == 1),
      true = pri_city_match,
      false = pri_city_norm
    )
  ) %>% 
  select(
    -pri_city_match,
    -pri_match_abb,
    -pri_match_dist
  )
```

These two steps _drastically_ improve the consistency in city names.

```{r city_progress, echo=FALSE}
progress <- 
  progress_table(
    str_to_upper(palr$lob_city),
    palr$lob_city_norm,
    palr$lob_city_swap,
    str_to_upper(palr$pri_city),
    palr$pri_city_norm,
    palr$pri_city_swap,
    compare = c(valid_city, extra_city)
  ) %>% 
  separate(
    col = stage,
    sep = "_",
    into = c("type", "stage"),
    extra = "merge"
  ) %>% 
  group_by(type) %>% 
  mutate(stage = as_factor(stage))
```

```{r plot_prop, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_col(aes(fill = type), position = "dodge") +
  scale_fill_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0.99) +
  facet_wrap(~type) +
  coord_cartesian(ylim = c(0.75, 1.00)) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Pennsylvania City Normalization Progress",
    x = "Stage",
    y = "Percent Valid"
  )
```

```{r plot_distinct, echo=FALSE}
progress %>% 
  select(
    type,
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  ) +
  facet_wrap(~type)
```

## Export

```{r write_proc}
proc_dir <- here("pa", "lobby", "data", "processed")
dir_create(proc_dir)
write_csv(
  x = palr,
  path = glue("{proc_dir}/pa_lobbyists.csv"),
  na = ""
)
```

