---
title: "Virginia Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("va", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # print markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  aws.s3, # aws cloud storage
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # read html pages
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Virginia [Department of Elections (DOE)][doe] State
Board of Elections (SBE).

> The State Board of Elections is the regulatory board for the Department of
Elections. It is authorized to prescribe standard forms for voter registration
and elections, and to supervise, coordinate, and adopt regulations governing the
work of local electoral boards, registrars, and officers of election.

[doe]: https://www.elections.virginia.gov/index.html

The SBE has candidates and committees file ten types of reports:

```{r raw_scheds, echo=FALSE}
"https://www.elections.virginia.gov/candidatepac-info/" %>% 
  str_c("campaign-finance-disclosure-forms/") %>% 
  read_html() %>% 
  html_node(".schedules_table") %>% 
  html_table() %>% 
  as_vector() %>% 
  enframe(NULL) %>% 
  separate(
    col = value,
    into = c("sched", "desc"),
    sep = "–"
  ) %>% 
  mutate(across(.fns = str_trim)) %>% 
  mutate(across(desc, str_trunc, width = 80)) %>% 
  kable(col.names = c("Schedule", "Description"))
```

We are interested in schedule's A and B.

## Download

The data itself is hosted on an SBE server and organized in monthly directories.

```{r raw_list}
sbe_url <- "https://apps.elections.virginia.gov/SBE_CSV/CF/"
sbe_list <- read_html(sbe_url)
```

```{r raw_show, echo=FALSE}
sbe_table <- sbe_list %>% 
  html_node("pre") %>% 
  html_text() %>% 
  str_remove("(.*)\\]") %>% 
  str_remove("10/24/2018  1:45 PM\\s+168 web.config") %>% 
  str_replace_all(
    pattern = "(?<=<dir> (\\d{4}|\\d{4}_\\d{2}))(?=(\\s+)?\\d{1,2}/)", 
    replacement = "\n"
  ) %>%
  read_table(col_names = FALSE) %>% 
  unite(datetime, 1:3, sep = " ") %>% 
  rename(type = X4, dir = X5) %>% 
  mutate(across(datetime, mdy_hms)) %>% 
  filter(str_detect(dir, "_"))

sbe_table %>% 
  select(-type) %>% 
  mutate(
    dir = md_link(
      dir,
      paste0(
        "https://apps.elections.virginia.gov/SBE_CSV/CF/", 
        sbe_table$dir
      )
    )
  ) %>% 
  kable()
```

From this list of directories, we can extract each `<href>` HTML tag. These
tags provide the directory names.

```{r sbe_dirs_get, results='asis'}
raw_dir <- dir_create(here("va", "contribs", "data", "raw"))
sbe_base <- "https://apps.elections.virginia.gov"
sbe_dirs <- sbe_list %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("\\d_")
```

```{r sbe_dirs_list, echo=FALSE}
sbe_dirs %>% 
  tail() %>% 
  md_code() %>% 
  md_bullet()
```

Then we can navigate to each of these directories and look for file names
containing "ScheduleA" or "ScheduleB" (these file names are somewhat 
inconsistent).

We also want all of the `Report.csv` files for each month, these files contain
the information on the reports filed by the committees which contains the
contributions.

```{r sbe_names_get}
sbe_names <- character()
for (dir in sbe_dirs) {
  sbe_names <- sbe_base %>% 
    str_c(dir) %>% 
    read_html() %>% 
    html_nodes("a") %>% 
    html_attr("href") %>% 
    str_subset("(Schedule(A|B))|Report") %>% 
    append(sbe_names)
}
```

```{r sbe_paths_list, echo=FALSE}
sbe_names %>% 
  tail() %>% 
  md_code() %>% 
  md_bullet()
```

Using these server paths, we can build local paths to save each file to.

```{r raw_paths_build}
raw_urls <- str_c(sbe_base, sbe_names)
raw_names <- basename(str_replace(raw_urls, "/(?=[^/]*$)", "_"))
raw_paths <- path(raw_dir, raw_names)
```

```{r raw_paths_list, echo=FALSE}
raw_paths %>%
  tail() %>% 
  str_remove("(.*)(?=/va)") %>% 
  md_code() %>% 
  md_bullet()
```

If these paths don't already exist, we can download them now. Each file needs to
be read so that erroneous double quotations can be removed using regular 
expressions.

```{r raw_download}
for (i in seq_along(raw_urls)) {
  if (file_exists(raw_paths[i])) {
    next("file already downloaded")
  } else {
    download.file(raw_urls[i], raw_paths[i])
    x <- read_lines(raw_paths[i])
    if (str_starts(x[1], "#")) {
      next("file already fixed")
    } else {
      x <- str_replace_all(x, "(?<!^|,|\r\n)\"(?!,|\r\n|$)", "'")
      x <- c("### file fixed", x)
      write_lines(x, raw_paths[i])
    }
  }
}
```

## Read

We can now read all of the schedule A and B files into a single dataframe.

```{r sched_paths}
sched_paths <- str_subset(raw_paths, "Schedule")
```

```{r raw_read}
vac <- map_df(
  .x = sched_paths,
  .f = read_delim,
  .id = "source_file",
  delim = ",",
  comment = "###",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    IsIndividual = col_logical(),
    TransactionDate = col_character(),
    Amount = col_double(),
    TotalToDate = col_double()
  )
)
```

We can also identify the source file for each record.

```{r raw_source}
vac$source_file <- basename(sched_paths[as.integer(vac$source_file)])
vac <- relocate(vac, source_file, .after = last_col())
```

```{r raw_rename, echo=FALSE}
vac <- vac %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  rename(
    rpt_id = report_id,
    last_name = last_or_company_name,
    fil_id = committee_contact_id,
    emp_name = name_of_employer,
    occupation = occupation_or_type_of_business,
    emp_place = primary_city_and_state_of_employment_or_business,
    state = state_code,
    zip = zip_code,
    addr1 = address_line1,
    addr2 = address_line2,
    date = transaction_date,
    total = total_to_date,
    sched_a = schedule_a_id,
    sched_b = schedule_b_id,
    valuation = valuation_basis,
    product = product_or_service
  ) %>% 
  select(
    -report_uid
  )
```

The `date` column is a mix of `%m/%d/%Y` and `%Y-%m-%d` formats.

```{r raw_dates}
vac_mdy <- str_which(vac$date, "^\\d{2}/")
vac$date[vac_mdy] <- as.character(mdy(vac$date[vac_mdy]))
vac$date <- parse_datetime(vac$date)
```

Each contribution has a unique ID number for either schedule A or B. We can
create a single shared ID column.

```{r rpt_id}
vac <- vac %>% 
  mutate(con_id = coalesce(sched_a, sched_b)) %>% 
  select(-sched_a, -sched_b)
```

## Join

Each contribution is reported by the committee to which is was made. The
information on those reports, including who filed them and when, is contained
in a separate file every month.

```{r rpt_paths}
rpt_paths <- str_subset(raw_paths, "Report")
```

```{r rpt_read}
rpt <- map_df(
  .x = rpt_paths,
  .f = read_delim,
  .id = "source_file",
  delim = ",",
  comment = "###",
  col_types = cols(
    .default = col_character(),
    IsStateWide = col_logical(),
    IsGeneralAssembly = col_logical(),
    IsLocal = col_logical(),
    FilingDate = col_datetime(),
    StartDate = col_datetime(),
    EndDate = col_datetime(),
    IsFinalReport = col_logical(),
    AmendmentCount = col_integer(),
    ElectionCycleStartDate = col_datetime(),
    ElectionCycleEndDate = col_datetime(),
    BalanceLastReportingPeriod = col_double(),
    NoActivity = col_logical(),
    DateOfReferendum = col_datetime(),
    SubmittedDate = col_datetime(),
    DueDate = col_datetime(),
    IsXmlUpload = col_logical()
  )
)
```

```{r rpt_source}
rpt$source_file <- basename(rpt_paths[as.integer(rpt$source_file)])
rpt <- relocate(rpt, source_file, .after = last_col())
```

```{r rpt_rename, echo=FALSE}
rpt <- rpt %>% 
  clean_names("snake") %>% 
  rename_all(str_replace, "committee_", "fil_") %>% 
  rename_all(str_replace, "submitter_", "fil_") %>% 
  remove_empty("cols") %>% 
  remove_constant(na.rm = TRUE) %>% 
  rename(
    rpt_id = report_id,
    is_stwide = is_state_wide,
    is_assembly = is_general_assembly,
    cand_name = candidate_name,
    is_final = is_final_report,
    addr1 = address_line1,
    addr2 = address_line2,
    state = state_code,
    zip = zip_code,
    is_amend = is_amendment,
    amend_count = amendment_count,
    office = office_sought,
    elect_start = election_cycle_start_date,
    elect_end = election_cycle_end_date,
    balance = balance_last_reporting_period,
    referendum_date = date_of_referendum
  ) %>% 
  select(
    -report_uid,
    -is_xml_upload,
    -balance
  )
```

The information on these reports can then be joined using the report ID.

```{r rpt_join}
vac <- left_join(
  x = vac, y = rpt,
  by = "rpt_id",
  suffix = c("_con", "_fil")
)
```

We identify the columns common to both by either being from a contributor
(`con_*`) or the filer (`fil_*`).

```{r rpt_prefix}
vac <- rename_prefix(
  df = vac,
  suffix = c("_con", "_fil")
)
```

## Explore

There are `r comma(nrow(vac))` columns and `r ncol(vac)` columns.

```{r glimpse}
glimpse(vac)
tail(vac)
```

## Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(vac, count_na)
```

We can flag any record missing a key variable, like a date or name, needed to
identify a transaction and all of it's parties.

```{r na_flag}
key_vars <- c("con_name", "date", "amount", "fil_name")
vac <- vac %>% 
  unite(
    col = con_name,
    prefix, first_name, middle_name, last_name, suffix,
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  mutate(across(where(is_character), na_if, "")) %>% 
  relocate(con_name, .after = last_col()) %>% 
  flag_na(all_of(key_vars))
```

There are `r comma(sum(vac$na_flag))` records missing a key variable.

```{r na_sum}
sum(vac$na_flag)
vac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can flag any record duplicated at least once, aside from the supposedly
unique contribution ID.

```{r dupe_find}
dupe_file <- path(dirname(raw_dir), "dupes.csv")
if (file_exists(dupe_file)) {
  dupes <- read_csv(
    file = dupe_file,
    col_types = cols(
      rpt_id = col_character(),
      con_id = col_character()
    )
  )
  vac <- left_join(vac, dupes)
} else {
  d1 <- duplicated(select(vac, -con_id), fromLast = FALSE)
  flush_memory()
  d2 <- duplicated(select(vac, -con_id), fromLast = TRUE)
  flush_memory()
  vac <- mutate(vac, dupe_flag = d1 | d2)
  rm(d1, d2)
}
```

```{r dupe_count}
percent(mean(vac$dupe_flag), 0.01)
```

```{r dupe_view}
vac %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(date, con_name)
```

### Categorical

```{r n_distinct}
col_stats(vac, n_distinct)
```

```{r echo=FALSE}
explore_plot(vac, is_individual)
explore_plot(vac, fil_type) + scale_x_truncate()
explore_plot(vac, party, nbar = 4)
explore_plot(vac, office) + scale_x_truncate()
```

### Amounts

```{r ammount_summary}
summary(vac$amount)
mean(vac$amount <= 0)
```

```{r hist_amount, echo=FALSE}
vac %>%
  filter(amount >= 1, !is.na(amount)) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Virginia Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: VA SBE",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
vac <- mutate(vac, year = year(date))
```

```{r date_range}
min(vac$date)
sum(vac$year < 2000)
max(vac$date)
sum(vac$date > today())
```

```{r bar_year, echo=FALSE}
vac %>% 
  count(year) %>% 
  mutate(even = !is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2009, 2020, by = 1)) +
  coord_cartesian(xlim = c(2009, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Virginia Contributions by Year",
    caption = "Source: VA SBE",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

There are `r comma(sum(vac$year == 2009))` records with a `year` value of 2009,
which is before the lowest date of 2012 we would expect.

All of these records came from the same two reports made in January and April
of 2014. An exact same number of records from each report.

```{r year_source}
vac %>% 
  filter(year == 2009) %>% 
  count(rpt_id, con_source_file, fil_source_file, sort = TRUE)
```

These transactions were made in 2009 and weren't reported until 2014.

```{r year_filed}
vac %>% 
  filter(year == 2009) %>% 
  count(date, filing_date, sort = TRUE)
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `*_addr` variables, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

We will create separate tables of unique, normalized addresses.

```{r con_addr_full}
con_addr_norm <- vac %>% 
  select(starts_with("con_addr")) %>% 
  distinct() %>% 
  unite(
    col = con_addr_full,
    everything(),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    con_addr_norm = normal_address(
      address = con_addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-con_addr_full)
```

```{r con_addr_view, echo=FALSE}
con_addr_norm
```

This table can be joined to the original contributions.

```{r con_addr_join}
vac <- left_join(vac, con_addr_norm)
```

This process is repeated for filer addresses, of which there are far fewer.

```{r fil_addr_norm}
fil_addr_norm <- vac %>% 
  select(starts_with("fil_addr")) %>% 
  distinct() %>% 
  unite(
    col = fil_addr_full,
    everything(),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    fil_addr_norm = normal_address(
      address = fil_addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-fil_addr_full)
```

```{r fil_addr_join}
vac <- left_join(vac, fil_addr_norm)
```

```{r address_view}
vac %>% 
  select(contains("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
vac <- mutate(
  vac, across(
    .cols = ends_with("zip"),
    .fns = list(norm = normal_zip),
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  vac$con_zip,
  vac$con_zip_norm,
  vac$fil_zip,
  vac$fil_zip_norm,
  compare = valid_zip
)
```

### State

The state variables do not need to be normalized.

```{r state_check}
prop_in(vac$con_state, valid_state)
prop_in(vac$fil_state, valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
vac <- mutate(
  vac, across(
    .cols = ends_with("city"),
    .fns = list(norm = normal_city),
    abbs = usps_city,
    states = c("VA", "DC", "VIRGINIA"),
    na = invalid_city,
    na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r con_city_swap}
con_city_swap <- vac %>% 
  count(
    con_city_norm, 
    con_state, 
    con_zip_norm,
    sort = TRUE
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "con_state" = "state",
      "con_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(con_city_norm, city_match),
    match_dist = str_dist(con_city_norm, city_match),
    con_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = city_match,
      false = con_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb,
    -n
  )
```

```{r echo=FALSE}
con_city_swap %>% 
  filter(con_city_norm != con_city_swap)
```

```{r con_city_join}
vac <- left_join(vac, con_city_swap)
```

Repeat the process for the filer cities.

```{r fil_city_swap}
fil_city_swap <- vac %>% 
  count(
    fil_city_norm, 
    fil_state, 
    fil_zip_norm,
    sort = TRUE
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "fil_state" = "state",
      "fil_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(fil_city_norm, city_match),
    match_dist = str_dist(fil_city_norm, city_match),
    fil_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = city_match,
      false = fil_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb,
    -n
  )
```

```{r echo=FALSE}
fil_city_swap %>% 
  filter(fil_city_norm != fil_city_swap)
```

```{r fil_city_join}
vac <- left_join(vac, fil_city_swap)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- vac %>% 
  mutate(
    con_city_refine = con_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(con_city_refine != con_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "con_city_refine" = "city",
      "con_state" = "state",
      "con_zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
nrow(good_refine)
good_refine %>%
  count(
    con_state, 
    con_zip_norm, 
    con_city_swap, 
    con_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
vac <- vac %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(con_city_refine = coalesce(con_city_refine, con_city_swap))
```

#### Progress

```{r con_city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(vac$con_city),
  vac$con_city_norm,
  vac$con_city_swap,
  vac$con_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

```{r fil_city_progress, echo=FALSE}
progress_table(
  str_to_upper(vac$fil_city),
  vac$fil_city_norm,
  vac$fil_city_swap,
  compare = many_city
) %>% 
  mutate(stage = as_factor(stage)) %>% 
  kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(vac$con_city, many_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Virginia City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Virginia City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
vac <- vac %>% 
  select(
    -con_city_norm,
    -con_city_swap,
    con_city_clean = con_city_refine
  ) %>% 
    select(
    -fil_city_norm,
    fil_city_clean = fil_city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(vac, 100))
```

1. There are `r comma(nrow(vac))` records in the database.
1. There are `r comma(sum(vac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(vac$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("va", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "va_contribs_clean.csv")
write_csv(vac, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
s3_head <- head_object(s3_path, "publicaccountability")
(s3_size <- as_fs_bytes(attr(s3_head, "content-length")))
unname(s3_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(vac)),
  type = md_code(map_chr(vac, typeof)),
  def = c(
    "Filed report ID number",
    "Unique Filer ID",
    "Contributor first name",
    "Contributor middle name",
    "Contributor last name",
    "Contributor name prefix",
    "Contributor name suffix",
    "Contributor employer name",
    "Contributor occipation",
    "Employer location",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor state abbreviation",
    "Contributor ZIP+4 code",
    "Flag for individual contributor",
    "Date contribution was made",
    "Amount of contribution or correction",
    "Total amount contributed to date",
    "In-kind contribution valuation method",
    "In-kind contribution product",
    "Contribution source file name",
    "**Unique** contribution ID",
    "**Unique** filer code",
    "Filing committee name",
    "Filing committee type",
    "Candidate name",
    "Flag for statewide race",
    "Flag for general assembly race",
    "Flag for local race",
    "Candidate political party",
    "Reporting year",
    "Date report filed",
    "Date reporting period began",
    "Date reporting period ended",
    "Committee street address",
    "Committee secondary address",
    "Committee city name",
    "Committee state abbreviation",
    "Committee ZIP+4 code",
    "Type of report filed",
    "Flag for final election report",
    "Flag for amended report",
    "Amended report version count",
    "Filer phone number",
    "Filer email address",
    "Election cycle date/month",
    "Date election began",
    "Date election ended",
    "Office sought by receiving candidate",
    "District in which election held",
    "Flag for no activity reported",
    "Date referendum held",
    "Date report filed",
    "Unique account ID hash of filer",
    "Date report due with state",
    "Report source file name",
    "Combined full contributor name",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized contributor combined address",
    "Normalized committee combined address",
    "Normalized contributor 5-digit ZIP code",
    "Normalized committee 5-digit ZIP code",
    "Normalized contributor city name",
    "Normalized committee city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
