---
title: "Washington Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("wa", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  jsonlite, # parse json data
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("wa/contribs/docs/wa_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

```{r raw_meta}
wa_meta <- fromJSON("https://data.wa.gov/api/views/kv7h-kjye")
```

```{r echo=FALSE}
wa_create <- as_datetime(wa_meta$createdAt)
wa_update <- as_datetime(wa_meta$rowsUpdatedAt)
```

[Contributions][con] for the state of Washington can be obtained from the 
state's [Public Disclosure Commission][pdc] on their [Open Data portal][odp].
The data is titled "Contributions to Candidates and Political Committees" and
was created on `r format(wa_create, "%B %d, %Y")` and was last updated at
`r format(wa_update, "%B %d, %Y")`.

#### Description

> This dataset contains cash and in-kind contributions, (including unpaid loans)
made to Washington State Candidates and Political Committees for the last 10
years as reported to the PDC on forms C3, C4, Schedule C and their electronic
filing equivalents. It does not include loans which have been paid or forgiven,
pledges or any expenditures.
>
> For candidates, the number of years is determined by the year of the election,
not necessarily the year the contribution was reported. For political
committees, the number of years is determined by the calendar year of the
reporting period.
>
> Candidates and political committees choosing to file under "mini reporting"
are not included in this dataset. See WAC 390-16-105 for information regarding
eligibility.
>
> This dataset is a best-effort by the PDC to provide a complete set of records
as described herewith and may contain incomplete or incorrect information. The
PDC provides access to the original reports for the purpose of record
verification.
>
> Descriptions attached to this dataset do not constitute legal definitions;
please consult RCW 42.17A and WAC Title 390 for legal definitions and additional
information political finance disclosure requirements.
>
> CONDITION OF RELEASE: This publication constitutes a list of individuals
prepared by the Washington State Public Disclosure Commission and may not be
used for commercial purposes. This list is provided on the condition and with
the understanding that the persons receiving it agree to this statutorily
imposed limitation on its use. See RCW 42.56.070(9) and AGO 1975 No. 15.

[con]: https://data.wa.gov/d/kv7h-kjye
[pdc]: http://pdc.wa.gov/
[odp]: https://data.wa.gov/

#### Dictionary

```{r echo=FALSE}
wa_cols <- wa_meta$columns %>% 
  select(fieldName, dataTypeName, description)
wa_cols %>% 
  mutate(across(description, str_replace_all, "[:blank:]", " ")) %>% 
  mutate(across(description, str_replace_all, "\n", "<br/>")) %>% 
  mutate(across(fieldName, md_code)) %>% 
  kable()
```

## Download

```{r raw_dir}
raw_dir <- dir_create(here("wa", "contribs", "data", "raw"))
raw_tsv <- path(raw_dir, path_ext_set(wa_meta$resourceName, "tsv"))
```

```{r raw_download}
if (!file_exists(raw_tsv)) {
  wa_head <- GET(
    url = "https://data.wa.gov/api/views/kv7h-kjye/rows.tsv",
    query = list(accessType = "DOWNLOAD"),
    write_disk(path = raw_tsv),
    progress("down")
  )
}
```

## Read

```{r raw_read}
wac <- read_delim(
  file = raw_tsv,
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    id = col_integer(),
    election_year = col_integer(),
    amount = col_double(),
    receipt_date = col_date("%m/%d/%Y")
  )
)
```

## Explore

There are `r comma(nrow(wac))` rows of `r ncol(wac)` columns. Each record
represents a single `r str_to_lower(wa_meta$metadata$rowLabel)`.

```{r glimpse}
glimpse(wac)
tail(wac)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(wac, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("receipt_date", "contributor_name", "amount", "filer_name")
wac <- flag_na(wac, all_of(key_vars))
sum(wac$na_flag)
```

```{r na_view}
wac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any entirely duplicate rows. To keep memory usage low with such
a large data frame, we will split our data into a list and check each element of
the list. For each chunk, we will write the duplicate `id` to a text file.

```{r}
prop_distinct(wac$id)
```

```{r dupe_write}
dupe_file <- here("il", "contribs", "data", "dupe_ids.txt")
if (!file_exists(dupe_file)) {
  tmp <- file_temp(ext = "rds")
  write_rds(wac, file = tmp)
  file_size(tmp)
  wa_id <- split(wac$id, wac$receipt_date)
  was <- wac %>%
    select(-id) %>% 
    group_split(receipt_date)
  if (file_exists(tmp)) {
    rm(wac)
    Sys.sleep(5)
    flush_memory(2)
  }
  pb <- txtProgressBar(max = length(was), style = 3)
  for (i in seq_along(was)) {
    if (nrow(was[[i]]) < 2) {
      next
    }
    d1 <- duplicated(was[[i]], fromLast = FALSE)
    d2 <- duplicated(was[[i]], fromLast = TRUE)
    dupe_vec <- d1 | d2
    rm(d1, d2)
    if (any(dupe_vec)) {
      write_lines(
        x = wa_id[[i]][dupe_vec], 
        file = dupe_file, 
        append = file_exists(dupe_file),
        na = ""
      )
    }
    rm(dupe_vec)
    was[[i]] <- NA
    wa_id[[i]] <- NA
    if (i %% 100 == 0) {
      Sys.sleep(2)
      flush_memory()
    }
    setTxtProgressBar(pb, i)
  }
  rm(was, wa_id)
  Sys.sleep(5)
  flush_memory()
  wac <- read_rds(tmp)
}
```

```{r dupe_add}
dupe_id <- tibble(
  id = as.integer(read_lines(dupe_file, skip_empty_rows = TRUE)),
  dupe_flag = TRUE
)
wac <- left_join(wac, dupe_id, by = "id")
wac <- mutate(wac, across(dupe_flag, Negate(is.na)))
```

`r percent(mean(wac$dupe_flag), 0.1)` of rows are duplicates.

```{r dupe_view}
wac %>% 
  filter(dupe_flag) %>% 
  count(receipt_date, contributor_name, amount, filer_name, sort = TRUE)
```

### Categorical

```{r distinct_count}
col_stats(wac, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(wac, type)
```

### Amounts

```{r amount_summary}
summary(wac$amount)
mean(wac$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(wac[c(which.max(wac$amount), which.min(wac$amount)), ])
```

```{r hist_amount, echo=FALSE}
wac %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Washington Contributions Amount Distribution",
    caption = "Source: WA PRC",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `receipt_date` with `lubridate::year()`

```{r date_year}
wac <- mutate(wac, receipt_year = year(receipt_date))
```

```{r date_range}
min(wac$receipt_date)
sum(wac$receipt_year < 2000)
max(wac$receipt_date)
sum(wac$receipt_date > today())
```

```{r bar_year, echo=FALSE}
wac %>%
  count(receipt_year) %>% 
  filter(receipt_year >= 2004, receipt_year <= 2021) %>% 
  mutate(even = is_even(receipt_year)) %>% 
  ggplot(aes(x = receipt_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Washington Contributions by Year",
    caption = "Source: WA PRC",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

The `contributor_address` variable is already sufficiently normalized.

```{r}
sample(wac$contributor_address, 5)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
wac <- wac %>% 
  mutate(
    zip_norm = normal_zip(
      zip = contributor_zip %>% 
        str_remove_all("-\\d{4}$"),
      na_rep = TRUE
    )
  )
```

```{r}
wac %>% 
  count(contributor_zip, zip_norm, sort = TRUE) %>% 
  filter(contributor_zip != zip_norm)
```

```{r zip_progress}
progress_table(
  wac$contributor_zip,
  wac$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
wac <- wac %>% 
  mutate(
    state_norm = normal_state(
      state = contributor_state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
wac %>% 
  count(contributor_state, state_norm, sort = TRUE) %>% 
  filter(contributor_state != state_norm)
```

```{r state_progress}
progress_table(
  wac$contributor_state,
  wac$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- wac %>% 
  distinct(contributor_city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = contributor_city, 
      abbs = usps_city,
      states = c("WA", "DC", "WASHINGTON"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
wac <- left_join(
  x = wac,
  y = norm_city,
  by = c(
    "contributor_city",
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- wac %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
wac <- wac %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(wac$contributor_city),
  wac$city_norm,
  wac$city_swap,
  wac$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(1, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(wac$contributor_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Washington City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Washington City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
wac <- wac %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(city_clean, state_clean, .before = zip_clean)
```

```{r clean_glimpse}
glimpse(sample_n(wac, 50))
```

## Conclude

1. There are `r comma(nrow(wac))` records in the database.
1. There are `r comma(sum(wac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wac$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("wa", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "wa_contribs_20040101-20210922.csv")
write_csv(wac, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
