---
title: "Wyoming Stimulus"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("wy", "stimulus", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  jsonlite, # remote browser
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

> The Wyoming Legislature created **three grant programs** during their May
special session to distribute $325 million in federal CARES Act funding to
Wyoming small-business owners who have experienced hardship related to COVID-19.
These grant programs are administered by the Wyoming Business Council. Below is
information about each grant award, including the name of the business awarded a
grant, grant amounts, program identification and date the award was distributed.
For additional information about grant program specifics, including eligibility,
please click [here](https://wyomingbusiness.org/wyobizrelief).

As the Wyoming Economic Development Agency explains:

> The Wyoming Legislature crafted three grant programs during a special session
May 15-16 to distribute $325 million in federal CARES Act funding for Wyoming
small-business owners who have experienced hardship related to COVID-19.
Governor Mark Gordon signed the bill on May 20.
>
> The Wyoming Business Council is working to stand up the three relief programs
as quickly as possible while ensuring the application process is secure,
accessible and easy to use.
>
> Applications for the first grant program -- for businesses with 50 or fewer
employees -- opened June 8.

The Wyoming Business Interruption Stipend is the primary program in the data:

> This $50 million program provides grants up to $50,000. Applicants must be
headquartered in Wyoming, employ 50 or fewer people and have established their
business before any public health orders were issued in Wyoming. Businesses must
show costs of business interruption due to closures related to federal, state or
local orders. Visit our FAQ for more information.

## Read

```{r raw_get}
wys <- type_convert(
  df = fromJSON("http://www.wyopen.gov/wbc/json"),
  col_types = cols(
    .default = col_character(),
    doccreatedt = col_date(),
    lineamount = col_double()
  )
)
```

```{r raw_tweak}
wys <- mutate(
  .data = as_tibble(wys), 
  across(where(is.character), str_to_upper)
)
```

## Explore

```{r glimpse}
glimpse(wys)
tail(wys)
```

### Missing

There are no missing values.

```{r na_count}
col_stats(wys, count_na)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag}
sum(duplicated(wys) | duplicated(wys, fromLast = TRUE))
```

### Categorical

```{r distinct_count}
col_stats(wys, n_distinct)
```

### Amounts

```{r ammount_summary}
summary(wys$lineamount)
mean(wys$lineamount <= 0)
```

Most borrowers took the maximum $50,000 loan amount.

```{r hist_amount, echo=FALSE}
wys %>%
  ggplot(aes(lineamount)) +
  geom_histogram(fill = dark2["purple"], bins = 10) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar, breaks = seq(0, 5e4, by = 1e4)) +
  labs(
    title = "Wyoming Stimulus Amount Distribution",
    caption = "Source: WyOpen",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`. However, all
loans were created in 2020.

```{r date_year}
wys <- mutate(wys, year = year(doccreatedt))
```

```{r date_range}
min(wys$doccreatedt)
sum(wys$year < 2000)
max(wys$doccreatedt)
sum(wys$doccreatedt > today())
```

## Wrangle

The existing geographic variables do not need to be cleaned.

```{r geo_check}
prop_in(wys$state, valid_state) %>% percent(0.01)
prop_in(wys$city, valid_city) %>% percent(0.01)
```

We should add the disbursing government state abbreviation.

```{r geo_add}
wys <- mutate(wys, govt = "WY", .before = agency)
```

## Conclude

1. There are `r comma(nrow(wys))` records in the database.
1. There are `r comma(sum(wys$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wys$na_flag))` records missing key variables.
1. Geographic variables are clean, a source government was added.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("wy", "stimulus", "data", "clean"))
clean_path <- path(clean_dir, "wy_stimulus_clean.csv")
write_csv(wys, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wys)),
  type = md_code(map_chr(wys, typeof)),
  def = c(
    "Unique loan ID",
    "Disbursing government",
    "Disbursing agency name",
    "Borowing vendor name",
    "Fund source (CARES Act)",
    "Disburing program (BIS)",
    "Date loan document created",
    "Vendor state",
    "Loan amount",
    "Vendor city",
    "Calendar year"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
