---
title: "Hawaii Expenditures"
author: "Kiernan Nicholls, Yanqi Xu, and Aarushi Sahejpal"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  tidytext, #unnest tokens
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  aws.s3, # aws cloud storage
  scales, # format strings
  rvest, # read html pages
  RSocrata, # read SODA API
  knitr, # knit documents
  stringdist, # token distance
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_campfin}
pacman::p_load_gh("irworkshop/campfin")
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is provided by the Hawaii Campaign Spending Committee's Socrata open data portal. From the
[Primer metadata page][03], we can find more information. The Records Reporting System RSN is 38832.
The file was created on December 12, 2019 and last updated February 8, 2023. This file represents expenditures made by Hawaii state and county candidates From November 8, 2006 Through December 31, 2022.

[03]: https://opendata.hawaii.gov/dataset/expenditures-made-by-hawaii-state-and-county-candidates

## Import

The data can be directly read using `RSocrata::read.socrata()`.

```{r raw dir}
raw_dir <- dir_create(here("state","hi", "expends", "data", "raw"))
```


```{r download, eval=FALSE}
# hie <- as_tibble(read.socrata("https://data.hawaii.gov/resource/3maa-4fgr.json"))
raw_url <- "http://hicscdata.hawaii.gov/api/views/3maa-4fgr/rows.csv?accessType=DOWNLOAD"
raw_path <- path(raw_dir,"hi_exp_download.csv")

if (this_file_new(raw_path)) {
  download.file(raw_url, raw_path, method = "libcurl")
}
```

Then we can make some general changes to the structure and format of the data frame.

```{r read}
hie <- read_csv(dir_ls(raw_dir)) %>% clean_names()

hie <- hie %>% 
  mutate_if(
    is_character, 
    str_to_upper
  ) %>% 
  mutate(
    amount = as.double(amount)
  ) 
```

## Explore

```{r glimpse}
head(hie)
tail(hie)
glimpse(hie)
```

### Distinct

```{r n_distinct}
col_stats(hie, n_distinct)
```

We can use `campfin::explore_plot()` and/or `ggplot2::geom_bar()` to explore the distribution of
distinct categorical variables.

```{r vendor_bar, echo=FALSE}
explore_plot(
  data = hie,
  var = vendor_type,
) + labs(  title = "HI Expend Vendor Types",
  caption = "Source: HI CRS")
```

```{r category_bar, echo=FALSE, fig.height=10}
hie %>%
  count(expenditure_category, sort = TRUE) %>%
  mutate(p = n/sum(n)) %>%
  ggplot(aes(reorder(expenditure_category, desc(p)), p)) +
  geom_col(aes(fill = expenditure_category)) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "HI Expend Types",
    caption = "Source: HI CRS",
    x = "Expenditure Type",
    y = "Percent"
  )
```

```{r office_bar, echo=FALSE}
hie %>%
  count(office, sort = TRUE) %>%
  mutate(p = n/sum(n)) %>%
  ggplot(aes(reorder(office, desc(p)), p)) +
  geom_col(aes(fill = office)) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "HI Expend Office",
    caption = "Source: HI CRS",
    x = "Office Sought",
    y = "Percent"
  )
```


```{r state_bar, echo=FALSE}
explore_plot(
  data = hie,
  var = in_out_state
) + labs( title = "HI Expends In State",
  caption = "Source: HI CRS")
```

```{r county_bar, echo=FALSE}
explore_plot(
  data = filter(hie, !is.na(county)),
  var = county
) + labs(  title = "HI Expends County",
  caption = "Source: HI CRS")
```

```{r authorized_bar, echo=FALSE}
explore_plot(
  data = filter(hie, !is.na(authorized_use)),
  var = authorized_use
) + coord_flip() +
  labs(
  title = "HI Expends Use",
  caption = "Source: HI CRS")
```

```{r words_bar, echo=FALSE, fig.height=10}
hie %>% 
  unnest_tokens(word, purpose_of_expenditure) %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  head(30) %>% 
  ggplot(
    mapping = aes(
      x = reorder(word, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "HI Expend Purpose",
    caption = "Source: HI CRS",
    x = "Word",
    y = "Percent"
  )
```

### Duplicates

We can use `janitor::get_dupes()` to create a separate table of duplicate rows, then flag those
rows on the original data frame.

```{r get_dupes, collapse=TRUE}
hie <- flag_dupes(hie, everything())
tabyl(hie$dupe_flag)
```

### Missing

We flag the fields that miss the "city" field. 

```{r count_na}
col_stats(hie, count_na)

key_vars <- c("city","party")

hie <- flag_na(hie, all_of(key_vars))
tabyl(hie$na_flag)
```

### Ranges

#### Amount

```{r summary_amount, collapse=TRUE}
summary(hie$amount)
sum(hie$amount < 0)
```

```{r amount_hist, echo=FALSE}
hie %>% 
  ggplot(aes(amount)) +
  geom_histogram() +
  scale_x_continuous(
    trans = "log10",
    labels = dollar
  ) +
  labs(
    title = "HI Expend Amount",
    caption = "Source: HI CRS",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_hist_party, echo=FALSE}
hie %>% 
  filter(party %in% c("DEMOCRAT", "NON-PARTISAN","INDEPENDENT","REPUBLICAN")) %>% 
  ggplot(aes(amount)) +
  geom_histogram(aes(fill = party)) +
  scale_x_continuous(
    trans = "log10",
    labels = dollar
  ) +
  scale_fill_manual(
    guide = FALSE,
    values = c(
      "dodgerblue",
      "mediumpurple",
      "grey10",
      "red2"
    )
  ) +
  labs(
    title = "HI Expend Amount",
    caption = "Source: HI CRS",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~party)
```

```{r amount_box_party, echo=FALSE}
hie %>% 
  filter(party %in% c("DEMOCRAT", "NON-PARTISAN","INDEPENDENT","REPUBLICAN")) %>% 
  ggplot(
    mapping = aes(
      x = party,
      y = amount
    )
  ) +
  geom_boxplot(
    mapping = aes(fill = party),
    varwidth = TRUE,
    outlier.alpha = 0.01
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar
  ) +
  scale_fill_manual(
    guide = FALSE,
    values = c(
      "dodgerblue",
      "mediumpurple",
      "red2",
      "grey10"
    )
  ) +
  labs(
    title = "HI Expend Amount",
    caption = "Source: HI CRS",
    x = "Amount",
    y = "Count"
  )
```


```{r amount_cat_box, echo=FALSE}
top_category <- hie %>% 
  count(expenditure_category, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  filter(p > 0.05) %>% 
  pull(expenditure_category)

hie %>% 
  filter(expenditure_category %in% top_category) %>% 
  ggplot(
    mapping = aes(
      x = reorder(expenditure_category, amount),
      y = amount
    )
  ) +
  geom_boxplot(
    mapping = aes(fill = expenditure_category),
    varwidth = TRUE,
    outlier.alpha = 0.01
  ) +
  scale_fill_brewer(
    type = "qual",
    palette = "Set1",
    guide = FALSE
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar
  ) +
  coord_flip() +
  labs(
    title = "HI Expend Amount",
    caption = "Source: HI CRS",
    x = "Amount",
    y = "Count"
  )
```

#### Date


To better explore the distribution of dates and track expenditures, we will create a `year_clean`
variable from `date` using `lubridate::year()`.

```{r add_year}
hie <- hie %>% mutate(date = as.Date(date, format = "%m/%d/%Y"))
hie <- mutate(hie, year_clean = year(date))
```


```{r date_range, collapse=TRUE}
min(hie$date)
max(hie$date)
```
We can see the expenditures naturally increase in frequency every other year, during the elections.

```{r year_count}
hie %>%
  ggplot(aes(year_clean)) +
  geom_bar() +
  labs(
    title = "HI Expends per Year",
    caption = "Source: HI CRS",
    x = "Year",
    y = "Count"
  )
```

```{r year_amount}
is_even <- function(x) x %% 2 == 0
hie %>% 
  mutate(election_year = is_even(year_clean)) %>% 
  group_by(year_clean, election_year) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(year_clean, mean)) +
  geom_col(aes(fill = election_year)) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "HI Expend Mean Amount per Year",
    caption = "Source: HI CRS",
    fill = "Election Year",
    x = "Amount",
    y = "Mean Amount"
  )
```

```{r}
hie %>% 
  mutate(
    month = month(date),
    election_year = is_even(year_clean)
  ) %>%
  group_by(month, election_year) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(month, mean)) +
  scale_y_continuous(labels = dollar) +
  geom_line(size = 2, aes(color = election_year)) +
  labs(
    title = "HI Expend Mean Amount over Year",
    caption = "Source: HI CRS",
    fill = "Election Year",
    x = "Amount",
    y = "Mean Amount"
  )
```

## Wrangle

To improve the searchability of the database, we can perform some functional data cleaning and
text normalization, using the `campfin::normal_*()` functions, which wrap around `stringr::str_*()`
functions.

### Address

```{r norm_address}
hie <- hie %>% 
  unite(
    col = address_combine,
    address_1, address_2,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_combine,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r view_address}
hie %>% 
  select(
    address_1,
    address_2,
    address_norm
  ) %>% head(10)
```

### ZIP

```{r pre_zip, collapse=TRUE}
n_distinct(hie$zip_code)
prop_in(hie$zip_code, valid_zip)
sum(hie$zip_code %out% valid_zip)
```

```{r normal_zip}
hie <- hie %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip_code,
      na_rep = TRUE
    )
  )
```

```{r post_zip, collapse=TRUE}
n_distinct(hie$zip_norm)
prop_in(hie$zip_norm, valid_zip)
sum(hie$zip_norm %out% valid_zip)
```

### State

`r percent(prop_in(hie$state, valid_state))` of `state` values are valid.

```{r pre_state, collapse=TRUE}
n_distinct(hie$state)
prop_in(hie$state, valid_state)
sum(hie$state %out% valid_state)
```

### City

#### Normal

```{r pre_city, collapse=TRUE}
n_distinct(hie$city)
prop_in(hie$city, valid_city)
sum(unique(hie$city) %out% valid_city)
```

```{r view_city}
hie %>% 
  count(city, sort = TRUE) %>% 
  filter(city %out% valid_city)
```

```{r}
hie <- hie %>% 
  mutate(
    city_norm = normal_city(
      city = city,
      abbs = usps_city,
      states = c("HI", "HAWAII", "DC"),
      na = invalid_city,
      na_rep = TRUE
    )
  )

n_distinct(hie$city_norm)
prop_in(hie$city_norm, valid_city)
sum(unique(hie$city_norm) %out% valid_city)
```

#### Swap

```{r swap_city, collapse=TRUE}
hie <- hie %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "zip_norm" = "zip", 
      "state" = "state"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_dist = stringdist(city_norm, city_match),
    city_swap = if_else(
      condition = match_dist == 1,
      true = city_match,
      false = city_norm
    )
  )

mean(hie$match_dist, na.rm = TRUE)
max(hie$match_dist, na.rm = TRUE)
sum(hie$match_dist == 1, na.rm = TRUE)
n_distinct(hie$city_swap)
prop_in(hie$city_swap, valid_city)
sum(unique(hie$city_swap) %out% valid_city)
```

```{r view_swap}
hie %>% 
  count(state, city_swap, sort = TRUE) %>% 
  filter(city_swap %out% valid_city) %>% 
  drop_na()
```

```{r fix_city}
hie$city_swap <- hie$city_swap %>% 
  str_replace("^HON$", "HONOLULU") %>% 
  na_if("HI")

prop_in(hie$city_swap, valid_city)

hie <- hie %>% rename(city_clean = city_swap)
```


## Conclude

1. There are `r comma(nrow(hie))` records in the database.
1. There are `r sum(hie$dupe_flag)` duplicate records, flagged with `dupe_flag`.
1. Ranges for `amount` and `date` are both reasonable.
1. There are no missing records of importance.
1. Consistency issues in geographic values have been improved.
1. The 5-digit `zip_norm` variable has been created with `campfin::normal_zip(hie$zip_code)`
1. The 4-gitit `year_clean` variable has been created with `lubridate::year(hie$date)`
1. Every record has a payer, payee, date, and amount.

## Export

```{r}
proc_dir <- here("state","hi", "expends", "data", "processed")
dir_create(proc_dir)

hie <- hie %>% 
  select(
    -zip_code,
    -address_1,
    -address_2,
    -address_combine,
    -city_norm,
    -city_match,
    -match_dist
  )
```

## Export

```{r}
# lookup <- read_csv("state/hi/expends/data/hi_city_lookup.csv") %>% select(1:2)
# hie <- left_join(hie, lookup)
progress_table(hie$city_raw,
               hie$city_norm,
               hie$city_clean, compare = valid_city)
write_csv(
  x = hie,
  path = glue("{proc_dir}/hi_expends_20230127.csv"),
  na = ""
)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
