---
title: "Arizona Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("az", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  mdbr, # read mdb files
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("az/contribs/docs/az_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Arizona campaign finance database was obtained by the Investigative 
Reporting Workshop via a state records request pursuant to Ariz. Rev. Stat. Ann.
Secs. 39-121 to 39-122. The request was made via email to Renada Fisher,
Elections Customer Service Manager for the Arizona Secretary of State on January
13, 2021. She supplied a Microsoft Access database file via the secure file
sharing site Serv-U. That file was download and then uploaded to the IRW server,
where it can be downloaded.

```{r raw_dir}
raw_dir <- dir_create(here("az", "contribs", "data", "raw"))
raw_obj <- "20201230-CFS_Export.mdb"
raw_mdb <- path(raw_dir, raw_obj)
raw_aws <- path("csv", raw_obj)
```

```{r raw_download}
if (!file_exists(raw_mdb)) {
  save_object(
    object = raw_aws,
    bucket = "publicaccountability",
    file = raw_mdb,
    overwrite = FALSE,
    show_progress = TRUE
  )
}
```

## Read

The tables in this Microsoft Access database can be converted to flat, comma
delimited text files which can then be read as data frames and joined together
into a single, searchable file. The `mdbr` R package is needed for this 
conversion. That packages uses the [MDB Tools][mdbt] software, which needs to
be installed separately on your system. For example, Debian users can install
MDB Tools from the apt repository from the command line.

[mdbt]: https://github.com/mdbtools/mdbtools

``` bash
sudo apt install mdbtools
```

```{r raw_tables}
az_tables <- mdb_tables(raw_mdb)
md_bullet(md_code(az_tables))
```

The `Transactions` table contains the bulk data, every financial transaction
conducted by campaign committees in the state of Arizona. The other tables
contain the codes needed to identify the parties details of each transaction.

We can read all of these tables as data frames into a list.

```{r raw_read}
az <- map(
  .x = az_tables, 
  .f = read_mdb, 
  file = raw_mdb, 
  na = c("N/A", "", "NA")
)
```

```{r raw_name}
az <- map(az, clean_names, "snake")
names(az) <- make_clean_names(az_tables)
```

## Join

```{r fn_no_time, echo=FALSE}
no_time <- function(x) {
  x <- unique(x)
  if (inherits(x, "POSIXct")) {
    no_hour <- isTRUE(length(na.omit(unique(hour(x)))) == 1)
    no_min <- isTRUE(length(na.omit(unique(minute(x)))) == 1)
    all(no_min, no_hour)
  } else {
    FALSE
  }
}
```

```{r fn_fix_cons, echo=FALSE}
fix_cols <- function(.data) {
  .data %>% 
    remove_constant() %>% 
    remove_empty("cols") %>% 
    rename_with(~str_replace(., "transaction", "txn")) %>% 
    rename_with(~str_replace(., "committee", "cmte")) %>% 
    rename_with(~str_replace(., "candidate", "cand")) %>% 
    rename_with(~str_replace(., "category", "cat")) %>% 
    rename_with(~str_replace(., "report", "rpt")) %>% 
    rename_with(~str_replace(., "name", "nm")) %>% 
    mutate(across(where(no_time), as_date)) %>% 
    mutate(across(where(is_character), na_if, "N/A"))
}
```

```{r fix_map}
az <- map(az, fix_cols)
```

```{r join_cmte_nms}
cmte_nms <- az$names %>% 
  select(nm_id, cmte_nm = last_nm)
```

```{r join_cmtes}
az_cmte <- az$committees %>% 
  # keep address and candidate info
  rename_with(~str_replace(., "physical", "cmte")) %>% 
  select(
    cmte_id, nm_id, starts_with("cmte"), 
    cand_party_id, cand_office_id,
    cand_is_incumbent, cand_cycle_id
  ) %>% 
  rename(cmte_zip = cmte_zip_code) %>% 
  # add office name from id
  left_join(az$offices, by = c("cand_office_id" = "office_id")) %>% 
  select(-cand_office_id, cand_office = office_nm) %>% 
  # add party name from id
  left_join(az$parties, by = c("cand_party_id" = "party_id")) %>% 
  select(-cand_party_id, cand_party = party_nm) %>% 
  # add cycle year from id
  left_join(az$cycles[, 1:2], by = c("cand_cycle_id" = "cycle_id")) %>% 
  select(-cand_cycle_id, cand_cycle = cycle_nm) %>% 
  # add the committee name
  left_join(cmte_nms, by = "nm_id") %>% 
  relocate(cmte_nm, .after = cmte_id) %>% 
  select(-nm_id)
```

```{r txn_income_type}
income_txn_type_id <- az$transaction_types %>% 
  left_join(az$income_expense_neutral) %>% 
  filter(income_expense_neutral == "Income") %>% 
  pull(txn_type_id)
```

```{r join_names}
az$names <- az$names %>% 
  rename(zip = zip_code) %>% 
  left_join(az$entity_types, by = "entity_type_id") %>% 
  select(-nm_group_id, -entity_type_id) %>% 
  rename(entity_type = entity_type_nm)
```

```{r join_all}
azc <- az$transactions %>% 
  filter(txn_type_id %in% income_txn_type_id) %>% 
  select(1, 4:7) %>% 
  left_join(az_cmte, by = "cmte_id") %>% 
  inner_join(az$names, by = "nm_id") %>% 
  relocate(nm_id, .before = last_nm) %>% 
  relocate(cmte_id, .before = cmte_nm)
```

```{r join_rm_gc}
rm(az, az_cmte, income_txn_type_id, cmte_nms)
flush_memory(1)
```

## Explore

There are `r comma(nrow(azc))` rows of `r ncol(azc)` columns. Each record
represents a single contribution made from an entity to a political committee.

```{r glimpse}
glimpse(azc)
tail(azc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(azc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("date", "last", "amount", "committee")
azc <- flag_na(azc, all_of(key_vars))
sum(azc$na_flag)
```

```{r na_view}
azc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
azc <- flag_dupes(azc, -id)
sum(azc$dupe_flag)
```

```{r dupe_view}
azc %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r distinct_count}
col_stats(azc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(azc, type)
```

### Amounts

```{r amount_summary}
summary(azc$amount)
mean(azc$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(azc[c(which.max(azc$amount), which.min(azc$amount)), ])
```

```{r hist_amount, echo=FALSE}
azc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Arizona Contributions Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
azc <- mutate(azc, year = year(date))
```

```{r date_range}
min(azc$date)
sum(azc$year < 2000)
max(azc$date)
sum(azc$date > today())
```

```{r bar_year, echo=FALSE}
azc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
azc <- azc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
azc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
azc <- azc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  azc$zip,
  azc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
azc <- azc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
azc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  azc$state,
  azc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- azc %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("AZ", "DC", "ARIZONA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
azc <- left_join(
  x = azc,
  y = norm_city,
  by = c(
    "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- azc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
azc <- azc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(azc$city_raw),
  azc$city_norm,
  azc$city_swap,
  azc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(azc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
azc <- azc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(azc, 50))
```

1. There are `r comma(nrow(azc))` records in the database.
1. There are `r comma(sum(azc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(azc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("az", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "az_contribs_clean.csv")
write_csv(azc, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(azc)),
  type = md_code(map_chr(azc, typeof)),
  def = c(
    ""
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
