---
title: "Oklahoma Expenditures"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ok", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("ok/expends/docs/ok_expends_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

The [Oklahoma Ethics Commission][oec] provides a [data download page][ddp] where
users can download campaign finance records by year.

> ### Data Download
> 
> This page provides comma separated value (CSV) downloads of contribution,
loan, and expenditure data for each reporting year in a zipped file format.
These files can be downloaded and imported into other applications (Microsoft
Excel, Microsoft Access, etc.)
>
> This data is extracted from the state of Oklahoma database as it existed as of
9/9/2021  12:08 AM

> ### Downloading Contribution and Expenditure Data
> 
> You can access the Campaign Finance Data Download page to download
contribution and expenditure data for import into other applications such as
Microsoft Excel or Access. A weekly batch process is run that captures the
year-to-date information for the current year. The data is available for each
calendar year. The file is downloaded in CSV format.

The OEC also provides a [PDF file layout key][flk].

|   |Field Name     |Description                                            |
|:--|:--------------|:------------------------------------------------------|
|A  |RECEIPT ID     |This is the Receipt internal ID. This ID is unique.    |
|B  |ORG ID         |Unique ID of the receiving candidate or committee.     |
|C  |RECEIPT TYPE   |This is the Receipt Type.                              |
|D  |RECEIPT DATE   |Receipt Date                                           |
|E  |RECEIPT AMOUNT |Receipt Amount                                         |
|F  |DESCRIPTION    |This is the description provided for the receipt.      |
|G  |SOURCE TYPE    |Type of entity that is the source of the Receipt.      |
|H  |FIRST NAME     |Source First Name                                      |
|I  |MIDDLE NAME    |Source Middle Initial or Name if provided.             |
|J  |LAST NAME      |Source Last Name                                       |
|K  |SUFFIX         |Source Name Suffix                                     |
|L  |SPOUSE NAME    |Source Spouse Name                                     |
|M  |ADDRESS 1      |Source , PO Box, or other directional information      |
|N  |ADDRESS 2      |Source Suite/Apartment number                          |
|O  |CITY           |Source City                                            |
|P  |STATE          |Source State                                           |
|Q  |ZIP            |Source  Zip Code                                       |
|R  |FILED DATE     |Receipt Filed Date                                     |
|S  |COMMITTEE TYPE |Indicates Type of receiving committee                  |
|T  |COMMITTEE NAME |This is the name of the receiving committee.           |
|U  |CANDIDATE NAME |This is the name of the receiving candidate            |
|V  |AMENDED        |Y/N indicator to show if an amendment was filed...     |
|W  |EMPLOYER       |Source's employer...                                   |
|X  |OCCUPATION     |The Source's occupation... used for Individual donors. |

[oec]: https://guardian.ok.gov/PublicSite/Homepage.aspx#
[ddp]: https://guardian.ok.gov/PublicSite/DataDownload.aspx
[flk]: https://guardian.ok.gov/PublicSite/Resources/PublicDocuments/OKReceiptsAndTransfersInFileLayout.pdf

## Download

The annual ZIP archives provided by OEC have unique URLs and can be downloaded.

```{r raw_dir}
ok_url <- "https://guardian.ok.gov/PublicSite/Docs/BulkDataDownloads/"
raw_url <- str_c(ok_url, glue("{2014:2022}_ExpenditureExtract.csv.zip"))
raw_dir <- dir_create(here("ok", "contribs", "data", "raw"))
raw_zip <- path(raw_dir, basename(raw_url))
```

```{r echo=FALSE}
md_order(md_code(raw_url))
```

```{r raw_download}
if (!all(file_exists(raw_zip))) {
  download.file(raw_url, raw_zip)
}
```

```{r echo=FALSE}
file_info(raw_zip) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(1, basename))
```

```{r raw_unzip}
raw_csv <- map_chr(raw_zip, unzip, exdir = raw_dir)
```

## Fix

The double-quotes (`"`) in this file are not properly escaped. We can read the
lines of each text file and replace double-quotes in the middle of "columns"
with _two_ double-quotes, which can be properly ignored when reading the data.

```{r fix.csv}
fix_csv <- path_temp(basename(raw_csv))
for (i in seq_along(raw_csv)) {
  message(basename(raw_csv[i]))
  read_lines(raw_csv[i]) %>% 
    # double quote in middle of string
    str_replace_all("(?<!^|,)\"(?!,|$)", r"("")") %>% 
    write_lines(fix_csv[i])
  flush_memory()
}
```

## Read

```{r raw-read}
oke <- read_delim(
  file = fix_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = TRUE,
  locale = locale(date_format = "%m/%d/%Y"),
  col_types = cols(
    .default = col_character(),
    `Expenditure Date` = col_date(),
    `Expenditure Amount` = col_double(),
    `Filed Date` = col_date()
  )
)
```

```{r clean-names}
oke <- clean_names(oke, case = "snake")
```

## Explore

There are `r comma(nrow(oke))` rows of `r ncol(oke)` columns. Each record
represents a single expenditure from a campaign to a vendor.

```{r glimpse}
glimpse(oke)
tail(oke)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(oke, count_na)
```

There are two columns for the recipient name; one for candidates and one for
committees. Neither column is missing any values without a corresponding value
in other column.

```{r}
oke %>% 
  group_by(committee_type) %>% 
  summarise(
    no_cand_name = prop_na(candidate_name),
    no_comm_name = prop_na(committee_name)
  )
```

Of the other key variables, only a few hundred `last_name` values are missing.

```{r key_vars}
key_vars <- c(
  "expenditure_date", "last_name", "expenditure_amount", 
  "candidate_name", "committee_name"
)
```

```{r cand_comm_name}
prop_na(oke$candidate_name[!is.na(oke$committee_name)])
prop_na(oke$committee_name[!is.na(oke$candidate_name)])
```

We can flag any record missing a key variable needed to identify a transaction.

```{r flag_na}
count_na(oke$expenditure_date)
count_na(oke$expenditure_amount)
count_na(oke$last_name)
oke <- mutate(oke, na_flag = is.na(last_name) | is.na(expenditure_amount))
sum(oke$na_flag)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
oke <- flag_dupes(oke, -expenditure_id)
sum(oke$dupe_flag)
mean(oke$dupe_flag)
```

```{r dupe-view}
oke %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(expenditure_date)
```

### Categorical

```{r distinct-count}
col_stats(oke, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(oke, purpose)
explore_plot(oke, committee_type)
explore_plot(oke, amended)
```

### Amounts

```{r amount-round}
# fix floating point precision
oke$expenditure_amount <- round(oke$expenditure_amount, digits = 2)
```

```{r amount-summary}
summary(oke$expenditure_amount)
mean(oke$expenditure_amount <= 0, na.rm = TRUE)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(oke[c(which.max(oke$expenditure_amount), which.min(oke$expenditure_amount)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
oke %>%
  ggplot(aes(expenditure_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Oklahoma Expenditures Amount Distribution",
    caption = "Source: Oklahoma Ethics Commission",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
oke <- mutate(oke, expenditure_year = year(expenditure_date))
```

```{r date-range}
min(oke$expenditure_date)
sum(oke$expenditure_year < 2000)
max(oke$expenditure_date)
sum(oke$expenditure_date > today())
```

It's common to see an increase in the number of expenditures in elections years.

```{r bar-year, echo=FALSE}
oke %>% 
  count(expenditure_year) %>% 
  mutate(even = is_even(expenditure_year)) %>% 
  ggplot(aes(x = expenditure_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oklahoma Expenditures by Year",
    caption = "Source: Oklahoma Ethics Commission",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
addr_norm <- oke %>% 
  distinct(address_1, address_2) %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address-view}
addr_norm
```

```{r address-join}
oke <- left_join(oke, addr_norm, by = c("address_1", "address_2"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
oke <- oke %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  oke$zip,
  oke$zip_norm,
  compare = valid_zip
)
```

### State

The state values to not need to be normalized.

```{r state.check}
prop_in(oke$state, valid_state)
unique(what_out(oke$state, valid_state))
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- oke %>% 
  distinct(city, state, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("OK", "DC", "OKLAHOMA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
oke <- left_join(
  x = oke,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- oke %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
oke <- oke %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(oke$city),
  oke$city_norm,
  oke$city_swap,
  oke$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(oke$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Oklahoma City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oklahoma City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
oke <- oke %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(oke, 1000))
```

1. There are `r comma(nrow(oke))` records in the database.
1. There are `r comma(sum(oke$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(oke$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(oke$expenditure_date), "-")
max_dt <- str_remove_all(max(oke$expenditure_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("ok", "expends", "data", "clean"))
clean_csv <- path(clean_dir, glue("ok_expends_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(oke, clean_csv, na = "")
write_rds(oke, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict-make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(oke)),
  type = md_code(map_chr(oke, typeof)),
  def = c(
    ""
  )
)
```

```{r dict-md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
