---
title: "California Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ca", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Instead of getting California campaign expenditure data from the Secretary of
State's official [Cal-Access database][cala], we will instead be using the clean
version published by the [California Civic Data Coalition (CCDC)][ccdc].

>  The coalition was formed in 2014 by Ben Welsh and Agustin Armendariz to lead
the development of open-source software that makes California's public data
easier to access and analyze. The effort has drawn hundreds of contributions
from developers and journalists at dozens of news organizations.
> 
> Our primary focus is refining CAL-ACCESS, the jumbled, dirty and difficult
government database that tracks campaign finance and lobbying activity in
California politics.

The CCDC provides a [single file][ex] for state campaign expenditures. The page
contains a list of variables found in the data along with their definitions.

```{r raw_dict, echo=FALSE}
cae_dict <- read_csv(here("ca", "expends", "ca_exp_dict.tsv"))
cae_dict$column <- md_code(cae_dict$column)
kable(cae_dict)
```

[cala]: http://cal-access.sos.ca.gov/
[ccdc]: https://www.californiacivicdata.org
[ex]: https://calaccess.californiacivicdata.org/documentation/raw-files/expn-cd/

## Download

We can download the CCDC expenditures file directly, ignoring all the other
data typically included in the Cal-Access database.

```{r raw_download}
raw_dir <- here("ca", "expends", "data", "raw")
raw_source <- "https://s3-us-west-2.amazonaws.com/django-calaccess" 
raw_url <- str_c(raw_source, "latest", "expn_cd.csv", sep = "/")
raw_path <- path(raw_dir, basename(raw_url))
if (!file_exists(raw_path)) {
  download.file(raw_url, raw_path)
}
```

> ### Errata
> The raw data provided by the state contains errors in how some values are
escaped, quoted and delimited. The result is a small number of records in the
latest download were lost when we prepared files for download. However, only 638
records out of 7,219,803 total rows were lost (you can download them [here][e]).

[e]: https://s3-us-west-2.amazonaws.com/django-calaccess/2020-08-26_09-50-41/expn_cd.errors.csv

```{r raw_errata, echo=FALSE}
errata_url <- str_c(
  raw_source, "2020-08-26_09-50-41", 
  "expn_cd.errors.csv", sep = "/"
)
cae_errata <- read_csv(errata_url)
print(cae_errata)
```

## Read

Once downloaded, we can read the expenditures file directly, using the variable
dictionary to define the column types.

```{r raw_read}
# 7,217,483 rows
cae <- vroom(
  file = raw_path,
  delim = ",",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    AMEND_ID = col_integer(),
    AMOUNT = col_double(),
    CUM_OTH = col_double(),
    CUM_YTD = col_double(),
    EXPN_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    FILING_ID = col_integer(),
    LINE_ITEM = col_integer()
  )
)
```

The expenditures data does not include any identifying information on the filers
of the forms, the committees making the expenditures. There is are three
variables for _candidate_ names, but the vast majority are empty.

```{r raw_missing}
prop_na(cae$cand_naml)
```

```{r}
cae <- cae %>% 
  select_if(~prop_na(.) < 0.95) %>% 
  select(-starts_with("cand_")) %>% 
  rename(date = expn_date)
```

To get this information, we can download the cover page data from the 
[campaign-disclosure forms][cvr]. These forms contains the filer names and
addresses. We can also download this file from the CCDC and read the same way.

[cvr]: https://calaccess.californiacivicdata.org/documentation/raw-files/cvr-campaign-disclosure-cd/

```{r fil_download}
cvr_url <- str_c(raw_source, "cvr_campaign_disclosure_cd.csv", sep = "/")
cvr_path <- path(raw_dir, basename(cvr_url))
if (!file_exists(cvr_path)) {
  download.file(cvr_url, cvr_path)
}
```

```{r cvr_read}
cvr <- vroom(
  file = cvr_path,
  delim = ",",
  escape_double = TRUE,
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    AMEND_ID = col_integer(),
    FILING_ID = col_integer(),
    RPT_DATE = col_date("%m/%d/%Y %H:%M:%S %p")
  )
)
```

The CVR data has `r comma(nrow(cvr))` rows and `r ncol(cvr)` columns, only a few
of which are useful to us. We also only want those records identifying those
who have filed expenditure data. We also want to ensure only one version of
filer information for each unique `filing_id`.

```{r cvr_filter}
cvr <- cvr %>% 
  select(filing_id, amend_id, starts_with("filer"), rpt_date) %>% 
  filter(filing_id %in% cae$filing_id) %>% 
  select_if(~prop_na(.) < 0.95)
```

We can then use this new data to identify the parties in the expenditures table
using the unique `filing_id` and `amend_id`. We will remove the useless 

```{r cvr_join}
cae <- left_join(cae, cvr)
```

## Explore

```{r glimpse}
glimpse(cae)
tail(cae)
```

### Missing

Columns vary in their degree of missing values. We can flag any values missing
from key variables used to identify the transaction.

```{r na_count}
col_stats(cae, count_na)
```

```{r na_flag}
cae <- cae %>% 
  # unite names
  unite(
    col = payee_cmbn, 
    starts_with("payee_nam"), 
    sep = " ", 
    remove = FALSE,
    na.rm = TRUE
  ) %>%
    unite(
    col = filer_cmbn, 
    starts_with("filer_nam"), 
    sep = " ", 
    remove = FALSE,
    na.rm = TRUE
  ) %>%
  # use report date for those missing
  mutate(date = coalesce(date, rpt_date)) %>% 
  # flag missing records
  flag_na(date, payee_cmbn, amount, filer_cmbn) %>% 
  # remove combined names
  select(-ends_with("_cmbn"))
```

```{r na_prop}
mean(cae$na_flag)
```

```{r na_view}
cae %>% 
  filter(na_flag) %>% 
  select(date, payee_naml, amount, filer_naml) %>% 
  sample_frac()
```

```{r na_recount}
cae %>% 
  filter(na_flag) %>% 
  select(date, payee_naml, amount, filer_naml) %>% 
  col_stats(count_na)
```

### Duplicates

```{r dupe_write}
dupe_file <- here("ca", "expends", "dupes.txt")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  n <- 1e5 # rows per chunk
  nr <- nrow(cae)
  # split file into chunks
  cas <- split(
    x = arrange(cae, payee_naml), 
    f = rep(x = 1:ceiling(nr/n), each = n, length.out = nr)
  )
  pb <- txtProgressBar(max = length(cas), style = 3)
  for (i in seq_along(cas)) {
    # check dupes from both ends
    d1 <- duplicated(cas[[i]], fromLast = FALSE)
    d2 <- duplicated(cas[[i]], fromLast = TRUE)
    # append to disk
    write_lines(d1 | d2, dupe_file, append = TRUE)
    rm(d1, d2)
    cas[[i]] <- NA
    Sys.sleep(1); flush_memory()
    setTxtProgressBar(pb, value = i)
  }
  rm(cas)
}
```

```{r dupe_read}
cae <- mutate(cae, dupe_flag = as.logical(read_lines(dupe_file)))
```

After all that work, there are `r comma(sum(cae$dupe_flag))` duplicate records

```{r dupe_view}
cae %>% 
  filter(dupe_flag) %>% 
  select(date, payee_naml, amount, filer_naml)
```

### Categorical

```{r distinct_count}
col_stats(cae, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(cae, form_type)
explore_plot(cae, entity_cd)
explore_plot(cae, office_cd)
explore_plot(cae, juris_cd)
```

### Amounts

```{r ammount_summary}
summary(cae$amount) %>% map_chr(dollar)
mean(cae$amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
cae %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "California Expenditures Amount Distribution",
    caption = "Source: californiacivicdata.org",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
cae <- mutate(cae, year = year(date))
```

There are a number of dates with impossibly low or high values.

```{r date_range}
min(cae$date, na.rm = TRUE)
sum(cae$year < 1999, na.rm = TRUE)
max(cae$date, na.rm = TRUE)
sum(cae$date > today(), na.rm = TRUE)
```

```{r date_flag}
cae <- mutate(cae, date_flag = date > today() | year < 1999 | is.na(date))
percent(mean(cae$date_flag, na.rm = TRUE), 0.001)
```

```{r bar_year, echo=FALSE}
cae %>% 
  filter(!date_flag) %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "California Expenditures by Year",
    caption = "Source: californiacivicdata.org",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
cae <- mutate(
  .data = cae, across(
    .cols = ends_with("_zip4"),
    .fns = list(norm = normal_zip),
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  cae$payee_zip4,
  cae$payee_zip4_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
cae <- mutate(
  .data = cae, across(
    .cols = ends_with("_st"),
    .fns = list(norm = normal_state),
    abbreviate = TRUE,
    na_rep = TRUE,
  )
)
```

```{r state_view}
cae %>% 
  filter(payee_st != payee_st_norm) %>% 
  count(payee_st, payee_st_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  cae$payee_st,
  cae$payee_st_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
cae <- mutate(
  .data = cae, across(
    .cols = ends_with("_city"),
    .fns = list(norm = normal_city),
    abbs = usps_city,
      states = c("CA", "DC", "CALIFORNIA"),
      na = invalid_city,
      na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
cae <- cae %>% 
  left_join(
    y = zipcodes,
    by = c(
      "payee_st_norm" = "state",
      "payee_zip4_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(payee_city_norm, city_match),
    match_dist = str_dist(payee_city_norm, city_match),
    payee_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = payee_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- cae %>% 
  mutate(
    payee_city_refine = payee_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(payee_city_refine != payee_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "payee_city_refine" = "city",
      "payee_st_norm" = "state",
      "payee_zip4_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    payee_st_norm, 
    payee_zip4_norm, 
    payee_city_swap, 
    payee_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
cae <- cae %>% 
  left_join(good_refine) %>% 
  mutate(payee_city_refine = coalesce(payee_city_refine, payee_city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(cae$payee_city),
  str_to_upper(cae$filer_city),
  cae$payee_city_norm,
  cae$filer_city_norm,
  cae$payee_city_swap,
  cae$payee_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(cae$payee_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "California City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "California City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
cae <- cae %>% 
  select(
    -payee_city_norm,
    -payee_city_swap,
    payee_city_clean = payee_city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(cae, 50))
```

1. There are `r comma(nrow(cae))` records in the database.
1. There are `r comma(sum(cae$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(cae$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ca", "expends", "data", "clean"))
clean_path <- path(clean_dir, "ca_expends_clean.csv")
write_csv(cae, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
