---
title: "Puerto Rico Contributions"
author: "Kiernan Nicholls & Aarushi Sahejpal"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("pr", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("pr/contribs/docs/pr_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Puerto Rico [Oficina del Contralor Electoral][oce],
the Comptroller's Office (OCE).

> Descripción: La Ley Núm. 222 de 18 de noviembre de 2011, según enmendada,
mejor conocida como “Ley para la Fiscalización del Financiamiento de Campañas
Políticas en Puerto Rico”, crea la Oficina del Contralor Electoral, a fin de
establecer el marco legal y administrativo que regirá la legalidad, examen y
evaluación de donativos y gastos para fines electorales.

Translation via Google Translate: 

> _Description: Act No. 222 of November 18, 2011, as amended, better known as the
“Law for the Control of Political Campaign Financing in Puerto Rico,” creates
the Office of the Electoral Comptroller, in order to establish the legal
framework and administrative that will govern the legality, examination and
evaluation of donations and expenses for electoral purposes._

[oce]: http://oce.pr.gov/

The data is accessed through the [Puerto Rico OpenData portal][odp]. The data
was first uploaded on September 28, 2016 and last updated February 14, 2019.

There are 166,000 rows and 11 columns (translated):

* `Candidate`: Name of the candidate, party or campaign committee.
* `Candidature`: Description of the candidate, party or campaign committee.
* `Acronym`: Acronym of the political party or type of committee:
    * "CI" = Independent Candidate
    * "CGI" = Independent Expenditure Committee
* `Amount`: Amount of money contributed.
* `Method`: Donation form.
* `Donor`: Full name of the donor.
* `City`: City where the donor resides.
* `Date`: Date the donation was generated (day / month / year).
* `Event`: Description of the election year or event.
* `Zip Code`: Donor zip code.
* `Location 1`: Donation coordinates based on the donor's zip code.

[odp]: https://data.oce.pr.gov

## Download

```{r raw_dir}
raw_dir <- dir_create(here("pr", "contribs", "data", "raw"))
raw_csv <- path(raw_dir, "Donaciones.csv")
```

```{r raw_download}
if (!file_exists(raw_csv)) {
  GET(
    url = "https://data.oce.pr.gov/api/views/kdwd-nb6g/rows.csv",
    query = list(accessType = "DOWNLOAD"),
    write_disk(path = tmp <- file_temp(ext = "csv")),
    progress(type = "down")
  )
}
```

## Read

```{r raw_read}
prc <- read_delim(
  file = raw_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    Cantidad = col_double(),
    Fecha = col_date("%d/%m/%Y")
  )
)
```

```{r clean_name}
prc <- clean_names(prc, case = "snake")
```

We can also rename the columns to their English translation.

```{r}
# per Google Translate
names(prc)[1:9] <- c(
  "candidate", # candidato
  "candidacy", # candidatura
  "acronym",   # siglas
  "amount",    # cantidad
  "method",    # metodo
  "donor",     # donante
  "city",      # ciudad
  "date",      # fecha
  "event"      # evento
)
```

## Explore

There are `r comma(nrow(prc))` rows of `r ncol(prc)` columns. Each record
represents a single contribution from an individual to a political candidate.

```{r glimpse}
glimpse(prc)
tail(prc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(prc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("date", "donor", "amount", "candidate")
prc <- flag_na(prc, all_of(key_vars))
mean(prc$na_flag)
```

```{r na_view}
prc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
prc <- flag_dupes(prc, everything())
mean(prc$dupe_flag)
```

```{r dupe_view}
prc %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(date)
```

### Categorical

```{r distinct_count}
col_stats(prc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(prc, acronym)
explore_plot(prc, method) + scale_x_wrap()
explore_plot(prc, event) + scale_x_wrap()
```

### Amounts

```{r amount_summary}
summary(prc$amount)
mean(prc$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(prc[c(which.max(prc$amount), which.min(prc$amount)), ])
```

```{r hist_amount, echo=FALSE}
prc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 10) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Puerto Rico Contributions Amount Distribution",
    caption = "Source: Oficina del Contralor Electoral",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
prc <- mutate(prc, year = year(date))
```

```{r date_range}
min(prc$date)
sum(prc$year < 2000)
max(prc$date)
sum(prc$date > today())
```

```{r bar_year, echo=FALSE}
prc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Puerto Rico Contributions by Year",
    caption = "Source: Oficina del Contralor Electoral",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
prc <- prc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip_code,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  prc$zip_code,
  prc$zip_norm,
  compare = valid_zip
)
```

### State

There is no state information but we can add the state associated with each
normalized `zip_code` value.

```{r state_match}
prc <- left_join(
  x = prc,
  y = zipcodes[, 2:3],
  by = c("zip_norm" = "zip")
)
```

```{r state_rename}
prc <- rename(prc, state_match = state)
count(prc, state_match, sort = TRUE)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- prc %>% 
  distinct(city, state_match, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("PR", "DC", "PUERTO RICO"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_match" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
prc <- left_join(
  x = prc,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state_match", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- prc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_match" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_match, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
prc <- prc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(prc$city),
  prc$city_norm,
  prc$city_swap,
  prc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
  mutate(across(1, md_code)) %>% 
  kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(prc$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Puerto Rico City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Puerto Rico City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
prc <- prc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(city_clean, state_match, zip_clean, .before = last_col())
```

```{r clean_glimpse}
glimpse(sample_n(prc, 50))
```

1. There are `r comma(nrow(prc))` records in the database.
1. There are `r comma(sum(prc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(prc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("pr", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "pr_contribs_20110101-20181231.csv")
write_csv(prc, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
