---
title: "Michigan Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mi", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `zip`
1. Create a `year` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  gluedown, # printing markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  rvest, # read html pages
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

This data is obtained from the Michigan Board of Elections (BOE) Campaign
Finance Reporting (CFR) system. The data is provided as annual ZIP archives.

```{r key_page, echo=FALSE}
key_url <- "https://miboecfr.nictusa.com/cfr/dumpall/cfrdetail/ReadMe_CONTRIBUTIONS.html"
key_page <- read_html(key_url)
```

```{r key_desc, results='asis', echo=FALSE}
key_page %>% 
  html_node("p") %>% 
  html_text() %>% 
  md_quote()
```

```{r key_vars, results='asis', echo=FALSE}
key_page %>% 
  html_node("table") %>% 
  html_table() %>% 
  as_tibble() %>% 
  mutate(
    X1 = md_code(X1),
    X2 = str_trim(str_squish(str_trunc(X2, 100)))
  ) %>% 
  md_table(col.names = c("Variable", "Description"))
```

## Import

### Download

```{r download_raw}
raw_dir <- dir_create(here("mi", "contribs", "data", "raw"))
raw_base <- "https://miboecfr.nictusa.com/cfr/dumpall/cfrdetail/"
raw_page <- read_html(raw_base)
raw_urls <- raw_page %>% 
  html_node("table") %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("contributions") %>% 
  str_c(raw_base, ., sep = "/")
raw_paths <- path(raw_dir, basename(raw_urls))
if (!all(this_file_new(raw_paths))) {
  download.file(raw_urls, raw_paths)
}
```

```{r download_raw2, include=FALSE, eval=FALSE}
write_zip <- function(url, dir) {
  path <- fs::path(dir, basename(url))
  r <- httr::GET(url, httr::write_disk(path, overwrite = TRUE))
}
map(raw_urls, write_zip, raw_dir)
```

### Read

```{r mic_names}
mic_names <- str_split(read_lines(raw_paths[1])[1], "\t")[[1]]
mic_names <- mic_names[-length(mic_names)]
mic_names[length(mic_names)] <- "runtime"
```

```{r read_tsv, eval=FALSE}
mic <- vroom(
  file = raw_paths,
  delim = "\t",
  skip = 1,
  col_names = mic_names,
  col_types = cols(
    .default = col_character(),
    page_no = col_integer(),
    doc_stmnt_year = col_integer(),
    received_date = col_date_usa(),
    amount = col_double(),
    aggregate = col_double(),
    runtime = col_skip()
  )
)
```

```{r map_read}
mic <- map_dfr(
  .x = raw_paths,
  .f = read_delim,
  escape_double = FALSE,
  escape_backslash = FALSE,
  delim = "\t",
  skip = 1,
  col_names = mic_names,
  col_types = cols(
    .default = col_character(),
    page_no = col_integer(),
    doc_stmnt_year = col_integer(),
    received_date = col_date_usa(),
    amount = col_double(),
    aggregate = col_double(),
    runtime = col_skip()
  )
)
```

## Explore

```{r glimpse}
head(mic)
tail(mic)
glimpse(mic)
```

```{r summary_amount}
summary(mic$amount)
```

```{r amount_histogram, echo=FALSE}
mic %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Michigan Contribution Amount Distribution",
    subtitle = "from 1998 to 2020",
    caption = "Source: Michigan Board of Elections",
    x = "Amount",
    y = "Count"
  )
```

```{r year_add}
mic <- mutate(mic, received_year = year(received_date))
```

```{r date_range}
min(mic$received_date, na.rm = TRUE)
sum(mic$received_year < 1998, na.rm = TRUE)
max(mic$received_date, na.rm = TRUE)
sum(mic$received_date > today(), na.rm = TRUE)
```

```{r year_bar, echo=FALSE}
mic %>% 
  count(received_year) %>% 
  ggplot(aes(received_year, n)) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  coord_cartesian(xlim = c(1998, 2020)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Michigan Contributions Recieved by Year",
    subtitle = "from 1998 to 2020",
    caption = "Source: Michigan Board of Elections",
    x = "Year Recieved",
    y = "Count"
  )
```

## Wrangle

### Address

```{r address_norm}
mic <- mic %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
mic %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

```{r zip_norm}
mic <- mic %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mic$zip,
  mic$zip_norm,
  compare = valid_zip
)
```

### State

```{r state_norm}
mic <- mic %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_progress}
progress_table(
  mic$state,
  mic$state_norm,
  compare = valid_state
)
```

### City

```{r city_norm}
mic <- mic %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("MI", "DC", "MICHIGAN"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r city_swap}
mic <- mic %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  )
```

```{r city_progress}
progress_table(
  mic$city_raw,
  mic$city_norm,
  mic$city_swap,
  compare = valid_city
)
```

## Conclude

1. There are `nrow(df)` records in the database.
1. There are `sum(mic$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `sum(mic$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 5-digit `zip_norm` variable has been created with `campfin::normal_zip(mic$zip)`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r create_proc_dir}
proc_dir <- dir_create(here("mi", "contribs", "data", "processed"))
```

```{r write_clean}
mic %>% 
  write_csv(
    path = glue("{proc_dir}/df_type_clean.csv"),
    na = ""
  )
```
