---
title: "Connecticut Businesses"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "ct", "Businesses", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/ct/business/docs/ct_business_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

Businesses registrations in Connecticut are provided online as part of a
collaboration between the [CT Data Collaborative][dc] and the Office of The
Secretary of State.

[dc]: http://ctbusiness.ctdata.org/#/

> In partnership with the Connecticut Secretary of the State, The Connecticut
Data Collaborative is excited to present multiple tools to explore business
registration data from January 1980 to May 2021. This is the first time this
data has been made open and accessible. Here you will find six ways to explore
the data.

## Download

We can obtain business [registration data by month][mn] from a public Google 
Drive folder. This folder can be downloaded as a single ZIP archive and the
CSV files can then be extracted locally.

[mn]: https://www.ctdata.org/sots-monthly-registrations

```{r}
self_name <- function(x) {
  stopifnot(is_character(x))
  setNames(x, x)
}
```

```{r raw-dir}
raw_dir <- dir_create(here("state", "ct", "business", "data", "raw"))
raw_csv <- dir_ls(raw_dir, glob = "*.csv")
raw_zip <- dir_ls(raw_dir, glob = "*.zip")
if (length(raw_csv) == 0) {
  raw_csv <- unzip(raw_zip, exdir = raw_dir, junkpaths = TRUE)
  raw_csv <- self_name(fs_path(sort(str_subset(raw_csv, "csv$"))))
}
```

## Read

The monthly CSV files can be read together into a list.

```{r raw-read}
ctb <- map(
  .x = raw_csv,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    annual_report_due_date = col_date(),
    date_registration = col_date(),
    woman_owned_organization = col_logical(),
    veteran_owned_organization = col_logical(),
    minority_owned_organization = col_logical(),
    organization_is_lgbtqi_owned = col_logical(),
    create_dt = col_datetime(),
    began_transacting_in_ct = col_date(),
    date_of_organization_meeting = col_date()
  )
)
```

That list of monthly registrations can be combined into a single data frame.

```{r raw-bind}
ctb <- ctb |> 
  bind_rows(.id = "source_file") |> 
  relocate(source_file, .after = last_col()) |> 
  mutate(across(source_file, basename))
```

```{r echo=FALSE}
ctb <- ctb |>
  relocate(
    office_jurisdiction_1,
    office_jurisdiction_2,
    office_jurisdiction_3,
    office_jurisdiction_4,
    .after = office_jurisdiction
  ) |> 
    relocate(
    mailing_jurisdiction_1,
    mailing_jurisdiction_2,
    mailing_jurisdiction_3,
    mailing_jurisdiction_4,
    .after = mailing_jurisdiction
  )
```

We are going to keep the first 30 columns, dropping mostly empty and redundant
data at the

```{r raw-select}
ctb <- ctb[, c(1:30, 58)]
```

## Explore

There are `r comma(nrow(ctb))` rows of `r ncol(ctb)` columns. Each record
represents a single business registered in Connecticut.

```{r glimpse}
glimpse(ctb)
tail(ctb)
```

### Missing

There are no rows missing a name or creation date.

### Duplicates

There are no duplicate rows.

### Categorical

```{r distinct-count}
col_stats(ctb, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(ctb, business_type)
explore_plot(ctb, status)
explore_plot(ctb, citizenship)
```

```{r lgl-count}
ctb |> 
  select(where(is.logical)) |> 
  map(~round(prop.table(table(.)), 3))
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
ctb <- mutate(ctb, year_registration = year(date_registration))
```

```{r date-range}
min(ctb$date_registration)
max(ctb$date_registration)
```

```{r bar-year, echo=FALSE}
ctb %>% 
  count(
    y = year(date_registration), m = month(date_registration)
  ) %>% 
  mutate(
    lbl = as_factor(paste(y, m, sep = "."))
  ) |> 
  ggplot(aes(x = lbl, y = n)) +
  geom_col(aes(fill = as.character(y))) + 
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = NULL) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut Businesses by Year",
    caption = "Source: DT Data Collab.",
    x = "Year Made",
    y = "Count",
    fill = "Year"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
addr_norm <- ctb %>% 
  distinct(billingstreet) %>% 
  mutate(
    address_norm = normal_address(
      address = billingstreet,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address-view}
addr_norm
```

```{r address-join}
ctb <- left_join(ctb, addr_norm, by = "billingstreet")
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
ctb <- ctb %>% 
  mutate(
    zip_norm = normal_zip(
      zip = billingpostalcode,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  ctb$billingpostalcode,
  ctb$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state-norm}
ctb <- ctb %>% 
  mutate(
    state_norm = normal_state(
      state = billingstate,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state-view}
ctb %>% 
  filter(billingstate != state_norm) %>% 
  count(billingstate, state_norm, sort = TRUE)
```

```{r state-progress}
progress_table(
  ctb$billingstate,
  ctb$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- ctb %>% 
  distinct(billingcity, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = billingcity, 
      abbs = usps_city,
      states = c("CT", "DC", "CONNECTICUT"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = billingcity) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
ctb <- left_join(
  x = ctb,
  y = norm_city,
  by = c(
    "billingcity" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- ctb %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
ctb <- ctb %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ctb$billingcity),
  ctb$city_norm,
  ctb$city_swap,
  ctb$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(ctb$billingcity, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.98, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Connecticut City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
ctb <- ctb %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(ctb, 1000))
```

1. There are `r comma(nrow(ctb))` records in the database.
1. There are `r comma(sum(ctb$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ctb$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(ctb$date_registration), "-")
max_dt <- str_remove_all(max(ctb$date_registration), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("state", "ct", "business", "data", "clean"))
clean_csv <- path(clean_dir, glue("ct_businesses_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(ctb, clean_csv, na = "")
write_rds(ctb, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
