---
title: "Utah Expenditures Data Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ut", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

## Data

Campaign expenditures are from the Utah [Financial Disclosures website][udw].

[udw]: https://disclosures.utah.gov/

> #### Reporting Contributions and Expenditures:
> The general requirement for contribution reporting is to report each
contribution you receive within 31 days of receiving it. However, there are
additional requirements that must be followed: ...
> 
> * All expenditures must be reported by the reporting deadline for each
reporting period.
> * All additional contributions received during a reporting period must be
reported by the reporting deadline.
> * All filing reports must be filled by no later than 11:59 p.m. on the day of
the reporting deadline. Failure to file a report on time may result in a fine.

## Download

Using the website's [advance search feature][adv], users can search along a
name, entity type, and report year. The site gives ambiguous directions, but
we can obtain _all_ records for a year and entity type if we leave the name
blank and click on the top record.

>  Please follow these steps:
> * Select an Entity Type
> * Select a Report Year
> * Click Find
> * In the top row of results, click the year under the far right column --
"Download Data by Year"

We make these direct HTTP requests using `httr::GET()`.

[adv]: https://disclosures.utah.gov/Search/AdvancedSearch

```{r raw_dir}
raw_dir <- dir_create(here("ut", "expends", "data", "raw"))
```

Search results are categorized across eight entity types for every year from
1998 to 2020.

* PCC = Candidates & Office Holders
* CORP = Corporation
* ELECT = Electioneering
* INDEXP = Independent Expenditures
* LABOR = Labor Organizations
* PAC = Political Action Committee
* PIC = Political Issues Committee
* PARTY = Political Party

We can first scrape all the entity types and report years from the search site's
dropdown menus.

```{r search_years}
sch_url <- "https://disclosures.utah.gov/Search/AdvancedSearch"
ut_search <- read_html(sch_url)
report_years <- ut_search %>% 
  html_nodes("#ReportYear > option") %>% 
  html_attr("value") %>% 
  as.integer()
```

```{r}
min(report_years)
max(report_years)
```

```{r search_types}
entity_types <- ut_search %>% 
  html_nodes("#EntityType > option") %>% 
  html_attr("value") %>% 
  str_subset("\\w")
```

There are `r length(entity_types) * length(report_years)` combinations of these
types and years, however many of these combinations have no search results.

```{r search_combn}
length(entity_types) * length(report_years)
```

We can make a `for` loop to run through each combination and request data from
the server with `httr::GET()`. For any combination without any results, a HTML
file is returned instead of the CSV file we want; these files can be detected
and deleted, leaving us with only files containing expenditure records.

```{r search_get, eval=FALSE}
rpt_url <- str_c(sch_url, "/GenerateReport")
for (y in report_years) {
  for (t in entity_types) {
    p <- path(raw_dir, glue("{t}-{y}.csv"))
    if (file_exists(p)) {
      next() # skip is file exists
    } else { # save if file needed
      q <- list(ReportYear = y, EntityType = t)
      g <- GET(rpt_url, query = q, write_disk(p, TRUE))
      if (str_detect(headers(g)[["content-type"]], "csv")) {
      } else { # delete if html or txt
        file_delete(p) 
      }
      Sys.sleep(runif(1, 5, 10)) 
    }
  }
}
```

We are left with `r length(dir_ls(raw_dir))` CSV files with data.

```{r entity_list}
raw_info <- dir_info(raw_dir)
nrow(raw_info)
sum(raw_info$size)
as_tibble(raw_info) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```
## Read

```{r raw_read}
utc <- read_delim(
  file = dir_ls(raw_dir)
)
```

## Explore

There are `r scales::comma(nrow(utc))` rows of `r ncol(utc)` columns.

```{r glimpse}
glimpse(utc)
tail(utc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(utc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
utc <- utc %>% flag_na(date, last, amount, committee)
sum(utc$na_flag)
```

```{r na_view}
utc %>% 
  filter(na_flag) %>% 
  select(date, last, amount, committee)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
utc <- flag_dupes(utc, -id)
sum(utc$dupe_flag)
```

```{r dupe_view}
utc %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, committee)
```

### Categorical

```{r distinct_count}
col_stats(utc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(utc, type)
```

### Amounts

```{r amount_summary}
summary(utc$amount)
mean(utc$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(utc[c(which.max(utc$amount), which.min(utc$amount)), ])
```

```{r hist_amount, echo=FALSE}
utc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Utah Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
utc <- mutate(utc, year = year(date))
```

```{r date_range}
min(utc$date)
sum(utc$year < 2000)
max(utc$date)
sum(utc$date > today())
```

```{r bar_year, echo=FALSE}
utc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Utah Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
utc <- utc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
utc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
utc <- utc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  utc$zip,
  utc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
utc <- utc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
utc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  utc$state,
  utc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
utc <- utc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("UT", "DC", "UTAH"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
utc <- utc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- utc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
utc <- utc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(utc$city_raw),
  utc$city_norm,
  utc$city_swap,
  utc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(utc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Utah City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Utah City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
utc <- utc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(utc, 50))
```

1. There are `r comma(nrow(utc))` records in the database.
1. There are `r comma(sum(utc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(utc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ut", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ut_contribs_clean.csv")
write_csv(utc, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
s3_head <- head_object(s3_path, "publicaccountability")
(s3_size <- as_fs_bytes(attr(s3_head, "content-length")))
unname(s3_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(utc)),
  type = md_code(map_chr(utc, typeof)),
  def = c(
    ""
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
