---
title: "North Dakota Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nd", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  aws.s3, # upload to AWS
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts can be downloaded from the [North Dakota transparency portal][tran],
a site run by the [state Office of Managment and Budget (OMB)][omb].

The Transparency [FAQ] page provides more information on the site and data:

> #### About the Data
> This portal was established by the 2009 Legislative Assembly, North Dakota
Century Code 54-44.1-18, Searchable Database of Expenditures, and provides a
single point of reference to view State and University System expenditures.
Providing direct access to the state's accounting and payroll systems'
checkbook, the North Dakota Transparency portal is a tool that opens the books
of government. It is a powerful way for the public to access information about
the State of North Dakota’s payment activities as well as providing salary
information on state employees. The portal, created by the Office of Management
and Budget (OMB), will make state financial activity more understandable and
create additional accountability on how tax dollars are utilized.  A tutorial
has been provided to assist users on how to best use the portal.

> #### What information is included?
> All payments and payroll transactions made through the State of North Dakota's
PeopleSoft accounting system are included in the transparency portal. North
Dakota’s Transparency portal contains payment information for all state
agencies, the legislature, courts, and University system.

> #### What information is excluded?
> As directed by law, the website does not include aid payments to individuals,
organizations, and local assistance payments.  Also, the names of some state
employees have been redacted to protect their safety and welfare. For example,
undercover officers in the Bureau of Criminal Investigation are not include to
protect the agent’s cover.  All transactions where the vendor’s name is
protected by law will list the account description as the vendor instead.

> #### How often is the data updated?
> Payment and salary information is provided from data recorded in the state’s
PeopleSoft accounting system and is updated on a monthly basis, usually around
the middle of the month for the fiscal period just ended.

> #### How do I download information?
> In the left menu, you will see a “Mass Download” hyperlink.  By clicking on
that link, you will have the option to export detailed checkbook data for a
biennium to a CSV or Microsoft Excel file.

The site also provides a glossary:

```{r glossary, echo=FALSE}
gloss <- read_html("https://www.nd.gov/omb/public/transparency/glossary")
gloss <- tibble(
  term = html_text(html_nodes(gloss, "h2")),
  def = html_text(html_nodes(gloss, "p"))[-18]
)
kable(gloss, col.names = c("Term", "Definition"))
```

[tran]: https://www.nd.gov/omb/public/transparency
[omd]: https://www.nd.gov/omb/
[FAQ]: https://www.nd.gov/omb/public/transparency/faq

## Download

Per the instructions found in the [FAQ], we can download files from the portal's
mass download page. The OMB provides data in Excel files by fiscal year.

```{r raw_list}
raw_home <- "https://www.nd.gov/omb/public/transparency/mass-download"
raw_urls <- read_html(raw_home) %>% 
  html_nodes(".field-items") %>% 
  html_nodes("li > a") %>% 
  html_attr("href") %>% 
  str_c("https://www.nd.gov", .)
```

```{r raw_dir}
raw_dir <- dir_create(here("nd", "contracts", "data", "raw"))
raw_paths <- path(raw_dir, basename(raw_urls))
```

```{r raw_download}
for (i in seq_along(raw_urls)) {
  if (file_exists(raw_paths[i])) {
    next()
  } else {
    download.file(raw_urls[i], raw_paths[i])
  }
}
```

```{r raw_info}
raw_info <- dir_info(raw_dir)
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## Read

The transactions are spread into multiple sheets in each Excel workbook. We can
write a new function that identifies the sheets and reads them together.

```{r read_all_sheets}
read_all_sheets <- function(path) {
  path %>%
    readxl::excel_sheets() %>% # identify sheet names
    purrr::set_names() %>% # create a vector
    purrr::map_df( # read all sheets together
      .f = readxl::read_excel,
      .name_repair = janitor::make_clean_names,
      path = path
    )
}
```

```{r raw_read}
ndc <- map_df(
  .x = raw_paths,
  .f = read_all_sheets
)
```

Across the different sheets and files, some columns have different names and
need to be coalesced into a single column.

```{r raw_combine}
ndc <- ndc %>% 
  mutate(fund = coalesce(fund, fund_description)) %>% 
  mutate(across(date, as_date)) %>% 
  select(-fund_description) %>% 
  rename(
    category = expenditure_category,
    description = detailed_expenditure_category
  )
```

Note that this database does _not_ contain street addresses for businesses.

```{r raw_head}
comma(nrow(ndc))
head(ndc)
```

## Join

The Investigative Reporting Workshop initially requested this data using the
North Dakota Open Records Statute, N.D.C.C. §44-04-18 et seq. The Director of
the Central Services Division provided a single Excel file to satisfy our
request, but pointed us to the more comprehensive public data.

```{r req_read}
req_path <- path(raw_dir, "ND July+21+2020+Open+Records+Request.xlsx")
ndrr <- read_all_sheets(req_path)
ndrr <- mutate(ndrr, across(date, as_date))
```

This requested data has half as many rows and slightly different columns.

```{r raw_nrow}
comma(nrow(ndrr))
head(ndrr)
```

Importantly, for our interests, the requested data contains street addresses for
the business receiving the payment. The Public Accountability project uses
addresses as a way to identify the parties to a transaction. Unfortunately,
there are no unique vendor IDs that can be used to join the two databases. The
business names are not quite identical.

First, we will count all of the vendors in the public database.

```{r}
a <- ndc %>% 
  count(vendor, city, state, sort = TRUE) %>% 
  select(-n)
```

Next, we will count the vendors in the requested data after normalizing the
addresses.

```{r}
b <- ndrr %>% 
  rename(vendor = name, state = st, zip = postal) %>% 
  # address and zip are new
  count(vendor, city, state, address_1, address_2, zip, sort = TRUE) %>% 
  select(-n)
```

Then, we will find all vendors, cities, and state values common between both
databases. This will allow us to add the addresses and postal codes that are
missing in the raw data. Some vendors have multiple addresses with the same
city and state. We can only give each business name a single address, so we will
use the most common.

```{r}
c <- a %>% 
  inner_join(b, by = c("vendor", "city", "state")) %>% 
  group_by(vendor, city, state) %>% 
  slice(1)
```

After joining this address database with our transactions, we have the same
number of rows but have added the `address_*` and `zip` columns.

```{r}
dim(ndc)
ndc <- left_join(ndc, c)
dim(ndc)
```

Slightly more than half of rows have addresses, but it's better than nothing.

```{r}
percent(prop_na(ndc$address_1), 0.01)
```

## Explore

```{r glimpse}
glimpse(ndc)
tail(ndc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(ndc, count_na)
```

No rows are missing the names, date, or amount needed to identify a transaction.

A lot of records have an "exempt" city value.

```{r na_geo_view}
ndc %>% 
  count(city, state, sort = TRUE) %>% 
  add_prop()
```

```{r na_exempt}
ndc <- mutate(ndc, geo_exempt = (city == "Exempt") & is.na(state))
ndc$city[ndc$geo_exempt] <- NA_character_
prop_na(ndc$address_1[!ndc$geo_exempt])
```

### Duplicates

```{r dupe_flag}
d1 <- duplicated(ndc, fromLast = FALSE)
d2 <- duplicated(ndc, fromLast = TRUE)
if (sum(d1 | d2) > 0) {
  ndc <- mutate(ndc, dupe_flag = d1 | d2)
  sum(ndc$dupe_flag)
}
rm(d1, d2); flush_memory()
```

### Categorical

```{r distinct_count}
col_stats(ndc, n_distinct)
```

```{r distinct_plots}
explore_plot(ndc, category) + scale_x_truncate()
explore_plot(ndc, fund) + scale_x_truncate()
explore_plot(ndc, agency) + scale_x_truncate()
```

### Amounts

```{r amount_summary}
summary(ndc$amount)
mean(ndc$amount <= 0)
```

Here are the smallest and largest transactions.

```{r amount_minmax}
glimpse(ndc[c(which.min(ndc$amount), which.max(ndc$amount)), ])
```

```{r hist_amount, echo=FALSE}
ndc %>%
  filter(amount >= 1, amount <= 1e6) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "North Dakota Contracts Amount Distribution",
    caption = "Source: ND OMB",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
ndc <- mutate(ndc, year = year(date))
```

```{r date_range}
min(ndc$date)
sum(ndc$year < 2000)
max(ndc$date)
sum(ndc$date > today())
```

```{r bar_year, echo=FALSE}
ndc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  labs(
    title = "North Dakota Contracts by Year",
    caption = "Source: ND OMB",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

Street addresses are already abbreviated and normalized how we would.

```{r addr_check}
ndc %>% 
  count(address_1, address_2, sort = TRUE)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
ndc <- ndc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  ndc$zip,
  ndc$zip_norm,
  compare = valid_zip
)
```

### State

Almost all `state` values are valid 2-digit abbreviations and most of the
remaining are Canadian provinces or some foreign country we want to keep.

```{r state_check}
percent(prop_in(ndc$state, valid_state), 0.01)
```

```{r state_norm}
ndc %>% 
  count(state, sort = T) %>% 
  filter(state %out% valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ndc <- ndc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("ND", "DC", "NORTH DAKOTA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ndc <- ndc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c("state", "zip_norm" = "zip")
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- ndc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c("city_refine" = "city", "state", "zip_norm" = "zip")
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
ndc <- ndc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ndc$city_raw),
  ndc$city_norm,
  ndc$city_swap,
  ndc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ndc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "North Dakota City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Dakota City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
ndc <- ndc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(ndc, 50))
```

1. There are `r comma(nrow(ndc))` records in the database.
1. There are `r comma(sum(ndc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ndc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nd", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "nd_contracts_clean.csv")
write_csv(ndc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=TRUE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    multipart = TRUE,
    show_progress = TRUE
  )
}
```

```{r s3_size}
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ndc)),
  type = md_code(map_chr(ndc, typeof)),
  def = c(
    "The two-year period spending made",
    "Fiscal year code spending made",
    "Date contract of payment made",
    "Spending agency name",
    "Receiving vendor name",
    "Vendor city name",
    "Vendor state abbreviation",
    "Spending category",
    "Description of payment made",
    "Dollar amount paid",
    "Amount fund source",
    "Vendor street address",
    "Vendor secondary address",
    "Vendor ZIP+4 code",
    "Exempt from geographic information",
    "Calendar year payment made",
    "Normalized 5-digit ZIP code",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
