---
title: "Missouri Lobbying Registration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mo", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, #read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The [Missouri Ethics Commission](https://jcope.ny.gov/lobbying-datasets) makes lobbyist expenditure data from 2004 to 2020 available for download. Note that data for 2019 onward exists in a separate search. The 2018 data and prior year's data exist in two separate files.

## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("mo", "lobby", "data", "raw","exp"))
```

### Read
The `xlsx` files we obtained actually contain two tabs, one for `individual/solicitation` and the other for `group`. The columns are slightly dififerent, and we will read them separately. 
```{r raw_read}
mo_lob_exp <- dir_ls(raw_dir) %>% str_subset("200.|201[^9]")  %>% map_dfr(read_xlsx, sheet = 1) %>% clean_names()
mo_lob_grp <- dir_ls(raw_dir) %>% str_subset("200.|201[^9]") %>% map_dfr(read_xlsx, sheet = 2) %>% clean_names()
```
## Explore

```{r glimpse}
head(mo_lob_exp)
tail(mo_lob_exp)
glimpse(sample_n(mo_lob_exp, 20))

head(mo_lob_grp)
tail(mo_lob_grp)
glimpse(sample_n(mo_lob_grp, 20))
```

### Missing
Not many fields have missing records. However, we can convert the `NULL` entries to _NAs_.
```{r na_count}
col_stats(mo_lob_exp, count_na)
col_stats(mo_lob_exp, count_na)

mo_lob_grp <- mo_lob_grp %>% 
  mutate(amend_reason = na_if(amend_reason, "NULL"),
         description = na_if(description, "NULL"))

mo_lob_exp <- mo_lob_exp %>% 
  mutate(amend_reason = na_if(amend_reason, "NULL"),
         description = na_if(description, "NULL"))
```


### Duplicates
Running the following commands shows that there are over a thousand entries with totally identical information. Since each of them contain `total_compensation` and `total_reimbursed` for a certain period, there should be only one such entry for each corresponding time period. We will note that in the `dupe_flag` column.
```{r dupe_flag}
mo_lob_exp <- flag_dupes(mo_lob_exp, dplyr::everything())
sum(mo_lob_exp$dupe_flag)
```

```{r dupe_flag grp}
mo_lob_grp <- flag_dupes(mo_lob_grp, dplyr::everything())
sum(mo_lob_grp$dupe_flag)
```


### Categorical
First, we will create a `year` variable.
```{r year}
mo_lob_exp <- mo_lob_exp %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y"),
         year = year(date))
```

```{r year}
mo_lob_grp <- mo_lob_grp %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y"),
         year = year(date))
```

#### No. of expenditures
Visualize number of expenses each year.
We can see that some of the year variable is not valid, with values greater than 2200. We will manually filter these entries out before visualizing the data. 
```{r year_bar, echo=FALSE}
exp_no_ind <- mo_lob_exp %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2004, 2019, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Missouri Lobbyist Expenditures by Year (Individuals and Solicitations)",
    caption = "Source: Missouri Ethics Commission",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

```{r year bar group}
exp_no_grp <- mo_lob_grp %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2004, 2019, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Missouri Lobbyist Expenditures by Year (Groups)",
    caption = "Source: Missouri Ethics Commission",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )

exp_no_grp + exp_no_ind + plot_layout(nrow = 2)
```


#### Expenditure Type

```{r client times, echo=FALSE}
mo_lob_exp %>% 
group_by(expenditure_type) %>% 
  summarize(median_exp = median(amount, na.rm = T)) %>% 
  arrange(desc(median_exp))  %>% 
  ggplot(aes(x = reorder(expenditure_type,median_exp), y = median_exp)) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  scale_y_continuous(labels = dollar ) +
  scale_x_discrete(labels = wrap_format(10)) +
    theme(legend.position = "bottom") +
  labs(
    title = "Median Expenses by Client Type",
    caption = "Source: Missouri Ethics Commission",
    fill = "Election Year",
    x = "Client Type",
    y = "Count"
  )
```

```{r expense type, echo=FALSE}
mo_lob_exp %>% filter(!dupe_flag) %>% 
  group_by(description) %>% 
  summarize(total = sum(amount)) %>% arrange(desc(total)) %>% top_n(8) %>% 
  ggplot(aes(x = reorder(description, total), y = total)) +
  geom_col() +
  scale_y_continuous(labels = dollar ) +
  scale_x_discrete(labels = wrap_format(10)) +
    theme(legend.position = "bottom") +
  labs(
    title = "Top Expense Types",
    caption = "Source: Missouri Ethics Commission",
    x = "Expense Description",
    y = "Total"
  )
```

```{r client type, echo=FALSE}
mo_lob_exp %>% filter(!dupe_flag) %>% 
  group_by(public_official) %>% 
  summarize(median = median(amount)) %>% arrange(desc(median)) %>% top_n(8) %>% 
  ggplot(aes(x = reorder(public_official, median), y = median)) +
  geom_col() +
  scale_y_continuous(labels = dollar ) +
  scale_x_discrete(labels = wrap_format(10)) +
    theme(legend.position = "bottom") +
  labs(
    title = "Top public officials targeted in lobbying expenditures (measured by median expenditure)",
    caption = "Source: Missouri Ethics Commission",
    x = "Public Officials",
    y = "Median expenditures made by lobbyists"
  )
```


### Continuous


First, let's take a look at the distribution of the amount by expenditure type
```{r}
mo_lob_exp %>% 
  filter(amount > 1, amount < 1000000) %>% 
  ggplot(aes(x = type, y = amount, fill = type)) +
  geom_violin(draw_quantiles = TRUE, scale = "width", ) +
  scale_y_continuous(labels = dollar, trans = "log10") +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(
    title = "Missouri Contribution Amount Distribution by Expenditure Type (to Individuals and Solicitations)",
    subtitle = "from 2004 to 2020",
    caption = "Source: Missouri Ethics Commission",
    x = "Expenditure Type",
    y = "Amount"
  )
```

```{r echo=FALSE}
mo_lob_grp %>% 
  filter(amount > 1, amount < 1000000) %>% 
  ggplot(aes(x = type, y = amount, fill = type)) +
  geom_violin(draw_quantiles = TRUE, scale = "width", ) +
  scale_y_continuous(labels = dollar, trans = "log10") +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(
    title = "Missouri Contribution Amount Distribution by Expenditure Type (to Groups)",
    subtitle = "from 2004 to 2020",
    caption = "Source: Missouri Ethics Commission",
    x = "Expenditure Type",
    y = "Amount"
  )
```

```{r histogram}
 mo_lob_exp %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  geom_vline(xintercept = median(mo_lob_exp$amount[mo_lob_exp$amount!= 0], na.rm = TRUE), linetype = 2) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Missouri Lobbying Expenditures Distribution",
    subtitle = "attributable expenses made by lobbyists from 2004 to 2020",
    caption = "Source: Missouri Ethics Commission",
    x = "Amount",
    y = "Count"
  )
```


## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
mo_lob_exp <- mo_lob_exp %>% 
    # combine street addr
  unite(
    col = lobbyist_address,
    starts_with("lobbyist_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
    unite(
    col = client_address,
    starts_with("client_address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate_at(.vars = vars(ends_with('address')), .funs = list(norm = ~ normal_address(.,abbs = usps_street,
      na_rep = TRUE))) %>% 
  select(-ends_with("address"))
```

```{r address_view}
mo_lob_exp %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10) %>% 
  glimpse()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mo_lob_exp <- mo_lob_exp %>% 
    mutate_at(.vars = vars(ends_with('zip')), 
              .funs = list(norm = ~ normal_zip(.))) %>% 
  rename(lobbyist_zip5 = lobbyist_zip_norm,
         client_zip5 = client_zip_norm)
```

```{r zip_progress}
progress_table(
  mo_lob_exp$lobbyist_zip,
  mo_lob_exp$lobbyist_zip5,
  mo_lob_exp$client_zip,
  mo_lob_exp$client_zip5,
  compare = valid_zip
)
```

### state

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
mo_lob_exp <- mo_lob_exp %>% 
   mutate_at(.vars = vars(ends_with('state')), .funs = list(norm = ~ normal_state(.,abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state)))
```

```{r state_view}
mo_lob_exp %>% 
  filter(lobbyist_state != lobbyist_state_norm) %>% 
  count(lobbyist_state, sort = TRUE)
```

```{r firm_state_view}
mo_lob_exp %>% 
  filter(client_state != client_state_norm) %>% 
  count(client_state, sort = TRUE)
```

```{r state_progress}
progress_table(
  mo_lob_exp$lobbyist_state,
  mo_lob_exp$lobbyist_state_norm,
  mo_lob_exp$client_state,
  mo_lob_exp$client_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good mo_lob_expart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mo_lob_exp <- mo_lob_exp %>% 
      mutate_at(.vars = vars(ends_with('city')), .funs = list(norm = ~ normal_city(.,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE)))

prop_in(mo_lob_exp$lobbyist_city_norm, valid_city, na.rm = T)
prop_in(mo_lob_exp$client_city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
mo_lob_exp <- mo_lob_exp %>% 
  left_join(
    y = zipcodes,
    by = c(
      "lobbyist_state_norm" = "state",
      "lobbyist_zip5" = "zip"
    )
  ) %>% 
  rename(lobbyist_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(lobbyist_city_norm, lobbyist_city_match),
    match_dist = str_dist(lobbyist_city_norm, lobbyist_city_match),
    lobbyist_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = lobbyist_city_match,
      false = lobbyist_city_norm
    )
  ) %>% 
  select(
    -lobbyist_city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap clients}
mo_lob_exp <- mo_lob_exp %>% 
  left_join(
    y = zipcodes,
    by = c(
      "client_state_norm" = "state",
      "client_zip5" = "zip"
    )
  ) %>% 
  rename(client_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(client_city_norm, client_city_match),
    match_dist = str_dist(client_city_norm, client_city_match),
    client_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = client_city_match,
      false = client_city_norm
    )
  ) %>% 
  select(
    -client_city_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  mo_lob_exp$lobbyist_city,
  mo_lob_exp$client_city,
  mo_lob_exp$lobbyist_city_norm,
  mo_lob_exp$client_city_norm,
  mo_lob_exp$lobbyist_city_swap,
mo_lob_exp$client_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Missouri City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "NY City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(mo_lob_exp, 20))
```

1. There are `r nrow(mo_lob_exp)` records in the database.
1. There are `r sum(mo_lob_exp$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(mo_lob_exp$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `YEAR` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("mo", "lobby", "data", "reg","clean"))
```

```{r write_clean}
write_csv(
  x = mo_lob_exp %>% rename(lobbyist_city_clean = lobbyist_city_swap) %>% rename( client_city_clean = client_city_swap),
  path = path(clean_dir, "mo_lob_exp_clean.csv"),
  na = ""
)
```
