---
title: "Missouri Lobbying Registration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mo", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.


## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The workshop obtained the MO lobbying registration data through a Sunshine Request to the Missouri Ethics Commission. The data is as current as Feb. 14, 2020. Note that the file structure changed in 2019, since which a table of all current lobbyists is available, so we will process these two datasets separately. 

## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("mo", "lobby", "data", "raw", "reg"))
```

### Read
```{r raw_read}
mo_lob <- read_xlsx(
  path = dir_ls(raw_dir, regexp = "Data.xlsx")
) %>% clean_names() %>% mutate_if(is.character, str_conv, encoding = "UTF-8") %>% mutate_if(is.character, str_to_upper)

mo_prin <- read_xlsx(
  path = dir_ls(raw_dir, regexp = "Data.xlsx"), sheet = 3
) %>% clean_names() %>% mutate_if(is.character, str_conv, encoding = "UTF-8") %>% mutate_if(is.character, str_to_upper)

mo_prin_new <- read_xlsx(
  path = dir_ls(raw_dir) %>% str_subset("2019"), sheet = 2
) %>% clean_names() %>% mutate_if(is.character, str_conv, encoding = "UTF-8") %>% mutate_if(is.character, str_to_upper)

mo_lob_new <- read_xlsx(
  path = dir_ls(raw_dir) %>% str_subset("2019")
) %>% clean_names() %>% mutate_if(is.character, str_conv, encoding = "UTF-8") %>% mutate_if(is.character, str_to_upper)
```


#### Year
Sometimes in the current lobbyist list, there will be 2019 and 2020 registrations while everything else is the same. We can safely flag the earlier ones as duplicates and go with the new registration when joining it with the principal data.
```{r create year}
mo_lob_reg <- mo_prin %>% 
  left_join(mo_lob, by = "lob_id")

mo_lob_reg <- mo_lob_reg %>% 
  mutate(year = year(rec_date))

mo_lob_reg$year %>% tabyl()
```

### Join
For years prior to 2019, we can see that each `lob_id` responds to one lobbyist, so it's a unique id, which we can use to join the `mo_lob` and `mo_prin` dataframes.

```{r}
mo_prin_new <- mo_prin_new %>% 
  mutate(year = year(relationship_from))

mo_prin_new <- mo_prin_new %>% 
  rename_at(.vars = vars(-c(mecid_l, p_name)), .funs = ~ str_c("p_",.))
```
From 2019 onward, however, each `mecid_l` can correspond to multiple years, for which we'll just aggregate with the earliest year. Note that some lobbyists terminate their status and re-register within a year as well.

```{r}
mo_min_date <- mo_lob_new %>% group_by(mecid_l) %>% summarize(date_min = min(date_registration))

mo_lob_new <- mo_lob_new %>% 
  left_join(mo_min_date, by = "mecid_l")

mo_lob_filter <- mo_lob_new %>% 
  filter(date_registration == date_min) %>% 
  select(-date_min)
```
We can see that one record `mecid_l` "L200260" doesn't have a corresponding record in the principal dataframe. As a result, we will just left join the `mo_lob_filter` with the `mo_prin_new` dataframe.
```{r}
mo_lob_reg_new <- mo_lob_filter %>% 
  left_join(mo_prin_new, by = "mecid_l")
```


### Column Specs
We can see that the `rec_date` and `ent_date` are date columns, while `term_date` is read as character. We'll convert it to POSIXct objects (dates). We can see that the date is formatted in Excel and we need to use `excel_numeric_to_date`
```{r date}
mo_lob_reg <- mo_lob_reg %>% 
  mutate(term_date = na_if(term_date, "NULL") %>% as.numeric() %>% excel_numeric_to_date(date_system = "modern"))

mo_lob_reg_new <- mo_lob_reg_new %>% 
  mutate(date_terminated = na_if(date_terminated, "NULL") %>% as.numeric() %>% excel_numeric_to_date(date_system = "modern"))
```

```{r na_if cols}
mo_lob_reg <- mo_lob_reg %>% 
  mutate_if(is.character, na_if, "NULL")

mo_lob_reg_new <- mo_lob_reg_new %>% 
  mutate_if(is.character, na_if, "NULL")
```

## Explore

```{r glimpse}
head(mo_lob_reg)
tail(mo_lob_reg)
glimpse(sample_n(mo_lob_reg, 20))

head(mo_lob_reg_new)
tail(mo_lob_reg_new)
glimpse(sample_n(mo_lob_reg_new, 20))
```

### Missing

```{r na_count}
col_stats(mo_lob_reg, count_na)
col_stats(mo_lob_reg_new, count_na)
```

```{r na_flag}
mo_lob_reg <- mo_lob_reg %>% flag_na(p_name, p_address)
sum(mo_lob_reg$na_flag)
```

### Duplicates
We can see there's no duplicate entry.
```{r dupe_flag}
mo_lob_reg <- flag_dupes(mo_lob_reg, dplyr::everything())
sum(mo_lob_reg$dupe_flag)

mo_prin_new<- flag_dupes(mo_prin_new, dplyr::everything())
mo_prin_new$dupe_flag %>% tabyl()
```

### Categorical

```{r n_dimo_lob_reginct}
col_stats(mo_lob_reg, n_distinct)
col_stats(mo_lob_reg_new, n_distinct)
```

It was not until 2007 that lobbyist were required to register every year. Prior to that a lobbyist could register and never have to register again.

#### Dates

```{r date_range}
min(mo_lob_reg$rec_date)
max(mo_lob_reg$rec_date)
min(mo_lob_reg$term_date, na.rm = T)
max(mo_lob_reg$term_date, na.rm = T)

min(mo_lob_reg_new$date_registration)
max(mo_lob_reg_new$date_registration)
min(mo_lob_reg_new$date_terminated, na.rm = T)
max(mo_lob_reg_new$date_terminated, na.rm = T)
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
mo_lob_reg <- mo_lob_reg %>% 
    # combine street addr
  unite(
    col = p_address_full,
    c(p_address, p_address2),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
    unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate_at(.vars = vars(ends_with('full')), .funs = list(norm = ~ normal_address(.,abbs = usps_street,
      na_rep = TRUE))) %>% 
  select(-ends_with("address"))

mo_lob_reg_new <- mo_lob_reg_new %>% 
    # combine street addr
  unite(
    col = p_address_full,
    c(p_address1, p_address2),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
    unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
    mutate_at(.vars = vars(ends_with('full')), .funs = list(norm = ~ normal_address(.,abbs = usps_street,
      na_rep = TRUE))) %>% 
  select(-ends_with("address"))
```

```{r address_view}
mo_lob_reg %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)

mo_lob_reg_new %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mo_lob_reg <- mo_lob_reg %>% 
  mutate(
    zip5 = normal_zip(
      zip = zip,
      na_rep = TRUE
    ),
    p_zip5 = normal_zip(
      zip = p_zip,
      na_rep = TRUE
    )
  )

mo_lob_reg_new <- mo_lob_reg_new %>% 
  mutate(
    zip5 = normal_zip(
      zip = zip,
      na_rep = TRUE
    ),
    p_zip5 = normal_zip(
      zip = p_zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mo_lob_reg$zip,
  mo_lob_reg$zip5,
  mo_lob_reg$p_zip,
  mo_lob_reg$p_zip5,
  compare = valid_zip
)

progress_table(
  mo_lob_reg_new$zip,
  mo_lob_reg_new$zip5,
  mo_lob_reg_new$p_zip,
  mo_lob_reg_new$p_zip5,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
prop_in(mo_lob_reg$state, valid_state, na.rm = T)
prop_in(mo_lob_reg$p_state, valid_state, na.rm = T)
prop_in(mo_lob_reg_new$state, valid_state, na.rm = T)
prop_in(mo_lob_reg_new$p_state, valid_state, na.rm = T)
```
We can see that the state fields in the new dataframe are clean and don't need extra cleaning.
```{r}
mo_lob_reg <- mo_lob_reg %>% 
  mutate_at(.vars = vars(ends_with('state')), .funs = list(norm = ~ normal_state(., na_rep = T)))
```

```{r state_view}
mo_lob_reg %>% 
  filter(state != state_norm) %>% 
  count(state, sort = TRUE)
```

```{r p_state_view}
mo_lob_reg %>% 
  filter(p_state != p_state_norm) %>% 
  count(p_state, sort = TRUE)
```

We can see that the "M)" in `p_state` should be MO based on the full address and city. We can manually change it.

```{r manual state}
mo_lob_reg$p_state_norm <- mo_lob_reg$p_state_norm %>% 
  str_replace("^M\\)$", "MO")
```


```{r state_progress}
progress_table(
  mo_lob_reg$state,
  mo_lob_reg$state_norm,
  compare = valid_state
)
```

### city

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good mo_lob_regart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
prop_in(mo_lob_reg$p_city, valid_city, na.rm = T)
prop_in(mo_lob_reg$city, valid_city, na.rm = T)

prop_in(mo_lob_reg_new$p_city, valid_city, na.rm = T)
prop_in(mo_lob_reg_new$city, valid_city, na.rm = T)

mo_lob_reg <- mo_lob_reg %>% 
  mutate_at(.vars = vars(ends_with('city')), .funs = list(norm = ~ normal_city(.,abbs = usps_city,
                                                                               states = usps_state,
                                                                               na = invalid_city,
                                                                               na_rep = TRUE)))
mo_lob_reg_new <- mo_lob_reg_new %>% 
  mutate_at(.vars = vars(ends_with('city')), .funs = list(norm = ~ normal_city(.,abbs = usps_city,
                                                                               states = usps_state,
                                                                               na = invalid_city,
                                                                               na_rep = TRUE)))

prop_in(mo_lob_reg$p_city_norm, valid_city, na.rm = T)
prop_in(mo_lob_reg$city_norm, valid_city, na.rm = T)
prop_in(mo_lob_reg_new$p_city_norm, valid_city, na.rm = T)
prop_in(mo_lob_reg_new$city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
mo_lob_reg <- mo_lob_reg %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip5" = "zip"
    )
  ) %>% 
 rename(city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1 | is.na(match_dist),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r}
mo_lob_reg_new <- mo_lob_reg_new %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
 rename(city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1 | is.na(match_dist),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```


```{r p_city_swap}
mo_lob_reg <- mo_lob_reg %>% 
  left_join(
    y = zipcodes,
    by = c(
      "p_state_norm" = "state",
      "p_zip5" = "zip"
    )
  ) %>% 
  rename(p_city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(p_city_norm, p_city_match),
    match_dist = str_dist(p_city_norm, p_city_match),
    p_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = p_city_match,
      false =p_city_norm
    )
  ) %>% 
  select(
    -p_city_match,
    -match_dist,
    -match_abb
  )
```

```{r new p_city_swap}
mo_lob_reg_new <- mo_lob_reg_new %>% 
  left_join(
    y = zipcodes,
    by = c(
      "p_state" = "state",
      "p_zip5" = "zip"
    )
  ) %>% 
  rename(p_city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(p_city_norm, p_city_match),
    match_dist = str_dist(p_city_norm, p_city_match),
    p_city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = p_city_match,
      false =p_city_norm
    )
  ) %>% 
  select(
    -p_city_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is close to 100%.

#### Progress
We can see the city normalization progress for 
```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  mo_lob_reg$city,
  mo_lob_reg$p_city,
  mo_lob_reg$city_norm,
  mo_lob_reg$p_city_norm,
  mo_lob_reg$city_swap,
mo_lob_reg$p_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))

kable(progress, digits = 3)
```

For the new data from 2019 onward.
```{r progress_print, echo=FALSE}
progress_new <- progress_table(
  mo_lob_reg_new$city,
  mo_lob_reg_new$p_city,
  mo_lob_reg_new$city_norm,
  mo_lob_reg_new$p_city_norm,
  mo_lob_reg_new$city_swap,
  mo_lob_reg_new$p_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress_new)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Missouri City Normalization Progress",
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "MO City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(mo_lob_reg, 20))

```

1. There are `r nrow(mo_lob_reg)` records in the database.
1. There are `r sum(mo_lob_reg$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(mo_lob_reg$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("mo", "lobby", "data", "processed", "reg","clean"))

mo_lob_reg <- mo_lob_reg %>% 
  select(-c(p_city_norm,
            city_norm
            )) %>% 
  rename(city_clean = city_swap,
            p_city_clean = p_city_swap)

mo_lob_reg_new <- mo_lob_reg_new %>% 
  select(-c(p_city_norm,
            city_norm
            )) %>% 
  rename(city_clean = city_swap,
            p_city_clean = p_city_swap)
```

```{r write_clean}
write_csv(
  x = mo_lob_reg,
  path = path(clean_dir, "mo_lob_reg_before2019.csv"),
  na = ""
)

write_csv(
  x = mo_lob_reg_new,
  path = path(clean_dir, "mo_lob_reg_after2019.csv"),
  na = ""
)
```
