---
title: "New York Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
# options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state", "ny", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Software

This data is processed using the free, open-source statistical computing
language R, which can be [installed from CRAN][cran] for various opperating
systems. For example, R can be installed from the apt package repository on
Ubuntu.

```bash
sudo apt update
sudo apt -y upgrade
sudo apt -y install r-base
```

[cran]: https://cran.r-project.org/

The following additional R packages are needed to collect, manipulate,
visualize, analyze, and communicate these results. The `pacman` package will
facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel
  knitr, # knit documents
  vroom, # read files fast
  furrr, # parallel map
  glue, # combine strings
  here, # relative storage
  httr, # http requests
  pryr, # memory usage
  fs # search storage 
)
```

This document should be run as part of the `us_spending` project, which lives as
a sub-directory of the more general, language-agnostic [`irworkshop/tap`][tap]
GitHub repository.

The `us_spending` project uses the [RStudio projects][rproj] feature and should
be run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data is obtained from the New York Office of the State Comptroller via
their Open Book portal, which includes:

> All State agency contracts, in effect 4/1/12 or later, which includes
contracts approved by the Office of the State Comptroller (OSC) and those that
don't require OSC approval.

> State authority contracts historically subject to approval by OSC (primarily
for the Thruway Authority and the Long Island Power Authority only).

## Import

We can import the file by first downloading with a `httr::GET()` request and
then reading into memory with `vroom::vroom()`.

```{r raw_path}
url <- "https://wwe2.osc.state.ny.us/transparency/contracts/contractresults.cfm"
raw_dir <- dir_create(here("state", "ny", "contracts", "data", "raw"))
raw_path <- path(raw_dir, "contracts-and-amendments.csv")
```

```{r raw_get}
if (!file_exists(raw_path)) {
  GET(
    url = url,
    write_disk(raw_path, overwrite = TRUE),
    query = list(
      DocType = "csv",
      ac = "",
      v = "",
      vo = "B",
      cn = "",
      selOrigDateChoiceOperator = "Before",
      txtOrigFromDate = format(Sys.Date(), "%m/%d/%Y"),
      txtOrigToDate = "",
      selCTDateChoice = "0",
      selCTDateChoiceOperator = "0",
      txtCTFromDate = "",
      txtCTToDate = "",
      selContractAmountChoice = "0",
      txtContractAmount1 = "",
      txtContractAmount2 = "",
      b = "Search",
      order = "VENDOR_NAME",
      sort = "ASC"
    )
  )
}
```

```{r raw_read}
nyc <- read_delim(
  file = raw_path,
  delim = ",",
  skip = 1,
  escape_double = FALSE,
  escape_backslash = FALSE,
  locale = locale(date_format = "%m/%d/%Y"),
  col_types = cols(
    .default = col_character(),
    `TRANSACTION AMOUNT` = col_double(),
    `START DATE` = col_date(),
    `END DATE` = col_date(),
    `TRANSACTION APPROVED/FILED DATE` = col_date()
  )
)
```

```{r raw_rename}
nyc <- nyc %>% 
  clean_names("snake") %>% 
  rename_all(str_remove, "transaction_") %>% 
  rename_all(str_remove, "_\\w+") %>% 
  mutate_at(vars(type), str_remove, "\\sContract$")
```

## Explore

```{r glimpse}
head(nyc)
tail(nyc)
glimpse(sample_n(nyc, 20))
```

The only variables missing from any records are the `start` and `end` dates.

```{r na_count}
col_stats(nyc, count_na)
```

There are also a handful of duplicate records, which we can flag.

```{r dupe_flag}
nyc <- flag_dupes(nyc, everything(), .check = TRUE)
sum(nyc$dupe_flag)
```

```{r dupe_view}
nyc %>% 
  filter(dupe_flag) %>% 
  select(start, department, amount, vendor)
```

Most columns are unique strings, but the `type` is pretty evenly split between
original data and amendments. We will include all types.

```{r n_distinct}
col_stats(nyc, n_distinct)
```

```{r bar-type}
explore_plot(nyc, type)
```

### Amounts

```{r ammount_summary}
enframe(map_chr(summary(nyc$amount), dollar))
mean(nyc$amount <= 0)
median(nyc$amount)
```

Here is the largest contract.

```{r amount_max}
glimpse(nyc[which.min(nyc$amount), ])
```

```{r hist_amount, echo=FALSE}
nyc %>%
  filter(amount > 100) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 30) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(2:8)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New York Contracts Amount Distribution",
    subtitle = "from 2012 to 2019",
    caption = "Source: NYS Comptroller",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add a 4 digit `year` variable taken from the approval date.

```{r date_year}
nyc <- mutate(nyc, year = year(approved))
```

There are a handful of dates before 2012, which is where the majority of the 
records begin.

```{r date_range}
min(nyc$approved)
mean(nyc$year < 2012)
max(nyc$approved)
sum(nyc$approved > today())
```

```{r bar_year, echo=FALSE}
nyc %>% 
  filter(year > 2006) %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Contracts by Year",
    caption = "Source: NYS Comptroller",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

There are no geographic variables, but we can add a 2-digit state abbreviation
for the spending agency. 

```{r state_add}
nyc <- mutate(nyc, state = "NY", .after = department)
```

## Update

```{r date_new}
old_date <- as.Date("2020-05-29")
nyc <- nyc %>% 
  filter(approved > old_date)
```

## Export

```{r clean_glimpse}
glimpse(sample_n(nyc, 20))
```

1. There are `r comma(nrow(nyc))` records in the database.
1. There are `r comma(sum(nyc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem mostly reasonable.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

```{r clean_write}
clean_dir <- dir_create(here("state", "ny", "contracts", "data", "clean"))
csv_dates <- paste(str_remove_all(c(old_date, Sys.Date()), "-"), collapse = "-")
clean_csv <- path(clean_dir, "ny_contracts_clean.csv")
write_csv(nyc, clean_csv, na = "")
(clean_size <- file_size(clean_csv))
mutate(file_encoding(clean_csv), across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
