---
title: "Consolidated Screening List"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "ofac_screen", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tabulizer, # read pdf tables
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("us/ofac_screen/docs/screening_list_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

A copy of the Consolidated Screening List (CSL) can be obtained from from the
[U.S. Treasury website.][csl]

> The Consolidated Screening List (CSL) is a list of parties for which the
United States Government maintains restrictions on certain exports, reexports,
or transfers of items. Below, under "Tools" are links to the CSL search engine,
downloadable CSL files, and the CSL Application Programming Interface (API), all
consisting of the consolidation of multiple export screening lists of the
Departments of Commerce, State, and Treasury.

We should note here the considerations listed on the Treasury website:

> ### Key Considerations:
> * These tools are a consolidation of multiple export screening lists of the
    Departments of Commerce, State, and the Treasury and may be used as an aid 
    to industry in conducting electronic screens of potential parties to 
    regulated transactions.
> * In the event that a company, entity, or person on the list appears to match
    a party potentially involved in your export transaction, additional due
    diligence should be conducted before proceeding.
> * There may be a strict export prohibition, requirement for seeking a license
    application, evaluation of the end-use or user to ensure it does not result 
    in an activity prohibited by any U.S. export regulations, or other 
    restriction.
> * Prior to taking further action, to ensure full compliance with all of the
    terms and conditions of the restrictions placed on the parties on this list, 
    the user must check the official publication of restricted parties in the 
    Federal Register. They should also check the official lists of restricted 
    parties maintained on the websites of the Departments of Commerce, State, 
    and the Treasury.

[csl]: https://www.trade.gov/consolidated-screening-list

## Download

We can directly download a copy of the list in tab-separated format.

> #### CSL Downloadable Files
> In the CSL downloadable files, there is a column which is titled "source."
This column indicates which specific federal agency’s proscribed party screening
list is the source for each entry on the spreadsheet. Blank data fields in the
CSL downloadable file are not applicable to the consolidated screening list in
the "source" column.
> * Instructions for Downloading the Consolidated Screening List 
> * CSV Download 
> * TSV Download (ASCII encoded)   
> * JSON

From the above instructions file, there is information on the format.

> #### Data Specifications:
> Each file contains 27 columns – one for each field. The first row consists of
the field names. Please note that some of the field names [e.g., source,
entity_number] are longer than the data length of those fields.

> #### Recent Changes to the Files:
> On July 1, 2015, both the CSV and TSV files have been updated to accommodate
changes in how the data is organized. Several of the field names have been
updated and some of the fields contain additional information. For example, the
City, State/Province, Postal Code, and Country have been combined into one
addresses field. If an entity has more than one Address, then they are all
included in the one field separated by semi-colons. This is true for alternate
names as well. Because multiple addresses and alternate names have been
combined, there are far fewer rows in the CSV/TSV files than previously.

```{r raw-dir}
raw_url <- "http://api.trade.gov/static/consolidated_screening_list/consolidated.tsv"
raw_dir <- dir_create(here("us", "ofac_screen", "data", "raw"))
raw_tsv <- path(raw_dir, basename(raw_url))
```

```{r raw-download}
if (!file_exists(raw_tsv)) {
  download.file(raw_url, raw_tsv)
}
```

## Dictionary

We can get a dictionary of columns from the instructions PDF.

```{r echo=FALSE}
dict <- file_temp(ext = "pdf")
download.file("https://legacy.trade.gov/CSL_Download_Instructions.pdf", dict)
d <- extract_tables(
  file = dict,
  pages = 2:3,
  columns = list(c(121, 247))
)
d[[1]] <- d[[1]] %>% 
  as_tibble() %>% 
  row_to_names(1) %>% 
  clean_names() %>% 
  na_if("") %>% 
  fill(column, field_name) %>% 
  group_by(column, field_name) %>% 
  summarise(description = paste(description, collapse = " "))
d[[2]] <- setNames(as_tibble(d[[2]]), names(d[[1]]))
d[[2]]$description <- str_replace_all(d[[2]]$description, "\r", " ")
d <- bind_rows(d[[1]], d[[2]])
d %>% 
  mutate(across(field_name, md_code)) %>% 
  kable()
```

## Read

```{r raw-read}
csl <- read_delim(
  file = raw_tsv,
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character()
  )
)
```

```{r clean-names}
csl <- clean_names(csl, case = "snake")
```

## Explore

There are `r comma(nrow(csl))` rows of `r ncol(csl)` columns. Each record
represents a single name of a person with whom the U.S. restricts business.

```{r glimpse}
glimpse(csl)
tail(csl)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(csl, count_na)
```

No rows are missing the `name` of the person or vehicle, but since the data is
so varied there is no consistency in other variables.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
csl <- flag_dupes(csl, -id)
sum(csl$dupe_flag)
```

There are no duplicate values in the data.

### Categorical

```{r distinct-count}
col_stats(csl, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(csl, type)
csl %>% 
  mutate(source = str_extract(source, "\\((.*)\\)")) %>% 
  explore_plot(source)
```

## Pivot

The original format of the CSL lists a single _primary_ name in the `name`
column and then a list of many _other_ names in the `alt_names` column. For our
purposes, we are interested in making these alternative names easily searchable.
When a person is hoping to find results on a given individual, we want to make
it possible for _any_ of their known names to return information.

We can do this by un-nesting each of the alternate names into their own copy of 
the original row, with the same additional information.

```{r alt-count}
sum(str_count(csl$alt_names, ";"), na.rm = TRUE)
```

```{r alt-unnest}
csl <- csl %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  unite(
    col = all_names, 
    name, alt_names,
    sep = "; ",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  mutate(across(all_names, str_split, ";\\s")) %>% 
  unnest(
    cols = all_names,
    keep_empty = TRUE
  )
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(csl, 1000))
```

1. There are `r comma(nrow(csl))` records in the database.
1. There are `r comma(sum(csl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(csl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean-dir}
clean_dir <- dir_create(here("us", "ofac_screen", "data", "clean"))
clean_csv <- path(clean_dir, glue("ofac_screening_list.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(csl, clean_csv, na = "")
write_rds(csl, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict-make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(csl)),
  type = md_code(map_chr(csl, typeof)),
  def = c(
    ""
  )
)
```

```{r dict-md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
