---
title: "Idaho Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("id", "lobbying", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

> All registered lobbyists must file financial reports showing the totals of all expenditures made
by the lobbyist or incurred on behalf of such lobbyist‘s employer (not including payments made
directly to the lobbyist) during the period covered by the report, the totals are segregated
according to financial category i.e., entertainment, food and refreshment, advertising, living
accommodations, travel, telephone, office expenses, and other expenses or services.

### Download

```{r create_raw_dir}
raw_dir <- here("id", "lobbying", "reg", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw}
idlr_base <- "https://sos.idaho.gov/elect/lobbyist"
idlr_urls <- c(
  "https://sos.idaho.gov/elect/lobbyist/2008/08lob_exp.xls",
  "https://sos.idaho.gov/elect/lobbyist/2009/2009_lobbyist_exp.xls",
  "https://sos.idaho.gov/elect/lobbyist/2010/2010_lobbyist_exp.xls",
  "https://sos.idaho.gov/elect/lobbyist/2011/2011_Lob_Exp.xls",
  "https://sos.idaho.gov/elect/lobbyist/2012/2012LobbyistExpensesAnnual.xls",
  "https://sos.idaho.gov/elect/lobbyist/2013/2013LobbyistExpensesAnnual.xlsx",
  "https://sos.idaho.gov/elect/lobbyist/2014/2014LobbyistExpensesAnnual.xlsx",
  "https://sos.idaho.gov/elect/lobbyist/2015/LobbyistExpensesAnnual-2015.xlsx",
  "https://sos.idaho.gov/elect/lobbyist/2016/LobbyistExpensesAnnual-2016.xlsx",
  "https://sos.idaho.gov/elect/lobbyist/2017/LobbyistExpensesAnnual-2017.xlsx",
  "https://sos.idaho.gov/elect/lobbyist/2018/Sorted-By-Lobbyist-2018.xlsx"
)

for (url in idlr_urls) {
  download.file(
    url = url,
    destfile = url2path(url, raw_dir)
  )
}
```

### Read

```{r read_raw}
idlr <- map_dfr(
  .x = dir_ls(raw_dir),
  .f = read_excel,
  col_types = "text"
)
```

