---
title: "District Of Columbia Businesses"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "dc", "businesses", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/dc/businesses/docs/dc_businesses_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

> All applicants for a Basic Business License operating from a commercial
location in the District of Columbia must provide a Certificate of Occupancy (C
of O) for the premise address from which the business activity is conducted in
order to demonstrate the activity does not conflict with building and zoning
codes. A certificate of occupancy is needed to occupy any structure other than a
single family dwelling. To include the following uses: two family flat,
apartment house and all commercial uses.

* `OBJECTID` (OID, OBJECTID)
* `PERMIT_NUMBER` (String, Permit Number, length: 30)
* `ISSUE_DATE` (Date, Issue Date, length: 8)
* `ADDRESS` (String, Address, length: 200)
* `WARD` (String, Ward, length: 10)
* `DESCRIPTION_OF_OCCUPANCY` (String, Description of Occupancy, length: 2000)
* `PERMISSION_GRANTED_TO` (String, Permission Granted To, length: 200)
* `TRADING_AS` (String, Trading As, length: 200)
* `FLOORS_OCCUPIED` (String, Floors Occupied, length: 200)
* `OCCUPANCY_LOAD` (String, Occupancy Load, length: 200)
* `PROPERTY_OWNER` (String, Property Owner, length: 200)
* `OWNER_ADDRESS` (String, Owner Address, length: 200)
* `B_Z_A_NUMBER` (String, Board of Zoning Adjustment Number, length: 50)
* `APPROVED_BUILDING_CODE_USE` (String, Approved Building Code Use, length: 50)
* `APPROVED_ZONING_USE` (String, Approved Zoning Use, length: 200)
* `EXPIRATION_DATE` (Date, Expiration Date, length: 8)
* `MAR_ID` (Double, Master Address Repository ID)
* `X_COORDINATE` (Double, X Coordinate)
* `Y_COORDINATE` (Double, Y Coordinate)
* `LATITUDE` (Double, Latitude)
* `LONGITUDE` (Double, Longitude)
* `DCS_LAST_MOD_DTTM` (Date, Database Last Modified Date, length: 8)
* `SHAPE` (Geometry, SHAPE)
* `APPLICATION_COMMENTS` (String, Application Comments, length: 4000)

## Download

```{r raw-dir}
raw_dir <- dir_create(here("state", "dc", "businesses", "data", "raw"))
raw_csv <- path(raw_dir, "Certificate_of_Occupancy.csv")
```

## Read

```{r raw-read}
dcb <- read_delim(
  file = raw_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  locale = locale(date_format = "%Y/%m/%d %H:%M:%S+%z"),
  col_types = cols(
    .default = col_character(),
    X = col_double(),
    Y = col_double(),
    ISSUE_DATE = col_date(),
    EXPIRATION_DATE = col_date(),
    X_COORDINATE = col_double(),
    Y_COORDINATE = col_double(),
    LATITUDE = col_double(),
    LONGITUDE = col_double(),
    DCS_LAST_MOD_DTTM = col_date()
  )
)
```

```{r clean-names}
dcb <- clean_names(dcb, case = "snake")
```

```{r}
dcb <- dcb %>% 
  select(-x, -y)
```

## Explore

There are `r comma(nrow(dcb))` rows of `r ncol(dcb)` columns. Each record
represents a single business registered in the District.

```{r glimpse}
glimpse(dcb)
tail(dcb)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(dcb, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("issue_date", "property_owner", "address")
dcb <- flag_na(dcb, all_of(key_vars))
sum(dcb$na_flag)
```

```{r na-view}
dcb %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
dcb <- flag_dupes(dcb, -objectid)
sum(dcb$dupe_flag)
```

There are no duplicate rows in this data.

### Categorical

```{r distinct-count}
col_stats(dcb, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(dcb, ward)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
dcb <- mutate(dcb, issue_year = year(issue_date))
```

```{r date-range}
min(dcb$issue_date)
sum(dcb$issue_year < 2000)
max(dcb$issue_date)
sum(dcb$issue_date > today())
```

```{r bar-year, echo=FALSE}
dcb %>% 
  count(issue_year) %>% 
  ggplot(aes(x = issue_year, y = n)) +
  geom_col(aes(fill = n), color = "black") + 
  scale_fill_viridis_c(option = "C", end = 0.8, guide = "none") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "District Of Columbia Businesses by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

Because all of these addresses are in DC by their very nature, we can use that
information to separate the very messy address string into it's component parts
by manually adding the `city` and `state` and removing those values from the
address. We will look more carefully for any kid of ZIP code and then clean it
into the 5-digit version.

```{r geo-split}
x <- dcb %>% 
  distinct(address) %>% 
  mutate(
    address_sep = str_remove(address, "(?<=\\d)-(?=$)"),
    city_sep = "WASHINGTON",
    state_sep = "DC",
    zip_norm = address_sep %>% 
      str_extract("\\d{5}(?:-?\\d{4})?$") %>% 
      normal_zip(na_rep = TRUE),
    # remove extracted strings from address
    address_sep = address_sep %>% 
      str_remove("(,\\s)?DC") %>% 
      str_remove("(,\\s)?WASHINGTON") %>% 
      str_remove("\\d{5}(?:-?\\d{4})?$") %>% 
      str_trim()
  )
```

```{r}
dcb <- left_join(dcb, x, by = "address")
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(dcb, 1000))
```

1. There are `r comma(nrow(dcb))` records in the database.
1. There are `r comma(sum(dcb$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dcb$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(dcb$issue_date), "-")
max_dt <- str_remove_all(max(dcb$issue_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("state", "dc", "businesses", "data", "clean"))
clean_csv <- path(clean_dir, glue("dc_businesses_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(dcb, clean_csv, na = "")
write_rds(dcb, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
