---
title: "New Jersey Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nj", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tabulizer, # read pdf tables
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data was provided by the [New Jersey Office of the State Comptroller][compt] via
[an open records request][req] filed by MuckRock News, under to the New Jersey
Open Public Records Act, on behalf of the Investigative Reporting Workshop.

The data was provided to MuckRock in the form of a single PDF file that was
then hosted on their website for download.

[compt]: https://www.nj.gov/comptroller/
[req]: https://www.muckrock.com/foi/new-jersey-229/state-contracts-office-of-the-state-comptroller-91636/

## Download

We can download the PDF data from MuckRock for processing.

```{r raw_download}
raw_dir <- dir_create(here("nj", "contracts", "data", "raw"))
pdf_url <- "https://cdn.muckrock.com/foia_files/2020/05/08/Morisy_Response.pdf"
pdf_path <- path(raw_dir, basename(pdf_url))
download.file(pdf_url, pdf_path)
```

## Read

Reading this data is not a very simple process. There are three different table
formats in the single PDF file:

1. 14-column, whitespace separated table on pages 1 through 9
2. 13-column, cell border separated table on pages 10 through 27
3. 15-column, cell border separated table on pages 28 through 65

The three table types take up different dimensions on the page. We can use
[Tabula](https://tabula.technology/) to identify the tables on the PDF pages
and convert them to CSV files that can be easily read as data frames and 
combined into a single file.

### Table 1

For the first table, we will use the `tabulizer` package to extract the
whitespace separated tables as a CSV file.

```{r t1_extract}
t1_dir <- dir_create(path_temp("t1"))
extract_tables(
  file = pdf_path,
  pages = 1:9,
  guess = TRUE,
  output = "tsv",
  method = "stream",
  outdir = t1_dir
)
```

The table extraction results in `r length(dir_ls(t1_dir))` CSV files.

```{r t1_list}
t1_paths <- dir_ls(t1_dir)
md_list(md_code(t1_paths))
```

These CSV files can be read as data frames in a list. By reading each data frame
without column headers, we can then convert the initial row to a single header
and combined the subsequent data frame list elements.

```{r t1_read}
t1 <- t1_paths %>% 
  map_df(
    read_tsv,
    col_names = FALSE,
    col_types = cols(.default = "c")
  ) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names(case = "snake") %>% 
  type_convert(
    col_types = cols(
      .default = col_character(),
      creation_date = col_date("%m/%d/%y"),
      budget_fiscal_year = col_integer(),
      quantity = col_number(),
      unit_price_amount = col_number()
    )
  )
```

```{r t1_rename, echo=FALSE}
t1 <- rename(
  .data = t1,
  fy = budget_fiscal_year,
  date = creation_date,
  po = purchase_order_number,
  vendor = vendor_name,
  contact = vendor_contact,
  addr = address_line_1,
  state = state_code,
  zip = zip_code,
  price = unit_price_amount,
  line = order_line_number,
  desc1 = order_line_description1,
  desc2 = order_line_description2
)
```

```{r t1_head}
head(t1)
```

### Table 2

Table two is in a different structure, with columns not in a fixed position
separated by whitespace but rather contained within a cell border line. We can
extract these tables using Tabula's "lattice" method.

```{r t2_extract}
t2_dir <- dir_create(path_temp("t2"))
extract_tables(
  file = pdf_path,
  pages = 10:27,
  guess = TRUE,
  output = "tsv",
  method = "lattice",
  outdir = t2_dir
)
```

```{r t2_list}
t2_paths <- dir_ls(t2_dir)
```

```{r t2_read}
t2 <- t2_paths %>% 
  map(
    read_tsv, 
    col_names = FALSE, 
    col_types = cols(.default = "c")
  ) %>% 
  map(remove_empty, "cols") %>% 
  map_df(set_names, LETTERS[1:13]) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names("snake") %>% 
  type_convert(
    col_types = cols(
      .default = col_character(),
      sent_date = col_date("%b %d, %Y"),
      quantity = col_number(),
      unit_cost = col_number()
    )
  )
```

```{r t2_rename, echo=FALSE}
t2 <- t2 %>% 
  select(
    -org_name,
    -department_name
  ) %>% 
  rename(
    po = po_number,
    date = sent_date,
    desc1 = item_description,
    price = unit_cost,
    vendor = vendor_name,
    id = vendor_nbr,
    addr = address_line_1
  )
```

```{r t2_head}
head(t2)
```

### Table 3

```{r t3_extract}
t3_dir <- dir_create(path_temp("t3"))
extract_tables(
  file = pdf_path,
  pages = 28:65,
  guess = TRUE,
  output = "tsv",
  method = "lattice",
  outdir = t3_dir
)
```

```{r t3_list}
t3_paths <- dir_ls(t3_dir)
```

```{r t3_read}
t3 <- t3_paths %>% 
  map_df(
    read_tsv,
    col_names = FALSE,
    col_types = cols(.default = "c")
  ) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names(case = "snake") %>% 
  mutate(across(
    .cols = where(is.character), 
    .fns = ~str_replace(., "\r", " ")
  )) %>% 
  type_convert(
    col_types = cols(
      .default = col_character(),
      sent_date = col_date("%b %d, %Y"),
      quantity = col_number(),
      unit_cost = col_number()
    )
  )
```

```{r t3_rename, echo=FALSE}
t3 <- t3 %>% 
  select(
    -org_name,
    -department_name
  ) %>% 
  rename(
    po = po_number,
    date = sent_date,
    desc1 = item_description,
    price = unit_cost,
    vendor = vendor_name,
    id = vendor_nbr,
    addr = address_line_1,
    contact = contact_name,
    email = email_address
  )
```

```{r t3_head}
head(t3)
```

### Bind

Once each of the table types is read into separate data frames, we can bind them
all into a single data frame once column names are aligned.

```{r raw_bind}
njc <- bind_rows(
  a = t1,
  b = t2,
  c = t3,
  .id = "type"
)
```

```{r raw_case}
njc <- mutate(
  .data = njc,
  across(
  .cols = where(is.character),
  .fns = str_to_upper
  )
)
```

The three tables stem from format changes between years.

```{r raw_year_count}
count(njc, type, year = year(date))
```

Once bound together, the differences between existing columns creates some
columns filled with `NA` values for those records belonging to a table type
without said column.

## Explore

```{r glimpse}
glimpse(njc)
tail(njc)
```

### Missing

The only missing values stem from the column differences so no individual
records need to be flagged.

```{r na_count}
col_stats(njc, count_na)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag, warning=TRUE}
njc <- flag_dupes(njc, everything())
```

### Categorical

```{r distinct_count}
col_stats(njc, n_distinct)
```

```{r distinct_plots}
explore_plot(njc, type)
explore_plot(njc, vendor) + scale_x_truncate()
```

### Amounts

```{r ammount_summary}
summary(njc$price)
mean(njc$price <= 0)
```

Most contracts are for items with a relatively low unit price.

```{r hist_amount, echo=FALSE}
njc %>%
  ggplot(aes(price)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New Jersey Contracts Amount Distribution",
    caption = "Source: NJ Comptroller",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_max}
glimpse(njc[which.max(njc$price),])
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
njc <- mutate(njc, year = year(date))
```

```{r date_range}
min(njc$date)
sum(njc$year < 2000)
max(njc$date)
sum(njc$date > today())
```

```{r bar_year, echo=FALSE}
njc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2015:2020) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Jersey Contracts by Year",
    caption = "Source: NJ Comptroller",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
njc <- mutate(
  .data = njc,
  addr_norm = normal_address(
    address = addr,
    abbs = usps_street,
    na_rep = TRUE
  )
)
```

```{r address_view}
njc %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

The ZIP code does not need to be normalized.

```{r zip_check}
percent(prop_in(njc$zip, valid_zip))
```

### State

Only one small change needs to be made to the `state` variable.

```{r state_check}
percent(prop_in(njc$state, valid_state), 0.001)
```

```{r state_compare}
njc %>% 
  filter(state %out% valid_state) %>% 
  count(state, zip) %>% 
  left_join(zipcodes, by = "zip")
```

```{r state_fix}
njc <- mutate(njc, across(state, str_remove_all, "\\s"))
percent(prop_in(njc$state, valid_state), 0.001)
```

### City

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_pre}
many_city <- c(valid_city, extra_city)
percent(prop_in(njc$city, many_city), 0.01)
```

```{r city_norm}
njc <- mutate(
  .data = njc,
  city_norm = normal_city(
    city = city, 
    abbs = usps_city,
    states = c("NJ", "DC", "NEW JERSEY"),
    na = invalid_city,
    na_rep = TRUE
  )
)
```

```{r city_out_count}
njc %>% 
  filter(city_norm %out% many_city) %>% 
  count(city_norm, state, zip, sort = TRUE)
```

```{r city_replace}
njc <- mutate(
  .data = njc,
  city_norm = city_norm %>% 
    str_replace("^NY$", "NEW YORK") %>% 
    str_replace("(?<=\\s)PA$", "PARK") %>% 
    str_replace("(?<=\\s)VL$", "VALLEY") %>% 
    str_replace("(?<=\\s)LAK$", "LAKE") %>% 
    str_replace("^SOMSERSET$", "SOMERSET")
)
```

```{r city_post}
percent(prop_in(njc$city_norm, many_city), 0.01)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
njc <- rename_all(njc, ~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(njc, 20))
```

1. There are `r comma(nrow(njc))` records in the database.
1. There are `r comma(sum(njc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `price` and `date` seem reasonable.
1. There are `r comma(sum(njc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nj", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "nj_contracts_clean.csv")
write_csv(njc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(njc)),
  type = md_code(map_chr(njc, typeof)),
  def = c(
    "PDF Table type (year format)",
    "Fiscal year",
    "Date sent or created",
    "Purchase order number",
    "Vendor name",
    "Vendor contact name",
    "Vendor street address",
    "Vendor city name",
    "Vendor state abbreviation",
    "Vendor state 5-digit ZIP code",
    "Quantity of item purchased",
    "Unit price of item",
    "Order item line number",
    "Item description",
    "Secondary description",
    "Unique vendor ID",
    "Vendor contact email",
    "Calendar year of purchase",
    "Normalized street address",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
