---
title: "Florida Campaign Contributions Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("fl", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

## Download
Files (in Tab-delimited .txt) were downloaded from [Florida Department of State's Campaign Finance Database](https://dos.elections.myflorida.com/campaign-finance/contributions/#both):
The Accountability Project already hosts data from 01/01/2008 to 08/01/2020. We'll set the date range to 

```{r raw_dir}
raw_dir <- dir_create(here("fl", "contribs", "data", "raw"))
```


## Read
We can see that sometimes quotes are not closed and thus cause confusion for R to parse. After examining the use of quotes, we can see that they don't really  Therefore, we can safely remove the quote
```{r raw_read}
flc <- read_lines(file = dir_ls(raw_dir))
flc <- flc %>% str_remove_all('"') %>% read_delim(delim = "\t",
                                                  col_types = cols(.default = col_character(),
                                                                    Amount = col_double()))
```

To keep consistency with TAP's previous data structure, we will rename some columns. 
```{r rename}
flc <- flc %>% 
  rename(Candidate_Committee = `Candidate/Committee`,
         ContribName=`Contributor Name`,
         City_State_Zip=`City State Zip`,
         InkindDesc = `Inkind Desc`,
         Type = Typ)
```

## Explore

```{r glimpse}
glimpse(flc)
tail(flc)
```

### Missing

```{r na_count}
col_stats(flc, count_na)
```

```{r na_flag}
flc <- flc %>% flag_na(Address, ContribName)
sum(flc$na_flag)
```

```{r na_view}
flc %>% 
  filter(na_flag) %>% 
  select(Candidate_Committee,Date, Amount, Type, InkindDesc)
```

### Duplicates

```{r dupe_flag}
flc <- flag_dupes(flc, everything())
sum(flc$dupe_flag)
```

```{r dupe_view}
flc %>% 
  filter(dupe_flag) %>% 
  select(Candidate_Committee,Date, Amount, Type, InkindDesc)
```

### Categorical

```{r distinct_count}
col_stats(flc, n_distinct)
```

```{r distinct_plots}
explore_plot(flc, Type)
```

### Amounts

```{r ammount_summary}
summary(flc$amount)
mean(flc$amount <= 0)
```

```{r hist_amount, echo=FALSE}
flc %>%
  ggplot(aes(Amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Florida Campaign Contributions Amount Distribution",
    caption = "Source: FL Secretary of State",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
flc$Date <- as.Date(flc$Date,format = "%m/%d/%Y")
flc <- mutate(flc, YEAR = year(Date))
```

```{r date_range}
min(flc$date)
sum(flc$year < 2000)
max(flc$date)
sum(flc$date > today())
```

```{r bar_year, echo=FALSE}
flc %>% 
  count(YEAR) %>% 
  mutate(even = is_even(YEAR)) %>% 
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
flc <- flc %>% 
  mutate(
    Address_norm = normal_address(
      address = Address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
flc %>% 
  select(contains("Address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP
To normalize the City, State and ZIP fields, we will need to separate them first. 
```{r}
flc <- flc %>% 
  mutate(Zip= str_extract(City_State_Zip, "\\d{2,}"))
```

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
flc <- flc %>% 
  mutate(
    ZIP5 = normal_zip(
      zip = Zip,
      na_rep = TRUE
    )
  ) 
```

```{r zip_progress}
progress_table(
  flc$Zip,
  flc$ZIP5,
  compare = valid_zip
)
```

### State
```{r sep state}
flc <- flc %>% 
  mutate(State = str_extract(City_State_Zip, "(?<=,\\s)[A-Z]{2}(?=\\s)"))

flc$State[str_detect(flc$City_State_Zip,"F 3\\d*")] <- "FL"
```


Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
flc <- flc %>% 
  mutate(
    State_norm = normal_state(
      state = State,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
flc %>% 
  filter(State != State_norm) %>% 
  count(State, sort = TRUE)
```

```{r state_progress}
progress_table(
  flc$State,
  flc$State_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
```{r sep city}
flc <- flc %>% 
  mutate(City = str_extract(City_State_Zip,".+(?=,\\s)"))
```


#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
flc <- flc %>% 
  mutate(
    City_norm = normal_city(
      city = City, 
      abbs = usps_city,
      states = c("FL", "DC", "FLORIDA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
flc <- flc %>% 
  rename(City_raw = City) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "State_norm" = "state",
      "ZIP5" = "zip"
    )
  ) %>% 
  rename(City_match = city) %>% 
  mutate(
    match_abb = is_abbrev(City_norm, City_match),
    match_dist = str_dist(City_norm, City_match),
    City_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = City_match,
      false = City_norm
    )
  ) %>% 
  select(
    -City_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- flc %>% 
  mutate(
    City_refine = City_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(City_refine != City_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "City_refine" = "city",
      "State_norm" = "state",
      "ZIP5" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    State_norm, 
    ZIP5,
    City_swap, 
    City_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
flc <- flc %>% 
  left_join(good_refine) %>% 
  mutate(City_refine = coalesce(City_refine, City_swap))
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r check_filter}
flc_out <- flc %>% 
  filter(City_refine %out% c(valid_city, extra_city)) %>% 
  count(City_refine, State_norm, sort = TRUE) %>% 
  drop_na() %>% 
  filter(n > 1)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("fl", "contribs", "data", "api_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file
  )
} else {
  check <- pmap_dfr(
    .l = list(
      flc_out$City_refine, 
      flc_out$State_norm
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    path = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  c(valid_locality)
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city,valid_locality)
progress <- progress_table(
  str_to_upper(flc$City_raw),
  flc$City_norm,
  flc$City_swap,
  flc$City_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(flc$City_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Florida City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
flc <- flc %>% 
  select(
    -City_norm,
    -City_swap,
    -City_raw,
    -State,
    -Zip,
    City_clean = City_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(flc, 20))
```

1. There are `r comma(nrow(flc))` records in the database.
1. There are `r comma(sum(flc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(flc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("fl", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "fl_contribs_clean.csv")
write_csv(flc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload


```{r clean_upload, eval=FALSE}
s3write_using(
  x = flc,
  FUN = write_csv,
  object = "csv/sd_contracts.csv",
  bucket = "publicaccountability",
  na = ""
)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(flc)),
  type = md_code(map_chr(flc, typeof)),
  def = c(
    "Recipient candidate or committee",
    "Date of contribution made",
    "Contribution amount",
    "Contributor type",
    "Contributor name",
    "Contributor street address",
    "Contributor City, State and ZIP code",
    "Contributor occupation",
    "Description of Inkind Contributions",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Florida Contribution Data Dictionary\n", dict_md),
  path = here("fl", "contribs", "fl_contribs_dict.md"),
)
```
