---
title: "Minnesota Contracts"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "mn", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/mn/contracts/docs/mn_contracts_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

Minnesota state contracts were retrieved via a state records requests filed by
students at American University.

## Download

The ZIP archive containing the requested data can be downloaded from the
Investigative Reporting Workshop's server.

```{r raw-dir}
raw_key <- "jla_class_data2021/fwdminnesotadatarequest.zip"
raw_dir <- dir_create(here("state", "mn", "contracts", "data", "raw"))
raw_zip <- path(raw_dir, raw_key)
```

```{r raw-download}
if (!file_exists(raw_zip)) {
  save_object(
    object = raw_key,
    bucket = "publicaccountability",
    file = raw_zip,
    show_progress = TRUE
  )
}
```

```{r}
raw_xls <- unzip(raw_zip, exdir = raw_dir)
```

## Read

The monthly excel files can be read together and combined into a single table.

```{r}
mnc <- map(
  .x = raw_xls, 
  .f = read_excel, 
  skip = 2, 
  col_names = FALSE,
  col_types = "text"
)
```

One file is missing column names, which can be added from another file.

```{r fix-names}
mnc[[4]] <- bind_rows(mnc[[1]][1, -3], mnc[[4]])
mnc <- map(
  .x = mnc,
  .f = function(dat) {
    dat <- dat |> 
      remove_empty("cols") |> 
      row_to_names(1)
    if (is.na(last(dat[[1]]))) {
      dat <- head(dat, -1)
    }
    return(dat)
  }
)
```

```{r bind-rows}
mnc <- mnc |> 
  set_names(basename(raw_xls)) |> 
  bind_rows(.id = "source_file") |>
  relocate(source_file, .after = last_col()) |> 
  clean_names(case = "snake") |> 
  type_convert(
    na = c("", "-"),
    col_types = cols(
      .default = col_character(),
      payment_day_dt = col_date_mdy(),
      payment_amt = col_double()
    )
  )
```

The `supplier` should be separated into two separate values for the ID and name.

```{r supp-split}
mnc <- mnc |> 
  separate(
    col = supplier,
    into = c("supplier_id", "supplier_name"),
    sep = "(?<=\\d)\\s-\\s"
  )
```

## Explore

There are `r comma(nrow(mnc))` rows of `r ncol(mnc)` columns. Each record
represents a single contract between the state and a vendor.

```{r glimpse}
glimpse(mnc)
tail(mnc)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(mnc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("payment_day_dt", "payment_amt", "supplier_name")
mnc <- flag_na(mnc, all_of(key_vars))
sum(mnc$na_flag)
```

```{r na-view}
# remove if no missing
if (sum(is.na(mnc$na_flag)) == 0) {
  mnc <- select(mnc, -na_flag)
}
```

There are no rows missing key values.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
mnc <- flag_dupes(mnc, everything())
sum(mnc$dupe_flag)
```

There are now duplicate rows.

### Categorical

```{r distinct-count}
col_stats(mnc, n_distinct)
```

### Amounts

```{r amount-round}
# fix floating point precision
mnc$payment_amt <- round(mnc$payment_amt, digits = 2)
```

```{r amount-summary}
summary(mnc$payment_amt)
mean(mnc$payment_amt <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(mnc[c(which.max(mnc$payment_amt), which.min(mnc$payment_amt)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
mnc %>%
  ggplot(aes(payment_amt)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Minnesota Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`.

```{r date-year}
mnc <- mutate(mnc, payment_day_yr = year(payment_day_dt))
```

```{r date-range}
min(mnc$payment_day_dt)
sum(mnc$payment_day_yr < 2000)
max(mnc$payment_day_dt)
sum(mnc$payment_day_dt > today())
```

All of these contracts are from the 12 months between November 2020 and October
2021.

```{r bar-year, echo=FALSE}
mnc %>% 
  count(y = payment_day_yr, m = month(payment_day_dt, TRUE, TRUE)) |> 
  mutate(lbl = as_factor(paste(y, m, sep = "."))) %>% 
  ggplot(aes(x = lbl, y = n)) +
  geom_col() + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Minnesota Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
mnc <- mnc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = supplier_address_postal_cd,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  mnc$supplier_address_postal_cd,
  mnc$zip_norm,
  compare = valid_zip
)
```

### State

The existing `supplier_address_state` column is already normalized. All invalid
abbreviations are from other countries.

```{r state-check}
prop_in(mnc$supplier_address_state, valid_state)
unique(what_out(mnc$supplier_address_state, valid_state))
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- mnc %>% 
  distinct(supplier_address_city, supplier_address_state, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = supplier_address_city, 
      abbs = usps_city,
      states = c("MN", "DC", "MINNESOTA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = supplier_address_city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "supplier_address_state" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
mnc <- left_join(
  x = mnc,
  y = norm_city,
  by = c(
    "supplier_address_city" = "city_raw", 
    "supplier_address_state", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- mnc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "supplier_address_state" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    supplier_address_state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
mnc <- mnc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(mnc$supplier_address_city),
  mnc$city_norm,
  mnc$city_swap,
  mnc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(mnc$supplier_address_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.50, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Minnesota City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Minnesota City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
mnc <- mnc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(city_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(mnc, 1000))
```

1. There are `r comma(nrow(mnc))` records in the database.
1. There are `r comma(sum(mnc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mnc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(mnc$payment_day_dt), "-")
max_dt <- str_remove_all(max(mnc$payment_day_dt), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("state", "mn", "contracts", "data", "clean"))
clean_csv <- path(clean_dir, glue("mn_contracts_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(mnc, clean_csv, na = "")
write_rds(mnc, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
