---
title: "Delaware Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("de", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  jsonlite, # read json files
  janitor, # clean data frames
  batman, # convert to logical
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Delaware contracts data can be obtained via the [Open Data portal][odp] from the
Office of Management and Budget (OMB) and Government Support Services (GSS).

> The Government Support Services (GSS) office within Office of Management of
Budget (OMB) administers statewide central contracts for use by state agencies,
school districts, and municipalities which includes their spend data.

The data was first published on June 15, 2016 and was updated as recently as
May 11, 2020. The data covers July 2012 to present and is updated monthly. Each
record is a single contract:

> Spend with a specific vendor for a specific contract for a specific spend
group (agency, school, or municipality) for the years reported

The OMB and GSS also provide a data dictionary [key as a PDF file][key], which
can also be scraped from the Open Data portal as JSON.

```{r}
meta <- fromJSON("https://data.delaware.gov/api/views/sifm-293u.json")
layout <- as_tibble(meta$columns)
layout %>% 
  select(fieldName, renderTypeName, description) %>% 
  mutate(across(fieldName, md_code)) %>% 
  kable(col.names = c("Field", "Type", "Description"))
```


[odp]: https://data.delaware.gov
[key]: https://data.delaware.gov/api/views/sifm-293u/files/03148284-a0ec-44dc-9c6f-f7571e674aac?download=true&filename=Data%20Dictionary%20for%20State%20Contracts%20-%20GSS%20Spend%20Data.pdf

## Read

We can read the database directly from the DE Open Data portal as a CSV file.

```{r raw_read}
dec <- vroom(
  file = "https://data.delaware.gov/api/views/sifm-293u/rows.csv",
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    `Total Spend` = col_number(),
    # all 1st of month 12:00 AM
    `Month of Spend` = col_character(),
    `Calendar Year` = col_integer(),
    `Calendar Month` = col_integer(),
    `FY` = col_integer()
  )
)
```

The `month_of_spend` variable is presented as a date time column, but every
variable is midnight on the first of the month. That makes this variable
redundant with `calendar_month` and can be removed.

```{r raw_month}
head(dec$month_of_spend, 2)
unique(day(mdy_hms(dec$month_of_spend)))
mean(month(mdy_hms(dec$month_of_spend)) == dec$calendar_month)
dec <- select(dec, -month_of_spend)
```

### Scrape

```{r}
url <- "https://contracts.delaware.gov"
html <- content(GET(url))
tbody <- html_nodes(html, "table")[[13]]
href <- html_attr(html_nodes(tbody, ".fontsmall > a"), "href")
ids <- str_subset(href, "i=\\d+$")
tbl <- as_tibble(html_table(tbody, header = TRUE))
de_scrape <- tbl %>% 
  na_if("N/A") %>% 
  clean_names("snake") %>% 
  mutate(across(ends_with("_date"), mdy)) %>% 
  mutate(across(ends_with("_available"), to_logical)) %>% 
  mutate(id = as.integer(str_extract(ids, "\\d+$")))
```

```{r}
html <- content(GET(paste(url, "contract_archive.asp", sep = "/")))
tbody <- html_nodes(html, "table")[[13]]
href <- html_attr(html_nodes(tbody, ".fontsmall > a"), "href")
ids <- str_subset(href, "i=\\d+$")
tbl <- as_tibble(html_table(tbody, header = TRUE))
archive <- tbl %>% 
  na_if("N/A") %>% 
  clean_names("snake") %>% 
  mutate(across(ends_with("_date"), mdy)) %>% 
  mutate(across(ends_with("_available"), to_logical)) %>% 
  mutate(id = as.integer(str_extract(ids, "\\d+$")))
rm(html, tbody, href, ids, tbl)
```

```{r}
de_scrape <- bind_rows(active = de_scrape, archive = archive, .id = "source")
count(de_scrape, source)
```

```{r}
mean(de_scrape$contract_number %in% dec$long_contract_number)
```

## Explore

```{r glimpse}
glimpse(dec)
tail(dec)
```

### Missing

Only one variable is missing a single value, not enough to justify a flag.

```{r na_count}
col_stats(dec, count_na)
```

```{r na_view}
dec %>% 
  filter(is.na(vendor_name)) %>% 
  select(calendar_month, agency, total_spend, vendor_name)
```

### Duplicates

```{r dupe_flag}
dec <- flag_dupes(dec, -id)
sum(dec$dupe_flag)
```

```{r dupe_view}
dec %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, committee)
```

### Categorical

```{r distinct_count}
col_stats(dec, n_distinct)
```

```{r distinct_plots}
explore_plot(dec, type)
```

### Amounts

```{r ammount_summary}
summary(dec$amount)
mean(dec$amount <= 0)
```

```{r hist_amount, echo=FALSE}
dec %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Delaware Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
dec <- mutate(dec, year = year(date))
```

```{r date_range}
min(dec$date)
sum(dec$year < 2000)
max(dec$date)
sum(dec$date > today())
```

```{r bar_year, echo=FALSE}
dec %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Delaware Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
dec <- dec %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
dec %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
dec <- dec %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  dec$zip,
  dec$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
dec <- dec %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
dec %>% 
  filter(state != state_norm) %>% 
  count(state, sort = TRUE)
```

```{r state_progress}
progress_table(
  dec$state,
  dec$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
dec <- dec %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("DE", "DC", "DELAWARE"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
dec <- dec %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- dec %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
dec <- dec %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(dec$city_raw),
  dec$city_norm,
  dec$city_swap,
  dec$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(dec$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Delaware City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Delaware City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
dec <- dec %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(dec, 20))
```

1. There are `r comma(nrow(dec))` records in the database.
1. There are `r comma(sum(dec$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dec$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("de", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "de_contracts_clean.csv")
write_csv(dec, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] commnand line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(dec)),
  type = md_code(map_chr(dec, typeof)),
  def = c(
    "Date contribution was made",
    "Contribution amount or correction",
    "Contributor name",
    "Contributor unique ID",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Contributor type",
    "Recipient committee name",
    "Recipient unique ID",
    "Contribution method",
    "Recipient election or office sought",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Delaware Contracts Data Dictionary\n", dict_md),
  path = here("de", "contracts", "de_contracts_dict.md"),
)
```
