---
title: "Rhode Island Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  RSelenium, # remote browsing
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & database
  refinr, # cluster and merge
  scales, # number formatting
  knitr, # knit documents
  here, # relative storage
  glue, # combine strings
  fs # search storage 
)

# fix conflict
here <- here::here
```

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data comes from the [Rhode Island Board of Elections][03]. Each records tracks a single expenditure
as reported by the campaign.

[03]: http://www.elections.ri.gov/finance/publicinfo/ "ri_boe"

### About

> The Expenditures tab allows you to run ad-hoc reports on expenditures filed on Summary of
Campaign Activity (CF-2) reports. The reports allow a certain degree of customization in that you
are able to specify certain filters to limit what appears on the reports.

## Import

### Download

```{r raw_dir}
# create a directory for the raw data
raw_dir <- here("state","ri", "expends", "data", "raw")
dir_create(raw_dir)
```

To download the file, we can create our own URL and navigate to the page using the `RSelenium`
package.The last update was `2019-07-30`, and this round of update covers the period from `2019-07-31` to `2022-11-04`. The next update should start from `2022-11-05`.

```{r download_raw, eval=FALSE}
url <- str_c(
  "http://ricampaignfinance.com/RIPublic/Reporting/ExpenditureReport.aspx?OrgID=0",
  "BeginDate=07/31/2019",
  "EndDate=11/04/2022",
  "LastName=",
  "FirstName=",
  "ContType=0",
  "State=",
  "City=",
  "ZIPCode=",
  "Amount=0",
  "ReportType=Expend",
  "CFStatus=F",
  "MPFStatus=F",
  "Level=S",
  "SumBy=Type",
  "Sort1=None",
  "Direct1=asc",
  "Sort2=None",
  "Direct2=asc",
  "Sort3=None",
  "Direct3=asc",
  "Site=Public",
  "Incomplete=A",
  "ContSource=CF",
  sep = "&"
)

# open the driver with auto download options
remote_driver <- rsDriver(
  port = 4440L,
  browser = "firefox",
  extraCapabilities = makeFirefoxProfile(
    list(
      browser.download.dir = raw_dir,
      browser.download.folderList = 2L,
      browser.helperApps.neverAsk.saveToDisk = "text/csv"
    )
  )
)

# navigate to the APOC download site
remote_browser <- remote_driver$client
remote_browser$navigate(url)

# click the export button
remote_browser$findElement("css", "#lnkExport")$clickElement()

# switch to pop up window
pop_up <- remote_driver$client$getWindowHandles()[[2]]
remote_driver$client$switchToWindow(windowId = pop_up)

# click the download option button
csv_button <- remote_browser$findElement("css", "#hypFileDownload")$clickElement()

# close the browser and driver
remote_browser$close()
remote_driver$server$stop()
```

### Read

```{r read_raw}
ri <- read_csv(
  file = dir_ls(raw_dir),
  col_types = cols(
    .default = col_character(),
    ExpDate = col_date("%m/%d/%Y"),
    Amount = col_double(),
    OSAP =  col_double(),
    ZeroedByCF7 = col_logical()
  )
)

ri <- ri %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper) %>% 
  mutate_if(is_character, str_squish)
```

## Explore

There are `r comma(nrow(ri))` records of `r comma(length(ri))` variables in the full database.

```{r glimpse}
glimpse(sample_frac(ri))
```

### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
col_stats(ri,n_distinct)
```

```{r plot_exp_bar, echo=FALSE}
ggplot(data = ri) + 
  geom_bar(mapping = aes(exp_desc)) +
  coord_flip()
```

```{r plot_rec_bar, echo=FALSE}
ggplot(data = ri) + 
  geom_bar(mapping = aes(receipt_desc)) +
  coord_flip()
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
ri %>% glimpse_fun(count_na)
```

We will flag any record missing a `full_name` or `address`.

```{r flag_na}
ri <- ri %>% mutate(na_flag = is.na(full_name) | is.na(address) | is.na(city_st_zip))
```

### Duplicates

We can use `janitor::get_dupes()` to create a table of records duplicated more than once.

```{r get_dupes, collapse=TRUE}
ri_dupes <- distinct(get_dupes(ri))
nrow(ri_dupes)
sum(ri_dupes$dupe_count)
```

We can then join this table back with the full dataset, flagging any duplicated rows with
`dupe_flag()`.

```{r join_dupes}
ri <- ri %>%
  left_join(ri_dupes) %>% 
  mutate(dupe_flag = !is.na(dupe_count))

rm(ri_dupes)
```

### Ranges

#### amounts

There are `r sum(ri$amount < 0)` records with `amount` values less than zero. 

```{r range_amount, collapse=TRUE}
summary(ri$amount)
sum(ri$amount < 0)
```

```{r amount_hist}
ri %>% 
  ggplot(aes(amount)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10", labels = dollar)
```

```{r amount_box_type, fig.height=10}
ri %>% 
  ggplot(aes(exp_desc, amount)) +
  geom_boxplot(varwidth = TRUE, outlier.alpha = 0.01) +
  scale_y_continuous(trans = "log10", labels = dollar) +
  coord_flip()
```

### Dates

There are `r sum(ri$exp_date > today())` records with an `exp_date` past `r today()`.

```{r range_date, collapse=TRUE}
summary(ri$exp_date)
sum(ri$exp_date > today())
```

```{r year_bar, echo=FALSE}
ri %>% 
  ggplot() +
  geom_bar(aes(year(exp_date)))
```

```{r month_line_desc, echo=FALSE}
ri %>% 
  group_by(exp_desc, month = month(exp_date)) %>% 
  summarize(median_amount = median(amount)) %>% 
  ggplot(aes(month, median_amount)) +
  geom_line(aes(color = exp_desc), size = 2) +
  scale_y_continuous(labels = scales::dollar)
```

## Wrangle

Before wrangling the data, we need to separate the `city_st_zip` variable into it's respective
parts.

```{r sep_geo}
ri <- ri %>% 
  separate(
    col = city_st_zip,
    into = c("city_sep", "state_zip"),
    sep = ",\\s",
    remove = FALSE
  ) %>% 
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s",
    remove = TRUE
  )
```

### Year

We can create an `exp_year` variable from the `exp_date` using `lubridate::year()` (after parsing
the string with `readr::col_date()`).

```{r add_year}
ri <- mutate(ri, exp_year = year(exp_date))
```

### Address

The `address` variable should be minimally cleaned using the `campfin::normal_address()` to
simplify text and expand abbreviations.

```{r norm_address}
ri <- ri %>% 
  mutate(
    address_clean = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )

ri %>% 
  filter(address_clean != address) %>% 
  select(address, address_clean) %>% 
  sample_n(10)
```

### ZIP

We can do the same to ZIP codes using `campfin::normal_zip()`.

```{r norm_zip, collapse=TRUE}
n_distinct(ri$zip_sep)
mean(na.omit(ri$zip_sep) %in% valid_zip)

ri <- ri %>% 
  mutate(
    zip_clean = normal_zip(
      zip = zip_sep, 
      na_rep = TRUE
    )
  )

mean(is.na(ri$zip_clean))
n_distinct(ri$zip_clean)
mean(na.omit(ri$zip_clean) %in% valid_zip)
n_distinct(ri$zip_clean[which(ri$zip_clean %out% valid_zip)])
```

### State

```{r norm_state, collapse=TRUE}
n_distinct(ri$state_sep)
mean(na.omit(ri$state_sep) %in% valid_state)
setdiff(ri$state_sep, valid_state)

ri <- ri %>% 
  mutate(
    state_clean = normal_state(
      state = state_sep,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )

n_distinct(ri$state_clean)
mean(na.omit(ri$state_clean) %in% valid_state)
setdiff(ri$state_clean, valid_state)
```

### City

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Normalize

```{r count_city, collapse=TRUE}
n_distinct(ri$city_sep)
mean(ri$city_sep %in% valid_city)
sum(unique(ri$city_sep) %out% valid_city)
```

```{r prep_city, collapse=TRUE}
norm_city <- ri %>% 
  distinct(city_sep, state_clean, zip_clean) %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep %>% str_replace("\\bPROV\\b", "PROVIDENCE"), 
      abbs = usps_city,
      states = c("RI", "DC", "RHODE ISLAND"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
n_distinct(ri$city_norm)
mean(ri$city_norm %in% valid_city)
sum(unique(ri$city_norm) %out% valid_city)
```

#### Swap
We can further improve normalization by comparing our normalized value against the expected value for that record’s state abbreviation and ZIP code. If the normalized value is either an abbreviation for or very similar to the expected value, we can confidently swap those two.


```{r match_dist, collapse=TRUE}
norm_city <- norm_city %>% 
  rename(city_raw = city_sep) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_clean" = "state",
      "zip_clean" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )

ri <- left_join(
  x = ri,
  y = norm_city,
  by = c(
    "city_sep" = "city_raw",
    "state_clean", 
    "zip_clean"
  )
)
```

```{r count_swaps}
ri %>% 
  filter(city_swap != city_norm) %>% 
  count(state_clean, city_sep, city_norm, city_swap) %>% 
  arrange(desc(n))
```

#### Refine

```{r view_refine}

good_refine <- ri %>% 
  filter(state_clean == "RI") %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_clean" = "state",
      "zip_clean" = "zip"
    )
  )

ri <- ri %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap)) %>% 
  rename(city_clean =city_refine)
```

Each step of the cleaning process reduces the number of distinct city values.

#### Progress

Our goal for normalization was to increase the proportion of city molues known
to be valid and reduce the total distinct molues by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  toupper(ri$city_sep),
  ri$city_norm,
  ri$city_swap,
  ri$city_clean,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of molid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ri$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Rhode Island City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Rhode Island City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```


Before exporting, we can remove the intermediary normalization columns and rename all added variables with the _clean suffix.
```{r}
ri <- ri %>% 
  select(
    -city_sep,
    -city_norm,
    -city_swap,
  ) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```


## Conclude

1. There are `r nrow(ri)` records in the database
1. There are `r sum(ri$dupe_flag)` records with duplicate values (flagged with `dupe_flag`)
1. The ranges for dates and amounts are reasonable
1. Consistency has been improved with `stringr` package and custom `campfin::normal_*()` functions
1. The five-digit `zip_clean` variable has been created with `campfin::normal_zip()`
1. The `year` variable has been created with `lubridate::year()`
1. There are `r sum(ri$na_flag)` records with some missing key value, 
`r percent(sum(ri$na_flag)/nrow(ri))` of the total records.

## Export

```{r write_clean}
proc_dir <- here("state","ri", "expends", "data", "processed")
dir_create(proc_dir)

ri %>% 
  write_csv(
    na = "",
    path = glue("{proc_dir}/ri_expends_clean.csv")
  )
```

