---
title: "Iowa Stimulus"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ia", "stimulus", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tabulizer, # scrape pdf tables
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Stimulus data comes from the Iowa Economic Development Authority.

> Governor Reynolds launched the Iowa Small Business Relief Program to provide
financial assistance to small businesses economically impacted by the COVID-19
pandemic. The program offers eligible small businesses grants ranging from
$5,000-$25,000 in addition to offering Iowa businesses a deferral of sales and
use or withholding taxes due and waiver of penalty and interest. The application
window closed March 31. IEDA will not open a new round of applications.

> $24 million was appropriated and disbursed for the Small Business Relief
program April 7-10. On April 23, Governor Reynolds announced additional funding
through the federal CARES Act to assist more eligible businesses that applied
during the March application period.

## Download

We can scrape the IEDA website to obtain the links of PDF reports.

```{r raw_dir}
raw_dir <- dir_create(here("ia", "stimulus", "data", "raw"))
```

```{r raw_get}
home <- "https://www.iowaeconomicdevelopment.com/aspx/general/dynamicpage.aspx"
raw_get <- GET(home, query = list(id = 205))
```

```{r raw_links}
raw_urls <- content(raw_get) %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("sbrg-recipientslist_\\d+.pdf") %>% 
  str_c("https://www.iowaeconomicdevelopment.com", .)
```

Those `r length(raw_urls)` PDF files can be downloaded locally.

```{r raw_download}
raw_paths <- path(raw_dir, basename(raw_urls))
if (!all(file_exists(raw_paths))) {
  download.file(raw_urls, raw_paths)
}
```

```{r raw_size}
sum(file_size(raw_paths))
```

## Read

Using the `tabulizer::extract_tables()` function, we can scrape the PDF pages
for the tables containing our data.

```{r raw_scrape}
mid_area <- c(50, 0, 700, 600)
ias <- rep(list(NA), length(raw_paths))
for (i in seq_along(raw_paths)) {
  # extract date from file name
  file_date <- mdy(str_extract(raw_paths[i], "\\d+"))
  # create list of table areas
  n_pgs <- get_n_pages(raw_paths[i])
  pg_areas <- rep(list(mid_area), n_pgs)
  # first and last have smaller tables
  pg_areas[[1]][1] <- 100
  pg_areas[[n_pgs]][1] <- 100
  # scrape all tables
  ias[[i]] <- raw_paths[i] %>% 
    extract_tables(area = pg_areas) %>% 
    map(row_to_names, 1) %>% 
    map_df(as_tibble) %>%
    set_names(c("business", "county", "amount")) %>% 
    # remove total row from end
    slice(-nrow(.)) %>%
    # filter out double headers
    filter(str_detect(amount, "\\d")) %>% 
    # parse numeric column
    mutate(across(3, parse_number)) %>% 
    # add row and date from file
    mutate(round = i, date = file_date, .before = 1)
}
```

The tables from each file can be combined into a single data frame.

```{r raw_bind}
ias <- bind_rows(ias)
ias <- mutate(ias, across(where(is.character), str_to_upper))
```

## Explore

```{r glimpse}
head(ias)
tail(ias)
```

### Missing

There are no missing values.

```{r na_count}
col_stats(ias, count_na)
```

### Duplicates

There are no duplicate records.

```{r dupe_flag}
any(duplicated(ias))
```

### Categorical

```{r distinct_count}
col_stats(ias, n_distinct)
```

```{r distinct_plots}
explore_plot(ias, county)
```

### Amounts

We see the minimum and maximum are in line with the statutory values.

```{r ammount_summary}
noquote(map_chr(summary(ias$amount), dollar))
```

```{r hist_amount, echo=FALSE}
ias %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Iowa Stimulus Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
ias <- mutate(ias, year = year(date))
```

```{r date_range}
min(ias$date)
max(ias$date)
sum(ias$date > today())
```

## Wrangle

We need to manually add the state, agency, and funding program name.

```{r agency_add}
ias <- ias %>% 
  mutate(
    .data = ias,
    .before = business,
    state = "IA",
    agency = "IOWA ECONOMIC DEVELOPMENT AUTHORITY"
  )
```

## Conclude

```{r clean_tail}
head(ias)
```

1. There are `r comma(nrow(ias))` records in the database.
1. There are `r comma(sum(ias$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ias$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ia", "stimulus", "data", "clean"))
clean_path <- path(clean_dir, "ia_stimulus_clean.csv")
write_csv(ias, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ias)),
  type = md_code(map_chr(ias, typeof)),
  def = c(
    "Stimulus funding round",
    "Date reported with file",
    "Funding state abbreviation",
    "Funding agency name",
    "Recipient business name",
    "Recipient county name",
    "Loan amount ($5,000 to $25,000)",
    "Year loan given"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
