---
title: "New York Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ny", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The New York state campaign finance contributions data can be obtained from the
[New York State Board of Elections (NYSBOE)][boe].

[boe]: https://www.elections.ny.gov/INDEX.html

From the NYSBOE's [view reports page][vr] we can obtain the raw database in ASCII
files. The NYSBOE also provides disclaimers on the data availability and accuracy.

[vr]: https://www.elections.ny.gov/CFViewReports.html

> ##### Data Availability
> This database contains all financial disclosure reports filed with NYSBOE from
July of 1999 to the present. Financial disclosure reports filed prior to the
1999 July Periodic report are either on file with the New York State Archives or
in storage with the New York State Board of Elections. For further information
or to obtain copies of these archived or stored filings, please call
1-800-458-3453. Each page costs 25¢ plus postage and copy orders must be
prepaid.
>
> Electronically filed disclosure reports are generally available in the
database on the day they are received. A small number of candidates and
committees are either statutorily exempt or have applied for and obtained
exemptions from electronic filing. These filers will continue filing on paper
and their disclosure reports will become available as they are manually entered
into the database by NYSBOE staff.

> ##### Data Accuracy
> The majority of financial disclosure reports filed at NYSBOE are entered into
the database directly from e-mail, diskette, CD or DVD filings submitted by
committee treasurers or candidates. The information contained in paper filings
will be entered into the database exactly as it appears on the forms. Because
database searches retrieve information exactly the way it is reported and then
entered into the database, search results may be inaccurate and/or incomplete.
This will occur, for example, if filers do not adhere to the required format, do
not use the proper codes, misspell words or leave items blank. Although NYSBOE
carefully reviews disclosure reports and requires treasurers to submit amended
reports as needed, there will necessarily be delays before the review process is
completed and the information in the database is corrected.

## Import

To import the contributions data into R, we first have to download two files:
(1) the filer data and (2) the filings data. Together, these two files can
identify all parties to contributions.
 
> ##### Database Files in ASCII Delimited Format
> Updated data files are uploaded during active filing periods after 4:00 P.M.
daily until the filing is complete.
> 
> Note: To match the filing data files to Filer Names by filer ID you will need
to [Download the Filer data file][f]. Commcand.zip is a zipped file containing
the data file (commcand.asc) in ASCII delimited and two text files. (filerec.txt
contains the data file layout - codes.txt explains the codes used in the data
file).
>
> All downloadable files are zipped files containing a data file in ASCII
delimited format and two text files. (`efsrecb.txt` contains the data file
layout - `efssched.txt` explains the different schedules as they apply to the
database).
>
> [Download Data file containing ALL filings][a]. Note: This file is a large
file (238, 994 KB) that contains over 6 million records. Do not attempt to
download this file unless you have a database to download the file to.

[f]: https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/commcand.zip
[a]: https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/ALL_REPORTS.zip

### Download

```{r raw_size}
sboe <- "https://cfapp.elections.ny.gov/NYSBOE/download/ZipDataFiles/"
con_url <- str_c(sboe, "ALL_REPORTS.zip")
url_file_size(con_url)
fil_url <- str_c(sboe, "commcand.zip")
url_file_size(fil_url)
```

```{r raw_download}
raw_dir <- dir_create(here("ny", "contribs", "data", "raw"))
con_zip <- path(raw_dir, basename(con_url))
fil_zip <- path(raw_dir, basename(fil_url))
if (!this_file_new(con_zip)) {
  download.file(fil_url, fil_zip)
  download.file(con_url, con_zip)
}
```

```{r unzip_path}
if (length(dir_ls(raw_dir)) != 9) {
  unzip(con_zip, exdir = raw_dir)
  unzip(con_zip, exdir = raw_dir)
}
```

### Read

We will have to read the `EFSRECB.TXT` file to determine the column names and
formats for the `ALL_REPORTS.out` file, which has no header. The file also
describes the specifications needed to parse the file, which can be passed as
arguments to the `vroom::vroom()`.

```{r efsrecb, echo=FALSE}
efsrecb <- read_lines(path(raw_dir, "EFSRECB.TXT"))
cat(efsrecb[44:50], sep = "\n")
efsrecb <- efsrecb[10:42] %>% 
  read_table(na = "") %>% 
  clean_names("snake") %>% 
  mutate(
    field = field %>% 
      str_to_lower() %>% 
      str_remove("_\\d+$")
  )

print(efsrecb, n = Inf)
```

```{r raw_fix}
con_file <- path(raw_dir, "ALL_REPORTS.out")
fix_file <- path(raw_dir, "FIX_REPORTS.out")
if (!file_exists(fix_file)) {
  read_lines(raw_file) %>% 
  str_conv(encoding = "ASCII") %>% 
  str_replace_all("(?<!\r)\n", " ") %>%
  str_remove_all(",(?!\")") %>% 
  str_replace_all("(?<!^|,|\r\n)\"(?!,|\r\n|$)", "\'") %>% 
  write_lines(fix_file)
}
```

```{r raw_read}
nyc <- read_delim(
  file = fix_file,
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_names = efsrecb$field,
  col_types = cols(
    .default = col_character(),
    date1 = col_date_usa(),
    date2 = col_date_usa(),
    check_date = col_date_usa(),
    amount = col_double(),
    amount2 = col_double(),
    crerec_date = col_datetime("%m/%d/%Y %H:%M:%S")
  )
)
```

To ensure the file was properly read, we can count the distinct values of a
discrete variable like `contrib_type_code`.

```{r raw_check}
comma(nrow(nyc))
count(nyc, freport_id)
```

This file contains _all_ campaign finance transactions. If we consult
`EFSSHCED.TXT`, we determine which record types correspond to contributions.

```{r efssched, echo=FALSE}
efssched <- read_lines(path(raw_dir, "EFSSCHED.TXT"))
efssched <- efssched %>% 
  extract(60:78) %>% 
  str_trim() %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("chkbox", "report_type"),
    sep = "\\s-\\s{2}",
    extra = "merge"
  ) %>% 
  mutate(contrib = str_detect(report_type, "Contrib"))
print(efssched, n = Inf)
```

Then we can filter our data frame to include only those records using the 
`transaction_code` variable.

```{r raw_filter}
nyc <- nyc %>% 
  filter(transaction_code %in% efssched$chkbox[efssched$contrib]) %>% 
  remove_empty("cols")
```

### Join

We will use the files in `commcand.zip` to build a useful data frame of filer
information, which can then be joined to our contributions.

```{r filerec, echo=FALSE}
filerec <- read_table(read_lines(path(raw_dir, "FILEREC.TXT"))[8:22])
print(filerec)
```

```{r codes_parse, echo=FALSE}
codes_all <- str_trim(read_lines(file = path(raw_dir, "CODES.TXT")))
codes_all <- na.omit(na_if(codes_all, ""))
codes_office <- codes_all %>% 
  extract(5:79) %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("office", "office_name"),
    sep = "\\s",
    extra = "merge"
  ) %>% 
  mutate(office = as.integer(office))

codes_district <- codes_all %>% 
  extract(81:83) %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("district", "district_name"),
    sep = "\\s{2,}",
    extra = "merge",
    convert = TRUE
  )

codes_comm <- codes_all %>% 
  extract(85:105) %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("committee_type", "committee_name"),
    sep = "\\s{2,}",
    extra = "merge",
    convert = TRUE
  )

codes_report <- codes_all %>% 
  extract(107:118) %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("report", "report_name"),
    sep = "\\s{2,}",
    extra = "merge",
    convert = TRUE
  )
```

```{r zip_path}
fil_path <- path(raw_dir, "COMMCAND.txt")
commcand <- read_delim(
  file = fil_path,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = str_to_lower(filerec$FIELD),
  col_types = cols(
    .default = col_character(),
    office = col_integer(),
    district = col_integer()
  )
)
```

```{r join_codes, echo=FALSE}
commcand <- commcand %>%
  # join codes
  left_join(codes_office) %>% 
  # replace codes
  mutate(office = office_name) %>% 
  # remove joined names
  select(-office_name) %>% 
  # repeat for committee codesd
  left_join(codes_comm) %>% 
  mutate(committee_type = committee_name) %>% 
  select(-committee_name)
```

```{r join_contribs}
prop_in(nyc$filer_id, commcand$filer_id)
nyc <- left_join(
  x = nyc,
  y = commcand,
  by = "filer_id",
  suffix = c("_con", "_rec")
)

nyc <- rename_prefix(
  df = nyc,
  suffix = c("_con", "_rec"),
  punct = TRUE
)
```

```{r raw_rename, echo=FALSE}
# rename for clarify and brevity
nyc <- nyc %>% 
  rename(
    rec_id = filer_id,
    tran_code = transaction_code,
    date = date1,
    contrib_type = contrib_type_code,
    con_corp = corp,
    con_first = first_name,
    con_mid = mid_init,
    con_last = last_name,
    con_addr = addr_1,
    descr = description,
    recpt_code = other_recpt_code,
    purpose1 = purpose_code1,
    purpose2 = purpose_code2,
    explan = explanation,
    xfer = xfer_type,
    rec_name = filer_name,
    rec_type = filer_type,
    comm_type = committee_type,
    district = district,
    treas_first = treas_first_name,
    treas_last = treas_last_name,
    rec_addr = address
  )
```

```{r raw_select, echo=FALSE}
nyc <- nyc %>% 
  # see EFSSCHED.txt
  # remove non contrib cols
  select(
    -date2,
    -contrib_type,
    -check_date,
    -amount2,
    -descr,
    -recpt_code,
    -purpose1,
    -purpose2,
    -explan,
    -xfer,
    -chkbox
  )
```

## Explore

```{r glimpse}
head(nyc)
tail(nyc)
glimpse(sample_n(nyc, 20))
```

### Missing

```{r na_count}
col_stats(nyc, count_na)
```

_Very_ few records are missing the date, amount, or the name of the contributor
or recipient. We can flag the few that are with `campfin::flag_na()`.

```{r na_flag}
nyc <- nyc %>% 
  unite(
    con_first, con_mid, con_last, con_corp,
    col = con_name,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  flag_na(
    date, con_name, amount, rec_name
  )

percent(mean(nyc$na_flag), 0.01)
```

### Categorical

```{r n_distinct}
col_stats(nyc, n_distinct)
```

```{r explore_plot, echo=FALSE, fig.keep=FALSE}
explore_plot(nyc, freport_id)
explore_plot(nyc, tran_code)
explore_plot(nyc, rec_type)
explore_plot(nyc, status)
explore_plot(nyc, comm_type, nbar = 4)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(nyc$amount)
mean(nyc$amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
nyc %>%
  filter(amount > 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New York Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: NYSBOE",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
nyc <- mutate(nyc, year = year(date))
```

```{r date_fix}
nyc$year %>% 
  str_replace_all("^([0-9])$", "200\\1") %>% 
  str_replace_all("^(1[0-9])$", "20\\1") %>% 
  str_replace_all("^99$", "1999") %>% 
  str_replace_all("^999$", "1999") %>% 
  str_replace("^2(\\d{2})$", "20\\1") %>% 
  str_replace("^10(\\d{2})$", "20\\1") %>%
  str_replace("^190(\\d{1})$", "200\\1") %>%
  str_replace("^1899$", "1999") %>% 
  as.integer() -> nyc$year
```

```{r date_invalid}
nyc$year[!between(nyc$year, 1990, 2020)] <- NA
```

```{r bar_year, echo=FALSE}
nyc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1999, 2020, by = 2)) +
  coord_cartesian(xlim = c(1999, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Contributions by Year",
    caption = "Source: NYSBOE",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
nyc <- mutate_at(
  .tbl = nyc,
  .vars = vars(ends_with("addr")),
  .funs = list(norm = normal_address),
  abbs = usps_street,
  na_rep = TRUE
)
```

```{r address_view}
nyc %>% 
  select(contains("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nyc <- mutate_at(
  .tbl = nyc,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_progress}
progress_table(
  nyc$rec_zip,
  nyc$rec_zip_norm,  
  nyc$con_zip,
  nyc$con_zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nyc <- mutate_at(
  .tbl = nyc,
  .vars = vars(ends_with("state")),
  .funs = list(norm = normal_state),
  na_rep = TRUE,
  valid = NULL
)
```

```{r state_view}
nyc %>% 
  filter(con_state != con_state_norm) %>% 
  count(con_state, con_state_norm, sort = TRUE)
```

```{r state_fix}
one_let <- str_detect(nyc$con_state_norm, "^(N|Y)$")
zip_in <- nyc$con_zip_norm %in% zipcodes$zip[zipcodes$state == "NY"]
nyc$con_state_norm[which(one_let & zip_in)] <- "NY"
one_let <- str_detect(nyc$rec_state_norm, "^(N|Y)$")
zip_in <- nyc$rec_zip_norm %in% zipcodes$zip[zipcodes$state == "NY"]
nyc$rec_state_norm[which(one_let & zip_in)] <- "NY"
```

```{r state_progress}
progress_table(
  nyc$con_state,
  nyc$con_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
nyc <- mutate_at(
  .tbl = nyc,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  states = c("NY", "DC", "NEW YORK"),
  na = invalid_city,
  na_rep = TRUE
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r con_city_swap}
nyc <- nyc %>% 
  left_join(
    y = zipcodes,
    by = c(
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(con_city_norm, city_match),
    match_dist = str_dist(con_city_norm, city_match),
    con_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = con_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r rec_city_swap}
nyc <- nyc %>% 
  left_join(
    y = zipcodes,
    by = c(
      "rec_state_norm" = "state",
      "rec_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(rec_city_norm, city_match),
    match_dist = str_dist(rec_city_norm, city_match),
    rec_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = rec_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r con_city_refine}
con_good_refine <- nyc %>% 
  mutate(
    con_city_refine = con_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(con_city_refine != con_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "con_city_refine" = "city",
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  )
```

```{r con_city_count, echo=FALSE}
con_good_refine %>%
  count(
    con_state_norm, 
    con_zip_norm, 
    con_city_swap, 
    con_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r con_city_join}
nyc <- nyc %>% 
  left_join(con_good_refine) %>% 
  mutate(con_city_refine = coalesce(con_city_refine, con_city_swap))
```

Then repeat the process for recipient cities.

```{r rec_city_refine}
rec_good_refine <- nyc %>% 
  mutate(
    rec_city_refine = rec_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(rec_city_refine != rec_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "rec_city_refine" = "city",
      "rec_state_norm" = "state",
      "rec_zip_norm" = "zip"
    )
  )
nyc <- nyc %>% 
  left_join(rec_good_refine) %>% 
  mutate(rec_city_refine = coalesce(rec_city_refine, rec_city_swap))
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r check_filter}
nyc_out <- nyc %>% 
  filter(con_city_refine %out% c(valid_city, extra_city)) %>% 
  count(con_city_refine, con_state_norm, sort = TRUE) %>% 
  drop_na() %>% 
  head(1000)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("ny", "contribs", "data", "api_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file
  )
} else {
  check <- pmap_dfr(
    .l = list(
      nyc_out$con_city_refine, 
      nyc_out$con_state_norm
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    path = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  c(valid_locality)
```

#### Progress

```{r con_city_remain}
many_city <- c(valid_city, extra_city, valid_locality)
nyc %>% 
  filter(con_city_refine %out% many_city) %>% 
  count(con_city_refine, sort = TRUE)
```

```{r con_city_replace}
nyc <- nyc %>% 
  mutate(
    con_city_refine = con_city_refine %>% 
      str_replace("^NY$", "NEW YORK") %>% 
      str_replace("^NYC$", "New York")
  )
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(nyc$con_city),
  nyc$con_city_norm,
  nyc$con_city_swap,
  nyc$con_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nyc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

```{r rec_city_progress}
progress_table(
  str_to_upper(nyc$rec_city),
  nyc$rec_city_norm,
  nyc$rec_city_swap,
  nyc$rec_city_refine,
  compare = many_city
)
```

## Conclude

```{r clean_select}
nyc <- nyc %>% 
  select(
    -ends_with("_city_norm"),
    -ends_with("_city_swap")
  ) %>% 
  rename_all(str_replace, "_refine", "_clean") %>% 
  rename_all(str_replace, "_norm", "_clean")
```

```{r clean_glimpse}
glimpse(sample_n(nyc, 20))
```

1. There are `r comma(nrow(nyc))` records in the database.
1. There are `r comma(sum(nyc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nyc$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("ny", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ny_contribs_clean.csv")
write_csv(nyc, clean_path, na = "")
file_size(clean_path)
guess_encoding(clean_path)
```
