---
title: "New York Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ny", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("ny/contribs/docs/ny_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The New York state campaign finance contributions data can be obtained from the
[New York State Board of Elections (NYSBOE)][boe]. From the BOE's 
[Campaign Finance page][cf] we can navigate to the [bulk data][bulk] download
page (from the Campaign Disclosure drop-down menu).

> The New York State Board of Elections (NYSBOE) makes this reporting system
available as an UNOFFICIAL resource for information related to candidates and
their respective ballot access and campaign disclosure filings.
> 
> The NYSBOE Public Reporting system logs and tracks the disclosure reports of
candidate, ballot access documents including petitions, certificates of
nomination, related documents, and their status...
> 
> ##### Bulk Download
> Disclosure Reports filed with the New York State Board Of Elections in csv
format. Option to download All or specified Filing Year and Disclosure Period.

> ##### Data Availability
> This database contains all financial disclosure reports filed with NYSBOE from
July of 1999 to the present. Financial disclosure reports filed prior to the
1999 July Periodic report are either on file with the New York State Archives or
in storage with the New York State Board of Elections. For further information
or to obtain copies of these archived or stored disclosure reports, please call
1-800-458-3453.
> 
> Electronically filed disclosure reports are generally available in the
database on the day they are received.

> ##### Data Accuracy
> Financial disclosure reports filed at NYSBOE are added to the database
directly by submitting from the Electronic Filing Application. Because database
searched retrieve information exactly the way it is reported, search results may
be inaccurate and/or incomplete. This will occur, for example, if filers do not
adhere to the required format, do not use the proper codes, misspell words or
leave items blank. Although NYSBOE carefully reviews disclosure reports and
requires treasurers to submit amended reports as needed, there will necessarily
be delays before the review process is completed and the information in the
database is corrected.

> #### Database Files in Comma Seprated Value (CSV) Format
> To match the disclosure report data files to Filer Names by filer ID you will
need to Download the Filer data file (Commcand.zip) is a zipped file containing
the data file (commcand.csv) in CSV and a PDF file
(FileFormatReferenceFiler.pdf) contains the data file layout.

> All downloadable files are zipped files containing eight data files in CSV
format (`STATE_CANDIDATE_ACTIVE.csv`, `STATE_CANDIDATE_INACTIVE.csv`,
`COUNTY_CANDIDATE_ACTIVE.csv`, `COUNTY_CANDIDATE_INACTIVE.csv`,
`STATE_COMMITTEE_ACTIVE.csv`, `STATE_COMMITTEE_INACTIVE.csv`,
`COUNTY_COMMITTEE_ACTIVE.csv`, `COUNTY_COMMITTEE_INACTIVE.csv`) and a PDF file
(`FileFormatReference.pdf`) contains the datafile layout.

> **Note:**
> Selected Report Year "All" and Report Type "All" contains large files. Do not
attempt to download these files unless you have a database to download the files
to.

[boe]: https://www.elections.ny.gov/
[cf]: https://www.elections.ny.gov/CampaignFinance.html
[bulk]: https://publicreporting.elections.ny.gov/DownloadCampaignFinanceData/DownloadCampaignFinanceData

### Download

The BOE website has instructions on how to use the website to select what files
are to be downloaded. We can use these instructions to try and replicate the
form process with `GET()` and `POST()` request from the httr package.

> #### Bulk Download
> Search produces all financial disclosure reports filed with the New York State
Board of Elections (NYSBOE) from July of 1999 to present in csv format.
Financial disclosure reports filed prior to the 1999 July Periodic report are
either on file with the New York State Archives or in storage with the New York
State Board of Elections. Electronically filed disclosure reports are generally
available in the database on the day they are received.

> ##### Data Type
> Click on the Data Type drop-down list and select either **Filer Data** to
download the data of Filers registered with State Offices or select **Disclosure
Report** to download All or specified Filing Year and Disclosure Period. If
**Disclosure Report** is selected then system displays **Report Year** and
**Report Type** dropdown fields.

> ##### Report Year
> Dropdown is visible only on selecting **Disclosure Report** in the **Data
Type** dropdown. Click on the **Report Year** drop-down list and select a year
to download disclosure report for the selected year.

> ##### Report Type
> Dropdown is visible only on selecting **Disclosure Report** in the **Data
Type** dropdown. Click on the **Report Type** drop-down list and select an
report type to download disclosure report for the selected year and for the
selected type.

> ##### Download
> Click on the **Download** button to download the Filer Data or Disclosure
report based on values selected in the Filter fields. System will download the
Zip file with filer data or Disclosure Reports
### Download

```{r raw_dir}
raw_dir <- dir_create(here("ny", "contribs", "data", "raw"))
fil_zip <- file_temp(ext = "zip")
```

```{r}
dcf <- "https://publicreporting.elections.ny.gov/DownloadCampaignFinanceData/"
dcf_get <- GET(url = str_c(dcf, "DownloadCampaignFinanceData"))
dcf_cookie <- cookies(dcf_get)
dcf_asp <- dcf_cookie$value
names(dcf_asp) <- dcf_cookie$name
```

```{r fil_get}
fil_get <- GET(
  url = str_c(dcf, "DownloadZipFile"),
  user_agent("https://publicaccountability.org/"),
  write_disk(fil_zip, overwrite = TRUE),
  set_cookies(dcf_asp),
  progress("down"),
  query = list(
    lstDateType = "--lstDateType",
    lstUCYearDCF = "--lstUCYearDCF",
    lstFilingDesc = "--lstFilingDesc"
  )
)
```

### Extract

There are three files in each of the two ZIP archives.

```{r zip_view}
zip_ls <- list(
  "ALL_REPORTS.zip" = unzip(con_zip, list = TRUE),
  "commcand.zip"    = unzip(fil_zip, list = TRUE)
)
```

```{r zip_view2, echo=FALSE}
zip_ls <- zip_ls %>% 
  bind_rows(.id = "source_zip") %>% 
  as_tibble() %>% 
  transmute(
    source_zip = fs_path(source_zip),
    file_name = fs_path(Name),
    size = fs_bytes(Length),
    date = Date
  )
zip_ls
```

If these files don't exist in our raw data directory, they can be extracted.

```{r zip_txt}
raw_txt <- path(raw_dir, zip_ls$file_name)
raw_txt <- setNames(raw_txt, basename(raw_txt))
```

```{r unzip_path}
if (!all(file_exists(raw_txt))) {
  con_txt <- unzip(con_zip, exdir = raw_dir)
  fil_txt <- unzip(fil_zip, exdir = raw_dir)
}
```

### About

We will have to read the `EFSRECB.TXT` file to determine the column names and
formats for the `ALL_REPORTS.out` file, which has no header. The file also
describes the specifications needed to parse the file, which can be passed as
arguments to `readr::read_delim()`.

```{r efsrecb, echo=FALSE}
efsrecb <- read_lines(raw_txt["EFSRECB.TXT"])
```

```{r efsrecb_spec, echo=FALSE}
out_args <- str_split(efsrecb[44:50], ":\\s", simplify = TRUE)
out_args[, 1] <- str_remove_all(out_args[, 1], "\\W")
out_args[, 2] <- str_to_upper(out_args[, 2])
colnames(out_args) <- c("", "Value")
kable(out_args)
```

```{r efsrecb_cols, echo=FALSE}
out_cols <- efsrecb[10:42] %>% 
  read_table(na = "") %>% 
  clean_names("snake") %>% 
  mutate(
    location = parse_integer(location),
    type = case_when(
      type == "CHAR" ~ "`<chr>`",
      type == "DATE" ~ "`<date>`",
      type == "FLOAT" ~ "`<dbl>`",
      type == "INTEGER" ~ "`<int>`",
    ),
    format = str_remove_all(format, "'"),
    field = make_clean_names(str_remove(field, "_\\d+$")),
    efs_import = (efs_import == "REQUIRED")
  )

out_cols %>% 
  mutate(
    across(1, md_code),
    across(5, as.character),
    across(4:5, ~str_replace_na(., "")),
  ) %>% 
  kable()
```

### Fix

```{r raw_fix}
fix_file <- path(raw_dir, "FIX_REPORTS.csv")
if (!file_exists(fix_file)) {
  read_lines(raw_txt["ALL_REPORTS.out"]) %>% 
    iconv(to = "UTF-8", sub = "byte") %>% 
    str_replace_all("(?<!\r)\n", " ") %>%
    str_remove_all(",(?!\")") %>% 
    str_replace_all("(?<!^|,|\r\n)\"(?!,|\r\n|$)", "\'") %>% 
    write_lines(fix_file)
}
```

### Read

```{r raw_read}
nyc <- read_delim(
  file = fix_file,
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_names = out_cols$field,
  na = c("", "NA", "N/A"),
  col_types = cols(
    .default = col_character(),
    date1 = col_date("%m/%d/%Y"),
    date2 = col_date("%m/%d/%Y"),
    check_date = col_date("%m/%d/%Y"),
    amount = col_double(),
    amount2 = col_double(),
    crerec_date = col_datetime("%m/%d/%Y %H:%M:%S")
  )
)
```

To ensure the file was properly read, we can count the distinct values of a
discrete variable like `contrib_type_code`.

```{r raw_check}
comma(nrow(nyc))
all(unique(nyc$freport_id) %in% LETTERS[1:12])
count(nyc, freport_id)
```

### Filter

This file contains _all_ campaign finance transactions. If we consult
`EFSSHCED.TXT`, we determine which record types correspond to contributions.

```{r read_sched}
efssched <- read_lines(raw_txt["EFSSCHED.TXT"])
```

```{r type_code, echo=FALSE}
type_codes <- str_trim(efssched[60:78]) %>% 
  enframe(name = NULL) %>% 
  separate(
    col = value,
    into = c("type_code", "report_type"),
    sep = "\\s-\\s{2}",
    extra = "merge"
  ) %>% 
  mutate(contrib = str_detect(report_type, "Contrib|Receipt"))
print(type_codes, n = Inf)
```

Then we can filter our data frame to include only those records using the 
`transaction_code` variable.

```{r raw_filter}
nyc <- nyc %>% 
  filter(transaction_code %in% type_codes$type_code[type_codes$contrib]) %>% 
  remove_empty("cols")
```

We can consult the `EFSSCHED.txt` to see what columns apply to contribution
type records, removing any which shouldn't belong.

```{r raw_select, echo=FALSE}
nyc <- nyc %>% 
  # see EFSSCHED.txt
  select(
    -date2,
    -contrib_type_code,
    -check_date,
    -amount2,
    -description,
    -other_recpt_code,
    -purpose_code1,
    -purpose_code2,
    -explanation,
    -xfer_type,
    -chkbox
  )
```

```{r rows check}
# ensure proper rows kept
rows_post <- nrow(nyc)
comma(rows_post)
```

### Filers

We will use the files in `commcand.zip` to build a useful data frame of filer
information, which can then be joined to our contributions.

```{r filerec, echo=FALSE}
filerec <- read_table(read_lines(raw_txt["FILEREC.TXT"])[8:22])
filerec <- clean_names(filerec, case = "snake")
filerec <- mutate(filerec, across(1, make_clean_names))
filerec %>% 
  mutate(across(1, md_code)) %>% 
  mutate(
    type = case_when(
      type == "CHAR" ~ "`<chr>`",
      type == "DATE" ~ "`<date>`",
      type == "FLOAT" ~ "`<dbl>`",
      type == "INTEGER" ~ "`<int>`",
    )
  ) %>% 
  kable()
```

We will also read codes from the `CODES.TXT` file and join them to the data.

```{r codes_read}
codes_all <- str_trim(read_lines(file = raw_txt["CODES.TXT"]))
```

```{r codes_parse, echo=FALSE}
codes_all <- na.omit(na_if(codes_all, ""))
codes_office <- read_table(
  file = codes_all[5:79], 
  col_names = c("office", "office_name"),
  col_types = cols(
    office = col_integer(),
    office_name = col_character()
  )
)

codes_district <- read_table(
  file = codes_all[81:83], 
  col_names = c("district", "district_name")
)

codes_comm <- read_table(
  file = codes_all[85:105], 
  col_names = c("committee_type", "committee_name")
)

codes_report <- read_table(
  file = codes_all[107:118], 
  col_names = c("report", "report_name")
)
```

```{r codes_print, echo=FALSE}
kable(codes_office)
kable(codes_district)
kable(codes_comm)
kable(codes_report)
```

```{r zip_path}
commcand <- read_delim(
  file = raw_txt["COMMCAND.txt"],
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = filerec$field,
  col_types = cols(
    .default = col_character(),
    office = col_integer(),
    district = col_integer()
  )
)
```

```{r join_codes, echo=FALSE}
commcand <- commcand %>%
  # join codes
  left_join(codes_office, by = "office") %>% 
  left_join(codes_comm, by = "committee_type") %>% 
  # rename columns
  mutate(committee_type = committee_name, office = office_name) %>% 
  # remove codes
  select(-committee_name, -office_name)
```

#### Wrangle

Before joining the filer data to the contributions, we can clean up the
address data of each filer. See the section on wrangling before for a more
detailed explanation of wrangling steps made on contributor data.

```{r fil_addr}
fil_addr <- commcand %>% 
  select(fil_id = filer_id, addr = address, city, state, zip)
```

```{r fil_addr_addr}
fil_addr <- fil_addr %>% 
  mutate(
    addr_norm = normal_address(
      address = addr,
      abbs = usps_street,
      na = invalid_city
    )
  )
```

```{r fil_addr_zip}
fil_addr <- fil_addr %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r fil_addr_st}
fil_addr <- fil_addr %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r fil_addr_city}
fil_addr <- fil_addr %>% 
  mutate(
    city_norm = normal_city(
      city = city %>% 
        str_replace("STA$", "STATION") %>% 
        str_replace("STATI$", "STATION"), 
      abbs = usps_city,
      states = c("NY", "DC", "NEW YORK"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r fil_addr_check}
prop_in(fil_addr$zip_norm, valid_zip)
prop_in(fil_addr$state_norm, valid_state)
prop_in(fil_addr$city, valid_city)
```

```{r fil_addr_clean}
fil_addr <- fil_addr %>% 
  select(fil_id, ends_with("norm")) %>% 
  rename_with(~str_c("fil", ., sep = "_"), .cols = -1)
```

#### Join

We can now add the filer data to our contribution data.

```{r join_check_id}
# all filers in data
prop_in(nyc$filer_id, commcand$filer_id)
```

```{r join_contribs}
nyc <- left_join(
  x = nyc,
  y = commcand,
  by = "filer_id",
  suffix = c("_con", "_fil")
)

nyc <- rename_prefix(
  df = nyc,
  suffix = c("_con", "_fil"),
  punct = TRUE
)
```

```{r echo=FALSE}
nyc <- mutate(nyc, across(where(is.character), str_squish))
```

```{r raw_rename, echo=FALSE}
# rename for clarify and brevity
nyc <- nyc %>% 
  rename_with(~str_replace(., "filer_", "fil_")) %>% 
  rename(
    txn_code = transaction_code,
    date = date1,
    con_addr = addr_1,
    comm_type = committee_type,
    treas_first = treas_first_name,
    treas_last = treas_last_name,
    fil_addr = address,
    fil_status = status
  )
```

## Explore

There are `r comma(nrow(nyc))` rows of `r ncol(nyc)` columns. Each record
represents a single contribution made from an individual or entity to a 
political committee or candidate.

```{r glimpse}
glimpse(nyc)
tail(nyc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(nyc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

We must first create a single variable with the combined contributor names.

```{r con_name_unite}
nyc <- nyc %>% 
  unite(
    col = con_name,
    first_name, mid_init, last_name, corp,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  relocate(con_name, .after = last_col()) %>% 
  mutate(across(con_name, na_if, ""))
```

```{r na_flag}
key_vars <- c("date", "con_name", "amount", "fil_name")
nyc <- flag_na(nyc, all_of(key_vars))
mean(nyc$na_flag)
```

A little over `r percent(mean(nyc$na_flag))` of records are missing a key
variable, mostly a name identifying the contributor.

```{r na_view}
nyc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r distinct_count}
col_stats(nyc, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(nyc, freport_id)
explore_plot(nyc, txn_code)
explore_plot(nyc, contrib_code)
explore_plot(nyc, fil_type)
explore_plot(nyc, fil_status)
explore_plot(nyc, comm_type)
explore_plot(nyc, office)
```

### Amounts

```{r amount_summary}
summary(nyc$amount)
mean(nyc$amount <= 0, na.rm = TRUE)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(nyc[c(which.max(nyc$amount), which.min(nyc$amount)), ])
```

```{r hist_amount, echo=FALSE}
nyc %>%
  filter(!is.na(amount), amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New York Contributions Amount Distribution",
    caption = "Source: NYSBOE",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nyc <- mutate(nyc, year = year(date))
```

```{r date_fix}
nyc$year %>% 
  str_replace_all("^([0-9])$", "200\\1") %>% 
  str_replace_all("^(1[0-9])$", "20\\1") %>% 
  str_replace_all("^99$", "1999") %>% 
  str_replace_all("^999$", "1999") %>% 
  str_replace("^2(\\d{2})$", "20\\1") %>% 
  str_replace("^10(\\d{2})$", "20\\1") %>%
  str_replace("^190(\\d{1})$", "200\\1") %>%
  str_replace("^1899$", "1999") %>% 
  as.integer() -> nyc$year
```

```{r date_invalid}
count_na(nyc$date)
nyc$year[!between(nyc$year, 1990, 2020)] <- NA
```

```{r date_range}
count_na(nyc$date)
min(nyc$date, na.rm = TRUE)
sum(nyc$year < 1999, na.rm = TRUE)
max(nyc$date, na.rm = TRUE)
sum(nyc$date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
nyc %>% 
  count(year) %>% 
  filter(year >= 1999) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1999, 2021, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Contributions by Year",
    caption = "Source: NYSBOE",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Filers

We can now re-join the normalized address variables for filers.

```{r fil_addr_rejoin}
nyc <- left_join(nyc, fil_addr, by = "fil_id")
nrow(nyc) == rows_post
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r con_addr_norm}
con_addr_norm <- nyc %>% 
  distinct(con_addr) %>% 
  mutate(
    con_addr_norm = normal_address(
      address = con_addr,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r addr_rejoin}
nyc <- left_join(nyc, con_addr_norm, by = "con_addr")
nrow(nyc) == rows_post
rm(con_addr_norm); flush_memory()
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nyc <- nyc %>% 
  mutate(
    con_zip_norm = normal_zip(
      zip = con_zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nyc$con_zip,
  nyc$con_zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
st_norm <- nyc %>% 
  distinct(con_state, con_zip_norm) %>% 
  mutate(
    con_state_norm = normal_state(
      # should be "N/A" like with city, etc
      state = na_if(con_state, "N/"),
      abbreviate = TRUE,
      na_rep = TRUE
    )
  )
```

```{r state_bad}
bad_state <- st_norm$con_state_norm %out% valid_state
ny_zip <- st_norm$con_zip_norm %in% zipcodes$zip[zipcodes$state == "NY"]
ny_rx <- str_detect(st_norm$con_state_norm, "^N|Y$")
sort(table(st_norm$con_state_norm[bad_state & ny_zip & ny_rx]))
# replace if bad, ny zip, starts with N or ends Y
st_norm$con_state_norm[bad_state & ny_zip & ny_rx] <- "NY"
```

```{r state_join}
nyc <- left_join(nyc, st_norm, by = c("con_state", "con_zip_norm"))
```

```{r state_view}
nyc %>% 
  filter(con_state != con_state_norm) %>% 
  count(con_state, con_state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  nyc$con_state,
  nyc$con_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- nyc %>% 
  distinct(con_city, con_state_norm, con_zip_norm) %>% 
  mutate(
    con_city_norm = normal_city(
      city = con_city %>% 
        str_replace("^S\\.I\\.$", "STATEN ISLAND"), 
      abbs = usps_city,
      states = c("NY", "DC", "NEW YORK"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(con_city_norm, city_match),
    match_dist = str_dist(con_city_norm, city_match),
    con_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = con_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
nyc <- left_join(
  x = nyc,
  y = norm_city,
  by = c(
    "con_city", 
    "con_state_norm", 
    "con_zip_norm"
  )
)
nrow(nyc) == rows_post
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- nyc %>% 
  mutate(
    con_city_refine = con_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(con_city_refine != con_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "con_city_refine" = "city",
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    con_state_norm, 
    con_zip_norm, 
    con_city_swap, 
    con_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
nyc <- nyc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(con_city_refine = coalesce(con_city_refine, con_city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(nyc$con_city),
  nyc$con_city_norm,
  nyc$con_city_swap,
  nyc$con_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nyc$con_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
nyc <- nyc %>% 
  select(
    -con_city_norm,
    -con_city_swap,
    con_city_clean = con_city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(nyc, 50))
```

1. There are `r comma(nrow(nyc))` records in the database.
1. There are `r comma(sum(nyc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nyc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ny", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ny_contribs_clean.csv")
write_csv(nyc, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
