---
title: "Arizona Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("az", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  mdbr, # read mdb files
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("az/contribs/docs/az_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Arizona campaign finance database was obtained by the Investigative 
Reporting Workshop via a state records request pursuant to Ariz. Rev. Stat. Ann.
Secs. 39-121 to 39-122. The request was made via email to Renada Fisher,
Elections Customer Service Manager for the Arizona Secretary of State on January
13, 2021. She supplied a Microsoft Access database file via the secure file
sharing site Serv-U. That file was download and then uploaded to the IRW server,
where it can be downloaded.

```{r raw_dir}
raw_dir <- dir_create(here("az", "contribs", "data", "raw"))
raw_obj <- "20201230-CFS_Export.mdb"
raw_mdb <- path(raw_dir, raw_obj)
raw_aws <- path("csv", raw_obj)
```

```{r raw_download}
if (!file_exists(raw_mdb)) {
  save_object(
    object = raw_aws,
    bucket = "publicaccountability",
    file = raw_mdb,
    overwrite = FALSE,
    show_progress = TRUE
  )
}
```

## Read

The tables in this Microsoft Access database can be converted to flat, comma
delimited text files which can then be read as data frames and joined together
into a single, searchable file. The `mdbr` R package is needed for this 
conversion. That packages uses the [MDB Tools][mdbt] software, which needs to
be installed separately on your system. For example, Debian users can install
MDB Tools from the apt repository from the command line.

[mdbt]: https://github.com/mdbtools/mdbtools

``` bash
sudo apt install mdbtools
```

```{r raw_tables}
az_tables <- mdb_tables(raw_mdb)
md_bullet(md_code(az_tables))
```

The `Transactions` table contains the bulk data, every financial transaction
conducted by campaign committees in the state of Arizona. The other tables
contain the codes needed to identify the parties details of each transaction.

We can read all of these tables as data frames into a list.

```{r raw_read}
az <- map(
  .x = az_tables, 
  .f = read_mdb, 
  file = raw_mdb, 
  col_types = TRUE,
  na = c("", "NA", "N/A", "<missing value>")
)
```

```{r raw_name}
az <- map(az, clean_names, "snake")
names(az) <- make_clean_names(az_tables)
```

## Prepare

```{r fn_no_time, echo=FALSE}
no_time <- function(x) {
  x <- unique(x)
  if (inherits(x, "POSIXct")) {
    no_hour <- isTRUE(length(na.omit(unique(hour(x)))) == 1)
    no_min <- isTRUE(length(na.omit(unique(minute(x)))) == 1)
    all(no_min, no_hour)
  } else {
    FALSE
  }
}
```

```{r fn_fix_cons, echo=FALSE}
fix_cols <- function(.data) {
  .data %>% 
    remove_constant() %>% 
    remove_empty("cols") %>% 
    rename_with(~str_replace(., "transaction", "txn")) %>% 
    rename_with(~str_replace(., "committee", "cmte")) %>% 
    rename_with(~str_replace(., "candidate", "cand")) %>% 
    rename_with(~str_replace(., "category", "cat")) %>% 
    rename_with(~str_replace(., "report", "rpt")) %>% 
    rename_with(~str_replace(., "name", "nm")) %>% 
    mutate(across(where(no_time), as_date))
}
```

First we will simplify each data frame by removing empty and constant columns
and abbreviating column names for ease of use.

```{r fix_map}
az <- map(az, fix_cols)
```

Now that each data frame is read and cleaned, we can prepare them for a join.

### Transactions

The most important table is `Transactions`. This table only identifies the
reporting committee and transaction partner using their unique ID number. This
table includes not just contribution _to_ a committee but also the expenses 
_from_ that committee, etc. Each transaction has it's own unique ID.

```{r echo=FALSE}
az$transactions <- select(az$transactions, c(1, 3:7))
az$transactions <- mutate(az$transactions, across(txn_date, as_date))
```

```{r txn_look}
az$transactions
```

Contributions from individuals are the most common transaction type, but there
are other types that are different kinds of income.

```{r txn_types}
az$transactions %>% 
  count(txn_type_id, sort = TRUE) %>% 
  left_join(az$transaction_types, by = "txn_type_id") %>% 
  add_prop(sum = FALSE)
```

Here, we are interested in _all_ income types, so we will use the 
`income_expense_neutral_id` variable associated with each `txn_type_*` to keep
only "Income" transactions.

```{r txn_ien}
az$income_expense_neutral
```

```{r txn_txn_filter}
az$transactions <- az$transactions %>% 
  # identify transaction type
  left_join(az$transaction_types, by = "txn_type_id") %>% 
  # identify if type is income
  left_join(az$income_expense_neutral, by = "income_expense_neutral_id") %>% 
  # keep only income types
  filter(income_expense_neutral == "Income") %>% 
  # remove ids and move cols
  select(-txn_type_id, -income_expense_neutral_id, -income_expense_neutral) %>% 
  relocate(txn_type_nm = txn_type_nm, .before = cmte_id)
```

```{r echo=FALSE}
az$transactions
```

Now we will have to use the `Names` and `Committees` tables to identify the
entities making the contributions and the committees receiving them.

### Committees

The `Committees` table contains all the information on political committees,
including their financial institutions, chairpersons, etc. We are only
interested in adding names and addresses to the transaction data so it can be
searched.

```{r cmte_view_pre}
az$committees
```

```{r cmte_addr_keep}
az$committees <- az$committees %>% 
  # keep address and candidate info
  rename_with(~str_replace(., "physical", "cmte")) %>% 
  select(
    cmte_id, nm_id, starts_with("cmte"), 
    cand_party_id, cand_office_id,
    cand_is_incumbent, cand_cycle_id,
    cand_other_party_nm
  ) %>% 
  rename(cmte_zip = cmte_zip_code)
```

We can then use the `Offices`, `Parties`, and `Cycles` tables to identify
candidate information by code. We will keep the names and remove the codes.

```{r cmte_code_join}
az$committees <- az$committees %>% 
  # add office name from id
  left_join(az$offices, by = c("cand_office_id" = "office_id")) %>% 
  # add party name from id
  left_join(az$parties, by = c("cand_party_id" = "party_id")) %>% 
  # add cycle year from id
  left_join(az$cycles[, 1:2], by = c("cand_cycle_id" = "cycle_id")) %>% 
  # remove codes
  select(
    -cand_office_id, cand_office = office_nm,
    -cand_party_id, cand_party = party_nm,
    -cand_cycle_id, cand_cycle = cycle_nm
  ) %>% 
  # combine "other" party and other col
  mutate(
    cand_party = cand_party %>% 
      na_if("Other") %>% 
      coalesce(cand_other_party_nm)
  ) %>% 
  select(-cand_other_party_nm)
```

This data still only identifies the committees by their unique `cmte_id`. We 
can use the `nm_id` variable to add the names for those committees.

```{r cmte_nm_select}
cmte_nms <- select(az$names, nm_id, cmte_nm = last_nm)
cmte_nms
```

```{r cmte_nm_join}
az$committees <- az$committees %>% 
  # add the committee name
  inner_join(cmte_nms, by = "nm_id") %>% 
  relocate(cmte_nm, .after = cmte_id) %>% 
  select(-nm_id)
```

```{r echo=FALSE}
az$committees
```

### Contributors

We can now identify the contributors by their name and address. Again, we will
join together other tables which identify the _type_ of contributor and the
county name if they're from Arizona.

```{r ctbr_select}
az$names <- az$names %>% 
  rename(zip = zip_code) %>% 
  # add entity type and county name
  left_join(az$entity_types, by = "entity_type_id") %>% 
  left_join(az$counties, by = "county_id") %>% 
  select(-nm_group_id, -entity_type_id, -county_id) %>% 
  rename(entity_type = entity_type_nm, county = county_nm) %>% 
  # keep names of contributors only
  filter(nm_id %in% az$transactions$nm_id)
```

```{r echo=FALSE}
az$names %>% 
  slice(1:10) %>% 
  remove_empty("cols")
```

## Wrangle

Before identifying the parties to each transaction, we can use their individual
tables to normalize the geographic variables search on the Accountability site.

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Fix

There are `r comma(sum(str_detect(az$names$zip, "\\$"), na.rm = TRUE))` records
with some kind of problem shifting the geographic variables out of place. For
these records, the `city` is in thr `address1` variable, the `state` is in the
`city` variable, and it appears as though the `zip` variable has been split in
half with the first two digits (85 is a comma Arizona ZIP code prefix) in the
`state` column and the last three converted somehow to a dollar value?

This problem has been confirmed to exist in the raw Microsoft Access database
and not be an error stemming from our conversion to a flat text file.

```{r bad_zip_dollar}
az$names %>% 
  filter(str_detect(zip, "\\$")) %>% 
  select(6:10)
```

We can attempt to fix these values to the best of our ability.

First, we will find the index of any record with a `zip` value containing a
dollar sight at the start.

```{r bad_dollar_find}
dollar_zip <- str_which(az$names$zip, "^\\$")
length(dollar_zip)
```

We can then create vectors of the city names and state abbreviations from the
wrong column.

```{r bad_city_state}
bad_city <- az$names$address1[dollar_zip]
bad_state <- az$names$city[dollar_zip]
```

Then, we will combine the two numbers from the `state` column of these bad
records the dollar value, after converting those dollar values to three digit
numbers.

```{r bad_zip_merge}
bad_zip <- str_c(
  az$names$state[dollar_zip],
  str_sub(
    string = str_pad(
      string = parse_number(x = az$names$zip[dollar_zip]) * 100, 
      width = 3, 
      pad = "0"
    ), 
    end = 3
  )
)
```

Only `r percent(prop_in(bad_zip, valid_zip), 0.1)` of these combined ZIP code
values are actually valid, but that is an improvement over the bad data. It
seems as though some accuracy was lost whenever these ZIP codes were 
accidentally converted to dollar values.

```{r bad_zip_check}
prop_in(az$names$zip[dollar_zip], valid_zip)
prop_in(bad_zip, valid_zip)
```

```{r bad_zip_fix_view}
mutate(count_vec(bad_zip), valid = value %in% valid_zip)
```

We can now take these three vectors and place them back into the proper columns.

```{r bad_cols_fix}
az$names$city[dollar_zip] <- bad_city
az$names$state[dollar_zip] <- bad_state
az$names$zip[dollar_zip] <- bad_zip
```

For these bad records, there is no street address value. The city values were
placed there instead and the addresses were not misplaced (as was the case with
the state and ZIP code values).

```{r bad_addr_remove}
az$names$address1[dollar_zip] <- NA
```

### Address

For the street `addresss*` variables, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

We will repeat the process once for committee addresses then for contributors.

```{r cmte_address_norm}
az$committees <- az$committees %>% 
  unite(
    col = cmte_address_full,
    matches("address\\d$"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    cmte_address_norm = normal_address(
      address = cmte_address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-cmte_address_full)
```

```{r nm_address_norm}
nm_norm_addr <- az$names %>% 
  distinct(address1, address2) %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r nm_address_view}
nm_norm_addr
```

```{r nm_address_join}
az$names <- left_join(az$names, nm_norm_addr)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r cmte_zip_good}
prop_in(az$committees$cmte_zip, valid_zip)
```

```{r zip_norm}
az$names <- az$names %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  az$names$zip,
  az$names$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r cmte_state_norm}
az$committees <- az$committees %>% 
  mutate(
    cmte_state_norm = normal_state(
      state = cmte_state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r nm_state_norm}
az$names <- az$names %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r nm_state_view}
az$names %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r nm_state_progress}
progress_table(
  az$names$state,
  az$names$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r nm_city_norm}
nm_norm_city <- az$names %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("AZ", "DC", "ARIZONA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r cmte_city_norm}
cmte_norm_city <- az$committees %>% 
  distinct(cmte_city, cmte_state_norm, cmte_zip) %>% 
  mutate(
    cmte_city_norm = normal_city(
      city = cmte_city, 
      abbs = usps_city,
      states = c("AZ", "DC", "ARIZONA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r nm_city_swap}
nm_norm_city <- nm_norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r nm_city_rejoin}
az$names <- left_join(
  x = az$names,
  y = nm_norm_city,
  by = c(
    "city" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

```{r cmte_city_swap}
cmte_norm_city <- cmte_norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "cmte_state_norm" = "state",
      "cmte_zip" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(cmte_city_norm, city_match),
    match_dist = str_dist(cmte_city_norm, city_match),
    cmte_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = cmte_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r cmte_city_rejoin}
az$committees <- left_join(
  x = az$committees,
  y = cmte_norm_city,
  by = c(
    "cmte_city", 
    "cmte_state_norm", 
    "cmte_zip"
  )
)
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r nm_city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(az$names$city),
  az$names$city_norm,
  az$names$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

```{r cmte_city_progress, echo=FALSE}
cmte_progress <- progress_table(
  str_to_upper(az$committees$cmte_city),
  az$committees$cmte_city_norm,
  az$committees$cmte_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(cmte_progress, digits = 3)
```

These intermediary columns can then be removed.

```{r nm_geo_clean}
az$names <- select(az$names, -city_norm, city_norm = city_swap)
```

```{r cmte_geo_clean}
az$committees <- select(
  .data = az$committees, 
  -cmte_city_norm, cmte_city_norm = cmte_city_swap
)
```

```{r echo=FALSE}
rm(nm_norm_city, cmte_norm_city)
flush_memory()
```

## Join

These tables, with their normalized geographic variables, can now be joined.

```{r comm_join}
az$transactions <- az$transactions %>% 
  left_join(az$committees, by = "cmte_id") %>% 
  relocate(cmte_id, .before = cmte_nm)
```

```{r ctbr_join}
az$transactions <- az$transactions %>% 
  left_join(az$names, by = "nm_id") %>% 
  relocate(nm_id, .before = last_nm)
```

```{r az_rename}
azc <- az$transactions
rm(az); flush_memory()
```

## Explore

There are `r comma(nrow(azc))` rows of `r ncol(azc)` columns. Each record
represents a single contribution made from an entity to a political committee.

```{r glimpse}
glimpse(azc)
tail(azc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(azc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("txn_id", "txn_date", "last_nm", "amount", "cmte_nm")
azc <- flag_na(azc, all_of(key_vars))
sum(azc$na_flag)
```

```{r na_view}
azc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("az", "contribs", "dupes.tsv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  azs <- azc %>% 
    select(-txn_id) %>% 
    group_split(cmte_zip)
  split_id <- split(azc$txn_id, azc$cmte_zip)
  pb <- txtProgressBar(max = length(azs), style = 3)
  for (i in seq_along(azs)) {
    d1 <- duplicated(azs[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(azs[[i]], fromLast = TRUE)
      dupes <- tibble(txn_id = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(azs)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("txn_id", "dupe_flag"),
  col_types = cols(
    txn_id = col_integer(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(azc)
azc <- left_join(azc, dupes, by = "txn_id")
azc <- mutate(azc, dupe_flag = !is.na(dupe_flag))
mean(azc$dupe_flag)
```

We can see that, despite unique `txn_id` values, there are duplicate records.

```{r dupe_view}
azc %>% 
  filter(dupe_flag) %>% 
  select(txn_id, all_of(key_vars)) %>% 
  arrange(last_nm)
```

### Categorical

```{r distinct_count}
col_stats(azc, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(azc, txn_type_nm) + scale_x_wrap()
explore_plot(azc, cand_office) + scale_x_wrap()
explore_plot(azc, cand_party) + scale_x_wrap()
explore_plot(azc, entity_type) + scale_x_wrap(15)
azc %>% 
  count(cand_cycle) %>% 
  ggplot(aes(cand_cycle, n)) +
  geom_col(aes(fill = n)) +
  scale_fill_viridis_c(option = "B", end = 0.75, guide = FALSE) +
  scale_x_continuous(breaks = 1994:2022) +
  scale_y_continuous(labels = comma) +
  labs(y = "Count")
```

### Amounts

```{r amount_summary}
summary(azc$amount)
mean(azc$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(azc[c(which.max(azc$amount), which.min(azc$amount)), ])
```

```{r hist_amount, echo=FALSE}
azc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Arizona Contributions Amount Distribution",
    caption = "Source: AZ SOS",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `txn_date` with `lubridate::year()`

```{r date_year}
azc <- mutate(azc, txn_year = year(txn_date))
```

```{r date_range}
min(azc$txn_date)
mean(azc$txn_year < 2000)
max(azc$txn_date)
sum(azc$txn_date > today())
```

```{r bar_year, echo=FALSE}
azc %>% 
  count(txn_year) %>% 
  mutate(even = is_even(txn_year)) %>% 
  ggplot(aes(x = txn_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1994, 2020, by = 2)) +
  coord_cartesian(xlim = c(1996, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona Contributions by Year",
    caption = "Source: AZ SOS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
azc <- rename_all(azc, ~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(azc, 50))
```

1. There are `r comma(nrow(azc))` records in the database.
1. There are `r comma(sum(azc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(azc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("az", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "az_contribs_clean.csv")
write_csv(azc, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
