---
title: "Virgina Expenditures Data Diary"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Prerequisites

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, warning=FALSE, error=FALSE}
pacman::p_load_gh("VerbalExpressions/RVerbalExpressions")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # text mining tools
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & databse
  batman, # parse logicals
  refinr, # cluster & merge
  scales, #format strings
  rvest, # scrape website
  skimr, # summary stats
  vroom, # quickly read
  glue, # combine strings
  gluedown, #markdown
  here, # locate storage
  fs # search storage 
)
```

```{r custom, echo=FALSE}
# fix conflict
here <- here::here
# # custom utility functions
# "%out%" <- Negate("%in%")
# print_all <- function(df) df %>% print(n = nrow(.)) 
# # source functions
# source(here("R", "code", "normalize_geo.R"))
# source(here("R", "code", "all_files_new.R"))
# # load data
# data("zipcode")
# zipcode <-
#   as_tibble(zipcode) %>% 
#   select(city, state, zip) %>% 
#   mutate(city = normalize_city(city))
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is provided by the Virginia Department of Elections (ELECT). From the campaign finance
reporting home page, we can navigate to "Download Campaign Finance Data" which takes us to the 
[`SBE_CSV/CF/`][03] subdirectory of the ELECT website.

On this page, there are additional subdirectories for each year from 1999 until 2012. There are
additional subdirectories for each month from January 2012 to June 2019.

The new update includes datasets from July 2019 to Nov. 2022.

Inside each subdirectory of `/SBE_CSV/CF/`, there are separate CSV files for each form submitted
by committees. Expenditure data is reported by candidates and committees using "Schedule D" forms,
as described by the [ELECT website][04]:

> Schedule D – Itemization of Expenditures Use to report all expenditures incurred by a committee. 

From the [instructions on how to fill our Schedule D forms][05], we know the following data is
included:

1. The full name of person or company paid and the complete mailing address of payee
2. The description of the item or service purchased
3. The name of the individual who authorized the expenditure
4. The date the expense was incurred
5. The amount of the expenditure

[03]: https://apps.elections.virginia.gov/SBE_CSV/CF/ "source"
[04]: https://www.elections.virginia.gov/candidatepac-info/campaign-finance-disclosure-forms/index.html
[05]: https://www.elections.virginia.gov/media/formswarehouse/Campaign-Finance/2018/CampaignFinance/Schedules/2014Schedule-D-Instructions.pdf "instructions"

## Import

Schedule D bulk downloads are saved as CSV files, which can easily be imported once downloaded.

### Download

In the yearly subdirectories (e.g., `/SBE_CSV/CF/2010/`), schedule D data is separated into two
files:

1. `SBE_CSV/CF/2010/ScheduleD.csv` (5MB)
2. `SBE_CSV/CF/2010/ScheduleD_PAC.csv` (115KB)

For years after 2011, the files are organized by month and are not separated:

1. `SBE_CSV/CF/2013_02/ScheduleD.csv`

We will start by downloading all the files separated by month from 2012 to January 2023. Next update should start February 2023. 

First we need to create the URLs for each year/month combination.

```{r expend_urls}
#sub_dirs <- unlist(map(2012:2019, str_c, str_pad(1:12, 2, side = "left", pad = "0"), sep = "_"))
sub_dirs <- unlist(map(2012:2023, str_c, str_pad(1:12, 2, side = "left", pad = "0"), sep = "_"))
exp_urls <- sort(glue("https://apps.elections.virginia.gov/SBE_CSV/CF/{sub_dirs}/ScheduleD.csv"))
rep_urls <- sort(glue("https://apps.elections.virginia.gov/SBE_CSV/CF/{sub_dirs}/Report.csv"))
exp_urls <- exp_urls[3:(length(sub_dirs)-11)]
rep_urls <- rep_urls[3:(length(sub_dirs)-11)]
head(exp_urls)

exp_urls %>% 
  tail() %>% 
  md_code() %>% 
  md_bullet()
```

Then we can download these files to our `/data/raw/single/` directory.

```{r raw_paths_build}
raw_dir <- here("state","va", "expends", "data", "raw", "single")
dir_create(raw_dir)

raw_names <- basename(str_replace(exp_urls, "/(?=[^/]*$)", "_"))
rep_names <- basename(str_replace(rep_urls, "/(?=[^/]*$)", "_"))
raw_paths <- path(raw_dir, raw_names)
rep_paths <- path(dir_create(here("state","va", "expends", "data", "raw", "meta")), rep_names)
```


```{r download, eval=FALSE}
library(httr)    

for (i in seq_along(raw_paths)) {
  wait = 1
  if (file_exists(raw_paths[i])) {
    next("file already downloaded")
  } else {
 ua <- "Mozilla/5.0 (Linux; Android 6.0; Nexus 5 Build/MRA58N) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/110.0.0.0 Mobile Safari/537.36"
    httr::GET(exp_urls[i],add_headers(`Connection` = "keep-alive", `User-Agent` = ua), write_disk(raw_paths[i]))
    x <- read_lines(raw_paths[i])
    if (str_starts(x[1], "#")) {
      next("file already fixed")
    } else {
      x <- str_replace_all(x, "(?<!^|,|\r\n)\"(?!,|\r\n|$)", "'")
      x <- c("### file fixed", x)
      write_lines(x, raw_paths[i])
      Sys.sleep(time = wait)
    }
  }
}

for (i in seq_along(rep_paths)) {
  wait = 0.5
  if (file_exists(rep_paths[i])) {
    next("file already downloaded")
  } else {
 ua <- "Mozilla/5.0 (Linux; Android 6.0; Nexus 5 Build/MRA58N) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/110.0.0.0 Mobile Safari/537.36"
    httr::GET(rep_urls[i],add_headers(`Connection` = "keep-alive", `User-Agent` = ua), write_disk(rep_paths[i]))
      Sys.sleep(time = wait)
    }
  }
}
```

## Fix
To properly read the file into R, we first have to do some simple string processing to the text file.

```{r fix_dir}
raw_info <- dir_info(raw_dir)
fix_dir <- dir_create(path(dirname(raw_dir), "fix"))
fix_eval <- length(dir_ls(fix_dir)) != nrow(raw_info)
```

```{r fix_loop_old, eval=FALSE}
# for old format files
for (f in raw_info$path) {
  n <- path(fix_dir, str_c("FIX", basename(f), sep = "_"))
  x <- read_lines(f, skip = 1)
  for (i in rev(seq_along(x))) {
    y <- i - 1
    if (y == 0) {
      next() # skip first
    } else if (str_starts(x[i], "\"\\d+\",") | str_ends(x[y], "\"(Y|N)\"")) {
      next() # skip if good
    } else { # merge if bad
      x[y] <- str_c(x[y], x[i])
      x <- x[-i] # remove bad
    }
  }
  x <- str_remove(x, '(?<=")"(?!,)')
  write_lines(x, n)
  message(basename(n))
}
```

```{r fix_loop_new, eval=FALSE}
# new format files
for (f in raw_info$path[1:length(raw_info$path)]) {
  n <- path(fix_dir, str_c("FIX", basename(f), sep = "_"))
  x <- read_lines(f, skip = 1)
  for (i in rev(seq_along(x))) {
    if (str_starts(x[i], "\\d+,\\d+,")) {
      next() # skip if good
    } else { # merge if bad
      x[i - 1] <- str_c(x[i - 1], x[i])
      x <- x[-i] # remove bad
    }
  }
  write_lines(x, n)
  message(basename(n))
}
```

```{r fix_info}
fix_info <- as_tibble(dir_info(fix_dir))
sum(fix_info$size)
fix_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```


### Read

Since all recent files are located in the same directory with the same structure, we can read them
all at once by using `purrr::map()` to apply `readr::read_csv()` to each file in the directory,
then binding each file into a single data frame using `dplyr::bind_rows()`.

We also need to read the related data on reports submitted by each committee, which contain names and addresses of each committee. The field we can join the report data by is `report_id`, which is unique. Each committee could have multiple report ids, but not vice versa.

```{r read fix}
va <- 
  dir_ls(fix_dir, glob = "*.csv") %>% 
  map_dfr(
    read_delim,
    delim = ",",
    na = c("NA", "N/A", ""),
    escape_double = FALSE,
    col_types = cols(
      .default = col_character(),
      IsIndividual = col_logical(),
      TransactionDate = col_date("%m/%d/%Y"),
      Amount = col_double()
    )
  ) %>% 
    clean_names()

va_meta <- rep_paths %>% map_dfr(read_delim, delim = ",",
                                   na = c("NA", "N/A", ""),
                                 escape_double = TRUE, col_types = cols(
                                   .default = col_character(), ReportYear=col_integer(),
                                 )) %>% clean_names()

va_prev <- dir_ls(here("state", "va","expends", "data", "previous")) %>% 
                    read_delim(
    delim = ",",
    na = c("NA", "N/A", ""),
    escape_double = FALSE,
    col_types = cols(.default = col_character(),
                     IsIndividual = col_logical(),
                      TransactionDate = col_date("%m/%d/%Y"),
                      Amount = col_double())) %>% clean_names()
```

The older files, separated by payee type, have a different structure and will have to be imported,
explored, and cleaned separated from the recent files.

## Explore

There are `r nrow(va)` records of `r length(va)` variables in the full database.

```{r glimpse}
glimpse(sample_frac(va))
glimpse(sample_frac(va_meta))
```

```{r meta streamline}
va_meta <- va_meta %>% select(report_id,candidate_name, committee_name,office_sought, city, address_line1, address_line2, address_line3, state_code, zip_code)

va_meta <- va_meta %>% rename_at(vars(5:10), ~paste0("com_",.))
```

```{r report}
va <- va %>% left_join(va_meta, by = "report_id")

va <- va %>% filter(schedule_d_id %out% va_prev$schedule_d_id)

va %>% filter(committee_name %>% is.na()) %>% nrow()
```

```{r}
va_prev <- va_prev %>% rename_at(vars(starts_with("committee")), ~sub("committee","com", .x)) %>% glimpse()

va_new <- va_prev %>% 
  rename(committee_contact_id = com_contact_id,
         committee_name = com_name) %>% 
  bind_rows(va)
```


### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
va %>% 
  map(n_distinct) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_distinct") %>% 
  mutate(prop_distinct = round(n_distinct / nrow(va), 4))
```

We can explore the distribution of the least distinct values with `ggplot2::geom_bar()`.

```{r type_bar, echo=FALSE}
va %>%
  filter(!is.na(is_individual)) %>%
  count(is_individual) %>% 
  mutate(percent = n/nrow(va)) %>% 
  ggplot() + 
  scale_y_continuous(labels = scales::percent) +
  geom_col(aes(is_individual, percent))
```

Or, filter the data and explore the most frequent discrete data.

```{r state_bar, echo=FALSE, fig.height=10}
va %>% 
  count(state_code, sort = TRUE) %>% 
  head(20) %>% 
  ggplot() + 
  geom_col(aes(reorder(state_code, n), n)) +
  coord_flip() +
  labs(x = "State", y = "count")
```

The `item_or_service` variable is an open-ended text field, so we can only analyze it by frequency
or word tokens.

```{r words_bar}
va %>% 
  unnest_tokens(word, item_or_service) %>% 
  mutate(word = str_to_lower(word)) %>%
  count(word, sort = TRUE) %>% 
  anti_join(stop_words) %>% 
  head(20) %>% 
  ggplot() + 
  geom_col(aes(reorder(word, n), n)) +
  coord_flip() +
  labs(x = "Word", y = "count")
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
va %>% 
  map(function(var) sum(is.na(var))) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_na") %>% 
  mutate(prop_na = n_na / nrow(va))
```

### Duplicates

We can use `janitor::get_dupes()` to create a table only including records that are completely
duplicated across every row, aside from the uniqe `schedule_d_id` variable.

```{r get_dupes}
va_dupes <- va %>% 
  select(-schedule_d_id) %>% 
  get_dupes() %>% 
  distinct() %>% 
  mutate(dupe_flag = TRUE)
```

There are `r nrow(va_dupes)` distinct duplicated records in this database, covering 
`r sum(va_dupes$dupe_count)` total records. It's entirely possible that two expenditures can be
made by the same committee, to the same payeee, of the same amount, on the same day, for the same
purpose. However, we will flag these records with `dupe_flag` nonetheless.

```{r view_dupes}
va_dupes %>% 
  tabyl(item_or_service) %>%
  as_tibble() %>% 
  arrange(desc(n)) %>% 
  mutate(cum_percent = cumsum(percent))
```

```{r join_dupes}
va <- va %>%
  left_join(va_dupes) %>% 
  mutate(dupe_flag = !is.na(dupe_flag))
```

### Ranges

It's important to ensure the ranges for continuous variables makes sense; that there aren't any
old or future dates or trillion dollar expenditures. We can explore these variables with
`ggplot2::ggplot()` functions and `base::summary()`.

#### Amounts

The expenditure `amount` variable contains the USD value of the expenditure and can reasonably
reach millions of dollars. This dataset contains `r sum(va$amount < 0, na.rm = T)` records with
an `amount` value less than zero, which sometimes indicate expenditure correction filings.
There are however `r sum(va$amount == 0, na.rm = T)` records with an `amount` value _of_ zero.

```{r}
summary(va$amount)
va %>% filter(amount == max(amount, na.rm = TRUE)) %>% glimpse()
```

We can visually explore the distribution and range of `amount` values using 
`ggplot2::geom_histogram()` and `ggplot2::geom_boxplot()`. 

Expenditures have a Log-normal distribution, with the vast majority of expenditures around the 
mean of \$1,385.

```{r}
va %>% 
  select(amount) %>% 
  mutate(oom = 10^(ceiling(log10(amount)))) %>% 
  count(oom) %>% 
  arrange(oom) %>%
  mutate(oom = as.ordered(oom)) %>% 
  ggplot() +
  geom_col(aes(oom, n)) +
  labs(
    title = "Distribution of VA Expenditures by Amount Order of Magnitude",
    x = "Order of Magnitude",
    y = "Count"
  )
```

```{r amount_hist_nonlog, echo=FALSE}
va %>% 
  ggplot() +
  geom_histogram(aes(amount)) +
  scale_x_continuous(labels = scales::dollar) +
  labs(
    title = "Distribution of VA Expenditure Amount",
    x = "Expenditure Amount (Log)",
    caption = "SOURCE: www.virginia.gov"
  )
```

We will have to transform the x-axis logarithmically to find patterns in the distribution.

```{r amount_hist, echo=FALSE}
va %>% 
  ggplot() +
  geom_histogram(aes(amount)) +
  scale_x_continuous(trans = "log10", labels = scales::dollar) +
  labs(
    title = "Distribution of VA Expenditure Amount",
    x = "Expenditure Amount (Log)",
    caption = "SOURCE: www.virginia.gov"
  )
```

We can use `ggplot2::facet_wrap()` to explore that distribution for both individual (candidate)
committees and more general issue committees.

```{r amount_hist_ind, echo=FALSE}
va %>% 
  filter(!is.na(is_individual)) %>% 
  ggplot() +
  geom_histogram(aes(amount)) +
  scale_x_continuous(trans = "log10", labels = scales::dollar) +
  facet_wrap(~is_individual, ncol = 1, labeller = ) +
  labs(
    title = "Distribution of VA Expenditure Amount",
    subtitle = "by Committee Type",
    x = "Expenditure Amount (Log)",
    caption = "SOURCE: www.virginia.gov"
  )
```

```{r amount_box_ind, echo=FALSE}
va %>% 
  filter(!is.na(is_individual)) %>% 
  ggplot(aes(x = is_individual, y = amount)) +
  geom_boxplot(varwidth = TRUE, outlier.alpha = 0.10) +
  scale_y_continuous(trans = "log10", labels = scales::dollar) +
  labs(
    title = "Range of VA Expenditure Amount",
    subtitle = "by Committee Type",
    x = "Individual Committee",
    y = "Expenditure Amount (Log)",
    caption = "SOURCE: www.virginia.gov"
  )
```

```{r mean_month_line, echo=FALSE}
va %>%
  group_by(
    election_year = year(transaction_date) %% 2 != 0,
    month = month(transaction_date)
  ) %>% 
  summarise(mean_amount = mean(amount)) %>% 
  ungroup() %>% 
  ggplot(
    mapping = aes(
      x = month, 
      y = mean_amount
    )
  ) +
  geom_line(mapping = aes(color = election_year), size = 2) +
  scale_x_continuous() +
  labs(
    title = "Mean Cost of VA Expenditures Over Year",
    x = "Month of Year",
    y = "Mean Expenditure Amount",
    caption = "SOURCE: www.virginia.gov"
  )
```

### Dates

The quasi-continuous variable `transaction_date` should also be explored for a reasonable
range. There are no expenditures made before `r min(va$transaction_date)` and
`r sum(va$transaction_date > today(), na.rm = TRUE)` expenditures reported as being made in the
future.

```{r summary_date, collapse=TRUE}
summary(va$transaction_date)
```

```{r n_year_bar, echo=FALSE}
va %>%
  group_by(year = year(transaction_date)) %>% 
  count() %>% 
  
  ggplot() +
  geom_col(aes(year, n)) +
  scale_x_continuous(breaks = 2012:2023) +
  labs(
    title = "Number of VA Expenditures Over Time",
    x = "Year",
    y = "Count",
    caption = "SOURCE: www.virginia.gov"
  )
```

```{r n_month_line, echo=FALSE}
va %>%
  group_by(
    election_year = year(transaction_date) %% 2 != 0,
    month = month(transaction_date)
  ) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(x = month, y = n, color = election_year), size = 2) +
  labs(
    title = "Number of VA Expenditures Over Year",
    x = "Month of Year",
    y = "Number of Expenditures",
    caption = "SOURCE: www.virginia.gov"

  )
```

## Wrangle

To improve the searchability of the data on the TAP website, we will endeavor to normalize
character strings, correct misspellings, flag undisambiguatable values. Original variables will
remain immutable, all records will be preserved, and manipulated versions of each variable take
the form `*_clean`.

### Year

Create a `transaction_year` variable from `transaction_date` using `lubridate::year()` after 
parsing the character string earlier with `readr::col_date()`.

```{r add_year}
va <- va %>% mutate(transaction_year = year(transaction_date))
```

### Address

The `address` variable should be minimally cleaned by removing punctuation and fixing white-space.

```{r str_normalize}
va <- va %>% 
 unite(
    col = address_full,
    starts_with("address_"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
  mutate(
    address_clean = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r com address norm}
va <- va %>% 
 unite(
    col = address_full,
    starts_with("com_address_"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  # normalize combined addr
  mutate(
    com_address_clean = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.


```{r clean_zipcodes}
va <- va %>% 
  mutate_at(
    .vars = vars(ends_with("zip_code")),
    .funs = list(norm = normal_zip),
    na_rep = TRUE
  )
```

### State

Using comprehensive list of state abbreviations in the Zipcodes database, we can isolate invalid
`state` values and manually correct them.

```{r valid_state, collapse=TRUE}
valid_state <- c(unique(campfin::zipcodes$state), "AB", "BC", "MB", "NB", "NS", "ON", "PE", "QC", "SK")
length(valid_state)
setdiff(valid_state, state.abb)
```

```{r view_states, collapse=TRUE}
setdiff(va$state_code, valid_state)
va <- va %>%
    mutate_at(
    .vars = vars(ends_with("state_code")),
    .funs = list(norm = normal_state),
    abbreviate = TRUE,
    na_rep = TRUE,
    valid = NULL
  )
```

### City

```{r city_start, collapse=TRUE}
valid_city <- unique(campfin::zipcodes$city)
n_distinct(va$city)
mean(va$city %in% campfin::zipcodes$city)
```

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Normalize

We will use the `prep_city()` function in the `/R` directory to normalize the strings, remove some
common `NA` values, and lop abbreviations off the end of the string.

```{r prep_city, collapse=TRUE}
va <- va %>% 
    mutate_at(
    .vars = vars(ends_with("city")),
    .funs = list(norm = normal_city),
    abbs = usps_city,
    states = c("VA", "VIRGINIA", "MA", "DC", "TX"),
    na = invalid_city,
    na_rep = TRUE
  )

n_distinct(va$city_norm)
```

#### Match

To disambiguate the city values, we will look at the _expected_ city name for a given ZIP code. We
can calculate the edit distance between each original value and the expected value.

```{r match_city}
va <- va %>%
  left_join(
    zipcodes,
    by = c(
      "state_code_norm" = "state",
      "zip_code_norm" = "zip"
    )
  ) %>%
  rename(city = city.x, city_match = city.y) %>%
  mutate(match_dist = stringdist(city_norm, city_match))
```


```{r view_dist}
summary(va$match_dist)
va %>% 
  count(perf_match = match_dist == 0) %>% 
  mutate(p = n/sum(n))
```

```{r view_match}
va %>% 
    filter(match_dist == 1) %>% 
    count(city_norm, city_match, sort = TRUE)
```

#### Swap

If the edit distance is less than or equal to two, we can fairly safely use the actual, expected
value instead of the original value.

```{r swap_city, collapse=TRUE}
va <- va %>% 
  mutate(
    city_swap = if_else(
      condition = match_dist <= 2, 
      true = city_match, 
      false = city_norm
    )
  )

n_distinct(va$city_swap)

va <- va %>% 
rename(com_city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "com_state_code_norm" = "state",
      "com_zip_code_norm" = "zip"
    )
  ) %>% 
  rename(com_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(com_city_norm,com_city_match),
    match_dist = str_dist(com_city_norm,com_city_match),
    com_city_swap = if_else(
      condition = !is.na(com_city_match) & (match_abb | match_dist == 1),
      true = com_city_match, 
      false = com_city_norm
    )
  ) %>% 
  select(
    -com_city_match,
    -match_abb,
    -match_dist
  ) %>% glimpse()
```

There are still `r sum(va$city_swap %out% valid_city, na.rm = T)` records with a `city_swap` value
not in our list of valid city names. Many, if not most, of these values are actually acceptable
city names that are simply not in our list.

```{r view_swap}
va %>% 
  filter(city_swap %out% valid_city) %>%
  count(city_swap, sort = TRUE)
```

#### Refine
The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- va %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_code_norm" = "state",
      "zip_code_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_code_norm, 
    zip_code_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
va <- va %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```


#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  toupper(va$city),
  va$city_norm,
  va$city_swap,
  va$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(va$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Virgnia City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Virginia City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```


## Export
1. There are `r comma(nrow(va))` records in the database.
1. There are `r comma(sum(va$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(va$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.
```{r}
proc_dir <- here("state","va", "expends", "data", "processed")
dir_create(proc_dir)

va <- va %>% 
  rename(city_clean = city_refine,
         state_clean = state_code_norm,
  com_state_clean = com_state_code_norm,
         zip_clean = zip_code_norm,
         com_zip_clean = com_zip_code_norm,
         com_city_clean = com_city_swap) %>% 
  select(
    -city_norm,
    -city_swap,
    -city_match,
    -com_city_norm
  )

va %>% 
  write_csv(
    na = "",
    path = str_c(proc_dir, "va_expends.csv", sep = "/")
  )
```
