---
title: "Baltimore Employee Earnings"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("md", "baltimore", "salary", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("md/baltimore/docs/md_baltimore_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

> This dataset represents gross salary of Baltimore City employees from fiscal
year 2011 through last fiscal year and includes employees who were employed on
June 30 of last fiscal year

## Download

```{r raw_dir}
raw_dir <- dir_create(here("md", "baltimore", "salary", "data", "raw"))
raw_csv <- path(raw_dir, "Baltimore_Employee_Salaries.csv")
```

```{r raw_download}
if (!file_exists(raw_csv)) {
  GET(
    write_disk(raw_csv),
    progress("down"),
    url = "https://opendata.arcgis.com/api/v3/datasets/afdaf8cca48a4bcea9282a781e9190a6_0/downloads/data",
    query = list(
      format = "csv",
      spatialRefId = "4326"
    )
  )
}
```

## Read

```{r raw_read}
bal <- read_delim(
  file = raw_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  trim_ws = TRUE,
  col_types = cols(
    .default = col_character(),
    ID = col_integer(),
    HireDate = col_date("%Y/%m/%d %H:%M:%S+%z"),
    AnnualSalary = col_double(),
    FiscalYear = col_number(),
    ObjectId = col_integer()
  )
)
```

```{r}
bal <- clean_names(bal, case = "snake")
```

```{r}
prop_distinct(bal$id)
unique(bal$fiscal_year)
```

## Explore

There are `r comma(nrow(bal))` rows of `r ncol(bal)` columns. Each record
represents a single employee of Baltimore earning a salary.

```{r glimpse}
glimpse(bal)
tail(bal)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(bal, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("first_name", "last_name", "annual_salary", "hire_date")
bal <- flag_na(bal, all_of(key_vars))
sum(bal$na_flag)
```

Every record missing a value has the first name of "BPD" and is missing a
`last_name` value.

```{r na_view}
bal %>% 
  filter(na_flag) %>% 
  count(first_name, agency_name, sort = TRUE)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
bal <- flag_dupes(bal, -id, -object_id)
sum(bal$dupe_flag)
```

There are no duplicate records, even ignore the unique `id` variables.

### Categorical

```{r distinct_count}
col_stats(bal, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(bal, job_title)
explore_plot(bal, agency_name)
```

### Amounts

```{r amount_summary}
summary(bal$annual_salary)
sum(bal$annual_salary <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(bal[c(which.max(bal$annual_salary), which.min(bal$annual_salary)), ])
```

```{r hist_amount, echo=FALSE}
bal %>%
  ggplot(aes(annual_salary)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Baltimore Salary Distribution",
    caption = "Source: Baltimore Open Data",
    x = "Amount",
    y = "Count"
  )
```

### Dates

```{r date_range}
min(bal$hire_date)
sum(bal$hire_date == min(bal$hire_date))
sum(bal$fiscal_year < 2000)
max(bal$hire_date)
sum(bal$hire_date > today())
```

```{r bar_year, echo=FALSE}
bal %>% 
  filter(hire_date > "1950-01-01") %>% 
  ggplot(aes(x = hire_date)) + 
  geom_histogram(bins = 50) +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y") +
  labs(
    title = "Baltimore Employees by Year Hired",
    caption = "Source: Baltimore Open Data",
    x = "Year Hired",
    y = "Count"
  )
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(bal, 50))
```

1. There are `r comma(nrow(bal))` records in the database.
1. There are `r comma(sum(bal$dupe_flag))` duplicate records in the database.
1. The range and distribution of `annual_salary` and `hire_date` are mostly
reasonable aside from a couple of outliers.
1. There are `r comma(sum(bal$na_flag))` records missing key variables.
1. No geographic variables are in the data.
1. The fiscal year variable can be used for the year.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("md", "baltimore", "salary", "data", "clean"))
clean_path <- path(clean_dir, "baltimore_salary_2011-2021.csv")
write_csv(bal, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
