---
title: "Florida Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("fl", "lobbying", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load_gh("kiernann/gluedown")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # read web pages
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained as tab-delinated files from the [Florida Lobbying Registration Office][lro] (LRO).

[lro]: https://floridalobbyist.gov/

### About

>  Delimited data files are made available below for compensation reports submitted online,
beginning in 2007. Data files for the last eight quarters will be retained for each branch. The
tab-delimited files below are in the (.TXT) format and can be imported into any word processor,
spreadsheet, or database program.

### Variables

The LRO provides a variable key with definitions for each column in the data sets.

[key]: https://floridalobbyist.gov/CompensationReportSearch/DownloadDataChart

```{r var_defs, echo=FALSE}
var_url <- "https://floridalobbyist.gov/CompensationReportSearch/DownloadDataChart"
read_html(var_url) %>% 
  html_node("table") %>% 
  html_table() %>% 
  mutate(
    `Data Element` = glue("`{make_clean_names(`Data Element`)}`"),
    `Definition` = str_trunc(Definition, width = 80)
  ) %>% 
  kable()
```

## Import

To create a single clean data file of lobbyist activity, we will first download each file locally
and read as a single data frame.

### Download

The data is separated into quarterly files by year. The URL for each file takes a consistent
format. With the `tidyr::expand_grid()` and `glue::glue()` functions, we can create a URL for all
bombinations of year, quarter, and branch.

```{r list_year_urls}
urls <- 
  expand_grid(
    year = 2008:2019,
    quarter = 1:4,
    branch = c("Executive", "Legislative")
  ) %>% 
  mutate(
    url = glue("https://floridalobbyist.gov/reports/{year}_Quarter{quarter}_{branch}.txt")
  )
```

```{r urls_view, echo=FALSE}
print(urls)
```

```{r}
urls <- pull(urls)
```

This creates `r n_distinct(urls)` distinct URLs, each corresponding to a separate file.

```{r print_urls, results='asis'}
md_bullet(head(urls), cat = TRUE)
```

We can download each TXT file to the `/fl/data/raw` directory.

```{r create_raw_dir}
raw_dir <- here("fl", "lobbying", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw, eval=FALSE}
if (!all_files_new(raw_dir, glob = "*.txt$")) {
  for (url in urls) {
    download.file(url, destfile = str_c(raw_dir, basename(url), sep = "/"))
  }
}
```

```{r list_files, echo=FALSE}
dir_info(raw_dir) %>% 
  mutate(path = str_remove(path, here())) %>% 
  select(
    path, 
    type,
    size,
    birth_time
  )
```

### Read

```{r read_raw}
fll <- dir_ls(raw_dir) %>% 
  vroom(
    delim = "\t",
    .name_repair = make_clean_names,
    escape_backslash = FALSE,
    escape_double = TRUE,
    id = "source_file",
    col_types = cols(
      .default = col_character(),
      REPORT_YEAR = col_double(),
      TOTAL_COMPENSATION_RANGE = col_factor(
        levels = c(
          "$0.00", 
          "$1.00-$49,999.00",
          "$50,000.00-$99,999.00", 
          "$100,000.00-$249,999.00", 
          "$250,000.00-$499,999.00", 
          "$500,000.00-$999,999.00",
          "$1,000,000.00"
        )
      )
    )
  )
```

```{r glimpse_raw}
head(fll)
tail(fll)
glimpse(sample_frac(fll))
```

## Wrangle

To improve the searchability of the database, we can normalize much of the data using the `campfin`
package. Much of the data is repeated for both the principal lobbyist and then their firm.

### Firm

#### Address

```{r address_norm}
packageVersion("tidyr")
fll <- fll %>%
  unite(
    starts_with("address_line"),
    col = "address_full",
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      add_abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view, echo=FALSE}
fll %>% 
  select(
    starts_with("address_line"),
    address_norm
  ) %>% 
  distinct() %>% 
  sample_frac()
```

#### Postal

```{r zip_norm}
fll <- fll %>% 
  mutate(
    zip_norm = normal_zip(
      zip = postal_code,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  fll$postal_code,
  fll$zip_norm,
  compare = valid_zip
)
```

#### State

```{r state_norm}
fll <- fll %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE
    )
  )
```

```{r state_progress}
progress_table(
  fll$state,
  fll$state_norm,
  compare = valid_state
)
```

#### City

```{r city_norm_swap}
fll <- fll %>% 
  rename(city_raw = city) %>% 
  mutate(
    city_norm = normal_city(
      city = city_raw,
      geo_abbs = usps_city,
      na_rep = TRUE
    )
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "zip_norm" = "zip",
      "state_norm" = "state"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_dist = str_dist(city_norm, city_match),
    match_abb = is_abbrev(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  )
```

```{r city_refine}
good_refine <- fll %>% 
  filter(state_norm == "FL") %>% 
  mutate(
    city_refine = city_swap %>% 
      refinr::key_collision_merge() %>% 
      refinr::n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_norm != city_refine) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "zip_norm" = "zip",
      "state_norm" = "state"
    )
  )
```

```{r view_refine, echo=FALSE}
good_refine %>% 
  count(state_norm, city_swap, city_refine, sort = TRUE)
```

```{r join_refine}
fll <- fll %>% 
  left_join(good_refine, by = names(fll)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

```{r city_progress}
progress_table(
  fll$city_raw,
  fll$city_norm,
  fll$city_swap,
  fll$city_refine,
  compare = valid_city
)
```

#### Phone

```{r phone_norm, echo=FALSE}
fll <- fll %>% 
  mutate(
    phone_norm = normal_phone(
      number = phone_number,
      format = "(%a) %e-%l",
      na_bad = FALSE,
      rm_ext = TRUE
    )
  )
```

```{r phone_view}
fll %>%
  filter(phone_number != phone_norm) %>% 
  select(
    phone_number,
    phone_norm
  ) %>% 
  drop_na() %>% 
  distinct() %>% 
  sample_frac()
```

### Principal

#### Address

```{r principal_address_norm}
fll <- fll %>%
  unite(
    starts_with("principal_address_line"),
    col = "principal_address_full",
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    principal_address_norm = normal_address(
      address = principal_address_full,
      add_abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-principal_address_full)
```

```{r principal_address_view, echo=FALSE}
fll %>% 
  select(
    starts_with("principal_address_line"),
    principal_address_norm
  ) %>% 
  distinct() %>% 
  sample_frac()
```

#### Postal

```{r principal_zip_norm}
fll <- fll %>% 
  mutate(
    principal_zip_norm = normal_zip(
      zip = principal_postal_code,
      na_rep = TRUE
    )
  )
```

```{r principal_zip_progress}
progress_table(
  fll$principal_postal_code,
  fll$principal_zip_norm,
  compare = valid_zip
)
```

#### State

```{r principal_state_norm}
fll <- fll %>% 
  mutate(
    principal_state_norm = normal_state(
      state = principal_state_name,
      abbreviate = TRUE,
      na_rep = TRUE
    )
  )
```

```{r principal_state_progress}
progress_table(
  fll$principal_state_name,
  fll$principal_state_norm,
  compare = valid_state
)
```

#### City

```{r principal_city_norm_swap}
fll <- fll %>% 
  mutate(
    principal_city_norm = normal_city(
      city = principal_city_name,
      geo_abbs = usps_city,
      na_rep = TRUE
    )
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "principal_zip_norm" = "zip",
      "principal_state_norm" = "state"
    )
  ) %>% 
  rename(principal_city_match = city) %>% 
  mutate(
    principal_match_dist = str_dist(principal_city_norm, principal_city_match),
    principal_match_abb = is_abbrev(principal_city_norm, principal_city_match),
    principal_city_swap = if_else(
      condition = principal_match_abb | principal_match_dist == 1,
      true = principal_city_match,
      false = principal_city_norm
    )
  )
```

```{r principal_city_progress}
progress_table(
  fll$principal_city_name,
  fll$principal_city_norm,
  fll$principal_city_swap,
  compare = valid_city
)
```

#### Phone

```{r principal_phone_norm}
fll <- fll %>% 
  mutate(
    principal_phone_norm = normal_phone(
      number = principal_phone_number,
      format = "(%a) %e-%l",
      na_bad = FALSE,
      rm_ext = TRUE
    )
  )
```

```{r principal_phone_view, echo=FALSE}
fll %>%
  filter(principal_phone_number != principal_phone_norm) %>% 
  select(
    principal_phone_number,
    principal_phone_norm
  ) %>% 
  drop_na() %>% 
  distinct() %>% 
  sample_frac()
```

### Firm

#### Address

```{r prime_firm_address_norm}
fll <- fll %>%
  unite(
    starts_with("prime_firm_address_line"),
    col = "prime_firm_address_full",
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    prime_firm_address_norm = normal_address(
      address = prime_firm_address_full,
      add_abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-prime_firm_address_full)
```

```{r prime_firm_address_view, echo=FALSE}
fll %>% 
  select(
    starts_with("prime_firm_address_line"),
    prime_firm_address_norm
  ) %>% 
  distinct() %>% 
  sample_frac()
```

#### Postal

```{r prime_firm_zip_norm}
fll <- fll %>% 
  mutate(
    prime_firm_zip_norm = normal_zip(
      zip = prime_firm_postal_code,
      na_rep = TRUE
    )
  )
```

```{r prime_firm_zip_progress}
progress_table(
  fll$prime_firm_postal_code,
  fll$prime_firm_zip_norm,
  compare = valid_zip
)
```

#### State

```{r prime_firm_state_norm}
fll <- fll %>% 
  mutate(
    prime_firm_state_norm = normal_state(
      state = prime_firm_state_name,
      abbreviate = TRUE,
      na_rep = TRUE
    )
  )
```

```{r prime_firm_state_progress}
progress_table(
  fll$prime_firm_state_name,
  fll$prime_firm_state_norm,
  compare = valid_state
)
```

#### City

```{r prime_firm_city_norm_swap}
fll <- fll %>% 
  mutate(
    prime_firm_city_norm = normal_city(
      city = prime_firm_city_name,
      geo_abbs = usps_city,
      na_rep = TRUE
    )
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "prime_firm_zip_norm" = "zip",
      "prime_firm_state_norm" = "state"
    )
  ) %>% 
  rename(prime_firm_city_match = city) %>% 
  mutate(
    prime_firm_match_dist = str_dist(prime_firm_city_norm, prime_firm_city_match),
    prime_firm_match_abb = is_abbrev(prime_firm_city_norm, prime_firm_city_match),
    prime_firm_city_swap = if_else(
      condition = prime_firm_match_abb | prime_firm_match_dist == 1,
      true = prime_firm_city_match,
      false = prime_firm_city_norm
    )
  )
```

```{r prime_firm_city_progress}
progress_table(
  fll$prime_firm_city_name,
  fll$prime_firm_city_norm,
  fll$prime_firm_city_swap,
  compare = valid_city
)
```

#### Phone

```{r prime_firm_phone_norm}
fll <- fll %>% 
  mutate(
    prime_firm_phone_norm = normal_phone(
      number = prime_firm_phone_number,
      format = "(%a) %e-%l",
      na_bad = FALSE,
      rm_ext = TRUE
    )
  )
```

```{r prime_firm_phone_view, echo=FALSE}
fll %>%
  filter(prime_firm_phone_number != prime_firm_phone_norm) %>% 
  select(
    prime_firm_phone_number,
    prime_firm_phone_norm
  ) %>% 
  drop_na() %>% 
  distinct() %>% 
  sample_frac()
```
