---
title: "New Hampshire Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nh", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contribution data can be obtained from the New Hampshire Secreyary of State's
[Campaign Finance System][cfs]. Each record represents a single monetary 
contribution or correction reported by the recieving committee or candidate.

Importantly, it seems as though not all contributions can be obtained from this
online portal: 

> NOTE: Use of the CFS is not mandatory; therefor, reports filed by candidates
and committees who have not registered and filed on the CFS are not searchable.
You may; however, search the "reports" and view them in pdf format.

[cfs]: https://cfs.sos.nh.gov/

## Download

To download the file, we will have to run an empty search on the CFS search
portal for all contributions from 2000 to `r today()`. This can be done manually
or with the an `httr::POST()` (provided you have the right cookies).

> Search for receipts reported on all campaign finance reports filed with the
Secretary of State, who have registered and filed their reports using the
Campaign Finance System (CFS).

> Search results are returned in a grid view and can be sorted within each
column by using the arrow at the top of the column or exported to your local
computer in Excel or CSV format, by clicking on the appropriate button in the
lower right hand corner of the grid.

```{r raw_dir}
raw_dir <- dir_create(here("nh", "contribs", "data", "raw"))
raw_path <- path(raw_dir, "ViewContributionsList.csv")
raw_gone <- !file_exists(raw_path)
```

```{r raw_download, eval=FALSE}
POST(
  url = "https://cfs.sos.nh.gov/Public/ViewReceipts",
  write_disk(raw_path, overwrite = FALSE),
  query = list(
    ContributorType = "",
    ContributionType = "",
    ContributorName = "",
    FirstName = "",
    Employer = "",
    FilingPeriodName = "",
    OccupationTitle = "",
    GabId = "",
    OfficeTypeCode = "",
    DistrictID = "",
    OfficeID = "",
    StartDate = "1/1/2000",
    EndDate = format(Sys.Date(), "%m/%d/%Y"),
    AmountRangeFrom = "",
    AmountRangeTo = "",
    OutOfState = "on",
    MemberId = "",
    PageIndex = "1",
    Election = "",
    AggAmountRangeFrom = "",
    CFID = "",
    AggAmountRangeTo = ""
  )
```

## Read

There are a handful of issues with the raw text file which can be fixed using
regular expressions.

```{r raw_fix}
# read lines as vector
nhl <- read_lines(raw_path)
# remove trailing comma from every line
nhl <- str_remove(nhl, ",$")
# add empty field for some ID at start
no_id <- str_starts(nhl, "\\d{1,2}/\\d{1,2}/\\d{4}")
nhl[no_id] <- str_replace(nhl[no_id], "^(.*)?$", ",\\1")
# collapse split line
nhl[53100] <- str_c(nhl[53100], nhl[53101])
nhl[53101] <- NA
# save to temp file
fix_path <- file_temp(ext = "csv")
write_lines(na.omit(nhl), fix_path)
```

Then this fixed temp file can be read and parsed into R.

```{r raw_read}
# 73,717
nhc <- read_delim(
  file = fix_path,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Transaction Date` = col_date("%m/%d/%Y %H:%M:%S %p"),
    `Amount` = col_double(),
    `Reporting Period` = col_date("%m/%d/%Y %H:%M:%S %p"),
    `Total Contribution Amount` = col_double()
  )
)
```

For convenience, the column names will be shorted and converted to snake case.

```{r raw_rename}
raw_names <- names(nhc)
nhc <- nhc %>% 
  rename(
    date = `Transaction Date`,
    con_type = `Contributor Type`,
    contributor = `Contributor Name`,
    geo_full = `Contributor Address`,
    recipient = `Receiving Registrant`,
    rec_type = `Receiving Registrant Type`,
    cycle = `Election Cycle`,
    period = `Reporting Period`,
    method = `Contribution Type`,
    total = `Total Contribution Amount`,
    in_kind = `In-kind Sub Category`,
    city_old = `Town/City`,
    state_old = `Town State`,
    occupation = `Occupation`,
    employer = `Employer Name`
  ) %>% 
  clean_names("snake")
```

We can also remove all excess white space from character variables.

```{r raw_trim}
nhc <- mutate(nhc, across(is.character, str_squish))
```

## Explore

```{r glimpse}
glimpse(nhc)
tail(nhc)
```

### Missing

```{r na_count}
col_stats(nhc, count_na)
```

_Very_ few records are missing a name, date, or amount.

```{r na_flag}
nhc <- nhc %>% flag_na(date, contributor, amount, recipient)
sum(nhc$na_flag)
```

```{r na_view}
nhc %>% 
  filter(na_flag) %>% 
  select(date, contributor, amount, recipient)
```

### Duplicates

There are however a large number of duplicate records. We can flag every record
that has duplicate values across every column.

```{r dupe_flag}
nhc <- flag_dupes(nhc, everything())
percent(mean(nhc$dupe_flag), 0.01)
```

```{r dupe_view}
nhc %>% 
  filter(dupe_flag) %>% 
  select(date, contributor, amount, recipient) %>% 
  arrange(date, contributor)
```

If we count the contributor and recipient names, we see most of these duplicates
are unitemized contributions handled through a PAC or service like ActBlue.

```{r dupe_count}
nhc %>% 
  filter(dupe_flag) %>% 
  count(contributor, recipient, sort = TRUE) %>% 
  mutate(p = n/sum(n))
```

### Categorical

```{r n_distinct}
col_stats(nhc, n_distinct)
```

```{r distinct_bars}
explore_plot(nhc, con_type)
explore_plot(nhc, rec_type)
explore_plot(nhc, cycle, nbar = 5)
explore_plot(nhc, method)
explore_plot(nhc, in_kind)
```


### Amounts

```{r ammount_summary}
summary(nhc$amount)
sum(nhc$amount <= 0.01)
```

```{r hist_amount, echo=FALSE}
nhc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New Hampshire Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nhc <- mutate(nhc, year = year(date))
```

```{r date_range}
min(nhc$date)
sum(nhc$year < 2000)
max(nhc$date)
sum(nhc$date > today())
```

```{r bar_year, echo=FALSE}
nhc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2016, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Hampshire Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

First, we notice that the vast majority of the existing `city_old` and 
`state_old` variables are missing.

```{r old_count}
prop_na(nhc$city_old)
prop_na(nhc$state_old)
nhc %>% 
  select(geo_full, city_old, state_old) %>% 
  filter(!is.na(city_old) & !is.na(state_old))
```

However, the city and state values are present in the full contributor address
line, which can be split into it's components with `tidyr::separate()`.

```{r old_sep}
nhc <- nhc %>% 
  mutate(across(geo_full, na_in, c("UnItemized", "Unitemized"))) %>% 
  separate(
    col = geo_full,
    into = c(glue("addr{1:10}"), "city", "state_zip"),
    sep = "\\s?,\\s",
    remove = FALSE,
    fill = "left",
    extra = "merge"
  ) %>% 
  unite(
    col = addr,
    starts_with("addr"),
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  mutate(across(addr, na_if, "")) %>% 
  separate(
    col = state_zip,
    into = c("state", "zip"),
    sep = "\\s(?=\\d)",
    remove = TRUE
  ) %>% 
  mutate(
    city = coalesce(city, city_old),
    state = coalesce(state, state_old)
  )
```

```{r old_view}
nhc %>% 
  select(geo_full, addr, city, state, zip) %>% 
  distinct() %>% 
  sample_n(20)
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
nhc <- nhc %>% 
  mutate(
    addr_norm = normal_address(
      address = addr,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
nhc %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nhc <- nhc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nhc$zip,
  nhc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nhc <- nhc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
nhc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  nhc$state,
  nhc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
nhc <- nhc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NH", "DC", "NEW HAMPSHIRE"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
nhc <- nhc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- nhc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
nhc <- nhc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(nhc$city_raw),
  nhc$city_norm,
  nhc$city_swap,
  nhc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nhc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New Hampshire City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Hampshire City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
nhc <- nhc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  select(-addr, -city, -state, -zip)
```

```{r clean_glimpse}
glimpse(sample_frac(nhc))
```

1. There are `r comma(nrow(nhc))` records in the database.
1. There are `r comma(sum(nhc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nhc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nh", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "nh_contribs_clean.csv")
write_csv(nhc, clean_path, na = "")
file_size(clean_path)
```

The encoding of the exported file should be UTF-8 or ASCII.

```{r clean_encode}
enc <- system2("file", args = paste("-i", clean_path), stdout = TRUE)
str_replace_all(enc, clean_path, basename)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(nhc)),
  type = md_code(map_chr(nhc, typeof)),
  def = c(
    "Recipient unique ID",
    "Date contribution was made",
    "Contributor type",
    "Contributor name",
    "Contributor full address",
    "Recipient committee name",
    "Recipient type",
    "Recipient office sought",
    "Election county",
    "Election cycle",
    "Contribution period reported",
    "Contribution method",
    "Contribution amount or correction",
    "Total ammount contributor given",
    "Contribution comments",
    "Contribution In-Kind category",
    "Original contributor city",
    "Original contributor state",
    "Contributor occupation",
    "Contributor employer name",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Separated & normalized combined street address",
    "Separated & normalized 5-digit ZIP code",
    "Separated & normalized state abbreviation",
    "Separated & normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# New Hampshire Contributions Data Dictionary\n", dict_md),
  path = here("nh", "contribs", "nh_contribs_dict.md"),
)
```

