---
title: "New York Lobbying Expenditure Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","ny", "lobby", "exp", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  pdftools, #wrangle pdf files
  readxl, #read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The [New York State Joint Commission on Public Ethics](https://reports.ethics.ny.gov/publicreports/Default.aspx) started a new system for lobbyist registration from 2019 on. 

For lobbying expenditure records, we will download the [lobbyist dataset][https://reports.ethics.ny.gov/publicreports/Default.aspx?r=01-LobbyistDataset] generated from the [public search query][https://reports.ethics.ny.gov/PublicQuery] page. Data is available from 2019 to 2023, with 2020 data being the same as 2019, and 2022 the same as 2021. In this update, we'll run the script to process 2019-2022.
## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("state","ny", "lobby", "exp", "data", "raw"))
```

### Read
Note that since there's not an easy way of determining date of expenditures, we create a `year` field that's the first year of the two-year report period. For example, if the record comes from the 2019-2020 report, the year field would be 2019.
```{r raw_read}
nyle <- read_csv(dir_ls(raw_dir, regexp =".+LobbyistDataset.+.csv"),id="path") %>% clean_names()

nyle <- nyle %>% rename(contractual_client = client1,
                        beneficial_client = bene1)
```

```{r}
nyle <- nyle %>% 
  mutate(year = str_extract(path, "\\d{4}")) %>% 
  select(-path)
```

We will convert the string columns to numeric.
```{r}
nyle <- nyle %>% 
  mutate(across(5:8, ~ str_remove(.,"^\\$") %>% str_remove(",") %>% as.numeric()))
```

## Explore

```{r glimpse}
head(nyle)
tail(nyle)
glimpse(sample_n(nyle, 20))
```

### Missing

```{r na_count}
col_stats(nyle, count_na)
```
There are many fields missing,m especially for fields in columns after `total_reimbursed`.

### Duplicates
Running the following commands shows that there are over a thousand entries with totally identical information when we exclude the time period we created.  We will note that in the `dupe_flag` column.

We will exclude all the 
```{r dupe_flag}
nyle <- flag_dupes(nyle, 1:(ncol(nyle)-1))
sum(nyle$dupe_flag)
```

### Categorical

```{r n_dinyleinct}
col_stats(nyle, n_distinct)
```

### Amounts


```{r plot_amt_nonlog, echo=FALSE}
nyle %>%
  filter(compensation >= 1 & !dupe_flag) %>% 
  ggplot(aes(compensation)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "NY Lobbying compensation,2019-2022",
    y = "Count",
    x = "Amount"
  )
```
We can take a look at the top compensated principal lobbyists. 
```{r}
nyle %>% 
  filter(compensation>1 & dupe_flag) %>% 
  group_by(principal_lobbyist) %>% 
  summarize(total_comp = sum(compensation)) %>% 
  arrange(desc(total_comp)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = total_comp, y = reorder(principal_lobbyist, total_comp))) + 
  geom_col() + 
  scale_x_continuous(
    labels = dollar
  ) +
  geom_col(fill = "#FAAB18") + 
    labs(
    title = "Top Lobbyists (by compensation)",
    caption = "Source: New York State Commission on Ethics and Lobbying in Government",
    x = "Compensation",
    y = "Principal Lobbyist"
  )
  
```

Similarly, we can look at which principal lobbyists spent the most on non-lobbying expenditures

```{r}

nyle %>% 
  filter(compensation>1 & dupe_flag) %>% 
  group_by(principal_lobbyist) %>% 
  summarize(total_exp = sum(non_lobbying_exp)) %>% 
  arrange(desc(total_exp)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = total_exp, y = reorder(principal_lobbyist, total_exp))) + 
  geom_col() + 
  scale_x_continuous(
    labels = dollar
  ) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) + 
    labs(
    title = "Top Lobbyists (by non-lobbying expenditures)",
    caption = "Source: New York State Commission on Ethics and Lobbying in Government",
    x = "Expenditures",
    y = "Principal lobbyist"
  )
```


## Conclude

```{r clean_glimpse}
glimpse(sample_n(nyle, 20))
```


1. There are `r nrow(nyle)` records in the database.
1. There are `r sum(nyle$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(nyle$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `YEAR` variable has been created with `lubridate::year()`.


## Export

```{r clean_dir}
clean_dir <- dir_create(here("state","ny", "lobby", "data", "processed","exp"))
```

```{r write_clean}
write_csv(
  x = nyle,
  path = path(clean_dir, "ny_lob_exp_2019-2022.csv"),
  na = ""
)
```
### Dictionary
