---
title: "New York Lobbying Registration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ny", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  pdftools, #wrangle pdf files
  readxl, #read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The [New York State Joint Commission on Public Ethics](https://reports.ethics.ny.gov/publicreports/Default.aspx) started a new system for lobbyist registration from 2019 on. For data between 2011 and 2018, see `nylr_reg_diary.Rmd` for data processing notes. 

For lobbying registration records, we will download the data generated from the [public search query][https://reports.ethics.ny.gov/PublicQuery] page. Go to [public search query][https://reports.ethics.ny.gov/PublicQuery] and select by year. Data is available from 2019 to 2023, with 2020 data being the same as 2019, and 2022 the same as 2021. In this update, we'll run the script to process the 2019-2022 data. 

There are two different types of clients.
> A Contractual Client means an individual or Organization that retains the services of a Lobbyist for the benefit of itself or another.

> A Beneficial Client is the specific individual or Organization on whose behalf and at whose request Lobbying Activity is conducted.

> An individual or Organization that lobbies on its own behalf is both the Beneficial Client and the Contractual Client.
> While a Contractual and Beneficial Client must be identified for every lobbying arrangement, the Contractual Client may also be the Beneficial Client.

> Lobbyist includes the lobbying firm or organization (the Principal Lobbyist for filing purposes) and every person identified on a Statement of Registration, as well as any equity partners, officers, or directors of the organization who operate out of the same geographic office as any person identified on a Statement of Registration.
## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("state","ny", "lobby", "reg", "data", "raw"))
```

### Read

```{r raw_read}
nylr <- dir_ls(raw_dir, regexp =".+Search.+.xlsx")  %>% map_dfr(read_xlsx) %>% clean_names()

nylr <- nylr %>% select(-associated_filings)

nylr <- nylr %>% rename(ind_lob = individual_lobbyist_s,
                        co_sub_lob = co_lobbyist_s_sub_lobbyist_s,
                        prin_lob = principal_lobbyist) %>% 
  mutate_all(str_remove, "^\r\n")
```

### Separate address
The name and address of principal lobbyists, clients and beneficial clients are all in the same field. We will first separate these fields.
```{r}
nylr <- nylr %>% 
  separate(
    col = prin_lob,
    into = c("prin_lob_name", "state_zip"),
    sep = "\r\n\r\n",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) 
  
 nylr <- nylr %>%  
  mutate(st_zip = str_extract(state_zip, "(?<=,\\s)[A-Z]{2}\\s\\d{5}"),
    prin_lob_state = str_extract(st_zip,"[A-Z]{2}"),     
    prin_lob_zip = str_remove(st_zip, "[A-Z]{2}") %>% str_trim(),
    addr_city = str_remove(state_zip, ",\\s[A-Z]{2}\\s\\d{5}.*\r\n") %>% str_trim(),
    prin_lob_city = str_extract(addr_city, "(?<=\r\n).+$"),
    prin_lob_addr = str_remove(addr_city,"\r\n.+$") %>% str_trim()
      # separate(col, into = c("a", "b"), sep="\r\n(?=[^\r\n]+$)")
         ) %>% 
   select(-c(state_zip, addr_city, st_zip))
```
We will repeat the process for `contractual_client` and `beneficial_client`.
```{r}
nylr <- nylr %>% 
  separate(
    col = contractual_client,
    into = c("cont_client_name", "state_zip"),
    sep = "\r\n\r\n",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) 
  
 nylr <- nylr %>%  
  mutate(st_zip = str_extract(state_zip, "(?<=,\\s).+\\s\\d{5}"),
    cont_client_state = str_extract(st_zip,".+(?=\\s\\d{5})"),     
    cont_client_zip = str_remove(st_zip, cont_client_state) %>% str_trim(),
    addr_city = str_remove(state_zip, ",\\s.+\\s\\d{5}.*"),
    cont_client_city = str_extract(addr_city, "(?<=\r\n)[^\r]+\r\n\r\n") %>% str_trim(),
    cont_client_addr = str_remove(addr_city, cont_client_city) %>% str_trim()
      # separate(col, into = c("a", "b"), sep="\r\n(?=[^\r\n]+$)")
         ) %>% 
   select(-c(state_zip, addr_city, st_zip))
```

```{r}
nylr <- nylr %>% 
  separate(
    col = beneficial_client,
    into = c("ben_client_name", "state_zip"),
    sep = "\r\n\r\n",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) 
  
 nylr <- nylr %>%  
  mutate(st_zip = str_extract(state_zip, "(?<=,\\s).+\\s\\d{5}"),
    #ben_client_state = str_extract(st_zip,".+(?=\\s\\d{5})"),     
    ben_client_state = str_extract(st_zip,campfin::rx_state),     
    ben_client_zip = str_remove(st_zip, ben_client_state) %>% str_extract(campfin::rx_zip) %>% str_trim(),
    addr_city = str_remove(state_zip, ",\\s.+\\s\\d{5}.*"),
    ben_client_city = str_extract(addr_city, "(?<=\r\n)[^\r]+(?=\r\n\r\n)") %>% str_trim() %>% replace(., .%in% c("TH FLOOR", "SUITE\\s*\\d*"), NA_character_),
    ben_client_addr = str_remove(addr_city, ben_client_city) %>% str_trim()
      # separate(col, into = c("a", "b"), sep="\r\n(?=[^\r\n]+$)")
         ) %>% 
   select(-c(state_zip, addr_city, st_zip))
   
nlyr <- nylr %>% 
     mutate(across(starts_with("ben_"), str_trim))
  
```

## Explore

```{r glimpse}
head(nylr)
tail(nylr)
glimpse(sample_n(nylr, 20))
```

### Missing

```{r na_count}
col_stats(nylr, count_na)
```
There are many fields missing,m especially for fields in columns after `total_reimbursed`.

### Duplicates
Running the following commands shows that there are over a thousand entries with totally identical information. Since each of them contain `total_compensation` and `total_reimbursed` for a certain period, there should be only one such entry for each corresponding time period. We will note that in the `dupe_flag` column.
```{r dupe_flag}
nylr <- flag_dupes(nylr, dplyr::everything())
sum(nylr$dupe_flag)
```

### Categorical

```{r n_dinylrinct}
col_stats(nylr, n_distinct)
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.


### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
nylr <- nylr %>% 
    mutate_at(.vars = vars(ends_with('addr')), .funs = list(norm = ~ normal_address(.,abbs = usps_street,
      na_rep = TRUE)))
```

```{r address_view}
nylr %>% 
  select(contains("addr")) %>% 
  distinct() %>% 
  sample_n(10) %>% 
  glimpse()
```

### ZIP

The zipcode fields are largely valid and canwoe can't really improve the accuracy with `zip_norm`, so we will leave them as is. 

```{r zip_progress}
progress_table(
  nylr$prin_lob_zip,
  nylr$ben_client_zip,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nylr <- nylr %>% 
   mutate_at(.vars = vars(ends_with('state')), .funs = list(norm = ~ normal_state(.,abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state)))
```

```{r state_view}
nylr %>% 
  filter(prin_lob_state != prin_lob_state_norm) %>% 
  count(prin_lob_state, sort = TRUE)
```

```{r firm_state_view}
nylr %>% 
  filter(ben_client_state != ben_client_state_norm) %>% 
  count(ben_client_state, sort = TRUE)
```

```{r state_progress}
progress_table(
  nylr$prin_lob_state,
  nylr$prin_lob_state_norm,
  nylr$ben_client_state,
  nylr$ben_client_state_norm,
  compare = valid_state
)
```

```{r}
nylr <- nylr %>% select(-prin_lob_state_norm)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good nylrart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
nylr <- nylr %>% 
      mutate_at(.vars = vars(ends_with('city')), .funs = list(norm = ~ normal_city(.,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE)))

prop_in(nylr$prin_lob_city_norm, valid_city, na.rm = T)
prop_in(nylr$ben_client_city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
nylr <- nylr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "prin_lob_state" = "state",
      "prin_lob_zip" = "zip"
    )
  ) %>% 
  rename(prin_lob_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(prin_lob_city_norm, prin_lob_city_match),
    match_dist = str_dist(prin_lob_city_norm, prin_lob_city_match),
    prin_lob_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = prin_lob_city_match,
      false = prin_lob_city_norm
    )
  ) %>% 
  select(
    -prin_lob_city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_swap clients}
nylr <- nylr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "ben_client_state_norm" = "state",
      "ben_client_zip" = "zip"
    )
  ) %>% 
  rename(ben_client_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(ben_client_city_norm, ben_client_city_match),
    match_dist = str_dist(ben_client_city_norm, ben_client_city_match),
    ben_client_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = ben_client_city_match,
      false = ben_client_city_norm
    )
  ) %>% 
  select(
    -ben_client_city_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  nylr$prin_lob_city,
  nylr$ben_client_city,
  nylr$prin_lob_city_norm,
  nylr$ben_client_city_norm,
  nylr$prin_lob_city_swap,
nylr$ben_client_city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New York City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "NY City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

### Separate individual lobbyists
We then need to separate individual lobbyists all retained by a single principal lobbyist.
```{r sep ind lob}
nylr <- nylr %>% 
  mutate(ind_lob = str_remove(ind_lob, "\r\n$")) %>% 
  separate_rows(ind_lob, sep = "\r\n")
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(nylr, 20))
```

## Rename

```{r}
nylr <- nylr  %>% 
  rename(prin_lob_city_clean = prin_lob_city_swap) %>% 
  rename( ben_client_city_clean = ben_client_city_swap) %>% 
  select(-c(prin_lob_city_norm, ben_client_city_norm)) %>% 
  rename_with(~str_replace(., "_norm", "_clean"))
```

1. There are `r nrow(nylr)` records in the database.
1. There are `r sum(nylr$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(nylr$na_flag)` records missing either recipient or date.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `YEAR` variable has been created with `lubridate::year()`.


## Export

```{r clean_dir}
clean_dir <- dir_create(here("state","ny", "lobby", "data", "processed","reg"))
```

```{r write_clean}
write_csv(
  x = nylr,
  path = path(clean_dir, "ny_lob_reg_2019-2022.csv"),
  na = ""
)
```
