---
title: "Pennsylvania Contributions"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("pa", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # format markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  aws.s3, # aws cloud storage
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  httr, # make http requests
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is from the [Pennsylvania Election and Campaign Finance System (ECF)][ecf].

[ecf]: https://www.dos.pa.gov/VotingElections/CandidatesCommittees/CampaignFinance/Pages/default.aspx

The ECF provides a [Full Campaign Finance Export][data]. From this page,
files are organized as annual directories containing files for contributions,
debt, expenditures, filer information, and receipts.

[data]: https://www.dos.pa.gov/VotingElections/CandidatesCommittees/CampaignFinance/Resources/Pages/FullCampaignFinanceExport.aspx

The ECF also provides a `readme.txt` file, which we can read to better
understand the data we will be downloading.

```{r vars_read}
pa_host <- "https://www.dos.pa.gov/VotingElections"
pa_dir <- "CandidatesCommittees/CampaignFinance/Resources/Documents"
readme_file <- "readmepriorto2022.txt"
readme_url <- paste(pa_host, pa_dir, readme_file, sep = "/")
```

This text file provides the column names and types for the each of the data
files included in the raw download.

```{r vars_print, echo=FALSE}
readme <- readme_url %>% 
  read_lines(skip = 4) %>% 
  str_c(collapse = "\r\n") %>% 
  str_split("\\W(?=\\w+\r\n-+)") %>% 
  pluck(1) %>%
  map(
    ~read_lines(.) %>% 
      str_trim() %>%
      str_remove("(?<=\\))(.*)$") %>% 
      enframe(name = NULL) %>% 
      separate(
        col = value, 
        into = c("col", "type"), 
        sep = "\t+",
        extra = "merge",
        fill = "right"
      ) %>% 
      #na_if("") %>% 
      drop_na() %>% 
      mutate(col = str_to_lower(col))
  ) %>% 
  set_names(c("contribs", "debt", "expense", "filer", "receipt"))
```

Here are the columns included in the contributions data set:

```{r vars_contrib}
readme$contribs %>% 
  mutate(col = as.character(md_code(col))) %>% 
  kable(col.names = c("Columns", "Type"))
```

There are no variables providing information on the recipients, those committees
which _filed_ the reports containing each contribution. That data is provided in
a separate file.

```{r vars_filer}
readme$filer %>% 
  mutate(col = as.character(md_code(col))) %>% 
  kable(col.names = c("Columns", "Type"))
```

## Import 

To import the files into R, we will have to first download the annual ZIP 
archive file containing all campaign finance transactions. Then we can extract
the contributions file and read them all into a single file for processing.

### Download

Each ZIP archive is simply named as the 4-digit year for the files within.
We can download each ZIP to the `/data/raw` directory.

```{r raw_download}
zip_names <- paste(2000:2022, "zip", sep = ".")
zip_urls <- paste(pa_host, pa_dir, zip_names, sep = "/")
raw_dir <- dir_create(here("state","pa", "contribs", "data", "raw"))
zip_paths <- path(raw_dir, zip_names)
fix_check <- here("state","pa", "contribs", "data", "fixed.txt")
if (length(dir_ls(raw_dir, regexp = "zip")) < 10) {
  for (i in seq_along(zip_urls)) {
  download.file(zip_urls[i], zip_paths[i])
  #file_delete(fix_check)
  }
}
```

Then we will unzip the annual directory from each archive.

```{r raw_unzip, eval=FALSE}
zip_files <- dir_ls(raw_dir, glob = "*.zip")

if (all_files_new(path = raw_dir, glob = "*.txt")) {
  for (i in seq_along(zip_files)) {
    unzip(
      zipfile = zip_files[i],
      #Matches the csv files that starts with contrib, and trim the "./ " from directory names
      files = grep("(contrib|filer).+", unzip(zip_files[i]), value = TRUE) %>% substring(3,),
      exdir = raw_dir
    )
  }
}
```

For each year, there is a file for contribution and a file for the information
on the recipients of those contributions, who file the reports containing the
data. We will identify the path of each file type in new vectors, which can then
be read together.

```{r con_move}
con_paths <- dir_ls(
  path = raw_dir, 
  recurse = TRUE, 
  regexp = "(C|c)ontrib[\\.|_]"
)

fil_paths <- dir_ls(
  path = raw_dir, 
  recurse = TRUE, 
  regexp = "(F|f)iler[\\.|_]"
) %>% 
c(dir_ls(path = here("state","pa", "expends", "data", "raw"),
                recurse = TRUE, 
  regexp = "(F|f)iler[\\.|_]"))
```

The file names are a little different year to year, but they all have the same
format.

```{r con_list, echo=FALSE}
con_paths %>% 
  str_replace("(.*)/R_campfin", "~") %>% 
  md_code() %>% 
  md_bullet()
```

### Fix

To properly read so many records, we need to first perform some manipulation of
the text files. Each "cell" of character type columns are surrounded in double
quotation marks (`"`) to help prevent misreading. However, some of the text in
these cells itself contains double quotes or newline characters (`\n`).

We need to read each file as a character string and use regular expressions to
identify these erroneous characters and remove or replace them.

```{r raw_fix}
# do not repeat if done
if (!file_exists(fix_check)) {
  # for all contrib and filer files
  for (f in c(con_paths, fil_paths)) {
    # read raw file
    read_file(f) %>% 
      # force conversion to simple
      iconv(to = "ASCII", sub = "") %>% 
      # replace non-carriage newline
      str_replace_all("(?<!\r)\n", " ") %>%
      # replace not-field double quotes
      str_replace_all("(?<!^|,|\r\n)\"(?!,|\r\n|$)", "\'") %>% 
      # replace non-delim commas
      str_remove_all(",(?!\"|\\d|\\.\\d+|-(\\d|\\.))") %>% 
      # overwrite raw file
      write_file(f)
    # check progress
    message(paste(basename(f), "done"))
    # clean garbage memory
    flush_memory()
  }
  # note this has done
  file_create(fix_check)
}
```

### Read

Now that each text file has been cleaned of irregularities, they can each be
properly read into R.

First, we will read all the annual contribution files into a single data frame
using `vroom::vroom()`. We need to use the column names and types listed in the
`readme.txt` file we downloaded earlier.

```{r raw_read}
pac <- map_df(
  .x = con_paths,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_names = readme$contribs$col,
  col_types = cols(
    .default = col_skip(),
    filerid = col_character(),
    eyear = col_integer(),
    cycle = col_integer(),
    section = col_character(),
    contributor = col_character(),
    address1 = col_character(),
    address2 = col_character(),
    city = col_character(),
    state = col_character(),
    zipcode = col_character(),
    occupation = col_character(),
    ename = col_character(),
    contdate1 = col_date("%Y%m%d"),
    contamt1 = col_double()
  )
)
```

Then we can read the fixed filer files to describe the recipients.

```{r fil_read}
filers <- map_df(
  .x = fil_paths,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_names = readme$filer$col,
  col_types = cols(
    .default = col_skip(),
    eyear = col_integer(),
    filerid = col_character(),
    filertype = col_character(),
    filername = col_character(),
    office = col_character(),
    district = col_character(),
    party = col_character(),
    address1 = col_character(),
    address2 = col_character(),
    city = col_character(),
    state = col_character(),
    zipcode = col_character(),
    county = col_character(),
    phone = col_character(),
  )
)
```

We only want to join a single filer to each contribution listed in the data.
We can group by the unique filer ID and a filing year and select only one
copy of the data.

```{r fil_slice}
nrow(filers)
filers <- filers %>% 
  group_by(filerid, eyear) %>% 
  slice(1) %>% 
  ungroup()
nrow(filers)
```

Now the filer information can be added to the contribution data with a 
`dplyr::left_join()` along the unique filer ID and election year.

```{r raw_join}
# 21,761,034
pac <- left_join(
  x = pac,
  y = filers,
  by = c("filerid", "eyear"),
  suffix = c("_con", "_fil")
)

rm(filers)

pac <- rename_prefix(
  df = pac,
  suffix = c("_con", "_fil"),
  punct = TRUE
)
```

```{r raw_rename}
pac <- pac %>% 
  rename_with(~str_replace(., "address", "addr")) %>% 
  rename(
    con_zip = con_zipcode,
    date = contdate1,
    amount = contamt1,
    fil_type = filertype,
    filer = filername,
    fil_zip = fil_zipcode,
    fil_phone = phone
  )
```

We will also add a temporary unique ID for each transaction.

```{r raw_id}
pac <- mutate(pac, tx = row_number())
```

```{r include=FALSE}
flush_memory()
```

## Explore

We should first check the top and bottom of the read data frame to ensure the
file was read correctly. This view also helps simply understand the format.

There are `r comma(nrow(pac))` rows of `r ncol(pac)` columns.

```{r glimpse}
glimpse(pac)
head(pac)
```

Checking the number of distinct values of a discrete variable is another good
way to ensure the file was read properly.

```{r read_check}
count(pac, fil_type)
```

### Missing

We should first check the number of missing values in each column.

```{r na_count}
col_stats(pac, count_na)
```

Any record missing a date, name, or amount should be flagged. These variables
are key to identifying transactions.

```{r na_flag}
key_vars <- c("date", "contributor", "amount", "filer")
pac <- flag_na(pac, all_of(key_vars))
percent(mean(pac$na_flag), 0.01)
```

```{r na_view}
pac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_n(10)
```

All of the records missing a value are missing a `date`.

```{r na_recount}
pac %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

We can check for records that are entirely duplicated across every variable
using `duplicated()`. This process is memory inefficient, so we will split our
data frame into a list of 100,000 row chunks and check each chunk at a time,
appending our duplicate rows to a local text file.

```{r dupe_id}
pac <- mutate(pac, tx = row_number())
```

```{r dupe_write}
dupe_file <- here("pa", "contribs", "dupes.txt")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  pac <- mutate(pac, group = str_sub(date, end = 7))
  pa_ids <- split(pac$tx, pac$group)
  pas <- pac %>% 
    select(-tx) %>% 
    group_split(group)
  pb <- txtProgressBar(max = length(pas), style = 3)
  pac <- select(pac, -group)
  flush_memory(1)
  for (i in seq_along(pas)) {
    d1 <- duplicated(pas[[i]], fromLast = FALSE) # check from front
    d2 <- duplicated(pas[[i]], fromLast = TRUE) # check from back
    dupes <- tibble(tx = pa_ids[[i]], dupe_flag = d1 | d2)
    dupes <- filter(dupes, dupe_flag == TRUE) # remove non dupes
    write_csv(dupes, dupe_file, append = TRUE) # append to disk
    rm(d1, d2, dupes); pas[[i]] <- NA # remove for memory
    Sys.sleep(10)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(pas, pb)
  flush_memory()
}
```

We can now read that file and join it against the contributions.

```{r dupe_read}
dupes <- read_csv(
  file = dupe_file,
  col_names = c("tx", "dupe_flag"),
  col_types = cols(
    tx = col_double(),
    dupe_flag = col_logical()
  )
)
comma(nrow(dupes))
```

```{r dupe_join}
pac <- left_join(pac, dupes, by = "tx")
pac <- mutate(pac, dupe_flag = !is.na(dupe_flag))
percent(mean(pac$dupe_flag), 0.01)
```

```{r echo=FALSE}
rm(dupes)
flush_memory()
```

```{r dupe_view}
pac %>% 
  filter(dupe_flag) %>% 
  select(tx, all_of(key_vars))
```

```{r}
pac <- pac %>% select(-tx)
```


### Amounts

The range and distribution of the contribution `amount` should be checked. We
also want to note what percentage of the values are zero or below.

```{r ammount_summary}
summary(pac$amount)
mean(pac$amount <= 0)
```

```{r amount_max}
glimpse(pac[c(which.min(pac$amount), which.max(pac$amount)), ])
```

```{r hist_amount, echo=FALSE}
pac %>%
  filter(!is.na(amount), amount >= 1, amount <= 1e6) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Pennsylvania Contributions Amount Distribution",
    subtitle = "from 2000 to 2022",
    x = "Amount",
    y = "Count"
  )
```

```{r violin_amount_party, echo=FALSE, eval=FALSE}
pac %>%
  filter(amount >= 1, amount <= 1e6) %>% 
  filter(party %in% most_common(party, 3) | is.na(party)) %>% 
  ggplot(aes(party, amount)) +
  geom_violin(
    mapping = aes(fill = party),
    trim = TRUE,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "area",
    adjust = 4
  ) +
  scale_fill_brewer(
    palette = "Dark2", 
    guide = FALSE,
    na.value = "#3d4043"
  ) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Pennsylvania Contributions Amount by Party",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year a contribution was made using `lubridate::year()`. 

```{r date_year}
pac <- mutate(pac, year = year(date))
```

There are a handful of missing or irregular dates.

```{r date_range}
percent(prop_na(pac$date), 0.01)
min(pac$date, na.rm = TRUE)
sum(pac$year < 2000, na.rm = TRUE)
max(pac$date, na.rm = TRUE)
sum(pac$date > today(), na.rm = TRUE)
```

The bulk of transactions occur between 2000 and 2022.

```{r bar_year, echo=FALSE}
pac %>% 
  count(year) %>%
  filter(!is.na(year)) %>% 
  mutate(odd = !is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = odd)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2022, by = 2)) +
  coord_cartesian(xlim = c(2000, 2022)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania Contributions by Year",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

```{r include=FALSE}
flush_memory()
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `*_addresss` variables, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r con_addr_norm}
norm_addr <- pac %>%
  count(con_addr1, con_addr2, sort = TRUE) %>% 
  select(-n) %>% 
  unite(
    col = con_addr_full,
    starts_with("con_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    con_addr_norm = normal_address(
      address = con_addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-con_addr_full)
```

```{r addr_join}
norm_addr
pac <- left_join(pac, norm_addr)
rm(norm_addr); flush_memory(1)
```

We will repeat the process for filer addresses.

```{r fil_addr_norm}
norm_addr <- pac %>% 
  count(fil_addr1, fil_addr2, sort = TRUE) %>% 
  select(-n) %>% 
  unite(
    col = fil_addr_full,
    starts_with("fil_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    fil_addr_norm = normal_address(
      address = fil_addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-fil_addr_full)
```

```{r fil_addr_join}
norm_addr
pac <- left_join(pac, norm_addr)
rm(norm_addr); flush_memory(1)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
pac <- mutate_at(
  .tbl = pac,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_progress}
progress_table(
  pac$con_zip,
  pac$con_zip_norm,
  compare = valid_zip
)
```

### State

There is no need to clean the two state variables.

```{r state_check}
prop_in(pac$con_state, valid_state)
prop_in(pac$fil_state, valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm_con}
con_norm_city <- pac %>% 
  count(con_city, con_state, con_zip_norm, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    across(
      .cols = con_city, 
      .fns = list(norm = normal_city), 
      abbs = usps_city,
      states = c("PA", "DC", "PENNSYLVANIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r city_norm}
fil_norm_city <- pac %>% 
  count(fil_city, fil_state, fil_zip_norm, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    across(
      .cols = fil_city, 
      .fns = list(norm = normal_city), 
      abbs = usps_city,
      states = c("PA", "DC", "PENNSYLVANIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r con_city_swap}
con_norm_city <- con_norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "con_state" = "state",
      "con_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(con_city_norm, city_match),
    match_dist = str_dist(con_city_norm, city_match),
    con_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = con_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r fil_city_swap}
fil_norm_city <- fil_norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "fil_state" = "state",
      "fil_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(fil_city_norm, city_match),
    match_dist = str_dist(fil_city_norm, city_match),
    fil_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = fil_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r include=FALSE}
flush_memory()
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- con_norm_city %>% 
  mutate(
    con_city_refine = con_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(con_city_refine != con_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "con_city_refine" = "city",
      "con_state" = "state",
      "con_zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    con_state, 
    con_zip_norm, 
    con_city_swap, 
    con_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
con_norm_city <- con_norm_city %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(con_city_refine = coalesce(con_city_refine, con_city_swap))
```

```{r include=FALSE}
flush_memory()
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r check_filter}
pac_out <- con_norm_city %>% 
  filter(con_city_refine %out% c(valid_city, extra_city)) %>% 
  count(con_city_refine, con_state, sort = TRUE) %>% 
  drop_na() %>% 
  head(1000)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("state","pa", "contribs", "data", "api_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file,
    col_types = cols(
      .default = col_character(),
      check_city_flag = col_logical()
    )
  )
} else {
  check <- pmap_dfr(
    .l = list(
      pac_out$con_city_refine, 
      pac_out$con_state
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    path = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  c(valid_locality)
```

```{r}
valid_locality <- c(valid_locality, "ABBOTT PARK", "RESEARCH TRIANGLE PARK")
```

#### Progress

```{r con_city_fix}
con_norm_city <- con_norm_city %>% 
  mutate(
    con_city_refine = con_city_refine %>% 
      na_if("ILLEGIBLE") %>% 
      str_replace("^PHILA$", "PHILADELPHIA") %>% 
      str_replace("^PGH$", "PITTSBURGH") %>% 
      str_replace("^NEW YORK CITY$", "NEW YORK") %>% 
      str_replace("^H\\sBURG$", "HARRISBURG") %>% 
      str_replace("^HBG$", "HARRISBURG") %>% 
      str_replace("^NYC$", "NEW YORK")
  )
```

```{r city_rejoin}
pac <- left_join(pac, con_norm_city)
pac <- left_join(pac, fil_norm_city)
```

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city, valid_locality)
progress <- progress_table(
  str_to_upper(pac$con_city),
  pac$con_city_norm,
  pac$con_city_swap,
  pac$con_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r city_final_prop}
prop_in(pac$con_city_refine, many_city)
prop_in(pac$fil_city_swap, many_city)
```

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(pac$con_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

```{r include=FALSE}
flush_memory()
```

## Conclude

```{r clean_select}
pac <- pac %>% 
  select(
    -con_city_norm,
    -con_city_swap,
    con_city_clean = con_city_refine
  ) %>% 
  select(
    -fil_city_norm,
    fil_city_clean = fil_city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  relocate(ends_with("city_clean"), .after = fil_addr_clean)
```

```{r clean_glimpse}
glimpse(sample_n(pac, 100))
```

1. There are `r comma(nrow(pac))` records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r percent(mean(pac$na_flag), 0.01)` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("state","pa", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "pa_contribs_clean.csv")
write_csv(pac, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(pac)),
  type = md_code(map_chr(pac, typeof)),
  def = c(
    "Filer unique filer ID",
    "Election year",
    "Election cycle",
    "Election section",
    "Contributor full name",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Date contribution made",
    "Contribution amount or correction",
    "Filer type",
    "Filer committee name",
    "Filer office sought",
    "District election held",
    "Filer political party",
    "Filer street address",
    "Filer secondary address",
    "Filer city name",
    "Filer 2-digit state abbreviation",
    "Filer ZIP+4 code",
    "County election held in",
    "Filer telephone number",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized contributor street address",
    "Normalized Filer street address",
    "Normalized Filer 5-digit ZIP code",
    "Normalized Filer state abbreviation",
    "Normalized Filer city name",
    "Normalized contributor city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
