---
title: "Utah Contributions"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ut", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  cli, # command line
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("ut/contribs/docs/ut_contribs_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

Campaign expenditures are from the Utah [Financial Disclosures website][udw].

[udw]: https://disclosures.utah.gov/

> #### Reporting Contributions and Expenditures:
> The general requirement for contribution reporting is to report each
contribution you receive within 31 days of receiving it. However, there are
additional requirements that must be followed: ...
> 
> * All expenditures must be reported by the reporting deadline for each
reporting period.
> * All additional contributions received during a reporting period must be
reported by the reporting deadline.
> * All filing reports must be filled by no later than 11:59 p.m. on the day of
the reporting deadline. Failure to file a report on time may result in a fine.

## Download

Using the website's [advance search feature][adv], users can search along a
name, entity type, and report year. The site gives ambiguous directions, but
we can obtain _all_ records for a year and entity type if we leave the name
blank and click on the top record.

>  Please follow these steps:
> * Select an Entity Type
> * Select a Report Year
> * Click Find
> * In the top row of results, click the year under the far right column --
"Download Data by Year"

We make these direct HTTP requests using `httr::GET()`.

[adv]: https://disclosures.utah.gov/Search/AdvancedSearch

```{r raw_dir}
raw_dir <- dir_create(here("ut", "contribs", "data", "raw"))
```

Search results are categorized across eight entity types for every year from
1998 to 2020.

* PCC = Candidates & Office Holders
* CORP = Corporation
* ELECT = Electioneering
* INDEXP = Independent Expenditures
* LABOR = Labor Organizations
* PAC = Political Action Committee
* PIC = Political Issues Committee
* PARTY = Political Party

We can first scrape all the entity types and report years from the search site's
dropdown menus.

```{r search_years}
search_url <- "https://disclosures.utah.gov/Search/AdvancedSearch/"
ut_search <- read_html(search_url)
rpt_years <- ut_search %>% 
  html_elements("#ReportYear option") %>% 
  html_attr("value")
```

```{r year_minmax}
min(rpt_years)
max(rpt_years)
```

```{r search_types}
entity_types <- ut_search %>% 
  html_elements("#EntityType option") %>% 
  html_attr("value") %>% 
  str_subset("\\w")
```

There are `r length(entity_types) * length(rpt_years)` combinations of these
types and years, however many of these combinations have no search results.

```{r search_combn}
length(entity_types) * length(rpt_years)
```

We can make a `for` loop to run through each combination and request data from
the server with `httr::GET()`. For any combination without any results, a HTML
file is returned instead of the CSV file we want; these files can be detected
and deleted, leaving us with only files containing expenditure records.

```{r}
for (yr in rpt_years) {
  cli_h2("Year: {yr}")
  for (et in entity_types) {
    combo_csv <- path(raw_dir, glue("{et}-{yr}.csv"))
    if (!file_exists(combo_csv)) {
      # request the year and type file
      combo_get <- GET(
        url = str_c(search_url, "GenerateReport"),
        write_disk(combo_csv),
        query = list(
          ReportYear = yr,
          EntityType = et
        )
      )
      # delete the saved file if it isn't a CSV
      combo_mime <- headers(combo_get)[["content-type"]]
      if (!str_detect(combo_mime, "application/csv")) {
        file_delete(combo_csv)
        cli_alert_danger("{et} deleted")
      } else {
        cli_alert_success("{et} saved")
      }
    } else {
      cli_alert_success("{et} exists")
    }
  }
  Sys.sleep(runif(1, 1, 2)) 
}
```

We are left with `r length(dir_ls(raw_dir))` CSV files with data.

```{r entity_list}
raw_info <- dir_info(raw_dir, glob = "*.csv")
nrow(raw_info)
sum(raw_info$size)
raw_csv <- raw_info$path
```

## Fix

Most columns in the comma-separated file are enclosed in double-quotation marks
(`"`) to escape any commas _within_ the columns. However, double-quotes within
strings are not escaped with doubling or a backslash. We can use regular
expressions to find quotes not being used to enclose a column and replace them
with a single-quote mark.

```{r}
tmp_dir <- dir_create(path(dirname(raw_dir), "tmp"))
tmp_csv <- path(tmp_dir, basename(raw_csv))
for (i in seq_along(raw_csv)) {
  if (!file_exists(tmp_csv[i])) {
    y <- read_lines(raw_csv[i])
    # quote not after comma and before non-quote/comma
    y <- str_replace_all(y, '(?<!,|^)\"(?=[^\",])', "'")
    y <- str_replace_all(y, '(?<=[^\",])\"(?!,|$)', "'")
    y <- str_replace(y, '",""","', '","","')
    write_lines(y, file = tmp_csv[i])
  }
}
```

## Read

The fixed files can be read together into a list of tables.

```{r raw_read}
utc <- map(
  tmp_csv,
  read_delim,
  delim = ",",
  na = c("", "NA", "na", "N/A"),
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    TRAN_ID = col_integer(),
    TRAN_DATE = col_date("%m/%d/%Y"),
    TRAN_AMT = col_double()
  )
)
```

The second column in each file contains the name of the entity filing the report
on which the contribution was reported. The name of that second column is thus
different depending on the entity type of that particular file. We can rename
the second column in every file to `ENTITY_NAME`.

```{r}
utc <- map(utc, rename, ENTITY_NAME = 2)
```

Now every table has the same names and can be combined into a single table.

```{r}
names(utc) <- basename(tmp_csv)
utc <- bind_rows(utc, .id = "FILE_NAME")
```

We will identify the source of each row by the entity type and report year from
the file name.

```{r entity_join}
file_entity <- extract(
  data = tibble(FILE_NAME = basename(tmp_csv)),
  col = FILE_NAME,
  into = c("ENTITY_TYPE", "REPORT_YEAR"),
  regex = "(\\w+)-(\\d{4}).csv",
  remove = FALSE,
  convert = TRUE
)
```

```{r raw_bind}
utc <- utc %>% 
  left_join(file_entity, by = "FILE_NAME") %>% 
  select(-FILE_NAME)
```

This data includes both `Contribution` and `Expenditure` transaction types. We
can filter the data to only include the former.

```{r raw_filter}
utc <- filter(utc, TRAN_TYPE == "Contribution")
```

```{r raw_rename}
utc <- clean_names(utc, case = "snake")
```

Finally, three columns use an `X` to indicate a `TRUE` value and blank `NA` to
indicate false.

```{r binary_view}
utc %>% 
  select(filed, inkind, loan) %>% 
  distinct()
```

We can convert these quasi-logical columns to their proper data type for 
mathematical analysis.

```{r binary_convert}
utc <- utc %>% 
  mutate(across(c(filed, inkind, loan), ~!is.na(.)))
```

## Explore

There are `r comma(nrow(utc))` rows of `r ncol(utc)` columns. Each record
represents a single contribution made from an entity to a political campaign
or committee.

```{r glimpse}
glimpse(utc)
tail(utc)
```

We can see that despite data existing since 1998, contributions aren't really
reported in bulk until 2008.

```{r echo=FALSE}
row_count <- utc %>% 
  count(entity_type, report_year, sort = TRUE)

et_levels <- row_count %>% 
  group_by(entity_type) %>% 
  summarise(n = sum(n)) %>% 
  arrange(desc(n)) %>% 
  pull(entity_type)

row_count$entity_type <- factor(
  x = row_count$entity_type,
  levels = et_levels
)

row_count %>% 
  ggplot(aes(x = report_year, y = n)) +
  geom_col(aes(fill = entity_type), position = position_stack(reverse = TRUE)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1, byrow = TRUE))+
  labs(
    title = "Utah Contributions by Entity Type and Report Year",
    subtitle = "Combonation determines in which file contribution obtained",
    x = "Report Year",
    y = "Count"
  )
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(utc, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("tran_date", "name", "tran_amt", "entity_name")
utc <- flag_na(utc, all_of(key_vars))
sum(utc$na_flag)
```

Of the four key variables, only a few are missing the `name` of the individual
making the contribution.

```{r na_view}
utc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
utc <- flag_dupes(utc, -tran_id)
sum(utc$dupe_flag)
```

Despite unique `tran_id` values, there are `r comma(sum(utc$dupe_flag))` fully
duplicated values (same date, amount, names, etc.).

```{r dupe_view}
utc %>% 
  filter(dupe_flag) %>% 
  select(tran_id, all_of(key_vars)) %>% 
  arrange(tran_date)
```

### Categorical

```{r distinct_count}
col_stats(utc, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(utc, report) + scale_x_wrap()
explore_plot(utc, entity_type)
```

### Amounts

```{r amount_summary}
summary(utc$tran_amt)
mean(utc$tran_amt <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(utc[c(which.max(utc$tran_amt), which.min(utc$tran_amt)), ])
```

```{r hist_amount, echo=FALSE}
utc %>%
  ggplot(aes(tran_amt)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Utah Contributions Amount Distribution",
    caption = "Source: UT SOS",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `tran_date` with `lubridate::year()`

```{r date_year}
utc <- mutate(utc, tran_year = year(tran_date))
```

```{r date_range}
min(utc$tran_date)
sum(utc$tran_year < 2000)
max(utc$tran_date)
sum(utc$tran_date > today())
```

```{r bar_year, echo=FALSE}
utc %>% 
  count(tran_year) %>% 
  mutate(even = is_even(tran_year)) %>% 
  ggplot(aes(x = tran_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Utah Contributions by Year",
    caption = "Source: UT SOS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
norm_addr <- utc %>% 
  distinct(address1, address2) %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view, echo=FALSE}
norm_addr
```

```{r address_rejoin}
utc <- left_join(utc, norm_addr, by = c("address1", "address2"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
utc <- utc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  utc$zip,
  utc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
utc <- utc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
utc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  utc$state,
  utc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- utc %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("UT", "DC", "UTAH"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city_rejoin}
utc <- left_join(
  x = utc,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- utc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
utc <- utc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

Some misspellings and abbreviations remain than can be manually fixed.

```{r many_city}
many_city <- c(valid_city, extra_city)
utc %>% 
  count(city_refine, sort = TRUE) %>% 
  filter(city_refine %out% many_city)
```

```{r city_manual}
utc <- utc %>% 
  mutate(
    city_refine = city_refine %>% 
      str_replace("^SALT LAKE CIT$", "SALT LAKE CITY") %>% 
      str_replace("^SALT LAKE CTY$", "SALT LAKE CITY") %>% 
      str_replace("^SLC$", "SALT LAKE CITY") %>% 
      str_replace("^WEST VALLEY C$", "WEST VALLEY CITY") %>% 
      str_replace("^WVC$", "WEST VALLEY CITY")
  )
```

Some other common _valid_ city names need to be recognized before we check our
progress.

```{r}
many_city <- c(many_city, "NIBLEY", "ST GEORGE")
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(utc$city),
  utc$city_norm,
  utc$city_swap,
  utc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>%
  mutate(across(1, md_code)) %>% 
  kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(utc$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Utah City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Utah City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
utc <- utc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(utc, 50))
```

1. There are `r comma(nrow(utc))` records in the database.
1. There are `r comma(sum(utc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `tran_amt` and `tran_date` seem reasonable.
1. There are `r comma(sum(utc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `tran_year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ut", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ut_contribs_2000-20210827.csv")
write_csv(utc, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
