---
title: "Wisconsin Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wi", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # print markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is from the Wisconsin Secretary of State's Campaign Finance System (CIFS).

> Wyoming's Campaign Finance Information System (WYCFIS) exists to provide a
mechanism for online filing of campaign finance information and to provide full
disclosure to the public. This website contains detailed financial records and
related information that candidates, committees, organizations and parties are
required by law to disclose.

## Import

Using the CIFS [contribution search portal][search], we can run a search for all
contributions from "All Filing Periods" from the dates 2000-01-01 to 
`r today()`. Those search results need to be manually exported as the
`ReceiptsList.csv` file.

> To view contributions to a committee, go to the CFIS Home Page, on the left
hand side, and click View Receipts. A pop up letting you know that this
information cannot be used for solicitation purposes will appear – click
Continue. Type in a committee’s ID in the field titled ID, or start typing the
name of the candidate in the Registrant Name field and click on the correct
committee name when it appears. Type in any additional information you would
like to search for, including a name of a contributor or amount of contribution.
To view all contributions, remove the filing period by clicking, in the Filing
Period Name field, and scroll all the way to the top and select All Filing
Periods. Click Search and all of the contributions fitting your search criteria
will appear. If you would like to export these into Excel, scroll all the way to
the bottom and on the right hand side, click the XLS icon.

[search]: https://cfis.wi.gov/Public/Registration.aspx?page=ReceiptList#

Infuriatingly, the site only lets users export 65,000 records at a time. We
have manually exported 91 files.

```{r raw_dir}
raw_dir <- dir_create(here("wi", "contribs", "data", "raw"))
raw_files <- as_tibble(dir_info(raw_dir))
sum(raw_files$size)
raw_files %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```

The files can be read into a single data frame with `vroom::vroom()`.

```{r raw_read}
wic <- vroom(
  file = raw_files$path,
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(.default = "c")
)
```

We can check the number of rows against the total reported by our empty search.
We can also count the number of distinct values from a discrete column.

```{r raw_check}
nrow(wic) == 5866891 # check col count
count(wic, ContributorType) # check distinct col
prop_na(wic$`72 Hr. Reports`) # empty column
```

The file appears to have been read correctly. We just need to parse, rename, and
remove some of the columns.

```{r raw_names}
raw_names <- names(wic)[c(-19, -21)]
```

```{r raw_parse}
wic <- wic %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  select(-x72_hr_reports) %>% 
  mutate(across(transaction_date, mdy)) %>% 
  mutate(across(contribution_amount, parse_double)) %>% 
  mutate(across(segregated_fund_flag, parse_logical)) %>% 
  rename(
    date = transaction_date,
    period = filing_period_name,
    con_name = contributor_name,
    amount = contribution_amount,
    addr1 = address_line1,
    addr2 = address_line2,
    state = state_code,
    emp_name = employer_name,
    emp_addr = employer_address,
    con_type = contributor_type,
    rec_name = receiving_committee_name,
    seg_fund = segregated_fund_flag
  )
```

## Explore

```{r glimpse}
glimpse(wic)
tail(wic)
```

### Missing

Very few important records are missing a value.

```{r na_count}
col_stats(wic, count_na)
```

We can flag these few records with `campfin::flag_na()`.

```{r na_flag}
wic <- wic %>% flag_na(date, con_name, amount, rec_name)
percent(mean(wic$na_flag), 0.0001)
```

```{r na_view}
wic %>% 
  filter(na_flag) %>% 
  select(date, con_name, amount, rec_name)
```

### Duplicates

Quite a few more records are duplicated. While it's possible for the same person
to make a contribution of the same amount on the same day, we will still flag
these records with `campfin::flag_dupes()`.

```{r dupe_flag}
wic <- flag_dupes(wic, everything())
percent(mean(wic$dupe_flag), 0.01)
```

```{r dupe_view}
wic %>% 
  filter(dupe_flag) %>% 
  select(date, con_name, amount, rec_name)
```

### Categorical

```{r n_distinct}
col_stats(wic, n_distinct)
```

```{r n_count}
count(wic, seg_fund, sort = TRUE)
count(wic, con_type, sort = TRUE)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(wic$amount)
mean(wic$amount <= 0)
```

```{r hist_amount, echo=FALSE}
wic %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Wisconsin Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can use `lubridate::year()` to add a calendar year from the date.

```{r date_year}
wic <- mutate(wic, year = year(date))
```

The `date` and new `year` columns are very clean.

```{r date_range}
min(wic$date)
sum(wic$year < 2000)
max(wic$date)
sum(wic$date > today())
```

```{r bar_year, echo=FALSE}
wic %>%
  count(year) %>% 
  filter(year >= 2008) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2008, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Wisconsin Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
wic <- wic %>% 
  unite(
    starts_with("addr"),
    col = addr_full,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address_view}
wic %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
wic <- wic %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  wic$zip,
  wic$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
wic <- wic %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
wic %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  wic$state,
  wic$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
wic <- mutate(
  .data = wic,
  city_norm = normal_city(
    city = city, 
    abbs = usps_city,
    states = c("WI", "DC", "WISCONSIN"),
    na = invalid_city,
    na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
wic <- wic %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    # check for abb or small diff
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      # if nan and either condition
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- wic %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
nrow(good_refine)
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
wic <- wic %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

There are one or two more very common values we can adjust by hand.

```{r city_fix}
wic <- wic %>% 
  mutate(
    city_refine = city_refine %>% 
      na_if("WORK AT HOME") %>% 
      na_if("WI") %>% 
      str_replace("^NYC$", "NEW YORK") %>% 
      str_replace("TRIANGLE PA$", "TRIANGLE PARK")
  )
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city, "RESEARCH TRIANGLE PARK")
progress <- progress_table(
  str_to_upper(wic$city_raw),
  wic$city_norm,
  wic$city_swap,
  wic$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(wic$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Wisconsin City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Wisconsin City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Export

```{r clean_select}
wic <- wic %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename(city = city_raw)
```

```{r clean_glimpse}
glimpse(sample_n(wic, 20))
```

1. There are `r comma(nrow(wic))` records in the database.
1. There are `r comma(sum(wic$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wic$na_flag))` records missing a key variable.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

```{r clean_dir}
clean_dir <- dir_create(here("wi", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "wi_contribs_clean.csv")
write_csv(wic, clean_path, na = "")
file_size(clean_path)
guess_encoding(clean_path)
```

## Dictionary

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wic)),
  # raw = c(md_code(raw_names), rep("", 7)),
  type = md_code(map_chr(wic, rlang::as_label)),
  def = c(
    "Date contribution was made",
    "Election during which contribution was made",
    "Contributor full name",
    "Contribution amount or correction",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor 2-digit state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Contributor employer address",
    "Contributor type",
    "Recipient committee name",
    "Recipient ethics & campaign finance ID",
    "Contribution condiut (method)",
    "Recipient election office sought",
    "Comment (typically check date)",
    "PAC segregated fund sourced",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Wisconsin Contributions Data Dictionary\n", dict_md),
  path = here("wi", "contribs", "wi_contribs_dict.md"),
)
```

