---
title: "Kansas Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ks", "lobby", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # scrape html pages
  glue, # combine strings
  here, # relative storage
  httr, # http queries
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Kansas Secretary of State's Office][sos].

[sos]: https://sos.kansas.gov/

As described on the SOS website:

> Lobbyists are required to register with the Secretary of State’s office if
they meet the qualifications outlined in K.S.A. 46-222. Individuals who meet one
of the following criteria in a calendar year must register as a lobbyist:
>
> *  receives compensation to lobby;
> * serves as the primary representative of an organization, individual or other
> entity to lobby, regardless of whether compensation is received; or
> * expends a total of $100 or more for lobbying activities in a calendar year.
> 
> For more information on lobbying activities in Kansas, please contact the
Kansas Governmental Ethics Commission. For assistance with the Kansas Lobbyist
Center, please contact the Elections Division at 785-296-4561.

## Import

We can use `httr::GET()` to query the [SOS lobbyist directory][lob] and scrape
then scrape the HTML table that's returned with `rves::html_table()`.

[lob]: http://www.sos.ks.gov/elections/elections_lobbyists.html

```{r raw_get}
url <- "http://www.kssos.org/elections/lobbyist_directory_display.aspx?"
response <- GET(url, query = list(SearchBy = "Lobbyist", LobbyYear = "2019"))
kslr <- content(response) %>% 
  html_node("table") %>% 
  html_table(fill = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names)
```

Or, we can download the file manually and read it directly...

```{r raw_dir}
raw_dir <- here("ks", "lobby", "data", "raw")
dir_create(raw_dir)
```

```{r raw_read}
kslr <- dir_ls(raw_dir, glob = "*.html$") %>% 
  read_html() %>% 
  html_node("table") %>% 
  html_table(fill = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>% 
  mutate_all(str_squish) %>% 
  na_if("") %>% 
  select(-starts_with("x")) %>% 
  mutate(client = "") %>% 
  mutate_at(vars(registration_date), parse_date, "%m/%d/%Y")
```

As you can see, the clients of each lobbyist are not listed as a separate
column, but are instead listed as in the first `name` _under_ each lobbyist.

```{r raw_view}
kslr
```

Looping from the bottom to top, we can check every row and attempt to
concatinate each client name into a new client column.

```{r raw_fix}
for (i in nrow(kslr):1) {
  if (is.na(kslr$name[i])) {
    next
  } else {
    if (str_sub(kslr$name[i], end = 1) == "*") {
      kslr$client[i-1] <- str_c(kslr$name[i], kslr$client[i], collapse = "#")
      kslr[i, 1:7] <- NA
    }
  }
}
```

Then, we can split this new column into a list-column and use
`tidyr::pivot_longer()` to create a new row for every client with the lobbyist
repeated in each row.

```{r raw_unnest}
kslr <- kslr %>% 
  drop_na(name) %>% 
  mutate(client = client %>% str_remove("\\*\\s") %>% str_split("\\*\\s")) %>% 
  unnest_longer(client) %>% 
  distinct()
```

## Wrangle

The single `city_state_zip` needs to be separated into three columns with
`tidyr::separate()`.

```{r geo_split}
kslr <- kslr %>% 
  separate(
    col = city_state_zip,
    into = c("city_sep", "state_zip"),
    sep = ",\\s",
    remove = TRUE
  ) %>% 
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s",
    remove = TRUE
  )
```

From these three new columns, we can see almost all rows already contain
valid values.

```{r geo_check_pre}
prop_in(kslr$city_sep, c(valid_city, extra_city))
prop_in(kslr$state_sep, valid_state)
prop_in(kslr$zip_sep, valid_zip)
```

```{r geo_normal}
kslr <- kslr %>%
  rename(address = addr_1_addr_2) %>% 
  mutate_at(vars(address), normal_address, abbs = usps_street) %>% 
  mutate_at(vars(city_sep), normal_city, abbs = usps_city) %>% 
  mutate_at(vars(state_sep), normal_state, na_rep = TRUE) %>% 
  mutate_at(vars(zip_sep), normal_zip, na_rep = TRUE)
```

```{r geo_check_post}
prop_in(kslr$city_sep, c(valid_city, extra_city))
prop_in(kslr$state_sep, valid_state)
prop_in(kslr$zip_sep, valid_zip)
```

## Explore

```{r glimpse}
head(kslr)
tail(kslr)
glimpse(sample_frac(kslr))
```

```{r}
col_stats(kslr, count_na)
```

## Export

```{r}
proc_dir <- here("ks", "lobby", "data", "processed")
dir_create(proc_dir)
write_csv(
  x = kslr,
  path = glue("{proc_dir}/ks_lobbyists.csv"),
  na = ""
)
```

