---
title: "North Carolina Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nc", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
data set row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # print markdown
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http requests
  fs # search storage 
)
```

```{r conflicts, include=FALSE}
filter <- dplyr::filter
here <- here::here
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

North Carolina contributions data is obtained from the [North Carolina State
Board of Elections (NCSBE)][ncsbe]

> The State Board of Elections (State Board) is the state agency charged with
the administration of the elections process and campaign finance disclosure and
compliance.

Each record reflects a single contribution (receipt) made to a candidate or
committee.

[ncsbe]: https://www.ncsbe.gov/index.html

## Download

To download the data, we must run searches on the [NCSBE search portal][search].
Searching for all records at once returns an error, but we can ask for all
contributions from a given calendar year and export annual files.

[search]: https://cf.ncsbe.gov/CFTxnLkup/AdvancedSearch/

```{r raw_export, eval=FALSE}
# does not work
date_from <- "01/01/2000"
param_file <- here("nc", "contribs", "ncsbe_param.txt")
raw_param <- glue::glue(readr::read_lines(param_file))
raw_url <- "https://cf.ncsbe.gov/CFTxnLkup/ExportResults/"
raw_export <- paste(raw_url, raw_param, sep = "?")
for (y in 2000:2020) {
  date_to <- sprintf("12/31%s", y)
  raw_path <- path(raw_dir, sprintf("transinq_results-%s.csv", y))
  POST(raw_export, write_disk(raw_path, overwrite = TRUE))
  message(y)
}
```

```{r raw_dir}
raw_dir <- dir_create(here("nc", "contribs", "data", "raw"))
raw_files <- as_tibble(dir_info(raw_dir))
raw_files %>% 
  select(path, size, modification_time) %>% 
  mutate_at(vars(path), basename)
```

## Read

The annual files can be read into a single data frame with `vroom::vroom().`

```{r raw_read}
ncc <- vroom(
  file = raw_files$path,
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  na = c("", "NA", "Not Available"),
  col_types = cols(
    .default = col_character(),
    `Date Occured` = col_date_usa(),
    `Amount` = col_double()
  )
)
```

We can check the number of distinct values in a discrete variables to ensure the
file was properly read into memory.

```{r raw_check}
count(ncc, `Form of Payment`)
```

For conveyance we will rename the variables.

```{r raw_names}
raw_names <- names(ncc)
ncc <- ncc %>% 
  rename(
    con_name = `Name`,
    con_addr1 = `Street Line 1`,
    con_addr2 = `Street Line 2`,
    con_city = `City`,
    con_state = `State`,
    con_zip = `Zip Code`,
    con_job = `Profession/Job Title`,
    con_emp = `Employer's Name/Specific Field`,
    con_type = `Transction Type`,
    rec_name = `Committee Name`,
    rec_id = `Committee SBoE ID`,
    rec_addr1 = `Committee Street 1`,
    rec_addr2 = `Committee Street 2`,
    rec_city = `Committee City`,
    rec_state = `Committee State`,
    rec_zip = `Committee Zip Code`,
    report = `Report Name`,
    date = `Date Occured`,
    account = `Account Code`,
    method = `Form of Payment`,
    candidate = `Candidate/Referendum Name`
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

## Explore

Since the empty searches returned the yearly contributions in alphabetical
order, the top of each file has empty name records. We will sort this single
data frame chronologically.

```{r raw_arrange}
ncc <- arrange(ncc, date)
```

```{r glimpse}
glimpse(ncc)
tail(ncc)
```

### Missing

The first thing we notice are the high number of missing records.

```{r na_count}
col_stats(ncc, count_na)
```

This is likely due to the lack of individual information for those contributions
reported as aggrigates.

```{r na_view}
ncc %>% 
  filter(is.na(con_state)) %>% 
  select(con_name, con_state) %>% 
  count(con_name, sort = TRUE) %>% 
  mutate(p = n/sum(n))
```

We can flag any kind of non-aggrigate record missing a name or number.

```{r na_flag}
ncc <- ncc %>% flag_na(date, con_name, amount, rec_name)
percent(mean(ncc$na_flag), 0.001)
```

### Duplicates

We can also flag any entirely duplicate rows. To keep memory usage low with such
a large data frame, we will split our data into a list and check each element of
the list. For each chunk, we will write the duplicate row positions to a local
file.

```{r dupe_write}
dupe_file <- here("nc", "contribs", "dupes.txt")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  n <- 1e5 # rows per chunk
  nr <- nrow(ncc)
  # split file into chunks
  ncs <- split(ncc, rep(1:ceiling(nr/n), each = n, length.out = nr))
  for (i in seq_along(ncs)) {
    # check dupes from both ends
    d1 <- duplicated(ncs[[i]], fromLast = FALSE)
    d2 <- duplicated(ncs[[i]], fromLast = TRUE)
    # append to disk
    write_lines(d1 | d2, dupe_file, append = TRUE)
    rm(d1, d2)
    ncs[[i]] <- NA
    message(percent(i/length(ncs)))
    Sys.sleep(10)
    gc(reset = TRUE, full = TRUE)
  }
  rm(ncs)
}
```

Then we can read this file as a logical vector and add it as a variable.

```{r dupe_add}
dupe_vec <- as.logical(read_lines(dupe_file))
ncc <- mutate(ncc, dupe_flag = dupe_vec)
rm(dupe_vec)
gc(reset = TRUE, full = TRUE)
```

A _huge_ percentage of the overall records in this database are duplicates.
This stems from the use of aggregated records reported millions of timnes by a
handful of recipients.

```{r dupe_view}
percent(mean(ncc$dupe_flag))
ncc %>% 
  filter(dupe_flag) %>% 
  count(date, con_name, amount, rec_name, rec_id, report, sort = TRUE)
```

### Categorical

```{r n_distinct}
col_stats(ncc, n_distinct)
```

```{r distinct_bars, echo=FALSE}
explore_plot(ncc, con_type)
explore_plot(ncc, method)
explore_plot(ncc, declaration)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(ncc$amount)
mean(ncc$amount <= 0, na.rm = TRUE)
mean(ncc$amount <= 5, na.rm = TRUE) # itemized
```

```{r hist_amount, echo=FALSE}
ncc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "North Carolina Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can add the 4-digit calendar year from the `date` with `lubridate::year()`.

```{r date_year}
ncc <- mutate(ncc, year = year(date))
```

Aside from a handful of missing values, the range of dates is clean.

```{r date_range}
prop_na(ncc$date)
min(ncc$date, na.rm = TRUE)
sum(ncc$year < 2000, na.rm = TRUE)
max(ncc$date, na.rm = TRUE)
sum(ncc$date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
ncc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_unite}
ncc <- ncc %>% 
  unite(
    col = con_addr_norm,
    starts_with("con_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  unite(
    col = rec_addr_norm,
    starts_with("rec_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  )
```

```{r address_norm}
ncc <- mutate_at(
  .tbl = ncc,
  .vars = vars(ends_with("addr_norm")),
  .funs = normal_address,
  abbs = usps_street,
  na = invalid_city,
  na_rep = TRUE
)
```

```{r address_move}
ncc <- relocate(ncc, ends_with("addr_norm"), .after = last_col())
```

```{r address_view}
ncc %>% 
  select(starts_with("con_addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
ncc <- mutate_at(
  .tbl = ncc,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_progress}
progress_table(
  ncc$con_zip,
  ncc$con_zip_norm,
  ncc$rec_zip,
  ncc$rec_zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
ncc <- mutate_at(
  .tbl = ncc,
  .vars = vars(ends_with("state")),
  .funs = list(norm = normal_state),
  valid = valid_state,
  na_rep = TRUE
)
```

```{r state_view}
ncc %>% 
  filter(con_state != con_state_norm) %>% 
  count(con_state, con_state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  ncc$con_state,
  ncc$con_state_norm,
  ncc$rec_state,
  ncc$rec_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ncc <- mutate_at(
  .tbl = ncc,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  states = c("NC", "DC", "NORTH CAROLINA"),
  na = invalid_city,
  na_rep = TRUE
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ncc <- ncc %>% 
  left_join(
    y = zipcodes,
    by = c(
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(con_city_norm, city_match),
    match_dist = str_dist(con_city_norm, city_match),
    con_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = con_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- ncc %>% 
  mutate(
    con_city_refine = con_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(con_city_refine != con_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "con_city_refine" = "city",
      "con_state_norm" = "state",
      "con_zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
nrow(good_refine)
good_refine %>%
  count(
    con_state_norm, 
    con_zip_norm, 
    con_city_swap, 
    con_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
ncc <- ncc %>% 
  left_join(good_refine) %>% 
  mutate(con_city_refine = coalesce(con_city_refine, con_city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ncc$con_city),
  ncc$con_city_norm,
  ncc$con_city_swap,
  ncc$con_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ncc$con_city, many_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "North Carolina City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Export

```{r clean_select}
ncc <- ncc %>% 
  select(-con_city_norm, -con_city_swap) %>% 
  rename_all(~str_replace(., "_(norm|refine|swap)", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(ncc, 20))
```

1. There are `r comma(nrow(ncc))` records in the database.
1. There are `r comma(sum(ncc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ncc$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

```{r clean_dir}
clean_dir <- dir_create(here("nc", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "nc_contribs_clean.csv")
write_csv(ncc, clean_path, na = "")
file_size(clean_path)
guess_encoding(clean_path)
```

## Dictionary

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ncc)),
  # raw = c(md_code(raw_names), rep("", 7)),
  type = md_code(map_chr(ncc, rlang::as_label)),
  def = c(
    "Contributor full name",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor 2-digit state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Contributor type",
    "Recipient committee name",
    "Recipient unique ID",
    "Recipient street address",
    "Recipient secondary address",
    "Recipient city name",
    "Recipient 2-digit state abbreviation",
    "Recipient ZIP+4 code",
    "Election contribution reported for" ,
    "Date contribution was made",
    "Contribution amount or correction",
    "Contribution method",
    "Contribution purpose",
    "Recipient candidate or referendum",
    "Support or oppose declaration",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized contributor street address",
    "Normalized recipient street address",
    "Normalized contributor 5-digit ZIP code",
    "Normalized recipient 5-digit ZIP code",
    "Normalized contributor 2-digit state abbreviation",
    "Normalized recipient 2-digit state abbreviation",
    "Normalized recipient city name",
    "Normalized contributor city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# North Carolina Contributions Data Dictionary\n", dict_md),
  path = here("nc", "contribs", "wi_contribs_dict.md"),
)
```
