---
title: "Alaksa Lobbyists"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ak", "lobbying", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http queries
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Alaska Public Offices Commission][apoc] (APOC).

> #### Mission
> To encourage the public's confidence in their elected and appointed officials
by administering Alaska's disclosure statutes and publishing financial
information regarding the activities of election campaigns, public officials,
lobbyists and lobbyist employers.

On the APOC home page, we can see that lobbying disclosure data can be searched.

> #### [Lobbying Disclosure](https://aws.state.ak.us/ApocReports/Lobbying/)
> Search reports from Lobbyists and Employers of Lobbyists.

[apoc]: http://doa.alaska.gov/apoc/home.html

## Import

From the APOC website, we can query the database for any year and download the
data locally. 

```{r raw_dir}
raw_dir <- here("state","ak", "lobby", "data", "raw")
dir_create(raw_dir)
```

We could also attemp this with `httr::GET()` (but not now).

```{r raw_get, eval=FALSE}
response <- GET(
  url = "https://aws.state.ak.us/ApocReports/Lobbying/LORForms.aspx", 
  query = list(
    exportAll = "True", 
    exportFormat = "CSV", 
    isExport = "True"
  )
)
aklr <- content(response)
```

```{r list_files}
raw_file <- dir_ls(raw_dir)
length(raw_file) == 1
```

The raw file can be read using `vroom::vroom()`.

```{r raw_read}
aklr <- vroom(
  file = raw_file,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  guess_max = 0,
  num_threads = 1,
  col_types = cols(
    .default = col_character(),
    Result = col_double(),
    `Report Year` = col_integer(),
    Submitted = col_date_mdy(),
    `Other Services Performed` = col_logical(),
    `Administrative Lobbying` = col_logical(),
    `Legislative Lobbying` = col_logical(),
    `Start Date` = col_date_mdy(),
    `Compensation Amount` = col_double(),
    `Reimbursement Of Expenses` = col_logical(),
    `Other Compensation` = col_logical(),
    `Not Qualified As Lobbyist` = col_logical(),
    `Date Qualified As Lobbyist` = col_date_mdy()
  )
)
```

We know the file properly read as the number of rows is equal to the number of
distinct `report` values, a unique row number variable.

```{r raw_count}
n_distinct(aklr$result) == nrow(aklr)
```

```{r raw_repair}
aklr <- aklr %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  remove_empty("rows") %>%
  filter(!is.na(result))
```

```{r raw_count2}
n_distinct(aklr$result) == nrow(aklr)
```

## Explore

```{r glimpse}
head(aklr)
tail(aklr)
glimpse(sample_frac(aklr))
```

### Missing

_Very_ few records are missing some of the variables we need to identify a
lobbyist.

```{r na_count}
col_stats(aklr, count_na)
```

We can flag those rows with `campfin::na_flag()`.

```{r na_flag}
aklr <- aklr %>% 
  flag_na(last_name, address, employer_contact_last_name, start_date)
sum(aklr$na_flag)
```

### Duplicate

There are no duplicate records that need to be flagged.

```{r dupe_count, warning=TRUE}
aklr <- flag_dupes(aklr, -result)
```

### Categorical

```{r cat_distinct}
col_stats(aklr, n_distinct)
```

```{r plot_comp_name, echo=FALSE}
explore_plot(aklr, compensation_name) + coord_flip()
```

```{r plot_reimburse, echo=FALSE}
explore_plot(aklr, reimbursement_of_expenses)
```

### Continuous

### Amounts

```{r plot_comp_amount, echo=FALSE, fig.height=9}
aklr %>% 
  filter(compensation_name %in% most_common(aklr$compensation_name, 4)) %>% 
  ggplot(aes(compensation_amount)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  scale_x_continuous(labels = dollar, trans = "log10") + 
  scale_y_continuous(labels = comma) +
  facet_wrap(~compensation_name, scales = "free", ncol = 1)
```

### Dates

```{r year_add}
aklr <- mutate(aklr, start_year = year(start_date))
```

```{r date_range}
min(aklr$start_date, na.rm = TRUE)
max(aklr$start_date, na.rm = TRUE)
```

```{r plot_year_count, echo=FALSE}
aklr %>%
  count(start_year) %>% 
  mutate(start_year = as.character(start_year)) %>% 
  ggplot(aes(start_year, n)) +
  geom_col(fill = dark2["purple"]) +
  scale_fill_discrete(guide = FALSE)
```

## Wrangle

### Address

```{r address_norm}
aklr <- aklr %>% 
  mutate_at(
    .vars = vars(ends_with("address")),
    .funs = list(norm = normal_address),
    abbs = usps_street,
    na_rep = TRUE
  )
```

### ZIP

```{r zip_norm}
aklr <- aklr %>% 
  mutate_at(
    .vars = vars(ends_with("zip")),
    .funs = list(norm = normal_zip),
    na_rep = TRUE
  )
```

```{r zip_progress}
progress_table(
  aklr$zip,
  aklr$zip_norm,
  compare = valid_zip
)
progress_table(
  aklr$employer_contact_zip,
  aklr$employer_contact_zip_norm,
  compare = valid_zip
)
```

### State

```{r state_norm}
aklr <- aklr %>% 
  mutate_at(
    .vars = vars(ends_with("state_region")),
    .funs = list(norm = normal_state),
    abbreviate = TRUE,
    na_rep = TRUE,
    valid = NULL
  )
```

```{r state_progress}
progress_table(
  aklr$state_region,
  aklr$state_region_norm,
  compare = valid_state
)
progress_table(
  aklr$employer_contact_state_region,
  aklr$employer_contact_state_region_norm,
  compare = valid_state
)
```

```{r state_view_post}
count(aklr, state_region, state_region_norm, sort = TRUE)
```

```{r state_alberta}
aklr <- mutate(
  .data = aklr,
  state_region_norm = state_region_norm %>% 
    str_replace("^ALBERTA$", "AB")
)
```

### City

```{r city_norm}
aklr <- aklr %>% 
  mutate_at(
    .vars = vars(ends_with("city")),
    .funs = list(norm = normal_city),
    abbs = usps_city,
    states = c("AK", "DC", "ALASKA"),
    na = invalid_city,
    na_rep = TRUE
  )
```

```{r city_swap}
aklr <- aklr %>%
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_region_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(city_match) & (match_abb | match_dist == 1),
      true = city_match, 
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_abb,
    -match_dist
  )
```

```{r city_progress}
progress_table(
  str_to_upper(aklr$city_raw),
  aklr$city_norm,
  aklr$city_swap,
  compare = c(valid_city, extra_city)
)
```

```{r employer_city_swap}
aklr <- aklr %>%
  left_join(
    y = zipcodes,
    by = c(
      "employer_contact_state_region_norm" = "state",
      "employer_contact_zip_norm" = "zip"
    )
  ) %>% 
  rename(employer_city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(
      abb = employer_contact_city_norm, 
      full = employer_city_match
    ),
    match_dist = str_dist(
      a = employer_contact_city_norm, 
      b = employer_city_match
    ),
    employer_contact_city_swap = if_else(
      condition = !is.na(employer_city_match) & (match_abb | match_dist == 1),
      true = employer_city_match, 
      false = employer_contact_city_norm
    )
  ) %>% 
  select(
    -employer_city_match,
    -match_abb,
    -match_dist
  )
```

```{r employer_city_progress}
progress_table(
  str_to_upper(aklr$employer_contact_city),
  aklr$employer_contact_city_norm,
  aklr$employer_contact_city_swap,
  compare = c(valid_city, extra_city)
)
```

## Export

```{r proc_dir}
clean_dir <- dir_create(here("state","ak", "lobby", "data", "clean"))
```

```{r proc_export}
aklr %>% 
  select(
    -ends_with("city_norm"),
    -contains("match")
  ) %>% 
  write_csv(
    path = path(clean_dir, "ak_lobb_reg.csv"),
    na = ""
  )
```

