---
title: "Nebraska Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("kiernann/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # text analysis
  tidytext, # text analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

## Data

Data is obtained by from the Nebraska Accountability and Disclosure Commission's (NADC) 
[Open Data portal][03]. From this portal, we will download the [Campaign Statements Data][04].

> A weekly export of the campaign filings based upon the paper records filed with the Nebraska
Accountability and Disclosure Commission.

The [NADC home][04] page explains the role and resources provided by the Commission.

> This site includes the campaign statements filed with the Commission by candidates for state and
local offices, political action committees or PACs, political parties, and ballot question
committees. Campaign statements disclose the amount and sources of contributions to the committee
and the amount purposes of expenditures by the committee. The site also includes Reports of
Contributions filed with the Commission by corporations, unions, limited liability companies,
limited partnerships, and by industry, trade or professional associations. Reports of Contributions
disclose contributions and expenditures made to support or oppose candidates or ballot questions.
Forms, instructional materials and a list of committees may be accessed through this site. To
access the information described in this paragraph, click on the tab above marked CAMPAIGN FINANCE.

Navigating to the [campaign finance section of the website][05], the exact filing requirements are
layed out.

> Candidates for state and local office are subject to the campaign finance provisions of the
Nebraska Political Accountability and Disclosure Act (NPADA). Those seeking election to federal
office are not subject to the NPADA. Once a candidate raises, receives or expends $5,000 or more in
a calendar year, he or she is required to register the committee with the Commission by filing a
Statement of Organization and paying a filing fee of $100. Once the candidate committee is formed,
the committee must file periodic campaign statements.

specifically, there are **three** kinds of filings that must be made by general candidates:

1. > Statement of Organization - This document registers the Committee with the Commission. It
discloses the name of the committee, the name of the candidate, the office sought and the name of
the treasurer.

2. > Campaign Statement - this document provides a summary of the financial activity of the
committee. It discloses the name of contributors of more than \$250 and the amounts of the
contributions. It discloses expenditures of more than \$250.

3. > Report of Late Contributions - this document discloses contributions of $1,000 or more
received by the committee during 14 days immediately preceding the election. Information reported
as late contributions are also reported on subsequent campaign statements.

Aside from general candidates, the NADC also explains that political committees must also file.

> There are **two** types of committees that are commonly referred to as political action
committees or PACs. One type of PAC is a political committee sponsored by a corporation, union or
other type of organization. Another type of PAC is a political committee which is not sponsored by
another group. PACs are required to register with the Commission and file periodic campaign
statements disclosing receipts and expenditures.

1. > State and county political party committees with certain levels of financial activity,
register with the Commission and file periodic campaign statements.

2. > A ballot question committee is a political committee formed for the purpose of supporting or
opposing the qualification, passage or defeat of a ballot question. The committee must register
with the Commission within 10 days after raising, receiving or expending $5,000 or more in a
calendar year. Once registered, ballot questions committees file periodic campaign statements
disclosing receipts and expenditures.

Finally, regarding the specific data in question, the NADC identifies the source:

> The campaign filings displayed on this website are based upon the paper records filed with the
Nebraska Accountability and Disclosure Commission. While every effort has been made to ensure
accurate data input, errors may occur. Anyone noting an error is encouraged to contact this office.

In any discrepancy between the information displayed on this website and the official records on file with the Commission, the official records of the Commission take precedence.

## Import

The campaign statements data is provided as a series of text files organized in a relational
database management system (DRMS).

### Download

The data is provided as a ZIP file, which can be downloaded locally.

```{r raw_dir}
raw_dir <- here("ne", "expends", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw}
zip_url <- "http://www.nebraska.gov/nadc_data/nadc_data.zip"
zip_file <- str_c(raw_dir, basename(zip_url), sep = "/")
download.file(
  url = zip_url,
  destfile = zip_file
)
```

There are `r nrow(unzip(zip_file, list = T))` files contained in a single folder (`/nadc_data`) 
inside the ZIP archive.

```{r list_raw, results='asis', echo=FALSE}
zip_contents <- unzip(zip_file, list = TRUE)$Name
for (file in zip_contents) {
    cat("*", file, "\n")
}
```

We can then unzip the archive to the same `/raw` directory.

```{r unzip_raw}
unzip(
  zipfile = zip_file,
  exdir = raw_dir
)
```

The archive contains a single folder, inside which are the many TXT files we need.

```{r nadc_dir}
nadc_dir <- here("ne", "expends", "data", "raw", "nadc_data")
```

### Read

The first file in the folder, `DATE_UPDATED.TXT` contains the time the files were last updated.

```{r date_updated, collapse=TRUE}
read_lines(file = glue("{nadc_dir}/DATE_UPDATED.TXT")) %>% 
  str_remove_all("[:alpha:]") %>% 
  as_datetime() %>% 
  subtract(now())
```

To work with the relational database managment system, we are going to read every text file into a 
single list comprised of separate data frames by using `purrr::map()` and `readr::read_delim()`.

```{r map_read}
ne <- 
  dir_ls(
    path = nadc_dir,
    type = "file",
    glob = "*.txt$"
  ) %>% 
  map(
    read_delim,
    delim = "|",
    escape_double = FALSE,
    escape_backslash = FALSE,
    col_types = cols(.default = "c")
  ) %>% 
  map(clean_names)

# set names equal to form
names(ne) <- tools::file_path_sans_ext(basename(names(ne)))
```

Before we can prepare the file for indexing, we need to locate the exact data we're interested in.
The content, structure, and relationship of each file is described in the
`/nadc_data/nadc_tables.rtf` rich text file. The files correspond to the _forms_ through which
candidates and committees report their contributions, expenditures, etc. Each type of filter uses a
different form, so all expenditures are split among a few files (corresponding to a few forms).
Each form typically also has a section (schedule) where _expenditures_ are reported.

#### Form B-1

[Form B-1][06] covers Candiate and Ballot Question Committees, where Section 1 of Schedule D asks
the candidate/committee to:

> List all payees who were paid more than $250 during this reporting period. If multiple payments
to the same payee totaled more than $250 throughout this reporting period, those expenditures must
be listed. Reporting period refers to your entry on Page 1 under Item 4. Expenditures to the same
payee over separate reporting periods should not be accumulated. Expenditures to the same payee
must be listed under the same name. If the committee reimburses the candidate or engages the
services of an advertising agency or another agent of the committee for expenses they incurred on
behalf of the committee, list the payments the committee made to the candidate or agent and also
list the payments which were made by the candidate or agent on behalf of the committee. (E.g., If
the candidate makes payments to a newspaper for advertising and is reimbursed by the committee,
report the payments made to the candidate but also list the payments made by the candidate to the
newspaper. Include the name of the newspaper, and the date of each of the expenditures by the
candidate and list the amount only in the “purpose” box along with the description of the
expenditure.)

The form also defines the NADC conception of expenditure:

> Expenditure: A payment, donation, loan, pledge, or promise of payment of money or anything of
ascertainable monetary value for goods, materials, services, or facilities in assistance of, or in
opposition to, the nomination or election of a candidate, or the qualification, passage, or defeat
of a ballot question.  (An offer or tender of an expenditure is not an expenditure if expressly and
unconditionally rejected or returned.)  Expenditure shall include a contribution or a transfer of
anything of ascertainable monetary value for purposes of influencing the nomination or election of
any candidate or the qualification, passage or defeat of a ballot question.

```{r head_b1d1}
head(ne$formb1d)
```

#### Form B-2

[form B-2][07] covers Political Party Committees, where Item 5 of that form asks the party to:

> List expenditures of any amount which were made to or for the benefit of candidates or ballot
questions. If your committee made multiple contributions in support or opposition to the same
candidate or ballot question committee, also list the total accumulated contributions given for the
election. Expenditures made in support of more than one candidate or ballot question committee must
be allocated (E.g. the purchase of a newspaper ad advocating support or opposition to more than one
candidate or ballot question).

```{r head_b2b}
head(ne$formb2b)
```

#### Form B-4

[form B-4][08] covers Independent Committees, where Section 1 of Schedule B asks the committee to: 

> Indicate the total amount of expenditures, for or against each Nebraska candidate or ballot
question (B.Q.) or other allowed committee (including a political party), during this reporting
period and the total for the calendar year to date. An expenditure made in support of more than one
candidate or ballot question must be apportioned reasonably among the candidates or ballot
questions supported or opposed. "Expenditure" for the purpose of this schedule includes direct and
in-kind contributions, loans, independent expenditures, and all other disbursements of funds made
in support of or in opposition to Nebraska state and local candidates and ballot questions.

```{r head_b4b1}
head(ne$formb4b1)
```

#### Bind

We can bind these three data frames together to create a single data frame of expenditures by
candidates, ballot committees, political committees, and party committees. Then we can clean and
parse this combined data frame.

Before we bind these tables together, we need to rename some variables and add additional
information from other forms so we can identify the parties on both sides of each expenditure.

```{r fix_b1d}
fixed_b1d <- ne$formb1d %>% 
  # add from b1
  select(-committee_name)
```

```{r fix_b2b}
fixed_b2b <- ne$formb2b %>%
  remove_empty("cols") %>% 
  rename(
    committee_payee_id = committee_id_expenditure_is_for,
    # this file has last cols shifted left
    committee_payee_name = report_id,
  )
```

```{r fix_b4b1}
fixed_b4b1 <- ne$formb4b1 %>%
  remove_empty("cols") %>% 
  rename(
    committee_payee_id = committee_expenditure_id,
    # this file has last cols shifted left
    committee_payee_name = report_id,
  ) %>% 
  select(
    -form_id_number
  )
```

```{r bind_expends}
ne_exp <- 
  bind_rows(
    fixed_b1d, 
    fixed_b2b, 
    fixed_b4b1
  ) %>% 
  na_if("0") %>% 
  mutate(
    date_received = parse_date(date_received, "%m/%d/%Y"),
    expenditure_date = parse_date(expenditure_date, "%m/%d/%Y"),
    oppose_clean = parse_logical(support_oppose),
    amount_clean = parse_number(coalesce(amount, in_kind)),
    inkind_clean = !is.na(in_kind),
    payee_name_clean = coalesce(payee_name, committee_payee_name)
  )
```

## Explore

```{r glimpse}
head(ne_exp)
tail(ne_exp)
glimpse(sample_frac(ne_exp))
```

### Missing

```{r glimpse_na}
glimpse_fun(ne_exp, count_na)
```

Since the combined data frame comes from three different sources, it's difficult to identify what
are expected missing values and those that should be flagged. `na_flag` will equal `TRUE` when a 
record has either (1) no `amount_clean` (`amount` or `in_kind`) _or_ (2) no `payee_name_clean`
(`payee_name` for individuals or `payee_committee_name` for others).

```{r flag_na}
sum(is.na(ne_exp$payee_name) & is.na(ne_exp$committee_payee_name))

ne_exp <- ne_exp %>% 
  mutate(na_flag = is.na(amount_clean) | is.na(payee_name_clean))

sum(ne_exp$na_flag)
percent(mean(ne_exp$na_flag))
```

### Duplicates

We can use `campfin::flag_dupes()` to create a new `dupe_flag` logical variable, identifying every
duplicate row _after_ the initial occurance.

```{r get_dupes, collapse=TRUE}
ne_exp <- flag_dupes(ne_exp)
sum(ne_exp$dupe_flag)
```

```{r view_dupes}
filter(ne_exp, dupe_flag)
```

### Categorical

```{r glimpse_distinct}
glimpse_fun(ne_exp, n_distinct)
```

```{r words_bar, echo=FALSE, fig.height=10}
ne_exp %>% 
  unnest_tokens(word, expenditure_purpose) %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  head(25) %>% 
  drop_na() %>% 
  ggplot(
    mapping = aes(
      x = reorder(word, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Nebraska Expenditure Purpose",
    x = "Word",
    y = "Count"
  )
```

```{r nature_bar, echo=FALSE}
explore_plot(
  palette = "Dark2",
  nbar = 5,
  data = ne_exp,
  var = nature_of_expenditure,
  title = "Nebraska Expenditure Nature",
  x = "Word",
  y = "Count"
)
```

### Continuous

#### Amounts

```{r amount_range, collapse=TRUE}
summary(ne$amount)
sum(ne$amount < 0, na.rm = TRUE)
```

```{r}
ne_exp %>% 
  ggplot(aes(amount_clean)) +
  geom_histogram() +
  scale_x_continuous(
    trans = "log10",
    labels = dollar
  ) +
  labs(
    title = "Nebraska Expenditure Amount Distribution",
    x = "Word",
    y = "Count"
  )
```

#### Dates

```{r}
ne_exp <- mutate(ne_exp, expenditure_year = year(expenditure_date))
```

```{r date_range, collapse=TRUE}
min(ne_exp$expenditure_date)
sum(ne_exp$expenditure_year < 1999)
max(ne_exp$expenditure_date)
sum(ne_exp$expenditure_date > today())
```

```{r flag_dates, collapse=TRUE}
ne_exp <- ne_exp %>% 
  mutate(
    date_flag = or(expenditure_year < 1999, expenditure_year > 2019),
    date_clean = ifelse(date_flag, NA, expenditure_date)
  )

sum(ne_exp$date_flag)
```

## Wrangle

There is very little consistency in the format of `payee_address`.

```{r sample_address}
sample(
  x = na.omit(ne_exp$payee_address),
  size = 20
)
# Mon Aug  5 17:19:45 2019 ------------------------------
```

## Conclude 

## Export

```{r work_dir}
work_dir <- here("ne", "expends", "data", "working")
dir_create(work_dir)
```

```{r write_clean}
ne_exp %>% 
  write_csv(
    na = "",
    path = glue("{work_dir}/ne_expends_working.csv")
  )
```

[01]: https://github.com/irworkshop/accountability_datacleaning "tap"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "rproj"
[03]: http://www.nebraska.gov/government/open-data/ "portal"
[04]: http://www.nadc.nebraska.gov/index.html "nadc"
[05]: http://www.nadc.nebraska.gov/cf/index.html "nadc_cf"
[06]: http://www.nadc.nebraska.gov/docs/B-1-2018.doc "b-1"
[07]: http://www.nadc.nebraska.gov/docs/B-2-2018.doc "b-2"
[08]: http://www.nadc.nebraska.gov/docs/B-4-2018.doc "b-4"
