---
title: "Nebraska Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ne", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  textreadr, # read text files
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained by from the Nebraska Accountability and Disclosure Commission's
(NADC) [Open Data portal][03]. From this portal, we will download the 
[Campaign Statements Data][04].

> A weekly export of the campaign filings based upon the paper records filed
with the Nebraska Accountability and Disclosure Commission.

The [NADC home][04] page explains the role and resources provided by the 
Commission.

> This site includes the campaign statements filed with the Commission by
candidates for state and local offices, political action committees or PACs,
political parties, and ballot question committees. Campaign statements disclose
the amount and sources of contributions to the committee and the amount purposes
of expenditures by the committee. The site also includes Reports of
Contributions filed with the Commission by corporations, unions, limited
liability companies, limited partnerships, and by industry, trade or
professional associations. Reports of Contributions disclose contributions and
expenditures made to support or oppose candidates or ballot questions. Forms,
instructional materials and a list of committees may be accessed through this
site. To access the information described in this paragraph, click on the tab
above marked CAMPAIGN FINANCE.

Navigating to the [campaign finance section of the website][05], the exact
filing requirements are laid out.

> Candidates for state and local office are subject to the campaign finance
provisions of the Nebraska Political Accountability and Disclosure Act (NPADA).
Those seeking election to federal office are not subject to the NPADA. Once a
candidate raises, receives or expends $5,000 or more in a calendar year, he or
she is required to register the committee with the Commission by filing a
Statement of Organization and paying a filing fee of $100. Once the candidate
committee is formed, the committee must file periodic campaign statements.

specifically, there are **three** kinds of filings that must be made by general
candidates:

1. > Statement of Organization - This document registers the Committee with the
Commission. It discloses the name of the committee, the name of the candidate,
the office sought and the name of the treasurer.

2. > Campaign Statement - this document provides a summary of the financial
activity of the committee. It discloses the name of contributors of more than
\$250 and the amounts of the contributions. It discloses expenditures of more
than \$250.

3. > Report of Late Contributions - this document discloses contributions of
$1,000 or more received by the committee during 14 days immediately preceding
the election. Information reported as late contributions are also reported on
subsequent campaign statements.

Aside from general candidates, the NADC also explains that political committees
must also file.

> There are **two** types of committees that are commonly referred to as
political action committees or PACs. One type of PAC is a political committee
sponsored by a corporation, union or other type of organization. Another type of
PAC is a political committee which is not sponsored by another group. PACs are
required to register with the Commission and file periodic campaign statements
disclosing receipts and expenditures.

1. > State and county political party committees with certain levels of
financial activity, register with the Commission and file periodic campaign
statements.

2. > A ballot question committee is a political committee formed for the purpose
of supporting or opposing the qualification, passage or defeat of a ballot
question. The committee must register with the Commission within 10 days after
raising, receiving or expending $5,000 or more in a calendar year. Once
registered, ballot questions committees file periodic campaign statements
disclosing receipts and expenditures.

Finally, the NADC identifies the source of the data:

> The campaign filings displayed on this website are based upon the paper
records filed with the Nebraska Accountability and Disclosure Commission. While
every effort has been made to ensure accurate data input, errors may occur.
Anyone noting an error is encouraged to contact this office.

In any discrepancy between the information displayed on this website and the
official records on file with the Commission, the official records of the
Commission take precedence.

[01]: https://github.com/irworkshop/accountability_datacleaning
[03]: http://www.nebraska.gov/government/open-data/
[04]: http://www.nadc.nebraska.gov/index.html
[05]: http://www.nadc.nebraska.gov/cf/index.html
[06]: http://www.nadc.nebraska.gov/docs/B-1-2018.doc
[07]: http://www.nadc.nebraska.gov/docs/B-2-2018.doc
[08]: http://www.nadc.nebraska.gov/docs/B-4-2018.doc

## Download

We can download the ZIP archive locally.

```{r raw_download}
raw_dir <- dir_create(here("ne", "expends", "data", "raw"))
zip_url <- "http://www.nebraska.gov/nadc_data/nadc_data.zip"
zip_path <- path(raw_dir, basename(zip_url))
if (!file_exists(zip_path)) {
  download.file(zip_url, zip_path)
}
```

Then we can extract the contents to our `raw/data/` directory.

```{r zip_info}
file_size(zip_path)
(zip_content <- zip_path %>% 
  unzip(list = TRUE) %>% 
  as_tibble(.name_repair = tolower) %>% 
  mutate(across(length, as_fs_bytes)) %>% 
  mutate(across(name, as_fs_path)))
```

```{r zip_extract}
raw_paths <- as_fs_path(unzip(
  zipfile = zip_path, 
  exdir = raw_dir,
  junkpaths = TRUE
))
```

The first file in the folder, `DATE_UPDATED.TXT` contains the time the files were last updated.

```{r zip_date}
date_updated <- read_file(file = path(raw_dir, "DATE_UPDATED.TXT"))
print(date_updated)
date_updated <- as_datetime(str_extract(date_updated, "(?<=:\\s).*"))
date_updated - now()
```

## Read

Before we can prepare the file for indexing, we need to locate the exact data
we're interested in. The content, structure, and relationship of each file is
described in the `nadc_tables.rtf` rich text file. The files correspond to the
_forms_ through which candidates and committees report their contributions,
expenditures, etc. Each type of filter uses a different form, so all
expenditures are split among a few files (corresponding to a few forms). Each
form typically also has a section (schedule) where _expenditures_ are reported.

```{r tales_read}
ne_tables <- read_rtf(path(raw_dir, "nadc_tables.rtf"))
```

Using this file, we can grasp what information is contained on what forms and
which tables we will need to combine to identify the parties, date, and amount
of every campaign expenditure.

Form A1 contains all of the names and addresses of political, candidate, and
ballot question committees. This form has no transactions.

* Date Received
* Committee ID Number
* Committee Name
* Committee Address
* Committee City
* Committee State
* Committee Zip
* Committee Type
    * C = Candidate Committee
    * B = Ballot Question
    * P = Political Action Committee
    * T = Political Party Committee
    * I or R = Independent Reporting Committee
    * S = Separate Segregated Political Fund Committee

Then the various types of transactions for those committee types are reported
on subsequent forms:

* Form **B1D** – Form B-1 Schedule D Section 1, Expenditures
* Form **B2B** – Form B-2 Expenditures:
* Form **B4B1** – Form B-4 Schedule B Section 1
* Form **B6EXPEND** – Form B-6 Expenditures:
* Form **B9B** – Form B-9 Schedule B, expenditure
* Form **B10EXP** – Form B-10 expenditures
* Form **B11** – Report of Late Independent Expenditure

When we read all these sub-forms together and join them against the committee
statements then we can identify all transactions

### Form B

```{r read_b1}
b1 <- read_delim(
  file = path(raw_dir, "formb1.txt"),
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Date Received` = col_date_usa(),
    `Expenditure Date` = col_date_usa(),
    `Amount` = col_double(),
    `In-Kind` = col_double()
  )
)
```

```{r fix_b1d, echo=FALSE}
b1 <- b1 %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish))
```

### Form B-1D

> List all payees who were paid more than $250 during this reporting period. If
multiple payments to the same payee totaled more than $250 throughout this
reporting period, those expenditures must be listed. Reporting period refers to
your entry on Page 1 under Item 4. Expenditures to the same payee over separate
reporting periods should not be accumulated.  Expenditures to the same payee
must be listed under the same name. If the committee reimburses the candidate or
engages the services of an advertising agency or another agent of the committee
for expenses they incurred on behalf of the committee, list the payments the
committee made to the candidate or agent and also list the payments which were
made by the candidate or agent on behalf of the committee.

```{r vars_b1d}
md_order(ne_tables[141:149])
```

```{r read_b1d}
b1d <- read_delim(
  file = path(raw_dir, "formb1d.txt"),
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Date Received` = col_date_usa(),
    `Expenditure Date` = col_date_usa(),
    `Amount` = col_double(),
    `In-Kind` = col_double()
  )
)
```

```{r fix_b1d, echo=FALSE}
b1d <- b1d %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  select(
    comm_id = committee_id,
    payee_name, payee_address, 
    date = expenditure_date,
    amount, in_kind
  )
```

```{r amount_b1d}
b1d <- b1d %>% 
  # combine and indicate amount
  mutate(across(5:6, na_if, "0")) %>% 
  mutate(
    amount = coalesce(amount, in_kind),
    in_kind = !is.na(in_kind)
  )
```

```{r show_b1d, echo=FALSE}
print(b1d)
```

### Form B-2B

[form B-2][07] covers Political Party Committees, where Item 5 of that form asks the party to:

> List expenditures of any amount which were made to or for the benefit of
candidates or ballot questions. If your committee made multiple contributions in
support or opposition to the same candidate or ballot question committee, also
list the total accumulated contributions given for the election. Expenditures
made in support of more than one candidate or ballot question committee must be
allocated (E.g. the purchase of a newspaper ad advocating support or opposition
to more than one candidate or ballot question).

```{r lines_b2b}
b2b <- read_lines(path(raw_dir, "formb2b.txt"))
b2b[1] <- str_remove(b2b[1], "Report ID\\|")
```

```{r read_b2b}
b2b <- read_delim(
  file = b2b,
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Date Received` = col_date_usa(),
    `Support/Oppose` = col_logical(),
    `Expenditure Date` = col_date_usa(),
    `Amount` = col_double(),
    `Line ID` = col_double()
  )
)
```

```{r fix_b1d, echo=FALSE}
b2b <- b2b %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  select(
    comm_id = committee_id,
    payee_id = committee_id_expenditure_is_for,
    date = expenditure_date,
    amount
  )
```

```{r show_b1d, echo=FALSE}
print(b2b)
```

### Form B-4B1

[form B-4][08] covers Independent Committees, with Section 1 of Schedule B: 

> Indicate the total amount of expenditures, for or against each Nebraska
candidate or ballot question (B.Q.) or other allowed committee (including a
political party), during this reporting period and the total for the calendar
year to date. An expenditure made in support of more than one candidate or
ballot question must be apportioned reasonably among the candidates or ballot
questions supported or opposed. "Expenditure" for the purpose of this schedule
includes direct and in-kind contributions, loans, independent expenditures, and
all other disbursements of funds made in support of or in opposition to Nebraska
state and local candidates and ballot questions.

```{r lines_b2b}
b4b1 <- read_lines(path(raw_dir, "formb4b1.txt"))
b4b1[1] <- str_remove(b4b1[1], "Report ID\\|")
```

```{r read_b2b}
b4b1 <- read_delim(
  file = b4b1,
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Date Received` = col_date_usa(),
    `Support/Oppose` = col_logical(),
    `Expenditure Date` = col_date_usa(),
    `Amount` = col_double(),
    `Expense Category` = col_double()
  )
)
```

```{r fix_b1d, echo=FALSE}
b4b1 <- b4b1 %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  select(
    comm_id = committee_id,
    payee_id = committee_expenditure_id,
    date = expenditure_date,
    amount
  )
```

```{r show_b1d, echo=FALSE}
print(b4b1)
```

### Form B-11

> A Report of Late Independent Expenditure is filed on time if it is postmarked
or personally delivered on or before the due date, or if it is sent by facsimile
in such time as to cause it to be received by the Commission Office on the date
on which the report is due.  If this report is sent by facsimile, the sender
should call the Commission Office immediately thereafter to ensure it was
received.

```{r read_b11}
b11 <- read_delim(
  file = path(raw_dir, "formb11.txt"),
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Postmark Date` = col_date_usa(),
    `Date Received` = col_date_usa(),
    `Expenditure Date` = col_date_usa(),
    `Amount` = col_double(),
    `Date Last Revised` = col_date_usa()
  )
)
```

```{r fix_b1d, echo=FALSE}
b11 <- b11 %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  select(
    comm_id = committee_id,
    starts_with("recipient"),
    date = expenditure_date,
    amount
  ) %>% 
  rename_with(~str_replace(., "recipient", "payee"))
```

```{r show_b1d, echo=FALSE}
print(b11)
```

### Join

```{r read_a1}
a1 <- read_delim(
  file = path(raw_dir, "forma1.txt"),
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Date Received` = col_date_usa(),
    `Postmark Date` = col_date_usa(),
    `Oppose Ballot Question` = col_logical(),
    `Date over Theshold` = col_date_usa(),
    `Date Dissolved` = col_date_usa(),
    `Date of Next Election` = col_date_usa(),
    `Won General` = col_logical(),
    `Won Primary` = col_logical()
  )
)
```

```{r fix_a1, echo=FALSE}
a1 <- a1 %>% 
  clean_names("snake") %>% 
  mutate(across(where(is.character), str_squish)) %>% 
  select(starts_with("committee")) %>% 
  rename_with(~str_replace(., "^committee_", "comm_")) %>% 
  rename(comm_id = comm_id_number)
```

```{r print_a1, echo=FALSE}
```

```{r bind_bx}
bx <- bind_rows(b1d, b2b, b4b1)
```

```{r join_ab}
bx <- left_join(bx, a1, by = "comm_id")
ax <- rename_with(a1, ~str_replace(., "comm_", "payee_"))
ax <- select(ax, -payee_type)
bx <- left_join(bx, ax)
```

## Explore

```{r glimpse}
glimpse(nee)
tail(nee)
```

### Missing

```{r na_count}
col_stats(nee, count_na)
```

```{r na_flag}
nee <- nee %>% flag_na(date, last, amount, committee)
sum(nee$na_flag)
```

```{r na_view}
nee %>% 
  filter(na_flag) %>% 
  select(date, last, amount, committee)
```

### Duplicates

```{r dupe_flag}
nee <- flag_dupes(nee, -id)
sum(nee$dupe_flag)
```

```{r dupe_view}
nee %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, committee)
```

### Categorical

```{r distinct_count}
col_stats(nee, n_distinct)
```

```{r distinct_plots}
explore_plot(nee, type)
```

### Amounts

```{r ammount_summary}
summary(nee$amount)
mean(nee$amount <= 0)
```

```{r hist_amount, echo=FALSE}
nee %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Nebraska Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nee <- mutate(nee, year = year(date))
```

```{r date_range}
min(nee$date)
sum(nee$year < 2000)
max(nee$date)
sum(nee$date > today())
```

```{r bar_year, echo=FALSE}
nee %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nebraska Expenditures by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
nee <- nee %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
nee %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nee <- nee %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nee$zip,
  nee$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nee <- nee %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
nee %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  nee$state,
  nee$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
nee <- nee %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NE", "DC", "NEBRASKA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
nee <- nee %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- nee %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
nee <- nee %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(nee$city_raw),
  nee$city_norm,
  nee$city_swap,
  nee$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(nee$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Nebraska City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nebraska City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
nee <- nee %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(nee, 50))
```

1. There are `r comma(nrow(nee))` records in the database.
1. There are `r comma(sum(nee$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(nee$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ne", "expends", "data", "clean"))
clean_path <- path(clean_dir, "ne_expends_clean.csv")
write_csv(nee, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(nee)),
  type = md_code(map_chr(nee, typeof)),
  def = c(
    "Date contribution was made",
    "Contribution amount or correction",
    "Contributor name",
    "Contributor unique ID",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Contributor type",
    "Recipient committee name",
    "Recipient unique ID",
    "Contribution method",
    "Recipient election or office sought",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Nebraska Contracts Data Dictionary\n", dict_md),
  path = here("ne", "expends", "ne_expends_dict.md"),
)
```
