---
title: "Utah Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ut", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data can be obtained from the State of Utah [Division of Purchasing 
and General Services][utdp] via the state's [Open Data portal][odp]. The file
was originally uploaded September 12, 2018 and last updated on December 15, \
2019. The Open Data [data file][odf] has the cryptic title "Latest 6 Years" and
no additional metadata.

After contacting the Division of Purchasing, it was explained that this file
actually contains almost all payments made by the state accounting system, 
including biweekly payroll, jury duty compensation, etc.

[utdp]: https://purchasing.utah.gov/
[odp]: https://opendata.utah.gov
[odf]: https://opendata.utah.gov/dataset/Latest-6-Years/545s-5tnq

## Read

The data file is large and must be downloaded before it can be read.

```{r raw_path}
raw_dir <- dir_create(here("ut", "contracts", "data", "raw"))
raw_url <- "https://opendata.utah.gov/api/views/545s-5tnq/rows.tsv"
raw_path <- path(raw_dir, basename(raw_url))
```

```{r raw_download}
if (!file_exists(raw_path)) {
  download.file(
    url = raw_url,
    destfile = raw_path
  )
}
file_size(raw_path)
```

Then the file can be read as a single, large data frame with `vroom::vroom()`.

```{r raw_read}
utc <- vroom(
  file = raw_path,
  delim = "\t",
  na = c("Not Applicable", "Not Provided", "NA", ""),
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    FISCAL_PERIOD = col_integer(),
    CALC_posting_date = col_date("%m/%d/%Y %H:%M:%S %p"),
    CALC_FISCAL_PERIOD_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    CALC_amount = col_double(),
    fiscal_year = col_integer(),
    type = col_integer()
  )
)
```

```{r raw_rename, echo=FALSE}
utc <- rename_all(utc, str_remove, "calc_")
utc <- utc %>% 
  relocate(amount, .before = fiscal_period) %>% 
  select(-fiscal_period_date, -cat3) %>% 
  rename(
    date = posting_date,
    department = org1,
    agency = org2,
    division = org3,
    category = cat1,
    subcategory = cat2
  )
```

After reading the entire file, we can filter out any records not directly
related to contracts using the `contract_name` and `vendor_name` values.
After this filter, most rows are removed and we are left only with valid
contracts, most of which have a `vendor_name`.

```{r raw_filter}
nrow(utc)
utc <- filter(utc, !is.na(contract_name))
count(utc, vendor_name, sort = TRUE)
nrow(utc)
```

## Explore

```{r glimpse}
glimpse(utc)
tail(utc)
```

### Missing

After this filter, we are left with very few records missing any key values.

```{r na_count}
col_stats(utc, count_na)
```

### Duplicates

Flagging duplicates in a file this size requires breaking the data frame into
chunks and writing the duplicate rows to a local file.

```{r dupe_flag}
dupe_file <- file_temp(ext = "txt")
n <- 1e5 # rows per chunk
nr <- nrow(utc)
# split file into chunks
uts <- utc %>% 
  select(-ends_with("id")) %>% 
  split(rep(1:ceiling(nr/n), each = n, length.out = nr))
pb <- txtProgressBar(0, length(uts), style = 3)
for (i in seq_along(uts)) {
  # check dupes from both ends
  d <- as.integer(duplicated(uts[[i]], fromLast = TRUE))
  # append to disk
  write_lines(d, dupe_file, append = TRUE)
  uts[[i]] <- NA
  setTxtProgressBar(pb, i)
  Sys.sleep(10)
  flush_memory()
}
rm(uts)
```

This local file can be read, checked, and added as a new variable if needed.

```{r dupe_read}
dupes <- as.logical(as.integer(read_lines(dupe_file)))
percent(mean(dupes), 0.1)
utc <- mutate(utc, dupe_flag = dupes)
utc %>% 
  filter(dupe_flag) %>% 
  select(id, date, agency, amount, vendor_name) %>% 
  arrange(date)
rm(dupes)
```

A fairly significant amount of these rows are seemingly duplicated at least once
across every (non-id) variable.

### Categorical

There are a handful of categorical values who's distribution can be explored.

```{r distinct_count}
col_stats(utc, n_distinct)
```

```{r echo=FALSE}
flush_memory()
```

```{r distinct_plots, echo=FALSE}
explore_plot(utc, department) + scale_x_truncate()
explore_plot(utc, agency) + scale_x_truncate()
explore_plot(utc, category) + scale_x_truncate()
explore_plot(utc, subcategory) + scale_x_truncate()
explore_plot(utc, description) + scale_x_truncate()
explore_plot(utc, fund1) + scale_x_truncate()
explore_plot(utc, fund2) + scale_x_truncate()
```

### Amounts

We should also check the distribution of contract `amount` values.

```{r ammount_sumary}
noquote(map_chr(summary(utc$amount), dollar))
```

A significant amount of the `amount` values are negative.

```{r amount_negative}
percent(mean(utc$amount < 0), 0.1)
```

Most of the negative values are from purchasing cards. Negative values sometimes
represent corrections or deletions. We will leave them as-is.

```{r amount_neg_view}
utc %>% 
  filter(amount < 0) %>%
  count(description, sort = TRUE) %>% 
  add_prop()
```

The amounts are logarithmically-normally distributed with reasonable ranges.

```{r hist_amount, echo=FALSE}
utc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Utah Contracts Amount Distribution",
    x = "Amount",
    y = "Count"
  )
```

Here are the smallest and largest contracts:

```{r amount_minmax}
glimpse(utc[which.max(utc$amount), ])
glimpse(utc[which.min(utc$amount), ])
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
utc <- mutate(utc, year = year(date))
```

The date range is very clean.

```{r date_range}
min(utc$date)
sum(utc$year < 2013)
max(utc$date)
sum(utc$date > today())
```

The number of contracts/payments per year is very consistent.

```{r bar_year, echo=FALSE}
utc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2013:2019) +
  theme(legend.position = "bottom") +
  labs(
    title = "Utah Contracts by Year",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

There are no geographic variables, but we can manually add "UT" for all state
agencies.

```{r state_add}
utc <- mutate(utc, state = "UT", .before = agency)
```

## Conclude

1. There are `r comma(nrow(utc))` records in the database.
1. There are `r comma(sum(utc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(utc$na_flag))` records missing key variables.
1. There are no geographic variables to normalize, `state` was added.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r echo=FALSE}
flush_memory()
```

```{r clean_dir}
clean_dir <- dir_create(here("ut", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "ut_contracts_clean.csv")
vroom_write(utc, clean_path, na = "", delim = ",")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(utc)),
  type = md_code(map_chr(utc, typeof)),
  def = c(
    "Unique contract/payment ID",
    "Date contract posted",
    "Contrat paid out amount",
    "Fiscal period (month)",
    "Fiscal batch order",
    "Fiscal year ordered",
    "Semi-unique transaction ID",
    "Transasction description",
    "Reference ID",
    "Contract type (all 1)",
    "Contract name",
    "Contract number",
    "Spending department (`org1)",
    "Spending state (all UT)",
    "Spending agency (`org2`)",
    "Spending division/use (`org3`)",
    "Contract category (`cat1`)",
    "Contract subcategory (`cat2`)",
    "Contract fund source (`fund1`)",
    "Contract fund secondary source (`fund2`)",
    "Unique vendor ID",
    "Vendor name",
    "Flag indicating duplicate record",
    "Calendar year spent"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
