---
title: "COVID-19 Economic Injury Disaster Loans Diary"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "eidl", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## About

The Small Business Administration's [Economic Injury Disaster Loans][eidl] is a
program typically given to relieve the economic strain of natural disasters on
local economies. The loans are meant to cover operating expenses that could
have been met had the disaster not occurred. The loans have a 3.75% interest
rate for businesses and 2.75% for nonprofits over a 30 year term.

The EIDL program was expanded during the COVID-19 pandemic:

> In response to the Coronavirus (COVID-19) pandemic, small business owners,
including agricultural businesses, and nonprofit organizations in all U.S.
states, Washington D.C., and territories can apply for an Economic Injury
Disaster Loan. The EIDL program is designed to provide economic relief to
businesses that are currently experiencing a temporary loss of revenue due to
coronavirus (COVID-19).

The type of loans offered for COVID-19 applicants differ slightly:

> #### Who can use an SBA disaster loan?
> 
> _COVID-19 EIDL_: If you are a small businesses, nonprofit organization of any
size, or a U.S. agricultural business with 500 or fewer employees that have
suffered substantial economic injury as a result of the Coronavirus (COVID-19)
pandemic, you can apply for the COVID-19 EIDL. This loan applies to all
businesses based in any U.S. state, territory, or the District of Columbia.
>
> _ALL OTHER DISASTERS_: Businesses of all sizes located in declared disaster
areas, private nonprofit organizations, homeowners, and renters affected by
declared disaster, including civil unrest and natural disasters such as
hurricanes, flooding, wildfires, etc.

> #### How to use an SBA disaster loan
> 
> ###### _COVID-19 EIDL_
> * For working capital
> * Normal operating expenses (i.e. continuation of health care benefits, rent,
utilities, and fixed debt payments).
>
> ###### _ALL OTHER DISASTERS_
> * Losses not covered by insurance or funding from the Federal Emergency
Management Agency (FEMA) for both personal and business.
> * Business operating expenses that could have been met had the disaster not
occurred.

There are _two_ type of loans offered through the EIDL program. Regular loans
are not forgivable while the _Advance_ loans can be forgiven:

> EIDL Advance was a grant program offered together with the economic injury
loan program. The amount of the grant was determined by the number of employees
indicated on the EIDL application: $1,000/employee, up to a maximum of $10,000.
> 
> * EIDL Advance does not have to be repaid.
> * Recipients did not have to be approved for an EIDL loan to receive the EIDL
Advance.
> * The amount of the loan Advance was deducted from total loan eligibility.
> * Businesses who received an EIDL Advance in addition to the Paycheck
Protection Program (eidl) loan will have the amount of the EIDL Advance
subtracted from the forgiveness amount of their eidl loan.

[eidl]: https://www.sba.gov/funding-programs/disaster-assistance

## Data

In late November 2020, the Small Business administration was ordered by a 
federal court to release detailed data on the recipients of Paycheck Protection
Program and Economic Disaster Injury Loans. On December 1, the SBA complied with
this request and published the data on their [Box page][box]. The EIDL data is
separated into regular and advance loan folders. These can both be downloaded
locally.

[box]: https://sba.app.box.com/s/5myd1nxutoq8wxecx2562baruz774si6

We can download the ZIP archives fro Box locally.

```{r raw_dir}
raw_dir <- dir_create(here("us", "covid", "eidl_covid", "data", "raw"))
raw_zip <- path(raw_dir, "120120 EIDL Data.zip")
adv_zip <- path(raw_dir, "120120 EIDL Advance Data.zip")
```

We can extract the CSV files from each archive.

```{r eval=FALSE}
raw_paths <- unzip(raw_zip, exdir = raw_dir, junkpaths = TRUE)
(raw_info <- raw_paths %>%
  map_df(unzip, list = TRUE) %>%
  as_tibble() %>%
  clean_names() %>%
  filter(length > 212))
```

```{r eval=FALSE}
for (i in seq_along(raw_paths)) {
  raw_paths[i] <- unzip(
    zipfile = raw_paths[i],
    files = raw_info$name[i],
    exdir = raw_dir,
    overwrite = FALSE
  )
}
```

The process can be repeated for the advance loan files.

```{r eval=FALSE}
adv_paths <- unzip(adv_zip, exdir = raw_dir, junkpaths = TRUE)
(adv_info <- adv_paths %>%
  map_df(unzip, list = TRUE) %>%
  as_tibble() %>%
  clean_names() %>%
  filter(length > 212))
```

```{r eval=FALSE}
for (i in seq_along(adv_paths)) {
  adv_paths[i] <- unzip(
    zipfile = adv_paths[i],
    files = adv_info$name[i],
    exdir = raw_dir,
    overwrite = FALSE
  )
}
```

```{r echo=FALSE}
raw_paths <- dir_ls(raw_dir, regexp = "EIDL through.*csv$")
adv_paths <- dir_ls(raw_dir, regexp = "EIDL Advance.*csv$")
```

### Dictionary

We can download an Excel file from the Treasury Depeartment that contains some
definitions for the variables in the data.

```{r dict_get}
raw_names <- read_names(raw_paths[1])
dict_file <- path_temp("IDD-v2.0.xlsx")
dict_url <- "https://www.publicdebt.treas.gov/files/data-transparency/"
download.file(
  url = str_c(dict_url, basename(dict_file)),
  destfile = dict_file
)
```

```{r dict_read}
dict <- read_excel(
  path = dict_file,
  sheet = "D2-Award (Financial Assistance)",
  .name_repair = make_clean_names,
  skip = 2
)
```

```{r dict_trim, echo=FALSE}
dict <- dict %>% 
  filter(str_to_upper(data_element_label) %in% raw_names) %>% 
  select(
    column = data_element_label,
    grouping,
    example_value,
    type = valid_data_type,
    length = max_field_length,
    definition
  )
```

```{r dict_print}
dict %>% 
  select(column, definition) %>% 
  mutate(across(column, md_code)) %>% 
  kable()
```

## Read

```{r echo=FALSE}
eidl_names <- c(
  "action_type",
  "action_date",
  "assist_type",
  "record_type",
  "fain",
  "award_amend",
  "uri",
  "correct_delete",
  "fy_qt_correct",
  "sai_num",
  "awardee_name",
  "awardee_uuid",
  "addr1",
  "addr2",
  "addr3",
  "city",
  "state",
  "zip5",
  "zip4",
  "country",
  "foreign_city",
  "foreign_prov",
  "foreign_post",
  "congress_dist",
  "business_type",
  "fund_agency",
  "fund_sub",
  "fund_office",
  "award_agency",
  "award_sub",
  "award_office",
  "cfda_num",
  "prime_cd",
  "prime_county",
  "prime_zip9",
  "prime_foreign",
  "prime_congress",
  "award_desc",
  "period_start",
  "period_now",
  "fed_obligation",
  "nonfed_amount",
  "face_value",
  "og_loan_sub",
  "biz_fund_id"
)
```

```{r raw_read}
eidl <- map_df(
  .x = c(raw_paths, adv_paths),
  .f = read_delim,
  .id = "source_file",
  delim = ",",
  skip = 1,
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = eidl_names,
  col_types = cols(
    .default = col_character(),
    action_date = col_date("%Y%m%d"),
    record_type = col_integer(),
    award_amend = col_integer(),
    period_start = col_date("%Y%m%d"),
    period_now = col_date("%Y%m%d"),
    fed_obligation = col_double(),
    nonfed_amount = col_double(),
    face_value = col_double(),
    og_loan_sub = col_double()
  )
)
```

```{r eval=FALSE}
if (interactive()) {
  eidl <- eidl %>% 
    remove_constant() %>% 
    remove_empty("cols")
}
```

```{r raw_source}
eidl <- eidl %>% 
  relocate(source_file, .after = last_col()) %>% 
  mutate(advance_loan = str_detect(source_file, "Advance"))
```

## Explore

There are `r comma(nrow(eidl))` rows of `r ncol(eidl)` rows. Each record
represents a disaster loan made to a small business.

```{r glimpse}
nrow_pre <- nrow(eidl)
glimpse(eidl)
tail(eidl)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(eidl, count_na)
```

We can flag any record missing a variable needed to identify the transaction.

```{r na_amount_nonfed}
unique(eidl$nonfed_amount)
```

```{r na_amount_view}
eidl %>% 
  select(fed_obligation, face_value, og_loan_sub) %>% 
  sample_n(20)
```

```{r na_amount_coalesce}
eidl <- mutate(eidl, amount = coalesce(fed_obligation, face_value))
```

```{r na_flag}
key_vars <- c("action_date", "awardee_name", "award_office", "amount")
eidl <- flag_na(eidl, all_of(key_vars))
```

```{r na_view}
sum(eidl$na_flag)
sum(is.na(eidl$awardee_name))
eidl %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("us", "covid", "eidl_covid", "dupes.tsv")
```

We flag these records by first identifying rows by a unique ID, then splitting
the data into chunks along a discrete variable and checking each chunk for 
duplicate rows and saving any duplicate rows to a file.

```{r dupe_id}
eidl <- mutate(eidl, id = row_number(), .after = last_col())
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  eidls <- eidl %>% 
    select(-id, -fain, ) %>%
    group_split(action_date)
  split_id <- split(eidl$id, eidl$action_date)
  n_done <- 0
  pb <- txtProgressBar(max = nrow(eidl), style = 3)
  for (i in seq_along(eidls)) {
    n_new <- nrow(eidls[[i]])
    d1 <- duplicated(eidls[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(eidls[[i]], fromLast = TRUE)
      dupes <- tibble(voter_id = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, dupe_file, append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    eidls[i] <- list(NA)
    flush_memory(1)
    n_done <- n_done + n_new
    setTxtProgressBar(pb, value = n_done)
  }
  rm(eidls)
}
```

This file of duplicates can then be read and joined against the original data.

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = dupe_file,
  col_names = c("id", "dupe_flag"),
  col_types = cols(
    id = col_integer(),
    dupe_flag = col_logical()
  )
)
```

```{r dupe_join}
nrow(eidl)
eidl <- left_join(eidl, dupes, by = "id")
eidl <- mutate(eidl, dupe_flag = !is.na(dupe_flag))
sum(eidl$dupe_flag)
eidl <- select(eidl, -id)
```

We can see that, despite a supposedly unique `fain`, there are duplicate rows.

```{r dupe_view}
eidl %>% 
  filter(dupe_flag) %>% 
  select(fain, award_amend, all_of(key_vars), city) %>% 
  arrange(action_date, awardee_name)
```

### Categorical

```{r distinct_count}
col_stats(eidl, n_distinct)
```

```{r distinct_plots, fig.height=2.5}
explore_plot(eidl, assist_type)
explore_plot(eidl, award_desc) + scale_x_truncate(40)
explore_plot(eidl, business_type)
explore_plot(eidl, fund_office)
```

### Amounts

```{r amount_summary}
summary(eidl$amount)
mean(eidl$amount <= 0)
```

Most loan `amounts` are one of a few round integer values.

```{r amount_count}
eidl %>% 
  count(fed_obligation, face_value, sort = TRUE) %>% 
  add_prop() %>% 
  mutate(cp = cumsum(p))
```

```{r hist_amount, echo=FALSE}
eidl %>%
  filter(amount > 500) %>% 
  mutate(is_face = !is.na(face_value)) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "United States EIDL Amount Distribution",
    subtitle = "by Federal Obligation (TRUE) or Face Value (FALSE)",
    caption = "Source: SBA EIDL",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~is_face, scales = "free_y", ncol = 1)
```

```{r amount_map, echo=FALSE}
eidl %>% 
  group_by(state) %>%
  summarise(total = sum(amount)) %>% 
  mutate(across(state, usa::state_convert, to = "name")) %>% 
  left_join(usa::facts, by = c("state" = "name")) %>% 
  mutate(per_capita = total / population) %>% 
  usmap::plot_usmap(regions = "states", data = ., values = "per_capita") +
  scale_fill_viridis_c() +
  theme(legend.position = "right") +
  labs(
    title = "COVID-19 EIDL Loan Amount per capita by State",
    fill = "Amount per capita"
  )
```

### Dates

We can add the calendar year from `action_date` with `lubridate::year()`.

```{r date_year}
eidl <- mutate(eidl, action_year = year(action_date))
```

All of the loans in this database were distributed between April and November.

```{r date_range}
min(eidl$action_date)
max(eidl$action_date)
```

```{r bar_year, echo=FALSE}
eidl %>% 
  count(eweek = epiweek(action_date)) %>% 
  ggplot(aes(x = eweek, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(0, 50, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "2020 COVID-19 EIDL Loans CDC Epidemiological Week",
    caption = "Source: SBA EIDL",
    x = "Week of Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

We will also manually add the disbursing office and governing body.

```{r office_codes}
office_codes <- tribble(
  ~fund_office, ~office,
  "732990", "Office of Disaster Assistance",
  "737610", "Office of Entrepreneurial Development"
)
```

```{r office_join}
eidl <- eidl %>% 
  mutate(govt = "US", agency = "Small Business Administration") %>% 
  left_join(office_codes)
```

```{r office_count}
count(eidl, fund_office, office)
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_file}
addr_file <- path(dirname(raw_dir), "addr.tsv")
```

```{r address_norm}
if (file_exists(addr_file)) {
  norm_addr <- read_tsv(addr_file)
} else {
  norm_addr <- eidl %>%
    select(addr1, addr2) %>% 
    distinct() %>% 
    unite(
      col = addr_full,
      everything(),
      na.rm = TRUE,
      remove = FALSE,
    ) %>% 
    mutate(
      addr_norm = normal_address(
        address = addr_full,
        abbs = usps_street,
        na_rep = TRUE
      )
    ) %>% 
    select(-addr_full)
  write_tsv(distinct(norm_addr), addr_file)
}
```

```{r address_view, echo=FALSE}
sample_n(norm_addr, 10)
```

```{r address_join}
eidl <- left_join(eidl, norm_addr)
```

### ZIP

The business ZIP codes are already split into ZIP and ZIP+4 variables.

```{r zip_progress}
progress_table(
  eidl$zip5,
  compare = valid_zip
)
```

### State

All of the `state` values are already valid.

```{r state_progress}
progress_table(
  eidl$state,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- eidl %>% 
  count(city, state, zip5, sort = TRUE) %>% 
  select(-n) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = "DC",
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c("state", "zip5" = "zip")
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  rename(city = city_raw)
```


#### Progress

```{r city_rejoin}
eidl <- left_join(eidl, norm_city, by = c("city", "state", "zip5"))
```

```{r city_change}
eidl <- mutate(
  .data = eidl,
  city_swap = city_swap %>% 
    str_replace("^NEW YORK CITY$", "NEW YORK")
)
```

```{r city_remain}
many_city <- c(valid_city, extra_city)
eidl %>% 
  filter(city_swap %out% many_city) %>% 
  count(city_swap, state, sort = TRUE) %>% 
  add_prop(sum = TRUE)
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(eidl$city),
  eidl$city_norm,
  eidl$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(eidl$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "EIDL City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "EIDL City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
eidl <- eidl %>% 
  select(
    -city_norm,
    city_clean = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(eidl, 50))
```

1. There are `r comma(nrow(eidl))` records in the database.
1. There are `r comma(sum(eidl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(eidl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "covid", "eidl_covid", "data", "clean"))
clean_path <- path(clean_dir, "us_eidl_covid.csv")
if (nrow(eidl) != nrow_pre) {
  stop("rows have been removed or added")
} else {
  write_csv(eidl, clean_path, na = "")
}
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
