---
title: "Idaho Expenditures"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("id", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `zip`
1. Create a `year` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  snakecase, # column naming
  lubridate, # datetime strings
  jsonlite, # from json data
  magrittr, # pipe opperators
  gluedown, # printing markdown
  janitor, # dataframe clean
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  rvest, # read html pages
  vroom, # read files fast
  glue, # combine strings
  httr, # http requests
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][repo] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[repo]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

This data is obtained from the Idaho [Secretary of State (SoS)][sos]. The data is provided for the years 1994 through 2018. 
For data after 2018, we can download them from the new data portal.

[sos]: https://sos.idaho.gov/elect/finance/downloads.html
[data]: https://miboecfr.nictusa.com/cfr/dumpall/cfrdetail/

## Import

To import the data for processing, we will have to download each archive file
and read them together into a single data frame object.

### Download

We will scrape the download page for every archive link, then downloading each
to a local directory.

```{r download_raw}
raw_dir <- dir_create(here("state","id", "expends", "data", "raw"))
raw_base <- "https://sos.idaho.gov/elect/finance/downloads.html"
raw_page <- read_html(raw_base)
raw_urls <- raw_page %>% 
  html_node("table") %>% 
  html_nodes("a") %>% 
  html_attr("href") %>% 
  str_subset("exp") %>%
  str_subset("^2") %>% 
  str_c(dirname(raw_base), ., sep = "/") %>% 
  rev()
raw_paths <- path(raw_dir, basename(str_replace(raw_urls, "(?<=\\d)/", "_")))
if (!all(file_exists(raw_dir))) {
  download.file(raw_urls, raw_paths)
}
```

### Read

Each file has a different column order and names. We will first use 
`purrr::map()` to use `readxl::read_excel()` and create a list of data frames.

```{r read_excel}
ide <- map(
  .x = raw_paths,
  .f = read_excel,
  col_types = "text"
)
```

Since the variety of column names is so great, we will have to chain together
a number of `stringr::str_replace()` functions to create some consistency.

```{r rename_fun}
consistent_names <- function(nm) {
  nm %>% 
    to_snake_case() %>% 
    str_replace("recipient_", "recip_") %>% 
    str_replace("contrib_", "recip_") %>% 
    str_replace_all("money", "amount") %>% 
    str_replace("recip_name", "recip_last") %>% 
    str_replace("^name$", "committee") %>% 
    str_replace("^first_name$", "cand_first") %>% 
    str_replace("^last_name$", "cand_last") %>% 
    str_replace("^middle_name$", "cand_mi") %>% 
    str_replace("suf$", "suffix") %>% 
    str_replace("_mid$", "_mi") %>% 
    str_replace("middle", "mi") %>% 
    str_replace("cand_party", "party") %>% 
    # str_replace("_cp$", "_type") %>% 
    str_remove("_name$") %>% 
    str_replace("zipcode", "zip") %>% 
    str_replace("_st$", "_state") %>% 
    str_replace("mailing", "address") %>% 
    str_replace("line_1", "address_1") %>% 
    str_replace("line_2", "address_2") %>% 
    str_replace("explanation", "description") %>% 
    str_remove("^recip_") %>% 
    # str_remove("^contributing_") %>% 
    str_remove("^exp_") %>% 
    str_remove("^expen_") %>% 
    str_remove("^expense_") %>% 
    str_remove("^expenditure_") %>% 
    str_replace("^address$", "address_1") %>% 
    str_replace("^election_type$", "election") %>% 
    str_replace("_one", "_1") %>% 
    str_replace("_two", "_2") %>% 
    str_replace("_three", "_3") %>% 
    str_replace("^committee_company$", "company") %>% 
    str_replace("^company_committee$", "company")
}
```

```{r rename_all}
ide <- map(ide, rename_with, consistent_names)
```

Now that each individual data frame has similar column names, we can use
`dplyr::bind_rows()` to bind all `r length(ide)` data frames together.

```{r bind_rows}
names(ide) <- basename(raw_paths)
ide <- bind_rows(ide, .id = "source_file")
ide <- relocate(ide, source_file, .after = last_col())
```

Then, we can use `readr::type_convert()` to parse our character columns.

```{r parse_double}
ide$amount <- parse_double(ide$amount)
```

```{r purpose_fix}
ide <- ide %>% 
  unite(
    col = purpose_codes2,
    matches("purpose_code_\\d$"),
    sep = "|",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  unite(
    col = purpose2,
    matches("purpose_code_\\d_description$"),
    sep = "|",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  mutate(across(contains("purpose"), na_if, "")) %>% 
  mutate(
    purpose_codes = coalesce(purpose_codes, purpose_codes2),
    purpose_desc = coalesce(purpose, purpose2),
    .keep = "unused"
  ) %>% 
  relocate(purpose_desc, .after = purpose_codes)
```

We also need to reorder and recode these variables to be consistent across each
year.

```{r reorder, echo=FALSE}
ide <- ide %>% 
  mutate(
    last = coalesce(last, company)
  ) %>% 
  select(
    committee,
    cand_first,
    cand_mi,
    cand_last,
    cand_suffix,
    party,
    office,
    district,
    method = type,
    amount,
    date,
    type = cp,
    last,
    first,
    mi,
    suffix,
    address_1,
    address_2,
    city,
    state,
    zip,
    country,
    source_file
  )
```

```{r recode}
ide <- ide %>% 
  mutate(across(where(is.character), str_to_upper)) %>% 
  mutate(
    office = office %>% 
      str_replace("REPRESENTATIVE", "REP."),
    cand_suffix = cand_suffix %>% 
      str_remove_all("[:punct:]"),
    party = party %>% 
      str_replace("OTHER", "OTH") %>% 
      str_replace("REPUBLICAN", "REP") %>% 
      str_replace("DEMOCRATIC", "DEM") %>% 
      str_replace("INDEPENDENT", "IND") %>% 
      str_replace("LIBERTARIAN", "LIB") %>% 
      str_replace("CONSTITUTION", "CON") %>% 
      str_replace("NON-PARTISAN", "NON"),
    method = method %>% 
      str_replace("^I$", "IN KIND") %>% 
      str_replace("^R$", "REPAYMENT")
  )
```

```{r check}
count(ide, type)
```

```{r dates}
count_na(ide$date) # 452
slash_dates <- str_which(ide$date, "\\d+/\\d+/\\d{4}")
ide$date[slash_dates] <- as.character(mdy(ide$date[slash_dates]))
excel_dates <- str_which(ide$date, "[:punct:]", negate = TRUE)
ide$date[excel_dates] <- ide$date[excel_dates] %>% 
  as.numeric() %>% 
  excel_numeric_to_date() %>% 
  as.character()

ide$date <- as_date(ide$date)
count_na(ide$date) # 457
```

### 2020

The Excel files only cover data up until 2018. For newer data, we need to use
the new API. We can request all available, 1000 rows at a time.

```{r}
body <- list(
  dates = c("2020-01-01T05:00:00.000Z", "2022-03-17T04:00:00.000Z"), 
  activity_types = list(donate = FALSE, spend = TRUE, file = FALSE), 
  amounts = c(1, 200000), 
  campaign_types = list(Candidate = TRUE, PAC = TRUE),
  ch = "summary", 
  contest_years = c(2020, 2022), 
  order = "date desc", 
  page = 1, 
  limit = 1000
)
```

```{r new_data}
new_rds <- here("id", "expends", "data", "new_data.rds")
if (!file_exists(new_rds)) {
  new <- tibble()
  more_pg <- TRUE
  pg <- 1
  while (more_pg) {
    message(pg)
    # fetch data ----
    body$page <- pg
    a <- POST(
      url = "https://sunshine.sos.idaho.gov/api/es/finances/get_activity.json",
      body = toJSON(body, auto_unbox = TRUE)
    )
    b <- content(a, simplifyDataFrame = TRUE)
    c <- as_tibble(b$output)
    # check progress ----
    new <- bind_rows(new, c)
    more_pg <- (nrow(new)) < b$output_meta$count_total_rows
    pg <- pg + 1
  }
  write_rds(new, new_rds)
} else {
  new <- read_rds(new_rds)
}
```

We need to select and rename all of the columns that match between our new data
and the old Excel files.

```{r new_rename}
new <- new %>% 
  select(
    committee = from_display_name,
    party = from_party_code,
    office = from_office_name,
    district = from_district_name,
    method = spend_type,
    amount,
    date,
    type = from_entity_type,
    last = to_display_name,
    address_1 = to_address,
    city = to_city,
    state = to_state,
    zip = to_zip,
  ) %>% 
  mutate(
    across(everything(), ~na_if(str_squish(str_to_upper(.)), "")),
    across(amount, parse_double),
    across(date, as_date),
    source_file = "get_activity.json"
  )
```

```{r new_recode}
new <- new %>% 
  mutate(
    party = party %>% 
      recode(
        "REPUBLICAN" = "REP",
        "DEMOCRATIC" = "DEM",
        "INDEPENDENT" = "IND",
        "LIBERTARIAN" = "LIB",
        "CONSTITUTION" = "CON",
        "UNAFFILIATED" = "NON"
      )
  )
```

```{r new_join}
setdiff(names(new), names(ide))
ide <- bind_rows(ide, new)
```

## Explore

```{r glimpse}
glimpse(ide)
head(ide)
tail(ide)
```

We should first identify which columns are missing the kinds of key information
we need to properly identify all parties to a contribution. We can do this
with `campfin::flag_na()` after creating a new

```{r count_na}
col_stats(ide, count_na)
```

```{r flag_na}
ide <- ide %>% 
  # combine cand and comm names in new col
  mutate(recip = coalesce(cand_last, committee)) %>% 
  flag_na(last, recip, date, amount)

sum(ide$na_flag)
mean(ide$na_flag)
```

```{r}
ide <- select(ide, -recip)
```

Records that are entirely duplicated at least once across all columns should
also be identified with `campfin::flag_dupes()`. The first occurrence of the
record is not flagged, but all subsequent duplicates are. Not all these records
are true duplicates, since it is technically possible to make the same
contribution to the same person on the same day for the same amount.

```{r flag_dupes}
ide <- flag_dupes(ide, everything(), .check = TRUE)
sum(ide$dupe_flag)
mean(ide$dupe_flag)
ide %>% 
  filter(dupe_flag) %>% 
  count(committee, last, date, amount, sort = TRUE)
```

### Categorical

```{r n_distinct}
col_stats(ide, n_distinct)
```

```{r bar_office, echo=FALSE}
explore_plot(
  data = ide,
  var = office
) + 
  theme(axis.text.x = element_text(angle = 10, vjust = 0.7)) +
  labs(
    title = "Idaho Expenditures by Office",
    caption = "Source: Idaho Elections Office"
  )
```

```{r bar_party, echo=FALSE}
explore_plot(
  data = ide,
  var = party
) +
  labs(
  title = "Idaho Expenditures by Office",
  caption = "Source: Idaho Elections Office"
  )
```

```{r type_party, echo=FALSE}
explore_plot(
  data = filter(ide, !is.na(type)),
  var = type
) +
  labs(
    title = "Idaho Expenditures by Office",
    caption = "Source: Idaho Elections Office"
  )
```

### Amounts

```{r amount_summary}
summary(ide$amount)
mean(ide$amount <= 0, na.rm = TRUE)
```

```{r amount_minmax}
glimpse(ide[c(which.min(ide$amount), which.max(ide$amount)), ])
```

```{r amount_histogram, echo=FALSE}
ide %>%
  filter(amount > 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Idaho Expenditures Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: Idaho Elections Office",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_violin, echo=FALSE}
ide %>%
  filter(
    amount > 1,
    amount < 1e6,
    party %in% most_common(ide$party, 4)
  ) %>% 
  ggplot(aes(x = party, y = amount)) +
  geom_violin(aes(fill = party), adjust = 2) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Idaho Expenditures Amount Distribution",
    subtitle = "from 2000 to 2019, by Party",
    caption = "Source: Idaho Elections Office",
    x = "Amount",
    y = "Count"
  )
```

### Dates

```{r year_add}
ide <- mutate(ide, year = year(date))
```

```{r date_range}
min(ide$date, na.rm = TRUE)
max(ide$date, na.rm = TRUE)
ide <- mutate(ide, date_flag = date > today() | year < 1999 | is.na(date))
count_na(ide$date) # 457
sum(ide$date_flag) # 489 = 32
mean(ide$date_flag)
```

```{r date_replace}
x <- ide$date[ide$date_flag & !is.na(ide$date)]
x <- str_replace(x, "^202(?=[^2])", "201")
x <- str_replace(x, "^19([^9])", "199")
x <- str_replace(x, "^2([^2])", "20")
x <- str_replace(x, "2061", "2016")
x[which(x > today() | year(x) < 1999)] <- NA
ide$date[ide$date_flag & !is.na(ide$date)] <- as_date(x)
ide <- mutate(
  .data = ide,
  date_flag = date > today() | year < 1999 | is.na(date),
  year = year(date)
)
count_na(ide$date) # 484
sum(ide$date_flag) # 484
```

```{r}
ide %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Idaho Expenditures by Year",
    caption = "Source: Idaho Elections Office",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
addr_norm <- ide %>% 
  distinct(address_1, address_2) %>% 
  unite(
    col = address_full,
    starts_with("address_"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address-view}
addr_norm
```

```{r address-join}
ide <- left_join(ide, addr_norm, by = c("address_1", "address_2"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
ide <- ide %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  ide$zip,
  ide$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state-norm}
ide <- ide %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state-view}
ide %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state-progress}
progress_table(
  ide$state,
  ide$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- ide %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("ID", "DC", "IDAHO"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
ide <- left_join(
  x = ide,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- ide %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
ide <- ide %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ide$city),
  ide$city_norm,
  ide$city_swap,
  ide$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(ide$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Idaho City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Idaho City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
ide <- ide %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(ide, 1000))
```

1. There are `r comma(nrow(ide))` records in the database.
1. There are `r comma(sum(ide$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ide$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(ide$date, na.rm = TRUE), "-")
max_dt <- str_remove_all(max(ide$date, na.rm = TRUE), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("id", "expends", "data", "clean"))
clean_csv <- path(clean_dir, glue("id_expends_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(ide, clean_csv, na = "")
write_rds(ide, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
