---
title: "Arkansas Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browsing
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & database
  vroom, # ready many files
  knitr, # knit documents
  here, # relative storage
  fs # search storage 
)
```

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_campfin}
pacman::p_load_current_gh("kiernann/campfin")
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the Arkansas Secretary of State's [Financial Disclosure portal][03].

> This page provides comma separated value (CSV) downloads of contribution, expenditure, and loan
data for each reporting year in a zipped file format. These files can be downloaded and imported
into other applications (Microsoft Excel, Microsoft Access, etc.) This data is extracted from the
Arkansas Campaign Finance database as it existed as of 07/29/2019 11:11 AM.

The AK SOS also provides a [data key][04].

[03]: https://financial-disclosures.sos.arkansas.gov/index.html#/dataDownload
[04]: https://financial-disclosures.sos.arkansas.gov//CFISAR_Service/Template/KeyDownloads/Expenditures,%20Debts,%20and%20Payments%20to%20Workers%20File%20Layout%20Key.pdf

### Download

We can only download the data by navigating to the website and manually clicking the "Download
File" button next to each expenditures file. We can automate this process somewhat with the 
RSelenium package.

```{r raw_dir}
raw_dir <- here("ar", "expends", "data", "raw")
dir_create(raw_dir)
```

```{r remote_download, eval=FALSE}
# open the driver with auto download options
remote_driver <- rsDriver(
  port = 4444L,
  browser = "firefox",
  extraCapabilities = makeFirefoxProfile(
    list(
      browser.download.dir = raw_dir,
      browser.download.folderList = 2L,
      browser.helperApps.neverAsk.saveToDisk = "text/csv"
    )
  )
)

# navigate to the FL DOE download site
remote_browser <- remote_driver$client
remote_browser$navigate("https://financial-disclosures.sos.arkansas.gov/index.html#/dataDownload")

link_2019 <- "tbody.md-body:nth-child(4) > tr:nth-child(1) > td:nth-child(3) > a:nth-child(1)"
link_2018 <- "tbody.md-body:nth-child(5) > tr:nth-child(1) > td:nth-child(3) > a:nth-child(1)"
link_2017 <- "tbody.md-body:nth-child(8) > tr:nth-child(1) > td:nth-child(3) > a:nth-child(1)"

remote_browser$findElement("css", link_2019)$clickElement()
remote_browser$findElement("css", link_2018)$clickElement()
remote_browser$findElement("css", link_2017)$clickElement()

# close the browser and driver
remote_browser$close()
remote_driver$server$stop()
```

### Read

```{r}
ar <- 
  vroom(
    file = dir_ls(raw_dir),
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    col_types = cols(.default = "c")
  ) %>% 
  clean_names() %>%
  mutate_at(
    vars(expenditure_amount), 
    parse_double
  ) %>% 
  mutate_at(
    vars(expenditure_date, filed_date),
    parse_date, format = "%m/%d/%Y %H:%M:%S %p"
  )
```

## Explore

```{r}
head(ar)
tail(ar)
glimpse(sample_frac(ar))
```

