---
title: "Indiana Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("in", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts can be exported from the [State of Indiana Public Contract Search][s]
found on the [Indiana Transparency Portal][t]. The portal limits the number of
returned search results to 1,000 so queries must be split from January through
June and from July through December. Search results can then be exported as
Microsoft Excel files.

[s]: https://fs.gmis.in.gov/psp/guest/SUPPLIER/ERP/c/SOI_CUSTOM_APPS.SOI_PUBLIC_CNTRCTS.GBL
[t]: https://www.in.gov/itp/

```{r raw_dir}
raw_dir <- dir_create(here("in", "contracts", "data", "raw"))
raw_paths <- dir_ls(raw_dir)
```

## Read

The Excel files can then all be read into a single data frame using the
`readxl::read_excel()` and `purrr::map_df()` functions.

```{r raw_read}
inc <- map_df(
  .x = dir_ls("~/Downloads/in_contracts/"),
  .f = readxl::read_excel,
  .name_repair = make_clean_names
)
```

The date columns must be parsed after the fact. We can also remove empty and
useless columns.

```{r raw_parse}
inc <- inc %>% 
  mutate(across(ends_with("date"), parse_date, "%m/%d/%Y")) %>% 
  remove_empty("cols") %>% 
  remove_constant()
```

## Explore

```{r glimpse}
glimpse(inc)
tail(inc)
```

### Missing

```{r na_count}
col_stats(inc, count_na)
```

```{r na_flag}
inc <- inc %>% flag_na(start_date, agency, amount, vendor_name)
sum(inc$na_flag)
```

```{r na_view}
inc %>% 
  filter(na_flag) %>% 
  select(start_date, agency, amount, vendor_name)
```

### Duplicates

```{r dupe_flag}
inc <- flag_dupes(inc, everything())
sum(inc$dupe_flag)
```

```{r dupe_view}
inc %>% 
  filter(dupe_flag) %>% 
  select(start_date, agency, amount, vendor_name)
```

### Categorical

```{r n_distinct}
col_stats(inc, n_distinct)
add_prop(count(inc, action_type, sort = TRUE))
```

### Continuous

#### Amounts

```{r ammount_summary}
noquote(map_chr(summary(inc$amount), dollar))
mean(inc$amount <= 0)
```

```{r hist_amount, echo=FALSE}
inc %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Indiana Contracts Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
inc <- mutate(inc, start_year = year(start_date))
```

```{r date_range}
min(inc$start_date)
sum(inc$start_year < 2000)
max(inc$start_date)
sum(inc$start_date > today())
```

```{r}
add_prop(count(inc, start_year))
```

```{r bar_year, echo=FALSE}
inc %>% 
  count(start_year) %>% 
  mutate(even = is_even(start_year)) %>% 
  ggplot(aes(x = start_year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  coord_cartesian(xlim = c(2015, 2020)) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Indiana Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

The only geographic variable in need of cleaning is the ZIP code which contains
a handful of ZIP+4 codes, which we can trim to create 5-digit ZIP codes.

```{r zip_norm}
inc <- mutate(inc, across(zip_code, normal_zip))
prop_in(inc$zip_code, valid_zip)
```

## Conclude

1. There are `r comma(nrow(inc))` records in the database.
1. There are `r comma(sum(inc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(inc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `start_year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("in", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "in_contracts_clean.csv")
write_csv(inc, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, basename))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(inc)),
  type = md_code(map_chr(inc, typeof)),
  def = c(
    "Contract amount",
    "Unique contract ID",
    "Contract amendment number",
    "Contract action type (New, Amendment, Renewal)",
    "Contract start date",
    "Contract end date",
    "Spending agency name",
    "Recieving vendor name",
    "Vendor 5-digit ZIP code",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year contract started"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Indiana Contracts Data Dictionary\n", dict_md),
  path = here("in", "contracts", "in_contracts_dict.md"),
)
```
