---
title: "Maine Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("me", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  snakecase, # string convert
  gluedown, # print markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  batman, # parse logical
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # read html pages
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

```{r include=FALSE}
filter <- dplyr::filter
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

## Import

### Download

```{r raw_dir}
raw_dir <- dir_create(here("me", "contribs", "data", "raw"))
raw_info <- as_tibble(dir_info(raw_dir))
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```

### Read

#### Fix

```{r fix_dir}
fix_dir <- dir_create(path(dirname(raw_dir), "fix"))
fix_eval <- length(dir_ls(fix_dir)) != nrow(raw_info)
```

```{r fix_loop_old, eval=fix_eval}
# for old format files
for (f in raw_info$path[1:10]) {
  n <- path(fix_dir, str_c("FIX", basename(f), sep = "_"))
  x <- read_lines(f, skip = 1)
  for (i in rev(seq_along(x))) {
    y <- i - 1
    if (y == 0) {
      next() # skip first
    } else if (str_starts(x[i], "\"\\d+\",") | str_ends(x[y], "\"(Y|N)\"")) {
      next() # skip if good
    } else { # merge if bad
      x[y] <- str_c(x[y], x[i])
      x <- x[-i] # remove bad
    }
  }
  x <- str_remove(x, '(?<=")"(?!,)')
  write_lines(x, n)
  message(basename(n))
}
```

```{r fix_loop_new, eval=fix_eval}
# new format files
for (f in raw_info$path[11:length(raw_info$path)]) {
  n <- path(fix_dir, str_c("FIX", basename(f), sep = "_"))
  x <- read_lines(f, skip = 1)
  for (i in rev(seq_along(x))) {
    if (str_starts(x[i], "\\d+,\\d+,")) {
      next() # skip if good
    } else { # merge if bad
      x[i - 1] <- str_c(x[i - 1], x[i])
      x <- x[-i] # remove bad
    }
  }
  write_lines(x, n)
  message(basename(n))
}
```

```{r fix_info}
fix_info <- as_tibble(dir_info(fix_dir))
sum(fix_info$size)
fix_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```

#### Compare

```{r read_names}
old_names <- read_names(path(raw_dir, "CON_2008.csv.csv"))
new_names <- read_names(path(raw_dir, "CON_2019.csv.csv"))
```

The files come in two structures. For files from 2008 to 2017, there are `r
length(me08)` variables. For the newer files, 2018 and 2019, there are `r
length(me19)` variables.

```{r print_names, echo=FALSE, collapse=TRUE}
print(old_names)
print(new_names)
```

```{r names_in}
old_names <- old_names %>% 
  str_replace("^MI$", "Middle Name") %>% 
  str_replace("\\bInfo\\b", "Information")
to_snake_case(old_names) %in% to_snake_case(new_names)
```

We can read each type of file into a separate data frame and then combine the
two. Any of the new variables. There are 4 rows with a field containing double
quoation marks. This causes `readr::read_delim()` to incorectly shift values
right one column. We have no choice but to filter out these 4 records or
manually edit the text file.

```{r raw_read_old}
me_old_format <- fix_info$path[1:10] %>% 
  map(
    read_delim,
    delim = ",",
    escape_backslash = FALSE,
    escape_double = FALSE,
    col_names = old_names,
    col_types = cols(
      .default = col_character(),
      ReceiptAmount = col_number(),
      ReceiptDate = col_date_usa(),
      FiledDate = col_date_usa()
    )
  ) %>% 
  bind_rows(.id = "id") %>% 
  clean_names() %>% 
  remove_empty("cols") %>% 
  left_join(
    tibble(
      id = as.character(1:10), 
      file = basename(fix_info$path[1:10])
    )
  ) %>% 
  select(-id)
```

```{r raw_fix_null}
x <- read_lines(fix_info$path[13])
write_lines(x[-344], fix_info$path[13])
```

```{r raw_read_new}
me_new_format <- fix_info$path[11:13] %>% 
  map(
    read_delim,
    delim = ",",
    escape_backslash = FALSE,
    escape_double = TRUE,
    col_names = new_names,
    col_types = cols(
      .default = col_character(),
      `Receipt Amount` = col_number(),
      `Receipt Date` = col_date_usa(),
      `Filed Date` = col_date_usa()
    ) 
  ) %>% 
  bind_rows(.id = "id") %>% 
  clean_names() %>% 
  remove_empty("cols") %>% 
  left_join(
    tibble(
      id = as.character(1:3), 
      file = basename(fix_info$path[11:13])
    )
  ) %>% 
  select(-id)
```

```{r raw_bind}
mec <- 
  bind_rows(me_old_format, me_new_format) %>% 
  rename(emp_info_req = employment_information_requested) %>% 
  rename_all(str_remove, "receipt_") %>% 
  rename_all(str_remove, "_name") %>% 
  mutate_at(vars(emp_info_req, amended), to_logical) %>% 
  mutate_at(vars(file), basename) %>% 
  filter(!is.na(amended))
```

```{r raw_check}
count(mec, emp_info_req)
```

## Explore

```{r glimpse}
glimpse(mec)
tail(mec)
```

### Missing

```{r na_count}
mec_missing <- col_stats(mec, count_na)
```

Recipients are divided into committees and candidates. To better flag records
missing _either_ type, we will `coalesce()` the two into a single variable. We 
can also `unite()` the four contributor name columns.

```{r na_unite}
mec <- mec %>% 
  mutate(recipient = coalesce(candidate, committee)) %>% 
  unite(
    col = contributor,
    first, middle, last, suffix,
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  mutate(across(is.character, na_if, ""))
```

After uniting and coalescing the contributor and recipient columns, only
`r sum(mec$na_flag)` records are missing a name, date, or ammount.

```{r na_flag}
mec <- mec %>% flag_na(date, contributor, amount, recipient)
sum(mec$na_flag)
```

### Duplicates

If we ignore the supposedly (quasi) unique `id` variable, there are a number of
otherwise completely duplicated records. We can flag them with 
`campfin::flag_na()`

```{r dupe_flag, warning=TRUE}
mec <- flag_dupes(mec, -id)
percent(mean(mec$dupe_flag), 0.01)
```

```{r dupe_view}
mec %>% 
  filter(dupe_flag) %>% 
  select(date, contributor, amount, recipient)
```

### Categorical

```{r n_distinct}
col_stats(mec, n_distinct)
```

```{r}
explore_plot(mec, committee_type)
explore_plot(mec, office)
explore_plot(mec, forgiven_loan) + 
  scale_x_discrete(label = function(x) str_trunc(x, 20)) +
  labs(caption = paste(percent(prop_na(mec$forgiven_loan), 0.001), "NA"))
explore_plot(mec, election_type)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(mec$amount)
percent(mean(mec$amount <= 0, na.rm = TRUE), 0.01)
```

```{r hist_amount, echo=FALSE}
mec %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Maine Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

We can add the calendar year from the `date` column with `lubridate::year()`.

```{r date_year}
mec <- mutate(mec, year = year(date))
```

The new `year` and `date` columns are very clean.

```{r date_range}
min(mec$date)
sum(mec$year < 2000)
max(mec$date)
sum(mec$date > today())
```

```{r bar_year, echo=FALSE}
mec %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maine Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
mec <- mec %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
mec %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mec <- mec %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mec$zip,
  mec$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
mec <- mec %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
mec %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  mec$state,
  mec$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mec <- mec %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("ME", "DC", "MAINE"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
mec <- mec %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- mec %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
nrow(good_refine)
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
mec <- mec %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_remain}
many_city <- c(valid_city, extra_city)
mec %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, state_norm, sort = TRUE)
```

```{r city_add}
many_city <- c(many_city, "ARROWSIC", "WEST BATH", "VEAZIE")
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(mec$city_raw),
  mec$city_norm,
  mec$city_swap,
  mec$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(mec$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Maine City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maine City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
mec <- mec %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(mec, 20))
```

1. There are `r comma(nrow(mec))` records in the database.
1. There are `r comma(sum(mec$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mec$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("me", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "me_contribs_clean.csv")
write_csv(mec, clean_path, na = "")
file_size(clean_path)
guess_encoding(clean_path)
```

```{r clean_encode}
enc <- system2("file", args = paste("-i", clean_path), stdout = TRUE)
str_replace_all(enc, clean_path, basename)
```

## Dictionary

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(mec)),
  type = md_code(map_chr(mec, typeof)),
  # old = c(md_code(old_names), rep("", length(mec) - length(old_names))),
  # new = c(md_code(new_names), rep("", length(mec) - length(new_names))),
  def = c(
    "Recipient unique ID",
    "Contribution amount",
    "Date contribution was made",
    "Contributor full name",
    "Contributor first name",
    "Contributor middle name",
    "Contributor last name",
    "Contributor name suffix",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor 2-digit state abbreviation",
    "Contributor ZIP+4 code",
    "Contribution unique ID",
    "Date contribution filed",
    "Contribution type",
    "Contribution source",
    "Recipient committee type",
    "Recipient commttee name",
    "Recipient candidate name",
    "Contribution amended",
    "Contribution description",
    "Contributor employer name",
    "Contributor occupation",
    "Occupation comment",
    "Employer information requested",
    "Source file name",
    "Legacy recipient ID",
    "Recipient office sought",
    "Recipient district election",
    "Report contribution listed on",
    "Forgiven loan reason",
    "Election type",
    "Combined type recipient name",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Maine Contributions Data Dictionary\n", dict_md),
  path = here("me", "contribs", "me_contribs_dict.md"),
)
```
