---
title: "Vermont Parcels"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("vt", "parcels", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  jsonlite, # parse JSON data
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

```{r echo=FALSE}
options(knitr.kable.NA = "")
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("vt/parcels/docs/vt_parcels_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is from the [Vermont Parcel Program][vpp] run by the 
[Vermont Center for Geographic Information][vcgi].

> Parcel mapping in Vermont is undertaken and overseen by each municipality, who
may perform annual map updates on their own or hire vendors to maintain. As part
of the Parcel Program, VCGI coordinates the collection of the best available
digital parcel data from municipalities and publishes them in a uniform GIS
format joined with the Tax Department’s [annual grand list][agl]. VCGI also
develops the [data standards][ds], [templates][tem], [digital parcel mapping
guidelines][pmg], and guidance resources by which Vermont digital GIS parcel
data are organized and made publicly available.
> 
> The Parcel Program is an outgrowth of the [Statewide Property Mapping
Project][spmp] which included a three-year (2017-2019) funded effort to create
or update digital GIS parcel data for all of Vermont’s municipalities. The
Project successfully completed that work in December 2019, resulting in a freely
available GIS layer for all Vermont municipalities joined with their [annual
grand list][agl] via a parcel’s unique [School Property Account Number
(SPAN)][span]. The parcel Program now supports the ongoing collection and
publishing of municipal parcel data in GIS format statewide, continuing to use
the SPAN number as a unique parcel identifier.

[vpp]: https://vcgi.vermont.gov/data-and-programs/parcel-program
[vcgi]: https://vcgi.vermont.gov/
[agl]: https://tax.vermont.gov/municipal-officials/data-request
[ds]: https://vcgi.vermont.gov/document/vt-gis-parcel-data-standard
[tem]: https://vcgi.vermont.gov/document/geodatabase-template-parcel-data-standard
[pmg]: https://vcgi.vermont.gov/document/vermont-gis-parcel-mapping-guideline
[spmp]: https://legislature.vermont.gov/statutes/section/19/001/00044
[span]: https://tax.vermont.gov/home/tax-learning-center/glossary#s

### Dictionary

```{r echo=FALSE}
vt_meta <- fromJSON("https://opendata.arcgis.com/api/v3/datasets/09cf47e1cf82465e99164762a04f3ce6_0")
vt_dict <- as_tibble(vt_meta$data$attributes$fields)
vt_dict %>% 
  select(name, alias, type, length) %>% 
  mutate(
    across(name, md_code),
    across(type, str_remove, "^esriFieldType")
  ) %>% 
  kable()
```

## Download

The parcel list can be downloaded directly from the Vermont Open Geodata Portal.

```{r raw-head}
id <- "09cf47e1cf82465e99164762a04f3ce6_0"
vt_head <- HEAD(
  url = glue("https://opendata.arcgis.com/api/v3/datasets/{id}/downloads/data"),
  query = list(
    format = "csv",
    spatialRefId = 32145
  )
)
```

```{r raw-name}
vt_info <- headers(vt_head)
vt_name <- str_extract(vt_info[["content-disposition"]], '(?<=").*?(?=")')
```

```{r raw-dir}
raw_dir <- dir_create(here("vt", "parcels", "data", "raw"))
raw_csv <- path(raw_dir, vt_name)
```

```{r raw-download}
if (!file_exists(raw_csv)) {
  GET(
    url = vt_head$url,
    write_disk(raw_csv), 
    progress("down")
  )
}
```

## Read

The tabular parcel list can be read as a data frame from the CSV file.

```{r raw-read}
vtp <- read_delim(
  file = raw_csv,
  delim = ",",
  guess_max = 1e6,
  escape_backslash = FALSE,
  escape_double = FALSE
)
```

```{r clean-names}
vtp <- clean_names(vtp, case = "snake")
```

## Explore

There are `r comma(nrow(vtp))` rows of `r ncol(vtp)` columns. Each record
represents a single property parcel in the state of Vermont.

```{r glimpse}
glimpse(vtp)
tail(vtp)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(vtp, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("real_flv", "year", "addrgl1", "owner1")
vtp <- flag_na(vtp, all_of(key_vars))
mean(vtp$na_flag)
```

```{r na-view}
vtp %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
vtp <- flag_dupes(vtp, -objectid)
sum(vtp$dupe_flag)
```

```{r dupe-view}
vtp %>% 
  filter(dupe_flag) %>% 
  select(objectid, all_of(key_vars))
```

### Categorical

```{r distinct-count}
col_stats(vtp, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(vtp, proptype)
explore_plot(vtp, town)
explore_plot(vtp, sourcetype)
explore_plot(vtp, editmethod)
explore_plot(vtp, matchstat)
explore_plot(vtp, cat)
explore_plot(vtp, rescode)
explore_plot(vtp, equipcode)
explore_plot(vtp, hsdecl)
explore_plot(vtp, expdesc)
explore_plot(vtp, statute)
```

### Amounts

```{r amount-round}
vtp$real_flv <- round(vtp$real_flv, digits = 2)
```

```{r amount-summary}
summary(vtp$real_flv)
mean(vtp$real_flv <= 0, na.rm = TRUE)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(vtp[c(which.max(vtp$real_flv), which.min(vtp$real_flv)), ])
```

```{r hist-amount, echo=FALSE}
vtp %>%
  filter(real_flv >= 100, real_flv <= 1e7) %>% 
  ggplot(aes(real_flv)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:8)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Vermont Parcels Amount Distribution",
    caption = "Source: VCGI",
    x = "Amount",
    y = "Count"
  )
```

### Dates

There is an existing `year` variable.

```{r date-range}
min(vtp$year)
sum(vtp$year < 2000)
max(vtp$year)
sum(vtp$year > year(today()))
```

```{r bar-year, echo=FALSE}
vtp %>% 
  count(year) %>% 
  filter(year > 2009) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col() + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2017:2021) +
  theme(legend.position = "bottom") +
  labs(
    title = "Vermont Parcels by Year",
    caption = "Source: VCGI",
    x = "GIS Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

The variables being cleaned in this data refer mostly to the mailing address of
the property owner. The address of the parcel itself is only found in the
`e911addr` variable. The `town` variable similarly refers to the legal town
locality of the property, with the `citygl` variable referring to the city of
the owner's mailing address, etc.

### Address

The address variables are already sufficiently normalized.

```{r echo=FALSE}
vtp %>% 
  sample_n(10) %>% 
  distinct(addrgl1, addrgl2)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
vtp <- vtp %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zipgl,
      na_rep = TRUE
    )
  )
```

```{r zip-progress, echo=FALSE}
progress_table(
  vtp$zipgl,
  vtp$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state-norm}
vtp <- vtp %>% 
  mutate(
    state_norm = stgl %>%
      str_remove_all("\\.") %>% 
      str_replace("^CANADA$", "CN") %>% 
      str_replace("^QUEBEC$", "QC") %>% 
      str_replace("^QUE?(\\s+)?CANADA$", "QC") %>% 
      str_replace("^N Y$", "NY") %>% 
      str_remove("\\s+CANADA$") %>% 
      normal_state(abbreviate = TRUE, na_rep = TRUE) %>% 
      str_replace("^(\\w)\\s(\\w)", "\\1\\2")
  )
```

```{r state-view}
vtp %>% 
  filter(stgl != state_norm) %>% 
  count(stgl, state_norm, sort = TRUE)
```

```{r state-progress, echo=FALSE}
progress_table(
  vtp$stgl,
  vtp$state_norm,
  compare = valid_state
)
```

We are also going to manually add the state of each property address, which
should be Vermont in all cases.

```{r state-add}
vtp <- mutate(vtp, state_prop = "VT")
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- vtp %>% 
  distinct(citygl, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = citygl, 
      abbs = usps_city,
      states = c("VT", "DC", "VERMONT"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
vtp <- left_join(
  x = vtp,
  y = norm_city,
  by = c(
    "citygl", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- vtp %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
vtp <- vtp %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r}
many_city <- c(valid_city, extra_city)
```

```{r check_filter}
vtp_out <- vtp %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, state_norm, sort = TRUE) %>% 
  drop_na() %>% 
  head(100)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("vt", "parcels", "data", "city_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file,
    col_types = cols(
      .default = col_character(),
      check_city_flag = col_logical()
    )
  )
} else {
  check <- pmap_dfr(
    .l = list(
      vtp_out$city_refine, 
      vtp_out$state_norm
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    file = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
length(valid_locality)
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  c(valid_locality)

length(valid_locality)
```

```{r}
many_city <- c(many_city, valid_locality)
```

```{r}
many_city <- c(
  many_city, "BROWNINGTON", "WEYBRIDGE", "ST GEORGE", "GEORGIA", "PANTON"
)
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(vtp$citygl),
  vtp$city_norm,
  vtp$city_swap,
  vtp$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(vtp$citygl, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Vermont City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Vermont City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
vtp <- vtp %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(city_clean, state_clean, .before = zip_clean) %>% 
  relocate(state_prop, .before = city_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(vtp, 50))
```

1. There are `r comma(nrow(vtp))` records in the database.
1. There are `r comma(sum(vtp$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(vtp$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean-dir}
clean_dir <- dir_create(here("vt", "parcels", "data", "clean"))
clean_path <- path(clean_dir, "vt_parcels_2017-20211011.csv")
write_csv(vtp, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
