---
title: "Vermont Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("vt", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("vt/voters/docs/vt_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Vermont statewide voter list was obtained via a records request made to the
Secretary of State's Records Office on April 5, 2021. After signing an affidavit
asserting our intentions to distribute the data on a non-commercial basis, the
records request was fulfilled.

The data was provided to Kiernan Nicholls as attachments to an email.

> As requested, enclosed you will find a compressed zip file of the Vermont
Statewide Voter File (as of April 9, 2021) with a printed Excel list of the
fields that are created in text delimited format.
>
> You requested to receive a one-time dispatch of the Vermont Statewide Voter
Checklist. If you have any concerns, please contact me at
lelonie.oatway@vermont.gov.

Two files were attached to the email.

1. `4.9.2021Statewidevoters (81).zip`
2. `2020 Field Listing-Statewide Checklist.xls`

```{r raw_dir}
raw_dir <- dir_create(here("vt", "voters", "data", "raw"))
raw_zip <- path(raw_dir, "4.9.2021Statewidevoters (81).zip")
file_size(raw_zip)
raw_xls <- here("vt", "voters", "2020 Field Listing-Statewide Checklist.xls")
```

## About

```{r echo=FALSE}
xls_about <- read_excel(
  path = raw_xls,
  range = "A4:B43",
  col_names = c("Row", "Column")
)
kable(xls_about)
```

## Extract

The ZIP archive contains one text file of the same name.

```{r echo=FALSE}
unzip(raw_zip, list = TRUE) %>% 
  mutate(
    across(Name, md_code),
    across(Length, fs_bytes)
  ) %>% 
  kable()
```

```{r extract}
raw_txt <- unzip(raw_zip, exdir = raw_dir)
file_size(raw_txt)
```

## Read

```{r raw_read}
vtv <- read_delim(
  file = raw_txt,
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", " ", "NA"),
  col_types = cols(
    .default = col_character(),
    `Year of Birth` = col_integer(),
    `Date of Registration` = col_date_mdy(),
    `Date last Voted` = col_date_mdy()
  )
)
```

```{r raw_rename}
vtv <- vtv %>% 
  # remove empty end col
  remove_empty("cols") %>%
  # rename variables
  clean_names("snake") %>% 
  rename(
    address_1      = legal_address_line_1,
    address_2      = legal_address_line_2,
    city           = legal_address_city,
    state          = legal_address_state,
    zip            = legal_address_zip,
    mail_address_1 = mailing_address_line_1,
    mail_address_2 = mailing_address_line_2,
    mail_city      = mailing_address_city,
    mail_state     = mailing_address_state,
    mail_zip       = mailing_address_zip,
    reg_date       = date_of_registration,
    birth_year     = year_of_birth
  ) %>% 
  # remove voting history
  select(-ends_with("participation"))
```

## Explore

There are `r comma(nrow(vtv))` rows of `r ncol(vtv)` columns. Each record
represents a single voter registered in the state of Vermont as of April 5.

```{r glimpse}
glimpse(vtv)
tail(vtv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(vtv, count_na)
```

No rows are missing the last name or registration date used for identification.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
vtv <- flag_dupes(vtv, -voter_id)
sum(vtv$dupe_flag)
```

Ignoring the supposedly unique `voter_id`, we do find a number of duplicate
voters (born in the same year, living at the same address).

```{r dupe_view}
vtv %>% 
  filter(dupe_flag) %>% 
  select(voter_id, last_name, first_name, address_1, birth_year)
```

### Categorical

```{r distinct_count}
col_stats(vtv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(vtv, status)
explore_plot(vtv, town_of_registration)
explore_plot(vtv, county)
```

### Dates

There are thousands of voters with a registration date of January 1, 1900. These
should be `NA` values, not real dates. Similarly, a few hundred have the date
of January 1st of the year 1910. The same is true for the birth year.

```{r echo=FALSE}
og_date <- vtv$reg_date
```

```{r}
vtv %>% 
  count(
    y = str_sub(reg_date, 1, 3),
    y2 = str_sub(birth_year, 1, 3)
  ) %>% 
  ggplot(aes(y, n)) + 
  geom_col()
```


```{r date_min}
count_na(vtv$reg_date)
min(vtv$reg_date, na.rm = TRUE)
# compare two dates
sum(vtv$reg_date == "1900-01-01", na.rm = TRUE)
sum(vtv$reg_date == "1900-01-02", na.rm = TRUE)
```

```{r dates_na}
# bad dates
vtv$reg_date <- na_if(vtv$reg_date, "1900-01-01")
vtv$reg_date <- na_if(vtv$reg_date, "1901-01-01")
vtv$reg_date <- na_if(vtv$reg_date,   "01-01-01")
vtv$reg_date <- na_if(vtv$reg_date, "1910-01-01")
```

```{r bates_min}
# fix one bad date
vtv$reg_date[which(vtv$reg_date == "998-09-01")] <- "1998-09-01"
min(vtv$reg_date, na.rm = TRUE)
```

```{r}
sum(vtv$birth_year == 1900)
sum(vtv$birth_year == 1901)
vtv$birth_year <- na_if(vtv$birth_year, 1900)
```

```{r date_max}
max(vtv$reg_date, na.rm = TRUE)
```

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
vtv <- mutate(vtv, reg_year = year(reg_date))
```

```{r bar_reg_year, echo=FALSE}
vtv %>% 
  filter(!is.na(reg_year)) %>% 
  count(reg_year) %>% 
  mutate(
    elect_type = case_when(
      reg_year %% 4 == 0 ~ "Pres.",
      reg_year %% 2 == 0 ~ "Mid.",
      TRUE ~ NA_character_
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = elect_type)) + 
  scale_fill_manual(
    values = c("#1B9E77", "#D95F02"),
    na.value = "#666666"
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 4)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina Voters Registered by Year",
    caption = "Source: NC SOS",
    fill = "Election Year",
    x = "Year Registered",
    y = "Count"
  )
```

```{r hist_birth_year, echo=FALSE}
vtv %>% 
  filter(!is.na(birth_year)) %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(
    fill = dark2["purple"],
    binwidth = 3
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2005, by = 5)) +
  coord_cartesian(xlim = c(1900, 2005)) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina Voters by Birth Year",
    caption = "Source: NC SOS",
    x = "Birth Year",
    y = "Count"
  )
```

## Missing

Now that we've removed some invalid dates, we can flag all the rows missing any
name or date.

```{r}
vtv <- vtv %>% 
  mutate(
    na_flag = !complete.cases(first_name, reg_date, birth_year),
    .before = dupe_flag
  )
```

Quite a large percentage of rows were using these blatantly invalid registration
dates or birth years; `r percent(mean(vtv$na_flag))` of all rows are missing a
name or date.

## Wrangle

Almost all the geographic variables are sufficiently clean.

```{r geo_check}
prop_in(vtv$city, c(valid_city, extra_city))
prop_in(vtv$state, valid_state)
prop_in(vtv$zip, valid_zip)
```

The cities that can't be easily identified as valid are mostly legal 
distinctions between the town, city, and village of a jurisdiction. For 
[example][barre]:

> Barre is the most populous city in Washington County, Vermont, United States.
As of the 2010 census, the municipal population was 9,052. Popularly referred to
as "Barre City", it is almost completely surrounded by "Barre Town", which is a
separate municipality.

[barre]: https://en.wikipedia.org/wiki/Barre_(city),_Vermont

```{r city_bad}
vtv %>% 
  filter(city %out% c(valid_city, extra_city)) %>% 
  count(city, sort = TRUE)
```

The street addresses are in the proper USPS abbreviated format.

```{r addr_sample}
sample(vtv$address_1, 10)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(vtv, 50))
```

1. There are `r comma(nrow(vtv))` records in the database.
1. There are `r comma(sum(vtv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(vtv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("vt", "voters", "data", "clean"))
clean_path <- path(clean_dir, "vt_voters_2021-04-09.csv")
write_csv(vtv, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
