---
title: "Arkansas Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ar", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Arizona contracts data can be obtained from the [state transparecy website][tr].

[tr]: https://transparency.arkansas.gov/

> Transparency.Arkansas.gov is a resource for citizens, legislators, news media,
scholars, and nonprofit groups who want to know where the state, their city,
town, and school district get their money, and how they spend it. It was created
as a result of [Act 303](https://transparency.arkansas.gov/assets/pdf/Act303.pdf) 
which was passed in the 2011 General Session requiring the Department of Finance
and Administration to create a free website, accessible to the public that
contains state financial information.

> State Agencies negotiate contracts with vendors to provide services or goods
at set prices. Contract information is available. Choose a link to learn more
about contracts awarded in Arkansas.

> Three types of transactions are displayed on this website:
> 1. Standard contracts in excess of $25,000 and
> 1. Construction contracts in excess of $20,000
> 1. Single purchase orders in excess of $25,000
>
> Data about these contracts is displayed on this website. Actual contracts are
displayed as PDF documents. Some contract data is considered private or
protected under state and federal laws. That data will be summarized to provide
that privacy while also providing the reader with accurate details. Contracts
are updated daily.

## Glossary

There are a few terms in the data which can be found in the [site glossary][gl]:

[gl]: https://transparency.arkansas.gov/glossary.html

> DBA means: Doing Business As.Vendors often have more than one name, so Vendor DBA is used to help identify the correct business.

> The 12-month period beginning July 1 and ending June 30 used by the state government for accounting purposes. Fiscal year designation is based on the year in which it ends [e.g., fiscal year 2013 (FY-2013) runs from July 1, 2012 to June 30, 2013]. 

> A vendor is an individual or company that buys and sells goods or services to the State of Arkansas. 

## Download

The data is downloaded as ZIP archives. They can be downloaded manually on the
[search page][sp] by selecting a fiscal year from the drop down menu.

[sp]: https://www.ark.org/dfa/transparency/contracts.php

```{r raw_dir}
raw_dir <- dir_create(here("ar", "contracts", "data", "raw"))
```

Or, we can use an `httr::GET()` request for each year. This will change the
URL and ask for the appropriate archive from the server.

```{r raw_download}
if (length(dir_ls(raw_dir)) == 0) {
  for (year in 2013:2020) {
    get_path <- path(raw_dir, glue("contracts_overview_{year}.zip"))
    GET(
      url = glue("https://www.ark.org/dfa/transparency_{year}/contracts.php"),
      write_disk(get_path, overwrite = TRUE),
      query = list(
        # need new hash each session
        ina_sec_csrf = "6d2673540b4be801b5b58748b935dd2c",
        `do:export_contracts` = "",
        report = "contracts",
        level = "1",
        # keep as 2011
        fiscal_year = 2011,
        current_tab = "overview",
        type = "csv"
      )
    )
  }
}
```

```{r raw_info}
zip_info <- as_tibble(dir_info(raw_dir))
zip_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, ~as_fs_path(basename(.))))
```

## Extract

Each ZIP archive contains two files:

1. `Main_Contracts.csv`
2. `Purchase_Orders_Against_Contracts.csv`

We will extract each of these files to the `data/raw/` directory and add the
fiscal year to the file name.

```{r ext_loop}
for (z in zip_info$path) {
  # extract csv files
  out_files <- unzip(z, exdir = raw_dir)
  # determine year from zip
  year <- str_extract(z, "\\d{4}")
  # create new file names
  out_new <- str_c(c("purchase_orders-", "main_contracts-"), year, ".csv")
  out_new <- path(raw_dir, out_new)
  # rename with year
  file_move(out_files, out_new)
}
```

## Read

Each file type contains different columns. Some overlap and some do not. We can
read both file types into the same data frame, as the key mapping variables are
present in both.

```{r raw_names}
md_bullet(read_names(out_new[1])) # purchase orders
md_bullet(read_names(out_new[2])) # main contracts
```

```{r raw_paths}
raw_paths <- dir_ls(raw_dir, glob = "*.csv")
md_bullet(md_code(basename(raw_paths)))
```

```{r raw_read}
arc <- map_df(
  .x = raw_paths,
  .f = read_delim,
  .id = "file",
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character(),
    `Fiscal Year` = col_integer(),
    `Amount Ordered` = col_double(),
    `Amount Spent` = col_double()
  )
)
```

The added `file` variable lists the source file, but we will create a second
variable that explicitly describes whether the record is a contract or purchase
order.

```{r raw_source}
arc <- mutate(
  .data = arc,
  type = if_else(
    condition = str_detect(file, "order"), 
    true = "Purchase", 
    false = "Contract")
)
```


If the files were all read properly, there should only be years in the 
`fiscal_year` column.

```{r raw_check}
count(arc, `Fiscal Year`)
```

We have to do some parsing and renaming after the fact to account for slight
differences between the two types of files. Purchase orders have a single
`release_date` variable (when the purchase was made), whereas main contracts 
have both `contract_start_date` and `contract_end_date`; we will have to combine
these two variables into a single date to represent the transaction.

```{r raw_parse, warning=TRUE}
arc <- arc %>% 
  clean_names("snake") %>% 
  mutate(across(ends_with("date"), parse_date)) %>% 
  mutate(across(contract_value, parse_double)) %>% 
  mutate(across(file, basename)) %>% 
  mutate(date = coalesce(release_date, contract_start_date)) %>% 
  unite(
    col = vendor_dba, 
    contains("dba"), 
    sep = " ", 
    na.rm = TRUE, 
    remove = TRUE
  ) %>% 
  mutate(across(vendor_dba, na_if, ""))
```

For purchase orders, almost half the agency names are simply "STATEWIDE 
CONTRACT" but the `po_agency_name` variable contains useful information.

```{r raw_agency_order}
arc %>% 
  filter(str_detect(file, "order")) %>% 
  count(agency, po_agency_name, sort = TRUE) %>% 
  mutate(p = n/sum(n))
```

The `po_agency_name` variable is not present for main contracts, so we will have
to merge these two columns in some kind of meaningful way.

```{r raw_agency_contract}
arc %>% 
  filter(str_detect(file, "contract")) %>% 
  count(agency, po_agency_name, sort = TRUE) %>% 
  mutate(p = n/sum(n))
```

```{r raw_agency_merge}
arc <- arc %>% 
  mutate(
    agency = coalesce(po_agency_name, agency),
    agency_id = coalesce(po_agency_value, agency_id),
  ) %>% 
  select(
    -po_agency_name,
    -po_agency_value
  )
```

## Explore

Combining these two file types results in a structure where matching names are
aligned and missing columns are filled with `NA` values. This is most noticeable
at the `tail()`.

```{r glimpse}
glimpse(arc)
tail(arc)
```

### Missing

If we count the number of missing values per column, we can see a lot of the
values from the columns found only in one type of file are missing.

```{r na_count}
col_stats(arc, count_na)
```

But we can flag those key values that should be present in both file types.

```{r na_flag}
arc <- arc %>% flag_na(date, agency, amount_spent, vendor_name)
percent(mean(arc$na_flag), 0.01)
```

```{r na_view}
arc %>% 
  filter(na_flag) %>% 
  select(date, agency, amount_spent, vendor_name, type) %>% 
  distinct()
```

### Duplicates

There are no duplicate records in this database.

```{r dupe_flag, warning=TRUE}
arc <- flag_dupes(arc, everything())
```

### Categorical

```{r n_distinct}
col_stats(arc, n_distinct)
```

```{r distinct_plot}
explore_plot(arc, type)
explore_plot(arc, document_category)
explore_plot(arc, type_of_contract)
```

### Amounts

```{r amount_summary}
mean(arc$amount_ordered != arc$amount_spent)
noquote(map_chr(summary(arc$amount_spent), dollar))
mean(arc$amount_spent <= 0)
```

The maximum `amount_spent` is `r dollar(max(arc$amount_spent))`. In this value,
we can see an enormous discrepancy between the `amount_ordered` and the
`amount_spent`.

```{r amount_max}
glimpse(arc[which.max(arc$amount_spent), ])
```

Turns out there are quite a few records with enormous difference between the 
amount ordered and spent.

```{r amount_dif}
arc %>% 
  mutate(amount_dif = amount_spent - amount_ordered) %>% 
  select(agency, contains("amount"), material_group) %>% 
  arrange(amount_dif)

arc %>% 
  mutate(amount_dif = amount_spent - amount_ordered) %>% 
  select(agency, contains("amount"), material_group) %>% 
  arrange(desc(amount_dif))
```

```{r hist_amount, echo=FALSE}
arc %>%
  filter(amount_spent >= 1) %>% 
  ggplot(aes(amount_spent)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:8)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Arkansas Contracts Amount Distribution",
    caption = "Source: Transparency Arkansas",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
arc <- mutate(arc, year = year(date))
```

```{r date_range}
prop_na(arc$date)
min(arc$date, na.rm = TRUE)
sum(arc$year < 2000, na.rm = TRUE)
max(arc$date, na.rm = TRUE)
sum(arc$date > today(), na.rm = TRUE)
arc$year <- na_if(arc$year, 9999)
```

```{r bar_year, echo=FALSE}
arc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  coord_cartesian(xlim = c(2008, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arkansas Contracts by Year",
    caption = "Source: Transparency Arkansas",
    x = "Year Made",
    y = "Count"
  )
```

## Export

1. There are `r comma(nrow(arc))` records in the database.
1. There are `r comma(sum(arc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(arc$na_flag))` records missing key variables.
1. There are no geographic variables to be normalized.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ar", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "ar_contracts_clean.csv")
write_csv(arc, clean_path, na = "")
file_size(clean_path)
```

The encoding of the exported file should be UTF-8 or ASCII.

```{r clean_encode}
enc <- system2("file", args = paste("-i", clean_path), stdout = TRUE)
str_replace_all(enc, clean_path, basename)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(arc)),
  type = md_code(map_chr(arc, typeof)),
  both = FALSE,
  def = c(
    "Source file",
    "Transaction's fiscal year",
    "Spending agency name",
    "Contract category",
    "Tracking contract number",
    "Purchase order category",
    "Contract sub-type",
    "Vendor name",
    "Vendor \"doing business as\"",
    "Contract start date",
    "Contract end date",
    "Total contract value",
    "Amount initially ordered",
    "Amount finally spent",
    "Spending agency ID",
    "Releasing agency ID",
    "Purchase order date",
    "Type of record (Contract or purchase)",
    "Unified transaction date",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",    
    "Calendar year from unified date"
  )
)
```

```{r dict_md, echo=FALSE}
dict_raw$both[c(1:3, 5:6, 8, 13:14, 18:22)] <- TRUE
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Overlaped", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Arkansas Contracts Data Dictionary\n", dict_md),
  path = here("ar", "contracts", "ar_contracts_dict.md"),
)
```

