---
title: "Florida Lobbyist Registration"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("fl", "lobbying", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load_gh("kiernann/gluedown")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained as tab-delinated files from the [Florida Lobbying Registration Office][lro] (LRO).

[lro]: https://floridalobbyist.gov/

> Legislative and Executive branch lobbyist/principal registration datasets can be downloaded free
of charge. Each dataset contains basic information about lobbyists, the principals they register to
represent, and any associated firm information. Click on the File Definition Description link below
to view the file format. The datasets are updated daily.

```{r download_key}
key_url <- "https://floridalobbyist.gov/reports/disk%20file%20definition.pdf?cp=0.3379601757893852"
download.file(key_url, destfile = url2path(key_url, here("fl", "lobbying", "reg", "docs")))
```

## Import

### Download

The LRO provides the datasets in tab-delimited format.

> The tab-delimited files below are in the (.TXT) format and can be imported into any word
processor, spreadsheet, or database program.
> 
> * [Legislative Lobbyist File](https://floridalobbyist.gov/reports/llob.txt)
> * [Executive Lobbyist File](https://floridalobbyist.gov/reports/llob.txt)

We can download these two files to our raw directory.

```{r create_raw_dir}
raw_dir <- here("fl", "lobbying", "reg", "data", "raw")
dir_create(raw_dir)
```

```{r download_raw}
llob_url <- "https://floridalobbyist.gov/reports/llob.txt"
llob_file <- url2path(llob_url, raw_dir)
download.file(url = llob_url, destfile = llob_file)

elob_url <- "https://floridalobbyist.gov/reports/elob.txt"
elob_file <- url2path(elob_url, raw_dir)
download.file(url = elob_url, destfile = elob_file)
```

### Read

We can read both files at once with the `vroom::vroom()` function.

```{r read_raw}
fllr <- vroom(
  file = dir_ls(raw_dir),
  .name_repair = make_clean_names,
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  id = "source",
  skip = 2,
  col_types = cols(
    .default = col_character(),
    `Eff Date` = col_date_usa(),
    `WD Date` = col_date_usa(),
  )
)
```

The original file contains _three_ different types of data, with the type specified in the second
row of the spreadsheet.

```{r col_data_types, results='asis', echo=FALSE}
read_lines(llob_file)[[2]] %>% 
  str_split("\t+") %>% 
  extract2(1) %>% 
  na_if("") %>% 
  na.omit() %>% 
  md_list(cat = TRUE)
```

This resuled in most column names being repeated for each entity type, with `vroom::vroom()`
and `janitor::make_clean_names()` appending each repeated name with a unique digit. We will replace
these unique digits with a meaningful prefix identifying the entity type.

```{r list_names}
fllr_names <- names(fllr)
fllr_prefix <- c("lobby", "client", "firm")
```

```{r change_names, message=FALSE, warning=TRUE}
fllr_names[02:14] <- str_c(fllr_prefix[1], str_remove(fllr_names[02:14], "_(.*)$"), sep = "_")
fllr_names[15]    <- str_c(fllr_prefix[2], str_remove(fllr_names[15],    "^(.*)_"), sep = "_")
fllr_names[16:24] <- str_c(fllr_prefix[2], str_remove(fllr_names[16:24], "_(.*)$"), sep = "_")
fllr_names[26:36] <- str_c(fllr_prefix[3], str_remove(fllr_names[26:36], "_(.*)$"), sep = "_")
which_address <- str_which(fllr_names, "address")
fllr_names[which_address] <- str_c(fllr_names[which_address], c(1:3, 1:2, 1:2))
all(fllr_names == tidy_names(fllr_names))
```

We can see how this process made the variable names much more useful.

```{r check_names, echo=FALSE}
tibble(
  origial = names(fllr),
  fixed = fllr_names
) %>%
  sample_n(10) %>% 
  arrange(fixed)
```

So we can overwrite the orignal names with this new vector.

```{r replace_names}
fllr <- set_names(fllr, fllr_names)
```

## Explore

```{r glimpse}
head(fllr)
tail(fllr)
glimpse(sample_frac(fllr))
```

### Missing

```{r glimpse_na}
glimpse_fun(fllr, count_na)
```

```{r flag_na}
fllr <- flag_na(fllr, ends_with("name"))
sum(fllr$na_flag)
mean(fllr$na_flag)
```

### Duplicates

```{r flag_dupes}
fllr <- flag_dupes(fllr, everything())
sum(fllr$na_flag)
mean(fllr$na_flag)
```

### Dates

The database only contains registrants for the current year.

```{r add_year}
fllr <- mutate(fllr, client_year = year(client_eff))
unique(fllr$client_year == year(today()))
```

```{r date_range, collapse=TRUE}
min(fllr$client_eff)
max(fllr$client_eff)
```
