---
title: "United States Plum Book"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "plum_book", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  pdftools, # read pdf files
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("us/plum_book/docs/us_plum_book.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Government Publishing Office (GPO) released the [_Plum Book_][pb], a list of
federal positions often subject to noncompetitive (political) appointment. This
book is physically printed, published as a PDF, and sometimes as a text file.

> Every four years, just after the Presidential election, the United States
Government Policy and Supporting Positions, commonly known as the Plum Book, is
published, alternately, by the Senate and the House. The Plum Book is used to
identify presidentially appointed positions within the Federal Government.

[pb]: https://www.govinfo.gov/app/details/GPO-PLUMBOOK-2020

> |Summary          |                                                         |
> |:----------------|:--------------------------------------------------------|
> |Collection       |Additional Government Publications                       |
> |Publication Name |United States Government Policy and Supporting Positions |
> |Date Issued      |December 1, 2020                                         |
> |Congress         |116th Congress, 2nd Session                              |
> |Page Range       |1-221                                                    |
> |Committees       |Committee on Oversight and Government Reform             |

### Foreword

> This publication contains data (as of June 30, 2020) on over 9,000 Federal
civil service leadership and support positions in the legislative and executive
branches of the Federal Government that may be subject to noncompetitive
appointment (e.g., positions such as agency heads and their immediate
subordinates, policy executives and advisors, and aides who report to these
officials). The duties of many such positions may involve advocacy of
Administration policies and programs and the incumbents usually have a close and
confidential working relationship with the agency head or other key officials.
>
> Following are the major categories of positions listed:
>
> * Executive Schedule and salary-equivalent positions paid at the rates
> established for levels I through V of the Executive Schedule;
> * Senior Executive Service (SES) "General" positions;
> * Senior Foreign Service positions;
> * Schedule C positions excepted from the competitive service by the
> President, or by the Director, Office of Personnel Management, because of the
> confidential or policy-determining nature of the position duties; and
> * Other positions at the GS–14 and above level excepted from the competitive
> civil service by law because of the confidential or policy-determining nature
> of the position duties.
> 
> See Appendix 2 for more details on SES appointments and Appendix 3 for more
details on Schedule C appointments. Additional information on the positions
listed and the Federal salary schedules under which they are paid is provided in
the appendices. The Legend on the following page shows the codes and acronyms
used in this publication.

### Legend

The _Plum Book_ contains a legend section describing it's contents.

#### Position Location

> Listed are the cities, States/Provinces and foreign countries in which the
positions are located. Countries and cities (or other subdivisions) are shown
for overseas posts. Note that "Washington, DC" includes positions in the
entire metropolitan area and therefore may include certain cities and counties
in the States of Maryland and Virginia.

#### Name of Incumbent

> Listed are the names of individuals serving under other than career
appointments. The phrase "CareerIncumbent" is shown for positions incumbered by
career appointees. The term "Vacant" is shown for positionsthat were not
occupied on June 30, 2020, the "as of" date of this publication.
>
> Note the law requires "member" positions in certain agencies (e.g., boards,
committees, and commissions)be filled on a bipartisan basis. For such positions,
the following letter codes are shown in parentheses followingthe name of the
incumbent:
> 
> * (D) = Democrat
> * (I) = Independent
> * (R) = Republican

#### Type of Appointment 

> Listed are letter codes that denote the type of appointment under which the
position incumbent isserving. Note that several categories of positions can be
filled by more than one type of appointment, e.g., SESpositions listed in this
publication may be filled by using career, noncareer, limited emergency, or
limited termappointment authorities. Therefore, no "Type of Appointment" is
shown for such positions when they arevacant.
>
> * PAS = Presidential Appointment with Senate Confirmation
> * PA = Presidential Appointment (without Senate Confirmation)
> * CA = Career Appointment
> * NA = Noncareer AppointmentEA=Limited Emergency Appointment
> * TA = Limited Term Appointment
> * SC = Schedule C Excepted AppointmentXS=Appointment Excepted by Statute

#### Pay Plan

> Listed are letter codes that denote the Federal salary schedule or pay system
under which the position incumbents are going to be paid. Tables showing the
salary range for major pay systems are contained in Appendix 4.
>
> * AD = Administratively Determined Rates
> * ES = Senior Executive Service
> * EX = Executive Schedule
> * FA = Foreign Service Chiefs of Mission
> * FE = Senior Foreign Service
> * FP = Foreign Service Specialist
> * GS = General Schedule
> * PD = Daily Pay Rate* (per diem)
> * SL = Senior Level
> * TM = Federal Housing Finance Board Merit Pay
> * VH = Farm Credit Administration Pay Plan
> * WC = Without Compensation*
> * OT = Other Pay Plan* (all those not listed separately)
> 
> \* Although not pay plans, these codes are shown for information purposes

#### Level, Grade, or Pay

> Listed are numerical and letter codes that denote the level, grade or salary
of the position incumbered:
> 
> * Levels I through V of the Executive Schedule
> * Grades 1 through 15 of the General Schedule
> * Annual Salary in Dollars
> * Daily Pay Rate in Dollars
> 
> If there is no entry in this column, the position does not have an established
level, grade, or pay rate. Forexample, the pay rate for Senior Executive Service
and Senior Level positions is "personal," i.e., attached tothe incumbent, not
the position. The pay rate for each new appointee is set by the appointing
authority(usually the agency head) on a case-by-case basis. Annual salary
schedules and pay ranges for such positionsare shown in Appendix 4.

#### Tenure

> Listed are the terms or durations of the appointment in years. If there is no
entry in this column, theappointment has no fixed term.

#### Expires

> Listed are the expiration dates for time-limited appointments. If there is no
entry in this column, theincumbent is not serving under a time-limited
appointment. However, many member positions on agencyadvisory boards, councils,
and commissions are filled initially for a fixed term, but the incumbents may
(andoften do) serve beyond the expiration date until they are replaced. In such
cases, no expiration date is shown.

## Download

The Government Publishing Office (GPO) releases a ZIP archive containing all the
files related to the latest edition of the "Plum Book." This file can be
downloaded and extracted locally.

```{r raw_url}
raw_url <- "https://www.govinfo.gov/content/pkg/GPO-PLUMBOOK-2020.zip"
raw_dir <- dir_create(here("us", "plum_book", "data", "raw"))
raw_zip <- path(raw_dir, basename(raw_url))
```

```{r raw_download}
if (!file_exists(raw_zip)) {
  download.file(raw_url, raw_zip)
}
```

```{r raw_extract}
raw_all <- unzip(raw_zip, exdir = raw_dir)
```

We are interested in the Microsoft Excel spreadsheets in the `xsl/` directory.

```{r raw_list}
raw_xls <- str_subset(raw_all, "xlsx$")
names(raw_xls) <- basename(raw_xls)
```

## Read

These four excel files correspond to four employer types:

* [Legislative Branch][X1]
* [Executive Branch - Executive Office of the President][X2]
* [Executive Branch - Departments][X3]
* [Independent Agencies and Government Corporations][X4]

[X1]: https://www.govinfo.gov/content/pkg/GPO-PLUMBOOK-2020/xls/GPO-PLUMBOOK-2020-6.xlsx
[X2]: https://www.govinfo.gov/content/pkg/GPO-PLUMBOOK-2020/xls/GPO-PLUMBOOK-2020-7.xlsx
[X3]: https://www.govinfo.gov/content/pkg/GPO-PLUMBOOK-2020/xls/GPO-PLUMBOOK-2020-8.xlsx
[X4]: https://www.govinfo.gov/content/pkg/GPO-PLUMBOOK-2020/xls/GPO-PLUMBOOK-2020-9.xlsx

```{r raw_read}
pb <- map_dfr(
  .x = raw_xls, 
  .f = read_xlsx,
  .id = "src_file",
  col_types = "text"
)
```

```{r raw_fix}
pb <- pb %>% 
  clean_names(case = "snake") %>% 
  relocate(src_file, .after = last_col()) %>% 
  mutate(across(c(agcy_name, org_name), str_to_title)) %>% 
  mutate(across(expires, ~excel_numeric_to_date(as.numeric(.))))
```

## Explore

There are `r comma(nrow(pb))` rows of `r ncol(pb)` columns. Each record
represents a single contribution made to a political campaign or committee.

```{r glimpse}
glimpse(pb)
tail(pb)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(pb, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("agcy_name", "title", "name_of_incumbent")
pb <- flag_na(pb, all_of(key_vars))
sum(pb$na_flag)
```

```{r na_view}
pb %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
pb <- flag_dupes(pb, -id_number)
sum(pb$dupe_flag)
```

```{r dupe_view}
pb %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r distinct_count}
col_stats(pb, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(pb, agcy_name) + scale_x_wrap(10)
explore_plot(pb, org_name_level) + scale_x_wrap(10)
explore_plot(pb, type_of_appt) + scale_x_wrap(10)
explore_plot(pb, pay_plan) + scale_x_wrap(10)
explore_plot(pb, level_grade_or_pay) + scale_x_wrap(10)
explore_plot(pb, tenure) + scale_x_wrap(10)
```

## Wrangle

We will manually add the date this edition of the _Plum Book_ was published.

```{r date_add}
pb <- pb %>% 
  mutate(
    date = mdy("Dec 1, 2021"),
    year = year(date)
  )
```

We can also separate the `location` column into the city and state.

```{r location_split}
pb <- pb %>% 
  separate(
    col = location,
    into = c("city", "state"),
    sep = ",\\s",
    extra = "merge",
    remove = FALSE
  ) %>% 
  relocate(city, state, .after = last_col()) %>% 
  mutate(
    country = na_if(str_remove(state, "^[:upper:]{2}$"), ""),
    country = if_else(is.na(country) & !is.na(state), "United States", country),
    state = na_if(str_extract(state, "^[:upper:]{2}$"), "")
  )
```

## Conclude

1. There are `r comma(nrow(pb))` records in the database.
1. There are `r comma(sum(pb$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(pb$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "plum_book", "data", "clean"))
clean_path <- path(clean_dir, "us_plum_book.csv")
write_csv(pb, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
