---
title: "New Jersey Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nj", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("nj/voters/docs/nj_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The list of voters in New Jersey is from the Department of State's
[Division of Elections][doe] on their [Voter Information Portal][vip].

[doe]: https://nj.gov/state/elections/index.shtml
[vip]: https://nj.gov/state/elections/vote.shtml

This data was requested by the Investigative Reporting Workshop on September 18,
2020 and uploaded raw format to our AWS server.

```{r raw_dir}
raw_obj <- "svl-2020-0918-370128.zip"
raw_aws <- path("csv", "svl-2020-0918-370128.zip")
raw_dir <- dir_create(here("nj", "voters", "data", "raw"))
raw_zip <- path(raw_dir, raw_obj)
```

```{r raw_download}
if (!file_exists(raw_zip)) {
  save_object(
    object = raw_aws,
    bucket = "publicaccountability",
    file = raw_zip,
    show_progress = TRUE,
    overwrite = FALSE
  )
}
```

### Extract

The single archive file contains separate archives by county, each of those
containing a single comma-delimited text file. We can extract the text file
from each of these archives.

```{r raw_content}
zip_files <- mutate(
  .data = unzip(raw_zip, list = TRUE),
  across(Length, as_fs_bytes)
)
```

```{r echo=FALSE}
kable(zip_files)
```

```{r raw_unzip1}
raw_paths <- unzip(
  zipfile = raw_zip,
  exdir = raw_dir,
  junkpaths = TRUE,
  files = str_subset(
    string = zip_files$Name, 
    pattern = "^_", 
    negate = TRUE
  )
)
```

```{r raw_county}
county_zip <- str_subset(raw_paths, "zip$")
raw_paths <- character(length(county_zip))
csv_dir <- dir_create(here("nj", "voters", "data", "csv"))
```

```{r raw_unzip2}
for (i in seq_along(county_zip)) {
  aa <- unzip(county_zip[i], list = TRUE)
  raw_paths[i] <- unzip(
    zipfile = county_zip[i],
    exdir = csv_dir,
    junkpaths = TRUE,
    files = str_subset(
      string = aa$Name, 
      pattern = "^_", 
      negate = TRUE
    )
  )
}
```

```{r echo=FALSE}
raw_paths <- as_fs_path(raw_paths)
```

### About

The archive file also contains a PDF file describing the layout of these text
files and some of the codes used.

##### Voter Status

|code |status                        |
|:----|:-----------------------------|
|A    |Active                        |
|AD   |Active Need ID                |
|AF   |Active Federal Election Only  |
|IF   |Inactive Confirmation         |
|ID   |Inactive Confirmation Need ID |
|P    |Pending                       |
|PD   |Pending Need ID               |
|P7   |Pending-17                    |
|7D   |Pending-17 Need ID            |

##### Party Code

|code |party                   |
|:----|:-----------------------|
|CNV  |Conservative Party      |
|DEM  |Democratic              |
|GRE  |Green Party             |
|LIB  |Libertarian Party       |
|NAT  |Natural Law Party       |
|RFP  |Reform Party            |
|REP  |Republican              |
|SSP  |Socialist Party         |
|CON  |U.S. Constitution Party |
|UNA  |Unaffiliated            |

##### File Layout

|FIELD NAME      |FIELD DATA                   |FIELD TYPE       |
|:---------------|:----------------------------|:----------------|
|`displayId`     |Voter ID                     |TEXT             |
|`leg_id`        |Legacy Voter  ID             |TEXT             |
|`party`         |Party                        |TEXT             |
|`status`        |Status                       |TEXT             |
|`last`          |Last Name                    |TEXT             |
|`first`         |First Name                   |TEXT             |
|`middle`        |Middle Name                  |TEXT             |
|`suffix`        |Suffix                       |TEXT             |
|`dob`           |Date of Birth                |DATE(MM/DD/YYYY) |
|`street_num`    |Residence Street Number      |TEXT             |
|`street_pre`    |Residence Street Prefix      |TEXT             |
|`street_post`   |Residence Street  Post       |TEXT             |
|`street_base`   |Residence Base Street  Name  |TEXT             |
|`street_suff`   |Residence Street Suffix      |TEXT             |
|`street_name`   |Residence Street  Name       |TEXT             |
|`apt_unit`      |Residence Apt/Unit           |TEXT             |
|`city`          |Residence City               |TEXT             |
|`zip`           |Residence Zip                |TEXT             |
|`county`        |County                       |TEXT             |
|`municipality`  |Municipality                 |TEXT             |
|`ward`          |Ward                         |TEXT             |
|`district`      |District                     |TEXT             |
|`congressional` |Congressional District       |TEXT             |
|`legislative`   |Legislative District         |TEXT             |
|`freeholder`    |Freeholder District          |TEXT             |
|`school`        |School District              |TEXT             |
|`fire`          |Fire District                |TEXT             |

### Read

Using the record layout given in the PDF file, we can read each of the 
`r length(raw_paths)` text files into a single data frame.

```{r raw_read}
njv <- map_df(
  .x = raw_paths,
  .f = read_delim,
  .id = "source_file",
  delim = ",",
  guess_max = 0,
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    dob = col_date()
  )
)
```

```{r raw_source}
njv <- njv %>%
  relocate(source_file, .after = last_col()) %>% 
  mutate(source_file = basename(raw_paths)[as.integer(source_file)])
```

### Old

The Investigative Reporting Workshop previously requested this data in November
of 2019. The data received had a slightly different structure from the data from
2020. This includes the lack of a statewide voter ID and the presence of the
voter registration date. The 2020 data contained the same legacy county ID used
in 2019, meaning we can use this ID to identify the registration date for any
voter found in the 2019 database.

```{r old_file}
old_aws <- path("csv", "nj_voters.csv")
old_csv <- path(dirname(raw_dir), "nj_voters_old.csv")
```

```{r old_download}
if (!file_exists(old_csv)) {
  save_object(
    object = old_aws,
    bucket = "publicaccountability",
    file = old_csv,
    show_progress = TRUE,
    overwrite = FALSE
  )
}
```

```{r old_read}
njo <- read_csv(
  file = old_csv,
  col_types = cols(
    .default = col_character(),
    DATE_REGISTERED = col_date("%m/%d/%Y"),
    BIRTH_DATE = col_date("%m/%d/%Y")
  )
)
```

```{r old_dates}
prop_distinct(njo$VOTER_ID)
prop_in(njv$leg_id, njo$VOTER_ID)
old_info <- njo %>% 
  select(
    county = county_name,
    leg_id = VOTER_ID,
    reg_date = DATE_REGISTERED,
    sex = SEX
  )
```

```{r old_bind}
rm(njo)
flush_memory(2)
```

```{r old_join_date}
njv <- left_join(
  x = njv,
  y = old_info,
  by = c(
    "county",
    "leg_id"
  )
)
```

## Explore

There are `r comma(nrow(njv))` rows of `r ncol(njv)` columns. Each column
represents a single voter registered in the state of New Jersey.

```{r glimpse}
glimpse(njv)
tail(njv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(njv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
nm_vars <- c("first", "middle", "last", "suffix")
njv <- flag_na(njv, c(first, last))
sum(njv$na_flag)
```

There are `r sum(njv$na_flag)` records missing either the first or last name.
It would appear a lot of these voters come from countries with conventions
for naming different than the typical given name/surname combination.

```{r na_view}
njv %>% 
  filter(na_flag) %>% 
  select(all_of(nm_vars)) %>% 
  sample_n(20)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("nj", "voters", "dupes.tsv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  njs <- njv %>% 
    select(-displayId, -leg_id) %>% 
    group_split(district)
  split_id <- split(njv$displayId, njv$district)
  pb <- txtProgressBar(max = length(njs), style = 3)
  for (i in seq_along(njs)) {
    d1 <- duplicated(njs[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(njs[[i]], fromLast = TRUE)
      dupes <- tibble(displayId = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(njs)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("displayId", "dupe_flag"),
  col_types = cols(
    displayId = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(njv)
njv <- left_join(njv, dupes, by = "displayId")
njv <- mutate(njv, dupe_flag = !is.na(dupe_flag))
sum(njv$dupe_flag)
```

We can see that, despite supposedly unique IDs, there are duplicate voters.

```{r dupe_view}
njv %>% 
  filter(dupe_flag) %>% 
  select(displayId, all_of(nm_vars), street_name, city) %>% 
  arrange(last)
```

### Categorical

```{r distinct_count}
col_stats(njv, n_distinct)
```

```{r distinct_plots, echo=FALSE}
njv %>% 
  filter(!is.na(party)) %>% 
  count(party, sort = TRUE) %>% 
  filter(
    party %in% c(
      "Democratic",
      "Unaffiliated",
      "Republican",
      "Libertarian",
      "U.S. Constitution Party",
      "Green Party"
    )
  ) %>% 
  ggplot(aes(reorder(party, -n), n)) +
    geom_col(aes(fill = party)) +
    scale_y_continuous(labels = comma) +
    scale_fill_manual(
      guide = FALSE,
      values = c(
        "Democratic" = "#3333FF",
        "Unaffiliated" = "#666666",
        "Republican" = "#E81B23",
        "Libertarian" = "#FED105",
        "U.S. Constitution Party" = "#A356DE",
        "Green Party" = "#17aa5c"
      )
    ) +
    labs(y = "count", x = "party")
explore_plot(njv, status)
explore_plot(njv, county)
```

### Dates

A significant amount of voters have a registration date of January 1, 1901. This
seems to be the default date for any voter missing a real date. This is common
practice in voter data. We can remove these dates.

```{r date_bad}
sum(njv$reg_date == "1901-01-01", na.rm = TRUE)
njv$reg_date[which(njv$reg_date == "1901-01-01")] <- NA
```

We can then create variables identifying the year of birth and registration.

```{r year_birth}
njv <- mutate(
  .data = njv, 
  birth_year = year(dob),
  reg_year = year(reg_date)
)
```

```{r year_reg_bar, echo=FALSE}
njv %>% 
  count(reg_year) %>% 
  filter(reg_year > 1940, reg_year < 2021) %>% 
  mutate(
    type = case_when(
      reg_year %% 4 == 0 ~ "pres.",
      reg_year %% 2 == 0 ~ "mid.",
      TRUE ~ "none"
    )
  ) %>% 
  ggplot(aes(reg_year, n)) +
  geom_col(aes(fill = type)) +
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1940, 2020, by = 4)) +
  labs(
    title = "New Jersey Voters by Date of Registration",
    x = "Registration Year",
    y = "Count"
  )
```

```{r year_birth_bar, echo=FALSE}
njv %>%
  filter(dob > "1900-01-01", dob < today()) %>% 
  ggplot(aes(dob)) +
  geom_histogram(bins = 60, fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "New Jersey Voters by Date of Birth",
    x = "Date of Birth",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

The `zip` variable is already as clean as we can expect. 

```{r zip_check}
prop_in(njv$zip, valid_zip)
```

We will manually add a variable indicating these voters live in New Jersey.

```{r state_add}
njv <- mutate(njv, state = "NJ")
```

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
njv <- njv %>% 
  unite(
    col = address_norm,
    street_num, street_pre, street_post, street_name, apt_unit,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  relocate(address_norm, .after = last_col()) %>% 
  mutate(across(address_norm, normal_address))
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- njv %>% 
  distinct(city, state, zip) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("NJ", "DC", "NEW JERSEY"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c("state", "zip")
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  rename(city = city_raw)
```

```{r city_rejoin}
njv <- left_join(
  x = njv,
  y = norm_city,
  by = c("city", "state", "zip")
)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
njv <- njv %>% 
  select(
    -city_norm,
    city_clean  = city_swap
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(state, .after = city_clean)
```

```{r clean_glimpse}
glimpse(sample_n(njv, 50))
```

1. There are `r comma(nrow(njv))` records in the database.
1. There are `r comma(sum(njv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(njv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nj", "voters", "data", "clean"))
clean_path <- path(clean_dir, "nj_voters_clean.csv")
write_csv(njv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
