---
title: "North Carolina Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nc", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  batman, # na and lgl parse
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("nc/voters/docs/nc_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

North Carolina [voter registration][voter] data can be downloaded] from the
AWS S3 [file transfer page][ftp] of the [State Board of Elections][sbe].

> Access various types of North Carolina voter registration data on this page.
View current voter-level registration records or snapshots of voter registration
records across 15+ years. You may also view current group-level voter
demographic counts or election-specific voter demographic counts over the last
20+ years.

> The **"ncvoter"** files below contain the most up-to-date publicly available
information for individuals registered, pre-registered, or formerly registered
to vote in North Carolina, as well as individuals who have attempted to register
or have steps left uncompleted in the registration or preregistration process.

> **What’s included?** Voter registration status (e.g., active, inactive) and
reason are included. Voter demographics, birth state, address(es), and phone
number (when available) are included. Voter county, precinct, and district
information (e.g., NC Senate and House, judicial) are included. When applicable,
county commissioner and additional local districts (e.g. sanitation, school) are
included. See file layout for full list of variables.

> **What’s not included?** Voter birthdate, Social Security number, and driver’s
license number are confidential under state law and are not included. Voter
history is not included; but Voter Registration Data files are linkable to Voter
History Data files via NCID or county and voter registration number.

> **When are the files updated?** The files are updated weekly on Saturday
mornings. After each election, it may take multiple weeks for all 100 counties
to finalize their voter registration.

### About

The NCSBE file transfer page has a public data disclosure and file layout key.

```{r about_use}
readme_public_data <- s3read_using(
  FUN = read_lines,
  object = "data/ReadMe_PUBLIC_DATA.txt",
  bucket = "dl.ncsbe.gov"
)
```

```{r echo=FALSE, results='asis'}
md_quote(md_indent(readme_public_data))
```

```{r about_layout}
layout_ncvoter <- s3read_using(
  FUN = read_lines,
  object = "data/layout_ncvoter.txt",
  bucket = "dl.ncsbe.gov"
)
```

```{r echo=FALSE, results='asis'}
md_quote(md_indent(layout_ncvoter[1:14]))
```

#### Layout

The column layout table tells us what columns are included and what type of data
they contain.

```{r about_cols}
raw_layout <- read_fwf(
  file = layout_ncvoter[19:89],
  col_positions = fwf_positions(
    start = c(1, 25, 44),
    end = c(24, 43, NA),
    col_names = c("name", "data type", "description")
  )
)
```

```{r echo=FALSE}
raw_layout %>%
  rename(type = 2) %>% 
  mutate(
    .after = 2,
    length = as.integer(str_extract(type, "(?<=\\()\\d+(?=\\))")),
    type = case_when(
      name == "drivers_lic" ~ "<lgl>",
      type == "int" ~ "<int>",
      str_starts(type, "char") ~ "<fct>",
      str_starts(type, "var") ~ "<chr>",
      TRUE ~ type
    )
  ) %>% 
  mutate(across(1:2, md_code)) %>% 
  kable()
```

#### Codes

There are four additional tables describing the codes found in the data.

```{r about_fun_layout}
read_layout <- function(string, first, last, col_types = "cc") {
  string <- str_trim(string)
  read_fwf(
    file = string[seq(first, last)],
    col_types = col_types,
    col_positions = fwf_positions(
      start = c(1, 20),
      end = c(19, NA),
      col_names = str_split(
        string = string[first - 2],
        pattern = "\\s{2,}",
        simplify = TRUE
      )
    )
  )
}
```

##### Status

There are five codes for voter status.

```{r about_status_codes, echo=FALSE}
codes_status <- read_layout(layout_ncvoter, 97, 101)
```

```{r echo=FALSE}
kable(codes_status)
```

##### Race and Ethnicity

There are eight codes for the voter's race and three codes for ethnicity.

```{r about_race_codes, echo=FALSE}
codes_race <- read_layout(layout_ncvoter, 109, 116)
```

```{r echo=FALSE}
kable(codes_race)
```

```{r about_ethnicty_codes, echo=FALSE}
codes_ethnicity <- read_layout(layout_ncvoter, 124, 126)
```

```{r echo=FALSE, echo=FALSE}
kable(codes_ethnicity)
```

##### County

There are 100 counties in North Carolina, each with a numeric code.

```{r about_county_codes, echo=FALSE}
codes_county <- read_layout(
  string = layout_ncvoter, 
  col_types = "ci",
  first = 134, 
  last = 233 
)
```

```{r echo=FALSE}
codes_county
```

[ftp]: https://dl.ncsbe.gov/?prefix=data/
[sbe]: https://www.ncsbe.gov/
[voter]: https://www.ncsbe.gov/results-data/voter-registration-data

## Download

The `data/ncvoter_Statewide.zip` object can be downloaded and extracted locally.

```{r raw_dir}
raw_dir <- dir_create(here("nc", "voters", "data", "raw"))
raw_obj <- "ncvoter_Statewide.zip"
raw_zip <- path(raw_dir, raw_obj)
```

```{r raw_download}
if (!file_exists(raw_zip)) {
  save_object(
    object = path("data", raw_obj),
    bucket = "dl.ncsbe.gov",
    file = raw_zip,
    show_progress = TRUE
  )
}
```

```{r raw_unzip}
raw_txt <- path(raw_dir, unzip(raw_zip, list = TRUE)$Name)
if (!file_exists(raw_txt)) {
  unzip(raw_zip, exdir = raw_dir)
}
```

## Read

Using the column layout file, we can read this text file as a data frame.

```{r raw_read}
ncv <- vroom(
  file = raw_txt,
  delim = "\t",
  na = c("", " "),
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    county_id = col_integer(),
    absent_ind = col_skip(),
    name_prefx_cd = col_skip(),
    birth_age = col_integer(),
    registr_dt = col_date("%m/%d/%Y"),
    dist_2_abbrv = col_skip(),
    dist_2_desc = col_skip(),
    birth_year = col_integer()
  )
)
```

```{r raw_parse}
unique(ncv$drivers_lic)
ncv <- mutate(ncv, across(drivers_lic, to_logical))
ncv <- mutate(ncv, across(res_street_address, str_squish))
```

```{r old_save, echo=FALSE, eval=FALSE}
old_obj <- "nc_voters.csv"
old_csv <- path(raw_dir, old_obj)
if (!file_exists(old_csv)) {
  save_object(
    object = path("csv", old_obj),
    bucket = "publicaccountability",
    file = old_csv,
    show_progress = TRUE
  )
}
nco <- vroom(
  file = old_csv,
  delim = ",",
  quote = "",
  num_threads = 1,
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    county_id = col_integer(),
    absent_ind = col_skip(),
    name_prefx_cd = col_skip(),
    birth_age = col_integer(),
    registr_dt = col_date("%m/%d/%Y"),
    dist_2_abbrv = col_skip(),
    dist_2_desc = col_skip(),
    birth_year = col_integer()
  )
)
comma(nrow(nco))
comma(nrow(ncv))
nco <- nco %>% 
  # remove added columns
  select(-YEAR, -CITY_CLEAN) %>% 
  # remove voters in 2020 data
  filter(ncid %out% ncv$ncid)
comma(nrow(nco))
all(names(nco) %in% names(ncv))
ncv <- bind_rows(
  "dec2020" = ncv,
  "jun2019" = nco,
  .id = "source_age"
)
ncv <- relocate(ncv, source_age, .after = last_col())
rm(nco)
flush_memory(1)
```

## Explore

There are `r comma(nrow(ncv))` rows of `r ncol(ncv)` columns. Each record
represents a single voter once registered in the state of North Carolina,
represented by the unique `ncid` variable.

```{r glimpse}
prop_distinct(ncv$ncid)
glimpse(ncv)
tail(ncv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(ncv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("registr_dt", "first_name", "last_name", "birth_year")
ncv <- flag_na(ncv, all_of(key_vars))
sum(ncv$na_flag)
```

```{r na_view}
ncv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

```{r na_recount}
ncv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("nc", "voters", "dupes.tsv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  ncs <- ncv %>% 
    select(-ncid, -voter_reg_num) %>% 
    group_split(county_id)
  split_id <- split(ncv$ncid, ncv$county_id)
  pb <- txtProgressBar(max = length(ncs), style = 3)
  for (i in seq_along(ncs)) {
    d1 <- duplicated(ncs[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(ncs[[i]], fromLast = TRUE)
      dupes <- tibble(ncid = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(ohs)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("ncid", "dupe_flag"),
  col_types = cols(
    ncid = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(ncv)
ncv <- left_join(ncv, dupes, by = "ncid")
ncv <- mutate(ncv, dupe_flag = !is.na(dupe_flag))
sum(ncv$dupe_flag)
```

We can see that, despite unique IDs, there _are_ duplicate voters.

```{r dupe_view}
ncv %>% 
  filter(dupe_flag) %>% 
  select(ncid, all_of(key_vars), zip_code) %>% 
  arrange(last_name)
```

### Categorical

```{r distinct_count}
col_stats(ncv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3, warning=FALSE}
explore_plot(ncv, voter_status_desc)
explore_plot(ncv, voter_status_reason_desc) + scale_x_truncate(n = 15)
explore_plot(ncv, race_code)
explore_plot(ncv, ethnic_code)
ncv %>%
    count(party_cd, sort = TRUE) %>%
    mutate(p = .data$n/sum(.data$n)) %>%
    head(8) %>%
    ggplot(aes(reorder(party_cd, desc(.data$p)), .data$p)) +
    geom_col(ggplot2::aes(fill = party_cd)) +
    scale_y_continuous(labels = percent) +
    scale_fill_manual(
      guide = FALSE,
      values = c(
        DEM = "#3333FF",
        UNA = "#666666",
        REP = "#E81B23",
        LIB = "#FED105",
        CST = "#A356DE",
        GRE = "#17aa5c"
      )
    ) +
    labs(y = "Percent", x = "Party")
explore_plot(ncv, gender_code) + 
  scale_fill_manual(
    guide = FALSE,
    values = c(
      `U` = "#666666",
      `M` = "#7570B3",
      `F` = "#1B9E77"
    )
  )
```

### Dates

There are thousands of voters with a registration date of January 1, 1900. These
should be `NA` values, not real dates.

```{r date_min}
count_na(ncv$registr_dt)
min(ncv$registr_dt, na.rm = TRUE)
sum(ncv$registr_dt == "1900-01-01", na.rm = TRUE)
ncv$registr_dt[which(ncv$registr_dt == "1900-01-01")] <- NA
```

```{r date_max}
max(ncv$registr_dt, na.rm = TRUE)
sum(ncv$registr_dt > today(), na.rm = TRUE)
```

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
ncv <- mutate(ncv, registr_yr = year(registr_dt))
```

```{r bar_reg_year, echo=FALSE}
ncv %>% 
  filter(!is.na(registr_yr)) %>% 
  count(registr_yr) %>% 
  mutate(
    elect_type = case_when(
      registr_yr %% 4 == 0 ~ "Pres.",
      registr_yr %% 2 == 0 ~ "Mid.",
      TRUE ~ NA_character_
    )
  ) %>% 
  ggplot(aes(x = registr_yr, y = n)) +
  geom_col(aes(fill = elect_type)) + 
  scale_fill_manual(
    values = c("#1B9E77", "#D95F02"),
    na.value = "#666666"
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 4)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina Voters Registered by Year",
    caption = "Source: NC SOS",
    fill = "Election Year",
    x = "Year Registered",
    y = "Count"
  )
```

```{r hist_birth_year, echo=FALSE}
ncv %>% 
  filter(!is.na(birth_year)) %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(
    fill = dark2["purple"],
    binwidth = 3
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2005, by = 5)) +
  coord_cartesian(xlim = c(1900, 2005)) +
  theme(legend.position = "bottom") +
  labs(
    title = "North Carolina Voters by Birth Year",
    caption = "Source: NC SOS",
    x = "Birth Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

The `state_cd` and `zip_code` variables are already pretty normal, as we would
expect. Although there is a significant amount of missing values from each. 

```{r zip_check}
prop_in(ncv$zip_code, valid_zip)
prop_na(ncv$zip_code)
```

```{r state_count}
ncv %>% 
  count(state_cd, sort = TRUE) %>% 
  add_prop()
```

All of these voters missing `state_cd` or `zip_code` are also missing the city
name and have the `res_street_address` of "REMOVED".

```{r geo_vars}
geo_vars <- c("res_street_address", "res_city_desc", "state_cd", "zip_code")
```

```{r geo_view}
ncv %>% 
  select(all_of(key_vars), all_of(geo_vars)) %>% 
  filter(is.na(state_cd))
```

Almost all of these voters have a `voter_status_desc` of "REMOVED" which must
result in their address also being removed.

```{r geo_na_status}
ncv %>% 
  group_by(no_state = is.na(state_cd)) %>% 
  count(voter_status_desc, sort = TRUE) %>% 
  mutate(p = prop.table(n)) %>% 
  group_split() %>% 
  as.list()
```

We can makes these values `NA`.

```{r address_removed}
ncv <- mutate(ncv, across(res_street_address, na_if, "REMOVED"))
```

Aside from this, no additional normalization needs to be done. The addresses
are already in proper USPS format.

```{r address_count}
count(ncv, res_street_address, sort = TRUE)
```

The `res_city_desc` is also extremely clean, already above our 99% threshold.

```{r city_check}
many_city <- c(valid_city, extra_city, "ELON", "WINSTON-SALEM")
prop_in(ncv$res_city_desc, many_city)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(ncv, 50))
```

1. There are `r comma(nrow(ncv))` records in the database.
1. There are `r comma(sum(ncv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ncv$na_flag))` records missing key variables.
1. Geographic variables are already sufficiently normalized.
1. The 4-digit `registr_yr` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nc", "voters", "data", "clean"))
clean_path <- path(clean_dir, "nc_voters_clean.csv")
write_csv(ncv, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
