---
title: "Texas Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("tx", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  batman, # convert to logical
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data
Data is obtained from the [Texas Ethics Commission (TEC)][03]. According to 
[a TEC brochure][04],

> tatutory duties of the Ethics Commission are in Chapter 571 of the Government
Code.  The agency is responsible for administering these  laws:  (1)  Title  15,
Election  Code, concerning political  contributions  and expenditures, and
political advertising...

> The Ethics Commission serves as a repository of required disclosure statements
for state officials,  candidates,political  committees, lobbyists, and certain
district and county judicial officers.

Data is obtained from the [Campaign Finance section of the TEC website][05]. An
entire database can be downloaded as [a ZIP file][06]. The contents of that ZIP
and the layout of the files within are outlined in the 
[`CFS-ReadMe.txt` file][07].

> This zip package contains detailed information from campaign finance reports
filed electronically with the Texas Ethics Commission beginning July 1, 2000.
Flat File Architecture Record Listing -- Generated 06/11/2016 12:38:08 PM

```{r read_key}
readme_url <- "https://www.ethics.state.tx.us/data/search/cf/CFS-ReadMe.txt"
readme <- read_lines(here("tx", "expends", "CFS-ReadMe.txt"))
```

At the top of this file is a table of contents.

```{r}
read_table(readme[seq(13, 47, 2)][-2]) %>% 
  unite(`File Contents`, 2:3, sep = " ") %>% 
  mutate(
    `File Name(s)` = `File Name(s)` %>% 
      str_split(",\\s") %>% 
      map(md_code) %>% 
      map_chr(str_c, collapse = ", "),
    `File Contents` = str_trunc(`File Contents`, width = 30)
  ) %>% 
  kable()
```

From this table, we know the ExpendData record (`contribs_##.csv`) contains the
data we want.

> Expenditures - Schedules F/G/H/I - Expenditures from special pre-election
(formerly Telegram) reports are stored in the file `expn_t`. They are kept
separate from the expends file to avoid creating duplicates, because they are
supposed to be re-reported on the next regular campaign finance report.

```{r expend_cols, echo=FALSE}
x <- readme[499:535]
x[32] <- paste(str_trim(x[32], "right"), str_trim(x[33], "left"))
write_lines(x[-33], tmp <- file_temp())
exp_data <- read_fwf(
  file = tmp,
  na = "XXX",
  col_types = "icccic",
  col_positions = fwf_widths(
    widths = c(4, 46, 11, 31, 6, NA),
    col_names = c("Pos", "Field", "Type", "Mask", "Len", "Description") 
  )
)
exp_data %>% 
  mutate(across(Field, ~md_code(make_clean_names(.)))) %>% 
  kable()
```

The ExpendCategory record is a small table explaining the expenditure category
codes used.

```{r expend_codes, echo=FALSE}
write_lines(readme[545:547], tmp <- file_temp())
exp_cat <- read_fwf(
  file = tmp,
  na = "XXX",
  col_types = "icccic",
  col_positions = fwf_widths(
    widths = c(4, 46, 11, 31, 6, NA),
    col_names = c("Pos", "Field", "Type", "Mask", "Len", "Description") 
  )
)
exp_cat %>% 
  mutate(across(Field, ~md_code(make_clean_names(.)))) %>% 
  kable()
```

[03]: https://www.ethics.state.tx.us/search/cf/
[04]: https://www.ethics.state.tx.us/data/about/Bethic.pdf
[05]: https://www.ethics.state.tx.us/search/cf/
[06]: https://www.ethics.state.tx.us/data/search/cf/TEC_CF_CSV.zip
[07]: https://www.ethics.state.tx.us/data/search/cf/CFS-ReadMe.txt

### Download

```{r raw_paths}
raw_dir <- dir_create(here("tx", "expends", "data", "raw"))
zip_url <- "https://www.ethics.state.tx.us/data/search/cf/TEC_CF_CSV.zip"
zip_file <- path(raw_dir, basename(zip_url))
```

If the file hasn't been downloaded yet, do so now.

```{r download_raw}
if (!file_exists(zip_file)) {
  download.file(
    url = zip_url, 
    destfile = zip_file,
    method = "curl"
  )
}
```

### Extract

There are `r nrow(unzip(zip_file, list = T))` CSV files inside the ZIP archive.

```{r zip_list}
zip_contents <- 
  unzip(zip_file, list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>%
  mutate(across(length, as_fs_bytes)) %>% 
  filter(str_detect(name, "expend_\\d{2}"))
```

```{r zip_exp}
zip_expends <- str_subset(zip_contents$name, "expend_\\d{2}.csv")
```

If the files haven't been extracted, we can do so now.

```{r unzip_zip}
zip_expends <- as_fs_path(unzip(
  zipfile = zip_file,
  files = zip_expends,
  exdir = raw_dir
))
```

### Read 

The TEC provides a helpful [record layout key][08] describing the structure of
their flat files. We can use the details in this key to properly read the files
into R.

> The CSV file contains comma-delimited records –one line per record. Each
record consists of fields separated by commas.The following characters
constitute the permitted list. The space characterand commaarenotin this list.
`! @ # $ % * -_ + : ; . / 0-9 A-Z a-z`

> If a raw data field contains any character other than these permitted
characters, then the field is surrounded by double-quotesin the CSV. Space is
notin the above list–meaning that data containing spaces will be double-quoted.
Raw field data containing double-quotes will have doubled double-quotes in the
CSV encoding.In both raw dataand CSV encoding, new lines are represented with
the escape notation `\n`.

We can use this information as the arguments to `vroom::vroom()` and read all 8
files at once into a single data frame.

[08]: https://www.ethics.state.tx.us/data/search/cf/CampaignFinanceCSVFileFormat.pdf

```{r read_csv, collapse=TRUE}
txe <- vroom(
  file = zip_expends,
  .name_repair = make_clean_names,
  na = c("", "NA", "N/A", "UNKNOWN"),
  delim = ",",
  col_names = TRUE,
  escape_double = TRUE,
  escape_backslash = FALSE,
  num_threads = 1,
  locale = locale(tz = "US/Central"),
  col_types = cols(
    .default = col_character(),
    receivedDt = col_date("%Y%m%d"),
    expendDt = col_date("%Y%m%d"),
    expendAmount = col_double()
  )
)
```

```{r raw_rename, echo=FALSE}
txe <- txe %>% 
  # shorten column names
  rename_with(str_remove, everything(), "_name") %>% 
  rename_with(str_remove, everything(), "_cd") %>% 
  rename_with(str_remove, everything(), "_street") %>% 
  rename_with(str_replace, everything(), "ident", "id") %>% 
  rename(
    form = form_type,
    schedule = sched_form_type,
    report_id = report_info_id,
    received = received_dt,
    info_flag = info_only_flag,
    id = expend_info_id,
    date = expend_dt,
    amount = expend_amount,
    describe = expend_descr,
    category = expend_cat,
    description = expend_cat_descr,
    politics_flag = political_expend,
    reimburse_flag = reimburse_intended_flag,
    corp_flag = src_corp_contrib_flag,
    liveexp_flag = capital_livingexp_flag,
    payee_type = payee_persent_type,
    vendor = payee_organization,
    last = payee_last,
    suffix = payee_suffix,
    first = payee_first,
    prefix = payee_prefix,
    addr1 = payee_addr1,
    addr2 = payee_addr2,
    city = payee_city,
    state = payee_state,
    county = payee_county,
    country = payee_country,
    zip = payee_postal_code,
    region = payee_region
  ) %>% 
  # remove empty cols
  select(-record_type, -payee_short, -country, -county) %>% 
  # convert from Y/N to logical
  mutate(across(liveexp_flag, na_out, c("Y", "N"))) %>% 
  mutate(across(ends_with("_flag"), to_logical))
```

## Explore

```{r glimpse}
glimpse(txe)
tail(txe)
```

### Missing

Columns vary in their degree of missing values.

```{r na_glimpse}
col_stats(txe, count_na)
```

We can use `campfin::flag_na()` to create a new `na_flag` variable to identify
any record missing one of the values needed to identify the transaction.

We will have to create a temporary single variable with names for both 
individual and entity payees.

```{r na_flag, collapse=TRUE}
txe <- txe %>%
  mutate(payee = coalesce(last, vendor)) %>% 
  flag_na(payee, date, amount, filer) %>% 
  select(-payee)
```

```{r na_count}
percent(mean(txe$na_flag), 0.01)
```

```{r na_view}
txe %>% 
  filter(na_flag) %>% 
  select(last, vendor, date, amount, filer) %>% 
  distinct() %>% 
  sample_frac()
```

### Duplicates

We can also create a new `dupe_flag` variable to identify any record which is
duplicated at least once across all variables.

```{r flag_dupes}
d1 <- duplicated(select(txe, -id), fromLast = FALSE)
d2 <- duplicated(select(txe, -id), fromLast = TRUE)
txe <- mutate(txe, dupe_flag = d1 | d2)
percent(mean(txe$dupe_flag), 0.01)
rm(d1, d2); flush_memory()
```

```{r dupe_view}
txe %>% 
  filter(dupe_flag) %>% 
  select(last, vendor, date, amount, filer)
```

Much of these duplicate variables are also missing values

```{r dupe_na_mean}
percent(mean(txe$na_flag[txe$dupe_flag]), 0.01)
```

```{r dupe_na_view}
txe %>% 
  filter(dupe_flag) %>% 
  select(last, vendor, date, amount, filer) %>% 
  col_stats(count_na)
```

### Categorical

```{r distinct_view}
col_stats(txe, n_distinct)
```

```{r distinct_plot, echo=FALSE}
explore_plot(txe, filer_type)
explore_plot(txe, payee_type)
explore_plot(txe, category)
```

```{r distinct_lgl}
txe %>% 
  select(ends_with("_flag")) %>% 
  map_dbl(mean) %>% 
  enframe(
    name = "lgl_var",
    value = "prop_true"
  ) %>% 
  kable(digits = 2)
```

### Amounts

The `amount` value ranges from a 
`r dollar(min(txe$amount, na.rm = TRUE))` minimum to
`r dollar(max(txe$amount, na.rm = TRUE))`, with only 
`r sum(txe$amount < 0, na.rm = TRUE)` records having a value less than 
\$0.

```{r summary_amount, collapse=TRUE}
noquote(map_chr(summary(txe$amount), dollar))
sum(txe$amount <= 0, na.rm = TRUE)
```

The logarithm of `expend_amount` is normally distributed around the median value of `r dollar(median(txe$expend_amount, na.rm = TRUE))`.

```{r amount_histogram, echo=FALSE}
txe %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(bins = 30, fill = dark2["purple"]) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  geom_vline(xintercept = median(txe$amount, na.rm = TRUE)) +
  labs(
    title = "Texas Expenditure Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: Texas Ethics Commission"
  )
```

We can explore the distribution and range of `expend_amount` by expenditure category and filer type
to better understand how Texans are spending money during different kinds of campaigns.

```{r amount_violin_what, echo=FALSE}
txe %>% 
  filter(
    category %in% most_common(category),
    amount > 1e0, 
    amount < 1e4
  ) %>% 
  ggplot(
    mapping = aes(
      x = reorder(category, X = amount, FUN = median, na.rm = TRUE), 
      y = amount
    )
  ) +
  geom_violin(
    draw_quantiles = c(0.25, 0.50, 0.75),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
    mapping = aes(fill = category)
  ) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  labs(
    title = "Texas Expenditure Amount Distribution",
    caption = "Source: Texas Ethics Commission",
    x = "Expenditure Category",
    y = "Expenditure Amount"
  )
```

```{r amount_violin_who, echo=FALSE}
txe %>% 
  filter(
    filer_type %in% most_common(filer_type),
    amount > 1e0, 
    amount < 1e4
  ) %>% 
  ggplot(
    mapping = aes(
      x = reorder(filer_type, X = amount, FUN = median, na.rm = TRUE), 
      y = amount
    )
  ) +
  geom_violin(
    draw_quantiles = c(0.25, 0.50, 0.75),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
    mapping = aes(fill = filer_type)
  ) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  labs(
    title = "Texas Expenditure Amount Distribution",
    caption = "Source: Texas Ethics Commission",
    x = "Filer Type",
    y = "Expenditure Amount"
  )
```
  
### Dates

To better explore and search the database, we will create a `year` 
variable from `date` using `lubridate::year()`

```{r add_year}
txe <- mutate(txe, year = year(date))
```

The date range is fairly clean, with `r sum(txe$date > today(), na.rm = TRUE)` 
values after `r today()` and only `r sum(txe$year < 2000, na.rm = TRUE)` 
before the year 2000.

```{r date_range, collapse=TRUE}
percent(prop_na(txe$date), 0.01)
min(txe$date, na.rm = TRUE)
sum(txe$year < 2000, na.rm = TRUE)
max(txe$date, na.rm = TRUE)
sum(txe$date > today(), na.rm = TRUE)
```

We can see that the few expenditures in 1994 and 1999 seem to be outliers, with
the vast majority of expenditures coming from 2000 through 2019. We will flag
these records.

```{r count_year}
count(txe, year, sort = FALSE) %>% print(n = 23)
```

```{r year_bar, echo=FALSE}
txe %>% 
  filter(!is.na(year), year > 1999) %>%
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Texas Expenditure Count per Year",
    caption = "Source: Texas Ethics Commission",
    x = "Year Made",
    y = "Count",
    fill = "Election Year"
  ) +
  theme(legend.position = "bottom")
```

```{r amount_line_month, echo=FALSE}
txe %>% 
  mutate(
    month = month(date),
    even = is_even(year)
  ) %>% 
  group_by(even, month) %>% 
  summarise(median = median(amount)) %>% 
  ggplot(aes(month, median)) +
  geom_line(aes(color = even), size = 2) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  labs(
    title = "Texas Expenditure Median Amount by Month",
    caption = "Source: Texas Ethics Commission",
    x = "Month",
    y = "Median Expenditure",
    color = "Election Year"
  ) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_y_continuous(labels = dollar) +
  theme(legend.position = "bottom")
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addr1` and `addr2` variables, the `campfin::normal_address()`
function will force consistence case, remove punctuation, and abbreviate
official USPS suffixes.

```{r address_norm}
txe <- txe %>% 
  unite(
    col = addr_full,
    starts_with("addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address_view}
txe %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
txe <- txe %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  txe$zip,
  txe$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
txe <- txe %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
txe %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  txe$state,
  txe$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
txe <- txe %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("TX", "DC", "TEXAS"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
txe <- txe %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- txe %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
txe <- txe %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(txe$city_raw),
  txe$city_norm,
  txe$city_swap,
  txe$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(txe$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "{State} City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "{State} City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
txe <- txe %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(txe, 20))
```

1. There are `r comma(nrow(txe))` records in the database.
1. There are `r comma(sum(txe$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(txe$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("tx", "expends", "data", "clean"))
clean_path <- path(clean_dir, "tx_expends_clean.csv")
write_csv(txe, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
