---
title: "New Hampshire Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  snakecase, # change string case
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  batman, # rep(NA, 8) Batman!
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http query
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

New Hampshire campaign finance data from before 2016 can only be ontained in hand-written PDF
format. Data from after 2016 can be obtained from the [NH Campaign Finance System (CFS)][03].

[03]: https://cfs.sos.nh.gov/ "source"

### About

The [CFS FAQ page][04] defines the data we will be downloading:

>  RSA 664:2, IX defines "expenditure" as follows: the disbursement of money or thing of value or
the making of a legally binding commitment to make such a disbursement in the future or the
transfer of funds by a political committee to another political committee or to a candidate for the
purpose of promoting the success or defeat of a candidate or candidates or measure or measures.
"Expenditures" includes disbursement constituting independent expenditures, as defined in paragraph
XI. It dos not include: (a) the candidate's filing fee or his/her expenses for personal travel and
subsistence; (b) activity designed to encourage individuals to register to vote or to vote, if that
activity or communication does not mention a clearly identified candidate; (c) any communication by
any membership organization or corporation to its members or stockholders, if the primary purpose
of that membership organization or corporation is not for the purpose of promoting the success or
defeat of a candidate or candidate and measure or measures; or (d) any communication by a political
committee member that is not made for the purpose of promoting the success or defeat of a candidate
or candidates or measure or measures.  

> **What is an independent expenditure?**  
> RSA 664:2, XI defines "Independent Expenditures" as follows: Expenditure that pay for the
development and distribution of a communication that expressly advocates the election or defeat of
a clearly identified candidate or candidates or the success or defeat of a measure or measures,
which are made without cooperation or consultation with any candidate, or any authorized committee
or agent of such candidate, and which are not made in concert with, or at the request or suggestion
of, any candidates, or any authorized committee or agent of such candidate...
> 
> **Is there a separate independent expenditure report?**
> 
> There is no specific form. The information submitted must be in accordance with RSA 664:6, IV-a.
Any political committee whose independent expenditure, in aggregate, exceeds $500, shall file an
itemized statement which shall be received by the Secretary of State not later than 48 hours after
such expenditure is made. NOTE: In addition to this 48 hour notification, the independent
expenditure will be reported on the next report of receipts and expenditures that is due.

[04]: https://cfs.sos.nh.gov/Public/FAQ#Expenditures "FAQ"

## Import

To download the bulk data format, one needs to navigate to the [Expenditures search page][05]. From
there, remove "2020 General Election" from the "Election Cycle" drop down menu. Enter "01/01/2010"
in the "Transaction Date Range" input box. After searching with these parameters, download the file
by clicking the "CSV" button at the bottom of the page.

We will automate this using the `RSelenium` package.

```{r create_raw}
raw_dir <- here("nh", "expends", "data", "raw")
dir_create(raw_dir)
```

```{r remote_download, eval=FALSE}
remote_driver <- rsDriver(
  port = 4444L,
  browser = "firefox",
  extraCapabilities = makeFirefoxProfile(
    list(
      browser.download.dir = raw_dir,
      browser.download.folderList = 2L,
      browser.helperApps.neverAsk.saveToDisk = "text/csv"
    )
  )
)

# navigate to the NH download site
remote_browser <- remote_driver$client
remote_browser$navigate("https://cfs.sos.nh.gov/Public/ExpensesList")

# chose "All" from elections list
cycle_menu <- "/html/body/div[1]/div[3]/table/tbody/tr/td[4]/div[2]/table[1]/tbody/tr[3]/td/table/tbody/tr[6]/td[2]/select/option[1]"
remote_browser$findElement("xpath", cycle_menu)$clickElement()

# enter Jan 1 2008 as start date
remote_browser$findElement("css", "#dtStartDate")$sendKeysToElement(list("01/01/2008"))
remote_browser$findElement("css", "#dtEndDate")$sendKeysToElement(list(format(today(), "%m/%d/%Y")))

# click search button
remote_browser$findElement("css", "#btnSearch")$clickElement()

csv_button <- "td.bgfooter:nth-child(2) > a:nth-child(2)"
remote_browser$findElement("css", csv_button)$clickElement()

# close the browser and driver
remote_browser$close()
remote_driver$server$stop()
```

[05]: https://cfs.sos.nh.gov/Public/ExpensesList

```{r read_raw}
nh <- 
  read_csv(
    file = glue("{raw_dir}/ViewExpenditureList.csv"),
    col_types = cols(
      .default = col_character(),
      `Transaction Date` = col_date("%m/%d/%Y %H:%M:%S %p"),
      `Expenditure Amount` = col_double()
    )
  )
```

```{r read_problems}
problems(nh) %>% 
  filter(expected != "15 columns") %>% 
  distinct()
```

We will remove completely empty rows, clean names, uppcercase characters variables, and separate
some columns into their true underlying variables.

```{r format_raw}
nh <- nh %>%
  remove_empty("rows") %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper) %>% 
  separate(
    col = reporting_period, 
    remove = FALSE,
    into = c("reporting_date", "reporting_type"), 
    sep = "\\s-\\s"
  ) %>% 
  mutate(reporting_date = parse_date(reporting_date, "%m/%d/%Y")) %>% 
  separate(
    col = office,
    remove = FALSE,
    into = c("office_clean", "district_clean"),
    sep = "\\s-\\s",
    convert = TRUE
  )
```

## Explore

There are `r nrow(nh)` records of `r length(nh)` variables in the full database.

```{r glimpse}
head(nh)
tail(nh)
glimpse(sample_frac(nh))
```

### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
glimpse_fun(nh, n_distinct)
```

```{r payee_type_bar, echo=FALSE}
nh %>% 
  ggplot(aes(payee_type)) +
  geom_bar() +
  coord_flip()
```

```{r registrant_type_bar, echo=FALSE}
nh %>% 
  ggplot(aes(registrant_type)) +
  geom_bar() +
  coord_flip()
```

```{r office_bar, echo=FALSE}
nh %>% 
  ggplot(aes(office_clean)) +
  geom_bar() +
  coord_flip()
```

```{r county_bar, echo=FALSE}
nh %>% 
  ggplot(aes(county)) +
  geom_bar() +
  coord_flip()
```

```{r cycle_bar, echo=FALSE}
nh %>% 
  mutate(
    election_cycle = if_else(
      condition = election_cycle %in% c(
        "2018 ELECTION CYCLE", 
        "2020 ELECTION CYCLE", 
        "2016 ELECTION CYCLE"
      ),
      true = election_cycle,
      false = "OTHER"
    )
  ) %>% 
  ggplot(aes(election_cycle)) +
  geom_bar()
```

```{r report_type_bar, echo=FALSE}
nh %>% 
  filter(!is.na(reporting_type)) %>% 
  ggplot(aes(reporting_type)) +
  geom_bar()
```

```{r expend_type_bar, echo=FALSE}
nh %>% 
  ggplot(aes(expenditure_type)) +
  geom_bar() +
  coord_flip()
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
glimpse_fun(nh, count_na)
```

We will flag any records with missing values in the key variables used to identify an expenditure.

```{r na_flag}
nh <- flag_na(nh, payee_name)
sum(nh$na_flag)
```

### Duplicates

```{r get_dupes, collapse=TRUE}
nh <- flag_dupes(nh, everything())
sum(nh$dupe_flag)
```

### Ranges

#### Amounts

```{r range_amount}
summary(nh$expenditure_amount)
sum(nh$expenditure_amount <= 0)
```

```{r amount_hist}
nh %>% 
  ggplot(aes(expenditure_amount)) +
  geom_histogram() +
  scale_y_log10() +
  scale_x_continuous(labels = scales::dollar)
```

```{r amount_box_to}
nh %>% 
  ggplot(aes(payee_type, expenditure_amount)) +
  geom_boxplot(varwidth = TRUE) +
  scale_y_continuous(labels = scales::dollar, trans = "log10") +
  coord_flip()
```

```{r amount_box_from}
nh %>% 
  ggplot(aes(registrant_type, expenditure_amount)) +
  geom_boxplot(varwidth = TRUE) +
  scale_y_continuous(labels = scales::dollar, trans = "log10") +
  coord_flip()
```

```{r amount_box_how}
nh %>% 
  ggplot(aes(expenditure_type, expenditure_amount)) +
  geom_boxplot(varwidth = TRUE) +
  scale_y_continuous(labels = scales::dollar, trans = "log10") +
  coord_flip()
```

### Dates

```{r range_date, collapse=TRUE}
summary(nh$transaction_date)
sum(nh$transaction_date > today())
```

```{r year_bar}
nh %>% 
  ggplot(aes(year(transaction_date))) +
  geom_bar()
```

```{r}
nh %>% 
  group_by(month = month(transaction_date)) %>% 
  summarise(median_amount = median(expenditure_amount)) %>% 
  ggplot(aes(x = month, y = median_amount)) +
  geom_line(size = 2)
```

## Wrangle

### Year

Add a `transaction_year` variable from `transaction_date` using `lubridate::year()`.

```{r add_year}
nh <- nh %>% mutate(transaction_year = year(transaction_date))
```

```{r plot, year, echo=FALSE}
nh %>% 
  count(transaction_year) %>% 
  ggplot(aes(transaction_year, n)) +
  geom_col()
```

### Separate

The `payee_address` string contains the street address, city name, state abbreviation, and ZIP
code.

```{r sample_address}
sample(nh$payee_address, 10) %>% cat(sep = "\n")
```

We can use a combination of `tidyr::separate()` and `tidyr::unite()` to split this address
into four distinct columns:

1. `payee_address_sep`
1. `payee_city_sep`
1. `payee_state_sep`
1. `payee_zip_sep`

```{r separate_address}
nh <- nh %>% 
  separate(
    col = payee_address,
    into = c(glue("payee_split_address{1:10}"), "payee_city_sep", "payee_state_zip_sep"),
    sep = ",\\s",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) %>% 
  unite(
    starts_with("payee_split_address"),
    col = payee_address_sep,
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  separate(
    col = payee_state_zip_sep,
    into = c("payee_state_sep", "payee_zip_sep"),
    sep = "\\s(?=\\d)",
    remove = TRUE,
    convert = FALSE
  )
```

### Address

```{r normal_address}
nh <- nh %>% 
  mutate(
    payee_address_norm = normal_address(
      address = payee_address_sep,
      add_abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
nh %>% 
  select(payee_address_sep, payee_address_norm)
```

### ZIP Code

```{r zip_norm}
nh <- nh %>% 
  mutate(
    payee_zip_norm = normal_zip(
      zip = payee_zip_sep,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nh$payee_zip_sep,
  nh$payee_zip_norm,
  compare = valid_zip
)
```

### State

```{r state_norm}
nh <- nh %>% 
  mutate(
    payee_state_norm = normal_state(
      state = payee_state_sep,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

No changes were made.

```{r state_progress}
progress_table(
  nh$payee_state_sep,
  nh$payee_state_norm,
  compare = valid_state
)
```

### City

```{r city_norm}
nh <- nh %>% 
  mutate(
    payee_city_norm = normal_city(
      city = payee_city_sep,
      geo_abbs = usps_city,
      st_abbs = c("NH", "DC"),
      na_rep = TRUE
    )
  ) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "payee_state_norm" = "state",
      "payee_zip_norm" = "zip"
    )
  ) %>% 
  rename(
    payee_city_match = city
  ) %>% 
  mutate(
    match_abb = is_abbrev(payee_city_norm, payee_city_match),
    match_dist = str_dist(payee_city_norm, payee_city_match),
    payee_city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = payee_city_match,
      false = payee_city_norm
    )
  )
```

```{r city_progress}
progress_table(
  nh$payee_city_sep,
  nh$payee_city_norm,
  nh$payee_city_swap,
  compare = valid_city
)
```

## Conclude

```{r conclude_vals}
min_amount <- scales::dollar(min(nh$expenditure_amount, na.rm = TRUE))
max_amount <- scales::dollar(max(nh$expenditure_amount, na.rm = TRUE))

min_date <- as.character(min(nh$transaction_date, na.rm = TRUE))
max_date <- as.character(max(nh$transaction_date, na.rm = TRUE))
```

1. There are `r nrow(nh)` records in the database
1. There are `r sum(nh$dupe_flag)` records with duplicate rows(flagged with `dupe_flag`)
1. The `expenditure_amount` values range from `r min_amount` to `r max_amount`; 
the `transaction_date` values range from `r min_date` to `r max_date`
1. Consistency has been improved with `stringr` package and custom `normalize_*()` functions
1. The ZIP code and state abbreviation have been extracted fromt the `address` variable
1. The `transaction_year` variable has been created with `lubridate::year()`
1. There are `r sum(nh$na_flag)` records with missing `payee_name` values

## Export

```{r write_clean}
dir_proc <- here("nh", "expends", "data", "processed")
dir_create(dir_proc)

write_csv(
  x = nh,
  path = glue("{dir_proc}/nh_expends_clean.csv"),
  na = ""
)
```

