---
title: "Washington Voters"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 95)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("wa", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tabulizer, # read pdf tables
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  readxl, # read excel files
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::dr_here(show_reason = FALSE)
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Washington voter registration database (VRDB) can be obtained for free from
the Washington Secretary of State. The latest VRDB can be [obtained here][erq]
with an email address. The state also provides [an FAQ][faq] about the VRDB.

> The Secretary of State's Office maintains one statewide list of voters that
serves as the official list of registered voters for Washington. In January
2002, the Secretary of State asked the Legislature to authorize the creation of
a statewide voter registration database. The Legislature and Governor approved
the request. That same year Congress passed the Help America Vote Act, which
required states to develop a centralized voter registration database. In
compliance with the Help America Vote Act, the Washington State Voter
Registration Database was launched in January 2006.

[erq]: https://www.sos.wa.gov/elections/vrdb/extract-requests.aspx
[faw]: https://www.sos.wa.gov/elections/vrdb/vrdbfaq.aspx

## Download

The latest file was requested by Kiernan Nicholls on November 20, 2020.

> We have processed your request for the Washington State Voter Registration
Database. You can find the link to download that file below. This link will only
be active for 48 hours. If you’re unable to download the file during that time,
please resubmit your request or contact us at the contact information below and
we will send you a new link to the file.
>
> <https://www.sos.wa.gov/_assets/elections/8736776113.zip>

Since the URL used to access the data provided to the Workshop expires shortly,
a copy was archive on the Wayback Machine for reproducibility.

```{r raw_dir}
raw_dir <- dir_create(here("wa", "voters", "data", "raw"))
wbm_url <- "https://web.archive.org/web/20201123193808/"
raw_url <- "https://www.sos.wa.gov/_assets/elections/8736776113.zip"
raw_zip <- path(raw_dir, basename(raw_url))
```

```{r raw_download, eval=FALSE}
if (!file_exists(raw_zip)) {
  download.file(
    url = str_c(wbm_url, raw_url),
    destfile = raw_zip
  )
}
```

The contents of the downloaded archive can be extracted locally.

```{r raw_unzip}
raw_paths <- unzip(raw_zip, exdir = raw_dir)
```

## About

> You have downloaded the State Voter Registration Database (VRDB) Extract. This
extract includes all publically available data provided by RCW 29A.08.710 which
includes voter's name, address, political jurisdiction, gender, date of birth,
voting record, date of registration, and registration number. No other
information from voter registration records or files is available for public
inspection or copying.

> As provided by state law, the voter registration data contained in the file
you download may be used for political purposes only. Voter registration data
may not be used for commercial purposes. (RCW 29A.08.720, RCW 29A.08.740 and RCW
42.56.070(9))

### Format

> Included are two zipped text files: Registered voter files and Voting History.
Also included is a `Districts_Precincts` file in Microsoft Excel format that
includes all electoral districts in the state and the precincts that make them
up.
>
> The files included are compressed and will need to be uncompressed. Use WinZip
or similar software to unzip the files, or (with Windows XP or higher)
double-click to unzip the files.
>
> The data are in vertical bar-delimited format. We no longer offer Microsoft
Access format because the files are too large.
>
> The `Dflag` column was removed from the file starting on June 1, 2018. The
column was not used and therefore unnecessary for inclusion in the file.
> 
> The `CountyVoterID` column was removed from the file starting on August 1,
2019. The column is no longer used as Washington has transitioned to a unified
VoterID.

### Layout

```{r pdf_layout, echo=FALSE}
raw_cols <- extract_tables(
  file = raw_paths[6],
  pages = 1:2,
  output = "data.frame",
  header = FALSE,
  area = list(
    c(353, 29, 717, 529),
    c(56, 24, 230, 536)
  )
)

raw_cols <- raw_cols %>% 
  bind_rows() %>% 
  as_tibble() %>% 
  set_names(c("position", "name", "length", "type", "comments"))

kable(raw_cols)
```

### Districts

```{r xls_district}
wa_dist <- read_excel(
  path = raw_paths[1],
  .name_repair = make_clean_names
)
```

```{r echo=FALSE}
print(wa_dist)
```

## Read

```{r raw_txt}
raw_txt <- path(raw_dir, "202011_VRDB_Extract.txt")
```

The raw text file uses vertical pipe delimitation to separate columns.

```{r raw_read}
# 5,236,435
wav <- read_delim(
  file = raw_txt,
  delim = "|",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    Birthdate = col_date(),
    Registrationdate = col_date(),
    LastVoted = col_date()
  )
)
```

### Old

We want to keep the most comprehensive voter data possible, so we will use the
file previously requested by 

```{r old_path}
old_csv <- path(raw_dir, "wa_voters_old.csv")
old_aws <- "s3://publicaccountability/csv/wa_voters.csv"
old_head <- head_object(old_aws)
attr(old_head, "date")
```

```{r old_save}
if (!file_exists(old_csv)) {
  save_object(
    object = "csv/wa_voters.csv",
    bucket = "publicaccountability",
    file = old_csv,
    show_progress = TRUE
  )
}
```

```{r old_read}
wao <- read.csv(
  file = old_csv,
  na.strings = c("", " "),
  stringsAsFactors = FALSE,
  colClasses = "character"
)
```

```{r old_compare}
nrow(wao)
nrow(wav)
```

```{r old_ids}
head(wav$StateVoterID)
head(wao$StateVoterID)
```

```{r old_fix_id}
wao <- mutate(wao, across(StateVoterID, str_remove, "^WA0"))
prop_in(wao$StateVoterID, wav$StateVoterID)
```

```{r old_trim}
wao <- as_tibble(wao) %>% 
  filter(StateVoterID %out% wav$StateVoterID) %>% 
  select(-CITY_CLEAN, -YEAR, -BIRTHYEAR, -Title, -CountyVoterID) %>% 
  rename(RegStUnitNum = RegUnitNum) %>% 
  type_convert(
    col_types = cols(
      .default = col_character(),
      Birthdate = col_date("%m/%d/%Y"),
      Registrationdate = col_date("%m/%d/%Y"),
      LastVoted = col_date("%m/%d/%Y")
    )
  )
```

```{r old_diffnames}
setdiff(names(wav), names(wao))
```

### Bind

```{r bind_old}
wav <- wav %>% 
  bind_rows(wao, .id = "source_file") %>% 
  relocate(source_file, .after = last_col()) %>% 
  mutate(source_file = c("old", "new")[as.integer(source_file)])
```

```{r bind_recode}
wav <- mutate(wav, across(StatusCode, str_sub, end = 1))
```

```{r bind_count}
nrow(wav)
```

```{r raw_rename, echo=FALSE}
wav <- wav %>% 
  clean_names("snake") %>% 
  rename(
    voter_id = state_voter_id,
    name_first = f_name,
    name_mid = m_name,
    name_last = l_name,
    name_suf = name_suffix,
    birth_date = birthdate,
    st_num = reg_st_num,
    st_frac = reg_st_frac,
    st_name = reg_st_name,
    st_type = reg_st_type,
    unit_type = reg_unit_type,
    st_pre = reg_st_pre_direction,
    st_post = reg_st_post_direction,
    st_unit = reg_st_unit_num,
    city = reg_city,
    state = reg_state,
    zip = reg_zip_code,
    leg_district = legislative_district,
    congress_district = congressional_district,
    reg_date = registrationdate,
  )
```

## Explore

There are `r scales::comma(nrow(wav))` rows of `r ncol(wav)` columns. Each row
is a registered voter in the state of Washington.

```{r glimpse}
glimpse(wav)
tail(wav)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(wav, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("voter_id", "name_last", "name_first", "reg_date", "birth_date")
wav <- flag_na(wav, all_of(key_vars))
sum(wav$na_flag)
```

There are a small number of voters missing either a first or last name.

```{r na_view}
wav %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("wa", "voters", "dupes.tsv")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  was <- wav %>% 
    select(-voter_id) %>% 
    group_split(county_code)
  split_id <- split(wav$voter_id, wav$county_code)
  pb <- txtProgressBar(max = length(was), style = 3)
  for (i in seq_along(was)) {
    d1 <- duplicated(was[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(was[[i]], fromLast = TRUE)
      dupes <- tibble(voter_id = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, dupe_file, append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(fls)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = dupe_file,
  col_names = c("voter_id", "dupe_flag"),
  col_types = cols(
    voter_id = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(wav)
wav <- left_join(wav, dupes, by = "voter_id")
wav <- mutate(wav, dupe_flag = !is.na(dupe_flag))
sum(wav$dupe_flag)
```

We can see that there are duplicate voters in the database.

```{r dupe_view}
wav %>% 
  filter(dupe_flag) %>% 
  select(voter_id, all_of(key_vars), gender, source_file) %>% 
  arrange(reg_date)
```

Most of these duplicates were found in the old data we combined.

```{r dupe_old}
wav %>% 
  filter(dupe_flag) %>% 
  count(source_file, sort = TRUE)
```

### Categorical

```{r distinct_count}
col_stats(wav, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=2}
explore_plot(wav, gender)
explore_plot(wav, county_code)
explore_plot(wav, leg_district)
wav %>% 
  count(congress_district) %>% 
  mutate(across(congress_district, as.integer)) %>% 
  filter(congress_district > 1) %>% 
  ggplot(aes(congress_district, n)) + 
  geom_col(fill = dark2["purple"]) + 
  scale_x_continuous(breaks = 1:10)
explore_plot(wav, absentee_type)
explore_plot(filter(wav, mail_state != "WA"), mail_state) +
  labs(x = "mail_state (excluding WA)")
explore_plot(wav, status_code)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
wav <- mutate(
  .data = wav, 
  reg_year = year(reg_date),
  birth_year = year(birth_date)
)
```

```{r date_range}
min(wav$reg_date, na.rm = TRUE)
sum(wav$reg_date == "1900-01-01", na.rm = TRUE)
max(wav$reg_date, na.rm = TRUE)
```

```{r date_unborn}
wav %>% 
  select(all_of(key_vars)) %>% 
  filter(reg_date < birth_date)
```

```{r bar_birth_year, echo=FALSE}
wav %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(binwidth = 1, fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2010, by = 10)) +
  coord_cartesian(xlim = c(1900, 2010)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Washington Voters by Birth Year",
    x = "Year Made",
    y = "Count"
  )
```

```{r bar_reg_year, echo=FALSE}
wav %>% 
  filter(!is.na(reg_year)) %>% 
  count(reg_year) %>% 
  mutate(
    type = if_else(
      condition = reg_year %% 4 == 0,
      true = "pres.",
      false = if_else(
        condition = reg_year %% 2 == 0,
        true = "mid",
        false = "non"
      )
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = type)) +
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 10)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Washington Voters by Registration Year",
    fill = "Election Year Type",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

The various address component variables are already almost entirely in valid
USPS format, so it's easiest to just map each component variable, rather than
combine them into a single string and try to use `normal_address()`.

```{r address_norm}
wav %>% 
  select(st_num, st_frac, st_pre, st_name, st_type, st_post, st_unit)
```

```{r address_abb}
count(wav, st_type, sort = TRUE) %>% add_prop(sum = TRUE)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
wav <- wav %>% 
  mutate(
    zip_clean = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  wav$zip,
  wav$zip_clean,
  compare = valid_zip
)
```

### State

As we would expect, all the registered voters in Washington live in Washington.

```{r state_count}
count(wav, state, sort = TRUE)
```

### City

The `city` values are already quite normal, with the only unknown cities being
perfectly valid shorthands for neighborhoods and undesignated places

```{r city_count}
many_city <- c(valid_city, extra_city)
wav %>% 
  count(city, sort = TRUE) %>% 
  filter(city %out% many_city)
```

There are four common cities using abbreviations that can be replaced.

```{r city_unabb}
wav$city[str_which(wav$city, "^SPOKANE VLY")]   <- "SPOKANE VALLEY"
wav$city[str_which(wav$city, "^BAINBRIDGE IS")] <- "BAINBRIDGE ISLAND"
wav$city[str_which(wav$city, "^NINE MILE FLS")] <- "NINE MILE FALLS"
wav$city[str_which(wav$city, "^LAKE FOREST P")] <- "LAKE FOREST PARK"
```

There are also five common cities with unknown, valid names.

```{r city_add}
many_city <- c(many_city, 
  "CAMANO ISLAND", "SEATAC", "SEDRO-WOOLLEY", "EAST WENATCHE", "TULALIP"
)
```

With those 9 changes, we are over 99% valid.

```{r city_check}
percent(prop_in(wav$city, many_city), 0.01)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(wav, 50))
```

1. There are `r comma(nrow(wav))` records in the database.
1. There are `r comma(sum(wav$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wav$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("wa", "voters", "data", "clean"))
clean_path <- path(clean_dir, "wa_voters_clean.csv")
write_csv(wav, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

```{r clean_conv}
wav <- read_lines(clean_path)
wav <- iconv(wav, to = "ASCII", sub = "")
write_lines(wav, clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
