---
title: "United States Economic Injury Disaster Loans Diary"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "eidl", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Economic Injury Disaster Loans (EIDL) are a program of the Small Business
Administration. EIDL loans are typically given to relieve the economic strain of
natural disasters on local economies.

The Investigative Reporting Workshop has obtained a copy of EIDL loan 
disbursement from Investigative Reporters and Editors, Inc. (IRE), which they
originally obtained via a Freedom of Information Act request. The raw data set
can be purchased from IRE for $150.

> NICAR receives this data as an Access database from the SBA through a FOIA
request; we do not reformat or clean the data in any way. For the Access file we
change the names of the tables to make them easier to use; for the CSV format we
simply export the tables as comma-delimited files with double quote text
qualifiers...
> 
> Disaster loans from the Small Business Administration are the primary form of
federal assistance for non-farm, private-sector disaster losses. For this
reason, the disaster loans program is the only form of SBA assistance not
limited to small businesses. Disaster loans from SBA help homeowners, renters,
businesses of all sizes and nonprofit organizations finance their rebuilding.
The data identifies the borrower, the disaster, the amount and, for business
borrowers, whether the loan was paid in full or deemed uncollectible.

The file from IRE is obtained as a single ZIP archive.

```{r raw_zip}
raw_dir <- dir_create(here("us", "eidl", "data", "raw"))
raw_zip <- path(raw_dir, "SBADisaster_CSV_update.zip")
file_size(raw_zip)
```

We can extract the contents and move them to the raw directory.

```{r raw_extract}
unzip(raw_zip, exdir = raw_dir)
sba_dir <- path(raw_dir, "SBADisaster_CSV_update")
raw_files <- dir_ls(sba_dir) %>% 
  # str_subset(negate = TRUE)
  file_move(raw_dir)
dir_delete(sba_dir)
```

The ZIP archive contains a README file which tells us how to properly load data.

```{r raw_readme}
readme <- read_lines(path(raw_dir, "Readme_CSV.txt"))
md_quote(readme[4])
md_quote(readme[14:18])
```

## Read

Per the README, we can read the `Disaster.csv` file to get loan data.

```{r raw_read}
eidl <- vroom(
  file = path(raw_dir, "Disaster.csv"),
  escape_backslash = FALSE,
  escape_double = TRUE,
  delim = ",",
  na = c("", "NA"),
  .name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
    LOANTYPE = col_factor(),
    APPROVALAMT = col_double(),
    LOANAPPVDT = col_date("%Y%m%d")
  )
)
```

```{r raw_rename, echo=FALSE}
eidl <- rename(
  .data = eidl,
  loan_type = loantype,
  diaster_id = disaster_nbr,
  borrower = borrowername,
  address = borroweraddress,
  date = loanappvdt,
  amount = approvalamt
)
```

```{r raw_parse}
count(eidl, pif_ind)
count(eidl, chgoff_ind)
eidl <- mutate(eidl, across(12:13, ~!is.na(.)))
```

## Explore

```{r glimpse}
glimpse(eidl)
tail(eidl)
```

### Missing

No records are missing key values that need to be flagged.

```{r na_count}
col_stats(eidl, count_na)
```

### Duplicates

```{r dupe_flag}
d1 <- duplicated(eidl, fromLast = FALSE)
d2 <- duplicated(eidl, fromLast = TRUE)
eidl <- mutate(eidl, dupe_flag = d1 | d2)
sum(eidl$dupe_flag)
rm(d1, d2); flush_memory()
```

```{r dupe_view}
eidl %>% 
  filter(dupe_flag) %>% 
  select(date, borrower, amount, diaster_id) %>% 
  arrange(date)
```

### Categorical

```{r distinct_count}
col_stats(eidl, n_distinct)
```

```{r distinct_plots}
explore_plot(eidl, loan_type)
explore_plot(eidl, pif_ind)
explore_plot(eidl, chgoff_ind)
```

### Amounts

```{r ammount_summary}
summary(eidl$amount)
mean(eidl$amount <= 0)
```

```{r hist_amount, echo=FALSE}
eidl %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "United States EIDL Amount Distribution",
    caption = "Source: SBA EIDL",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
eidl <- mutate(eidl, year = year(date))
```

```{r date_range}
min(eidl$date)
mean(eidl$year < 2000)
max(eidl$date)
sum(eidl$date > today())
```

```{r bar_year, echo=FALSE}
eidl %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1987, 2018, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "United States EIDL Loans by Year",
    caption = "Source: SBA EIDL",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
eidl <- mutate(
  .data = eidl,
  address_norm = normal_address(
    address = address,
    abbs = usps_street,
    na_rep = TRUE
  )
)
```

```{r address_view}
eidl %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
eidl <- mutate(
  .data = eidl,
  zip_norm = normal_zip(
    zip = zip,
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  eidl$zip,
  eidl$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
eidl <- mutate(eidl, state_norm = state)
state_match <- eidl %>% 
  count(state_norm, zip_norm, sort = TRUE) %>% 
  left_join(zipcodes, by = c("zip_norm" = "zip")) %>% 
  filter(state_norm %out% valid_state) %>% 
  drop_na() %>% 
  count(state_norm, state, sort = TRUE) %>% 
  select(state_norm, state_match = state)
```

```{r state_match}
eidl <- eidl %>% 
  left_join(state_match) %>% 
  mutate(state_norm = coalesce(state_match, state_norm)) %>% 
  select(-state_match)
```

```{r state_progress}
progress_table(
  eidl$state,
  eidl$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
eidl <- mutate(
  .data = eidl,
  city_norm = normal_city(
    city = city, 
    abbs = usps_city,
    states = "DC",
    na = invalid_city,
    na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
eidl <- eidl %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- eidl %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
eidl <- eidl %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Almost all of the remaining unknown cities are in Guam, the Virgin Islands, or
similar territories whose cities are not in our list. These are places often
affected by natural disasters.

```{r city_remain}
many_city <- c(valid_city, extra_city)
eidl %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, state_norm, sort = TRUE) %>% 
  add_prop(sum = TRUE)
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(eidl$city_raw),
  eidl$city_norm,
  eidl$city_swap,
  eidl$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(eidl$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "EIDL City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "EIDL City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
eidl <- eidl %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(eidl, 50))
```

1. There are `r comma(nrow(eidl))` records in the database.
1. There are `r comma(sum(eidl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(eidl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "eidl", "data", "clean"))
clean_path <- path(clean_dir, "us_eidl_ire-2018.csv")
write_csv(eidl, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(eidl)),
  type = md_code(map_chr(eidl, typeof)),
  def = c(
    "Whether the loan went to a business (2) or a home (1)",
    "Disaster declaration numbers (to join with the DISLOOK)",
    "The name of business/homeowner(s) receiving loan guarantees",
    "The mailing address of the business/home",
    "The city of the business/home",
    "The state of the business/home",
    "The ZIP code of the business/home",
    "The date the loan was approved for an SBA guarantee: YYYYMMDD.",
    "The amount of the loan guaranteed by the SBA",
    "The Standard Industry Classification code",
    "The North American Industrial Classification codes",
    "Whether or not the loan was fully paid (yes or blank)",
    "Whether or not the loan was charged off (yes or blank)",
    "Flag indicating completely duplicate record",
    "Calendar year the loan was approved",
    "Normalized borrower mailing address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-letter state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
