---
title: "Arkansas Lobbying Expenditure Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ar", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Expenditures

Data is obtained from the [Arkansas Secretary of State][sos]. We can export expenditure by year with the [Financial Disclosure Archive Search][download] for 2008-2017 data. From 2017 onwards, the data exists in a different system.

> This tool allows you to access archived Financial Disclosure reports filed both on paper and online with the Secretary of State's office, including Statement of Financial Interest Reports, Extra Income Disclosures, Disclosure by the Members of the General Assembly, Contribution and Expenditure Reports, Political Action Committee Reports and Lobbyist Reports.

[sos]: https://www.sos.arkansas.gov/elections/financial-disclosure/
[download]: https://www.sos.arkansas.gov/lobbyist_search/index.php/search/advanced/new

```{r raw_dir}
raw_dir <-dir_create(here("ar", "lobby", "data", "raw"))
```


### Fix

To properly read so many records, we need to first perform some manipulation of
the text files. Each "cell" of character type columns are surrounded in double
quotation marks (`"`) to help prevent misreading. Howver, some of the text in
these cells itself contains double quotes or newline characters (`\n`).

We need to read each file as a character string and use regular expressions to
identify these erroneous characters and remove or replace them.

```{r raw_fix}
fix_check <- here("ar", "lobby", "data", "fixed.txt")
  fixed_dir <- dir_create(path(raw_dir, "fixed"))
# do not repeat if done
if (!file_exists(fix_check)) {
  # for all contrib and filer files
  for (file in dir_ls(raw_dir, glob = "*.xls")) {
    # read raw file
    read_file(file) %>% 
      # force conversion to simple
      str_conv(encoding = "ASCII") %>% 
      # replace newlines followed by a non-numeric character with a "/" 
      str_replace_all("\n(?=[:alpha:])","/") %>% 
      # replace newlines followed by a $ with a "\t"
      str_replace_all("\n\\$","\t") %>% 
      # replace all newlines NOT followed by "20xx"  with a space
      str_replace_all("\n(?!20)", " ") %>% 
      # overwrite raw file
      write_file(path(fixed_dir, basename(file)))
    # check progress
    message(paste(basename(file), "done"))
    # clean garbage memory
    gc()
  }
  # note this has done
  file_create(fix_check)
}
```


### Import


```{r raw_read}
arle <- map_dfr(dir_ls(fixed_dir), 
            read_delim,
            delim = "\t", 
            escape_double = FALSE, 
            trim_ws = TRUE, 
            col_types = cols(.default = col_character())) %>% 
  clean_names()
```

We'll do some cleaning with regard to column types.
```{r col type}
arle <- arle %>% 
  mutate(filing_date = as.Date(filing_date, format = "%m-%d-%y")) %>% 
  # get rid of the dollar signs and commas in the amount collumns
  mutate_at(.vars = vars(ends_with("amount")), .funs = list(norm = ~ str_remove_all(.,"\\$|,"))) %>% 
  mutate_at(.vars = vars(ends_with("norm")), .funs = as.numeric)
```


### Explore

```{r raw_glimpse}
head(arle)
tail(arle)
glimpse(sample_frac(arle))
```

As we can see from `tail()`, the last two rows still need to be removed.


#### Missing

```{r n_distinct}
col_stats(arle, count_na)
col_stats(arle, n_distinct)
```

We can see that for some lobbying firms, the payment to each lobbyist is not specified.

There are no duplicate rows in the database.
#### Duplicates
There are many duplicate records in the dataset and identical records are BOTH flagged with `dupe_flag`.
```{r duplicated}
arle <- flag_dupes(arle, dplyr::everything())
sum(arle$dupe_flag)
```

```{r}
write_csv(
  x = arle,
  path = path(proc_dir, "ar_lobby_exp_early.csv"),
  na = ""
)
```

