---
title: "Mississippi PPP Breakdown"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 4
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/ms/",
  fig.width = 12,
  fig.height = 6,
  dpi = 300
)
if (!interactive()) {
  options(width = 95)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "covid", "ppp", "docs"))
```

## Data

This is an analysis of Paycheck Protection Program loans in Mississippi.

On December 1, 2020 the Small Business Administration was ordered by the court
to release more detailed data on loans made through the Paycheck Protection
Program (PPP), a major facet of the 2020 CARES Act to provide stimulus funding
during the ongoing COVID-19 pandemic. This detailed release came after the SBA
initially refused to publish any database, then released only partial data on
the largest loan recipients. The full database now contains all recipient names,
addresses, and exact loan amounts.

> #### PPP Is A Delegated Loan Making Process
> PPP loans are not made by SBA. PPP loans are made by lending institutions and
then guaranteed by SBA. Accordingly, borrowers apply to lenders and self-certify
that they are eligible for PPP loans. The self- certification includes a good
faith certification that the borrower has economic need requiring the loan and
that the borrower has applied the affiliation rules and is a small business. The
lender then reviews the borrower’s application, and if all the paperwork is in
order, approves the loan and submits it to SBA...

> #### Cancelled Loans Do Not Appear In The PPP Loan Data
The public PPP data includes only active loans. Loans that were cancelled for
any reason are not included in the public data release.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  patchwork, # combine plots
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  usmap, # plot us maps
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::dr_here(show_reason = FALSE)
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Read

We can read the normalized PPP data from the IRW S3 server.

```{r clean_find}
ppp_dir <- here("us", "covid", "ppp")
ppp_file <- path(ppp_dir, "data", "clean", "sba_ppp_full.csv")
```

```{r clean_down}
if (!file_exists(ppp_file)) {
  save_object(
    object = "csv/sba_ppp_full.csv",
    bucket = "publicaccountability",
    file = ppp_file
  )
}
```

```{r clean_read}
ppp <- vroom(
  file = ppp_file,
  col_types = cols(
    naics_code = col_character(),
    zip = col_character()
  )
)
```

```{r clean_arrange}
ppp <- arrange(ppp, date_approved)
```

We can add additional variables to identify which records are in Mississippi.

```{r clean_state_id}
ppp <- ppp %>% 
  mutate(
    is_ms = !is.na(state_clean) & state_clean == "MS",
    ms_lab = ifelse(is_ms, "Miss.", "All")
  )
```

We can then create a separate data frame of Mississippi loans only.

```{r clean_filter}
ms_ppp <- filter(ppp, is_ms)
```

## Explore

There are `r scales::comma(nrow(ms_ppp))` rows of `r ncol(ms_ppp)` columns.

Mississippi accounts for `r percent(nrow(ms_ppp)/nrow(ppp), 0.1)` of the total 
PPP loans, and `r percent(sum(ms_ppp$loan_amount)/sum(ppp$loan_amount), 0.1)` of 
the total amount disbursed. This is close to the 
`r percent(usa::facts$population[25]/sum(usa::facts$population), 0.1)` of the
US population that lives in the state.

```{r glimpse}
glimpse(ms_ppp)
tail(ms_ppp)
```

### Missing

Very few columns are missing a key value needed to identify the transaction.

```{r na_count}
col_stats(ms_ppp, count_na)
```

```{r key_vars}
id_vars <- c(
  "lender", "loan_amount", "date_approved", "jobs_reported", "business_name",
  "address_clean", "city_clean", "state_clean", "zip"
)
```

```{r na_view}
ms_ppp %>% 
  select(all_of(id_vars)) %>% 
  slice(which(!complete.cases(.))) %>% 
  filter(!is.na(jobs_reported))
```

### Duplicates

There are no duplicate records in the Mississippi subset.

```{r dupe_count}
sum(ms_ppp$dupe_flag)
sum(ppp$dupe_flag)
```

## Jobs

We can also explore the number of _jobs_ reportedly retained. These numbers were
self reported and a significant amount of applications reported zero jobs or
omitted a number altogether, although the number of missing jobs reportedly
saved was smaller in Mississippi than the country at large.

```{r jobs_missing}
ppp %>% 
  group_by(ms_lab) %>% 
  summarise(
    no_job = mean(is.na(jobs_reported)),
    zero_jobs = mean(jobs_reported == 0, na.rm = TRUE)
  )
```

These loans, with missing or invalid jobs reported, does not appear to be tied
to the time the loan was approved and probably isn't related to confusion.

```{r jobs_missing_time, echo=FALSE}
ppp %>% 
  group_by(week = epiweek(date_approved)) %>% 
  summarise(
    `No Jobs` = mean(is.na(jobs_reported)),
    `Zero Jobs` = mean(jobs_reported == 0, na.rm = TRUE)
  ) %>% 
  pivot_longer(
    cols = !week
  ) %>% 
  ggplot(aes(x = week, y = value)) +
  geom_col(aes(fill = name)) +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks = 14:32) +
  labs(
    title = "Loans Missing Reported Jobs Over Time",
    fill = "Type",
    x = "Week",
    y = "Proportion"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom")
```

There's even one business which reported _negative_ six jobs retained.

```{r jobs_negative}
ppp %>% 
  filter(jobs_reported < 0) %>% 
  select(all_of(id_vars))
```

Most loans did go to small businesses saving less than **25** jobs.

```{r jobs_small}
ppp %>% 
  group_by(is_ms) %>% 
  summarise(small_jobs = mean(jobs_reported < 25, na.rm = TRUE))
```

```{r hist_jobs, echo=FALSE, fig.height=10}
ppp %>%
  filter(jobs_reported > 0, jobs_reported <= 25) %>% 
  ggplot(aes(jobs_reported)) +
  geom_histogram(aes(fill = ms_lab), binwidth = 1) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  labs(
    title = "Mississippi Voters Amount Distribution",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1)
```

`r sum(ms_ppp$jobs_reported == 500, na.rm = TRUE)` businesses in Mississippi
reportedly retained the max of `r max(ms_ppp$jobs_reported, na.rm = TRUE)` jobs.

```{r jobs_max}
ms_ppp %>% 
  filter(jobs_reported == max(jobs_reported, na.rm = TRUE)) %>% 
  select(all_of(id_vars)) %>% 
  arrange(desc(loan_amount)) %>% 
  print(n = Inf)
```

From this list, we can see an incredible range of loan amounts for the same 500
jobs retained, from 
`r dollar(min(ms_ppp$loan_amount[ms_ppp$jobs_reported == 500], na.rm = TRUE))`
to 
`r dollar(max(ms_ppp$loan_amount[ms_ppp$jobs_reported == 500], na.rm = TRUE))`.

```{r head_tail, echo=FALSE}
head_tail <- function(data, n = 10) {
  list(
    head = head(data, n = n), 
    tail = tail(data, n = n)
  )
}
```

This leads us to investigate the dollars spent to save each job. There are
some businesses in Mississippi which took hundreds of thousands to save one
or two jobs and others which took only a couple thousand to save over a hundred
jobs.

```{r jobs_dollar_add}
ppp <- mutate(
  .data = ppp,
  .after = jobs_reported,
  dollar_job = loan_amount/jobs_reported
)
```

```{r}
top_bottom <- function(.data, n = 10) {
  nr = nrow(x = .data)
  top = seq(from = 1, to = n)
  bot = seq(from = nr - (n - 1), to = nr)
  .data[c(top, bot), ]
}
```

```{r jobs_dollar_minmax, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(!is.na(jobs_reported) & jobs_reported > 0) %>% 
  filter(loan_amount > 10, jobs_reported != 500) %>% 
  arrange(desc(dollar_job)) %>% 
  select(id_vars, dollar_job) %>% 
  relocate(dollar_job, .after = jobs_reported) %>% 
  top_bottom(10) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address_clean, 
      paste(city_clean, state_clean, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(date_approved, format, "%b, %d")) %>% 
  mutate(across(loan_amount, dollar, 0.01)) %>% 
  mutate(across(dollar_job, dollar, 0.01)) %>% 
  kable(
    col.names = c(
      "Lending Bank",
      "$ Mil.",
      "Date",
      "Jobs",
      "Cost Per",
      "Business Name",
      "Business Address"
    )
  )
```

The distribution of loan amount per job retained is similar for both Mississippi
and the United States, both with a spike of loans at $21,000 per job.

```{r jobs_dollar_hist, echo=FALSE}
ppp %>% 
  filter(dollar_job < 25000, dollar_job > 0) %>% 
  ggplot(aes(dollar_job)) +
  geom_histogram(aes(fill = ms_lab), bins = 50) +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(labels = dollar, n.breaks = 10) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Dollars Loaned per Job Reportedly Retained",
    x = "Dollars/Job",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~ms_lab,
    ncol = 1,
    scales = "free_y"
  )
```

A much greater proportion of these loans reporting roughly $20,000 reported
having saved a single job, mostly due to being a sole proprietorship,
independent contractor, or self-employed individual.

```{r jobs_20k_one, echo=FALSE}
ppp %>% 
  mutate(is_20k = between(dollar_job, 20500, 21000)) %>% 
  filter(!is.na(is_20k)) %>% 
  group_by(is_20k) %>% 
  summarise(one_job = mean(jobs_reported == 1, na.rm = TRUE))
```

```{r jobs_20k_sole, echo=FALSE}
solo_types <- c(
  "Sole Proprietorship",
  "Independent Contractors",
  "Self-Employed Individuals"
)
ppp %>% 
  mutate(is_20k = between(dollar_job, 20500, 21000)) %>% 
  filter(!is.na(is_20k)) %>% 
  group_by(is_20k) %>% 
  summarise(
    sole_prop = percent(mean(business_type %in% solo_types, na.rm = TRUE), 0.1)
  )
```

## Amounts

```{r amount_summary}
summary(ms_ppp$loan_amount)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
ms_ppp %>% 
  filter(
    loan_amount == min(loan_amount) | loan_amount == max(loan_amount)
  ) %>% 
  select(all_of(id_vars))
```

Overall, Mississippi has a similar loan amount distribution to the United States
at large.

```{r hist_amount, echo=FALSE, fig.height=10}
ppp %>%
  filter(loan_amount >= 100, loan_amount <= 100000) %>% 
  ggplot(aes(loan_amount)) +
  geom_histogram(aes(fill = ms_lab), bins = 50) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Mississippi Voters Amount Distribution",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1)
```

```{r hist_amount_log, echo=FALSE, fig.height=10}
ppp %>%
  filter(loan_amount >= 100) %>% 
  ggplot(aes(loan_amount)) +
  geom_histogram(aes(fill = ms_lab), bins = 50) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Mississippi Voters Amount Distribution",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1)
```

```{r amount_most, echo=FALSE}
ms_ppp %>% 
  arrange(desc(loan_amount)) %>% 
  select(all_of(id_vars)) %>% 
  head(20) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address_clean, 
      paste(city_clean, state_clean, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(business_name, str_to_title)) %>% 
  mutate(across(loan_amount, ~dollar(./1e6))) %>% 
  mutate(across(date_approved, format, "%b, %d")) %>% 
  kable(
    col.names = c(
      "Lending Bank",
      "$ Mil.",
      "Date",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

## Business Type

There are `r n_distinct(ppp$business_type)` different kinds of business types.

The PPP was ostensibly designed to help small businesses, but we can see that
most loans nationwide were given to Corporations. In Mississippi, a greater
percentage of loans were given to LLC's and sole proprietorship.

```{r biz_type, echo=FALSE}
ppp %>% 
  filter(!is.na(business_type)) %>% 
  count(is_ms, business_type) %>% 
  ggplot(aes(x = reorder(business_type, n), y = n)) +
  geom_col(aes(fill = business_type)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 0)) +
  scale_y_continuous(labels = comma) +
  scale_fill_discrete(guide = FALSE) +
  labs(
    title = "PPP Loans Issued by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~is_ms,
    ncol = 1,
    scales = "free_y"
  )
```

The distribution of loan amounts by business type is roughly similar between
loans to businesses in Mississippi and the country at large. The only major 
discrepancies come from the business types with very few loans in the state.

```{r biz_amount_violin, echo=FALSE}
ppp %>% 
  filter(
    !is.na(business_type), 
    loan_amount > 100,
    business_type %in% most_common(business_type, 8)
  ) %>% 
  ggplot(aes(x = reorder(business_type, loan_amount), y = loan_amount)) +
  geom_violin(
    mapping = aes(fill = ms_lab, alpha = ms_lab),
    draw_quantiles = c(0.50),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
    position = "identity"
  ) +
  scale_alpha_manual(values = c(1, 0.5), guide = FALSE) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(
    breaks = c(1 %o% 10^(2:7)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Amounts by Business Type",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r biz_jobs_count, echo=FALSE}
ppp %>% 
  filter(!is.na(business_type), 
         business_type %in% most_common(business_type, 8)) %>% 
  group_by(business_type, ms_lab) %>% 
  summarise(jobs = sum(jobs_reported, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(business_type, jobs), y = jobs)) +
  geom_col(aes(fill = business_type)) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 0)
    }
  ) +
   labs(
    title = "Jobs Retained by PPP in Miss. and the United States",
    subtitle = "Total Reported Jobs Retained by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Total Jobs Reported"
  )
```

```{r biz_jobs_dollar, echo=FALSE}
ppp %>% 
  filter(!is.na(business_type), 
         business_type %in% most_common(business_type, 8)) %>% 
  group_by(business_type, ms_lab) %>% 
  summarise(
    jobs_dollar = sum(jobs_reported, na.rm = TRUE)/(sum(loan_amount)/1e6)
  ) %>% 
  ggplot(aes(x = reorder(business_type, jobs_dollar), y = jobs_dollar)) +
  geom_col(aes(fill = business_type)) +
  facet_wrap(~ms_lab, ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
   labs(
    title = "Jobs Retained by PPP in Miss. and the United States",
    subtitle = "Total Jobs per Million Loaned Dollars by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Jobs/Dollar"
  )
```

```{r biz_dollar_jobs, echo=FALSE}
ppp %>% 
  filter(!is.na(business_type), 
         business_type %in% most_common(business_type, 8)) %>% 
  group_by(business_type, ms_lab) %>% 
  summarise(
    dollar_job = sum(loan_amount)/sum(jobs_reported, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = reorder(business_type, dollar_job), y = dollar_job)) +
  geom_col(aes(fill = business_type)) +
  facet_wrap(~ms_lab, ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 0)
    }
  ) +
   labs(
    title = "Jobs Retained by PPP in Miss. and the United States",
    subtitle = "Dollars per Reported Job Retained by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Jobs/Dollar"
  )
```

## Business Industry

```{r naics_download, echo=FALSE}
nurl <- "https://www.census.gov/eos/www/naics/2017NAICS/6-digit_2017_Codes.xlsx"
download.file(nurl, destfile = naics_file <- file_temp(ext = "xlsx"))
```

```{r naics_read, echo=FALSE}
naics_codes <- read_excel(
  path = naics_file,
  range = "A3:B1059",
  col_names = c("naics_code", "business_industry"),
  col_types = "text"
)
```

```{r naics_join, echo=FALSE}
ppp <- left_join(ppp, naics_codes)
ppp <- relocate(ppp, business_industry, .after = naics_code)
```

```{r}
ppp %>% 
  filter(business_industry %in% most_common(business_industry, 10)) %>% 
  group_by(ms_lab) %>% 
  count(business_industry, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(x = reorder(business_industry, p), y = p)) +
  geom_col(aes(fill = ms_lab), position = "dodge") +
  scale_y_continuous(labels = percent) +
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Count Proportion by Business Industry and State",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r}
ppp %>% 
  filter(business_industry %in% most_common(business_industry, 30)) %>% 
  ggplot(aes(x = ms_lab, y = loan_amount)) +
  geom_violin(
    mapping = aes(fill = ms_lab), 
    position = "dodge",
    draw_quantiles = c(0.50),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
  ) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = dollar, trans = "log10") +
  facet_wrap(~business_industry, ncol = 10)
```

## Business Ownership

```{r echo=FALSE}
str_empty_na <- function(x) {
  str_remove(str_replace_na(x), "NA")
}
```

### Race & Ethnicity

```{r race_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state_clean == "MS" & !is.na(state_clean)) %>% 
  count(race_ethnicity, sort = TRUE) %>% 
  mutate(
    p = if_else(race_ethnicity != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Race", "US Count", "", "MS Count", ""))
```

### Gender

```{r gender_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state_clean == "MS" & !is.na(state_clean)) %>% 
  count(gender, sort = TRUE) %>% 
  mutate(
    p = if_else(gender != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Gender", "US Count", "", "MS Count", ""))
```

### Veteran

```{r vet_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state_clean == "MS" & !is.na(state_clean)) %>% 
  count(veteran, sort = TRUE) %>% 
  mutate(
    p = if_else(veteran != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Veteran", "US Count", "", "MS Count", ""))
```

### Non-Profit

```{r nonprofit_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state_clean == "MS" & !is.na(state_clean)) %>% 
  count(non_profit, sort = TRUE) %>% 
  mutate(
    p = if_else(non_profit != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Non-Profit", "US Count", "", "MS Count", ""))
```

## Dates

```{r bar_day_n, echo=FALSE}
ppp %>% 
  count(ms_lab, date_approved) %>% 
  ggplot(aes(x = date_approved, y = n)) +
  geom_col(aes(fill = ms_lab), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1, 52, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loans Approved by Day of Year",
    caption = "Source: SBA",
    fill = "Election Year",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, ncol = 1, scales = "free_y")
```

```{r bar_day_amount, echo=FALSE}
ppp %>% 
  group_by(ms_lab, date_approved) %>% 
  summarise(sum = sum(loan_amount)/1e9) %>% 
  ggplot(aes(x = date_approved, y = sum)) +
  geom_col(aes(fill = ms_lab), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = function(x) dollar(x, suffix = "B")) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Amount Approved by Day of Year",
    caption = "Source: SBA",
    fill = "Election Year",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, ncol = 1, scales = "free_y")
```

## Lenders

## Businesses

## Geography

### County

```{r}
zcta <- "http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt"
zip_cross <- read_csv(file = zcta)
zip_cross <- zip_cross %>% 
  filter(STATE == 28) %>% 
  select(zip = 1, 2:3, zip_pop = ZPOP) %>% 
  unite(2:3, col = "fips", sep = "") %>% 
  inner_join(countypop)
```

```{r}
ms_county <- ms_ppp %>% 
  left_join(zip_cross) %>% 
  group_by(fips) %>% 
  summarise(
    n = n(),
    loan_sum = sum(loan_amount),
    loan_median = median(loan_amount),
    loan_mean = mean(loan_amount),
    loan_per = loan_sum/pop_2015
  ) %>% 
  distinct()
```

```{r map_county_n, echo=FALSE}
plot_usmap(
  regions = "counties", 
  include = "MS", 
  data = ms_county, 
  values = "n"
) +
  scale_fill_continuous(labels = dollar) +
  labs(
    title = "PPP Loans by MS County",
    fill = "# of Loans"
  )
```

```{r map_county_per, echo=FALSE}
plot_usmap(
  regions = "counties", 
  include = "MS", 
  data = ms_county, 
  values = "loan_per"
) +
  scale_fill_continuous(labels = dollar) +
  labs(
    title = "PPP Dollars per Capita by MS County",
    fill = "Dollars"
  )
```

```{r map_county_med, echo=FALSE}
plot_usmap(
  regions = "counties", 
  include = "MS", 
  data = ms_county, 
  values = "loan_median"
) +
  scale_fill_continuous(labels = dollar) +
  labs(
    title = "PPP Median Amount by MS County",
    fill = "Dollars"
  )
```

#### ZCTA

```{r zip_get, echo=FALSE}
shp_url <- "https://www.maris.state.ms.us/statewide/mstm/av/mszip10.zip"
shp_zip <- here("us", "covid", "ppp", "data", basename(shp_url))
download.file(shp_url, shp_zip)
shp_dir <- path_temp("shp")
```

```{r zip_unzip, echo=FALSE}
shp_files <- unzip(
  zipfile = shp_zip, 
  exdir = shp_dir
)
```

```{r zip_read, echo=FALSE}
library(rgdal)
ms_shp <- readOGR(dsn = shp_dir, layer = "mszip10")
ms_shp_df <- as_tibble(fortify(ms_shp))
```

```{r zip_poly, echo=FALSE}
ms_zip_poly <- tibble(
  id = as.character(seq(length(ms_shp$ZCTA5CE10)) - 1),
  zip = ms_shp$ZCTA5CE10
)
```

```{r zip_join, echo=FALSE}
ms_shp_df <- left_join(ms_shp_df, ms_zip_poly)
```

```{r zip_pop_read, echo=FALSE}
ms_zip_pop <- vroom(
  file = here("us", "covid", "ppp", "data", "ACSST5Y2019.S0101.csv"),
  skip = 1
)
```

```{r zip_pop_fix, echo=FALSE}
ms_zip_pop <- ms_zip_pop %>% 
  select(zip = 2, pop = 3) %>% 
  mutate(across(zip, str_remove_all, "^ZCTA5\\s"))
```

```{r zip_stat, echo=FALSE}
ms_zip_stat <- ms_ppp %>% 
  group_by(zip) %>% 
  summarise(
    n = n(),
    sum = sum(loan_amount),
    median = median(loan_amount)
  ) %>% 
  left_join(ms_zip_pop) %>% 
  mutate(
    n_per = n/(pop/1000),
    sum_per = sum/(pop/1000),
    med_per = median/(pop/1000)
  )

ms_shp_df_stat <- right_join(ms_zip_stat, ms_shp_df)
```

```{r zip_map_n, echo=FALSE}
zip_map_n <- ms_shp_df_stat %>%
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  coord_equal() +
  labs(
    title = "MS PPP Loans by ZCTA",
    fill = "Loans"
  ) +
  theme_void()
zip_map_npc <- ms_shp_df_stat %>% 
  filter(n_per < 100) %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = n_per)) +
  coord_equal() +
  labs(
    title = "MS PPP Loans per capita by ZCTA",
    fill = "Loans per 1,000"
  ) +
  theme_void()
zip_map_n + zip_map_npc
```

### Duplicate Addresses

```{r geo_count}
ms_geo <- ms_ppp %>% 
  select(all_of(id_vars)) %>% 
  group_by(address_clean, city_clean, state_clean, zip) %>% 
  mutate(num_loans = n(), group = cur_group_id(), .before = 1) %>% 
  arrange(desc(num_loans)) %>% 
  filter(num_loans > 1)
```

```{r}
ms_geo %>% 
  group_by(group, num_loans) %>% 
  summarise(mil_loan = sum(loan_amount)/1e6) %>% 
  arrange(desc(mil_loan))
```

```{r}
ms_geo %>% 
  group_by(group, num_loans) %>% 
  summarise(mil_loan = sum(loan_amount)/1e6) %>% 
  arrange(desc(num_loans))
```

```{r}
ms_geo %>% 
  group_by(group, num_loans) %>% 
  arrange(desc(num_loans)) %>% 
  select(id_vars)
```
