---
title: "Paycheck Protection Program Loans"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "covid", "ppp", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Paycheck Protection Loan Data is released by the Small Business Administration
on their public Box cloud storage server. The data released at the beginning of
June does _not_ contain exact loan amounts, only five ranges. For any loans
less than the smallest range of $150,000 loans are aggregated and provided as
separate files.

> #### Key Aspects of PPP Loan Data
> In releasing PPP loan data to the public, SBA is maintaining a balance between
providing transparency to American taxpayers and protecting small businesses’
confidential business information, such as payroll, and personally identifiable
information. Small businesses are the driving force of American economic
stability and are essential to America economic rebound from the pandemic. SBA
is committed to ensuring that any release of PPP loan data does not harm small
businesses or their employees...

> #### PPP Is A Delegated Loan Making Process
> PPP loans are not made by SBA. PPP loans are made by lending institutions and
then guaranteed by SBA. Accordingly, borrowers apply to lenders and self-certify
that they are eligible for PPP loans. The self- certification includes a good
faith certification that the borrower has economic need requiring the loan and
that the borrower has applied the affiliation rules and is a small business. The
lender then reviews the borrower’s application, and if all the paperwork is in
order, approves the loan and submits it to SBA...

> #### Cancelled Loans Do Not Appear In The PPP Loan Data
The public PPP data includes only active loans. Loans that were cancelled for
any reason are not included in the public data release.

## Download

We can download PPP loan data from the SBA Box server as a ZIP archive.

```{r raw_dir}
raw_dir <- dir_create(here("us", "covid", "ppp", "data", "raw"))
raw_zip <- path(raw_dir, "All Data by State.zip")
```

We can extract all files from the archive to a `data/raw/` directory.

```{r raw_unzip}
raw_path <- unzip(
  zipfile = raw_zip, 
  exdir = raw_dir,
  junkpaths = TRUE
)
```

```{r raw_relist, echo=FALSE}
raw_path <- dir_ls(raw_dir, glob = "*.csv")
```

Loan data is divided into one `foia_150k_plus.csv` file and
57 `foia_up_to_150k_*.csv` files for each state and territory.

```{r raw_info}
raw_dir %>% 
  dir_info(regexp = "csv$") %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename)) %>% 
  as_tibble()
```

## Read

We can read all these files into a single data frame using `purrr::map_df()`
and `readr::read_csv()`. There is a slight difference in columns across the
state files and single file; for large loans there is only a `loan_range` value
but for smaller loans broken down by state, there is a `loan_amount` instead.
When we merge these two files together, `NA` values will be used for any record
from a file without a given variable.

```{r raw_map}
ppp <- map_df(
  .x = raw_path, 
  .f = read_csv,
  .id = "file",
  na = c("", "N/A", "Unanswered"),
  col_types = cols(
    .default = col_character(),
    LoanRange = col_factor(),
    LoanAmount = col_double(),
    DateApproved = col_date_usa(),
    JobsRetained = col_integer()
  )
)
```

```{r raw_parse}
ppp <- ppp %>% 
  clean_names("snake") %>% 
  relocate(loan_amount, .before = loan_range) %>% 
  mutate(across(non_profit, ~!(is.na(.)))) %>% 
  mutate(across(file, basename)) %>% 
  relocate(file, .after = last_col())
```

## Explore

```{r glimpse}
glimpse(ppp)
tail(ppp)
```

### Missing

Variables differ in their degree of missing values.

```{r na_count}
col_stats(ppp, count_na)
```

No files are missing any values they aren't supposed to be missing. For loans
up to $150,000 (the ones aggregated by ZIP code in the individual state files),
there is no `business_name` variable.

### Duplicates

There are a small amount of duplicate records in the database, all of which can
be flagged with a new logical variable.

```{r dupe_flag}
d1 <- duplicated(ppp, fromLast = FALSE)
d2 <- duplicated(ppp, fromLast = TRUE)
ppp <- mutate(ppp, dupe_flag = d1 | d2)
percent(mean(ppp$dupe_flag), 0.01)
rm(d1, d2); flush_memory()
```

```{r dupe_view}
ppp %>% 
  filter(dupe_flag) %>% 
  select(loan_range, business_name, lender, date_approved)
```

### Categorical

```{r distinct_count}
col_stats(ppp, n_distinct)
```

```{r distinct_plots}
explore_plot(ppp, business_type) + scale_x_truncate()
explore_plot(ppp, gender)
explore_plot(ppp, veteran)
explore_plot(ppp, non_profit)
```

### Amounts

Since the amount values for loans over $150,000 are given as a range, we can't
combine them with the exact `loan_amount` given for aggregated records.

```{r ammount_summary}
summary(ppp$loan_amount)
mean(ppp$loan_amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
ppp %>%
  ggplot(aes(loan_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "United States PPP Loan Amount Distribution",
    subtitle = "For loans under $150,000 aggregated by ZIP code",
    caption = "Source: Small Business Administration",
    x = "aggregated Amount",
    y = "Count"
  )
```

```{r amount_range, echo=FALSE}
ppp %>% 
  filter(!is.na(loan_range)) %>% 
  count(loan_range) %>% 
  mutate(across(loan_range, fct_rev)) %>% 
  ggplot(aes(loan_range, n)) +
  geom_col(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "United States PPP Loan Range Distribution",
    subtitle = "For loans over $150,000 to exact businesses",
    caption = "Source: Small Business Administration",
    x = "Loan Range",
    y = "Count"
  )
```

We can combine these two variables into a new one using the minimum value
from ranges and the exact value from aggregates.

```{r amount_extract}
ppp <- ppp %>% 
  extract(
    # extract number
    col = loan_range, 
    into = "amount_min", 
    regex = "(\\d+(?:,\\d+)?)(?=-)",
    remove = FALSE,
  ) %>% 
  mutate(
    # add zeroes
    amount_min = case_when(
      str_detect(amount_min, ",") ~ amount_min,
      !str_detect(amount_min, ",") ~ str_c(amount_min, ",000,000")
    ),
    # combine with exact amounts
    amount_min = coalesce(parse_number(amount_min), loan_amount)
  )
```

```{r amount_hist_min, echo=FALSE}
ppp %>%
  ggplot(aes(amount_min)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(
    labels = comma,
    trans = "log10"
  ) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "United States PPP Loan Amount Distribution",
    subtitle = "Combined aggregated amount and minimum range amount",
    caption = "Source: Small Business Administration",
    x = "Amount with Minimum Range",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date_approved` with `lubridate::year()`

```{r date_year}
ppp <- mutate(ppp, year_approved = year(date_approved))
```

```{r date_range}
min(ppp$date_approved)
sum(ppp$year_approved < 2020)
max(ppp$date_approved)
sum(ppp$date_approved > today())
```

```{r bar_month, echo=FALSE}
ppp %>% 
  count(m = month(date_approved, label = TRUE)) %>% 
  ggplot(aes(x = m, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "United States PPP Loans by Month",
    caption = "Source: Small Business Administration",
    x = "Month Approved",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
ppp <- ppp %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
ppp %>% 
  select(starts_with("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

The `zip` variable is already extremely clean.

```{r zip_progress}
progress_table(ppp$zip, compare = valid_zip)
```

### State

The `state` variable is also entirely clean, aside from two values.

For these values of "XX" with a valid `zip` variable, we can use the matched
state abbreviation instead.

```{r state_match}
ppp %>% 
  filter(state %out% valid_state) %>% 
  count(state, zip, city, sort = TRUE) %>% 
  left_join(zipcodes, by = "zip")
```

```{r state_join}
state_match <- select(zipcodes, zip, state_norm = state)
ppp <- left_join(ppp, state_match, by = "zip")
ppp$state_norm[ppp$state != "XX"] <- NA
ppp <- mutate(ppp, state_norm = coalesce(state_norm, state))
ppp$state_norm <- str_replace(ppp$state_norm, "FI", "FL")
```

```{r state_progress}
sum(ppp$state == "XX")
sum(ppp$state_norm == "XX")
ppp %>% 
  filter(state == "XX") %>% 
  count(state_norm, sort = TRUE)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ppp <- ppp %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = "DC",
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ppp <- ppp %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- ppp %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
ppp <- ppp %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ppp$city_raw),
  ppp$city_norm,
  ppp$city_swap,
  ppp$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ppp$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "United States City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "United States City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
ppp <- ppp %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(ppp, 20))
```

1. There are `r comma(nrow(ppp))` records in the database.
1. There are `r comma(sum(ppp$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ppp$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.
The data frame will be split into two files, the original file types for loans
over and under $150,000.

```{r clean_paths}
clean_dir <- dir_create(here("us", "covid", "ppp", "data", "clean"))
clean_path <- path(clean_dir, "sba_ppp_loans.csv")
write_csv(ppp, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ppp)),
  type = md_code(map_chr(ppp, typeof)),
  def = c(
    "Aggregated loan amount (under $150,000)",
    "Loan range (over $150,000)",
    "Combined loan amount with range minimum",
    "Recipient business name",
    "Recipient business address",
    "Recipient business city name",
    "Recipient business state abbreviation",
    "Recipient business ZIP code",
    "North American Industry Classification System",
    "Recipient business type",
    "Recipient owner race or ethnicity",
    "Recipient owner gender",
    "Recipient owner veteran status",
    "Recipient business is non-profit",
    "Individual jobs retained by loan",
    "Date loan approved",
    "Lending institution name",
    "Loan recipient location code",
    "Source file name",
    "Flag indicating duplicate record",
    "Calendar year approved",
    "Normalized recipient address",
    "Normalized recipient state",
    "Normalized recipient city"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
