---
title: "COVID-19-Hit Nursing Home Disease Control History"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---
<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
options(knitr.kable.NA = '')
set.seed(5)
```

```{r create_docs_dir, echo=FALSE}
docs <- fs::dir_create(here::here("nursing_home", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://www.medicare.gov/nursinghomecompare/Data/About.html
[04]: https://wpinvestigative.github.io/nursing_homes_covid19/index.html# "Wash Post repo"
[05]: https://github.com/irworkshop/accountability_datacleaning/blob/master/nursing_home/cms_nursing_health_diary.md#Data "TAP CMS"

## Data
We'll join the dataset of nursing homes with reported cases of COVID-19 patients with CMS's Nursing Home Compare's disease control-related deficiencies. The list of nursing homes is compiled by [The Washington Post][04] and downloaded May 1, 2020. The nursing home history dataset was downloaded from [ the Centers for Medicare & Medicaid Services][03] and wrangled by the [Accountability Project][05] dated March 16, 2020.

There're two columns to join by, the CMS Provider Number and Provider Name. 
```{r read}
data_dir <- "../nursing_home/data"
nursing <- read_csv(dir_ls(data_dir,regexp = "nursing_infection.*"))
covid <- read_csv((dir_ls(data_dir,regexp = "U.S."))) %>% 
  clean_names() %>% 
  rename(provname = provider_name,
         provnum = cms_provider_number)
```

We can take a glimpse of the COVID-19 dataset.
```{r explore}
sample_frac(covid)
covid <- flag_dupes(covid, dplyr::everything())
```


```{r join}
infected_nursing <- covid %>% 
  left_join(nursing, by = c("provname","provnum","state"))
```

## Variable Encodings
Since not every nursing home with COVID-19 cases had a Emergency Preparedness dificiency related to disease-control, thus resulting in columns originally from the CMS table missing. We will create a column named `def_count` and mark records with these fields missing `0`, and all others 1, since they mean that there's indeed one count of dificiency from previous inspections.
```{r add def_count}
infected_explore <- infected_nursing %>% 
  # if the number of empty column is 0, there's 1 disease control deficiency, def_boolean = 1
  mutate(def_boolean = case_when(rowSums(is.na(.)) == 0 ~ 1,
                               rowSums(is.na(.)) != 0 ~ 0))
```

```{r total number}
# generate a raw count `def_count` of the def_boolean column
infected_table <- infected_explore %>% 
  group_by(provnum) %>% 
  summarize(def_count = sum(def_boolean))
# join the count back to explore so that we have all the information 
infected_table <- infected_table %>% 
  left_join(covid, by = "provnum") %>% 
  select(-x1)
```
We can see that COVID-19-Impacted nursing homes with most with the most previous disease control dificiencies are overwhelmingly for-profit.
```{r infected vis}
top20 <- infected_table %>% 
  arrange(desc(def_count)) %>% 
  head(20)

top20$provnum <- reorder(top20$provnum, top20$def_count)
top20$provname <- reorder(top20$provname, top20$def_count)

top20 %>% 
  ggplot(aes(x = provname,y = def_count,fill = ownership_type)) +
  geom_col() +
  scale_x_discrete(labels =  wrap_format(8))+
  scale_y_continuous()+
  scale_fill_brewer(palette = "Dark2") + 
  labs(
    title = "COVID-19-Impacted Nursing Homes'with Most Previous Disease Control Deficiencies",
    caption = "Source: CMS, The Washington Post",
    x = "Nursing Home Name",
    y = "Total disease-control-related deficiencies"
  ) +
  theme_bw()
```
### Group by Owners
The ownership data is obtained from [CMS][03] and wrangled by a staff member of the Investigative Reporting Workshop. The data diary can be accessed [here](https://github.com/irworkshop/accountability_datacleaning/blob/master/nursing_home/docs/cms_nursing_diary.md#Data).

Then we can join it to the joined table of COVID-19-impacted nursing homes' history of disease-control related deficiencies.
```{r load owners}
owner <- read_csv(dir_ls(data_dir, regexp = "owner")) 

owner <- owner %>% 
  select(provname, provnum,owner_name,owner_type,owner_percentage_clean, role_desc,association_date_clean)

infected_owner <- infected_explore %>% left_join(owner, by = c("provnum", "provname"))
```
Note that this table contains all owner information. 
We can group by owners this time and see whose nursing homes have the most disease-related deficiencies while having COVID-19 cases.
Note that here as long as the owner has a share in the nursing home facility, we'll count as once. This table doesn't take account of Before we make the join, we need to make sure that each owner is associated with each nursing home once. 
```{r explore owner}
owner_dedupe <- owner %>% flag_dupes(c(owner_name, provnum), .both=F) %>% filter(!dupe_flag)

infected_owner_dedupe <- infected_explore %>% left_join(owner_dedupe, by = c("provnum", "provname"))

infected_owner_table <- infected_owner_dedupe %>% 
  group_by(owner_name) %>% 
  summarize(def_count_by_owner = sum(def_boolean)) %>% 
  arrange(desc(def_count_by_owner)) %>% 
  left_join(infected_owner_dedupe, by= "owner_name") %>% 
  left_join(infected_table, by = names(infected_table) %>% setdiff("def_count")) %>% 
  rename(nursing_def_count = def_count)

infected_owner_dedupe %>% 
  group_by(owner_name) %>% 
  summarize(def_count_by_owner = sum(def_boolean)) %>% 
  arrange(desc(def_count_by_owner)) %>% 
          head(20) %>% 
    ggplot(aes(x = reorder(owner_name,def_count_by_owner),y = def_count_by_owner)) +
  geom_col(fill = "#66c2a5") +
  scale_x_discrete(labels =  wrap_format(8))+
  scale_y_continuous()+
  labs(
    title = "COVID-19-Impacted Nursing Homes with Most Previous Disease Control Deficiencies by Owners",
    caption = "Source: CMS, The Washington Post",
    x = "Owner Name",
    y = "Total number of disease-control-related deficiencies at owned nursing homes"
  ) +
  theme_bw()
  
```


### Export
There're several data products from the above joining and analysis. 

```{r proc dir}
# Set up processed directory
proc_dir <- dir_create(path(data_dir,"processed"))
```

1. Table 1: Table of all COVID-19-impacted nursing homes with one added column counting total numbers of disease control dificiencies
```{r glimpse infected_table}
glimpse(infected_table %>% sample_frac())
infected_table %>% write_csv(path = path(proc_dir,"infected_counts.csv"), na = "")
```
2. Table 2: Table of all COVID-19-impacted nursing homes and each of their disease-control related deficiencies detail table.
```{r}
glimpse(infected_table %>% sample_frac())
infected_explore %>% write_csv(path = path(proc_dir,"infected_details.csv"), na = "")
```
3. Table 3: This is essentially table 2, with additional owner information, types, first_association date etc.
Note that if a nursing home has one deficiency but multiple owner rows, it will show up multiple times.
```{r}
glimpse(infected_owner %>% sample_frac())

infected_owner %>% write_csv(path = path(proc_dir,"infected_owners.csv"), na = "")
```



