---
title: "Wyoming Lobbyists"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wy", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the Wyoming Secretary of State's office [Lobbying Center][03]. From there, we
can download "a zip file containing the lobbyist data files for the current period." We'll make a
`httr::GET()` request on the file to download write the raw ZIP archive to disk.

```{r raw_dir}
raw_dir <- here("state","wy", "lobby", "data", "raw")
dir_create(raw_dir)
```


```{r raw_get, eval=FALSE}
lob_url <- "https://lobbyist.wyo.gov/Lobbyist/Download.aspx"
lob_head <- headers(HEAD(lob_url))
lob_file <- str_remove(lob_head[["content-disposition"]], "(.*);\\sfilename=")
lob_path <- str_c(raw_dir, lob_file, sep = "/")
if (!this_file_new(lob_path)) {
  GET(lob_url, write_disk(lob_path, overwrite = TRUE))
  unzip(lob_path, exdir = raw_dir)
}
```


[03]: https://lobbyist.wyo.gov/Lobbyist/Default.aspx

## Import

The `schemaLobbyist.pdf` file outlines the relationship between the three text files included in the
ZIP archive. Using this guide, we can add the principal organizations to each lobbyist.

First, we will use `vroom::vroom()` to read the data frame of lobbyists.

```{r raw_read_lob}
lobs <- vroom(
  .name_repair = make_clean_names,
  file = glue("{raw_dir}/LOBBYIST.txt"),
  delim = "|",
  col_types = cols(
    .default = col_character(),
    REGISTRATION_DATE = col_date_usa(),
    EXPIRATION_DATE = col_date_usa(),
    TERMINATED_DATE = col_date_usa()
  )
)

lobs <- lobs %>% 
  rename(
    lob_id = lobbyist_id,
    lob_num = lobbyist_num,
    lob_status = status_id,
    lob_reg = registration_date,
    lob_exp = expiration_date,
    lob_badge = name_on_badge,
    lob_first = first_name,
    lob_middle = middle_name,
    lob_last = last_name,
    zip = postal_code,
    lob_email = email,
    lob_term = terminated_date
  )

lobs <- lobs %>%
  mutate(lob_status = equals(lob_status, "Active")) %>% 
  rename(lob_active = lob_status)
```

Then, we will read the `LOBBYIST_ORGANIZATION_XREF.txt` file to get the relational keys needed to
add the information from `LOBBYIST_ORGANIZATION.txt` to our lobbyists data frame.

```{r raw_read_xref}
xref <- vroom(
  .name_repair = make_clean_names,
  file = glue("{raw_dir}/LOBBYIST_ORGANIZATION_XREF.txt"),
  delim = "|",
  col_types = cols(
    .default = col_character()
  )
)

xref <- xref %>%
  remove_empty("cols") %>% 
  rename(
    xref_id = lobbyist_organization_xref_id,
    lob_id = lobbyist_id,
    org_id = lobbyist_organization_id
  )
```

```{r raw_read_orgs}
orgs <- vroom(
  .name_repair = make_clean_names,
  file = glue("{raw_dir}/LOBBYIST_ORGANIZATION.txt"),
  delim = "|",
  col_types = cols(
    .default = col_character()
  )
)

orgs <- orgs %>% 
  rename(
    org_id = lobbyist_organization_id,
    org_num = lobbyist_organization_num,
    org_name = name,
    zip = postal_code
  ) %>% 
  mutate(phone = phone %>% str_remove("\\|$"))
```

Finally, we can use `dplyr::*_join()` to combine these three tables into a single data frame with
the full record of a lobbyist and a client relationship.

```{r raw_join}
wylr <- lobs %>% 
  left_join(xref, by = "lob_id") %>% 
  left_join(orgs, by = "org_id", suffix = c("_lob", "_org")) %>% 
  rename_prefix(suffix = c("_lob", "_org"))
```

## Explore

```{r glimpse, echo=FALSE}
head(wylr)
tail(wylr)
glimpse(wylr)
```

Most columns do not contain any missing information. Only 
`r count_na(wylr$lob_phone) + count_na(wylr$lob_email)` lobbyists are missing either their phone
number of email address. These variables do not need to be flagged.

```{r col_missing}
col_stats(wylr, count_na)
```

As you'd expect, some columns are more distinct than others. In our original lobbyist and
organization tables, the respect `*_id` variables are 100% distinct, but lobbyists are repeated
for every client organization in our joined data frame.

```{r col_distinct}
col_stats(wylr, n_distinct)
```

All but one of the lobbyists listed in the database have an "Active" status. That lobbyist is the
only one to have a termination date, as we'd expect.

```{r active_count}
sum(!wylr$lob_active)
prop_na(wylr$lob_term)
prop_na(wylr$lob_term[wylr$lob_active])
```

## Wrangle

To improve the consistency and search ability of our accountability database, we will perform some
simple and **confident** manipulations to the original data and create new, normalized variables.

### Year

First, we will add the year in which each lobbyist was registered.

```{r year_add}
wylr$lob_reg <- mdy(wylr$lob_reg)
wylr$lob_exp <- mdy(wylr$lob_exp)
wylr <- mutate(wylr, lob_year = year(lob_reg))
```

```{r plot_reg_year, echo=FALSE}
wylr %>% 
  count(lob_year) %>%
  ggplot(aes(x = lob_year, y = n)) + 
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  labs(
    title = "Lobbyists Registered in Wyoming per Year",
    subtitle = "Carryover registrations in 2002",
    x = "Registration Year",
    y = "Lobbyists Registered"
  )
```

### Phone

We can use `campfin::normal_phone()` to convert the numeric phone numbers into an unambiguous
character format. This prevents the column from being read as a numeric variable.

```{r phone_norm}
wylr <- mutate_at(
  .tbl  = wylr,
  .vars = vars(ends_with("phone")),
  .funs = list(norm = normal_phone)
)
```

```{r phone_view, echo=FALSE}
wylr %>% 
  select(contains("phone")) %>% 
  distinct() %>% 
  sample_frac()
```

### Address

To normalize the street addresses, we will first `tidyr::unite()` each address column into a single
column and then pass that string to `campfin::normal_address()`.

```{r addr_unite_lob}
wylr <- wylr %>% 
  unite(
    starts_with("lob_addr"),
    col = "lob_addr_full",
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    lob_addr_norm = normal_address(
      address = lob_addr_full,
      abbs = usps_street
    )
  ) %>% 
  select(-ends_with("addr_full"))
```

```{r addr_unite_org}
wylr <- wylr %>% 
  unite(
    starts_with("org_addr"),
    col = "org_addr_full",
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    org_addr_norm = normal_address(
      address = org_addr_full,
      abbs = usps_street
    )
  ) %>% 
  select(-ends_with("addr_full"))
```

```{r}
wylr %>% 
  select(starts_with("lob_addr")) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

Our database uses 5-digit ZIP codes, so we can pass the original postal code variables to 
`campfin::normal_zip()` to trim the strings and try and repair and broken formats.

```{r zip_norm}
wylr <- mutate_at(
  .tbl  = wylr,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_view, echo=FALSE}
wylr %>% 
  select(contains("zip")) %>% 
  distinct() %>% 
  sample_frac() %>% 
  filter(lob_zip %out% valid_zip)
```

This makes out new ZIP variables very clean.

```{r zip_progress, echo=FALSE}
progress_table(
  wylr$lob_zip,
  wylr$lob_zip_norm,
  wylr$org_zip,
  wylr$org_zip_norm,
  compare = valid_zip
)
```

### State

This database contains a mix of full state names and 2-letter abbreviations; we can pass these
variables to `campfin::normal_state()` to try and convert them all the abbreviations.

```{r state_norm}
wylr <- mutate_at(
  .tbl  = wylr,
  .vars = vars(ends_with("state")),
  .funs = list(norm = normal_state),
  abbreviate = TRUE
)

wylr <- wylr %>% 
  mutate_at(
    .vars = vars(ends_with("state_norm")),
    .funs = str_replace, 
    "WY WY", "WY"
  )
```

```{r state_view, echo=FALSE}
wylr %>% 
  select(contains("state")) %>% 
  distinct() %>% 
  sample_frac()
```

```{r state_progress, echo=FALSE}
progress_table(
  wylr$lob_state,
  wylr$lob_state_norm,
  wylr$org_state,
  wylr$org_state_norm,
  compare = valid_state
)
```

### City

The city values are typically the hardest to normalize due to the variety of valid formats. Again,
the `campfin::normal_city()` function reduces inconsistencies and removes invalid values.

```{r city_norm}
wylr <- mutate_at(
  .tbl  = wylr,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  na = invalid_city
)
```

Then, we can compare these normalized values to the _expected_ values for that record's ZIP code.
If the two values are similar, we can confidently assume a typo was made and default to the
expected value.

```{r swap_city_org}
wylr <- wylr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "org_state_norm" = "state",
      "org_zip_norm" = "zip"
    )
  ) %>% 
  rename(org_city_match = city) %>% 
  mutate(
    org_match_abb = is_abbrev(org_city_norm, org_city_match),
    org_match_dist = str_dist(org_city_norm, org_city_match),
    org_city_swap = if_else(
      condition = org_match_abb | org_match_dist == 1,
      true = org_city_match,
      false = org_city_norm
    )
  ) %>% 
  select(
    -org_city_match,
    -org_match_abb,
    -org_match_dist
  )
```

Our relatively few city values were already very clean, but this process was able to make some
quick and easy improvements.

```{r city_progress, echo=FALSE}
progress_table(
  str_to_upper(wylr$lob_city),
  wylr$lob_city_norm,
  str_to_upper(wylr$org_city),
  wylr$org_city_norm,
  wylr$org_city_swap,
  compare = c(valid_city, extra_city)
)
```

Now we can remove the normalized city column in favor of our improved compared value.

```{r city_rename}
wylr <- wylr %>% 
  select(-org_city_norm) %>% 
  rename(org_city_norm = org_city_swap)
```

Even the few remaining values are actually valid and are just absent from our list.

```{r city_bad}
wylr %>% 
  filter(org_city_norm %out% valid_city) %>% 
  count(org_zip_norm, org_city_norm, sort = TRUE)
```

## Export

```{r proc_dir}
proc_dir <- here("state","wy", "lobby", "data", "processed")
dir_create(proc_dir)
```

```{r proc_write}
write_csv(
  x = wylr,
  path = glue("{proc_dir}/wy_lobby_reg.csv"),
  na = ""
)
```

