---
title: "Nebraska Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ne", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # print markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Nebraska Accountability and Disclosure Commission (NADC)][nadc].

> This site includes the campaign statements filed with the Commission by
candidates for state and local offices, political action committees or PACs,
political parties, and ballot question committees. Campaign statements disclose
the amount and sources of contributions to the committee and the amount purposes
of expenditures by the committee.

[nadc]: http://www.nadc.nebraska.gov/

The data itself can be downloaded from the [Campaign Statements Search][search]
page, where there is a [direct download link][dl].

[search]: http://www.nadc.nebraska.gov/ccdb/search.cgi
[dl]: http://www.nebraska.gov/nadc_data/nadc_data.zip

## Import

To import the file for processing, we will simply download the ZIP archive,
extract the appropriate files, and read them into R.

### Download

```{r raw_download}
raw_dir <- dir_create(here("ne", "contribs", "data", "raw"))
raw_url <- "http://www.nebraska.gov/nadc_data/nadc_data.zip"
raw_zip <- path(raw_dir, basename(raw_url))
download.file(raw_url, raw_zip)
as_tibble(unzip(raw_zip, list = TRUE))
```

```{r raw_unzip}
raw_files <- unzip(raw_zip, exdir = raw_dir)
zip_dir <- path(raw_dir, "nadc_data")
dir_ls(zip_dir) %>% file_move(raw_dir)
raw_all <- dir_ls(raw_dir, regexp = "txt$")
dir_delete(zip_dir)
file_delete(raw_zip)
```

### Read

The records of interest are spread across a number of seperate files, the 
descriptions of which are contained in the `nadc_tables.rtf` file. Furthermore,
the variables describing filing committees are contained in additional files.
We find read all these files and use a combination of `dplyr::left_join()` and
`dplyr::bind_rows()` to create a single table with all the data we want.

From reading the tables document, it seems like these files would contain the
relevant data:

* `FORMA1` – Statement of Organization of a Political Committee:
* `FORMB1` – Campaign Statement for Candidate and Ballot Question Committee:
* `FORMB1AB` – All Data from form B-1 schedule A and B, Contributions:
    * `Committee ID` – Use to link with `FORMb1`, along with Date Received
* `FORMB2` – From B-2 Campaign Statement Political Party Committees:
* `FORMB2A` – Form B-2 Contributions:
* `FORMB4` – Campaign Statement for Independent Committees:
* `FORMB4A` – Form B-4 Schedule A:
    * `Committee ID`  - Use to link with `FORMB4`, along with Date Received
* `FORMB5` – Report of Late contributions:
* `FORMB7` – Report of Political Contributions of Corporation, Union or Other:
* `FORMB72` – Form B-7 Direct contributions:
* `FORMB73` – Form B-7 Indirect contributions:
* `FORMB9` – Out of State Contribution Report:
* `FORMB9A` – Form B-9 Schedule A, contributions:
    * `Form B9 ID` – Use to link with `FORMB9`

First we will read _every_ file into a list of data frames with `purrr::mpa()`
and `vroom::vroom()`.

```{r all_read}
nadc <- map(
  .x = raw_all,
  .f = vroom,
  .name_repair = make_clean_names,
  delim = "|",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(.default = "c")
)
```

```{r all_spec_rm, echo=FALSE}
for (i in seq_along(nadc)) {
  attr(nadc[[i]], "spec") <- NULL
}
```

Then we can go about renaming, binding, and joining these various data frames.

```{r all_rename}
names(nadc) <- 
  names(nadc) %>% 
  basename() %>% 
  path_ext_remove() %>% 
  str_remove("form")
```

```{r all_glimpse}
glimpse(nadc)
```
