---
title: "Missouri Voters"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("mo", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Missouri Centralized Voter Registration (MCVR) data was obtained as an open
record pursuant to Section 115.157, RSMo. from the Elections Division, Office of
Secretary of State.

The MCVR data was provided as a ZIP archive through a file sharing site.

```{r raw_dir}
raw_dir <- dir_create(here("mo", "voters", "data", "raw"))
raw_zip <- path(raw_dir, "Files.zip")
raw_out <- as_fs_path(unzip(raw_zip, exdir = raw_dir))
```

The archive contains a README file to explain the data:

```{r readme_about, echo=FALSE, results='asis'}
readme <- read_lines(raw_out[3])
md_quote(readme[1:5])
```

The README also contains a disclaimer:

```{r readme_disclaim, echo=FALSE, results='asis'}
md_quote(readme[18:21])
```

And a record layout describing the columns of the file:

```{r readme_layout}
readme[31:83] %>% 
  str_replace("(?<=\\d)  ", "\t") %>% 
  str_replace("\t{2,}", "\t") %>% 
  read_tsv(col_names = c("Position", "Name", "Description")) %>% 
  mutate(across(Name, md_code)) %>% 
  select(-Position) %>% 
  kable()
```

## Read

We can read the tab-delimited file as a dataframe.

```{r raw_read}
mov <- read_delim(
  file = raw_out[1],
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "NA", "XXXXX"),
  col_types = cols(
    .default = col_character(),
    Birthdate = col_date_usa(),
    `Registration Date` = col_date_usa()
  )
)
```

There are 20 columns at the end of the dataframe containing all of the past
elections in which each person has voted. We are going to keep the most recent
election and then save all the columns as a separate data frame. This data frame
will be kept in a _long_ format, with a row for every election.

```{r voter_split, eval=FALSE}
hist_file <- path(dirname(raw_dir), "vote_history.csv")
if (file_exists(hist_file)) {
  vote_hist <- vroom(
    file = hist_file,
    col_types = cols(
      voter_id = col_character(),
      order = col_integer(),
      date = col_date(),
      election = col_character()
    )
  )
} else {
  vote_hist <- select(mov, `Voter ID`, starts_with("Voter History"))
  vote_hist <- pivot_longer(
    data = vote_hist,
    cols = starts_with("Voter History"),
    names_to = "order",
    values_to = "election"
  )
  vote_hist <- vote_hist %>% 
    clean_names("snake") %>% 
    filter(!is.na(election))
  vote_hist <- separate(
    data = vote_hist,
    col = election,
    sep = "(?<=\\d)\\s",
    into = c("date", "election")
  )
  vote_hist <- mutate(
    .data = vote_hist,
    order = as.integer(str_extract(order, "\\d+")),
    date = parse_date(date, "%m/%d/%Y")
  )
  write_csv(
    x = vote_hist,
    path = hist_file
  )
}
```

We can then remove the election columns.

```{r voter_remove}
mov <- rename(mov, `Last Election` = `Voter History 1`)
mov <- select(mov, -starts_with("Voter History"))
```

```{r clean_names}
mov <- clean_names(mov, "snake")
```

```{r raw_rename}
mov <- mov %>% 
  rename_all(str_remove, "residential_") %>% 
  rename_all(str_remove, "_new") %>% 
  rename(
    zip = zip_code,
    birth_date = birthdate,
    reg_date = registration_date
  )
```

## Explore

There are `r scales::comma(nrow(mov))` rows of `r ncol(mov)` columns.

```{r glimpse}
glimpse(mov)
tail(mov)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(mov, count_na)
```

No records are missing a name or registration date.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
voter_ids <- mov$voter_id
mov <- select(mov, -voter_id)
d1 <- duplicated(mov, fromLast = FALSE)
d2 <- duplicated(mov, fromLast = TRUE)
mov <- mov %>% 
  mutate(voter_id = voter_ids, .before = first_name) %>% 
  mutate(dupe_flag = d1 | d2)
```

```{r dupe_view}
percent(mean(mov$dupe_flag), 0.001)
mov %>% 
  filter(dupe_flag) %>% 
  select(voter_id, first_name, last_name, birth_date, reg_date)
```

### Categorical

```{r distinct_count}
col_stats(mov, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(mov, county)
explore_plot(mov, congressional)
explore_plot(mov, state_senate)
explore_plot(mov, voter_status)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
mov <- mutate(mov, reg_year = year(reg_date))
```

```{r date_range}
min(mov$reg_date)
sum(mov$reg_year < 2000)
max(mov$reg_date)
sum(mov$reg_date > today())
```

```{r bar_year, echo=FALSE}
mov %>% 
  count(reg_year) %>% 
  mutate(even = is_even(reg_year)) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1960, 2020, by = 20)) +
  coord_cartesian(xlim = c(1960, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Missouri Voters by Registration Year",
    caption = "Source: MDVR",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

We can create a single, unified normalized address field.

```{r addr_unite}
mov <- mov %>% 
  unite(
    col = address_norm,
    house_number:non_standard_address,
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  relocate(address_norm, .after = last_col())
```

```{r addr_view}
mov %>% 
  select(house_number:non_standard_address, address_norm) %>% 
  sample_n(20) %>% 
  remove_empty("cols")
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mov <- mov %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mov$zip,
  mov$zip_norm,
  compare = valid_zip
)
```

### State

As we would expect, all the Missouri voters live in Missouri.

```{r state_count}
count(mov, state, sort = TRUE)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mov <- mov %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("MO", "DC", "MISSOURI"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
mov <- mov %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c("state", "zip_norm" = "zip")
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- mov %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
mov <- mov %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(mov$city_raw),
  mov$city_norm,
  mov$city_swap,
  mov$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(mov$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Missouri City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Missouri City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
mov <- mov %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(mov, 50))
```

1. There are `r comma(nrow(mov))` records in the database.
1. There are `r comma(sum(mov$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mov$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("mo", "voters", "data", "clean"))
clean_path <- path(clean_dir, "mo_voters.csv")
write_csv(mov, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(mov)),
  type = md_code(map_chr(mov, typeof)),
  def = c(
    "Voter county registered in",
    "**Unique** Voter ID",
    "Voter first name",
    "Voter middle name",
    "Voter last name",
    "Voter name suffix",
    "House number on street",
    "House number suffix (letter)",
    "Address direction prefix",
    "Address street name",
    "Address street type",
    "Address direction suffix",
    "Address unit type",
    "Address unit number",
    "Non-standard full address",
    "Voter residential city",
    "Voter residential state",
    "Voter residential ZIP code",
    "Mailing address full",
    "Mailing city",
    "Mailing state",
    "Mailing ZIP code",
    "Voter date of birth",
    "Voter date registered",
    "Precinct number",
    "Precinct name",
    "Precinct split",
    "Township level name",
    "Ward number",
    "Congressional district",
    "State legislative district",
    "State senate district",
    "Voter status (active, inactive)",
    "Last election voted in",
    "Flag indicating duplicate record",
    "Year registered to vote",
    "Normalized combined address",
    "Normalized 5-digit ZIP code",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
