---
title: "Maryland Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("md", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Maryland [Campaign Reporting System][crs].

[crs]: https://campaignfinance.maryland.gov/Public/ViewReceipts?theme=vista

> A contribution is money or anything of value given to a political committee to
promote or assist in promoting the success or defeat of a candidate, political
party, or question. A contribution can be made by any individual or by an
entity, including a corporation or other business entity, a political club, a
federal committee, or a labor union...

> A person may contribute directly or indirectly no more than $6,000 to one
political committee, and a total of $24,000 to all political committees, during
the four-year cycle.

## Import

### Download

```{r raw_dir}
raw_dir <- dir_create(here("md", "contribs", "data", "raw"))
```

```{r raw_download}
raw_csv <- path(raw_dir, "ContributionsList.csv")
if (!file_exists(raw_csv)) {
  response <- GET(
    url = "https://campaignfinance.maryland.gov/Public/ExportCsv",
    write_disk(raw_csv, overwrite = TRUE),
    query = list(
      page = "1",
      orderBy = "~",
      filter = "~",
      `Grid-size` = 15,
      theme = "vista"
    )
  )
}
```

### Read

```{r raw_lines}
fix_csv <- file_temp(ext = "csv")
read_file(raw_csv) %>% 
  str_remove_all("(?<!((Electoral|Administrative|,),))\n") %>% 
  str_replace_all("\"", "\'") %>% 
  str_remove_all(",(?=\n|\n$)") %>% 
  write_lines(fix_csv)
```

```{r raw_read}
mdc <- read_delim(
  file = fix_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Contribution Date` = col_date_usa(),
    `Contribution Amount` = col_double()
  )
)
```

```{r raw_rename, echo=FALSE}
mdc <- remove_empty(mdc, "cols") 
names(mdc) <- c(
  "committee", "period", "date", "name", "address", "type", 
  "method", "amount", "employer", "occupation", "office", "fund"
)
```

## Wrangle

```{r address_split}
mdc <- mdc %>%
  mutate(
    address = address %>% 
      str_remove("-(\\s+|\\d{4})$") %>% 
      str_trim()
  ) %>% 
  separate(
    col = address,
    into = c(glue("address{1:10}"), "city", "state_zip"),
    sep = "\\s{2,}",
    extra = "merge",
    fill = "left"
  ) %>% 
  unite(
    starts_with("address"),
    col = "address",
    na.rm = TRUE,
    sep = " "
  ) %>% 
  mutate_if(is_character, na_if, "") %>% 
  separate(
    col = state_zip,
    into = c("state", "zip"),
    sep = "\\s(?=\\d)",
    extra = "merge",
    fill = "right"
  )
```

```{r}
zip_states <- str_which(mdc$state, "\\d+(?:-\\d+|\\s+)?$")
mdc$zip[zip_states] <- mdc$state[zip_states]
mdc$state[zip_states] %>% 
  str_remove_all("\\d+") %>% 
  str_remove_all("[:punct:]") %>% 
  str_trim() %>% 
  na_if("") -> mdc$state[zip_states]
```

## Explore

```{r glimpse}
head(mdc)
tail(mdc)
glimpse(sample_n(mdc, 20))
```

### Missing

```{r count_na}
col_stats(mdc, count_na)
```

```{r flag_na}
mdc <- mdc %>% flag_na(date, name, amount, committee)
percent(mean(mdc$na_flag), accuracy = 0.01)
sum(mdc$na_flag)
```

### Duplicates

```{r flag_dupes}
mdc <- flag_dupes(mdc, everything())
percent(mean(mdc$dupe_flag), accuracy = 0.01)
mdc %>% 
  filter(dupe_flag) %>% 
  select(date, name, amount, committee)
```

### Categorical

```{r n_distinct}
col_stats(mdc, n_distinct)
```

```{r type_bar}
explore_plot(
  data = mdc,
  var = type,
  title = "Maryland Contributions by Contributor Type"
)
```

```{r type_bar}
explore_plot(
  data = mdc,
  var = method,
  title = "Maryland Contributions by Method"
)
```

```{r type_bar}
explore_plot(
  data = filter(mdc, !is.na(occupation)),
  var = occupation,
  title = "Maryland Contributions by Contributor Occupation"
)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(mdc$amount)
mean(mdc$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
mdc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Maryland Contribution Amount Distribution",
    caption = "Source: Maryland Campaign Reporting System",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
mdc <- mutate(mdc, year = year(date))
```

```{r date_range}
min(mdc$date)
sum(mdc$year < 2000)
max(mdc$date)
sum(mdc$date > today())
```

```{r year_bar, echo=FALSE}
mdc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maryland Contribution by Year",
    caption = "Source: Maryland Campaign Reporting System",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Normalize

### Address

```{r address_norm}
mdc <- mdc %>% 
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
mdc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(20)
```

### ZIP

```{r zip_norm}
mdc <- mdc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  mdc$zip,
  mdc$zip_norm,
  compare = valid_zip
)
```

### State

```{r state_norm}
mdc <- mdc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_view}
mdc %>% 
  filter(state != state_norm) %>% 
  count(state, sort = TRUE)
```

```{r state_progress}
progress_table(
  mdc$state,
  mdc$state_norm,
  compare = valid_state
)
```

### City

```{r city_norm}
mdc <- mdc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("DF", "DC", "DATA FRAME"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r city_swap}
mdc <- mdc %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r refine_city}
good_refine <- mdc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r view_city_refines, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

```{r join_refine}
mdc <- mdc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

```{r city_progress}
progress_table(
  mdc$city,
  mdc$city_norm,
  mdc$city_swap,
  mdc$city_refine,
  compare = valid_city
)
```

## Conclude

1. There are `r nrow(mdc)` records in the database.
1. There are `r sum(mdc$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(mdc$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r proc_dir}
proc_dir <- dir_create(here("md", "contribs", "data", "processed"))
```

```{r write_clean}
write_csv(
  x = df,
  path = path(proc_dir, "df_type_clean.csv"),
  na = ""
)
```

