---
title: "Maryland Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("md", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Maryland [Campaign Reporting System (CRS)][crs].

[crs]: https://campaignfinance.maryland.gov/Public/ViewReceipts?theme=vista

> A contribution is money or anything of value given to a political committee to
promote or assist in promoting the success or defeat of a candidate, political
party, or question. A contribution can be made by any individual or by an
entity, including a corporation or other business entity, a political club, a
federal committee, or a labor union...

> A person may contribute directly or indirectly no more than $6,000 to one
political committee, and a total of $24,000 to all political committees, during
the four-year cycle.

## Download

As of now, this `GET()` request fails to return anything but an empty file. To
download the file manually, go to the [CRS search poral][search] and run a
query for everything from January 1st 2000 to January of the 2020.

This will return a list of 5,407,055 results which can be downloaded as a CSV
file to the disk.

[search]: https://campaignfinance.maryland.gov/Public/ViewReceipts?theme=vista

```{r raw_dir}
raw_dir <- dir_create(here("md", "contribs", "data", "raw"))
```

```{r raw_download}
raw_csv <- path(raw_dir, "ContributionsList.csv")
if (!file_exists(raw_csv)) {
  response <- GET(
    url = "https://campaignfinance.maryland.gov/Public/ExportCsv",
    write_disk(raw_csv, overwrite = TRUE),
    query = list(
      page = "1",
      orderBy = "~",
      filter = "~",
      `Grid-size` = 15,
      theme = "vista"
    )
  )
}
```

## Read

To properly read the file into R, we first have to do some simple string
processing to the text file.

```{r raw_lines}
fix_csv <- path(here("md", "contribs", "data"), "fix_file.csv")
if (!file_exists(fix_csv)) {
  read_file(raw_csv) %>% 
    str_remove_all("(?<!((Electoral|Administrative|,),))\n") %>% 
    str_replace_all("\"", "\'") %>% 
    str_remove_all(",(?=\n|\n$)") %>% 
    write_lines(fix_csv)
}
```

```{r raw_read}
# 5,407,055
mdc <- read_delim(
  file = fix_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "NA", "N/A", "NULL", "Not Included"),
  col_types = cols(
    .default = col_character(),
    `Contribution Date` = col_date_usa(),
    `Contribution Amount` = col_double()
  )
)
```

```{r raw_rename}
mdc <- mdc %>% 
  remove_empty("cols") %>% 
  rename(
    committee = `Receiving Committee`,
    period = `Filing Period`,
    date = `Contribution Date`,
    name = `Contributor Name`,
    address = `Contributor Address`,
    type = `Contributor Type`,
    method = `Contribution Type`,
    amount = `Contribution Amount`,
    employer = `Employer Name`,
    occupation = `Employer Occupation`,
    office = `Office`,
    fund = `Fundtype`
  )
```

## Separate

To better map the columns of the database, we will have to `separate()` the
`address` column into it's components.

```{r sep_address}
mdc <- mdc %>%
  mutate(
    address = address %>% 
      str_remove("-(\\s+|\\d+)$") %>% 
      # address in DC with quad in sep col
      str_replace("(\\s{2})(?=(NW|NE|SW|SE)\\.?\\s{2})", " ") %>% 
      str_replace("(\\s{2})(?=(N.W.|N.E.|S.W.|S.E.)\\.?\\s{2})", " ")
  ) %>% 
  separate(
    col = address,
    into = c(glue("split_addr{1:10}"), "city_sep", "state_zip"),
    sep = "\\s{2,}",
    remove = FALSE,
    extra = "merge",
    fill = "left"
  ) %>% 
  unite(
    starts_with("split_addr"),
    col = "address_sep",
    na.rm = TRUE,
    sep = " "
  ) %>% 
  mutate_if(is_character, na_if, "") %>% 
  separate(
    col = state_zip,
    into = c("state_sep", "zip_sep"),
    sep = "\\s(?=\\d)",
    extra = "merge",
    fill = "right"
  ) %>% 
  separate(
    col = name,
    into = c("last", "first"),
    sep = "\\s{2}",
    extra = "merge",
    fill = "right"
  )
```

This process does not work for the few addresses missing a state abbreviation or
some other geographic variable within the string. We can identify any extracted
ZIP codes in the new `state` column and adjust their position appropriately.

```{r sep_view}
mdc %>% 
  select("address", ends_with("_sep")) %>% 
  filter(state_sep %out% valid_state, !is.na(state_sep)) %>% 
  sample_frac()
```

```{r}
mdc <- mutate_if(mdc, is_character, na_if, "Null Zip Code")
```

```{r sep_fix}
# identify zips in state
zip_states <- str_which(mdc$state_sep, "\\d+(?:-\\d+|\\s+)?$")
# shift those cols left
mdc[zip_states, ] %>% select("address", ends_with("_sep"))
mdc$zip_sep[zip_states] <- mdc$state_sep[zip_states]
mdc$state_sep[zip_states] %>% 
  str_remove_all("\\d+") %>% 
  str_remove_all("[:punct:]") %>% 
  str_trim() %>% 
  na_if("") -> mdc$state_sep[zip_states]
mdc[zip_states, ] %>% select("address", ends_with("_sep"))
```

```{r}
addr_city <- which(str_starts(mdc$city_sep, "\\d") & is.na(mdc$address_sep))
mdc[addr_city, ] %>% select("address", ends_with("_sep"))
mdc$address_sep[addr_city] <- mdc$city_sep[addr_city]
mdc$city_sep[addr_city] <- NA
```

```{r}
addr_state <- which(str_starts(mdc$state_sep, "\\d") & is.na(mdc$address_sep))
mdc[addr_state, ] %>% select("address", ends_with("_sep"))
mdc$address_sep[addr_state] <- mdc$state_sep[addr_state]
mdc$state_sep[addr_state] <- NA
```

```{r}
city_state <- which(nchar(mdc$state_sep) > 2 & is.na(mdc$city_sep))
mdc[city_state, ] %>% select("address", ends_with("_sep"))
city_abb <- str_extract(mdc$state_sep[city_state], "(?<=\\s)[:upper:]{2}$")
city_name <- str_remove(mdc$state_sep[city_state], "\\s[:upper:]{2}$")
mdc$city_sep[city_state] <- city_name
mdc$state_sep[city_state] <- city_abb
mdc$state_sep[city_state] <- NA
```

```{r}
bad_st <- str_to_upper(mdc$state_sep) %out% valid_state
nan_st <- !is.na(mdc$state_sep)
add_st <- str_ends(mdc$city_sep, "\\d")
addr2_city <- which(bad_st & nan_st & add_st)

mdc$address_sep[addr2_city] <- mdc$address_sep[addr2_city] %>% 
  paste(mdc$city_sep[addr2_city]) %>% 
  str_remove("NA\\s")
mdc$city_sep[addr2_city] <- mdc$state_sep[addr2_city]
mdc$state_sep[addr2_city] <- str_extract(mdc$state_sep[addr2_city], "[:upper:]{2}$")
```

```{r}
mdc %>% 
  filter(
    str_to_upper(state_sep) %out% valid_state,
    !is.na(state_sep)
  ) %>% 
  select("address", ends_with("_sep"))
```

```{r}
mdc <- mutate_if(mdc, is_character, ~str_squish(str_trim(.)))
```

```{r echo=FALSE}
rm(
  zip_states,
  addr_city,
  addr_state,
  city_state,
  bad_st,
  nan_st,
  add_st,
  addr2_city
)
```

## Explore

```{r glimpse}
head(mdc)
tail(mdc)
glimpse(sample_n(mdc, 20))
```

### Missing

```{r count_na}
col_stats(mdc, count_na)
```

```{r flag_na}
mdc <- mdc %>% flag_na(date, last, amount, committee)
percent(mean(mdc$na_flag), 0.01)
sum(mdc$na_flag)
```

### Duplicates

There are quite a lot of duplicate records in the database. We can flag these
variables with `campfin::flag_dupes()` but not remove them.

```{r flag_dupes}
mdc <- flag_dupes(mdc, everything())
percent(mean(mdc$dupe_flag), 0.01)
mdc %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, committee)
```

### Categorical

```{r n_distinct}
col_stats(mdc, n_distinct)
```

```{r period_bar, echo=FALSE}
explore_plot(
  data = filter(mdc, !is.na(period)),
  var = period,
  title = "Maryland Contributions by Period"
) + scale_x_discrete(label = function(x) str_trunc(x, width = 15))
```

```{r type_bar, echo=FALSE}
explore_plot(
  data = filter(mdc, !is.na(type)),
  var = type,
  title = "Maryland Contributions by Contributor Type"
) + scale_x_discrete(label = function(x) str_trunc(x, width = 15))
```

```{r method_bar, echo=FALSE}
explore_plot(
  data = filter(mdc, !is.na(method)),
  var = method,
  title = "Maryland Contributions by Method"
) + scale_x_discrete(label = function(x) str_trunc(x, width = 15))
```

```{r occupation_bar, echo=FALSE}
explore_plot(
  data = filter(mdc, !is.na(occupation)),
  var = occupation,
  title = "Maryland Contributions by Contributor Occupation"
) + scale_x_discrete(label = function(x) str_trunc(x, width = 15))
```

```{r office_bar, echo=FALSE}
explore_plot(
  data = filter(mdc, !is.na(office)),
  var = office,
  title = "Maryland Contributions by Office"
) + scale_x_discrete(label = function(x) str_trunc(x, width = 15))
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(mdc$amount)
mean(mdc$amount <= 0)
```

```{r hist_amount, echo=FALSE}
mdc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Maryland Contributions Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
mdc <- mutate(mdc, year = year(date))
```

```{r date_range}
min(mdc$date)
sum(mdc$year < 2000)
max(mdc$date)
sum(mdc$date > today())
```

```{r year_bar, echo=FALSE}
mdc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 2)) +
  coord_cartesian(xlim = c(2005, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maryland Contribution by Year",
    caption = "Source: Maryland Campaign Reporting System",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
mdc <- mdc %>% 
  mutate(
    address_norm = normal_address(
      address = address_sep,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
mdc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
mdc <- mdc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip_sep,
      na_rep = TRUE,
      na = c("", "Null Zip Code")
    )
  )
```

```{r zip_progress}
progress_table(
  mdc$zip_sep,
  mdc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r}
mdc %>% 
  count(state_sep, sort = TRUE) %>% 
  filter(state_sep %out% valid_state)
```

```{r state_norm}
mdc <- mdc %>% 
  mutate(
    state_norm = normal_state(
      state = state_sep,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
mdc %>% 
  filter(state_sep != state_norm) %>% 
  count(state_sep, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  mdc$state_sep,
  mdc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
mdc <- mdc %>% 
  mutate(
    city_norm = normal_city(
      city = city_sep %>% str_remove("(?<=Baltimore)\\sCity$"), 
      abbs = usps_city,
      states = c("MD", "DC", "MARYLAND"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
mdc <- mdc %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- mdc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
mdc <- mdc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r check_filter}
mdc_out <- mdc %>% 
  filter(city_refine %out% c(valid_city, extra_city)) %>% 
  count(city_refine, state_norm, sort = TRUE) %>% 
  drop_na() %>% 
  slice(1:500)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("md", "contribs", "data", "api_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file
  )
} else {
  check <- pmap_dfr(
    .l = list(
      mdc_out$city_refine, 
      mdc_out$state_norm
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    path = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  c(valid_locality)
```

#### Progress

```{r city_remains}
many_city <- c(valid_city, extra_city, valid_locality)
mdc %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, sort = TRUE)
```

```{r city_fix}
mdc <- mdc %>% 
  mutate(
    city_refine = city_refine %>% 
      str_replace("^BALTO$", "BALTIMORE")
  )
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(mdc$city_sep),
  mdc$city_norm,
  mdc$city_swap,
  mdc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(mdc$city_sep, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Maryland City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Maryland City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
mdc <- mdc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(mdc, 20))
```

1. There are `r comma(nrow(mdc))` records in the database.
1. There are `r comma(sum(mdc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mdc$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("md", "contribs", "data", "clean"))
```

```{r clean_write}
md_file <- path(clean_dir, "md_contribs_clean.csv")
write_csv(mdc, path = md_file, na = "")
```

## Encoding

```{r check_encoding}
Encoding(md_file)
```

