---
title: "Nevada Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("nv", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("nv/voters/docs/nv_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The Nevada Secretary of State (NVSOS) office requires that one register for an
account to access "[bulk data download][bulk]" service page.

The process for downloaded a report is [outlined here][guide].

Create a report for "Full Unabridged Database Dump" of "Eligible Voter List" 
data.

> ### [Eligible Voter List][evl]
> 
> The Help America Vote Act of 2002 (42U.S.C. § 15301 et. Seq.) and NRS 293.675
require the Secretary of State (SoS) to establish and maintain an official
statewide voter registration list.
> 
> The statewide voter registration list is maintained in a single, uniform,
centralized and interactive computerized database administered and maintained by
the SoS.
> 
> Registration information held solely in a county or local database is not
considered part of the statewide voter registration list until it is uploaded to
the SoS statewide voter registration database. This upload and download of
registration information between the counties and the SoS does not take place in
real time. All uploaded voter registration information is made available within
24 hours of receipt. Registration information reflected in the statewide voter
registration list may not duplicate the registration information reflected in
the county voter registration records.
> 
> "Active registration" status means a current registration of a voter in the
official register, entitling such voter to voter in the manner provided pursuant
to NRS 293.017.
> 
> "Inactive voter" means a voter designated as inactive pursuant to NRS 293.530
whose registration has not been cancelled.
> 
> An "eligible voter" is a registered voter that is legally entitled to vote.
The statewide eligible voter list contains both active and inactive voters.
> 
> Voters with a Pend, Cancelled, P-17 or Void status will not be reflected in
the statewide eligible voter list.

### Report

The site allows users to define the format for their data download. The site
generated the following summary of our specified data format:

> If "Text File - Fixed Width" is selected your report results will be inserted
into a standard ASCII text file where each field starts at a specific "fixed"
position for each line. For more specific information about the report format
for custom built reports, including the position and data type of each field,
click the "View Selected File Structure" button on the General tab when you
view\edit your report.

This file structure report is an HTML page with a description and six tables.

>  Your report will generate 2 fixed width ASCII text file(s) compressed into
one Zip file named in the format `VoterList.43993.<Today's Date>.zip`*. Below
you will find the format of each file: ...

```{r about_file, echo=FALSE}
st_dir <- here("nv", "voters")
about_path <- path(st_dir, "File Format - Secretary of State, Nevada.html")
about <- read_html(x = about_path)
```

```{r about_tables, echo=FALSE}
about_tables <- about %>% 
  html_nodes(".entryform") %>% 
  html_table(fill = TRUE) %>% 
  map(as_tibble)
about_tables <- about_tables[[2]]
```

```{r about_tables_fix, echo=FALSE}
about_tables <- row_to_names(about_tables, row_number = 1)
about_names <- str_subset(html_text(html_nodes(about, "b")), "\\d")
```

```{r about_tables_print, echo=FALSE, results='asis'}
about_md <- str_split(kable(about_tables, align = "llrr"), "\n")
md_quote(about_md)
```

>  \*`<Today's Date>` will be the only variable in the file name that will change
and ensure your file names will be unique each time the report is ran. This date
will be the date the report is ran and will be in the format "MMddyyHHmmss".

### Age

> The data being reported off of is no more than 24 hours old. This data is
copied very late each night from live data to minimize the large burden of bulk
reporting on the production system.

### Variables

Definitions for few variables can be found in the "Result Field" tab:

1. County Status:
    > This is the status of the voter as determined by the respective county
    clerk/registrar. This status can be one of the following:
    > * Active – An eligible voter whose residence address is current.
    > * Inactive – A voter whose mail is returned from the address at which he
    > registered to vote and has not provided a County Elections Official 
    > with a new residence address. Inactive voters retain the legal right 
    > to vote, but need not be mailed election material.
    > * P-17 – A voter who was 17 years of age at the time their voter
    > registration was submitted, but who will be age 18 by election day and
    > who has met all other legal requirements to vote.
    
2. County Voter ID:
    > This is the unique identifier used by the county that this voter is
    registered with. This number is only unique for that county and not
    necessarily unique across the entire state-wide list. The numeric VoterID
    column that is included in each voter record in this data download will be
    unique across all voters for the entire state.

3. ID Required: 
    > A `Y` in this column indicates that the registered voter may vote at a 
    polling place only if the person presents to the election board officer at 
    the polling place official identification to establish identity as defined 
    in NAC 293.395. An `N` in this column indicates that the registered voter 
    does not need to provide identification.

4. Phone:
    > The phone number of the registered voter. If the voter has enrolled in the
    Please Don't Call list, the text "-PDC" will be appended to the end of the 
    phone number.

5. Registration Date:
    > It is possible that the registration date shown may not reflect the
    individual’s original registration date in Nevada due to movement between
    counties and other periodic updates to the particular voter registration 
    record.
    > 
    > If any questions arise regarding particular registration dates in regards 
    to specific individuals, please contact the respective county 
    clerk/registrar for clarification.

### Records

> Total number of records returned: 10,329,395

```{r rows_total}
total_rows <- 10329395 
```

> A record is one single entity or row from a database table. The "Total number
of records returned" displayed on the report preview page will be a summation of
all rows returned from each table you are reporting from. For example, if your
report queries for both Resident Agent and Corporation data from the
Corporations database, the number of records returned might be 1000, 700 of
which might be Corporation records and 300 being Resident Agent records.

## Download

Within seven days of running the report, the data can be downloaded from the
link provided to the account email address. The link will not work for anybody
not logged into that NVSOS account.

```{r raw_dir}
raw_url <- "https://www.nvsos.gov/yourreports/VoterList.43993.040921100238.zip"
raw_dir <- dir_create(here("nv", "voters", "data", "raw"))
raw_zip <- path(raw_dir, basename(raw_url))
file_size(raw_zip)
```

This URL contains the date the report was generated.

```{r raw_time}
report_time <- mdy_hms(str_extract(raw_url, "\\d+(?=\\.zip$)"))
with_tz(report_time, tzone = "PST")
```

This raw ZIP archive has been backed up to the IRW server.

```{r}
as_fs_bytes(object_size(
  object = path("IRW/raw_backup/nv", basename(raw_zip)), 
  bucket = "publicaccountability"
))
```

## Unzip

The provided ZIP archive contains two tables as fixed width text files. We are
only interested in the list of eligible voters.

```{r zip_list}
zip_txt <- as_tibble(unzip(raw_zip, list = TRUE))
raw_txt <- path(raw_dir, zip_txt$Name[1])
```

```{r echo=FALSE, echo=FALSE}
zip_txt %>% 
  mutate(
    across(Length, as_fs_bytes),
    across(Name, md_code)
  ) %>% 
  kable()
```

```{r zip_unzip}
if (!file_exists(raw_txt)) {
  raw_txt <- unzip(
    zipfile = raw_zip, 
    exdir = raw_dir,
    files = basename(raw_zip)
  )
}
```

## Columns

We can use the tables read from the HTML file, and described in the **About**
section above, to create (1) the column width table expected by `read_fwf()`,
and (2) the readr column type specification object. Two functions will
take the `Field Name`, `Data Type` and `Length`.

```{r as_fwf_width, echo=FALSE}
as_fwf_width <- function(.data) {
  fwf_widths(
    widths = as.integer(x = .data[[4]]),
    col_names = str_remove(
      string = .data[[1]], 
      pattern = "\\s\\(.*\\)"
    )
  )
}
```

```{r as_col_spec, echo=FALSE}
as_col_spec <- function(.data) {
  x <- .data[["Data Type"]]
  x <- case_when(
    str_detect(x, "varchar")  ~ "c",
    str_detect(x, "datetime") ~ "D",
    str_detect(x, "money")    ~ "d",
    str_detect(x, "bit")      ~ "l",
    str_detect(x, "int")      ~ "i",
    TRUE ~ "c"
  )
  as.col_spec(
    x = setNames(
      object = x,
      nm = str_remove(
        string = .data[["Field Name"]], 
        pattern = "\\s\\(.*\\)"
      )
    )
  )
}
```

```{r about_format}
raw_widths <- as_fwf_width(about_tables)
raw_types  <- as_col_spec(about_tables)
```

## Read

The eligible voters list can then be read into a list using `readr::read_fwf()` 
and the (1) width tables and (2) column type specifications.

```{r read_raw}
nvv <- read_fwf(
  file = raw_txt,
  col_positions = raw_widths,
  col_types = raw_types,
  locale = locale(
    date_format = "%m/%d/%Y",
    tz = "US/Pacific"
  )
)
```

The total number of rows read matches what we were told when exporting.

```{r rows_read_hist}
hist_wc <- system2(
  command = "wc", 
  args = c(path(raw_dir, zip_txt$Name[2]), "-l"), 
  stdout = TRUE
)
```

```{r rows_check}
nrow(nvv) + as.integer(str_extract(hist_wc, "\\d+")) == total_rows
```

```{r clean_names}
nvv <- nvv %>% 
  clean_names("snake") %>% 
  rename_with(~str_remove(., "residential_")) %>% 
  rename(reg_date = registration_date) %>% 
  mutate(across(id_required, `==`, "Y"))
```

## Explore

There are `r comma(nrow(nvv))` rows of `r ncol(nvv)` columns. Each record
represents a single voter registered in the state of Nevada.

```{r glimpse}
glimpse(nvv)
tail(nvv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(nvv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("first_name", "last_name", "birth_date", "reg_date")
nvv <- flag_na(nvv, all_of(key_vars))
sum(nvv$na_flag)
```

```{r na_view}
nvv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars), middle_name)
```

### Categorical

```{r distinct_count}
col_stats(nvv, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(nvv, county)
explore_plot(nvv, party)
explore_plot(nvv, congressional_district)
explore_plot(nvv, county)
```

### Dates

There are thousands of voters with a registration date of January 1, 1900. These
should be `NA` values, not real dates.

```{r date_min}
count_na(nvv$reg_date)
min(nvv$reg_date, na.rm = TRUE)
sum(nvv$reg_date == "1900-01-01", na.rm = TRUE)
nvv$reg_date[which(nvv$reg_date == "1900-01-01")] <- NA
```

```{r date_max}
max(nvv$reg_date, na.rm = TRUE)
sum(nvv$reg_date > today(), na.rm = TRUE)
```

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
nvv <- nvv %>% 
  mutate(
    reg_year = year(reg_date),
    birth_year = year(birth_date)
  )
```

```{r bar_reg_year, echo=FALSE}
nvv %>% 
  filter(!is.na(reg_year)) %>% 
  count(reg_year) %>% 
  mutate(
    elect_type = case_when(
      reg_year %% 4 == 0 ~ "Pres.",
      reg_year %% 2 == 0 ~ "Mid.",
      TRUE ~ NA_character_
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = elect_type)) + 
  scale_fill_manual(
    values = c("#1B9E77", "#D95F02"),
    na.value = "#666666"
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 4)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nevada Voters Registered by Year",
    caption = "Source: NV SOS",
    fill = "Election Year",
    x = "Year Registered",
    y = "Count"
  )
```

```{r hist_birth_year, echo=FALSE}
nvv %>% 
  filter(!is.na(birth_year)) %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(
    fill = dark2["purple"],
    binwidth = 3
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2005, by = 5)) +
  coord_cartesian(xlim = c(1900, 2005)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Nevada Voters by Birth Year",
    caption = "Source: NV SOS",
    x = "Birth Year",
    y = "Count"
  )
```

## Wrangle

```{r geo_check}
prop_in(nvv$zip, valid_zip)
prop_in(nvv$state, valid_state)
nvv$city <- str_to_upper(nvv$city)
prop_in(nvv$city, c(valid_city, extra_city))
sample(nvv$address_2, 10)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(nvv, 50))
```

1. There are `r comma(nrow(nvv))` records in the database.
1. There are `r comma(sum(nvv$dupe_flag))` duplicate records in the database.
1. The range and of birth and registration dates seem reasonable.
1. There are `r comma(sum(nvv$na_flag))` records missing key variables.
1. Geographic variables are formatted in the way we would expect.
1. The 4-digit year variables have been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("nv", "voters", "data", "clean"))
clean_path <- path(clean_dir, "nv_voters_clean.csv")
write_csv(nvv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
