---
title: "State Data"
author: "First Last"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("kiernann/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel file
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

### About

### Variables

## Import

### Download

```{r create_raw_dir}
raw_dir <- here("ms", "expends", "data", "raw")
dir_create(raw_dir)
```

### Read

```{r read_raw}
ms <- read_excel(
  path = dir_ls(raw_dir),
  .name_repair = make_clean_names,
  col_types = c("text", "text", "date", "text", "numeric")
)
```

## Explore

```{r glimpse}
head(ms)
tail(ms)
glimpse(sample_frac(ms))
```

### Missing

```{r glimpse_na}
glimpse_fun(ms, count_na)
```

```{r flag_na}
ms <- flag_na(ms, everything())
sum(ms$na_flag)
```

### Duplicates

```{r flag_dupes}
ms <- flag_dupes(ms, everything())
sum(ms$dupe_flag)
```

### Categorical

```{r glimpse_distinct}
glimpse_fun(ms, n_distinct)
```

### Continuous

#### Amounts

```{r summary_amount}
summary(ms$amount)
sum(ms$amount <= 0)
```

```{r amount_histogram, echo=FALSE}
ms %>%
  ggplot(aes(amount)) +
  geom_histogram() +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Mississippi Expenditures Amount Distribution",
    caption = "Source: Mississippi Secretary of State",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r add_year}
ms <- mutate(ms, year_received = year(date_received))
```

```{r date_range, collapse=TRUE}
min(ms$date_received)
max(ms$date_received)
sum(ms$date_received > today())
```

```{r clean_date}
ms <- mutate(ms, date_clean = date_received)
ms$date_clean[which(ms$date_received > today() | ms$year_received < 2016)] <- NA
ms <- mutate(ms, year_received = year(date_clean))
```

```{r year_bar, echo=FALSE}
ms %>% 
  count(year_received) %>%
  ggplot(aes(year_received, n)) +
  geom_col() +
  labs(
    title = "Mississippi Expenditures by Year",
    caption = "Source: Mississippi Secretary of State",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

```{r wrangle_upper}
ms <- as_tibble(map_if(ms, is_character, str_to_upper))
```

## Conclude

1. There are `r nrow(ms)` records in the database.
1. There are `r sum(ms$dupe_flag)` duplicate records.
1. There are `r sum(ms$na_flag)` records missing a value.
1. The range and distribution of the `amount` variable seems reasonable.
1. The `date_clean` variable has been created by removing `r sum(ms$date_received > today())` values from the future.
1. The database does not contain geographic data.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r create_proc_dir}
proc_dir <- here("ms", "expends", "data", "processed")
dir_create(proc_dir)
```

```{r write_clean}
write_csv(
  x = ms, 
  path = glue("{proc_dir}/ms_expends_clean.csv"),
  na = ""
)
```

