---
title: "Connecticut Contributions"
author: "Kiernna Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ct", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Campaign contribution data for the state of Connecticut is managed by the
[State Elections Enforcement Commission (SEEC)][seec]. The data can be obtained
from the SEEC [Campaign Reporting Information System (eCRIS)][ecris] public
[search portal][search].

> This page allows the public to search, browse and download information from
campaign finance reports filed by committees with the SEEC’s Disclosure and
Audit Unit. The term committees for purposes of this summary includes: Candidate
committees, Exploratory committees, Party committees and Political Action
committees (also known as PACs). We shall refer to all four distinct committee
types as political committees in order to accent the political nature of their
purpose in relationship to the financing of election campaigns for elective
public office in Connecticut.The Commission strives to offer fast and easy
public access to the filings by committees.

The SEEC also provides a [bulk download page][bulk] with annual CSV files.

> Receipts Data for Party, Political, Candidate and Exploratory Committees
(e-filed in eCRIS and paper report transactions were entered by the State
Election Enforcement Comission staff using a data entry module.)

[seec]: https://portal.ct.gov/seec
[ecris]: https://seec.ct.gov/Portal/eCRIS/eCRISlanding
[search]: https://seec.ct.gov/Portal/eCRIS/eCrisSearch
[bulk]: https://seec.ct.gov/Portal/eCRIS/CurPreYears

## Download

We can scrape this page and find any CSV files containing contribution receipts.

```{r raw_list}
# read the seec cris page
read_html("https://seec.ct.gov/Portal/eCRIS/CurPreYears") %>% 
  # grab all urls on the page
  html_nodes("a") %>% html_attr("href") %>% 
  # filter the receit csv files
  str_subset("Receipts.*(csv|CSV)$") -> raw_urls
```

These URLs will be downloaded to the local raw directory.

```{r raw_download}
raw_dir <- dir_create(here("ct", "contribs", "data", "raw"))
raw_paths <- path(raw_dir, basename(raw_urls))
for (i in seq_along(raw_urls)) {
  if (!file_exists(raw_paths[i])) {
    download.file(
      url = raw_urls[i], 
      destfile = raw_paths[i]
    )
  }
}
```

## Read

There are three types of files, with each varying slightly year to year:

1. Party and PAC receipts
    1. 2008-2013 (23 columns)
    2. 2014-2020 (35 columns)
2. Candidate exploratory committee receipts
    1. 2008-2013 (23 columns)
    2. 2014-2015 (37 columns)
    3. 2016-2020 (41 columns)
3. General committee receipts
    1. 2002-2007 (26 columns)

We will have to do some work to read these files into a single data frame and
match the columns.

### Older files

The oldest files have 26 columns with capitalized names. The 2006 receipts file
lacks column names, so we will read it last and then combine it separately to
the other years.

```{r old_files, results='asis'}
old_files <- str_subset(raw_paths, "Cifs_Receipts")
md_bullet(md_code(basename(old_files)))
```

```{r old_read}
ct_old <- map_df(
  .x = old_files[-2],
  .f = read_delim,
  .id = "source_file",
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    # FILED_TO_STATE_DATE = col_date_usa(),
    # SUBMITTED_DATE = col_date_usa(),
    # CONTRIB_DATE = col_date_usa(),
    CONTRIB_AMOUNT = col_double(),
    LOAN_AMOUNT = col_double(),
    INTEREST_EARNED_AMOUNT = col_double()
  )
)
```

```{r old_fix}
ct_2006 <- read_delim(
  file = old_files[2],
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = names(ct_old)[-1],
  col_types = spec(ct_old)
)
```

```{r old_identify}
ct_old <- relocate(ct_old, source_file, .after = last_col())
ct_old$source_file <- basename(old_files[as.integer(ct_old$source_file)])
ct_2006 <- mutate(ct_2006, source_file = basename(raw_paths[2]))
```

```{r old_bind}
ct_old <- ct_old %>% 
  bind_rows(ct_2006) %>% 
  clean_names("snake") %>% 
  mutate(across(source_file, basename))
```

These files include `r comma(nrow(ct_old))` columns of `r ncol(ct_old)` columns.

```{r echo=FALSE}
ct_old <- ct_old %>% 
  rename(
    committee = committee_name,
    fil_type = registration_type,
    fil_sub = committee_sub_type,
    trans_sub = trans_subcategoryde,
    trans_type = transaction_type_des,
    filed = filed_to_state_date,
    trans_name = trans_display_name,
    is_contrib = is_contribution,
    submitted = submitted_date,
    fil_id = contrib_comid,
    comm_to = committees_committeename,
    method = contrib_method_code,
    date = contrib_date,
    amount = contrib_amount,
    loan = loan_amount,
    inkind_desc = in_kind_contrib_desc,
    inkind_est = in_kind_est_value,
    interest = interest_earned_amount,
    occupation = occup_desc,
    occup_desc = other_occup_desc,
    first = first_name,
    middle = middle_name,
    last = last_name,
    zip = zip_code
  )
```

We can create a single column for the contributor name to make the later files.

```{r old_unite}
ct_old <- ct_old %>% 
  unite(
    col = contributor,
    first, middle, last,
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  relocate(
    contributor,
    .after = last_col()
  ) %>% 
  mutate(
    contributor = na_if(contributor, ""),
    date_clean = mdy(date)
  )
```

### Candidate files

After 2007, contribution receipts were divided between filer types. We can now
read all of the contributions made to candidate exploratory committees.

```{r cand_files, results='asis'}
cand_files <- str_subset(raw_paths, "Candidate")
md_bullet(md_code(basename(cand_files)))
```

```{r cand_read}
ct_cand <- map(
  .x = cand_files,
  .f = read_delim,
  delim = ",",
  na = c("", "NA", "N/A", "NULL", "NO"),
  escape_double = TRUE,
  col_types = cols(
    .default = col_character(),
    `Amount` = col_number(),
    # use lubridate::mdy() instead
    # `Transaction Date` = col_date_usa()
  )
)
```

The number of columns in a file increased over time.

```{r cand_length}
map_dbl(ct_cand, length)
```

However, all the columns present in the old files are present in the newer ones.
Only additional columns are added over time.

```{r cand_cols, echo=FALSE}
cand_names <- names(ct_cand[[1]])
map_dbl(ct_cand, ~mean(names(.) %in% cand_names))
```

That means we can safely `bind_rows()` and create a single dataframe.

```{r cand_bind}
ct_cand <- bind_rows(ct_cand, .id = "source_file")
ct_cand <- relocate(ct_cand, source_file, .after = last_col())
ct_cand$source_file <- basename(cand_files[as.integer(ct_cand$source_file)])
```

```{r cand_rename, echo=FALSE}
ct_cand <- ct_cand %>% 
  clean_names("snake") %>% 
  rename(
    contributor = contributor_name,
    office = office_sought,
    rec_type = receipt_type,
    fil_type = committee_type,
    date = transaction_date,
    filed = file_to_state,
    receipt = receipt_state,
    con_exec = contractor_executive_branch,
    con_leg = contractor_legislative_branch,
    source = data_source,
    refiled = refiled_electronically,
    address = street_address,
    event = fundraising_event,
    report = report_name,
    fil_id = committee_id,
    sec_letter = section_letter,
    sec_name = section_name,
    first = contributor_first_name,
    middle = contributor_middle_initial,
    last = contributor_last_name,
    report_id = report_id,
    method = payment_method
  )
```

```{r try_date}
try_date <- function(date) {
  out <- suppressWarnings(parse_date_time(date, orders = c("mdy", "dmy")))
  out <- as_date(out)
  return(out)
}
```

```{r cand_dates}
date_cols <- which(map_lgl(ct_cand, ~str_detect(.[1], "\\d+/\\d+/\\d")))
ct_cand <- mutate(ct_cand, date_clean = try_date(date))
```

### PAC files

After 2007, contribution receipts were divided between filer types. We can now
read all of the contributions made to candidate exploratory committees.

```{r pac_files, results='asis'}
pac_files <- str_subset(raw_paths, "PAC")
md_bullet(md_code(basename(pac_files)))
```

```{r pac_read}
ct_pac <- map(
  .x = pac_files,
  .f = read_delim,
  delim = ",",
  na = c("", "NA", "N/A", "NULL", "UNKNOWN"),
  escape_double = TRUE,
  col_types = cols(
    .default = col_character(),
    `Amount` = col_number(),
    # use lubridate::mdy() instead
    # `Transaction Date` = col_date_usa()
  )
)
```

The number of columns in a file increased over time.

```{r pac_length}
map_dbl(ct_pac, length)
```

However, all the columns present in the old files are present in the newer ones.
Only additional columns are added over time.

```{r pac_cols, echo=FALSE}
cand_names <- names(ct_pac[[1]])
map_dbl(ct_pac, ~mean(names(.) %in% cand_names))
```

That means we can safely `bind_rows()` and create a single dataframe.

```{r pac_bind}
ct_pac <- map(ct_pac, clean_names)
ct_pac <- bind_rows(ct_pac, .id = "source_file")
ct_pac <- relocate(ct_pac, source_file, .after = last_col())
ct_pac$source_file <- basename(cand_files[as.integer(ct_pac$source_file)])
```

```{r pac_rename, echo=FALSE}
ct_pac <- ct_pac %>% 
  rename(
    contributor = contributor_name,
    office = office_sought,
    rec_type = receipt_type,
    fil_type = committee_type,
    date = transaction_date,
    filed = file_to_state,
    receipt = receipt_state,
    con_exec = contractor_executive_branch,
    con_leg = contractor_legislative_branch,
    source = data_source,
    refiled = refiled_electronically,
    address = street_address,
    event = fundraising_event,
    report = report_name,
    fil_id = committee_id,
    sec_letter = section_letter,
    sec_name = section_name,
    first = contributor_first_name,
    middle = contributor_middle_initial,
    last = contributor_last_name,
    report_id = report_id,
    method = payment_method
  )
```

```{r pac_dates}
ct_pac <- mutate(ct_pac, date_clean = try_date(date))
```

### Bind files

Now we can combine these three different file type dataframes together. There
is sufficient overlap in the columns for our purposes.

```{r bind_names}
prop_in(names(ct_pac), names(ct_cand))
prop_in(names(ct_old), names(ct_cand))
intersect(names(ct_cand), names(ct_old))
```

```{r bind_rows}
ctc <- bind_rows(ct_cand, ct_pac, ct_old, .id = "source_type")
ctc <- relocate(ctc, source_type, .after = last_col())
ctc$source_type <- c("PAC", "CAND", "OLD")[as.numeric(ctc$source_type)]
ctc <- mutate(ctc, across(where(is.character), str_squish))
```

Here, we see how each of the three file types has 

```{r bind_view}
key_vars <- c("contributor", "date_clean", "amount", "committee")
ctc %>% 
  select(all_of(key_vars), source_type) %>% 
  sample_n(20)
```

## Explore

```{r glimpse}
head(ctc)
tail(ctc)
glimpse(sample_n(ctc, 20))
```

## Missing

While most of the files from 2008 to 2020 share the same ~20 columns, some of
the more recent files have an additional ~10 columns. Those files will
contribute empty columns for rows from the earlier files.

```{r na_count}
col_stats(ctc, count_na)
```

We know the variables like `first` and `last` exist for only the more recent
files but simply repeat the information from `contributor`.

Even from the main variables, quite a few are missing values.

```{r na_flag}
ctc <- flag_na(ctc, all_of(key_vars))
percent(mean(ctc$na_flag), 0.1)
```

```{r na_view}
ctc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

Most of these records missing values are missing a date.

```{r na_recount}
ctc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

Not much of these missing values were created when parsing the date.

```{r na_date}
prop_na(ctc$date)
prop_na(ctc$date_clean)
```

## Duplicates

```{r dupe_flag}
ctc <- flag_dupes(ctc, everything())
percent(mean(ctc$dupe_flag), 0.1)
```

```{r dupe_view}
ctc %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

## Amounts

```{r ammount_summary}
summary(ctc$amount)
mean(ctc$amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
ctc %>%
  filter(amount > 1, amount < 1e4) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 10) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Connecticut Contributions Amount Distribution",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_cents}
ctc %>% 
  count(cents = amount %% 1, sort = TRUE) %>% 
  add_prop()
```

## Dates

```{r date_year}
ctc <- mutate(ctc, year_clean = year(date_clean))
```

```{r date_range}
min(ctc$date_clean, na.rm = TRUE)
sum(ctc$year_clean < 2000, na.rm = TRUE)
max(ctc$date_clean, na.rm = TRUE)
sum(ctc$date_clean > today(), na.rm = TRUE)
ctc$year_clean[which(ctc$year_clean > 2020)] <- NA
ctc$year_clean[which(ctc$year_clean < 1998)] <- NA
```

```{r bar_year, echo=FALSE}
ctc %>% 
  filter(!is.na(year_clean)) %>% 
  count(year_clean) %>% 
  mutate(even = is_even(year_clean)) %>% 
  ggplot(aes(x = year_clean, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  coord_cartesian(xlim = c(1998, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut Contributions by Year",
    caption = "Source: SEEC eCRIS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
ctc <- ctc %>%
  mutate(
    address_norm = normal_address(
      address = address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
ctc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
ctc <- ctc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  ctc$zip,
  ctc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
ctc <- ctc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
ctc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  ctc$state,
  ctc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ctc <- ctc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("CT", "DC", "CONNECTICUT"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ctc <- ctc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- ctc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
ctc <- ctc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ctc$city_raw),
  ctc$city_norm,
  ctc$city_swap,
  ctc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ctc$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Connecticut City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
ctc <- ctc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(ctc, 50))
```

1. There are `r comma(nrow(ctc))` records in the database.
1. There are `r comma(sum(ctc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ctc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ct", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "ct_contribs_clean.csv")
write_csv(ctc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```
