---
title: "Wisconsin Expenditures"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wi", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # print markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is from the Wisconsin Secretary of State's Campaign Finance System (CIFS).

> Wyoming's Campaign Finance Information System (WYCFIS) exists to provide a
mechanism for online filing of campaign finance information and to provide full
disclosure to the public. This website contains detailed financial records and
related information that candidates, committees, organizations and parties are
required by law to disclose.

## Download

Using the CIFS [expenditures search portal][search], we can run a search for all
expenditures from "All Filing Periods" from the dates 1980-01-01 to 
`r today()`. Those search results need to be manually exported as the
`ReceiptsList.csv` file. The data includes updates till 2023-05-27. The next update should start 2023-05-28.

> To view expenditures made by a committee, go to the CFIS Home Page, on the left
hand side, and click View Expenses. A pop up letting you know that this
information cannot be used for solicitation purposes will appear – click
Continue. Type in a payer in the field titled ID. Type in any additional information you would
like to search for, including a name of a payer, payee or amount.
To view all expenditures, remove the filing period by clicking, in the Filing
Period Name field, and scroll all the way to the top and select All Filing
Periods. Click Search and all of the expenditures fitting your search criteria
will appear. If you would like to export these into Excel, scroll all the way to
the bottom and on the right hand side, click the XLS icon.

[search]: https://cfis.wi.gov/Public/Registration.aspx?page=ReceiptList#

Infuriatingly, the site only lets users export 65,000 records at a time. Kiernan 
wrote a scrip that will use Selenium to open a remote browser, submit a search
for all expenditures and download the row-limited files one by one. I adjusted the script
for expenditures.

```{r eval=FALSE}
source(
  file = here("state","wi", "expends", "docs", "scrape_wi_expends.R")
)
```

The files are downloaded to the `scrape/` directory.

```{r raw_dir}
raw_dir <- dir_create(here("state","wi", "expends", "data", "scrape"))
raw_info <- as_tibble(dir_info(raw_dir))
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
raw_csv <- raw_info$path
```

We should check the file names to ensure we were able to download every batch
of 65,000. If we count the distance between each of the sorted numbers in the
row ranges we should be left with only 1, 64999, and however many are in the 
last range (the only one below 65,000).

```{r}
row_range <- raw_csv %>% 
  str_extract(pattern = "(\\d+)-(\\d+)") %>% 
  str_split(pattern = "-") %>% 
  map(as.numeric) %>% 
  unlist() %>% 
  sort() 

sort(table(diff(row_range)))
```

## Read

The files can be read into a single data frame with `read_delim()`.

```{r raw_read}
wie <- read_delim( # 6,936,189
  file = raw_csv,
  delim = ",",
  escape_double = FALSE,
  escape_backslash = FALSE,
  col_types = cols(
    .default = col_character(),
    TransactionDate = col_date_mdy(),
    ContributionAmount = col_double(),
    ETHCFID = col_integer(),
    `72 Hr. Reports` = col_date_mdy(),
    SegregatedFundFlag = col_logical()
  )
)
```

We can check the number of rows against the total reported by our empty search.
We can also count the number of distinct values from a discrete column.

```{r raw_check}
nrow(wie) == 664843 # check col count
prop_na(wie[[length(wie)]]) # empty column
```

The file appears to have been read correctly. We just need to parse, rename, and
remove some of the columns.

```{r raw_names}
raw_names <- names(wie)[-length(wie)]
```

```{r name_clean}
wie <- wie %>% 
  clean_names("snake") %>% 
  select(-last_col()) # empty
```

## Explore

There are `r comma(nrow(wie))` rows of `r ncol(wie)` columns. Each record
represents a single contribution from an individual to a political committee.

```{r glimpse}
glimpse(wie)
tail(wie)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(wie, count_na)
```

We can flag any record missing a key variable needed to identify a transaction. These columns, "transaction_date", "registrant_name", "payee_name","amount", seem okay and not missing much info.

```{r na-flag}
key_vars <- c("transaction_date", "registrant_name", 
              "payee_name","amount")
wie <- flag_na(wie, all_of(key_vars))
sum(wie$na_flag)
```

Very, very few records are missing the contributor name.

```{r na-view}
wie %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any entirely duplicate rows. To keep memory usage low with such
a large data frame, we will split our data into a list and check each element of
the list. For each chunk, we will write the duplicate `id` to a text file.

```{r dupe_write}
wie <- wie %>% 
  group_split(
    q = quarter(transaction_date),
    y = year(transaction_date),
    .keep = FALSE
  ) %>% 
  map_dfr(
    .f = function(x) {
      message(x$transaction_date[1])
      if (nrow(x) > 1) {
        x <- flag_dupes(x, everything(), .check = FALSE)
        if (runif(1) > 0.75) {
          flush_memory(1)
        }
      } else {
        x$dupe_flag <- FALSE
      }
      return(x)
    }
  )
```

`r percent(mean(wie$dupe_flag, na.rm = TRUE), 0.1)` of rows are duplicates.

```{r dupe_view}
wie %>% 
  filter(dupe_flag) %>% 
  select(key_vars) %>% 
  arrange(transaction_date)
```

```{r dupe_count}
wie %>% 
  filter(dupe_flag) %>% 
  count(registrant_name, payee_name, transaction_date,
        sort = TRUE)
```

### Categorical

```{r distinct-count}
col_stats(wie, n_distinct)
```

```{r distinct-plots, echo=FALSE}
explore_plot(wie, expense_category) + coord_flip()
```

### Amounts

```{r amount-round}
wie$amount <- round(as.numeric(wie$amount), digits = 2)
```

```{r amount-summary}
summary(wie$amount)
mean(wie$amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(wie[c(
  which.max(wie$amount), 
  which.min(wie$amount)
), ])
```

```{r hist-amount, echo=FALSE}
wie %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Wisconsin Expenditures Amount Distribution",
    caption = "Source: WI CIFS",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
wie <- mutate(wie, transaction_year = year(transaction_date))
```

```{r date-range}
min(wie$transaction_date)
sum(wie$transaction_year < 2008)
max(wie$transaction_date)
sum(wie$transaction_date > today())
```

```{r bar-year, echo=FALSE}
wie %>% 
  count(transaction_year) %>% 
  mutate(even = is_even(transaction_year)) %>% 
  ggplot(aes(x = transaction_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2023, by = 2)) +
  theme(legend.position = "bottom") +
  coord_cartesian(xlim = c(2008, 2021)) +
  labs(
    title = "Wisconsin Expenditures by Year",
    caption = "Source: WI CIFS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address-norm}
addr_norm <- wie %>% 
  distinct(payee_address_line1,payee_address_line2) %>% 
  mutate(
    across(
      contains("address_"),
      list(anorm = normal_address),
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  unite(
    col = address_norm,
    ends_with("_anorm"),
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  )
```

```{r address-view}
addr_norm
```

```{r address-join}
wie <- left_join(wie, addr_norm, by = c("payee_address_line1","payee_address_line2"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
wie <- wie %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  wie$zip,
  wie$zip_norm,
  compare = valid_zip
)
```

### State

```{r state-upper}
wie$payee_state <- str_to_upper(wie$payee_state)
prop_in(wie$payee_state, valid_state)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- wie %>% 
  distinct(payee_city, payee_state, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = payee_city, 
      abbs = usps_city,
      states = c("WI", "DC", "WISCONSIN"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = payee_city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "payee_state" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
wie <- left_join(
  x = wie,
  y = norm_city,
  by = c(
    "payee_city" = "city_raw", 
    "payee_state", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- wie %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "payee_state" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    payee_state, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
good_refine <- distinct(good_refine)

wie <- wie %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(wie$city),
  wie$city_norm,
  wie$city_swap,
  wie$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(wie$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Wisconsin City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Wisconsin City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
wie <- wie %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(wie, 50))
```

1. There are `r comma(nrow(wie))` records in the database.
1. There are `r comma(sum(wie$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wie$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean-dir}
clean_dir <- dir_create(here("state","wi", "expends", "data", "clean"))
clean_path <- path(clean_dir, "wi_expends_2008-20230527.csv")
write_csv(wie, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
