---
title: "Kansas Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ks", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is retrieved from the [KanView Download Center][dlc]:

> KanView is an online solution that brings better visibility, openness, and
accountability to Kansas State Government. KanView contains data related to
Government spending and income, allowing Kansas taxpayers an inside look at
Government financial activity.

> The State of Kansas government operates on a fiscal year basis defined as July
1 through June 30. KanView presents data beginning with FY 2006 and each
completed fiscal year therafter. For detailed transactions please visit the
[Data Download Center][dlc].

The Download Center provides three types of data: Agency expenditures and
revenues, vendor payments, and employee compensation. We are interested in the
payments made from the state to outside vendors.

>  Vendor Payments: “Checkbook” payments made to vendors are available through
KanView. Download results will show payments made to vendors from all state
agencies and include the following key data elements: fiscal year, agency
(Business Unit), account description, funding, vendor name, document number,
payment date and amount. Due to data volumes, downloads are available on a
monthly or quarterly basis per each download request. Select the Fiscal Year,
then select either the quarter or month. If month is selected KanView ignores
any selection of a specific quarter.

[dlc]: http://kanview.ks.gov/DataDownload.aspx

## Download

Vendor payment data can be downloaded in files by fiscal year (July 1 to June
30). We can download the files manually from the Download Center, or we can
make a series of `POST()` requests via curl. 

```{r raw_dir}
raw_dir <- dir_create(here("ks", "contracts", "data", "raw"))
```

```{r}
raw_curl <- read_file(here("ks", "contracts", "raw_curl.txt"))
if (length(dir_ls(raw_dir)) == 0) {
  for (year in 2011:2019) {
    message("------", year, "------")
    raw_path <- path(raw_dir, glue("VendorData_{year}.csv"))
    raw_curl %>% 
      str_replace("(?<=VendorYearList\\=)\\d{4}", as.character(year)) %>% 
      system(intern = TRUE) %>% 
      write_lines(raw_path)
  }
}
```

```{r raw_info, results='asis'}
raw_info <- dir_info(raw_dir)
raw_info$path %>% 
  path.abbrev() %>% 
  md_code() %>% 
  md_bullet()
```

## Read

```{r raw_read}
ksc <- map_df(
  .x = raw_info$path,
  .f = read_delim,
  delim = ",",
  na = "",
  trim_ws = TRUE,
  guess_max = 0,
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    Fiscal_Year = col_integer(),
    # Payment_Date = col_date("%Y%m%d"),
    Amount = col_double(),
    X9 = col_skip()
  )
)
```

```{r raw_rename, echo=FALSE}
old_names <- names(ksc)
ksc <- ksc %>% 
  clean_names("snake") %>% 
  rename(
    fiscal = fiscal_year,
    agency = business_unit_descr,
    account = account_descr,
    fund = fund_code_descr,
    vendor = vendor_name,
    document = document_number,
    date = payment_date
  )
```

```{r raw_date}
x <- ksc$date
mdy_dates <- str_detect(x, "/")
x[mdy_dates] <- as.character(mdy(x[mdy_dates]))
x[!mdy_dates] <- as.character(ymd(x[!mdy_dates]))
ksc$date <- parse_date(x)
count_na(ksc$date)
rm(x)
```

## Explore

```{r glimpse}
glimpse(ksc)
tail(ksc)
```

### Missing

The only variable missing any values is the `fund` code.

```{r na_count}
col_stats(ksc, count_na)
```

This variable was not recorded from 2011 to 2013.

```{r na_year}
ksc %>% 
  group_by(fiscal) %>% 
  summarise(prop_na = percent(prop_na(fund)))
```

### Duplicates

```{r dupe_write}
tmp_files <- dir_ls(tempdir())
dupe_file <- if (any(str_detect(tmp_files, "dupes"))) {
  str_subset(tmp_files, "dupes")
} else {
  file_temp(pattern = "dupes")
}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  n <- 1e5 # rows per chunk
  nr <- nrow(ksc)
  # split file into chunks
  kss <- split(ksc, rep(1:ceiling(nr/n), each = n, length.out = nr))
  pb <- txtProgressBar(max = length(kss), style = 3)
  for (i in seq_along(kss)) {
    # check dupes from both ends
    d1 <- duplicated(kss[[i]], fromLast = FALSE)
    d2 <- duplicated(kss[[i]], fromLast = TRUE)
    # append to disk
    write_lines(d1 | d2, dupe_file, append = TRUE)
    rm(d1, d2)
    kss[[i]] <- NA
    Sys.sleep(10)
    flush_memory()
    setTxtProgressBar(pb, i)
  }
  rm(kss)
}
```

```{r dupe_read}
ksc <- mutate(ksc, dupe_flag = as.logical(read_lines(dupe_file)))
sum(ksc$dupe_flag)
```

```{r dupe_view}
ksc %>% 
  filter(dupe_flag) %>% 
  select(date, vendor, amount, agency)
```

### Categorical

```{r distinct_count}
col_stats(ksc, n_distinct)
```

```{r distinct_plots}
explore_plot(ksc, agency) + scale_x_truncate()
```

### Amounts

```{r ammount_summary}
summary(ksc$amount)
mean(ksc$amount <= 0)
```

```{r hist_amount, echo=FALSE}
ksc %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Kansas Vendor Payment Amount Distribution",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
ksc <- mutate(ksc, year = year(date))
```

```{r date_range}
min(ksc$date)
sum(ksc$year < 2000)
max(ksc$date)
sum(ksc$date > today())
```

```{r bar_year, echo=FALSE}
ksc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2011:2019) +
  theme(legend.position = "bottom") +
  labs(
    title = "Kansas Vendor Payments by Year",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

```{r state_add}
ksc <- mutate(ksc, agency_state = "KS", .after = agency)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(ksc, 20))
```

1. There are `r comma(nrow(ksc))` records in the database.
1. There are `r comma(sum(ksc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ksc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ks", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "ks_contracts_clean.csv")
write_csv(ksc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] comand line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ksc)),
  old = c(md_code(old_names), rep("", ncol(ksc) - length(old_names))),
  type = md_code(map_chr(ksc, typeof)),
  def = c(
    "Fiscal year paid",
    "Spending agency",
    "Agency state (KS)",
    "Funds source account",
    "Fund name",
    "Vendor name",
    "Document number",
    "Daid paid",
    "Payment amount",
    "Duplicate record",
    "Calendar year paid"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original Name", "Type", "Definition")
))
```
