---
title: "Texas Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
is_inter <- FALSE
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("tx", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  jsonlite, # convert json table
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Texas Ethics Commission (TEC)][tec]. According to 
[a TEC brochure][bro]:

> Statutory duties of the Ethics Commission are in Chapter 571 of the Government
Code. The agency is responsible for administering these laws: (1) Title 15,
Election Code, concerning political contributions and expenditures, and
political advertising...

> The Ethics Commission serves as a repository of required disclosure statements
for state officials, candidates, political committees, lobbyists, and certain
district and county judicial officers.

Data is ontained from the [Campaign Finance section of the TEC website][cf]. The
entire database can be downloaded as [a ZIP archive][zip]. The contents of that
ZIP and the layout of the files within are outlined in the
[`CFS-ReadMe.txt`][rm] file.

> This zip package contains detailed information from campaign finance reports
filed electronically with the Texas Ethics Commission beginning July 1, 2000.
Flat File Architecture Record Listing -- Generated 06/11/2016 12:38:08 PM

[tec]: https://www.ethics.state.tx.us/search/cf/
[bro]: https://www.ethics.state.tx.us/data/about/Bethic.pdf
[cf]: https://www.ethics.state.tx.us/search/cf/
[zip]: https://www.ethics.state.tx.us/data/search/cf/TEC_CF_CSV.zip
[rm]: https://www.ethics.state.tx.us/data/search/cf/CFS-ReadMe.txt

```{r read_key}
readme <- read_lines("https://www.ethics.state.tx.us/data/search/cf/CFS-ReadMe.txt")
```

At the top of this file is a table of contents.

```{r print_key, results='asis', echo=FALSE}
toc <- str_trim(readme[seq(13, 47, 2)][-2])
toc <- str_replace_all(toc, "\\s{2,}", "\t")
read_tsv(file = I(toc)) %>% 
  clean_names() %>% 
  mutate(
    file_name_s = file_name_s %>% 
      str_trunc(30) %>% 
      str_split(",\\s") %>% 
      map(md_code) %>% 
      map_chr(str_c, collapse = ", "),
    file_contents = str_trunc(file_contents, width = 50)
  ) %>% 
  md_table()
```

From this table, we know the "ContributionData" record (`contribs_##.csv`)
contains the data we want.

> Contributions - Schedules A/C - Contributions from special session and special
pre-election (formerly Telegram) reports are stored in the file cont_ss and
cont_t. These records are kept separate from the contribs files to avoid
creating duplicates, because they are supposed to be re-reported on the next
regular campaign finance report. Files: `contribs_##.csv`, `cont_ss.csv`, 
`cont_t.csv`

```{r print_contrib_data, results='asis', echo=FALSE}
x <- str_subset(readme[130:168], "^(\\s|-){10,}", negate = TRUE)
x <- str_trim(str_replace(x, "\\s(\\d)\\s", "0\\1 "))
con_abt <- read_fwf(
  file = I(x),
  skip = 1,
  col_positions = fwf_empty(
    file = I(x),
    col_names = str_split(
      string = x[1],
      pattern = "\\s{2,}|\\s(?=D)",
      simplify = TRUE
    )
  )
)

clean_names(con_abt) %>% 
  mutate_at(vars(field_name), ~md_code(make_clean_names(.))) %>% 
  md_table()
```

### Download

```{r raw_paths}
raw_dir <- dir_create(here("tx", "contribs", "data", "raw"))
zip_url <- "https://www.ethics.state.tx.us/data/search/cf/TEC_CF_CSV.zip"
zip_path <- path(raw_dir, basename(zip_url))
```

The ZIP file is fairly large, so check the file size before downloading.

```{r raw_size, collapse=TRUE}
# size of file
(zip_size <- url_file_size(zip_url))
```

If the file hasn't been downloaded yet, do so now.

```{r raw_download}
if (!file_exists(zip_path)) {
  download.file(zip_url, zip_path)
}
```

### Unzip

There are `r nrow(unzip(zip_path, list = T))` CSV files inside the ZIP archive.
We can list the content and extract only those pertaining to contributions.

```{r zip_list}
(zip_contents <- 
  unzip(zip_path, list = TRUE) %>% 
  as_tibble() %>% 
  clean_names() %>% 
  mutate(
    length = as_fs_bytes(length),
    date = as_date(date)
  ))

zip_contribs <- str_subset(zip_contents$name, "contribs_\\d{2}")
length(zip_contribs)
```

If the files haven't been extracted, we can do so now. There are 
`r length(zip_contribs)` contribution files to extract.

```{r zip_extract}
if (!all(file_exists(path(raw_dir, zip_contribs)))) {
  unzip(
    zipfile = zip_path,
    files = zip_contribs,
    exdir = raw_dir
  )
}

raw_paths <- path(raw_dir, zip_contribs)
```

### Read 

The `r length(raw_paths)` files can be read into a single data frame. We will
consult the `CFS-ReadMe.txt` file for the column types.

```{r raw_read}
txc <- read_delim(
  file = raw_paths,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    reportInfoIdent = col_integer(),
    receivedDt = col_date("%Y%m%d"),
    contributionInfoId = col_integer(),
    contributionDt = col_date("%Y%m%d"),
    contributionAmount = col_double()
  )
)
```

```{r clean_names}
txc <- clean_names(txc, case = "snake")
```

To ensure the file has been read correctly, we can check that a categorical
variable has very few distinct values.

```{r raw_check}
n_distinct(txc$record_type)
```

```{r flag_lgl}
yes_no <- function(x) x == "Y"
txc <- txc %>% 
  mutate(across(ends_with("_flag"), yes_no))
```

```{r include=FALSE}
flush_memory(2)
```

## Trim

Trim unused columns for memory space. Rejoin after the clean file is saved.

```{r}
txc <- txc %>% 
  select(
    filer_ident,
    filer_name,
    contribution_info_id,
    contribution_dt,
    contribution_amount,
    contributor_name_organization,
    contributor_name_last,
    contributor_name_first,
    contributor_street_city,
    contributor_street_state_cd,
    contributor_street_postal_code
  )
```

```{r include=FALSE}
Sys.sleep(10)
flush_memory(5)
Sys.sleep(10)
flush_memory(5)
```

## Explore

```{r nrow}
comma(nrow(txc))
```

```{r glimpse}
glimpse(txc[1:20, ])
tail(txc)
```

### Missing

```{r na_count, eval=FALSE}
col_stats(txc, count_na)
```

```{r key_vars}
key_vars <- c(
  "contribution_dt",
  "contributor_name_first",
  "contributor_name_last",
  "contribution_amount",
  "filer_name"
)
```

```{r na_flag}
txc <- txc %>% 
  mutate(
    contributor_name_any = coalesce(
      contributor_name_organization,
      contributor_name_last,
      contributor_name_first
    )
  ) %>% 
  flag_na(
    contribution_dt,
    contributor_name_any,
    contribution_amount,
    filer_name
  ) %>% 
  select(-contributor_name_any)

sum(txc$na_flag)
```

```{r na_view}
txc %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

```{r include=FALSE}
flush_memory(2)
```

### Duplicate

```{r dupe_write}
dupe_file <- here("tx", "contribs", "data", "dupes.txt")
if (!file_exists(dupe_file)) {
  # save copy to disc
  tmp <- file_temp(ext = "rds")
  write_rds(txc, file = tmp)
  file_size(tmp)
  # split file into chunks
  tx_id <- split(txc$contribution_info_id, txc$received_dt)
  txs <- txc %>%
    select(-contribution_info_id) %>% 
    group_split(received_dt)
  # remove from memory
  if (file_exists(tmp)) {
    rm(txc)
    Sys.sleep(5)
    flush_memory(2)
  }
  pb <- txtProgressBar(max = length(txs), style = 3)
  for (i in seq_along(txs)) {
    # check dupes from both ends
    if (nrow(txs[[i]]) > 1) {
      d1 <- duplicated(txs[[i]], fromLast = FALSE)
      d2 <- duplicated(txs[[i]], fromLast = TRUE)
      dupe_vec <- d1 | d2
      rm(d1, d2)
      # append dupe id to file
      if (any(dupe_vec)) {
        write_lines(
          x = tx_id[[i]][dupe_vec], 
          file = dupe_file, 
          append = file_exists(dupe_file)
        )
      }
      rm(dupe_vec)
    }
    txs[[i]] <- NA
    tx_id[[i]] <- NA
    if (i %% 100 == 0) {
      Sys.sleep(2)
      flush_memory(2)
    }
    setTxtProgressBar(pb, i)
  }
  rm(txs, tx_id)
  Sys.sleep(5)
  flush_memory(2)
  txc <- read_rds(tmp)
}
```

```{r dupes_read}
tx_dupes <- tibble(
  contribution_info_id = as.integer(read_lines(dupe_file)), 
  dupe_flag = TRUE
)
```

```{r dupes_join}
txc <- left_join(txc, tx_dupes, by = "contribution_info_id")
txc <- mutate(txc, dupe_flag = !is.na(dupe_flag))
```

```{r dupe_count}
mean(txc$dupe_flag)
```

```{r dupe_view, cache=TRUE}
txc %>% 
  filter(dupe_flag) %>% 
  select(contribution_info_id, all_of(key_vars)) %>% 
  arrange(contribution_dt)
```

### Categorical

```{r n_distinct, eval=FALSE}
col_stats(txc, n_distinct)
```

```{r explore_plot, echo=FALSE, eval=FALSE}
explore_plot(txc, form_type_cd)
explore_plot(txc, sched_form_type_cd)
explore_plot(txc, filer_type_cd)
explore_plot(txc, contributor_persent_type_cd)
```

```{r include=FALSE}
flush_memory(2)
```

### Amounts

`r percent(mean(txc$contribution_amount <= 0), 0.01)`  of contributions have a
`contribution_amount` less than or equal to zero.

```{r ammount_summary}
summary(txc$contribution_amount)
percent(mean(txc$contribution_amount <= 0), 0.01)
```

```{r hist_amount, echo=FALSE}
txc %>%
  filter(contribution_amount > 1, contribution_amount < 1e6) %>% 
  ggplot(aes(contribution_amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Texas Contributions Amount Distribution",
    subtitle = "from 2000 to 2021",
    caption = "Source: TEC",
    x = "Amount",
    y = "Count"
  )
```

```{r include=FALSE}
flush_memory(2)
```

### Dates

We can create a new `contribution_yr` variable from `contribution_dt`.

```{r date_year}
txc <- mutate(txc, contribution_yr = year(contribution_dt))
```

The `contribution_dt` column is very clean, with almost no dates out of the
expected range.

```{r date_range}
count_na(txc$contribution_dt)
min(txc$contribution_dt, na.rm = TRUE)
sum(txc$contribution_yr < 2000, na.rm = TRUE)
max(txc$contribution_dt, na.rm = TRUE)
sum(txc$contribution_dt > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
txc %>% 
  count(contribution_yr) %>% 
  mutate(even = is_even(contribution_yr)) %>% 
  ggplot(aes(x = contribution_yr, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  coord_cartesian(xlim = c(2000, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Texas Contributions by Year",
    caption = "Source: TEC",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

```{r include=FALSE}
flush_memory(2)
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

For this database, there are no street addresses.

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
txc <- txc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = contributor_street_postal_code,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  txc$contributor_street_postal_code,
  txc$zip_norm,
  compare = valid_zip
)
```

```{r include=FALSE}
flush_memory(2)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
st_norm <- txc %>% 
  distinct(contributor_street_state_cd) %>% 
  mutate(
    state_norm = normal_state(
      state = contributor_street_state_cd,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r state_join}
txc <- left_join(txc, st_norm, by = "contributor_street_state_cd")
rm(st_norm)
```

```{r state_view}
txc %>% 
  filter(contributor_street_state_cd != state_norm) %>% 
  count(contributor_street_state_cd, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  txc$contributor_street_state_cd,
  txc$state_norm,
  compare = valid_state
)
```

```{r include=FALSE}
flush_memory(2)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- txc %>% 
  distinct(contributor_street_city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = contributor_street_city, 
      abbs = usps_city,
      states = c("TX", "DC", "TEXAS"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

```{r include=FALSE}
flush_memory(2)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
norm_city <- norm_city %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  distinct()
```

```{r include=FALSE}
Sys.sleep(10)
flush_memory(2)
```

```{r city_rejoin}
txc <- left_join(
  x = txc,
  y = norm_city,
  by = c(
    "contributor_street_city", 
    "state_norm", 
    "zip_norm"
  )
)
rm(norm_city)
```

```{r include=FALSE}
Sys.sleep(10)
flush_memory(2)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/ZIP combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- txc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
txc <- txc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

```{r include=FALSE}
rm(good_refine)
Sys.sleep(10)
flush_memory(2)
```

#### Check

We can use the `campfin::check_city()` function to pass the remaining unknown
`city_refine` values (and their `state_norm`) to the Google Geocode API. The
function returns the name of the city or locality which most associated with
those values.

This is an easy way to both check for typos and check whether an unknown
`city_refine` value is actually a completely acceptable neighborhood, census
designated place, or some other locality not found in our `valid_city` vector
from our `zipcodes` database.

First, we'll filter out any known valid city and aggregate the remaining records
by their city and state. Then, we will only query those unknown cities which
appear at least ten times.

```{r check_filter}
many_city <- c(valid_city, extra_city)
tac_out <- txc %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, state_norm, sort = TRUE) %>% 
  drop_na() %>% 
  filter(n > 1) %>% 
  head(200)
```

Passing these values to `campfin::check_city()` with `purrr::pmap_dfr()` will
return a single tibble of the rows returned by each city/state combination.

First, we'll check to see if the API query has already been done and a file
exist on disk. If such a file exists, we can read it using `readr::read_csv()`.
If not, the query will be sent and the file will be written using
`readr::write_csv()`.

```{r check_send}
check_file <- here("tx", "contribs", "data", "api_check.csv")
if (file_exists(check_file)) {
  check <- read_csv(
    file = check_file
  )
} else {
  check <- pmap_dfr(
    .l = list(
      tac_out$city_refine, 
      tac_out$state_norm
    ), 
    .f = check_city, 
    key = Sys.getenv("GEOCODE_KEY"), 
    guess = TRUE
  ) %>% 
    mutate(guess = coalesce(guess_city, guess_place)) %>% 
    select(-guess_city, -guess_place)
  write_csv(
    x = check,
    path = check_file
  )
}
```

Any city/state combination with a `check_city_flag` equal to `TRUE` returned a
matching city string from the API, indicating this combination is valid enough
to be ignored.

```{r check_accept}
valid_locality <- check$guess[check$check_city_flag]
```

Then we can perform some simple comparisons between the queried city and the
returned city. If they are extremely similar, we can accept those returned
locality strings and add them to our list of accepted additional localities.

```{r check_compare}
valid_locality <- check %>% 
  filter(!check_city_flag) %>% 
  mutate(
    abb = is_abbrev(original_city, guess),
    dist = str_dist(original_city, guess)
  ) %>%
  filter(abb | dist <= 3) %>% 
  pull(guess) %>% 
  append(valid_locality)
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city, valid_locality)
progress <- progress_table(
  str_to_upper(txc$contributor_street_city),
  txc$city_norm,
  txc$city_swap,
  txc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
  mutate(across(1, md_code)) %>% 
  kable(digits = 4)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(txc$contributor_street_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Texas City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Texas City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
txc <- txc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(city_clean, state_clean, zip_clean, .before = last_col())
```

```{r clean_glimpse}
glimpse(sample_n(txc, 20))
```

1. There are `r comma(nrow(txc))` records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(txc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit year variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("tx", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "tx_contribs_2000-20210902.csv")
write_csv(txc, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
