---
title: "West Virginia Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("wv", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  batman, # parse logicals
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr2, # http requests
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contributions data can be obtained in bulk from the West Virginia Secretary of
State [Campaign Finance Reporting System (CFRS)][cfrs]. Each record contains
information on a contribution made to a candidate or committee. The CFRS also
provides a [record layout PDF][layout], which we have converted to a text file.

[cfrs]: https://cfrs.wvsos.gov/index.html#/index
[layout]: https://cfrs.wvsos.gov/CFIS_APIService/Template/KeyDownloads/Contributions%20and%20Loans%20File%20Layout%20Key.pdf

## Dictionary

```{r raw_layout, results='asis'}
key_path <- here("wv", "contribs", "record_layout.csv")
(dict_md <- kable(read_csv(key_path)))
```

```{r dict_write}
write_lines(
  x = c("# West Virginia Contracts Data Dictionary\n", dict_md),
  file = here("wv", "contribs", "wv_contribs_dict.md"),
)
```

## Download

The files can be downloaded with an `httr::GET()` request to the CFRS server.

```{r raw_get}
raw_dir <- dir_create(here("wv", "contribs", "data", "raw"))
wv_api <- "https://cfrs.wvsos.gov/CFIS_APIService/api"
```

```{r raw_list}
wv_ls <- request("https://cfrs.wvsos.gov/CFIS_APIService/api/") %>% 
  req_url_path_append("DataDownload", "GetCheckDatadownload") %>% 
  req_url_query(pageNumber = 1, pageSize = 50) %>% 
  req_perform() %>% 
  resp_body_json(
    check_type = FALSE,
    simplifyDataFrame = TRUE
  )
```

```{r echo=FALSE}
wv_ls %>% 
  mutate(across(NameOfFile, md_code)) %>% 
  kable()
```

```{r}
wv_ls <- wv_ls %>% 
  filter(TransactionType == "CON") %>% 
  mutate(FilePath = path(raw_dir, NameOfFile))
```

```{r}
for (i in seq(nrow(wv_ls))) {
  message(wv_ls$NameOfFile[i])
  if (!file_exists(wv_ls$FilePath[i])) {
    request("https://cfrs.wvsos.gov/CFIS_APIService/api/") %>% 
      req_url_path_append("DataDownload", "GetCSVDownloadReport") %>% 
      req_url_query(
        year = wv_ls$ElectionYear[i],
        transactionType = wv_ls$TransactionType[i],
        reportFormat = "csv",
        fileName = wv_ls$NameOfFile[i]
      ) %>% 
      req_perform(path = wv_ls$FilePath[i])
  }
}
```

```{r echo=FALSE}
raw_csv <- wv_ls$FilePath
```

```{r echo=FALSE}
file_info(raw_csv) %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```

## Read

While character columns are wrapped in double-quotes (`"`), any double-quotes
_within_ those columns are not escaped in any way. We will have to use regular
expressions to replace them with single-quotes (`'`).

```{r raw_fix}
fix_csv <- path_temp(basename(raw_csv))
for (i in seq_along(raw_csv)) {
  read_lines(raw_csv[i]) %>% 
    str_replace("Report$", "Report,") %>% 
    str_replace_all(",\"([A-z\\. ]+)\"[^,|\"]*", ",'\\1'") %>% 
    str_replace_all("\\s\"([A-z\\. ]+)\"[^,|\"]*", " '\\1'") %>% 
    str_replace_all("(?<!^|,)\"(?!,|$)", r"("""")") %>% 
    write_lines(fix_csv[i])
}
```

The fixed text files can be read into a single data frame.

```{r raw_read}
wvc <- read_delim(
  file = fix_csv,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = TRUE,
  na = c("", " "),
  col_types = cols(
    .default = col_character(),
    `Receipt Amount` = col_double(),
    `Receipt Date` = col_date("%m/%d/%Y %H:%M:%S %p"),
    `Filed Date` = col_datetime("%m/%d/%Y %H:%M:%S %p"),
    `Fundraiser Event Date` = col_datetime("%m/%d/%Y %H:%M:%S %p")
  )
)
```

```{r}
problems(wvc)
```

Column names can be converted to snake case and simplified.

```{r raw_rename}
wvc <- wvc %>% 
  clean_names("snake") %>% 
  mutate(across(c(amended, occupation_comment), to_logical))
```

We can check whether the files were properly read by counting the number of
distinct values in a discrete variable like the `contribution_type`.

```{r raw_check}
count(wvc, contribution_type)
```

## Explore

```{r glimpse}
glimpse(wvc)
tail(wvc)
```

### Missing

Columns range in their degree of missing values.

```{r na_count}
col_stats(wvc, count_na)
```

We should flag any record missing a key variable, those needed to identify a transaction and all parties, with `campfin::flag_na()`.

After combining these rows, we have no records missing key variables.

```{r key_vars}
key_vars <- c("receipt_date", "last_name", "receipt_amount", "candidate_name")
```

```{r na_flag}
wvc <- flag_na(wvc, all_of(key_vars))
mean(wvc$na_flag)
```

### Duplicates

We can also flag records that are entirely duplicated across every row, save for
the supposedly unique `id`.

```{r dupe_flag}
wvc <- flag_dupes(wvc, -receipt_id)
percent(mean(wvc$dupe_flag), 0.01)
```

```{r dupe_view}
wvc %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(receipt_date)
```

### Categorical

```{r distinct_cound}
col_stats(wvc, n_distinct)
```

```{r distinct_plot, echo=FALSE}
explore_plot(wvc, receipt_source_type) + scale_x_wrap()
explore_plot(wvc, receipt_type) + scale_x_wrap()
explore_plot(wvc, committee_type) + scale_x_wrap()
explore_plot(wvc, occupation) + scale_x_wrap()
explore_plot(wvc, report_name) + scale_x_wrap()
```

### Amounts

```{r}
wvc$receipt_amount <- round(wvc$receipt_amount, digits = 2)
```

The range of contribution amounts seems reasonable.

```{r amount_summary}
noquote(map_chr(summary(wvc$receipt_amount), dollar))
sum(wvc$receipt_amount <= 0.01)
```

Here is the largest contribution of `r dollar(max(wvc$receipt_amount))`:

```{r amount_max}
glimpse(wvc[which.max(wvc$receipt_amount), ])
```

```{r hist_amount, echo=FALSE}
wvc %>%
  ggplot(aes(receipt_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "West Virginia Contracts Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: WV CFRS",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `receipt_date` with `lubridate::year()`

```{r date_year}
wvc <- mutate(wvc, receipt_year = year(receipt_date))
```

The range of dates is completely normal.

```{r date_range}
min(wvc$receipt_date)
sum(wvc$receipt_year < 2016)
max(wvc$receipt_date)
sum(wvc$receipt_date > today())
```

```{r bar_year, echo=FALSE}
wvc %>% 
  count(receipt_year) %>% 
  mutate(even = is_even(receipt_year)) %>% 
  ggplot(aes(x = receipt_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2016, 2020, by = 1)) +
  theme(legend.position = "bottom") +
  labs(
    title = "West Virginia Contracts by Year",
    caption = "Source: WV CFRS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
addr_norm <- wvc %>%
  distinct(address1, address2) %>% 
  mutate(
    norm_address1 = normal_address(
      address = address1,
      abbs = usps_street,
      na_rep = TRUE
    ),
    norm_address2 = normal_address(
      address = address2,
      abbs = usps_street,
      na_rep = TRUE,
      abb_end = FALSE
    )
  ) %>% 
  unite(
    col = address_norm,
    starts_with("norm_address"),
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  ) %>% 
  mutate(across(address_norm, na_if, ""))
```

```{r address_view}
sample_n(addr_norm, 10)
```

```{r address_rejoin}
wvc <- left_join(wvc, addr_norm)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
wvc <- wvc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  wvc$zip,
  wvc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state-norm}
wvc <- wvc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state-view}
wvc %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state-progress}
progress_table(
  wvc$state,
  wvc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- wvc %>% 
  distinct(city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("WV", "DC", "WEST VIRGINIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
wvc <- left_join(
  x = wvc,
  y = norm_city,
  by = c(
    "city" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city-refine}
good_refine <- wvc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city-count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city-join}
wvc <- wvc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(wvc$city),
  wvc$city_norm,
  wvc$city_swap,
  wvc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>% 
    mutate(across(stage, md_code)) %>% 
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(wvc$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "West Virginia City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "West Virginia City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
wvc <- wvc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(wvc, 50))
```

1. There are `r comma(nrow(wvc))` records in the database.
1. There are `r comma(sum(wvc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wvc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean-dir}
clean_dir <- dir_create(here("wv", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "wv_contribs_20160301-20210630.csv")
write_csv(wvc, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict-make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wvc)),
  type = md_code(map_chr(wvc, typeof)),
  def = c(
    ""
  )
)
```

```{r dict-md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
