---
title: "Connecticut Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 95)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ct", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  textreadr, # read doc files
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::dr_here(show_reason = FALSE)
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The database of registered voters in Connecticut was obtained by Secretary of
State's office for a fee of $300 and delivered on compact disc.

> The Centralized Voter Registration System is utilized by all towns in
Connecticut. It is the exclusive means by which a town produces an official
voter registry list. The system includes information contained in voter
registration applications, indicates whether eligible voters participated in
past elections and primaries, and whether they voted in person or by absentee
ballot. This election history information is required to be updated by all towns
within 60 days after each election or primary.

> The cost of the file is $300 and can be paid by check or credit card; cash is
also accepted. We must receive payment before releasing the voter registry file.
If you have any questions you can contact us by phone at 860-509-6100 or by
email at LEAD@ct.gov.

## Download

The raw data from the disk can be downloaded from the Workshop's AWS server.

```{r aws_bucket}
aws_info <- get_bucket_df(
  bucket = "publicaccountability", 
  prefix = "FOR_REVIEW/ct_voters"
)
```

```{r aws_info, echo=FALSE}
aws_info <- aws_info %>% 
  as_tibble() %>% 
  transmute(
    path = as_fs_path(Key), 
    size = as_fs_bytes(Size), 
    modification_time = parse_datetime(LastModified)
  )
print(aws_info)
```

```{r raw_dir}
raw_dir <- dir_create(here("ct", "voters", "data", "raw"))
```

We will save each object to a local directory.

```{r aws_get}
for (key in aws_info$path[-1]) {
  p <- path(raw_dir, basename(key))
  if (!file_exists(p)) {
    save_object(
      object = key,
      bucket = "publicaccountability",
      file = p
    )
  }
}
```

```{r raw_info}
raw_info <- dir_info(raw_dir)
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## About

### Contents

The extract includes a Microsoft Word file describing the contents. 

> The extract is broken into 4 zipped files.  The towns are sequenced
alphabetically with a corresponding id tax town code(001-169).  The extract
files both fixed length and comma delimited and are assembled as follows:

| File    | Towns
|:--------|------------------------------------:|
| `FILE1` | Andover 001 thru East Hampton 042   |
| `FILE2` | East Hartford 043 thru 	Monroe 085  |
| `FILE3` | Montville 086 thru Sherman 127      |
| `FILE4` | Simsbury 128 thru Woodstock 169     |

### Codes

There's also a file to provide code descriptions.

```{r status_code, echo=FALSE}
(status_code <- tribble(
  ~CD_STATUS_CODE, ~CD_STATUS,
  "A", "Active",
  "I", "Inactive",
  "O", "Off"
))
```

```{r reason_code, echo=FALSE}
(reason_code <- tribble(
  ~CD_OFF_REASON, ~CD_OFF_DESC,
  "F", "Felon",
  "D", "Death",
  "C", "Move Out Of State",
  "V", "Voter Cancel",
  "M", "DMV",
  "R", "Notice-No Reply",
  "U", "Duplicate",
  "O", "Canvass - Moved Out",
  "T", "CVR Returned by Voter",
  "E", "ED-683 Return by Voter",
  "B", "Inactive 4 Yrs",
  "I", "Canvass - Inactive to Off"
))
```

```{r spec_code, echo=FALSE}
(spec_code <- tribble(
  ~CD_SPEC_CODE, ~CD_SPEC_STATUS,
  "A", "None –may be used for non displayable addresses",
  "M", "MILITARY",
  "I", "INSTITUTION",
  "O", "OVERSEAS",
  "X", "OTHER"
))
```

```{r elect_code, echo=FALSE}
(elect_code <- tribble(
  ~ELECT_TYPE_CODe, ~ELECT_TYPE_DESC,
  "S", "SPECIAL",
  "P", "PRIMARY",
  "R", "REFERENDUM",
  "E", "GENERAL ELECTION",
  "D", "DELEGATE CAUCAS",
  "T", "TOWN COMMITTEE PRIMARY"
))
```

```{r absent_flag, echo=FALSE}
(absent_flag <- tribble(
  ~ABSENTEE_FLAG, ~ABSENTEE_DESC,
  "Y", "VOTING USING ABSENTEE BALLOT",
  "N", "VOTING IN PERSON"
))
```

### Towns

The `town.txt` file contains the town names and corresponding numeric codes.

```{r town_codes, echo=FALSE}
town_codes <- read_lines(str_subset(raw_info$path, "town"))
(town_codes <- town_codes %>% 
  str_subset(",") %>% 
  str_replace("\\s+,\\s+", "\t") %>% 
  str_remove("\\s+,$") %>% 
  read_tsv(
    col_names = c("ID_TOWN", "NM_NAME"),
    col_types = cols(
      .default = col_character()
    )
  ))
```

### Columns

The `elctext.txt` file contains a layout key for the voter files.

> THIS FILE REPRESENTS THE RECORD LAYOUT FOR THE FOUR 4 FILE.  
> "PIC" IS THE LENGTH OF EACH FIELD.  
> "COMM#" IS A COMMA USED TO DELIMIT EACH FIELD. THIS WILL HELP IN CREATE AN
EXCEL SPREADSHEET.

```{r raw_cols, echo=FALSE}
raw_cols <- read_lines(path(raw_dir, "elctext.txt"))[9:94]
raw_cols <- raw_cols %>% 
  str_subset("COMM\\d+", negate = TRUE) %>% 
  read_fwf(
    col_positions = fwf_cols(
      column = c(13, 33),
      length = c(34, 46),
      desc = c(47, NA)
    )
  ) %>% 
  extract(
    col = length,
    into = "length",
    regex = "PIC X\\((\\d+)\\)",
    convert = TRUE
  )
```

The first 43 columns have unique descriptions.

```{r reg_cols}
raw_cols$column <- raw_cols$column %>% 
  make_clean_names(case = "snake") %>% 
  str_remove("ws_") %>% 
  str_remove("vtr_")

raw_cols %>% 
  mutate(across(column, md_code)) %>% 
  mutate(across(desc, str_to_sentence)) %>% 
  kable()
```

The last 60 columns are the same 3 columns repeated _up to_ 20 times. Each
group of 3 is the date of an election, the type of election, and whether or not
the voter voted absentee. The columns groups are ordered with the most recent
election to the first election in which that voter participated.

We need to create column names and lengths for these 60 new columns. We can add
these to the descriptions of the other columns.

```{r all_cols}
raw_cols <- bind_rows(
  raw_cols,
  tibble(
    column = paste(
      rep(c("date", "type", "abstee"), 20),
      rep(x = 1:20, each = 3), sep = "_elect"
    ),
    length = rep(c(10, 1, 1), length.out = length(column)),
    desc = ""
  )
)
```

## Read

Each file is both comma separated _and_ fixed width. Fixed width files are more
complicated to read but can be read faster and more safely as each column is at
the same location for every row. This makes it worth going through the effort to
read the file as fixed width and _ignore_ the commas.

To build the column position specification needed by `readr::read_fwf()`, we
can use the column lengths provided in the `elctext.txt` file. We have to 
shift the lengths of each column to account for the commas between them.

```{r col_positions}
col_length <- rep(list(c(NA, NA)), nrow(raw_cols))
for (i in seq_along(raw_cols$length)) {
  if (i == 1) {
    # col starting position
    col_length[[i]][1] <- 0
  } else {
    col_length[[i]][1] <- col_length[[i - 1]][2] + 1
  }
  # col ending position
  col_length[[i]][2] <- col_length[[i]][1] + raw_cols$length[i]
}
```

We can then separate the start and end positions of each column.

```{r col_vector}
col_start <- vapply(col_length, `[[`, double(1), 1) + 1
col_end   <- vapply(col_length, `[[`, double(1), 2)
```

These numbers can be converted to the data frame expected by a fixed width file.

```{r col_fwf}
raw_fwf <- fwf_positions(
  start = col_start,
  end = col_end,
  col_names = raw_cols$column
)
```

```{r echo=FALSE}
print(raw_fwf)
```

Each voter file is a compressed archive. We can extract them to a temporary
directory to facilitate the file reading.

```{r raw_unzip}
raw_paths <- str_subset(raw_info$path, "ZIP$")
tmp_paths <- character()
for (i in seq_along(raw_paths)) {
  tmp <- unzip(raw_paths[i], exdir = tempdir(), junkpaths = TRUE)
  tmp_paths <- append(tmp_paths, tmp)
}
```

Then, all the raw temporary files can be read at once into a single data frame.

```{r raw_read}
ctv <- vroom_fwf(
  file = tmp_paths,
  col_positions = raw_fwf,
  col_types = cols(
    .default = col_character(),
    dt_birth = col_date_usa(),
    dt_accept = col_date_usa()
  )
)
```

```{r raw_count}
comma(nrow(ctv))
```

### Old

The same Connecticut voter registration data was previously requested in the
summer of 2019, similarly processed, and uploaded to the IRW server.

```{r old_date}
old_head <- head_object("csv/ct_voters.csv", "publicaccountability")
attr(old_head, "last-modified")
```

We can downloaded that old file locally.

```{r old_path}
old_file <- path(raw_dir, "ct_voters_old.csv")
```

```{r old_save}
if (!file_exists(old_file)) {
  save_object(
    object = "csv/ct_voters.csv",
    bucket = "publicaccountability",
    file = old_file,
    show_progress = TRUE
  )
}
```

It can then be read into a data frame like the newer data.

```{r old_read}
cto <- vroom(
  file = old_file,
  delim = ",",
  col_select = 1:103,
  skip = 1,
  col_names = raw_cols$column,
  col_types = cols(
    .default = col_character(),
    dt_birth = col_date_usa(),
    dt_accept = col_date_usa()
  )
)
```

```{r old_count}
comma(nrow(cto))
max(cto$dt_accept, na.rm = TRUE)
```

### Join

We want the most comprehensive voter roll possible, so we will keep any voters
found in the older data but _not_ the newer data. We can use the unique
`id_voter` column to filter duplicates out.

```{r old_bind}
cto <- filter(cto, id_voter %out% ctv$id_voter)
cto$cd_status <- "I"
ctv <- ctv %>% 
  bind_rows(cto, .id = "source_file") %>% 
  arrange(id_voter) %>% 
  relocate(source_file, .after = dt_accept)
```

```{r bind_count}
comma(nrow(ctv))
```

### Trim

I am going to remove all but the most recent election and convert the voter
history in a more readable format.

```{r raw_trim}
vote_hist <- select(ctv, id_voter, 45:103)
ctv <- select(
  .data = ctv, 1:44,
  dt_last = date_elect1,
  cd_last = type_elect1,
  abstee_last = abstee_elect1
)
```

```{r hist_file}
hist_file <- path(dirname(raw_dir), "voter_history.tsv.xz")
```

Instead of 60 columns in groups of 3 per voter, we can create 20 rows of 3
columns, removing any empty rows in which they did not vote.

```{r hist_pivot}
if (!file_exists(hist_file)) {
  vote_hist <- pivot_longer(
    data = vote_hist,
    cols = !id_voter,
    names_to = c(".value", "elect"),
    names_sep = "_",
    values_drop_na = TRUE
  )
}
```

```{r hist_read}
if (!file_exists(hist_file)) {
  vote_hist <- mutate(
    .data = vote_hist,
    elect = as.integer(str_remove_all(elect, "\\D")),
    date = mdy(date),
    abstee = (abstee == "Y")
  )
}
```

This separate voter file can be written and compressed.

```{r hist_write}
if (!file_exists(hist_file)) {
  vroom_write(vote_hist, xzfile(hist_file))
}
```

```{r hist_size}
file_size(hist_file)
```

```{r hist_rm, echo=FALSE}
rm(vote_hist); flush_memory()
```

## Explore

There are `r scales::comma(nrow(ctv))` rows of `r ncol(ctv)` columns. Each row
is a single voter registered in Connecticut.

```{r glimpse}
glimpse(ctv)
tail(ctv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(ctv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("nm_last", "dt_birth", "dt_accept")
ctv <- flag_na(ctv, all_of(key_vars))
sum(ctv$na_flag)
```

```{r na_view}
ctv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

All these records are missing a voter birth date.

```{r na_recount}
ctv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("ct", "voters", "dupes.csv.xz")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  cts <- ctv %>% 
    select(-id_voter) %>% 
    group_split(town_id)
  split_id <- split(ctv$id_voter, ctv$town_id)
  pb <- txtProgressBar(max = length(cts), style = 3)
  for (i in seq_along(cts)) {
    d1 <- duplicated(cts[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(cts[[i]], fromLast = TRUE)
      dupes <- tibble(id_voter = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      vroom_write(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(cts)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_tsv(
  file = xzfile(dupe_file),
  col_names = c("id_voter", "dupe_flag"),
  col_types = cols(
    id_voter = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(ctv)
ctv <- left_join(ctv, dupes, by = "id_voter")
ctv <- mutate(ctv, dupe_flag = !is.na(dupe_flag))
sum(ctv$dupe_flag)
```

We can see that, despite unique IDs, there are duplicate voters.

```{r dupe_view}
ctv %>% 
  filter(dupe_flag) %>% 
  select(id_voter, all_of(key_vars), town_name) %>% 
  arrange(nm_last)
```

### Categorical

```{r distinct_count}
col_stats(ctv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(ctv, cd_status)
explore_plot(ctv, cd_spec_status)
explore_plot(ctv, cd_off_reason)
explore_plot(ctv, dist)
explore_plot(ctv, prec)
explore_plot(ctv, congress)
explore_plot(ctv, cd_party)
explore_plot(ctv, cd_gender)
explore_plot(ctv, abstee_last)
```

### Dates

It looks like January 1st, 1800 is the default year for voters missing a birth
date. We can remove these values.

```{r date_count}
count(ctv, dt_accept, sort = TRUE)
ctv$dt_accept[ctv$dt_accept == "1800-01-01"] <- NA
```

We can add the year registered from `dt_accept` with `lubridate::year()`.

```{r date_year}
ctv <- mutate(
  .data = ctv, 
  yr_accept = year(dt_accept),
  yr_birth = year(dt_birth)
)
```

```{r date_range}
min(ctv$dt_accept, na.rm = TRUE)
max(ctv$dt_accept, na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
ctv %>% 
  filter(!is.na(yr_accept)) %>% 
  count(yr_accept) %>% 
  mutate(even = is_even(yr_accept)) %>% 
  ggplot(aes(x = yr_accept, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 10)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut Voters by Registered Year",
    caption = "Source: CT SOS",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

```{r bar_birth_year, echo=FALSE}
ctv %>% 
  filter(!is.na(yr_birth)) %>% 
  ggplot(aes(x = yr_birth)) +
  geom_histogram(binwidth = 5, fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2005, by = 10)) +
  coord_cartesian(xlim = c(1900, 2005)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut Voters by Birth Year",
    caption = "Source: CT SOS",
    x = "Year Made",
    y = "Count"
  )
```

```{r bar_reg_year, echo=FALSE}
ctv %>% 
  filter(!is.na(yr_birth)) %>% 
  count(yr_birth) %>% 
  ggplot(aes(x = yr_birth, y = n)) +
  geom_col() + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2005, by = 10)) +
  coord_cartesian(xlim = c(1900, 2005)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Connecticut Voters by Birth Year",
    caption = "Source: CT SOS",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
addr_norm_full <- ctv %>% 
  select(ad_num, nm_street, ad_unit) %>% 
  distinct() %>% 
  unite(
    col = ad_full,
    everything(),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    ad_norm = normal_address(
      address = ad_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-ad_full)
```

```{r address_join}
ctv <- left_join(ctv, addr_norm_full)
```

```{r address_view}
ctv %>% 
  sample_n(10) %>% 
  select(ad_num, nm_street, ad_unit, ad_norm)
```

### ZIP

The data has separate `zip5` and `zip4` columns, so we only need to remove a
handful of invalid "00000" ZIP codes to normalize the data.

```{r zip_norm}
ctv <- mutate(ctv, zip_norm = na_if(zip5, "00000"))
```

```{r zip_progress}
progress_table(
  ctv$zip5,
  ctv$zip_norm,
  compare = valid_zip
)
```

### State

The `st` variable does not need to be normalized.

```{r state_check}
prop_in(ctv$st, valid_state)
```

### City

Similarly, the `town_name` column only needs to be converted to uppercase. 

```{r town_check}
ctv <- mutate(ctv, town_norm = str_to_upper(town_name))
prop_in(ctv$town_norm, valid_city)
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
ctv <- ctv %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  relocate(town_clean, .after = ad_clean)
```

```{r clean_glimpse}
glimpse(sample_n(ctv, 50))
```

1. There are `r comma(nrow(ctv))` records in the database.
1. There are `r comma(sum(ctv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ctv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ct", "voters", "data", "clean"))
clean_path <- path(clean_dir, "ct_voters_clean.csv")
vroom_write(ctv, clean_path, delim = ",", na = "", progress = TRUE)
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
raw_cols <- filter(raw_cols, desc != "")
new_cols <- tibble(
  column = names(ctv)[44:54],
  length = NA,
  desc = c(
    "Source file (new, old)",
    "Date of last election",
    "Type of last election",
    "Voted absentee last election",
    "Flag indicating missing name",
    "Flag indicating duplicate row",
    "Year voter registered",
    "Year voter born",
    "Full normalized address",
    "Uppercase town name",
    "Normalized ZIP code"
  )
)
```

```{r dict_bind, echo=FALSE}
all_cols <- bind_rows(raw_cols, new_cols)
all_cols$length <- as.character(all_cols$length)
all_cols$length[is.na(all_cols$length)] <- ""
all_cols <- add_column(all_cols, type = map_chr(ctv, typeof), .before = "desc")
all_cols <- mutate(all_cols, across(column, md_code))
```

```{r dict_length, echo=FALSE}
max_char <- function(x) {
  x <- iconv(as.character(x), to = "ASCII", sub = "")
  max(nchar(x), na.rm = TRUE)
}

all_cols$length[44:54] <- vapply(
  X = ctv[, 44:54],
  FUN = max_char,
  FUN.VALUE = double(1)
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = all_cols,
  format = "markdown",
  col.names = c("Column", "Length", "Type", "Definition")
))
```
