---
title: "Pennsylvania Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("pa", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # format markdown
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  httr, # make http requests
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is from the [Pennsylvania Election and Campaign Finance System (ECF)][ecf].

[ecf]: https://www.dos.pa.gov/VotingElections/CandidatesCommittees/CampaignFinance/Pages/default.aspx

The ECF provides a [Full Campaign Finance Export][data]. From this page,
files are organized as annual directories containing files for contributions,
debt, expenditures, filer information, and receipts.

[data]: https://www.dos.pa.gov/VotingElections/CandidatesCommittees/CampaignFinance/Resources/Pages/FullCampaignFinanceExport.aspx

The ECF also provides a `readme.txt` file, which can be 

```{r vars_read}
pa_host <- "https://www.dos.pa.gov/VotingElections"
pa_dir <- "CandidatesCommittees/CampaignFinance/Resources/Documents"
readme_file <- "readme.txt"
readme_url <- paste(pa_host, pa_dir, readme_file, sep = "/")
```

```{r vars_print, echo=FALSE}
readme <- readme_url %>% 
  read_lines(skip = 4) %>% 
  str_c(collapse = "\r\n") %>% 
  str_split("\\W(?=\\w+\r\n-+)") %>% 
  pluck(1) %>%
  map(
    ~read_lines(.) %>% 
      str_trim() %>%
      str_remove("(?<=\\))(.*)$") %>% 
      enframe(name = NULL) %>% 
      separate(
        col = value, 
        into = c("col", "type"), 
        sep = "\t+",
        extra = "merge",
        fill = "right"
      ) %>% 
      na_if("") %>% 
      drop_na() %>% 
      mutate(col = str_to_lower(col))
  ) %>% 
  set_names(c("contribs", "debt", "expense", "filer", "receipt"))


readme$contribs %>% 
  mutate(col = as.character(md_code(col))) %>% 
  kable()

readme$filer %>% 
  mutate(col = as.character(md_code(col))) %>% 
  kable()
```

## Import

To import the files into R, we will have to first download the annual ZIP 
archive file containing all campaign finance transactions. Then we can extract
the contributions file and read them all into a single file for processing.

### Download

Each ZIP archive is simply named as the 4-digit year for the files within.
We can download each ZIP to the `/data/raw` directory.

```{r raw_download}
zip_names <- paste(2000:2020, "zip", sep = ".")
zip_urls <- paste(pa_host, pa_dir, zip_names, sep = "/")
raw_dir <- dir_create(here("pa", "contribs", "data", "raw"))
zip_paths <- path(raw_dir, zip_names)
if (!all_files_new(raw_dir)) {
  download.file(zip_urls, zip_paths)
}
```

Then we will unzip the annual directory from each archive.

```{r raw_unzip}
if (all(dir_ls(raw_dir) %in% zip_paths)) {
  for (zip in zip_paths) {
    unzip(zip, exdir = raw_dir)
  }
}
```

For each year, there is a file for contribution and a file for the information
on the recipients of those contributions, who file the reports containing the
data. We will identify the path of each file type in new vectors, which can then
be read together.

```{r con_move}
con_paths <- dir_ls(
  path = raw_dir, 
  recurse = TRUE, 
  regexp = "contrib[\\.|_]"
)

fil_paths <- dir_ls(
  path = raw_dir, 
  recurse = TRUE, 
  regexp = "filer[\\.|_]"
)
```

### Fix

To properly read so many records, we need to first perform some manipulation of
the text files. Each "cell" of character type columns are surrounded in double
quotation marks (`"`) to help prevent misreading. Howver, some of the text in
these cells itself contains double quotes or newline characters (`\n`).

We need to read each file as a character string and use regular expressions to
identify these erroneous characters and remove or replace them.

```{r raw_fix}
fix_check <- here("pa", "contribs", "data", "fixed.txt")
# do not repeat if done
if (!file_exists(fix_check)) {
  # for all contrib and filer files
  for (file in c(con_paths, fil_paths)) {
    # read raw file
    read_file(file) %>% 
      # force conversion to simple
      str_conv(encoding = "ASCII") %>% 
      # replace non-carriage newline
      str_replace_all("(?<!\r)\n", " ") %>%
      # replace not-field double quotes
      str_replace_all("(?<!^|,|\r\n)\"(?!,|\r\n|$)", "\'") %>% 
      # replace non-delim commas
      str_remove_all(",(?!\"|\\d|\\.\\d+|-(\\d|\\.))") %>% 
      # overwrite raw file
      write_file(file)
    # check progress
    message(paste(basename(file), "done"))
    # clean garbage memory
    gc()
  }
  # note this has done
  file_create(fix_check)
}
```

### Read

Now that each text file has been cleaned of irregularies, they can each be
properly read into R.

First, we will read all the annual contribution files into a single data frame
using `vroom::vroom()`. We need to use the column names and types listed in the
`readme.txt` file we downloaded earlier.

```{r raw_read}
pac <- map_df(
  .x = con_paths,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_names = readme$contribs$col,
  col_types = cols(
    .default = col_skip(),
    filerid = col_character(),
    eyear = col_integer(),
    cycle = col_integer(),
    section = col_character(),
    contributor = col_character(),
    address1 = col_character(),
    address2 = col_character(),
    city = col_character(),
    state = col_character(),
    zipcode = col_character(),
    occupation = col_character(),
    ename = col_character(),
    contdate1 = col_date("%Y%m%d"),
    contamt1 = col_double()
  )
)
```

Then we can read the fixed filers files.

```{r read_fils}
filers <- map_df(
  .x = fil_paths,
  .f = read_delim,
  delim = ",",
  escape_backslash = FALSE, 
  escape_double = FALSE,
  col_names = readme$filer$col,
  col_types = cols(
    .default = col_skip(),
    eyear = col_integer(),
    filerid = col_character(),
    filertype = col_character(),
    filername = col_character(),
    office = col_character(),
    district = col_character(),
    party = col_character(),
    address1 = col_character(),
    address2 = col_character(),
    city = col_character(),
    state = col_character(),
    zipcode = col_character(),
    county = col_character(),
    phone = col_character(),
  )
)
```

```{r raw_join}
pac <- left_join(
  x = pac,
  y = filers,
  by = c("filerid", "eyear"),
  suffix = c("_con", "_fil")
)

pac <- rename_prefix(
  df = pac,
  suffix = c("_con", "_fil"),
  punct = TRUE
)
```

```{r raw_rename}
pac <- pac %>% 
  rename(
    con_zip = con_zipcode,
    date = contdate1,
    amount = contamt1,
    fil_type = filertype,
    filer = filername,
    fil_zip = fil_zipcode,
    fil_phone = phone
  )
```

## Explore

```{r glimpse}
head(pac)
tail(pac)
glimpse(sample_n(pac, 20))
```

### Missing

```{r na_count}
col_stats(pac, count_na)
```

```{r na_flag}
pac <- pac %>% flag_na(date, contributor, amount, filer)
percent(mean(pac$na_flag), 0.01)
```

### Categorical

```{r n_distinct}
col_stats(pac, n_distinct)
```

```{r distinct_plot, fig.keep=FALSE}
explore_plot(mutate(pac, cycle = as.factor(cycle)), cycle)
explore_plot(pac, section)
explore_plot(pac, fil_type)
explore_plot(filter(pac, !is.na(office)), office)
explore_plot(filter(pac, !is.na(party)), party)
```

### Continuous

#### Amounts

```{r ammount_summary}
summary(pac$amount)
mean(pac$amount <= 0)
```

```{r hist_amount, echo=FALSE}
pac %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Pennsylvania Contributions Amount Distribution",
    subtitle = "from 2000 to 2020",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
pac <- mutate(pac, year = year(date))
```

```{r date_range}
min(pac$date)
sum(pac$year < 2000)
max(pac$date)
sum(pac$date > today())
```

```{r bar_year, echo=FALSE}
pac %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
pac <- pac %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
pac %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
pac <- pac %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  pac$zip,
  pac$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
pac <- pac %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
pac %>% 
  filter(state != state_norm) %>% 
  count(state, sort = TRUE)
```

```{r state_progress}
progress_table(
  pac$state,
  pac$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
pac <- pac %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("PA", "DC", "PENNSYLVANIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
pac <- pac %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- pac %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
pac <- pac %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(pac$city_raw),
  pac$city_norm,
  pac$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(pac$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
pac <- pac %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean"))
```

```{r clean_glimpse}
glimpse(sample_n(pac, 20))
```

1. There are `r comma(nrow(pac))` records in the database.
1. There are `r comma(sum(pac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(pac$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("pa", "contribs", "data", "clean"))
```

```{r clean_write}
write_csv(
  x = pac,
  path = path(clean_dir, "pa_contribs_clean.csv"),
  na = ""
)
```


