---
title: "Oklahoma Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ok", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests,
  zip, # file compress
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("ok/voters/docs/ok_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Voter data for Oklahoma can be obtained from the [State Election Board][seb],
which provides [a website][virs] where one can download data after requesting 
access. Users must first register with the state by submitting a form with a
name, organization, and phone number. Kiernan Nicholls has registered on behalf
of the Investigative Reporting Workshop. Once registered, users are provided a
password to access and download the data from the Voter Information Requests
(VIRS) website.

> #### **Home Page**. Access and Manage Voter Information Requests
> Voter registration information available in download files from the State
> Election Board must be imported into user software to sort the information by
> ZIP Code, name, household or other ways as desired. Lists or labels cannot be
> produced directly from these files.  
> 
> Data is written in CSV format and is refreshed on weekends. A download of the
> data could take anywhere from several seconds up to 10 minutes (or longer),
> depending on the size of the file and the download speed of the user’s ISP.  
> 
> There are no associated fees, but all users must apply to the State Election
> Board for a user account.  
> 
> For more information including file content layouts, please refer to the
[readme PDF][readme].  


> ##### We suggest the following:
> 1. **Getting Started**: Register - To access this application, you must first
  register with the Oklahoma State Election Board. Complete and return the final
  page of this readme to gain access. Credentials will be emailed to you once 
  the request is approved.  
> 2. **Voter Information Requests**: Download - Log in to access and download 
  the voter information files. All files are zipped for faster processing. 
  However, due to the potentially large file sizes, downloads may take several 
  minutes.  
  
[seb]: https://oklahoma.gov/elections.html
[virs]: https://services.okelections.us/virs/
[readme]: https://services.okelections.us/virs/Documents/readme.pdf

### About

We learn a little about the data from the VIRS [`readme.pdf`][readme] file.

> Voter registration information includes:
> * Voter's precinct number
> * Registration address
> * Voter's name
> * Date of birth
> * Voter's ID number
> * Date of registration
> * Political affiliation
> * Mailing address (if different from registration address)
> * Status
> * Election dates and method of voting
> * School District
> * Municipality
> * Technology Center District

> Information on each voter may include a registration address and/or a mailing
address or even no address. Some addresses may not be satisfactory for mailing
purposes; some information may be incomplete or missing.

> Voters having a rural or physical location address will have this information
split among the street number, street name, etc., fields. This information will
need to be concatenated to be meaningful. Mailing address fields are blank
unless the mailing address is different from the registration address.

> Telephone numbers and Email addresses are NOT included.

> Statewide orders consist of separate files by county or district. Caution!
Large statewide, county or Congressional District files may be larger than 100MB
each and require several minutes to download.

### Columns

The [`readme.pdf`][readme] file also provides a table of file columns.

|ELEMENT NAME                    |NOTES                                        |
|:-------------------------------|:--------------------------------------------|
|`Precinct number`               |First two characters denote county number    |
|`Last name`                     |                                             |
|`First name`                    |                                             |
|`Middle name`                   |                                             |
|`Suffix`                        |Jr, Sr, II, etc.                             |
|`Voter identification number`   |Unique identifier of each voter registration |
|`Political affiliation`         |DEM, REP, LIB, IND                           |
|`Status`                        |(A) active, (I) inactive                     |
|`Street/house number`           |                                             |
|`Street Direction`              |E, SW, NE, etc.                              |
|`Street Name`                   |                                             |
|`Street Type`                   |Ave, Blvd, St, etc.                          |
|`Street Post Direction`         |                                             |
|`Building Number`               |Apartment or suite number, etc.              |
|`City/municipality`             |                                             |
|`ZIP Code`                      |In ##### or #####-#### format                |
|`Birth Date`                    |CCYYMMDD                                     |
|`Date of Registration`          |CCYYMMDD                                     |
|`Mailing street address1`       |Free form 30-character address line          |
|`Mailing street address 2`      |Free form 30-character address line          |
|`Mailing address city`          |                                             |
|`Mailing address state`         |                                             |
|`Mailing address ZIP Code`      |In ##### or #####-#### format                |
|`Municipality`                  |                                             |
|`Municipal Ward`                |                                             |
|`School District`               |                                             |
|`School Board District`         |                                             |
|`Technical Center District`     |                                             |
|`Tech. Center Board District`   |                                             |
|`Election date` (repeats 10)    |CCYYMMDD                                     |
|`Method of voting` (repeats 10) |(See below)                                  |

`Method of voting`: 
* IP = Voted in person at polling place
* AI = Voted absentee in person
* AB = Absentee
* PI = Physically Incapacitated
* CI = Absentee - Care of Physically Incapacitated
* EI = Absentee - Emergency Incapacitated
* MI = Absentee - Military
* OV = Absentee - Overseas

A table is provided for the columns in the `precincts.csv` file included with
each voter information download.

|ELEMENT NAME                   |NOTES                                     |
|:------------------------------|:-----------------------------------------|
|`Precinct number`              |First two characters denote county number |
|`Congressional district`       |                                          |
|`State Senate district`        |                                          |
|`State House district`         |                                          |
|`County Commissioner district` |                                          |
|`Polling place`                |                                          |

## Obtain

On the download page, users can chose statewide files voter registration and
history files sorted by congressional district, county, or state house/senate
district. We will download the statewide voter registration by county.

|DESCRIPTION            |TYPE               |DATE       |  BYTES|  RECORDS|
|:----------------------|:------------------|:----------|------:|--------:|
|Congressional District |Voter History      |2021-03-27 |  97.6M| 24319097|
|Congressional District |Voter Registration |2021-03-27 | 112.6M|  2274902|
|County                 |Voter History      |2021-03-27 |  92.6M| 24319097|
|County                 |Voter Registration |2021-03-27 | 105.4M|  2274902|
|State House District   |Voter History      |2021-03-27 |  95.8M| 24319097|
|State House District   |Voter Registration |2021-03-27 | 101.1M|  2274902|
|State Senate District  |Voter History      |2021-03-27 |  95.5M| 24319097|
|State Senate District  |Voter Registration |2021-03-27 | 103.8M|  2274902|

```{r raw_dir}
raw_dir <- dir_create(here("ok", "voters", "data", "raw"))
raw_zip <- path(raw_dir, "CTYSW_VR_20210330111644.zip")
```

This file was last updated on March 27 and downloaded on March 30.

```{r raw_size}
file_size(raw_zip)
```

```{r raw_time}
raw_zip %>% 
  str_remove_all("\\D") %>% 
  ymd_hms(tz = Sys.timezone())
```

```{r zip_list, echo=FALSE}
zip_list <- function(zipfile) {
  out <- zip::zip_list(zipfile = zipfile)
  out$compressed_size <- fs::fs_bytes(out$compressed_size)
  out$uncompressed_size <- fs::fs_bytes(out$uncompressed_size)
  out$permissions <- fs::fs_perms(out$permissions)
  tibble::as_tibble(out)
}
```

```{r raw_list}
csv_list <- zip_list(raw_zip)
nrow(csv_list)
sum(csv_list$uncompressed_size)
max(csv_list$timestamp)
```

```{r raw_unzip}
unzip(raw_zip, exdir = raw_dir)
raw_csv <- path(raw_dir, str_subset(csv_list$filename, "CTY\\d{2}_vr.csv"))
all(file_exists(raw_csv))
```

## Read

There are `r length(raw_csv)` CSV text files (one for each county) with voter
registration data. We can read the rows from each file into a single data frame.

```{r read_reg}
okv <- map_df(
  .x = raw_csv,
  .f = read_delim,
  .id = "source_file", 
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    DateOfBirth = col_date("%m/%d/%Y"),
    OriginalRegistration  = col_date("%m/%d/%Y")
  )
)
```

We can ensure all the records have been read by comparing our count of rows
against the number given in the download table.

```{r read_check}
nrow(okv) == 2274902
```

We can also read the table of precinct polling places.

```{r read_precincts}
precincts <- read_csv(
  file = path(raw_dir, "precincts.csv"),
  col_types = cols(
    .default = col_character()
  )
)
precincts <- clean_names(precincts)
```

We will remove the voter history columns from this file.

```{r vote_hist}
okv <- okv %>% 
  clean_names(case = "snake") %>% 
  select(
    -starts_with("voter_hist"),
    -starts_with("hist_method")
  ) %>% 
  relocate(source_file, .after = last_col()) %>% 
  mutate(source_file = basename(raw_csv)[as.integer(source_file)]) %>% 
  rename(
    birth_date = date_of_birth,
    reg_date = original_registration
  )
```

## Explore

There are `r comma(nrow(okv))` rows of `r ncol(okv)` columns. Each record
represents a single registered voter in the state of Oklahoma.

```{r glimpse}
glimpse(okv)
tail(okv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(okv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("first_name", "last_name", "birth_date")
okv <- flag_na(okv, all_of(key_vars))
sum(okv$na_flag)
```

```{r na_view}
okv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("ok", "voters", "dupes.csv")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  oks <- okv %>% 
    select(-voter_id) %>% 
    group_split(school)
  split_id <- split(okv$voter_id, okv$school)
  pb <- txtProgressBar(max = length(oks), style = 3)
  for (i in seq_along(oks)) {
    d1 <- duplicated(oks[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(oks[[i]], fromLast = TRUE)
      dupes <- tibble(ncid = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      write_csv(dupes, xzfile(dupe_file), append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(oks)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_csv(
  file = dupe_file,
  col_names = c("voter_id", "dupe_flag"),
  col_types = cols(
    voter_id = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(okv)
okv <- left_join(okv, dupes, by = "voter_id")
okv <- mutate(okv, dupe_flag = !is.na(dupe_flag))
sum(okv$dupe_flag)
```

We can see that, despite unique IDs, there are some duplicate voters. The IDs
for these duplicates are sequential.

```{r dupe_view}
okv %>% 
  filter(dupe_flag) %>% 
  select(voter_id, all_of(key_vars), zip) %>% 
  arrange(last_name)
```

### Categorical

```{r distinct_count}
col_stats(okv, n_distinct)
```

```{r distinct_plots, echo=FALSE}
okv %>% 
  count(polital_aff) %>% 
  add_prop() %>% 
  ggplot(aes(x = reorder(polital_aff, -n), y = p)) +
  geom_col(aes(fill = polital_aff)) +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(
    guide = FALSE,
    values = c(
      DEM = "#2171b5",
      IND = "#525252",
      REP = "#cb181d",
      IND = "#238b45"
    )
  ) +
  labs(
    title = "Oklahoma Voters by Political Affiliation",
    x = "Political Affiliation",
    y = "Percent"
  )
explore_plot(okv, status)
explore_plot(okv, muni) + scale_x_wrap()
```

### Dates

We can add the calendar year from `reg_date` with `lubridate::year()`.

```{r date_year}
okv <- okv %>% 
  mutate(
    birth_year = year(birth_date), 
    reg_year = year(reg_date)
  )
```

```{r bar_reg_year, echo=FALSE}
okv %>% 
  count(reg_year) %>% 
  filter(!is.na(reg_year)) %>% 
  mutate(
    type = if_else(
      condition = reg_year %% 4 == 0,
      true = "pres.",
      false = if_else(
        condition = reg_year %% 2 == 0,
        true = "mid.",
        false = "none"
      )
    )
  ) %>% 
  ggplot(aes(x = reg_year, y = n)) +
  geom_col(aes(fill = type)) + 
  scale_fill_manual(values = c("#1B9E77", "#666666", "#D95F02")) +
  scale_y_continuous(labels = comma, n.breaks = 6) +
  scale_x_continuous(breaks = seq(1956, 2020, by = 4)) +
  coord_cartesian(xlim = c(1956, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oklahoma Voters Registered by Year",
    caption = "Source: OH VIRS",
    fill = "Cycle Registered",
    x = "Year Registered",
    y = "Count"
  )
```

```{r bar_birth_year, echo=FALSE}
okv %>% 
  filter(!is.na(birth_year)) %>% 
  ggplot(aes(x = birth_year)) +
  geom_histogram(fill = dark2["purple"], binwidth = 1) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1900, 2004, by = 4)) +
  coord_cartesian(xlim = c(1900, 2004)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Oklahoma Voters by Birth Year",
    caption = "Source: OK VIRS",
    x = "Year Born",
    y = "Count"
  )
```

## Wrangle

Almost all city names and ZIP codes are known to be valid.

```{r geo_check}
prop_in(okv$city, c(valid_city, extra_city))
prop_in(okv$zip, valid_zip)
```

We can add the state abbreviation manually.

```{r state_add}
okv <- mutate(okv, state = "OK", .after = city)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(okv, 50))
```

1. There are `r comma(nrow(okv))` records in the database.
1. There are `r comma(sum(okv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(okv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ok", "voters", "data", "clean"))
clean_path <- path(clean_dir, "ok_voters_clean.csv")
write_csv(okv, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

```{r echo=FALSE}
rm(okv)
Sys.sleep(5)
flush_memory()
```

```{r}
okl <- read_lines(clean_path)
na <- non_ascii(clean_path, highlight = cli::bg_yellow)
okl[na$row]
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
