---
title: "Michigan Contracts"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "mi", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  jsonlite, # read json files
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  readxl, # read excel files
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/mi/contracts/docs/mi_contracts_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

Michigan state contracts are obtained from the state Department of Technology,
Management & Budget on that [Contract Connect website][1]. 

> #### Welcome to Contract Connect!
>
> Contract Connect is State of Michigan Procurement's primary website. Contract
Connect is where we connect the vendor community to the resources they need to
do bid on projects and do business with the State of Michigan. 
>
> #### About State of Michigan Procurement
>
> State of Michigan Procurement is the State of Michigan's central purchasing
office and serves as the Department of Technology, Management and Budget's
(DTMB) central purchasing authority. Procurement handles all aspects of
high-level, complex bids so the State's agencies can achieve their missions and
get the best value possible for Michigan's residents.
>
> Our policies and procedures for purchasing across State departments in
accordance with the Management and Budget Act.

## About

> #### DTMB-Procurement Contract List Explained
> 
> The DTMB-Procurement Contract List, which is updated on a weekly basis, is the
Department of Technology, Management and Budget’s (DTMB) contract portfolio.
> 
> The list includes contracts executed by DTMB-Procurement, as well as some
multi-year contracts executed by agencies under delegated authority. It is
important to note that the dollar amounts shown for each multi-year contract are
estimates only, and are not necessarily reflective of the actual spend, which is
typically less.

> The list contains the following information:
>
* The contract number,  
* A link to the actual contract (click the "Open Contract" or "Open MiDEAL 
  Contract" link)  
* The state department that requires the goods and services covered in the 
  contract  
* The name of the current vendor  
* A brief description of the goods and services covered in the contract  
* The multi-year spend estimate for the contract (the cumulative value of the 
  contract since its effective date). This is the maximum that can be spent; 
  not actual dollars spent.  
* The current expiration date for the contract, which may change if option 
  years are exercised.  

> Additional information:
>
* Most contracts include base years and option years. The initial contract 
  period (base years) is typically three to five years. The number of option 
  years varies by contract, and they may or may not be exercised. To determine
  whether a contract will be bid, or an option will be exercised, contact the 
  buyer on the cover sheet of the contract.  
* A contract may not be rebid because there will not be a future or on-going 
  need for the commodity or service.  
* Delegated authority: Contracts may be handled by departments other than
  Technology, Management & Budget due to delegated purchasing authority granted
  to them via statute or policy/procedure.  

[1]: https://www.michigan.gov/dtmb/procurement/contractconnect

## Download

The DTMB Central Procurement Contract List (Excel) can be installed as an Excel
file.

```{r raw-dir}
raw_url <- "https://www.michigan.gov/documents/web_contract_12037_7.xls"
raw_dir <- dir_create(here("state", "mi", "contracts", "data", "raw"))
raw_xls <- path(raw_dir, basename(raw_url))
```

```{r raw-download}
if (!file_exists(raw_xls)) {
  download.file(raw_url, raw_xls)
}
```

## Read

That Excel file can then be read as a data frame.

```{r raw-read}
mic <- read_excel(
  path = raw_xls,
  sheet = excel_sheets(raw_xls)[2],
  .name_repair = make_clean_names
)
```

```{r raw-dates}
mic <- mic %>% 
  mutate(
    across(where(is.POSIXct), as_date)
  )
```

## Explore

There are `r comma(nrow(mic))` rows of `r ncol(mic)` columns. Each record
represents a single contract between the state of Michigan and a vendor.

```{r glimpse}
glimpse(mic)
tail(mic)
```

### Missing

Columns vary in their degree of missing values.

```{r na-count}
col_stats(mic, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na-flag}
key_vars <- c("effective_date", "agency_name", "cumulative_value", "vendor_name")
mic <- flag_na(mic, all_of(key_vars))
sum(mic$na_flag)
```

```{r na-view}
mic %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe-flag}
mic <- flag_dupes(mic, -contract_number)
```

There are no duplicate records.

### Categorical

```{r distinct-count}
col_stats(mic, n_distinct)
```

```{r distinct-plots, echo=FALSE, fig.height=3}
explore_plot(mic, agency_name) + scale_x_wrap()
explore_plot(mic, buyer_name) + scale_x_wrap()
explore_plot(mic, extended_purchasing) + scale_x_wrap()
```

### Amounts

```{r amount-round}
# fix floating point precision
mic$cumulative_value <- round(mic$cumulative_value, digits = 2)
```

```{r amount-summary}
summary(mic$cumulative_value)
mean(mic$cumulative_value <= 0, na.rm = TRUE)
```

These are the records with the minimum and maximum amounts.

```{r amount-minmax}
glimpse(mic[c(which.max(mic$cumulative_value), which.min(mic$cumulative_value)), ])
```

The distribution of amount values are typically log-normal.

```{r hist-amount, echo=FALSE}
mic %>%
  ggplot(aes(cumulative_value)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Michigan Contracts Amount Distribution",
    caption = "Source: DTMB",
    x = "Amount",
    y = "Count"
  )
```

There are a number of contracts for a single U.S. dollar.

```{r amount-one}
mic %>% 
  filter(cumulative_value == 1)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
mic <- mutate(mic, effective_year = year(effective_date))
```

```{r date-range}
min(mic$effective_date)
sum(mic$effective_year < 2000)
max(mic$effective_date)
sum(mic$effective_date > today())
```

It's common to see an increase in the number of contracts in elections years.

```{r bar-year, echo=FALSE}
mic %>% 
  count(effective_year) %>% 
  ggplot(aes(x = effective_year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2022, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Michigan Contracts by Year",
    caption = "Source: DTMB",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

```{r state-add}
mic <- mutate(mic, agency_state = "MI")
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(mic, 1000))
```

1. There are `r comma(nrow(mic))` records in the database.
1. There are `r comma(sum(mic$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(mic$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(mic$effective_date), "-")
max_dt <- str_remove_all(max(mic$effective_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("state", "mi", "contracts", "data", "clean"))
clean_csv <- path(clean_dir, glue("mi_contracts_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(mic, clean_csv, na = "")
write_rds(mic, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
