---
title: "HHS Provider Relief Fund"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "covid", "hhs_relief", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data regarding the monetary relief given to healthcare providers is published by
the [Health Resources and Services Administration (HRSA)][hrsa]. The data is
updated every Tuesday and Thursday.

[hrsa]: https://www.hrsa.gov/

Per the [dataset description page][dd]:

> The bipartisan CARES Act and the Paycheck Protection Program and Health Care
Enhancement Act provided $175 billion in relief funds to hospitals and other
healthcare providers on the front lines of the coronavirus response. Payments
from the $50 billion general distribution, the $12 billion targeted allocation
to high impact areas, and the $10 billion targeted allocation to rural providers
are being distributed by the Department of Health and Human Services through the
Health Resources and Services Administration. This dataset represents the list
of providers that received a payment from the General Distribution, High Impact
Targeted Allocation and/or the Rural Targeted Allocation of the Provider Relief
Fund and who have attested to receiving one or more payments and agreed to the
Terms and Conditions as of May 13, 2020.
>
> In addition to the list of providers who received funds from the General
Distribution, HRSA has provided an additional list of providers
(https://data.cdc.gov/Administrative/Provider-Relief-Fund-COVID-19-High-Impact-Payments/b58h-s9zx).

[dd]: https://data.cdc.gov/Administrative/HHS-Provider-Relief-Fund/kh8y-3es6

Each record is a payment:

> Each row is a provider (aligned with billing TIN) that has received and
accepted at least one payment. If a provider received 2 payments and accepted
both, both payments are included. If a provider received two payments and only
accepted one payment, only the accepted payment is included. If and when
additional data are available, this data portal will be updated here.

## Read

```{r raw_read}
hhspr <-  vroom(
  file = "https://data.cdc.gov/api/views/kh8y-3es6/rows.csv",
  .name_repair = make_clean_names,
  col_types = cols(
    `Provider Name` = col_character(),
    State = col_character(), 
    City = col_character(),
    Payment = col_number()
  )
)
```

```{r raw_rename, echo=FALSE}
hhspr <- rename(hhspr, provider = provider_name)
```

## Explore

We will add a column with a constant agency name.

```{r name_add}
hhspr <- mutate(
  .data = hhspr, 
  .before = provider,
  agency = "Health Resources and Services Administration"
)
```

The data is a 5-column table with payment recipients arranged alphabetically.

```{r glimpse}
glimpse(hhspr)
tail(hhspr)
```

### Missing

No records are missing any values.

```{r na_count}
col_stats(hhspr, count_na)
```

### Amounts

The `payment` values range from `r dollar(min(hhspr$payment))` to
`r dollar(max(hhspr$payment))`.

```{r ammount_summary}
noquote(map_chr(summary(hhspr$payment), dollar))
mean(hhspr$payment <= 0) # no negatives
```

Here are the smallest and largest payments:

```{r amount_minmax}
hhspr[which.min(hhspr$payment), ]
hhspr[which.max(hhspr$payment), ]
```

The distribution of `payment` is logarithmically-normally distributed, a typical
distribution for large payments.

```{r hist_amount, echo=FALSE}
hhspr %>%
  ggplot(aes(payment)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "United States Expenditures Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

There are no dates in the database. We will have to manually add the current
calendar year.

```{r year_add}
hhspr <- mutate(hhspr, year = year(today()))
```

## Geography

All of the `state` values are valid abbreviations.

```{r state_check}
percent(prop_in(hhspr$state, valid_state), 0.01)
add_prop(count(hhspr, state, sort = TRUE))
```

_Most_ of the city values are known to be valid as well.

```{r city_check}
many_city <- c(valid_city, extra_city)
percent(prop_in(hhspr$city, many_city), 0.01)
```

The only different between the `city` column and our list of known cities is
the abbreviations.

```{r city_count}
hhspr %>% 
  filter(city %out% many_city) %>% 
  count(city, state, sort = TRUE) %>% 
  add_prop(n = n)
```

```{r city_norm}
hhspr <- mutate(hhspr, across(city, normal_city, abbs = usps_city))
percent(prop_in(hhspr$city, many_city), 0.01)
```

We can also add "US" as the state spending the money.

## Conclude

1. There are `r comma(nrow(hhspr))` records in the database.
1. There are `r comma(sum(hhspr$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` seems reasonable.
1. There are `r comma(sum(hhspr$na_flag))` records missing key variables.
1. Consistency in geographic data is high.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("us", "covid", "hhs_relief", "data", "clean"))
clean_path <- path(clean_dir, "hhs_provider_relief.csv")
write_csv(hhspr, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, path.abbrev))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(hhspr)),
  type = md_code(map_chr(hhspr, typeof)),
  def = c(
    "Distributing agency name",
    "Provider name associated with the billing TIN",
    "Provider state abbreviation",
    "Provider city name (with expanded abbreviations)",
    "The cumulative payment that the provider",
    "Current calendar year"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```
