---
title: "Colorado Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  snakecase, # change string case
  lubridate, # datetime strings
  tidytext, # text analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  batman, # parse NA and LGL
  scales, # text formatting
  vroom, # quick file read
  knitr, # knit documents
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_campfin}
pacman::p_load_current_gh("kiernann/campfin")
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Colorado campaign expenditures data comes courtesy of Colorado Campaign Finance Disclosure Website,
which is managed by the TRACER reporting system (**Tra**nsparency in **C**ontribution and 
**E**xpenditure **R**eporting). Files can be found on the [Data Download][03] page.

[03]: http://tracer.sos.colorado.gov/PublicSite/DataDownload.aspx "source"

### Access

> You can access the Campaign Finance Data Download page to download contribution and expenditure
data for import into other applications such as Microsoft Excel or Access. A weekly batch process
is run that captures the year-to-date information for the current year. The data is available for
each calendar year. The file is downloaded in CSV format.

> This page provides comma separated value (CSV) downloads of contribution/donation, expenditure,
and loan data for each reporting year in a zipped file format. These files can be downloaded and
imported into other applications (Microsoft Excel, Microsoft Access, etc.). This data is extracted
from the Department of State database as it existed as of  7/20/2019  3:01 AM

### Quality

In the [TRACER FAQ file][04], the Secretary of State explains:

[04]: http://tracer.sos.colorado.gov/PublicSite/FAQ.aspx

> The information presented in the campaign finance database is, to the best of the ability of the
Secretary of State, an accurate representation of the disclosure reports filed with the applicable
office.It is suggested that the information found from reports data-entered by the Secretary of
State or County Clerks (which includes reports filed prior to 2010) be cross-checked with the
original document or scanned image of the original document.
> 
> Beginning in 2010, all candidates, committees, and political parties who file disclosure reports
with the Secretary of State must do so electronically using the TRACER system. Therefore, all data
contained in the database dated January 2010 onward reflects that data as entered by the reporting
person or entity.
>
> Prior to 2010, filers had the option of filing manual disclosure reports. Therefore, some of the
information in the campaign finance database dated prior to 2010was submitted in electronic form by
the candidate, committee or party, and some of the information was data-entered from paper reports
filed with the appropriate office. Sometimes items which are not consistent with filing
requirements, such as missing names and addresses or contributions that exceed the allowable
limits, are displayed when data is viewed online. Incorrect entries in the database typically
reflect incorrect or incomplete entries on manually filed reports submitted to the Secretary of
State or County Clerk. If you believe that there is a discrepancy in data dated prior to January
2010, please contact the appropriate filing officer for that data—the Secretary of State for
statewide candidates, committees, and parties; or the County Clerk for county candidates and
committees.

### Variables

TRACER also provides a [spreadsheet key][05].

[05]: http://tracer.sos.colorado.gov/PublicSite/Resources/DownloadDataFileKey.pdf

## Import

To wrangle the expenditures files in R, we will download the data locally and read everything into
a single tabular data frame.

### Download

To download the **immutable** raw data files, we first have to create the URLs. Files are split
annually, with only the 4-digit year differing in each the URL.

```{r glue_urls}
co_exp_urls <- glue(
  "http://tracer.sos.colorado.gov/PublicSite/Docs/BulkDataDownloads/{2000:2019}_ExpenditureData.csv.zip"
)
```

```{r cat_urls, echo=FALSE, results = 'asis'}
for (url in co_exp_urls){
  cat("*", url, "\n")
}
```

If the files have not yet been downloaded to the Colorado `/data/raw` directory, we can do so now.

```{r download_raw}
raw_dir <- here("co", "expends", "data", "raw")
dir_create(raw_dir)
if (!all_files_new(raw_dir)) {
  for (url in co_exp_urls) {
    download.file(
      url = url,
      destfile = glue("{raw_dir}/{basename(url)}")
    )
  }
}
```

### Read

Reading these files into a single data frame is not easy. First, we will unzip each file.

```{r unzip}
zip_files <- dir_ls(raw_dir, glob = "*.zip")
if (!all_files_new(path = raw_dir, glob = "*.csv")) {
  for (i in seq_along(zip_files)) {
    unzip(
      zipfile = zip_files[i],
      exdir = raw_dir
    )
  }
}
```

Then we have to read the lines of each file (without separating the columns). We need to extract a
header from one file, remove the headers from the rest, and filter out any row with an unexpected
number of delimiters.

```{r read_lines}
# read as unlisted lines
co_lines <- dir_ls(raw_dir, glob = "*ExpenditureData.csv") %>% map(read_lines) %>% unlist()
# extract header line
col_names <- co_lines[1]
# remove other headers
co_lines <- co_lines[-str_which(co_lines, col_names)]
# count expected delims
n_delim <- str_count(co_lines[1], "\",")
# convert header line
col_names <- to_snake_case(unlist(str_split(col_names, ",")))
# remove if unexpected num of delims
co_lines <- co_lines[-which(str_count(co_lines, "\",") != n_delim)]
```

Then, we replace all comma delimiters with a `\v` (vertical tab) to use as the delimiter.

```{r collapse_lines}
co <- co_lines %>% 
  str_replace_all("\",", "\"\v") %>% 
  str_remove_all("\"") %>% 
  str_c(collapse = "\n") %>% 
  read_delim(
    delim = "\v",
    col_names = col_names,
    col_types = cols(
      .default = col_character(),
      expenditure_amount = col_double(),
      expenditure_date = col_date("%Y-%m-%d %H:%M:%S"),
      filed_date = col_date("%Y-%m-%d %H:%M:%S")
    )
  )

rm(col_names)
```

Then we should parse some quasi-logical values.

```{r parse_char}
co <- co %>% 
  mutate_if(is_character, str_to_upper) %>% 
  mutate_if(is_character, na_if, "UNKNOWN") %>%
  remove_empty("cols") %>%
  remove_empty("rows") %>% 
  mutate(
    amended = to_logical(amended),
    amendment = to_logical(amendment),
  )
```

And finally save the formatted single data frame to disc.

```{r write_raw}
proc_dir <- here("co", "expends", "data", "processed")
dir_create(proc_dir)

write_csv(
  x = co,
  path = glue("{proc_dir}/co_expends.csv"),
  na = ""
)
```

## Explore

```{r glimpse}
head(co)
tail(co)
glimpse(sample_frac(co))
```

### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
glimpse_fun(co, n_distinct)
```

We can use `ggplot::geom_col()` to explore the distribution of the least distinct categorical 
variables.

```{r expend_type_bar, echo=FALSE}
co %>%
  filter(!is.na(expenditure_type)) %>% 
  count(expenditure_type, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(reorder(expenditure_type, n), n)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "CO Expends Purpose",
    y = "Count", 
    x = "Type"
  )
```

```{r payment_type_bar, echo=FALSE}
co %>%
  filter(!is.na(payment_type)) %>% 
  count(payment_type, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(reorder(payment_type, p), p)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "CO Expends Type",
    y = "Count", 
    x = "Type"
  )
```

```{r disburse_type_bar, echo=FALSE}
co %>%
  filter(!is.na(disbursement_type)) %>% 
  count(disbursement_type, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(reorder(disbursement_type, p), p)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "CO Expends Disbursement Type",
    y = "Count", 
    x = "Type"
  )
```

```{r committee_type_bar, echo=FALSE}
co %>%
  filter(!is.na(committee_type)) %>% 
  count(committee_type, sort = TRUE) %>%
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(reorder(committee_type, p), p)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  labs(
    title = "CO Expends Committee Type",
    y = "Count", 
    x = "Type"
  )
```

```{r jurisdiction_bar, echo=FALSE}
co %>%
  filter(!is.na(jurisdiction)) %>% 
  mutate(statewide = equals(jurisdiction, "STATEWIDE")) %>% 
  ggplot(aes(x = statewide)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "CO Expends Statewide Jurisdiction",
    y = "Count", 
    x = "Statewide"
  )
```

```{r explanation_bar, fig.height=10, echo=FALSE}
co %>% 
  unnest_tokens(word, explanation) %>% 
  count(word, sort = TRUE) %>% 
  anti_join(stop_words) %>% 
  drop_na() %>% 
  head(30) %>% 
  ggplot(
    mapping = aes(
      x = reorder(word, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "CO Expends Explanation",
    subtitle = "Text Analysis of Sentences",
    x = "Word",
    y = "Count"
  )
```

### Missing

The variables also differ in their degree of missing values.

```{r}
glimpse_fun(co, count_na)
```

It's important to note that there are zero missing values in important rows like `co_id`,
`expenditure_amount`, or `expenditure_date`.

There are `r percent(count_na(co$last_name)/nrow(co))` of records missing a `last_name` value used
to identify every individual or entity. If the record has no name whatsoever, we will flag it with
a new `na_flag` variable.

```{r flag_na}
co <- co %>% 
  mutate(
    na_flag = is.na(last_name) & is.na(first_name) & is.na(mi)
  )
```

### Ranges

For continuous variables, we should check the ranges.

#### Amount

```{r amount_ranges, collapse=TRUE}
summary(co$expenditure_amount)
sum(co$expenditure_amount < 0)
```

From this summary, we can see the median of `r dollar(median(co$expenditure_amount))` and mean of
`r dollar(mean(co$expenditure_amount))` are reasonable, but the minimum and maximum should be
explored.

```{r glimpse_min}
glimpse(filter(co, expenditure_amount == min(expenditure_amount)))
```

```{r min_explanation, echo=FALSE, results='asis'}
cat(">", filter(co, expenditure_amount == min(expenditure_amount))$explanation)
```

```{r glimpse_max}
glimpse(filter(co, expenditure_amount == max(expenditure_amount)))
```


We can use `ggplot2:geom_histogram()` and `ggplot2:geom_boxplot()` to explore the distribution of
the amount.

```{r amount_hist_log}
co %>% 
  ggplot(aes(x = expenditure_amount)) +
  geom_histogram() +
  scale_x_continuous(
    trans = "log10",
    labels = dollar
  )
```

```{r amount_box_pay}
co %>% 
  ggplot(aes(y = expenditure_amount)) +
  geom_boxplot(aes(x = payment_type), outlier.alpha = 0.01) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar,
    breaks = c(0, 1, 10, 100, 1000, 1000000)
  ) +
  coord_flip() +
  labs(
    title = "CO Expends Amount",
    subtitle = "by Payment Type",
    x = "Payment Type",
    y = "Amount (log)"
  )
```

```{r amount_box_type, echo=FALSE}
top_types <- co %>% 
  count(expenditure_type, sort = TRUE) %>% 
  drop_na() %>% 
  pull(expenditure_type) %>% 
  extract(1:5)

co %>% 
  filter(expenditure_type %in% top_types) %>% 
  ggplot(aes(y = expenditure_amount)) +
  geom_boxplot(aes(x = expenditure_type), outlier.alpha = 0.01) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar,
    breaks = c(0, 1, 10, 100, 1000, 1000000)
  ) +
  coord_flip() +
  labs(
    title = "CO Expends Amount",
    subtitle = "by Top Expenditure Purposes",
    x = "Purpose",
    y = "Amount (log)"
  )
```

### Dates

From the minimum and maximum expenditure dates, we can see that something is wrong.

```{r check_date, collapse=TRUE}
min(co$expenditure_date)
max(co$expenditure_date)

sum(co$expenditure_date > today())
sum(co$expenditure_date < "2002-01-01")
```

First, we will create a new `expenditure_year` variable from the `expenditure_date` using
`lubridate::year()` (after parsing with `readr::col_date()`).

```{r mutate_year}
co <- co %>% mutate(expenditure_year = year(expenditure_date))
```

Then we can see that there are a handful of expenditures supposedly made before 2002 and after
2019.

```{r count_date}
co %>% 
  count(expenditure_year) %>% 
  print(n = n_distinct(co$expenditure_year))
```

We can flag these broken dates with a new `date_flag` variable.

```{r flag_date, collapse=TRUE}
co <- co %>% mutate(date_flag = !between(expenditure_date, as_date("2002-01-01"), today()))
sum(co$date_flag)
```

We can also explore the intersection of `expenditure_date` and `expenditure_anount`.

```{r amount_month_line, echo=FALSE}
co %>% 
  mutate(on_year = expenditure_year %% 2 == 0) %>% 
  group_by(on_year, month = month(expenditure_date)) %>% 
  summarise(med_amount = median(expenditure_amount)) %>% 
  ggplot(mapping = aes(x = month, y = med_amount)) +
  geom_line(mapping = aes(color = on_year), size = 2) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "CO Expends Amount",
    subtitle = "by Month and Election Year",
    color = "Election Year",
    x = "Month", 
    y = "Median Amount"
  )
```

```{r amount_type_bar, echo=FALSE}
co %>% 
  mutate(on_year = expenditure_year %% 2 == 0) %>% 
  group_by(on_year, expenditure_type) %>% 
  summarise(med_amount = median(expenditure_amount)) %>% 
  ggplot(mapping = aes(x = expenditure_type, y = med_amount)) +
  geom_col(mapping = aes(fill = on_year), position = "dodge") +
  scale_y_continuous(labels = dollar) +
  coord_flip() +
  labs(
    title = "CO Expends Amount",
    subtitle = "by Expenditure Purpose",
    color = "Election Year",
    x = "Purpose", 
    y = "Median Amount"
  )
```

## Wrangle

### Address

```{r norm_address}
co <- co %>% 
  mutate(
    address_clean = 
      paste(address_1, address_2) %>% 
      normal_address(
        add_abbs = usps,
        na_rep = TRUE
      )
  )
```

### ZIP

```{r norm_zip, collapse=TRUE}
mean(co$zip %in% geo$zip)

co <- co %>% 
  mutate(
    zip_clean = normal_zip(
      zip = zip,
      na_rep = TRUE,
    )
  )

# percent changed
mean(co$zip != co$zip_clean, na.rm = TRUE)
# percent valid
mean(co$zip_clean %in% geo$zip)
```

### State

The `state` values appear to already be trimmed, or are otherwise nonsense. We can make them `NA`.

```{r norm_state, collapse=TRUE}
n_distinct(co$state)
mean(co$state %in% geo$state)
setdiff(co$state, geo$state)
co$state[which(co$state %out% geo$state)] <- NA
```

### City

```{r}

```

