---
title: "Colorado Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","co", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
here::here() # where does this document knit?
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

Colorado campaign expenditures data comes courtesy of Colorado Campaign Finance
Disclosure Website, which is managed by the TRACER reporting system
(**Tra**nsparency in **C**ontribution and **E**xpenditure **R**eporting). Files
can be found on the [Data Download][down] page.

[down]: http://tracer.sos.colorado.gov/PublicSite/DataDownload.aspx

### Access

> You can access the Campaign Finance Data Download page to download
contribution and expenditure data for import into other applications such as
Microsoft Excel or Access. A weekly batch process is run that captures the
year-to-date information for the current year. The data is available for each
calendar year. The file is downloaded in CSV format.

> This page provides comma separated value (CSV) downloads of
contribution/donation, expenditure, and loan data for each reporting year in a
zipped file format. These files can be downloaded and imported into other
applications (Microsoft Excel, Microsoft Access, etc.). This data is extracted
from the Department of State database as it existed as of  7/20/2019  3:01 AM

In the [TRACER FAQ file][faq], the Secretary of State explains:

[faq]: http://tracer.sos.colorado.gov/PublicSite/FAQ.aspx

> The information presented in the campaign finance database is, to the best of
the ability of the Secretary of State, an accurate representation of the
disclosure reports filed with the applicable office.It is suggested that the
information found from reports data-entered by the Secretary of State or County
Clerks (which includes reports filed prior to 2010) be cross-checked with the
original document or scanned image of the original document.
> 
> Beginning in 2010, all candidates, committees, and political parties who file
disclosure reports with the Secretary of State must do so electronically using
the TRACER system. Therefore, all data contained in the database dated January
2010 onward reflects that data as entered by the reporting person or entity.
>
> Prior to 2010, filers had the option of filing manual disclosure reports.
Therefore, some of the information in the campaign finance database dated prior
to 2010was submitted in electronic form by the candidate, committee or party,
and some of the information was data-entered from paper reports filed with the
appropriate office. Sometimes items which are not consistent with filing
requirements, such as missing names and addresses or contributions that exceed
the allowable limits, are displayed when data is viewed online. Incorrect
entries in the database typically reflect incorrect or incomplete entries on
manually filed reports submitted to the Secretary of State or County Clerk. If
you believe that there is a discrepancy in data dated prior to January 2010,
please contact the appropriate filing officer for that data—the Secretary of
State for statewide candidates, committees, and parties; or the County Clerk for
county candidates and committees.

TRACER also provides a PDF [spreadsheet key][key].

|Field Name            |Description                                            |
|:---------------------|:------------------------------------------------------|
|`CO_ID`               |Alphanumeric committee ID for the recipient committee. |
|`CONTRIBUTION AMOUNT` |Contribution Amount.                                   |
|`CONTRIBUTION DATE`   |Contribution Receipt Date.                             |
|`LAST NAME`           |Last Name (or entity name) of Contributor              |
|`FIRST NAME`          |Contributor First Name                                 |
|`MI`                  |Contributor Middle Initial                             |
|`SUFFIX`              |Contributor Name Suffix                                |
|`ADDRESS 1`           |Street, PO Box, or other directional information       |
|`ADDRESS 2`           |Suite/Apartment number, directional information        |
|`CITY`                |City.                                                  |
|`STATE`               |State.                                                 |
|`ZIP`                 |Zip Code.                                              |
|`EXPLANATION`         |Description provided for the contribution.             |
|`RECORD ID`           |Contribution internal ID issued by the SOS.            |
|`FILED DATE`          |                                                       |
|`CONTRIBUTION TYPE`   |Contribution Type.                                     |
|`RECEIPT TYPE`        |                                                       |
|`CONTRIBUTOR TYPE`    |                                                       |
|`ELECTIONEERING`      |YES or blank                                           |
|`COMMITTEE TYPE`      |                                                       |
|`COMMITTEE NAME`      |Name of the recipient committee.                       |
|`CANDIDATE NAME`      |Name of the recipient candidate..                      |
|`EMPLOYER`            |Contributor's employer.                                |
|`OCCUPATION`          |Contributor's occupation in cases from list.           |
|`AMENDED`             |Amendment was filed for this record. (Y/N)             |
|`AMENDMENT`           |Record is an amendment to a previously file. (Y/N)     |
|`AMENDED RECORD ID`   |Original record ID that was amended.                   |
|`JURISDICTION`        |This is the jurisdiction name.                         |
|`OCCUPATION COMMENTS` |If ‘Other’ is chosen for the occupation.               |

[key]: http://tracer.sos.colorado.gov/PublicSite/Resources/DownloadDataFileKey.pdf

## Import

We can download the annual ZIP archives directly from TRACER. This data is extracted from the Department of State database as it existed as of  10/29/2022 early morning. Since the last update took place 

```{r raw_dir}
raw_dir <- dir_create(here("state","co", "contribs", "data", "raw"))
raw_base <- "http://tracer.sos.colorado.gov/PublicSite/Docs/BulkDataDownloads/"
raw_urls <- str_c(raw_base, glue("{2020:2022}_ContributionData.csv.zip"))
raw_paths <- path(raw_dir, basename(raw_urls))
for (f in raw_paths) {
if (!this_file_new(f)) {
  download.file(raw_urls, raw_paths)
}  
}

```

Some slight adjustments need to be made properly read the text files. Every cell
is surrounded in double-quotes (`"`), meaning any double-quotes _within_ a cell
result parsing errors. We can read the lines of each file and replace these
with single-quotes.

```{r raw_fix}

fix_dir <- dir_create(path(dirname(raw_dir), "fix"))
for (r in raw_paths) {
  f <- path(fix_dir, path_ext_remove(basename(r)))
  read_lines(r) %>% 
    str_replace_all('(?<!,(?!\\s)|\n|^)"(?!,(?!\\s)|\n|$)', "'") %>% 
    str_replace_all('\n(?!"\\d+",)', " ") %>% 
    write_lines(f)
  message(path.abbrev(f))
}
fix_paths <- dir_ls(fix_dir)
```

```{r raw_read}
coc <- vroom(
  file = fix_paths,
  delim = ",",
  id = "source",
  escape_double = TRUE,
  escape_backslash = FALSE,
  .name_repair = make_clean_names,
  col_types = cols( # from key
    .default = col_character(),
    ContributionAmount = col_double(),
    ContributionDate = col_datetime(),
    FiledDate = col_datetime(),
  )
)
```

```{r raw_move}
coc <- relocate(coc, source, .after = last_col())
```

We can ensure these files were properly read by counting the distinct values of 
a discrete variable like the binary `amended` variable. There should only be two
values.

```{r raw_check}
count(coc, amended, sort = TRUE)
```

These binary variable should be converted to logical.

```{r raw_parse}
old_names <- names(coc)
coc <- coc %>% 
  mutate(across(c(amended, amendment), .fns = magrittr::equals, "Y")) %>% 
  mutate(across(electioneering, ~!is.na(.))) %>% 
  mutate(across(source, basename)) %>% 
  mutate(across(ends_with("date"), as_date)) %>% 
  rename_all(str_remove, "contribution_") %>% 
  rename_all(str_remove, "_name") %>% 
  rename(
    amend_id = amended_record_id,
    cont_type = contributor_type,
    comm_type = committee_type
  )
```

## Explore

There are `r comma(nrow(coc))` records with `r ncol(coc)` columns.

```{r glimpse}
glimpse(coc)
tail(coc)
```

```{r, include=FALSE}
co_prev <- read_csv(dir_ls(raw_dir,regexp = "co_contribs_clean.csv"))
co_prev$filed_date <- co_prev$filed_date %>% as.Date(format == "%Y-%d-%m")
```

We'll look at the filing date range. We'll filter out anything filed before Oct 1, 2020 as those records are already in our previous update.
```{r}
max(coc$filed_date)
min(coc$filed_date)

coc <- coc %>% 
  filter(filed_date >= as.Date("2020-10-01"))
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(coc, count_na)
```

After creating a single contributor name variable, we can flag any record
missing a name, date, or amount.

```{r na_flag}
coc <- coc %>% 
  unite(
    col = contributor,
    first, mi, last,
    sep = " ",
    na.rm = TRUE,
    remove = FALSE
  ) %>% 
  relocate(contributor, .after = last_col()) %>% 
  mutate(across(contributor, na_if, "")) %>% 
  flag_na(date, contributor, amount, committee)
```

`r percent(mean(coc$na_flag), 0.1)` of records are missing some value.

```{r na_check}
mean(coc$na_flag)
```

All of these records are missing the contributor `last` name. 

```{r na_view}
coc %>% 
  filter(na_flag) %>% 
  select(date, contributor, amount, committee)
```

Practically all of these values are "Non-Itemized" contributions.

```{r na_type_count}
coc %>% 
  filter(na_flag) %>% 
  count(type, sort = TRUE) %>% 
  add_prop()
```

```{r na_type_bar, echo=FALSE}
coc %>% 
  count(na_flag, type, sort = TRUE) %>% 
  filter(type %in% most_common(type)) %>% 
  ggplot(aes(type, n)) +
  geom_col(aes(fill = type)) +
  facet_wrap(~na_flag, scales = "free", ncol = 1) +
  scale_y_continuous(trans = "log10", labels = comma) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  labs(
    title = "Colorado Contributions Missing Names by Type",
    x = "Contribution type",
    y = "Count"
  )
```

### Duplicates

```{r dupe_flag}
d1 <- duplicated(select(coc, -record_id), fromLast = FALSE)
d2 <- duplicated(select(coc, -record_id), fromLast = TRUE)
coc <- mutate(coc, dupe_flag = d1 | d2)
rm(d1, d2); flush_memory()
```

```{r dupe_count}
percent(mean(coc$dupe_flag), 0.01)
```

```{r dupe_view}
coc %>% 
  filter(dupe_flag) %>% 
  arrange(date) %>% 
  select(date, last, amount, committee) %>% 
  arrange(desc(date))
```

Most duplicate records are also missing a key value.

```{r dupe_na}
mean(coc$dupe_flag[coc$na_flag])
```

### Categorical

```{r n_distinct}
col_stats(coc, n_distinct)
```

```{r echo=FALSE}
explore_plot(coc, source) + scale_x_truncate()
explore_plot(coc, receipt_type) + scale_x_truncate()
explore_plot(coc, comm_type) + scale_x_truncate()
explore_plot(coc, comm_type) + scale_x_truncate()
explore_plot(coc, occupation) + scale_x_truncate()
explore_plot(coc, jurisdiction) + scale_x_truncate()
```

### Continuous

#### Amounts

```{r ammount_summary}
coc <- coc %>% mutate(amount = as.numeric(amount))
summary(coc$amount)
mean(coc$amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
coc %>%
  filter(amount > 1, amount < 1e6) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"], bins = 20) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Colorado Contributions Amount Distribution",
    subtitle = "from 2020 to 2022",
    caption = "Source: CO TRACER",
    x = "Amount",
    y = "Count"
  )
```

#### Dates

```{r date_year}
coc <- mutate(coc, year = year(date))
#coc$year[which(coc$year > 2020 | coc$year < 2000)] <- NA
```

```{r date_range}
min(coc$date, na.rm = TRUE)
sum(coc$year < 2000, na.rm = TRUE)
max(coc$date, na.rm = TRUE)
sum(coc$date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
coc %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(min(year(coc$date)), max(year(coc$date)))) +
  theme(legend.position = "bottom") +
  labs(
    title = "Colorado Contributions by Year",
    caption = "Source: CO TRACER",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
coc <- coc %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
coc %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
coc <- coc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  coc$zip,
  coc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
coc <- coc %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = NULL
    )
  )
```

```{r}
coc %>% 
  filter(state_norm %out% valid_state) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  coc$state,
  coc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
coc <- coc %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("CO", "DC", "COLORADO"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
coc <- coc %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  ) %>% 
  rename(city = city_raw)
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- coc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
coc <- coc %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_view}
many_city <- c(valid_city, extra_city, "PUEBLO WEST", "CASTLE PINES")
coc$city_refine <- na_if(coc$city_refine, "UNKNOWNCITY")
coc %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, state_norm, sort = TRUE)
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(coc$city),
  coc$city_norm,
  coc$city_swap,
  coc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(coc$city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Colorado City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Colorado City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
coc <- coc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_frac(coc))
```

1. There are `r comma(nrow(coc))` records in the database.
1. There are `r comma(sum(coc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(coc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("state","co", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "co_contribs_clean_20201001-20221028.csv")
write_csv(coc, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```
