---
title: "Data Diary"
subtitle: "State Data"
author: "First Last"
date: "` format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = FALSE,
  echo = TRUE,
  dfrning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
# install.packages("pacman")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  RSocrata, # read soda api
  janitor, # dataframe clean
  zipcode, # clean & database
  explore, # basic exploration
  batman, # parse yes & no
  refinr, # cluster & merge
  rvest, # scrape website
  knitr, # knit documents
  here, # relative storage
  fs # search storage 
)
```

```{r fix_fun, echo=FALSE}
# fix conflict
here <- here::here
# custom utility functions
"%out%" <- Negate("%in%")
print_all <- function(df) df %>% print(n = nrow(.)) 
# load data
data("zipcode")
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Describe _where_ the data is coming from. [Link to the data download][03] page if possible.

Describe the data set that is going to be cleaned. A file name, age, and unit of observation.

[03]: https://example.com "source"

### About

> If the publisher provides any information on the file, you can directly quote that here.

### Variables

Often the publisher will provide a dictionary to describe the variables in the data (and
potentially the key pairs between many relational tables). 
[Link to the dictionary](https://example.com).

`variable_name`:

> Directly quote the definition given for variables of interest.

## Import

### Download

Download raw, immutable data file.

```{r download_raw}
raw_dir <- here("df", "type", "data", "raw")
dir_create(raw_dir)
download.file(
  url = url,
  destfile = here(raw_dir, "file_name.zip")
)
```

### Unzip

If needed, unzip into the same directory

```{r unzip_raw}
unzip(
  zipfile = here(raw_dir, "file_name.zip"),
  exdir = raw_dir
)
```

### Read

```{r read_raw}
df <- read_delim(
  file = here(raw_dir, "file_name.csv"),
  delim = ",",
  col_types = cols(
    
  )
)
```

## Explore

There are `nrow(df)` records of `length(df)` variables in the full database.

```{r glimpse}
describe(df)
glimpse(sample_frac(df))
```

### Distinct

The variables range in their degree of distinctness.

```{r n_distinct}
df %>% 
  map(n_distinct) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_distinct") %>% 
  mutate(prop_distinct = round(n_distinct / nrow(df), 4)) %>%
  print(n = length(df))
```

We can explore the distribution of the least distinct values with `ggplot2::geom_bar()`.

```{r plot_var1_bar, echo=FALSE}
ggplot(data = df) + 
  geom_bar(mapping = aes(var))
```

Or, filter the data and explore the most frequent discrete data.

```{r plot_var2_bar, echo=FALSE}
df %>% 
  count(var2) %>% 
  arrange(desc(n)) %>% 
  slice(1:10) %>% 
  ggplot(mapping = aes(x = var)) + 
  geom_bar(mapping = aes(color = var))
```

### Missing

The variables also vary in their degree of values that are `NA` (missing).

```{r count_na}
df %>% 
  map(function(var) sum(is.na(var))) %>% 
  unlist() %>% 
  enframe(name = "variable", value = "n_na") %>% 
  mutate(prop_na = n_na / nrow(df)) %>% 
  print(n = length(df))
```

We will flag any records with missing values in the key variables used to identify an expenditure.

```{r na_flag}
df <- df %>% mutate(na_flag = is.na(var))
```

### Duplicates

```{r get_dupes, collapse=TRUE}
df_dupes <- df %>% 
  get_dupes(
    
    ) %>% 
  mutate(dupe_flag = TRUE)

nrow(df_dupes)
```

```{r flag_dupes, message=FALSE, dfrning=FALSE}
df <- df %>% 
  left_join(df_dupes, by = "id") %>% 
  mutate(dupe_flag = !is.na(dupe_flag))
```

### Ranges

Explore the continuous variables with `ggplot2::geom_histogram()` and `base::summary()`

#### Amounts

Below are the smallest and largest expenditures.

```{r glimpse_min_max}
glimpse(df %>% filter(amount == min(amount, na.rm = T)))
glimpse(df %>% filter(amount == max(amount, na.rm = T)))
```

### Dates

```{r date_future, collapse=TRUE}
max(df$date, na.rm = TRUE)
sum(df$date > today(), na.rm = T)
```

```{r date_past, collapse=TRUE}
min(df$date, na.rm = TRUE)
sum(year(df$date) < 2007, na.rm = T)
```

## Wrangle

### Year

Add a `year` variable from `date` after `col_date()` using `lubridate::year()`.

```{r add_year}
df <- df %>% mutate(year = year(date))
```

### Address

The `address` variable should be minimally cleaned by removing punctuation and fixing white-space.

```{r}
df <- df %>% 
  mutate(
    address_clean = address %>% 
      str_to_upper() %>% 
      str_replace("-", " ") %>% 
      str_remove_all("[:punct:]") %>% 
      str_trim() %>% 
      str_squish() %>% 
      na_if("") %>% 
      na_if("NA")
  )
```

### Zipcode

Use the `zipcodes::clean.zipcodes()` function to strips the ZIP+4 digits and adds leading zeroes to
three or four digit strings. We will also make some common invalid zips `NA`.

```{r zipcode_data, echo=FALSE}
source(here("R", "prep_city.R"))
zipcode <-
  as_tibble(zipcode) %>% 
  select(city, state, zip) %>% 
  mutate(city = prep_city(city))
```

```{r clean_zipcodes}
df <- df %>% 
  mutate(zip_clean = zip %>% 
           clean.zipcodes() %>% 
           na_if("00000") %>% 
           na_if("11111") %>% 
           na_if("99999")
  )

df$zip_clean[which(nchar(df$zip_clean) != 5)] <- NA
```

### State

Using comprehensive list of state abbreviations in the Zipcodes database, we can isolate invalid
`state` values and manually correct them.

```{r valid_state, collapse=TRUE}
valid_state <- c(unique(zipcode$state), "AB", "BC", "MB", "NB", "NL", "NS", "ON", "PE", "QC", "SK")
length(valid_state)
setdiff(valid_state, state.abb)
```

```{r view_states}
df %>% 
  filter(state %out% valid_state) %>% 
  filter(!is.na(state)) %>% 
  select(
    id,
    address,
    city,
    state,
    zip
  ) %>% 
  print_all()
```

If the `state` abbreviation is invalid, use the `city` and `zip` values for that record to manually
correct the abbreviation where possible. If it can't be manually corrected, make the value `NA`.

```{r fix_states}
df$state_clean <- df$state %>% 
  str_replace("DG", "DF") %>% # add match
  str_remove("[:punct:]") %>% 
  na_if("")   %>% # empty
  na_if("RE") %>% # requested
  na_if("99") %>% # df zip
```

### City

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Prep

```{r prep_city}
source(here("R", "prep_city.R"))
df <- df %>% 
  mutate(
    city_prep = prep_city(
      cities = city,
      na = read_lines(here("R", "na_city.csv")),
      abbs = c("DF", "OR", "ID", "DC", "BC")
    )
  )
```

#### Swap

```{r match_dist}
df <- df %>%
  left_join() %>%
  rename() %>%
  mutate(
    match_dist = stringdist(),
    city_swap = if_else(match_dist == 1)
  )
```

#### Refine

```{r valid_city}
valid_city <- unique(zipcode$city)
```

```{r view_refine}
df_refined <- df %>%
  filter(var == "") %>% 
  filter(match_dist != 1) %>% 
  mutate(
    city_refine = var %>% 
      key_collision_merge(dict = valid_city) %>% 
      n_gram_merge(numgram = 1),
    refined = (city_swal != city_refine)
  ) %>% 
  filter(refined) %>% 
  select(
    
  ) %>% 
  rename(
    
  )
```

#### Review

```{r most_refined}
df_refined %>% 
  count(city_swap, city_refine) %>% 
  arrange(desc(n))
```

```{r count_refined}
refined_values <- unique(df_refined$city_refine)
count_refined <- tibble(
  city_refine = refined_values, 
  refine_count = NA
)

for (i in seq_along(refined_values)) {
  count_refined$refine_count[i] <- sum(str_detect(df$city_swap, refined_values[i]), na.rm = TRUE)
}

swap_values <- unique(df_refined$city_swap)
count_swap <- tibble(
  city_swap = swap_values, 
  swap_count = NA
)

for (i in seq_along(swap_values)) {
  count_swap$swap_count[i] <- sum(str_detect(df$city_swap, swap_values[i]), na.rm = TRUE)
}
```

```{r distinct_refined}
df_refined %>% 
  left_join(count_swap) %>% 
  left_join(count_refined) %>%
  select(
    city_match,
    city_swap,
    city_refine,
    swap_count,
    refine_count
  ) %>% 
  mutate(diff_count = refine_count - swap_count) %>%
  mutate(refine_dist = stringdist(city_swap, city_refine)) %>%
  distinct() %>%
  arrange(city_refine) %>% 
  print_all()
```

```{r refine_fix}
df_refined$city_refine <- df_refined$city_refine %>% 
  str_replace("^BAD FIX$", "GOOD FIX") %>% 
  str_replace("^BAD FIX$", "ORIGINAL") %>% 
  na_if("NA")

refine_table <- df_refined %>% 
  select(id, city_refine)
```

#### Merge

```{r join_refine}
df <- df %>% 
  left_join(refine_table, by = "id") %>% 
  mutate(city_clean = coalesce(city_refine, city_swap))
```

Each step of the cleaning process reduces the number of distinct city values.

## Conclude

1. There are `nrow(df)` records in the database
1. There are `sum(df$dupe_flag)` records with duplicate filer, recipient, date, _and_ amount
(flagged with `dupe_flag`)
1. The ranges for dates and amounts are reasonable
1. Consistency in strings has been fixed with `city_prep()` and the `stringr` package
1. The five-digit `zip_clean` variable has been created with `zipcode::clean.zipcode()`
1. The `expenditure_year` variable has been created with `lubridate::year()`
1. There are `sum(is.na(df$name))` records with missing `name` values and ` sum(is.na(df$date))`
records with missing `date` values (both flagged with the `na_flag`)

## Write

```{r write_clean}
df %>% 
  select(
   
  ) %>% 
  write_csv(
    path = here(),
    na = ""
  )
```

