---
title: "Mississippi Voters"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ms", "voters", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("ms/voters/docs/ms_voters_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The voter registration list for Mississippi was received under an open records
request pursuant to Miss. Code R. ยง 10-7.2 and the Mississippi Public Records
Act of 1983. This file was sent to the Investigative Reporting Workshop in
October of 2020 after a $20 fee was paid.

## Download

The received file has been hosted on the IRW server, where it can be downloaded.

```{r raw_dir}
raw_dir <- dir_create(here("ms", "voters", "data", "raw"))
raw_path <- path(raw_dir, "ms_voters_ORIG.csv")
aws_path <- path("FOR_REVIEW", "ms_voters_ORIG.csv")
```

```{r save_object}
if (!file.exists(raw_path)) {
  save_object(
    object = aws_path,
    bucket = "publicaccountability",
    file = raw_path,
    show_progress = TRUE
  )
}
```

## Read

```{r raw_read}
msv <- read_delim(
  file = raw_path,
  delim = ",",
  quote = "", # no quotes!
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    EFFECTIVE_REGN_DATE = col_date_mdy(),
    DATE_VOTED = col_date_mdy()
  )
)
```

```{r}
if (all(is.na(msv[[length(msv)]]))) {
  # remove empty trailing column
  msv <- select(msv, -last_col())
}
# lowercase column names
msv <- clean_names(msv, case = "snake")
# remove excess spaces
msv <- mutate(msv, across(where(is.character), str_squish))
msv <- msv %>% rename(
  regn_date = effective_regn_date,
  res_address = residential_address
)
```

To ensure the file has been properly read, we can count the unique values of a
discrete variable. Given the subject matter, we should expect every record to
have "MS" in the `res_state` column (otherwise they wouldn't be voters in 
Mississippi).

```{r}
count(msv, res_state)
```

## Explore

There are `r comma(nrow(msv))` rows of `r ncol(msv)` columns. Each record
represents a single registered in the state of Mississippi.

```{r glimpse}
glimpse(msv)
tail(msv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(msv, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("first_name", "last_name", "regn_date", "status")
msv <- flag_na(msv, all_of(key_vars))
sum(msv$na_flag)
```

`r comma(sum(msv$na_flag))` voters are missing a key variable (mostly a first
name).

```{r na_view}
msv %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars))
```

Almost all of these records are from purged voters.

```{r}
msv %>% 
  filter(na_flag) %>% 
  count(status) %>% 
  add_prop()
```

```{r}
msv %>% 
  filter(na_flag, status == "Active") %>% 
  select(all_of(key_vars))
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_file}
dupe_file <- here("ms", "voters", "dupes.csv")
```

```{r dupe_write}
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  mss <- msv %>% 
    select(-mapping_value, -mapping_value) %>% 
    group_split(res_county)
  split_id <- split(msv$mapping_value, msv$res_county)
  pb <- txtProgressBar(max = length(mss), style = 3)
  for (i in seq_along(mss)) {
    d1 <- duplicated(mss[[i]], fromLast = FALSE)
    if (any(d1)) {
      d2 <- duplicated(mss[[i]], fromLast = TRUE)
      dupes <- tibble(mapping_value = split_id[[i]], dupe_flag = d1 | d2)
      dupes <- filter(dupes, dupe_flag == TRUE)
      write_csv(dupes, dupe_file, append = TRUE)
      rm(d2, dupes)
    }
    rm(d1)
    flush_memory(1)
    setTxtProgressBar(pb, i)
  }
  rm(mss)
}
```

```{r dupe_read}
file_size(dupe_file)
dupes <- read_csv(
  file = dupe_file,
  col_names = c("mapping_value", "dupe_flag"),
  col_types = cols(
    mapping_value = col_character(),
    dupe_flag = col_logical()
  )
)
dupes <- distinct(dupes)
```

```{r dupe_join}
nrow(msv)
msv <- left_join(msv, dupes, by = "mapping_value")
msv <- mutate(msv, dupe_flag = !is.na(dupe_flag))
sum(msv$dupe_flag)
```

We can see that, despite unique IDs, there _are_ duplicate voters.

```{r dupe_view}
msv %>% 
  filter(dupe_flag) %>% 
  select(mapping_value, all_of(key_vars), res_zip_code) %>% 
  arrange(last_name)
```

### Categorical

```{r distinct_count}
col_stats(msv, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(msv, res_city)
explore_plot(msv, res_county)
explore_plot(msv, status)
explore_plot(msv, cong)
```

### Dates

There are thousands of voters with a registration date before January 1, 1801. 
These should be `NA` values, not real dates.

```{r date_min}
count_na(msv$regn_date)
min(msv$regn_date, na.rm = TRUE)
sum(msv$regn_date == "1801-01-01", na.rm = TRUE)
sum(msv$regn_date < "1801-01-01", na.rm = TRUE)
msv$regn_date[which(msv$regn_date <= "1801-01-01")] <- NA
```

```{r date_max}
max(msv$regn_date, na.rm = TRUE)
sum(msv$regn_date > today(), na.rm = TRUE)
```

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
msv <- mutate(msv, regn_year = year(regn_date))
```

```{r bar_reg_year, echo=FALSE}
msv %>% 
  filter(!is.na(regn_year)) %>% 
  count(regn_year) %>% 
  mutate(
    elect_type = case_when(
      regn_year %% 4 == 0 ~ "Pres.",
      regn_year %% 2 == 0 ~ "Mid.",
      TRUE ~ NA_character_
    )
  ) %>% 
  ggplot(aes(x = regn_year, y = n)) +
  geom_col(aes(fill = elect_type)) + 
  scale_fill_manual(
    values = c("#1B9E77", "#D95F02"),
    na.value = "#666666"
  ) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 4)) +
  coord_cartesian(xlim = c(1950, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Mississippi Voters Registered by Year",
    fill = "Election Year Type",
    x = "Year Registered",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

```{r}
sample(msv$res_address, 5)
prop_in(msv$res_zip_code, valid_zip)
unique(msv$res_state)
prop_in(msv$res_city, c(valid_city, extra_city))
```

```{r}
msv <- mutate(msv, zip_clean = normal_zip(res_zip_code))
prop_in(msv$zip_clean, valid_zip)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(msv, 50))
```

1. There are `r comma(nrow(msv))` records in the database.
1. There are `r comma(sum(msv$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(msv$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ms", "voters", "data", "clean"))
clean_path <- path(clean_dir, "ms_voters_2020.csv")
write_csv(msv, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

```{r echo=FALSE}
rm(msv)
flush_memory()
```

```{r non_ascii}
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
