---
title: "Connecticut Payroll Data Diary"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---


<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("ct", "payroll", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel files
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [State of Connecitut's data portal][data portal]. According to the website, the data was created September 21, 2016 and Last Updated on February 19, 2020. It comes from the Office of the State Comptroller and is updated annually.

> Payroll Data, including Calendar Year 2015 through the most recent pay period. Data used in the Open Payroll application

[data portal]: https://data.ct.gov/Government/State-Employee-Payroll-Data-Calendar-Year-2015-thr/virr-yb6n

```{r raw_dir}
raw_dir <- dir_create(here("ct", "payroll", "data", "raw"))
data_dir <- here("ct", "payroll", "data")
```

### Download
```{r download, eval=FALSE}
ctp_url <- "https://data.ct.gov/api/views/virr-yb6n/rows.csv?accessType=DOWNLOAD"
wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose", 
      "--content-disposition", 
      url, 
      paste("-P", raw_dir)
    )
  )
}

if (!all_files_new(raw_dir)) {
  wget(ctp_url, raw_dir)
}
```



### Import

```{r raw_read}
ctp <- dir_ls(raw_dir) %>% read_csv() 
# change column names into snake case, i.e. snake_case
ctp <- ctp %>% clean_names()
```

### Column Specs
Some of the date columns are read as plain text, which can be changed by applying `as.Date`. 
```{r convert columns}
ctp <- ctp %>% 
  mutate_at(.vars = vars(c(orig_hire, check_dt)), as.Date, format = "%m/%d/%Y")
```

### Explore
```{r raw_glimpse}
head(ctp)
tail(ctp)
glimpse(sample_frac(ctp))
```

Then we can take a look at the _NA_ fields and number of distinct values for each column. 
```{r n_distinct}
col_stats(ctp, count_na)
col_stats(ctp, n_distinct) 
```
We'll use the `campfin:na()` function to flag the records without any names and title description
```{r flag na}
ctp <- ctp %>% 
  flag_na(first_name, agency)
```

#### Categorical
```{r year_plot, echo=FALSE}
ctp %>% 
  count(calendar_year) %>% 
  ggplot(aes(x = calendar_year, y = n)) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  labs(
    title = "Connecticut Staff Payroll Head Counts per Year",
    x = "Year",
    y = "Staffers",
    caption = "Source: Office of the State Comptroller via data.ct.gov"
  )
```

#### Continuous
```{r hist, echo=FALSE}
ctp %>% 
  filter(tot_gross > 1) %>% 
  ggplot(aes(tot_gross)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Connecticut State Employee Income Amount Distribution",
    subtitle = "Total Gross from 2016 to 2020",
    caption = "Source: Office of the State Comptroller via data.ct.gov",
    x = "Amount",
    y = "Count"
  )
```

### Wrangle
To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to
facilitate this process.


### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
prop_in(ctp$postal, valid_zip, na.rm = T)

ctp <- ctp %>% 
    mutate(zip5 = normal_zip(postal, na_rep = TRUE))
```

### State
Running the following code, we will find that the `state` column does not need further normalization.

```{r state_norm}
prop_in(ctp$state, valid_state, na.rm = T)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.
#### Normal

The `campfin::normal_city()` function is a good ctpart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ctp <- ctp %>% 
      mutate(city_norm = normal_city(city,abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE))

prop_in(ctp$city_norm, valid_city, na.rm = T)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap lobbyist}
ctp <- ctp %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```



After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  ctp$city_raw,
  ctp$city_norm,
  ctp$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Connecticut Payroll City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "CT City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(ctp, 20))
```

1. There are `r nrow(ctp)` records in the database.
1. There are `r sum(ctp$dupe_flag)` duplicate records in the database.
1. The range and distribution of `year` seems mostly reasonable except for a few entries.
1. There are `r sum(ctp$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.

## Export

```{r proc_dir}
proc_dir <- dir_create(here("ct", "payroll", "data", "processed"))
```


```{r write_clean}
write_csv(
  x = ctp %>% rename(city_clean = city_swap) %>% 
    select(-city_norm),
  path = path(proc_dir, "ct_payroll_clean.csv"),
  na = ""
)
```

