---
title: "South Carolina Lobbying Expenditure Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("sc", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.


## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  readxl, # read excel files
  rvest, # used to scrape website and get html elements
  tidyverse, # data manipulation
  stringdist, # calculate distances between strings
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  httr, # http queries
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rstudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
Lobbyist data is obtained from the [South Carolina State Ethics Commission][03].

> #### Welcome
> Registrations for both lobbyists and their respective lobbyist's principals are available online
for viewing. Disclosure for both lobbyists and their respective lobbyist's principals will also be
available at the conclusion of the first disclosure period, June 30, 2009, for the period, January
1, 2009 through May 31, 2009.

The [lobbying activity page][04], we can see the files that can be retrieved:


> #### Lobbying Activity
> Welcome to the State Ethics Commission Online Public Disclosure and Accountability Reporting
System for Lobbying Activity. Registrations for both lobbyists and their respective lobbyist's
principals are available online for viewing.
> 
> Disclosure for both lobbyists and their respective lobbyist's principals are available for the
period June 30, 2009 through the present.
> 
> These filings can be accessed by searching individual reports by lobbyist and lobbyist’s
principal names and by complete list of current lobbyist and lobbyist’s principal registrations.

> #### List Reports
View a list of lobbyists, lobbyists' principals or their contact information.
> 
> * [Lobbyists and Their Principals](https://apps.sc.gov/LobbyingActivity/SelectLobbyistGroup.aspx)
> * [Download Lobbyist Contacts (CSV file)](https://apps.sc.gov/LobbyingActivity/DisplayCsv.aspx)
> * [Individual Lobbyist Lookup](https://apps.sc.gov/LobbyingActivity/SearchLobbyistContact.aspx)
> * [Lobbyists' Principals and Their Lobbyists](https://apps.sc.gov/LobbyingActivity/SelectLobbyistPrincipalGroup.aspx)
> * [Download Lobbyist's Principal Contacts (CSV file)](https://apps.sc.gov/LobbyingActivity/DisplayCsv.aspx)
> * [Individual Lobbyist's Principal Lookup](https://apps.sc.gov/LobbyingActivity/SearchLPContact.aspx)
> * [Year End Compilation Report](https://apps.sc.gov/LobbyingActivity/CompilationReport.aspx)

This data diary handles processing of the lobbyist compensation and expenditure data, which can be accessed from the `Year End Compilation Report` by year.


## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("sc", "lobby", "data", "raw", "exp"))
```
### Download
We can see that though 2007 and 2008 were selectable in the drowdown menu, there's actually no data.

### Read
```{r raw_read}
read_exp <- function(file_path){
  df <- read_lines(file_path) %>% 
  str_replace_all("(?<!^|,)\"(?!$|,)", "'") %>% 
  read_delim(",", escape_backslash = FALSE, escape_double = FALSE)
  #get rid of the extraneous column 'X27'
  df <- df %>% select(-ncol(df))
  df <- df %>% 
    clean_names() %>% 
    mutate(year = str_extract(file_path, "\\d{4}"))
}

scle <- map_dfr(dir_ls(raw_dir), read_exp)
```


## Explore

```{r glimpse}
head(scle)
tail(scle)
glimpse(sample_n(scle, 20))
```


### Missing
There're no missing fields 
```{r na_count}
col_stats(scle, count_na)
```

### Duplicates
There `r sum(scle$dupe_flag)` duplicate columns. 
```{r dupe_flag}
scle <- flag_dupes(scle, dplyr::everything())
sum(scle$dupe_flag)
```

### Continuous
All the amount columns are character columns that contain special characters like "$" and ",".We will turn them into numeric columns.

```{r col type}
scle <- scle %>% 
  mutate_at(.vars = vars(-c(lobbyist, lobbyists_principal, year, dupe_flag)), .funs = str_remove_all,"\\$|,") %>% 
  mutate_at(.vars = vars(-c(lobbyist, lobbyists_principal, year, dupe_flag)), .funs = as.numeric) %>% 
  mutate_if(is.character, str_to_upper)
```

```{r plot top receipients ecoh}
scle%>% 
  group_by(lobbyists_principal) %>% 
  summarize(med = median(total_income_and_expenditures_calendar_year_total)) %>% 
  arrange(desc(med)) %>% 
  top_n(10) %>% 
  ggplot(aes(x = reorder(lobbyists_principal,med),
         y = med)) +
  geom_col(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  theme(legend.position = "none") +
  scale_x_discrete(labels = wrap_format(15)) +
  scale_y_continuous(labels = dollar) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Top 10 Highest-spending Lobbying Principals",
    subtitle = "(Measued by median total compensation and expenditures to lobbyists)",
    caption = "Source: South Carolina State Ethics Commission",
    x = "median expenses",
    y = 'dollar'
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(scle, 20))
```

1. There are `r nrow(scle)` records in the database.
1. There's no duplicate record in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(scle$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("sc", "lobby", "data", "processed","exp"))
```

```{r write_clean}
write_csv(
  x = scle,
  path = path(clean_dir, "sc_lob_exp_clean.csv"),
  na = ""
)
```
