
---
title: "Kentucky Contracts"
author: "Jennifer LaFleur/Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---
  
<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ky", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs, # local storage 
  stringi #string functions
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Kentucky contracts data were obtained via Public Records request by Nami Hijikata.


## Read

```{r read_data}
kyc <- read.csv("KY_Contracts.csv")
head(kyc)
```

There are two dates in the file. We used the year from start date as a separate year column. 

### Dates


```{r date_year}
kyc <- mutate(kyc, year = stri_sub(StartDate,-4))
```


```{r date_range}
prop_na(kyc$StartDate)
min(kyc$date, na.rm = TRUE)
sum(kyc$year < 2000, na.rm = TRUE)
max(kyc$date, na.rm = TRUE)
sum(kyc$date > today(), na.rm = TRUE)
kyc$year <- na_if(kyc$year, 9999)
```

```{r bar_year, echo=FALSE}
kyc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Kentucky Contracts by Year",
    x = "Start Year",
    y = "Count"
  )
```

Contracts have both `StartDate` and `EndDate`; we combine these two variables into a single date to represent the transaction. We also make the headers snake case. Replace the word NULL with actually NAs. Combine cab_name and dept_name.

```{r raw_parse, warning=TRUE}
kyc <- kyc %>% 
  clean_names("snake") %>% 
  replace(kyc == "NULL", NA) %>% 
  mutate(date = paste0(start_date," to ", end_date)) %>% 
  unite("full_agency", c(cab_name, dept_name), sep=", ", remove = FALSE, na.rm = TRUE) %>% 
  mutate(vend_city=toupper(vend_city))

```

## Explore


```{r glimpse}
glimpse(kyc)
tail(kyc)
```

### Missing

If we count the number of missing values per column, we can see a lot of the
values from the columns found only in one type of file are missing.

```{r na_count}
col_stats(kyc, count_na)
```


### Duplicates

There are no duplicate records in this database.

```{r dupe_flag, warning=TRUE}
kyc <- flag_dupes(kyc, everything())
```

### Categorical

```{r n_distinct}
col_stats(kyc, n_distinct)
```

```{r distinct_plot}
explore_plot(kyc, classification, na.rm = TRUE)
explore_plot(kyc, procurement_name, na.rm = TRUE)
```

### Amounts

```{r amount_summary}
mean(kyc$contract_amount)
noquote(map_chr(summary(kyc$contract_amount), dollar))
```


```{r hist_amount, echo=FALSE}
kyc %>%
  filter(contract_amount >= 1) %>% 
  ggplot(aes(contract_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:8)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Kentucky Contracts Amount Distribution",
    x = "Amount",
    y = "Count"
  )
```


## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `address` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
addr_norm <- kyc %>% 
  distinct(vend_address1) %>% 
  mutate(
    address_norm = normal_address(
      address = vend_address1,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
addr_norm
```

```{r address_join}
kyc <- left_join(kyc, addr_norm, by = "vend_address1")
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
kyc <- kyc %>% 
  mutate(
    zip_norm = normal_zip(
      zip = vend_zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  kyc$vend_zip,
  kyc$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
kyc <- kyc %>% 
  mutate(
    state_norm = normal_state(
      state = vend_state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
kyc %>% 
  filter(vend_state != state_norm) %>% 
  count(vend_state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  kyc$vend_state,
  kyc$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
norm_city <- kyc %>% 
  distinct(vend_city, state_norm, zip_norm) %>% 
  mutate(
    city_norm = normal_city(
      city = vend_city, 
      abbs = usps_city,
      states = c("KY", "DC", "KENTUCKY"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.


```{r city_swap}
norm_city <- norm_city %>% 
  rename(city_raw = vend_city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```


```{r city_rejoin}
kyc <- left_join(
  x = kyc,
  y = norm_city,
  by = c(
    "vend_city" = "city_raw", 
    "state_norm", 
    "zip_norm"
  )
)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/


```{r city_refine}
good_refine <- kyc %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
kyc <- kyc %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(kyc$vend_city),
  kyc$city_norm,
  kyc$city_swap,
  kyc$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(kyc$vend_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Kentucky City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Kentucky City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
kyc <- kyc %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw")) %>% 
  relocate(state_clean, zip_clean, .after = city_clean)
```

```{r clean_glimpse}
glimpse(sample_n(kyc, 50))
```


## Export

1. There are `r comma(nrow(kyc))` records in the database.
1. There are `r comma(sum(kyc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(kyc$na_flag))` records missing key variables.
1. There are no geographic variables to be normalized.
1. The 4-digit `year` variable has been created.

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ky", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "ky_contracts_clean.csv")
write_csv(kyc, clean_path, na = "")
file_size(clean_path)
```

The encoding of the exported file should be UTF-8 or ASCII.

```{r clean_encode}
enc <- system2("file", args = paste("-i", clean_path), stdout = TRUE)
str_replace_all(enc, clean_path, basename)
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(kyc)),
  type = md_code(map_chr(kyc, typeof)),
  both = FALSE,
  def = c(
     "link", 
     "contract_identifier",
     "classification",
     "classification_code",
     "cited_auth_description", 
     "contract_type_code",
     "doc_id",
     "branch_name",
     "full_agency",
     "cab_name",
     "dept_name",
     "reason_modification",
     "start_date",
     "end_date",
     "vend_cust_id",
     "vend_name",
     "vend_address1",
     "vend_city",
     "vend_state",
     "vend_zip",
     "procurement_name",
     "contract_amount",
     "year",
     "date",
     "address_clean",
     "city_clean",
     "state_clean",
     "zip_clean"
  )
)
```

```{r dict_md, echo=FALSE}
dict_raw$both[c(1:3, 5:6, 8, 13:14, 18:22)] <- TRUE
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Overlaped", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Kentucky Contracts Data Dictionary\n", dict_md),
  path = here("ky", "contracts", "ky_contracts_dict.md"),
)
```

