---
title: "Hawaii Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("hi", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data is obtained from the [Hawaii Awards & Notices Data System][hands]
where we can export active and archived contracts as an excel file.

> This page displays the notice of awards for procurements done by the State and
its municipalities. These are Award Notices ONLY, procurement records are
maintained by the agency that conducted the solicitation.

[hands]: https://hands.ehawaii.gov/hands/awards

### Read

The raw file can be read using `readxl::read_excel()`.

```{r raw_dir}
raw_dir <- dir_create(here("hi", "contracts", "data", "raw"))
raw_path <- path(raw_dir, "contract-awards-05142020045537AM.xlsx")
```

```{r raw_read}
hic <- read_excel(
  path = raw_path,
  skip = 3,
  .name_repair = make_clean_names
)
```

A few columns can be renamed for simplicity.

```{r raw_rename}
hic <- rename(
  .data = hic,
  date = date_awarded,
  amount = original_award_amount,
  id = solicitation_number
)
```

## Explore

```{r glimpse}
glimpse(hic)
tail(hic)
```

### Missing

A number of records are missing one of the names, dates, or numbers needed to
properly identify a transaction.

```{r na_count}
col_stats(hic, count_na)
```

We can flag these records with `campfin::flag_na()`.

```{r na_flag}
key_vars <- c("date", "department", "amount", "awardee")
hic <- hic %>% flag_na(all_of(key_vars))
sum(hic$na_flag)
```

```{r na_view}
hic %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac(1)
```

### Duplicates

A significant number of records are entirely duplicated across all columns,
including the supposedly unique contract `id` variable. We can flag these rows
as well.

```{r dupe_flag}
hic <- flag_dupes(hic, everything())
percent(mean(hic$dupe_flag), 0.01)
```

```{r dupe_view}
hic %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars))
```

### Categorical

```{r n_distinct}
col_stats(hic, n_distinct)
```

```{r distinct_bars}
explore_plot(hic, jurisdiction) + scale_x_truncate()
explore_plot(hic, department) + scale_x_truncate()
explore_plot(hic, category) + scale_x_truncate()
explore_plot(hic, island) + scale_x_truncate()
```

### Amounts

A fair number of records have an `amount` value equal or less than zero. This
is sometimes used to indicate a correction, but there is no way to tell.

```{r ammount_negative}
percent(mean(hic$amount <= 0), 0.01)
```

```{r amount_zero}
hic %>% 
  select(all_of(key_vars)) %>% 
  filter(amount <= 0)
```

We should also view the smallest and largest transactions for any discrepancies.

```{r amount_range}
noquote(map_chr(summary(hic$amount), dollar))
glimpse(hic[which.min(hic$amount), ]) # smallest
glimpse(hic[which.max(hic$amount), ]) # largest
```

Aside from those records equal to zero and a handful of fairly small `amount`
values, the overall distribution of `amount` looks to be what we'd expect.

```{r hist_amount, echo=FALSE}
hic %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Hawaii Contracts Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
hic <- mutate(hic, year = year(date))
```

A fair number of `date` values are `NA`, preventing us from creating the `year`.
There are also a handful of dates outside the range we would expect.

```{r date_range}
percent(prop_na(hic$date), 0.01)
min(hic$date, na.rm = TRUE)
sum(hic$year < 2000, na.rm = TRUE)
max(hic$date, na.rm = TRUE)
sum(hic$date > today(), na.rm = TRUE)
```

We can convert dates from the 22nd century to their 21st century equivalent.

```{r date_fix}
hic <- mutate(hic, across(date, ~as.Date(str_replace(., "^(21)", "20"))))
hic <- mutate(hic, across(date, ~as.Date(str_replace(., "^(18)", "19"))))
hic <- mutate(hic, year = year(date))
```

```{r bar_year, echo=FALSE}
hic %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  coord_cartesian(xlim = c(2006, 2020)) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Hawaii Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(hic, 20))
```

1. There are `r comma(nrow(hic))` records in the database.
1. There are `r comma(sum(hic$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(hic$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("hi", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "hi_contracts_clean.csv")
write_csv(hic, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, basename))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(hic)),
  type = md_code(map_chr(hic, typeof)),
  def = c(
    "Contract award date", 
    "Unique contract ID",
    "Contract title",
    "Contract award amount",
    "Awardee vendor name",
    "Awarding agency jurisdiction",
    "Awarding agency department",
    "Awarding agency division",
    "Contract category",
    "Vendor island",
    "Flag for missing key variables",
    "Flag for deplicate records",
    "Calendar year awarded"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Hawaii Contracts Data Dictionary\n", dict_md),
  path = here("hi", "contracts", "hi_contracts_dict.md"),
)
```

