---
title: "Louisiana Lobbying Expenditure Data Diary"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
   github_document:    
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("la", "lobbyists", "docs"))
```
## Project


The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.


Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:


1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved


## Objectives


This document describes the process used to complete the following objectives:


1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction


## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  readxl, # import excel files
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  patchwork, # position plots
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  vroom, #read deliminated files
  readxl #read excel files
)
```

```{r fix_fun, echo=FALSE, collapse = TRUE}
here <- here::here
```
This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.


The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

# Data sources
The data was obtained from the Louisiana Ethics Administration Program via a public records request. Expenditures are as current as December 9, 2020.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
raw_dir <- here("la","lobbyists","data","raw","exp")
reg_proc <- here("la","lobbyists","data","processed","reg")
```

# Reading
We will notice that some rows were delimited incorrectly, as a supposedly single rows is separated into two lines with the first row of the overflow line an invalid forward slash.

```{r}

la_lines <- read_lines(dir_ls(raw_dir, glob = "*.csv"))
la_cols <- str_split(la_lines[1], ",", simplify = TRUE)
la_lines <- la_lines[-1]

sum(str_detect(la_lines, "^\\D"))
#> 4373

for (i in rev(seq_along(la_lines))) {
  if (is.na(la_lines[i])) {
    next()
  }
  if (str_detect(la_lines[i], "^\\D")) {
    la_lines[i - 1] <- str_c(la_lines[i - 1], la_lines[i], collapse = ",")
    la_lines[i] <- NA_character_
  }
}

la_lines <- na.omit(la_lines)

la_exp <- read_csv(la_lines, col_names = la_cols) %>% clean_names() %>% 
  mutate_if(is.character, str_to_upper) %>% 
  rename(unique_id = fake_unique_id)
```


```{r}
la_exp %>% glimpse()
```

We'll clean the fields that contain strings like "NULL" a bit.
```{r}
la_exp <- la_exp %>% 
  mutate_if(is.character,na_if, "NULL")
```

##Duplicates
```{r flag_dupe}
la_exp <- flag_dupes(la_exp, dplyr::everything())
```

##Missing
```{r missing}
col_stats(la_exp, count_na)
```
We'll flag that instances where the lobbyist name is missing.
```{r}
la_exp <- la_exp %>% flag_na(last_name_first_name, report_period, amount)
```

## Join
```{r join reg}
la_reg <- read_csv(file.path(here("la", "lobbyists", "data", "processed", "reg"), "la_reg_clean.csv"))

la_exp <- la_exp %>% mutate(year = as.numeric(year(date_filed)))

la_reg <- la_reg %>% select(unique_id, first_name, middle, last_name, m_street_norm, m_city_clean, m_state_norm, m_zip,year_registered) %>% 
  rename(year = year_registered,
         lob_first_name = first_name,
         lob_middle = middle,
         lob_last_name = last_name)
la_reg <- flag_dupes(la_reg, dplyr::everything(), .both = FALSE)
  # unite(first_name, middle, col = lob_first_middle, na.rm = TRUE, remove = FALSE, sep = " ") %>% 
  # unite(last_name, lob_first_middle, col = lob_full_name, na.rm = TRUE, remove = FALSE, sep = ", ") %>% 
la_exp <- la_reg %>% filter(!dupe_flag) %>% select(-dupe_flag) %>% 
right_join(la_exp, by = c('unique_id', 'year'))
```

##Explore
```{r histogram distribution, echo=F}
la_exp %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  geom_vline(xintercept = median(la_exp$amount[la_exp$amount != 0], na.rm = TRUE), linetype = 2) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Louisiana Lobbying Expenditure Amount Distribution",
    caption = "Source: Louisiana Ethics Administration Program",
    x = "Amount",
    y = "Count"
  ) +
  theme_minimal()
```


```{r}
la_exp %>% 
  group_by(branch) %>% 
  ggplot(aes(branch)) +
  geom_bar(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  labs(
    title = "Louisiana Lobbying Expenditure Amount by Lobbying Branch",
    caption = "Source: Louisiana Ethics Administration Program",
    x = "Branch",
    y = "Count"
  )

```
### Top officials with highest total associated lobbying expenses
```{r echo=FALSE}
la_exp %>% 
  unite(first_name, last_name, sep = " ", na.rm = TRUE, col = "officials_legislators", remove = FALSE) %>% 
  group_by(officials_legislators) %>% 
  summarize(total_expenses = sum(amount)) %>% 
  arrange(desc(total_expenses)) %>% 
  filter(officials_legislators != "") %>% 
  head(10) %>% 
  ggplot(aes(x = officials_legislators, y = total_expenses)) +
  geom_col(fill =  RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  scale_x_discrete(labels = wrap_format(10)) +
  scale_y_continuous(labels = dollar) +
  theme_minimal() +
    labs(
    title = "Louisiana Officials or Legislator with the Most Total Lobbying Expenses Paid",
    caption = "Source: Louisiana Ethics Administration Program",
    x = "Official/Legislator",
    y = "Total Expenses"
  )
```

## Wrangle
### Date
We'll tweak the `report_period` column to follow the 4-digit format.
```{r year}
la_exp <- la_exp %>% 
  mutate(year = str_extract(str_replace(report_period, " 09", " 2009"), "\\d{4}"))
```
### Name
```{r name, echo=FALSE, eval=FALSE}
la_exp <- la_exp %>% mutate(first_middle = str_match(last_name_first_name, ",\\s([^,]*$)")[,2]) %>% 
  mutate(lob_last_name = str_remove(last_name_first_name, ",\\s([^,]*$)"),
         lob_middle = str_match(first_middle, "\\s([^\\s]*$)")[,2],
         lob_first = if_else(condition = is.na(lob_middle), true = first_middle, 
                             false = str_remove(first_middle, lob_middle)))

col_stats(la_exp, count_na)
```

## Writing
```{r}
clean_exp_dir <- here("la", "lobbyists", "data", "processed", "exp")
dir_create(clean_exp_dir)
la_exp %>%
  rename(m_zip5 = m_zip) %>% 
  write_csv(path = glue("{clean_exp_dir}/la_exp_clean.csv"),
            na = "")
```

