---
title: "Mississippi Lobbying Registration Data Diary"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project


The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.


Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:


1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved


## Objectives


This document describes the process used to complete the following objectives:


1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction


## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  rvest, # read html tables
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  vroom, #read deliminated files
  readxl #read excel files
)
```

```{r fix_fun, echo=FALSE, collapse = TRUE}
# fix conflict
here <- here::here
print_all <- function(df) df %>% print(n = nrow(.)) 
```
This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.


The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.


## Download
Set the download directory first.
```{r create raw_dir}
# create a directory for the raw data
raw_dir <- here("ms", "lobby", "data", "raw","reg")

dir_create(raw_dir)
```
According to [The Secretary of State’s Office] [03], 

> Mississippi law defines “lobbying” as:
 influencing or attempting to influence legislative or executive action through oral or
written communication;
 solicitation of others to influence legislative or executive action;
 paying or promising to pay anything of value directly or indirectly related to
legislative or executive action. (Miss. Code Ann. § 5-8-3(k)).  

> Mississippi law defines a “lobbyist’s client” as:
 an entity or person in whose behalf the lobbyist influences or attempts to influence
legislative or executive action.  

> Mississippi law defines a “lobbyist” as an individual who:
 is employed and receives payments, or who contracts for economic consideration,
including reimbursement for reasonable travel and living expenses, for the purpose of
lobbying;
 represents a legislative or public official or public employee, or who represents a
person, organization, association or other group, for the purpose of lobbying; or is a sole proprietor, owner, part owner, or shareholder in a business, who has a
pecuniary interest in legislative or executive action, who engages in lobbying
activities. (Miss. Code Ann. § 5-8-3(l)).  

> Mississippi law excludes certain individuals from the definition of “lobbyist” and
“lobbyist’s client” who are exempt from the registration and reporting requirements of the
Lobbying Law Reform Act of 1994. See, Miss. Code Ann. § 5-8-7 at Appendix D for further
information. 

[03]: https://www.sos.ms.gov/Elections-Voting/Documents/2019%20Lobbying%20Guide.pdf
This Rmd file documents the MS registration data only, whereas the expenditure data is wrangled in a separate data diary.

```{r lobbyists, eval = FALSE}

ms_url_1 <- glue("https://sos.ms.gov/elec/Config/Mississippi/Elections/Handlers/Export{c('Client','Lobbyist')}")
ms_url_2 <- glue("DirectoryToExcel.ashx?type=1&year={2009:2019}&lastName=&firstName=&clientName=, ")

ms_url <- ms_url_1 %>% map(str_c,ms_url_2) %>% unlist()

dest_file <- glue("{raw_dir}/ms_lobby_{c('cl','lb')}") %>% map(glue,"_{2009:2019}.xls") %>% unlist()

if (!all_files_new(raw_dir)) {
    download.file(
      url = ms_url,
      destfile = dest_file)
  Sys.sleep(1)
}

```

Then, we'll merge each dataset into a master dataset. Note that there is no date or year field in the individual databases, and we will need to create such fields.
[Miss.
Code Ann. §§ 5-8-1, et seq.][03] regulates lobbyists registration as such.  

> Current lobbying reporting processes applicable to all lobbyists and their respective
clients in the State of Mississippi are the result of the Lobbying Law Reform Act of 1994. (Miss.
Code Ann. §§ 5-8-1, et seq.). Unless excepted from the statutory definition of “lobbyist,” every
lobbyist and lobbyist’s client must file a registration statement with the Secretary of State’s
Office within five (5) calendar days after becoming a lobbyist or lobbyist’s client, or beginning
to lobby on behalf of a new client.   
> The Mississippi lobbying cycle begins on January 1 and
ends on December 31 of each calendar year and registration is required each calendar year.
All lobbying reports for the cycle are filed electronically with the Secretary of State’s
Office. Mississippi statute requires all registered lobbyists to file three (3) reports during the
lobbying cycle:  
  - Legislative Mid-Session Report due on February 25th  
  - Legislative End-of-Session Report due within ten (10) days after sine die  
  - Annual Report of Expenditures due no later than January 30th  
  (Miss. Code Ann. § 5-8-11(5)(6)).  
  Registered lobbyists’ clients file one (1) report, the Annual Report, with the Secretary
  of State’s Office during the lobbying cycle. 

## Reading
We discovered that the xls files are actually structured as html tables. We'll use the `rvest` package to read these files.
```{r read csv}
ms_cl_files <- list.files(raw_dir, pattern = "ms_lobby_cl.*", recursive = TRUE, full.names = TRUE)
ms_lb_files <- list.files(raw_dir, pattern = "ms_lobby_lb.*", recursive = TRUE, full.names = TRUE)
# Create function to read a html table
read_web_tb <- function(file){
  df <- read_html(file) %>% html_node("table") %>% html_table(header = T)
  return(df)
}

ms_lobby_cl <- ms_cl_files %>% map(read_web_tb) %>% 
                   bind_rows() %>% 
                   clean_names()


ms_lobby_lb <- ms_lb_files %>% map(read_web_tb) %>% 
                   bind_rows() %>% 
                   clean_names()
```
## Explore

### Duplicates

We'll use the `flag_dupes()` function to see if there are records identical to one another and flag the duplicates. A new variable `dupe_flag` will be created.

```{r flag dupe}
ms_lobby_lb <- flag_dupes(ms_lobby_lb, dplyr::everything())
ms_lobby_cl <- flag_dupes(ms_lobby_cl, dplyr::everything())
```
### Year
```{r create year}
ms_lobby_lb <- ms_lobby_lb %>% mutate (date = registration_date %>% as.Date(format = "%b %d %Y %I:%M%p"),
                                       year = year(date))
```


```{r }
ms_lobby_lb %>% 
  group_by(year) %>% 
  ggplot(aes(year)) +
  scale_x_continuous(breaks = 2010:2019) +
  geom_bar(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  labs(
    title = "Mississippi Lobbyists Registration by Year",
    caption = "Source: Mississippi Elections Division: Lobbying",
    x = "Year",
    y = "Count"
  )
```
### Missing
There's almost no empty fields in the two data frames. 
```{r count_na}
ms_lobby_cl  %>% col_stats(count_na)
ms_lobby_lb  %>% col_stats(count_na)
```
Few values are missing from the lobbyists database.

## Wrangling
We'll wrangle the two datasets to extract information such as address, city, ZIP, state, phone for both lobbyists and their clients, as well as authorization date. The lobbyists registry has the one-to-one relationship between lobbyists and clients, so we will use `ms_lobby_lb` as the main data frame and join the clients' information from the `ms_lobby_cl` data frame.

### Phone
```{r normal phone}
ms_lobby_cl <- ms_lobby_cl %>% mutate(telephone_norm = normal_phone(telephone))
ms_lobby_lb  <- ms_lobby_lb  %>% mutate(telephone_norm = normal_phone(telephone))
```

### Address
```{r normal address}
ms_lobby_lb <- ms_lobby_lb %>% 
  mutate(address_norm = normal_address(address = str_c(address_line1, address_line2, sep = " "),
      abbs = usps_city,
      na_rep = TRUE))

ms_lobby_cl <- ms_lobby_cl %>% 
  mutate(address_norm = normal_address(address = str_c(address_line1, address_line2, sep = " "),
      abbs = usps_city,
      na_rep = TRUE))
```


### ZIP 
The ZIP code fields need a little bit cleaning.
```{r client normal zip}
prop_in(ms_lobby_lb$postal_code, valid_zip, na.rm = TRUE) %>% percent()

ms_lobby_lb <- ms_lobby_lb %>% 
  mutate(zip5 = normal_zip(postal_code))
  
ms_lobby_cl <- ms_lobby_cl %>% 
  mutate(zip5 = normal_zip(postal_code))
```

### State
Running the following commands tells us the state fields are clean.
```{r clients clean state}
prop_in(ms_lobby_cl$state, valid_state, na.rm = TRUE) %>% percent()
prop_in(ms_lobby_lb$state, valid_state, na.rm = TRUE) %>% percent()
```
### City
The city fields in both data frames use upper-case letters and lower-case letters inconsistently. We'll convert everything to upper case.
```{r}
prop_in(ms_lobby_cl$city, valid_city, na.rm = TRUE) %>% percent()
prop_in(ms_lobby_lb$city, valid_city, na.rm = TRUE) %>% percent()
```

#### Normalize
```{r lb norm_city, collapse = TRUE}
ms_lobby_lb <- ms_lobby_lb %>% mutate(city_norm = normal_city(city = city,
                                            abbs = usps_city,
                                            states = c(valid_state),
                                            na = invalid_city,
                                            na_rep = TRUE))
n_distinct(ms_lobby_lb$city)
n_distinct(ms_lobby_lb$city_norm)

prop_in(ms_lobby_lb$city, valid_city, na.rm = TRUE)
prop_in(ms_lobby_lb$city_norm, valid_city, na.rm = TRUE)
```

```{r cl norm_city, collapse = TRUE}
ms_lobby_cl <- ms_lobby_cl %>% mutate(city_norm = normal_city(city = city,
                                            abbs = usps_city,
                                            states = c(valid_state),
                                            na = invalid_city,
                                            na_rep = TRUE))
n_distinct(ms_lobby_cl$city)
n_distinct(ms_lobby_cl$city_norm)

prop_in(ms_lobby_cl$city, valid_city, na.rm = TRUE)
prop_in(ms_lobby_cl$city_norm, valid_city, na.rm = TRUE)
```

#### Swap
Then, we will compare these normalized `city_norm` values to the _expected_ city value for that
vendor's ZIP code. If the [levenshtein distance][09] is less than 3, we can confidently swap these
two values.

[09]: https://en.wikipedia.org/wiki/Levenshtein_distance

```{r cl swap_city}
ms_lobby_cl <- ms_lobby_cl %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
  rename(city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )

prop_in(ms_lobby_cl$city_swap, valid_city, na.rm = TRUE) %>% percent()
```

```{r lb swap_city}
ms_lobby_lb <- ms_lobby_lb %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip5" = "zip"
    )
  ) %>% 
  rename(city_match = city.y,
         city = city.x) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = match_abb | match_dist == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )

prop_in(ms_lobby_lb$city_swap, valid_city, na.rm = TRUE) %>% percent()
```

Besides the `valid_city` vector, there is another vector of `extra_city` that contains other locales. We'll incorporate that in our comparison.

```{r valid_place check}
valid_place <- c(valid_city, extra_city) %>% unique()

progress_table(
  ms_lobby_cl$city,
  ms_lobby_cl$city_norm,
  ms_lobby_cl$city_swap,
  compare = valid_place
)

progress_table(
  ms_lobby_lb$city,
  ms_lobby_lb$city_norm,
  ms_lobby_lb$city_swap,
  compare = valid_place
)
```

This is a very fast way to increase the valid proportion in the lobbyist data frame to
`r percent(prop_in(ms_lobby_lb$city_swap, extra_city, na.rm = TRUE))` and reduce the number of distinct
_invalid_ values from `r length(setdiff(ms_lobby_lb$city_norm, valid_place))` to only
`r length(setdiff(ms_lobby_lb$city_swap, valid_place))`

Similarly, the valid proportion in the  clients data frame  was bumped up to
`r percent(prop_in(ms_lobby_cl$city_swap, extra_city, na.rm = TRUE))` and reduce the number of distinct
_invalid_ values from `r length(setdiff(ms_lobby_cl$city_norm, valid_place))` to only
`r length(setdiff(ms_lobby_cl$city_swap, valid_place))`

## Join
We'll join the two data frames together. Since there're no duplicate columns, we will delete the `dupe_flag` columns and add suffixes to each dataset's column names.
```{r join pre}
ms_lobby_lb$dupe_flag %>% tabyl()
ms_lobby_cl$dupe_flag %>% tabyl()

ms_lobby_cl <- ms_lobby_cl %>% 
  select(-c(dupe_flag,
            city_norm)) %>% 
  rename(city_clean = city_swap) %>% 
  rename_at(.vars = vars(-(starts_with("client"))),
    .funs = ~str_c("client_",.))
```

```{r join}
ms_lobby_reg <- ms_lobby_cl %>% 
  right_join(ms_lobby_lb,
            by = c("client_name" = "client_name",
            "client_certification_number" = "certification_number",
            "client_registration_date" = "registration_date"))
```


## Export

```{r writeean}
clean_dir <- here("ms", "lobby", "data", "processed","reg")
dir_create(clean_dir)
ms_lobby_reg %>% 
  select(-c(dupe_flag,
            city_norm)) %>% 
  mutate_if(is.character, str_to_upper) %>% 
  rename(city_clean = city_swap) %>% 
  write_csv(
    path = glue("{clean_dir}/ms_lobby_reg.csv"),
    na = ""
  )
```


