---
title: "Philadelphia City Employee Earnings "
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("pa", "philadelphia", "salary", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("pa/philadelphia/salary/docs/philadelphia_earnings.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

[Employee earnings][ee] for the city of Philadelphia, PA are found on the city's 
[open data portal][odp]. The data comes from the Office of Human Resources and 
Office of the Director of Finance. Data is pushed from the OnePhilly payroll
system to a SQL database, extracted via a python script and loaded to
DataBridge. The data starts Q2 of 2019 through the most recent fiscal quarter.

### Notes

> Data from Calendar Year (CY) 2019 Q2 onwards is extracted from the newer
OnePhilly payroll system and has a different data schema than the legacy payroll
system. Therefore, one cannot and should not compare the previous earnings data
(from 2016 to 2019 Q1) with the new data.

> Major changes in the data from 2019 Q2 onwards include various new
earnings-related fields, a more detailed look at gross pay, and quarter-to-date
gross pay totals instead of year-to-date totals. In the old earnings data (i.e.
prior to CY 2019 Q2), some of the gross pay fields were lumped into the
`ANNUAL_SALARY` field, without the ability to separate them out. With the new
OnePhilly system, we are now able to show these as their own fields. For this
reason, it may look like people make less than they used to but that’s not the
case - it’s just that we’ve separated out the fields for greater clarity and
transparency. Now, under the `BASE_SALARY` field, you can see someone’s salary
(if salaried) separate from any other additional compensation they might
receive. Given the increased scope of this data afforded through the new
compensation fields, we have also renamed the dataset from 'Employee Salaries'
to 'Employee Earnings.'

> Gross pay fields: Added all together, the following 5 fields equal an
employee’s gross pay. Please note these are quarter-to-date totals. To get
yearly amounts, add up the amounts from each quarter.
>
> 1. `base_gross_pay_QTD`
> 2. `overtime_gross_pay_QTD`
> 3. `longevity_gross_pay_QTD`
> 4. `post_separation_payments_QTD`
> 5. `miscellaneous_gross_pay_QTD`

> With the exception of the `BASE_GROSS_PAY_QTD` field, the remaining four
fields above are earnings in addition to an employee’s `BASE_SALARY.`
> 
> A small number of records related to individuals under the age of 18 were
removed from this dataset to ensure privacy.

[ee]: https://www.opendataphilly.org/showcase/city-employee-earnings
[odp]: https://www.opendataphilly.org/showcase/city-employee-earnings

### Fields

```{r echo=FALSE}
f <- GET(
  url = "https://us-api.knack.com/v1/scenes/scene_142/views/view_287/records/export/applications/550c60d00711ffe12e9efc64",
  query = list(
    type = "csv",
    format = "both",
    page = 1,
    rows_per_page = 10,
    representationdetails_id = "604284dc49a209001d746460",
    sort_field = "field_17",
    sort_order = "asc"
  )
)
```

```{r echo=FALSE}
f <- read_csv(
  file = content(f, as = "text", encoding = "UTF-8"),
  col_types = cols(
    .default = col_character()
  )
)
```

```{r echo=FALSE}
f %>% 
  relocate(Type, .before = Description) %>% 
  mutate(
    across(`Field Name`, md_code)
  ) %>% 
  kable()
```

## Download

We can make a `GET` request to query the SQL database and return a CSV file that
can be read into R and uploaded to TAP.

Making note of the changes described in Q2 and 2019, we can also get the
_archived_ salary data. The data scheme will differ slightly, so these two
files will have to be read separately and somehow combined.

```{r raw_dir}
raw_dir <- dir_create(here("pa", "philadelphia", "salary", "data", "raw"))
raw_csv <- path(raw_dir, "employee_earnings.csv")
```

```{r get_2019}
if (!file_exists(raw_csv)) {
  get_2019 <- GET(
    url = "https://phl.carto.com/api/v2/sql",
    write_disk(raw_csv),
    progress("down"),
    query = list(
      filename = "employee_earnings",
      format = "csv",
      q = "SELECT * FROM employee_earnings",
      skipfields = "cartodb_id,the_geom,the_geom_webmercator"
    )
  )
}
```

## Read

The CSV files can be read as a data frames and combined together. There are
fewer columns in the older data, so any record from that source will be missing
data for the newer columns.

```{r raw_read}
phl <- read_delim(
  file = raw_csv,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    calendar_year = col_integer(),
    quarter = col_integer(),
    base_salary = col_double(),
    overtime_gross_pay_qtd = col_double(),
    base_gross_pay_qtd = col_double(),
    longevity_gross_pay_qtd = col_double(),
    post_separation_gross_pay_qtd = col_double(),
    miscellaneous_gross_pay_qtd = col_double(),
    termination_year = col_integer()
  )
)
```

## Explore

There are `r comma(nrow(phl))` rows of `r ncol(phl)` columns. Each record
represents a single employee earning in the city of Philadelphia.

```{r glimpse}
glimpse(phl)
tail(phl)
```

### Amounts

We can create a new column containing the total pay received across the five
earning categories described in the notes above.

```{r}
phl <- phl %>% 
  mutate(
    total_gross_pay = sum(
      base_salary,
      overtime_gross_pay_qtd,
      longevity_gross_pay_qtd,
      post_separation_gross_pay_qtd,
      miscellaneous_gross_pay_qtd,
      na.rm = TRUE
    )
  )
```

```{r}
phl$total_gross_pay[phl$base_salary != 0 & phl$total_gross_pay == 0] <- NA
```

```{r amount_summary}
summary(phl$total_gross_pay)
mean(phl$total_gross_pay <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(phl[c(which.max(phl$base_salary), which.min(phl$base_salary)), ])
```

```{r hist_amount, echo=FALSE}
phl %>%
  filter(base_salary > 120) %>% 
  ggplot(aes(base_salary)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Philadelphia Salary Amount Distribution",
    subtitle = "Excluding yearly salaries below $150",
    caption = "Source: OnePhilly",
    x = "Amount",
    y = "Count"
  )
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(phl, count_na)
```

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
key_vars <- c("last_name", "title", "total_gross_pay")
phl <- flag_na(phl, all_of(key_vars))
sum(phl$na_flag)
```

```{r na_view}
phl %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars), salary_type)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
phl <- flag_dupes(phl, everything())
sum(phl$dupe_flag)
```

There are no duplicates in the newer version of the data.

### Categorical

```{r distinct_count}
col_stats(phl, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(phl, department_name)
explore_plot(phl, salary_type)
explore_plot(phl, employee_category)
explore_plot(phl, compulsory_union_code)
```

### Dates

```{r date_range}
min(phl$calendar_year)
max(phl$calendar_year)
```

```{r bar_year, echo=FALSE}
phl %>% 
  count(calendar_year, quarter) %>% 
  ggplot(aes(x = quarter, y = n)) +
  geom_col(aes(fill = quarter)) + 
  scale_fill_distiller(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Philadelphia Employees by Year and Quarter",
    caption = "Source: OnePhilly",
    x = "Quarter",
    y = "Count"
  ) + 
  facet_wrap(~calendar_year, nrow = 1)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(phl, 50))
```

1. There are `r comma(nrow(phl))` records in the database.
1. There are `r comma(sum(phl$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(phl$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("pa", "philadelphia", "salary", "data", "clean"))
clean_path <- path(clean_dir, "philadelphia_salary_2019-2021.csv")
write_csv(phl, clean_path, na = "")
(clean_size <- file_size(clean_path))
```

```{r echo=FALSE}
rm(phl)
```

```{r}
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
