---
title: "Minnesota Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("mn", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit state Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"
[03]: https://cfb.mn.gov/

# Registration

## Data

Data is obtained from the [Campaign Finance Public Disclosure Board (CFPB)][03]

> The Board's four major programs are campaign finance registration and disclosure, public subsidy
administration, **lobbyist registration and disclosure**, and economic interest disclosure by
public officials. The Board has six members, appointed by the Governor on a bi-partisan basis for
staggered four-year terms. The appointments must be confirmed by a three-fifths vote of the members
of each house of the legislature.

The CFPB provides an index of lobbyists in alphabetical order as well as a similar list for 
associations who have hired lobbyists. Each letter of the alphabet has a corresponding page listing
all the registered lobbyists (or associations) that start with that letter. Each name is linked to
their own page, on which the address is listed alongside a table of represented associations (or
lobbyists). 

I have written a separate script that goes through the complicated process of scraping each
individual lobbyist page, connecting that lobbyist to all their represented associations, and
combining these relationships into a single data frame. We can run that script now if needed.

```{r source}
raw_dir <- dir_create(here("mn", "lobby", "data", "raw"))
scrape_file <- dir_ls(raw_dir, glob = "*lob_scrape.csv$")
scrape_script <- here("mn", "lobby", "code", "scrape_mn_lobby.R")
if (!file_exists(path = scrape_file)) {
  source(file = scrape_script, local = TRUE)
}
```

## Import

```{r read_csv}
mnlr <- read_csv(
  file = scrape_file,
  col_types = cols(
    .default = col_character(),
    start = col_date(),
    end = col_date(),
    designated = col_logical()
  )
)
```

## Explore

```{r glimpse}
head(mnlr)
tail(mnlr)
glimpse(sample_frac(mnlr))
```

Very few records are missing the kind of values needed to identify the lobbyist relationships.

```{r count_na}
col_stats(mnlr, count_na)
```

## Wrangle

To improve the consistency and searchability of the database, we can perform some simple and 
confident text normalization.

### Phone

```{r phone_norm}
mnlr <- mutate(
  .data = mnlr,
  lb_tel_norm = normal_phone(lb_tel)
)
```

```{r phone_view, echo=FALSE}
mnlr %>% 
  select(contains("tel")) %>% 
  distinct() %>% 
  sample_frac()
```

### Address

```{r address_norm}
mnlr <- mutate_at(
  .tbl = mnlr,
  .vars = vars(ends_with("street")),
  .funs = list(norm = normal_address),
  abbs = usps_street
)
```

```{r address_view, echo=FALSE}
mnlr %>% 
  select(contains("street")) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

```{r zip_norm}
mnlr <- mutate_at(
  .tbl = mnlr,
  .vars = vars(ends_with("zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_view, echo=FALSE}
mnlr %>% 
  select(contains("zip")) %>% 
  distinct() %>% 
  sample_frac()
```

```{r zip_progress, echo=FALSE}
progress_table(
  mnlr$lb_zip,
  mnlr$lb_zip_norm,
  mnlr$a_zip,
  mnlr$a_zip_norm,
  compare = valid_zip
)
```

### State

```{r state_check}
prop_in(mnlr$lb_state, valid_state)
mnlr$lb_state <- na_out(mnlr$lb_state, valid_state)
prop_in(mnlr$a_state, valid_state)
```

### City

```{r city_normal}
mnlr <- mutate_at(
  .tbl = mnlr,
  .vars = vars(ends_with("city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  state = c("MN", "DC", "MINNESOTA"),
  na = invalid_city,
  na_rep = TRUE
)
```

```{r city_swap}
mnlr <- mnlr %>%
  left_join(
    y = zipcodes,
    by = c(
      "lb_state" = "state",
      "lb_zip_norm" = "zip"
    )
  ) %>% 
  rename(lb_city_match = city) %>% 
  mutate(
    lb_match_abb = is_abbrev(lb_city_norm, lb_city_match),
    lb_match_dist = str_dist(lb_city_norm, lb_city_match),
    lb_city_swap = if_else(
      condition = lb_match_abb | lb_match_dist == 1,
      true = lb_city_match,
      false = lb_city_norm
    )
  ) %>% 
  select(
    -lb_match_abb,
    -lb_match_dist,
    -lb_city_match
  )

# repeat for associations
mnlr <- mnlr %>% 
  left_join(
    y = zipcodes,
    by = c(
      "a_state" = "state",
      "a_zip_norm" = "zip"
    )
  ) %>% 
  rename(a_city_match = city) %>% 
  mutate(
    a_match_abb = is_abbrev(a_city_norm, a_city_match),
    a_match_dist = str_dist(a_city_norm, a_city_match),
    a_city_swap = if_else(
      condition = a_match_abb | a_match_dist == 1,
      true = a_city_match,
      false = a_city_norm
    )
  ) %>% 
  select(
    -a_match_abb,
    -a_match_dist,
    -a_city_match
  )
```

```{r city_progress}
progress <- 
  progress_table(
    str_to_upper(mnlr$lb_city),
    mnlr$lb_city_norm,
    mnlr$lb_city_swap,
    str_to_upper(mnlr$a_city),
    mnlr$a_city_norm,
    mnlr$a_city_swap,
    compare = valid_city
  ) %>%
  separate(
    col = stage, 
    into = c("type", "stage"), 
    sep = "_", 
    extra = "merge"
  ) %>% 
  mutate(stage = str_remove(as_factor(stage), "\\)$")) %>% 
  group_by(stage) 
```

```{r city_kable}
kable(progress, digits = 3)
```

```{r plot_prop, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_col(aes(fill = type), position = "dodge") +
  scale_fill_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0.99) +
  facet_wrap(~type) +
  coord_cartesian(ylim = c(0.5, 1.00)) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Pennsylvania City Normalization Progress",
    x = "Stage",
    y = "Percent Valid"
  )
```

```{r plot_distinct, echo=FALSE}
progress %>% 
  select(
    type,
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Pennsylvania City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  ) +
  facet_wrap(~type)
```

## Export

```{r write_csv}
proc_dir <- dir_create(here("mn", "lobby", "data", "processed"))
write_csv(
  x = mnlr,
  path = glue("{proc_dir}/mn_lobbyists.csv"),
  na = ""
)
```

# Expenditures

We can also obtain a report of lobbying expenditures made by the principal associations.

```{r read_expends}
exp_url <- "https://cfb.mn.gov/reports-and-data/self-help/data-downloads/lobbying/?download=102614850"
mnle <- read_csv(
  file = exp_url,
  col_types = cols(
    Principal = col_character(),
    `Entity ID` = col_character(),
    `Report Year` = col_double(),
    `General lobbying amount` = col_double(),
    `PUC lobbying amount` = col_double(),
    `Total spent` = col_double()
  )
)

mnle <- mnle %>% 
  rename(
    a_name  = `Principal`,
    a_id  = `Entity ID`,
    year  = `Report Year`,
    gen_amount  = `General lobbying amount`,
    lob_amount  = `PUC lobbying amount`,
    total_amount  = `Total spent`
  )
```

```{r plot_lob_amount}
mnle %>% 
  filter(total_amount > 1) %>% 
  ggplot(aes(x = total_amount)) +
  geom_histogram() +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "State Data Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: State Agency Name",
    x = "Amount",
    y = "Count"
  )
```

Since this database will be uploaded separately from the lobbyist registration containing the
phone number and addresses of principal clients, we will have to add these columns
so that the expenditure records will show up when this information is searched.

```{r expends_join}
a_info <- mnlr %>% 
  select(starts_with("a_")) %>% 
  select(a_id, a_website, a_contact, a_street_norm, a_city_norm, a_state, a_zip_norm)
mnle <- left_join(mnle, a_info, by = "a_id")
```

```{r}
col_stats(mnle, count_na)
```

```{r}
write_csv(
  x = mnle,
  path = glue("{proc_dir}/mn_lobby_expend.csv"),
  na = ""
)
```

