---
title: "Wyoming Campaign Expenditures Data Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  gluedown, # printing markdown
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  zipcode, # clean & database
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  rvest # scrape html
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

The data comes from the Wyoming Secretary of State. [Link to the data download][03].

In our last update, we downloaded 11 years worth of data (2008-2018) in 8 columns, Filer Type,	Filer Name,	Payee,	Purpose,	Date,	City,State & ZIP, Filing Status	and Amount. This time we'll download the data for the 2020 election cycle.

[03]: https://www.wycampaignfinance.gov/WYCFWebApplication/GSF_SystemConfiguration/SearchExpenditures.aspx "source"

### About

> Wyoming's Campaign Finance Information System (WYCFIS) exists to provide a mechanism for online filing of campaign finance information and to provide full disclosure to the public. This website contains detailed financial records and related information that candidates, committees, organizations and parties are required by law to disclose.
Wyoming requires all statewide candidates, candidate committees, political action committees, organizations and political parties to file electronically online using this system. 


## Import

### Download

Download raw, **immutable** data file. Go to [the download site](https://www.wycampaignfinance.gov/WYCFWebApplication/GSF_SystemConfiguration/SearchExpenditures.aspx), leave the fields blank, and click the "All" tab and hit "Search". After the table is populated, click "Export"

```{r raw_dir}
# create a directory for the raw data
raw_dir <- here("wy", "expends", "data", "raw")
dir_create(raw_dir)
```


### Read

```{r read_many, echo=FALSE}
wy <- 
  dir_ls(path = raw_dir) %>% 
  map(
    read_delim,
    delim = ",",
    escape_double = FALSE,
    escape_backslash = FALSE,
    col_types = cols(
      .default = col_character(),
      Date = col_date("%m/%d/%Y"),
      Amount = col_double()
    )  
    ) %>% 
  bind_rows() %>% 
  distinct() %>% 
  select(-starts_with("X")) %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper) %>% 
  map_if(is_character, str_replace_all, "\\\"", "\'") %>% 
  as_tibble()
```

## Explore

There are `nrow(wy)` records of `length(wy)` variables in the full database.

```{r glimpse}
head(wy)
tail(wy)
glimpse(wy)
```

### Distinct

The variables range in their degree of distinctness.


```{r n_distinct}
wy %>% col_stats(n_distinct)
```

We can explore the distribution of the least distinct values with `ggplot2::geom_bar()`.

```{r plot_bar, echo=FALSE}
ggplot(data = wy) +
  geom_bar(aes(filer_type))
```

Or, filter the data and explore the most frequent discrete data.

```{r plot_bar2, echo=FALSE}
wy %>% 
  group_by(payee)  %>% 
  summarize(total_spent = sum(amount)) %>% 
  head(10) %>% 
  ggplot(aes(x=reorder(payee,total_spent), y=total_spent)) + 
  geom_col(fill = dark2["purple"]) +
    labs(title = "Wyoming Campaign Expenditures",
       caption = "Source: Wyoming Secretary of State",
       x = "payee",
       y = "total expenditures") +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_discrete(labels = wrap_format(10)) +
  coord_flip() +
  theme_minimal()
```

### Missing

We will flag the entries with an empty `city_state_zip` column.

```{r count_na}
wy %>% col_stats(count_na)
```

We will flag any records with missing values in the key variables used to identify an expenditure.
There are `r sum(wy$na_flag)` columns in city_state_zip that are NAs.
```{r na_flag}
wy <- wy %>% flag_na(city_state_zip)
```

### Duplicates
There are no duplicates.
```{r get_dupes, collapse=TRUE}
wy_dupes <- flag_dupes(wy)
```

### Ranges

#### Amounts

```{r}
summary(wy$amount)
```

See how the campaign expenditures were distributed

```{r}
wy %>% 
  ggplot(aes(x = amount)) + 
  geom_histogram(fill = dark2[1]) +
  scale_x_continuous(
    trans = "log10", labels = dollar) +
  labs(title = "Wyoming Campaign Expenditures ")
```

Distribution of expenses by filer
```{r box_plot_by_type, echo=FALSE}
wy %>% 
  ggplot(
    mapping = aes(
      x = filer_type, 
      y = amount
    )
  ) +
  geom_boxplot(
    mapping  = aes(fill = filer_type), 
    varwidth = TRUE,
    outlier.alpha = 0.01
  ) +
  scale_fill_brewer(
    type    = "qual",
    palette = "Set1",
    guide   = FALSE
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = dollar
  ) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
  labs(
    title = "Wyoming Expenditure Amount Ranges",
    x     = "Expenditure Type",
    y     = "Amount",
    caption = "Source: Wyoming Secretary of State"
  )
```

### Dates
The dates seem to be reasonable, with records dating back to `r summary(wy$date)[1]` till `r summary(wy$date)[-1]` 
```{r}
summary(wy$date)
sum(wy$date > today())
```

### Year

Add a `year` variable from `date` after `col_date()` using `lubridate::year()`.

```{r add_year}
wy <- wy %>% mutate(year = year(date))
```

```{r year_count_bar, echo=FALSE}
wy %>% 
  mutate(on_year = is_even(year)) %>% 
  count(on_year, year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill=on_year)) +
  scale_x_continuous(position = "bottom", breaks = c(2019,2020)) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  labs(
    title = "Wyoming Expenditure Counts per Year",
    caption = "Source: Wyoming Secretary of State",
    x = "Year",
    y = "Count"
  )
  
```

```{r amount_year_bar, echo=FALSE}
wy %>% 
  mutate(on_year = is_even(year)) %>%
  group_by(year, on_year) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_col(aes(fill = on_year)) +
  scale_x_continuous(position = "bottom", breaks = c(2019,2020)) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2",
    guide = FALSE
  ) +
  scale_y_continuous(labels = scales::dollar) +
  labs(
    title = "Wyoming Expenditure Mean Amount by Year",
    caption = "Source: Wyoming Secretary of State",
    x = "Year",
    y = "Amount"
  ) 
```

```{r amount_month_line}
wy %>% 
  mutate(month = month(date)) %>% 
  mutate(on_year = is_even(year)) %>%
  group_by(on_year, month) %>% 
  summarize(mean = mean(amount)) %>% 
  ggplot(aes(month, mean)) +
  geom_line(aes(color = on_year), size = 2) +
  scale_y_continuous(labels = dollar) +
  scale_x_continuous(labels = month.abb, breaks = 1:12) +
  scale_color_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wyoming Mean Expenditure Amount by Month",
    caption = "Source: Wyoming Secretary of State",
    color = "Election Year",
    x = "Month",
    y = "Amount"
  )
```
## Wrangle
### Indexing
```{r}
wy <- tibble::rowid_to_column(wy, "id")
```

The lengths of city_state_zip column differ, and regular expressions can be used to separate the components.

The original data the city, state, and ZIP all in one column. The following code separates them. 

### Zipcode
First, we'll extract any numbers whose lengths range from 1 to 5 to `zip`, whose proportion of valid zip is pretty high and doesn't need further normalization. 

```{r}
wy <- wy %>% 
  mutate(
    zip = city_state_zip %>% 
      str_extract("\\d{2,5}") %>% 
      normal_zip(na_rep = TRUE))
sample(wy$zip, 10)
prop_in(wy$zip, valid_zip, na.rm = T)
```


### State

In this regex, state is considered to consist of two upper-case letters following a space, or two upper-case letters with a trailing space at the end.

```{r separate state}
wy <- wy %>% 
  mutate( state =
            trimws(str_extract(wy$city_state_zip, "\\s([A-Z]{2})\\s|^([A-Z]{2})\\s$")))
count_na(wy$state)
prop_in(wy$state, valid_state, na.rm = T)
```

The states are mostly valid and don't need to be cleaned. 

### City
First, we can get a list of incorporated cities and towns in Wyoming. The Wyoming State Archives provided the list in a web table. We use the `rvest` package to scrape the names of Wyoming cities and towns. http://wyoarchives.state.wy.us/index.php/incorporated-cities.

```{r}
wyoming_cities_page <- read_html("http://wyoarchives.state.wy.us/index.php/incorporated-cities")

wy_city <- wyoming_cities_page %>%  html_nodes("tr") %>% 
  html_text()

wy_city <- str_match(wy_city[2:100],"(^\\D{2,})\\r")[,2]
wy_city <- toupper(wy_city[!is.na(wy_city)])
```

```{r valid_city}
valid_city <- unique(c(wy_city,valid_city))
```

Cleaning city values is the most complicated. This process involves four steps:

1. Prepare raw city values by removing invalid data and reducing inconsistencies
1. Match prepared city values with the _actual_ city name of that record's ZIP code
1. swap prepared city values with the ZIP code match _if_ only 1 edit is needed
1. Refine swapped city values with key collision and n-gram fingerprints

#### Normalize
Find the cities before a comma first, if not, find the non-numeric string. 
``` {r two-step regex match}
wy <- wy %>% 
  mutate(
    city_raw = str_match(wy$city_state_zip,"(^\\D{3,}),")[,2]) 

wy <- wy %>% mutate(city_raw=ifelse(is.na(city_raw)==TRUE, 
               str_extract(city_state_zip, "[A-Z]{4,}"), paste(city_raw)))

wy$city_raw <- wy$city_raw %>% 
  str_replace("^ROCK$", "ROCK SPRING") 
```

```{r city_raw summary, collapse= TRUE}
count_na(wy$city_raw)
n_distinct(wy$city_raw)
prop_in(wy$city_raw, valid_city, na.rm = TRUE)
sum(unique(wy$city_raw) %out% valid_city)
```

`r sum(!is.na(wy$city_raw))` cities were found.
```{r prep_city}
wy <- wy %>% mutate(city_norm = normal_city(city_raw))
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r match_dist}
wy <- wy %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist) == 1,
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )

```

### Lookup
This process is manual lookup and correction
```{r }
wy$city_swap <- wy$city_swap %>% 
  str_replace("^CAS$", "CASPER") %>% 
  str_replace("^CA$", "CASPER") %>% 
  str_replace("^RS$","ROCK SPRINGS") %>% 
  str_replace("^AF$", "AFTON") %>% 
  str_replace("^M$", "MOUNTAIN VIEW") %>% 
  str_replace("^GR$", "GREEN RIVER") %>% 
  na_if("WY") %>% 
  str_replace(" WYOMING","") %>% 
  str_replace("^SLC$", "SALT LAKE CITY") %>% 
  str_replace("^COD$", "CODY") 

n_distinct(wy$city_swap)
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- wy %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state" = "state",
      "zip" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state, 
    zip, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
wy <- wy %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city, wy_city)
progress <- progress_table(
  str_to_upper(wy$city_raw),
  wy$city_norm,
  wy$city_swap,
  wy$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

```{r view_refine, echo=F, eval=FALSE}
# Get a table of cities that are not in the valid_city vector
wy_out <- wy %>% 
  count(state, city_swap, sort = TRUE) %>% 
  filter(city_swap %out% c(valid_city,wy_city)) %>% 
  drop_na()

# Fuzzy-matching city names with the names out of such list
prep_refine <- wy_out %>% 
  # Join on all names in the relevant state
  left_join(cities, by=c(state_clean = "state")) %>%
  # Calculate the distances, per original city name.
  group_by(city) %>%                
  mutate(dist = diag(adist(city_swap, city, partial=TRUE))) %>%
  # Append the agrepl result with the Levenshtein edit distance
  rowwise() %>%
  mutate(string_agrep = agrepl(city_swap, city, max.distance = 0.3)) %>%
  ungroup() %>%  
  # Only retain minimum distances
  group_by(city_swap) %>%   
  filter(dist == min(dist))

# Refine the entries where city_swap is six letter apart from a single match in cities (It's a rather safe switch, after examining the prep_refine table). Overcorrection can be manually changed later.
to_refine <- prep_refine %>% filter(n()==1) %>% filter(city_swap %in% prep_refine$city_swap[prep_refine$dist<6])
```

```{r join_refine, eval=FALSE, echo=FALSE}
wy_refined <- wy %>% left_join(to_refine, by = "city_swap") %>% 
  select(-n, -state_clean.y, -dist, - string_agrep) %>% 
   mutate(city_refine = if_else(
    condition = is.na(city) == TRUE,
    true = city_swap,
    false = city
  )) %>% select(-city)

```

Manually change the city_refine fields due to overcorrection/undercorrection.

```{r}
wy$city_refine <- wy$city_refine %>% 
  str_replace("^RIO VISTA$", "LAGO VISTA") %>% 
  str_replace("^OGEN$", "OGDEN") %>%
  str_replace("^ANNIPOLIS$", "ANNAPOLIS") %>% 
  str_replace("^LAR$", "LARAMIE") %>%
  str_replace("^LARA$", "LARAMIE") %>%
  str_replace("^CHE$", "CHEYENNE") %>%
  str_replace("^COLO SPGS$", "COLORADO SPRINGS") %>%
  str_replace("^WASHNGTON$", "WASHINGTON") %>% 
  str_replace("^WASHINGTON DC$", "WASHINGTON") %>% 
  str_replace("^ST.\\s", "SAINT " ) %>% 
  str_replace("^PINE$", "PINEDALE")
```

This process reduces the number of distinct city value by `r n_distinct(wy$city_raw) - n_distinct(wy$city_refine)`
```{r city_progress1, collapse=TRUE}
n_distinct(wy$city_raw)
n_distinct(wy$city_norm)
n_distinct(wy$city_swap)
n_distinct(wy$city_refine)
```


Each step of the cleaning process reduces the number of distinct city values.
There are `r sum(!is.na(wy$city_raw))` entries of cities identified in the original data matching the regex with `r n_distinct(wy$city_raw)` distinct values, after the swap and refine processes, there are `r sum(!is.na(wy$city_refine))` entries with `r n_distinct(wy$city_refine)` distinct values. 

## Conclude

1. There are `r nrow(wy)` records in the database
1. There are `r sum(wy$dupe_flag)` records with duplicate filer, recipient, date, _and_ amount
(flagged with `dupe_flag`)
1. The ranges for dates and amounts are reasonable
1. Consistency has been improved with `stringr` package and custom `normal_*()` functions.
1. The five-digit `zip_clean` variable has been created with `zipcode::clean.zipcode()`
1. The `year` variable has been created with `lubridate::year()`
1. There are `r sum(is.na(wy$filer_name))` records with missing `name` values and `r sum(is.na(wy$date))`
records with missing `date` values (both flagged with the `na_flag`)

## Export

```{r clean file}
wy <- wy %>% 
  rename(city_clean = city_refine) %>% 
  select(
    -city_raw,
    -city_norm,
    -city_swap,
    -id
  )
```

```{r write_clean}
clean_dir <- here("wy", "expends", "data", "processed")
clean_path <- glue("{clean_dir}/wy_expends_clean.csv")

dir_create(clean_dir)
wy %>% 
  write_csv(
    path = clean_path,
    na = ""
  )

file_size(clean_path)
file_encoding(clean_path)
```
## Upload

Using the `aws.s3` package, we can upload the file to the IRW server.

```{r upload,eval=FALSE}
s3_path <- path("csv", basename(clean_path))
put_object(
  file = clean_path,
  object = s3_path, 
  bucket = "publicaccountability",
  acl = "public-read",
  multipart = TRUE,
  show_progress = TRUE
)
as_fs_bytes(object_size(s3_path, "publicaccountability"))
```
## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(wy)),
  type = md_code(map_chr(wy, typeof)),
  def = c(
    "Type of filer",
    "Name of filer",
    "Payee name",
    "Expenditure purpose",
    "Expenditure date",
    "Expenditure city, state and zip",
    "Filing status",
    "Expenditure amount",
    "Flag for missing name, city or address",
    "Calendar year of expenditure date",
    "5-digit ZIP code",
    "2-letter state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write,echo=FALSE, eval=FALSE}
write_lines(
  x = c("# Wyoming Expenditure Data Dictionary\n", dict_md),
  path = here("wy", "exp", "wy_exp_dict.md"),
)
```


