---
title: "California Lobbyists"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here, collapse=TRUE}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the California Secretary of State's [Cal-Access database][03].

> Cal-Access provides financial information supplied by state candidates, donors, lobbyists, and
others. Simply start by selecting either Campaign Finance Activity, Lobbying Activity, or
Cal-Access Resources and we will guide you to the information.

The [Political Reform Division][04] (PRD) collects the data regarding expenditures made.

> The Political Reform Division administers provisions of California's Political Reform Act,
including... that the voters may be fully informed and the activities of lobbyists should be
regulated and their finances disclosed..."

> * Receive campaign disclosure statements (itemizing contributions received and expenditures made)
filed by individuals and committees raising or spending campaign funds to support or oppose state
candidates or ballot measures. (Local campaign committees file their itemized disclosure statements
with local filing officers).

### About

All California campaign finance data can be downloaded in a single file. As described on PRD
website:

> ### Raw Data for Campaign Finance and Lobbying Activity  
> In addition to presenting California campaign finance and lobbying activity on the user-friendly
[CAL-ACCESS website][05], the Secretary of State provides the raw data to allow people with 
technical expertise to create their own databases.
> 
> The raw data is presented in tab-delimited text files from corresponding tables in the CAL-ACCESS
database. Users can uncompress and extract the data with standard software such as PKZIP, WinZip,
or MacZip. The Secretary of State does not provide technical support to users who want to download
and analyze CAL-ACCESS data in this advanced way. However, the Secretary of State offers guides to
the CAL-ACCESS data structure and fields.

> ### Download Files  
* [Guides to CAL-ACCESS data structure and fields (ZIP)][06]
* [CAL-ACCESS raw data (ZIP)][07]
> 
> All CAL-ACCESS users should keep in mind that campaign finance and lobbying activity information
changes often. The raw data extracts are updated once a day. Campaign finance and lobbying activity
filings can also be obtained in hard copy by contacting the Secretary of State's Political Reform
Division.

[03]: https://www.sos.ca.gov/
[04]: https://www.sos.ca.gov/campaign-lobbying/about-political-reform-division/
[05]: http://cal-access.sos.ca.gov/
[06]: https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip
[07]: https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip

### Variables

Using the data key provided by Cal-Access, we can find the expenditure file and understand it's 
contents.

```{r check_key}
key_url <- "https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip"
url_file_size(key_url, format = TRUE)
```

If they ZIP file containing the documentation files has not yet been downloaded, we can do so now.

```{r download_key}
raw_dir <- dir_create(here("ca", "lobby", "data", "raw"))
key_file <- str_c(raw_dir, basename(key_url), sep = "/")
if (!this_file_new(key_file)) {
  download.file(
    url = key_url,
    destfile = key_file
  )
}
```

Before we unzip the file, we can view it's contents.

```{r list_key}
key_content <- key_file %>% 
  unzip(list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names)
```

```{r unzip_key}
key_dir <- str_c(raw_dir, "CalAccess-Documentation", sep = "/")
if (!dir_exists(key_dir)) {
  unzip(
    zipfile = key_file,
    exdir = raw_dir
  )
}
```

## Import

### Download

```{r zip_file}
zip_url <- "https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip"
zip_file <- str_c(raw_dir, basename(zip_url), sep = "/")
```

The ZIP file is extremelly large, and will take quite some time 

```{r file_size}
url_file_size(zip_url, format = TRUE)
if (requireNamespace("speedtest", quietly = TRUE)) {
  speedtest::spd_test()
}
```

If the most recent version of the file has not yet been downloaded, we can do so now.

```{r download_raw}
if (!this_file_new(zip_file)) {
  download.file(
    url = zip_url,
    destfile = zip_file
  )
}
```

We don't need to unzip every file, only those pertaining to lobbying.

```{r list_zip}
zip_content <- unzip(zip_file, list = TRUE) 
zip_lobby <- zip_content$Name[str_which(zip_content$Name, "LOBBY")]
cal_dir <- str_c(raw_dir, unique(dirname(zip_lobby)), sep = "/")
```

Then, if those files have not yet been unzipped, we can do so now.

```{r unzip_zip}
if (!dir_exists(cal_dir)) {
  unzip(
    zipfile = zip_file,
    files = zip_lobby,
    exdir = raw_dir
  )
}
```

### Read

Much of the data is split into multiple files. We will list all the files of a similar name and
read them into a single data frame with `purrr::map_dfr()` and `readr::read_tsv()`.

```{r warn_vroom}
if (packageVersion("vroom") < "1.0.2.9000") {
  stop("vroom version of at least 1.0.2.9000 is needed")
}
```

> Lobbyist contribution disclosure table. Temporary table used to generate disclosure table
(Lobbyist Contributions 3).

```{r read_conts}
lob_conts <- map_dfr(
  .x = dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_CONTRIBUTIONS\\d_CD.TSV$"
  ),
  .f = read_tsv,
  col_types = cols(
    FILER_ID = col_character(),
    FILING_PERIOD_START_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
    FILING_PERIOD_END_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
    CONTRIBUTION_DT = col_character(),
    RECIPIENT_NAME = col_character(),
    RECIPIENT_ID = col_character(),
    AMOUNT = col_double()
  )
) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

Every yearly relationship between lobbyists and their principal clients.

```{r read_emp_lob}
cal_emp_lob <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMP_LOBBYIST\\d_CD.TSV"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      SESSION_ID = col_double()
    )
  ) %>% 
  clean_names("snake")
```

Quarterly and annual employee compensation amounts.

```{r read_emp_total}
cal_emp_total <- map_df(
  .x = dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER\\d_CD.TSV$"
  ),
  .f = read_tsv,
  col_types = cols(
    .default = col_character(),
    SESSION_ID = col_double(),
    CURRENT_QTR_AMT = col_double(),
    SESSION_TOTAL_AMT = col_double(),
    SESSION_YR_1 = col_double(),
    SESSION_YR_2 = col_double(),
    YR_1_YTD_AMT = col_double(),
    YR_2_YTD_AMT = col_double(),
    QTR_1 = col_double(),
    QTR_2 = col_double(),
    QTR_3 = col_double(),
    QTR_4 = col_double(),
    QTR_5 = col_double(),
    QTR_6 = col_double(),
    QTR_7 = col_double(),
    QTR_8 = col_double()
  )
) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

```{r}
cal_emp_total <- cal_emp_total %>% 
  select(
    session_id,
    employer_id,
    interest_cd,
    interest_name,
    session_total_amt,
  )
```

```{r read_emp_firms}
cal_emp_firms <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER_FIRMS\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      SESSION_ID = col_double()
    )
  ) %>% 
  clean_names("snake") %>% 
  remove_empty("cols")
```

```{r read_emp_hist}
# empty file
cal_lob_hist <- 
dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_EMPLOYER_HISTORY_CD.TSV$"
  ) %>% 
  file_size()
```

```{r read_firms}
cal_firm_totals <- 
  dir_ls(
    path = cal_dir, 
    type = "file", 
    glob = "*LOBBYIST_FIRM\\d_CD.TSV$"
  ) %>% 
  vroom(
    col_types = cols(
      .default = col_character(),
      CURRENT_QTR_AMT = col_double(),
      SESSION_TOTAL_AMT = col_double(),
      SESSION_YR_1 = col_double(),
      SESSION_YR_2 = col_double(),
      YR_1_YTD_AMT = col_double(),
      YR_2_YTD_AMT = col_double(),
      QTR_1 = col_double(),
      QTR_2 = col_double(),
      QTR_3 = col_double(),
      QTR_4 = col_double(),
      QTR_5 = col_double(),
      QTR_6 = col_double(),
      QTR_7 = col_double(),
      QTR_8 = col_double()
    )
  ) %>% 
  clean_names("snake")
```

```{r}
cal_firm_totals <- cal_firm_totals %>% 
  select(
    session_id,
    firm_id,
    firm_name,
    session_total_amt,
  )
```

```{r}
lob <- read_delim(
  file = str_c(cal_dir, "CVR_LOBBY_DISCLOSURE_CD.TSV", sep = "/"),
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", "NA", "none", "NONE", "None"),
  col_types = cols(
    .default = col_character(),
    AMEND_ID = col_double(),
    RPT_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    FROM_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    THRU_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    CUM_BEG_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
    CUM_BEG_DT = col_date("%m/%d/%Y %H:%M:%S %p"),
    SIG_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
  )
)
```

## Join

```{r}
calr <- cal_emp_lob
```

## Explore

```{r}
head(cal)
tail(cal)
glimpse(sample_frac(cal))
```

