---
title: "Louisiana Salary"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("la", "salary", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel files
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Salary data for the state of Louisiana was obtained via records request pursuant
to the Louisiana Public Records Act from the Department of Administration. In
the email providing the data, more information was given:

> The information provided is for employees of executive branch agencies paid
through the LaGov HCM payroll system.  This does not include higher education or
agencies in the legislative and judicial branches of government.

> Gross wages and paid overtime is provided for fiscal year 2019; all other
information is reported as of the last day of fiscal year 2019 (06/30/2019).

> The Overtime tab includes every employee who received overtime pay at any
point during fiscal year 2019.  Overtime reported is overtime that is paid as
worked and it is included in the gross wages.

> The agency hire date is optional for entry in the LaGov HCM payroll system and
the use may vary by agency.  As such, this date may not exist for some
employees.  We cannot confirm the accuracy of this date (if populated) and we
cannot confirm it reflects the employee’s first date of state service.

## Read

Data was provided as an Excel file with multiple sheets containing the various
tables needed to identify all salary information. We can use 
`readxl::read_excel()` to read each individual sheet as a separate data frame.

```{r raw_dir}
raw_dir <- dir_create(here("la", "salary", "data", "raw"))
raw_path <- dir_ls(raw_dir)
raw_sheets <- excel_sheets(raw_path)
```

We can read each sheet as a data frame in a list.

```{r raw_list}
la_sheets <- list(
  emps = read_excel(raw_path, raw_sheets[1]),
  wages = read_excel(raw_path, raw_sheets[2]),
  overtime = read_excel(raw_path, raw_sheets[3]),
  hire_date = read_excel(raw_path, raw_sheets[4]),
  reference = read_excel(raw_path, raw_sheets[5], skip = 3)
)
```

```{r echo=FALSE}
la_sheets <- map(la_sheets, clean_names)
```

```{r list_glimpse}
glimpse(la_sheets)
```

Then, we can combine each sheet into a single data frame.

```{r list_rename, echo=FALSE}
la_sheets$emps <- la_sheets$emps %>% 
  mutate(across(pa, str_pad, 4, "left", "0")) %>% 
  rename(hours_worked = work_hours_per_pay_period)
la_sheets$wages <- la_sheets$wages %>% 
  rename(pa = pers_area) %>% 
  mutate(across(pa, str_pad, 4, "left", "0"))
la_sheets$overtime <- la_sheets$overtime %>% 
  rename(pa = pers_area) %>% 
  mutate(across(pa, str_pad, 4, "left", "0"))
la_sheets$hire_date <- la_sheets$hire_date %>% 
  mutate(across(pa, str_pad, 4, "left", "0")) %>% 
  rename(hire_date = agency_hire_date)
la_sheets$reference <- la_sheets$reference %>% 
  rename(pa = pers_area, area = personnel_area_name)
```

```{r list_join}
las <- la_sheets$emps %>% 
  left_join(la_sheets$wages) %>% 
  left_join(la_sheets$overtime) %>% 
  left_join(la_sheets$hire_date) %>% 
  left_join(la_sheets$reference) %>% 
  arrange(last_name, first_name) %>% 
  mutate(across(hire_date, as_date)) %>% 
  select(-pa)
```

The `area` variable taken from the `pa` variable and `reference` sheet contains
a single string with an agency abbreviation and agency or office. We can split
the abbreviation into a separate variable.

```{r raw_split}
las <- separate(
  data = las,
  col = area,
  into = c("agency_abb", "area"),
  sep = "\\s?-\\s?",
  remove = TRUE,
  fill = "left",
  extra = "merge"
)
```

## Explore

```{r glimpse}
glimpse(las)
tail(las)
```

### Missing

Most records are not missing any values and need not be flagged.

```{r na_count}
col_stats(las, count_na)
```

### Duplicates

There are only a handful of records duplicated across every variable.

```{r dupe_flag}
las <- flag_dupes(las, everything())
sum(las$dupe_flag)
```

```{r dupe_view}
las %>% 
  filter(dupe_flag) %>% 
  select(hire_date, last_name, gross_wages, area)
```

### Categorical

```{r distinct_count}
col_stats(las, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(las, employment_type)
explore_plot(las, agency_abb)
explore_plot(las, area) + scale_x_truncate()
```

### Amounts

There aren't any gross wages values at or under zero, as you'd expect.

```{r amount_summary}
summary(las$gross_wages)
mean(las$gross_wages <= 0, na.rm = TRUE)
```

Here is the highest paid public employee:

```{r amount_max}
glimpse(las[which.max(las$gross_wages), ])
```

The distribution of public employee salaries is reasonable.

```{r hist_amount, echo=FALSE}
las %>%
  filter(gross_wages < 2e5) %>% 
  ggplot(aes(gross_wages)) +
  geom_histogram(fill = dark2["purple"], bins = 30) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = seq(0, 2e5, by = 2.5e4),
    labels = dollar,
  ) +
  labs(
    title = "Louisiana Salary Amount Distribution",
    caption = "Source: LA DOA",
    x = "Gross Wages",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
las <- mutate(las, hire_year = year(hire_date))
```

```{r date_range}
min(las$hire_date, na.rm = TRUE)
max(las$hire_date, na.rm = TRUE)
sum(las$hire_date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
las %>% 
  filter(hire_year < 2020, hire_year > 1980) %>% 
  count(hire_year) %>% 
  mutate(even = is_even(hire_year)) %>% 
  ggplot(aes(x = hire_year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  labs(
    title = "Louisiana Employees Hired by Year",
    caption = "Source: LA DOA",
    x = "Year Hired",
    y = "Count"
  )
```

## Wrangle

There are no geographic variables to wrangle. We can manually add the state for
all spending agencies.

```{r state_add}
las <- mutate(las, agency_state = "LA", .after = agency_abb)
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(las, 20))
```

1. There are `r comma(nrow(las))` records in the database.
1. There are `r comma(sum(las$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(las$na_flag))` records missing key variables.
1. There are no geographic variables to normalize, `state` was added manually.
1. The 4-digit `hire_year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("la", "salary", "data", "clean"))
clean_path <- path(clean_dir, "la_salary_clean.csv")
write_csv(las, clean_path, na = "")
file_size(clean_path)
mutate(file_encoding(clean_path), across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(las)),
  type = md_code(map_chr(las, typeof)),
  def = c(
    "Unique personnel number",
    "Employee last name",
    "Emplyee first name",
    "Employment type (full, part)",
    "Full job title",
    "Annual salary before overtime or hours",
    "Estimated hours worked per pay period",
    "Reported wage type for `gross_wage`",
    "Gross wages paid, including overtime",
    "Overtime year-to-date",
    "Date hired at agency",
    "Personnel area agency",
    "Agency state, single value manually added",
    "Personnel area name",
    "Flag indicating duplicate record",
    "Calendar year hired at agency"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# Louisiana Salary Data Dictionary\n", dict_md),
  path = here("la", "salary", "la_salary_dict.md"),
)
```
