---
title: "Utah Lobbying Registration Data Diary"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

## Project


The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.


Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:


1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved


## Objectives


This document describes the process used to complete the following objectives:


1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction


## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  rvest, # read html tables
  httr, # interact with http requests
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  vroom, #read deliminated files
  readxl #read excel files
)
```

```{r fix_fun, echo=FALSE, collapse = TRUE}
# fix conflict
here <- here::here
```
This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.


The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Download
Set the download directory first.
```{r create raw_dir}
# create a directory for the raw data
raw_dir <- dir_create(here("ut", "lobby", "data", "raw","exp"))
```
According to [Utah Code 36-11 & Utah Code 36-11a] [03], 

Definition of a lobbyist:
> Generally, you are a lobbyist if you get paid to communicate with a public official, local official, or education
official for the purpose of influencing legislative, executive, local, or education action.

There are two types of lobbyists, a) state lobbyist b)local and education lobbyist.
> You are a state lobbyist if you lobby state legislators, elected state executive branch officials (such as the
governor), and non-elected officials within the state executive or legislative branch with certain decisionmaking powers.
> You are a local and education lobbyist if you lobby:  
 Elected members in local governments and non-elected officials within local governments that have
certain decision-making powers.
     Education officials, including elected members of the State Board of Education, State Charter School
      Board, local school boards, charter school governing boards, and non-elected officials within these
organizations that have certain decision-making powers.
If you lobby officials in both categories, register as a state lobbyist.

Definition of expenditure:

> An expenditure is something given to benefit an official. Common expenditures include purchasing meals or paying
 for an official’s travel to or lodging at a meeting. 

Reporting Requirements:
> A lobbyist is not required to file a quarterly financial report (Q1, Q2, Q3) if he or she
has not made an expenditure during that reporting period. All lobbyists – state, local, and education – are still
required to file the Quarter 4 (Year End) Report by January 10 of each year.

> 

https://elections.utah.gov/Media/Default/Lobbyist/2019%20Lobbyist/2019%20State%20Lobbyists%20Flowchart.pdf

[03]: https://elections.utah.gov/Media/Default/Lobbyist/2019%20Lobbyist/Lobbyist%20Frequently%20Asked%20Questions%202019%20(updated%20after%20session).pdf

This Rmd file documents the wrangling process of UT expedenditure data only, whereas the registration data is wrangled in a separate data diary.


 [Utah Lieutenant Governor's Office](https://lobbyist.utah.gov/Search/AdvancedSearch) makes available quarterly reports by lobbysits. We will see that only data from 2013 to 2019 is not empty.

## Reading
We can read the csv file here.
```{r read csv}
ut_lob_exp <- dir_ls(raw_dir, glob= "*.csv") %>% 
  map_dfr(read_csv, .id = "file") %>% clean_names() %>% mutate_if(is.character, str_to_upper) %>% 
  mutate(year = str_extract(file, "\\d{4}"))

col_stats(ut_lob_exp, count_na)
```

We can see that the geographical columns are overwhelmingly empty, and we will join such information from the registration data.


## Join
First, read the processed reg file.
```{r read reg}
reg_dir <- here("ut", "lobby", "data", "processed","reg")
ut_lob_reg <- dir_ls(reg_dir) %>% read_csv()
```
Also note that since the expenditure data doesn't contain the clients' information, and the address in the registration data's ID fields are likely for companies, since it differs within lobbyist from principal to principal.
```{r prep join}
ut_lob_reg <- ut_lob_reg %>% 
  unite(col = lobb,
    c(last_name, first_name),
    sep = ", ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  filter(!dupe_flag) %>% 
  select(zip5, city_clean)
```

```{r join}
ut_lob_exp_clean <- ut_lob_exp %>% 
  select(-file) %>% 
  left_join(ut_lob_reg, by = c(
    ""
  ))
```



## Export

```{r write clean}
clean_dir <- here("ut", "lobby", "data", "processed","exp")
dir_create(clean_dir)
ut_lob_exp %>% 
  select(-file) %>% 
  mutate_if(is.character, str_to_upper) %>% 
  write_csv(
    path = glue("{clean_dir}/ut_lobby_exp_clean.csv"),
    na = ""
  )
```


