---
title: "Delaware Contributions"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("de", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  batman, # parse na and lgl
  aws.s3, # upload to aws s3
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

This data is obtained from the [Campaign Finance Reporting System (CFRS)][cfrs].

On the [Contributions search page][04], the CFRS describes the data that can be
searched.

> An expenditure is defined as a gift, transfer, disbursement, debt incurred, or
promise of money or valuable thing by or on behalf of a political committee to
promote or assist in promoting the success or defeat of a candidate, political
party, or question at  an election. Expenditures must be election related; that
is, they must enhance the candidate's election chances, such that they would not
have been incurred if there had been no candidacy. Furthermore, expenditures,
including loans, may not be for the personal use of the candidate or any other
individual. Debt Incurred is any unpaid debt that the committee is obligated to
pay.

On the CFRS [FAQ page][search], the CFRS defines exactly what expenditures are
reported.

>  All receipts over $100 must be itemized and all sales items over $50 must be
itemized. All contributions from a political committee shall be itemized no
matter what amount. If the person who made the contribution is an organization,
and the total amount of contributions by such organization during the election
period exceeds $1,200, then the report shall also include the name and address
of one responsible party for the organization.

[cfrs]: https://cfrs.elections.delaware.gov/
[search]: https://cfrs.elections.delaware.gov/Public/ViewReceipts
[faq]: https://cfrs.elections.delaware.gov/Public/FAQ

## Download

We can use the search portal to find all record from 1975 to the current date.

```{r raw_dir}
raw_dir <- dir_create(here("state","de", "contribs", "data", "raw"))
raw_path <- path(raw_dir, "ViewContributionsList.csv")
```

First, we submit an `httr::POST()` request with all fields but the dates empty.

```{r raw_post, eval=FALSE}
de_post <- POST(
  url = "https://cfrs.elections.delaware.gov/Public/ViewReceipts",
  body = list(
    hdnTP = "",
    txtContributorName = "",
    txtFirstName = "",
    txtStreet = "",
    txtTown = "",
    MemberId = "",
    FilingYear = "",
    FilingPeriodName = "",
    ContributorType = "",
    ContributionType = "",
    ddlState = "",
    txtZipCode = "",
    txtZipExt = "",
    dtStartDate = "10/05/2020",
    dtEndDate = format(today(), "11/27/2022"),
    txtAmountRangeFrom = "",
    txtAmountRangeTo = "",
    ddlOffice = "",
    ddlCounty = "",
    ddlOfficeSought = "",
    ddljurisdiction = "",
    txtReceivingRegistrant = "",
    ddlEmployerOccupation = "",
    hdnFixedAssets = "",
    btnSearch = "Search",
    hdnddlOffice = "",
    hdnddlCounty = "",
    hdnddlOfficeSought = "",
    hdnddljurisdiction = ""
  )
)
```

Then, we use the cookies from that POST to make an `httr::GET()` request.

```{r raw_cookies, eval=FALSE}
de_cook <- cookies(de_post)$value
names(de_cook) <- cookies(de_post)$name
```

```{r raw_get, eval=FALSE}
if (!file_exists(raw_path) | file_size(raw_path) < "50M") {
  de_get <- GET(
    url = "https://cfrs.elections.delaware.gov/Public/ExportCSVNew",
    set_cookies(de_cook),
    write_disk(raw_path, overwrite = TRUE),
    query = list(
      `page` = "1",
      `orderBy` = "~",
      `filter` = "~",
      `Grid-size` = "15",
      `theme` = "vista"
    )
  )
}
```

## Read

Some errors with new line (`\n`) and double quote (`"`) characters needs to be
fixed with regular expressions and saved to a new text file.

```{r raw_fix}
fix_path <- file_temp(ext = "csv")
read_file(raw_path) %>%
  str_remove("(?<=Fixed Asset),") %>% 
  str_replace_all("(?<!,No|Asset)\n", " ") %>% 
  str_replace_all("\"", "\'") %>% 
  write_lines(fix_path)
```

The fixed file can be read with `readr::read_delim()`.

```{r raw_read}
# 304,202 rows
dec <- read_delim(
  file = fix_path,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_character(),
    `Contribution Date` = col_date_mdy(),
    `Contribution Amount` = col_double()
  )
)
```

For convenience, we can rename some character and remove empty columns.

```{r raw_clean}
old_names <- names(dec)
dec <- dec %>% 
  clean_names("snake") %>% 
  remove_empty("cols") %>% 
  mutate_at(vars(fixed_asset), to_logical) %>% 
  rename_all(str_remove, "contributor_") %>% 
  rename(
    contributor = name,
    recipient = receiving_committee,
    employer = employer_name,
    occupation = employer_occupation,
    method = contribution_type,
    period = filing_period,
  ) %>% 
  rename_all(str_replace, "address_line_", "addr") %>% 
  rename_all(str_remove, "contribution_")
```

Check the distinct values of a discrete variable to verify file reading.

```{r raw_check}
count(dec, fixed_asset)
```

## Explore

```{r glimpse}
glimpse(dec)
tail(dec)
```

### Missing

```{r na_count}
col_stats(dec, count_na)
```

Records missing a date, amount, or name need to be flagged.

```{r na_flag}
key_vars <- c("date", "contributor", "amount", "recipient")
dec <- flag_na(dec, all_of(key_vars))
percent(mean(dec$na_flag), 0.001)
```

All of these records are missing the contributor name.

```{r na_view}
dec %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

```{r na_recount}
dec %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  col_stats(count_na)
```

### Duplicates

The same can be done for records entirely duplicated more than once.

```{r dupe_flag}
dec <- flag_dupes(dec, everything())
percent(mean(dec$dupe_flag), 0.001)
```

```{r dupe_view}
dec %>% 
  filter(dupe_flag) %>% 
  select(all_of(key_vars)) %>% 
  arrange(date)
```

## Amounts

```{r ammount_summary}
summary(dec$amount)
mean(dec$amount <= 0, na.rm = TRUE)
```

```{r hist_amount, echo=FALSE}
dec %>%
  filter(!is.na(amount), amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Delaware Contributions Amount Distribution",
    subtitle = "from 2020 to 2022",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

## Dates

The calendar year can be added from the `date` column with `lubridate::year()`.

```{r date_year}
dec <- mutate(dec, year = year(date))
```

Aside from a handful of missing values, the `date` and `year` are fairly clean.

```{r date_range}
prop_na(dec$date)
min(dec$date, na.rm = TRUE)
sum(dec$year < 2000, na.rm = TRUE)
max(dec$date, na.rm = TRUE)
sum(dec$date > today(), na.rm = TRUE)
```

```{r bar_year, echo=FALSE}
dec %>% 
  count(year) %>% 
  filter(year > 2003) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2004, 2020, by = 2)) +
  coord_cartesian(xlim = c(2004, 2020)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Delaware Contributions by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
dec <- dec %>% 
  unite(
    col = addr_full,
    starts_with("addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    addr_norm = normal_address(
      address = addr_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-addr_full)
```

```{r address_view}
dec %>% 
  select(starts_with("addr")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
dec <- dec %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  dec$zip,
  dec$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
dec <- dec %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
dec %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  dec$state,
  dec$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
dec <- dec %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("DE", "DC", "DELAWARE"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
dec <- dec %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine] algorithms can be used to group similar strings and replace the
less common versions with their most common counterpart. This can greatly 
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- dec %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
nrow(good_refine)
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
dec <- dec %>% 
  left_join(good_refine, by = names(.)) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

A lot of the remaining invalid cities are really parts of the address.

```{r city_remain}
many_city <- c(valid_city, extra_city)
many_city <- c(many_city, usps_street$full, usps_street$abb, "DE")
dec %>% 
  filter(city_refine %out% many_city) %>% 
  count(city_refine, sort = TRUE)
```

```{r city_manual}
dec <- dec %>% 
  mutate(
    city_refine = city_refine %>% 
      str_replace("^WILM$", "WILMINGTON")
  )
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(dec$city_raw),
  dec$city_norm,
  dec$city_swap,
  dec$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(dec$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Delaware City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Delaware City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

```{r clean_select}
dec <- dec %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(dec, 100))
```

1. There are `r comma(nrow(dec))` records in the database.
1. There are `r comma(sum(dec$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(dec$na_flag))` records missing ....
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("state","de", "contribs", "data", "clean"))
clean_path <- path(clean_dir, "de_contribs_clean_20201005-20221127.csv")
write_csv(dec, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

The file strings need to be converted to ASCII.

```{r clean_conv}
read_lines(clean_path) %>% 
  iconv(to = "ASCII", sub = "byte") %>% 
  write_lines(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r s3_upload, eval=FALSE}
s3_path <- path("csv", basename(clean_path))
if (!object_exists(s3_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = s3_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
s3_head <- head_object(s3_path, "publicaccountability")
(s3_size <- as_fs_bytes(attr(s3_head, "content-length")))
unname(s3_size == clean_size)
```

## Dictionary

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(dec)),
  old = c(md_code(old_names), rep("", ncol(dec) - length(old_names))),
  type = md_code(map_chr(dec, typeof)),
  def = c(
    "Date contribution was made",
    "Contributor full name",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor 2-digit state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor type",
    "Contributor employer name",
    "Contributor occupation",
    "Contribution method",
    "Contribution amount or correction",
    "Unique campaign finance ID",
    "Recipient committee name",
    "Report filing period",
    "Office sought by recipient",
    "Fix asset flag",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized contributor street address",
    "Normalized contributor 5-digit ZIP code",
    "Normalized contributor 2-digit state abbreviation",
    "Normalized contributor city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original", "Type", "Definition")
))
```
