---
title: "Deleware Lobbyist Registration"
author: "Kiernan Nicholls & Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("state","de", "lobbying", "reg", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  RSelenium, # remote browser
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  httr, # http request
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data

Data is obtained from the [Deleware Public Integrity Commission][depic] (PIC). It was accessed on March 26
, 2023.

> Registration and authorization laws for those lobbying the General Assembly or State agencies,
whether paid or not. Quarterly expense reports required for direct expenditures on General Assembly
members or agency employees or officials. Lobbying activity reports with the Bill Number of
legislation and number or title of administrative action must be filed within 5 business days of
Lobbying. PIC submits weekly reports on that lobbying activity to the General Assembly while in
session.

[depic]: https://depic.delaware.gov/

## Import

From the PIC search page, we have the option to download the complete list of registered lobbyists.

> Enter the name of registered Delaware lobbyist to find employer/client list, lobbying expenses,
and address. Click magnifying glass to search.  
> [Download complete list](https://egov.delaware.gov/Lobs/Explore/DownloadReport?reportCode=LOB)

We can use `httr::GET()` to submit an HTTP request for the file. Then, the response content of that
request can be parsed as a CSV using `httr:content()` and `readr::read_csv()`.

Since some end dates are marked "Indefinite" when a lobbyist is still active, we transform the data to make the "indefinite" entries NAs instead when we are reading this column in as date.

```{r download_raw}
raw_url <- "https://pirs.delaware.gov/documents/EmployerAuthorizationCsv"
delr <- 
  GET(url = raw_url) %>%
  content(as = "raw") %>%
  read_csv(
    name_repair = make_clean_names,
    col_types = cols(
      .default = col_character(),
      lobbying_start_date = col_date_mdy(),
      lobbying_end_date = col_date_mdy()
    )
  )
```


Some of the variable names will be prepended with `lob_*` or truncated for simplicity and
clarity.

```{r names_change}
names(delr) <- str_replace(names(delr), "lobbyist_", "lob_") %>% 
  str_replace("employer_|empl_", "emp_") %>% 
  str_remove("lobbying_") %>% 
  str_remove("_code")
```

## Explore

We can `dplyr::glimpse()` to ensure our data was read and parsed properly.

```{r glimpse}
head(delr)
tail(delr)
glimpse(sample_frac(delr))
```

### Missing

There are no states missing key variables like `last_name` or `start_date`.

```{r glimpse_na}
col_stats(delr, count_na)
```

### Duplicates

There are a small number of duplicate records, which can be flagged with a new `dupe_flag` variable
using the `campfin::dupe_flag()` function.

```{r flag_dupes}
delr <- flag_dupes(delr, everything())
sum(delr$dupe_flag)
```

## Wrangle

To improve the searchability of the databse, we will normalize the variables for both lobbyist and
client.

### Address

For addressed, we will use `tidyr::unite()` to create a single variable, then normalize that new
variable with `campfin::normal_address()`.

```{r address_normal}
packageVersion("tidyr")
delr <- delr %>% 
  mutate(
    lob_address_norm = normal_address(
      address = lob_address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

The same process will be performed for `emp_address`.

```{r emp_address_normal}
delr <- delr %>% 
  mutate(
    emp_address_norm = normal_address(
      address = emp_address,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
delr %>% 
  select(starts_with("lob_address")) %>% 
  distinct() %>% 
  sample_frac()
```

### ZIP

```{r zip_normal}
delr <- mutate_at(
  .tbl = delr,
  .vars = vars(ends_with("_zip")),
  .funs = list(norm = normal_zip),
  na_rep = TRUE
)
```

```{r zip_progress}
progress_table(
  delr$lob_zip,
  delr$lob_zip_norm,
  delr$emp_zip,
  delr$emp_zip_norm,
  compare = valid_zip
)
```

### State

```{r state_normal}
delr <- mutate_at(
  .tbl = delr,
  .vars = vars(ends_with("_state")),
  .funs = list(norm = normal_state),
  na_rep = TRUE,
  valid = valid_state
)
```

```{r state_progress}
progress_table(
  delr$lob_state,
  delr$lob_state_norm,
  delr$emp_state,
  delr$emp_state_norm,
  compare = valid_state
)
```

```{r view_states}
select(delr, contains("state")) %>% 
  distinct() %>% 
  sample_frac()
```

### City

```{r normal_city}
delr <- mutate_at(
  .tbl = delr,
  .vars = vars(ends_with("_city")),
  .funs = list(norm = normal_city),
  abbs = usps_city,
  states = c("DE", "DC", "DELEWARE"),
  na = invalid_city,
  na_rep = TRUE
)
```

```{r lob_swap_city}
delr <- delr %>% 
  left_join(
    y = select(zipcodes, -state),
    by = c(
      "lob_zip_norm" = "zip"
    )
  ) %>% 
  rename(lob_city_match = city) %>% 
  mutate(
    lob_match_abb = is_abbrev(lob_city_norm, lob_city_match),
    lob_match_dist = stringdist(lob_city_norm, lob_city_match),
    lob_city_swap = if_else(
      condition = lob_match_abb | lob_match_dist == 1,
      true = lob_city_match,
      false = lob_city_norm
    )
  ) %>% 
  select(
    -lob_city_match,
    -lob_match_abb,
    -lob_match_dist
  )
```

```{r emp_swap_city}
delr <- delr %>% 
  left_join(
    y = select(zipcodes, -state),
    by = c(
      "emp_zip_norm" = "zip"
    )
  ) %>% 
  rename(emp_city_match = city) %>% 
  mutate(
    emp_match_abb = is_abbrev(emp_city_norm, emp_city_match),
    emp_match_dist = stringdist(emp_city_norm, emp_city_match),
    emp_city_swap = if_else(
      condition = emp_match_abb | emp_match_dist == 1,
      true = emp_city_match,
      false = emp_city_norm
    )
  ) %>% 
  select(
    -emp_city_match,
    -emp_match_abb,
    -emp_match_dist
  )
```

```{r city_progress}
progress_table(
  str_to_upper(delr$lob_city),
  delr$lob_city_norm,
  delr$lob_city_swap,
  str_to_upper(delr$emp_city),
  delr$emp_city_norm,
  delr$emp_city_swap,
  compare = valid_city
)
```

```{r city_view_bad}
delr %>% 
  filter(lob_city_swap %out% valid_city) %>% 
  count(
    lob_state_norm, 
    lob_zip_norm, 
    lob_city, 
    lob_city_norm,
    sort = TRUE
  )
```

### Year

```{r dates_range}
min(delr$start_date)
max(delr$start_date)
min(delr$end_date, na.rm = TRUE)
max(delr$end_date, na.rm = TRUE)
```

```{r year_add}
delr <- mutate(
  .data = delr,
  start_year = year(start_date),
  end_year = year(end_date)
)
```

## Conclude

1. There are `r nrow(delr)` records in the database.
1. There are `r sum(delr$dupe_flag)` duplicate records in the database.
1. There are zero records missing key date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 5-digit `zip_norm` variable has been created with `campfin::normal_zip()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r create_proc_dir}
proc_dir <- dir_create(here("state","de", "lobby", "data", "processed"))
```

```{r write_clean}
delr %>% 
  select(
    -lob_city_norm,
    -emp_city_norm,
  ) %>% 
  rename(
    lob_city_norm = lob_city_swap,
    emp_city_norm = emp_city_swap
  ) %>% 
  write_csv(
    path = glue("{proc_dir}/de_lobby_reg.csv"),
    na = ""
  )
```

