---
title: "{State} {Type}"
author: "First Last"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("hhs", "spending", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The data is obtained from the [USA Spending Award Data Archive][archive].

[archive]: https://www.usaspending.gov/#/download_center/award_data_archive

> Welcome to the Award Data Archive, which features major agencies’ award
transaction data for full fiscal years. They’re a great way to get a view into
broad spending trends and, best of all, the files are already prepared — you can
access them instantaneously.

## Download

If the zip archive has not been downloaded, we can do so now.

```{r download}
archive_url <- "https://files.usaspending.gov/award_data_archive/"
hhs_files <- glue("FY{2008:2020}_075_Contracts_Full_20200205.zip")
hhs_urls <- paste0(archive_url, hhs_files)
raw_dir <- dir_create(here("hhs", "data", "raw"))
hhs_paths <- path(raw_dir, hhs_files)
if (!all(file_exists(hhs_paths))) {
  download.file(hhs_urls, hhs_paths)
}
```

## Layout

The USA Spending website also provides a comprehensive data dictionary which
covers the many variables in this file.

```{r dictionary}
dict_file <- file_temp(ext = "xlsx")
download.file(
  url = "https://files.usaspending.gov/docs/Data_Dictionary_Crosswalk.xlsx",
  destfile = dict_file
)
dict <- read_excel(
  path = dict_file, 
  range = "A2:L414",
  na = "N/A",
  .name_repair = make_clean_names
)

hhs_names <- names(read_csv(last(hhs_paths), n_max = 0))
# get cols from hhs data
mean(hhs_names %in% dict$award_element)
dict <- dict %>% 
  filter(award_element %in% hhs_names) %>% 
  select(award_element, definition) %>% 
  mutate_at(vars(definition), str_replace_all, "\"", "\'") %>% 
  arrange(match(award_element, hhs_names))

dict %>% 
  head(10) %>% 
  mutate_at(vars(definition), str_trunc, 69) %>% 
  kable()
```


## Read

This archive file can be directly read as a data frame with `vroom::vroom()`.

```{r read}
hhs <- vroom(
  file = hhs_paths,
  delim = ",",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_types = cols(
    .default = col_skip(),
    action_date_fiscal_year = col_integer(),
    action_date = col_date(),
    modification_number = col_integer(),
    parent_award_agency_id = col_integer(),
    parent_award_agency_name = col_character(),
    total_dollars_obligated = col_double(),
    awarding_agency_code = col_integer(),
    awarding_agency_name = col_character(),
    awarding_sub_agency_code = col_integer(),
    awarding_sub_agency_name = col_character(),
    awarding_office_code = col_integer(),
    awarding_office_name = col_character(),
    recipient_name = col_character(),
    recipient_parent_name = col_character(),
    recipient_country_name = col_character(),
    recipient_address_line_1 = col_character(),
    recipient_city_name = col_character(),
    recipient_state_code = col_character(),
    recipient_zip_4_code = col_character(),
    recipient_phone_number = col_character(),
    primary_place_of_performance_city_name = col_character(),
    primary_place_of_performance_state_code = col_character(),
    primary_place_of_performance_zip_4 = col_character(),
    award_type_code = col_character(),
    award_description = col_character()
  )
)
```

```{r}
# properly read
count(hhs, awarding_sub_agency_name, sort = TRUE)
```

```{r}
# filter to only cdc
# cdc <- filter(cdc, parent_award_agency_id == 7523)
```


## Explore

```{r glimpse}
head(hhs)
tail(hhs)
glimpse(sample_n(hhs, 20))
```

```{r}
hhs %>%
  mutate(y = year(action_date), q = quarter(action_date)) %>%
  group_by(y, q) %>%
  summarise(sum = sum(total_dollars_obligated, na.rm = TRUE) / 1e9) %>% 
  ggplot(aes(x = q, y = sum)) +
  geom_col(aes(fill = sum)) +
  scale_fill_viridis_c(end = 0.75, guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  facet_wrap(~y, nrow = 1) +
  theme(
    panel.grid.minor.x = element_blank()
  ) +
  labs(
    title = "Department of Health and Human Services Spending",
    caption = "Source: USASpending.gov",
    x = "Quarter",
    y = "Obligated Spending (Billion USD)"
  )
```

## Conclude

```{r clean_glimpse}
# glimpse(sample_n(hhs, 20))
```

1. There are `r nrow(hhs)` records in the database.
1. There are `r sum(hhs$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(hhs$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("hhs", "spending", "data", "clean"))
```

```{r write_clean, eval=FALSE}
write_csv(
  x = hhs,
  path = path(clean_dir, "hhs_contracts_clean.csv"),
  na = ""
)
```

