---
title: "United States Broker-Dealers"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document:
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options:
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("us", "sec_brokers", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("us/sec_brokers/sec_broker-dealers_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

The Security Exchange Commission (SEC) requires brokers and dealers to register.

> Most "brokers" and "dealers" must register with the SEC and join a
"self-regulatory organization," or SRO. This section covers the factors that
determine whether a person is a broker or dealer. It also describes the types of
brokers and dealers that do not have to register with the SEC. Self-regulatory
organizations are described in Part III, below.

> ### A. Who is a "Broker"
> any person engaged in the business of effecting transactions in securities for
the account of others.

> ### B. Who is a "Dealer"
> any person engaged in the business of buying and selling securities for his
own account, through a broker or otherwise.

The SEC then hosts this list online for people to download.

```{r}
home <- read_html("https://www.sec.gov/help/foiadocsbdfoiahtm.html")
```

> ### What You Should Know About the Data File
> This ascii text file contains the Central Index Key (CIK) numbers, company
names, SEC reporting file numbers, and addresses (business addresses are
provided when mailing addresses are not available) of active broker-dealers who
are registered with the SEC.
> 
> The information in this file is **raw data** — data that is meant to be used
as input to another program. The data items are provided as a "tab delimited"
file. Although the file can be viewed in any program that accepts ASCII text
(for example, a word processor), the data fields are best viewed when imported
into a program that accepts delimited data, such as a spreadsheet. The record
layout and maximum field sizes are shown below for those who want to process the
data into another form.

```{r}
raw_cols <- home %>% 
  html_element("table") %>% 
  html_table()
```

```{r echo=FALSE}
kable(raw_cols)
raw_cols <- clean_names(raw_cols)
```

## Download

We are going to download _all_ of the files listed on this home page.

```{r}
down <- html_elements(home, "table")[[2]]
```

```{r}
raw_url <- down %>% 
  html_elements("a") %>% 
  html_attr("href") %>% 
  str_c("https://www.sec.gov", .)
```

```{r}
down_tbl <- down %>% 
  html_table() %>% 
  mutate(
    across(everything(), str_remove, "^\\w+:(\\s+)?"),
    across(Size, as_fs_bytes),
    url = raw_url
  )
```

```{r echo=FALSE}
down_tbl %>% 
  mutate(File = glue("[{File}]({url})")) %>% 
  select(-url) %>% 
  head() %>% 
  kable()
```

```{r raw-dir}
raw_dir <- dir_create(here("us", "sec_brokers", "data", "raw"))
raw_tsv <- path(raw_dir, basename(raw_url))
```

```{r raw-download}
if (!all(file_exists(raw_tsv))) {
  for (i in seq_along(raw_url)) {
    cli::cli_h2(basename(raw_url[i]))
    RETRY(
      verb = "GET",
      url = raw_url[i],
      pause_base = 2,
      quiet = FALSE,
      progress("down"),
      write_disk(
        path = raw_tsv[i], 
        overwrite = TRUE
      )
    )
  }
}
```

## Read

Some files vary in their line ending characters. We can quickly remove these.

```{r}
tmp_tsv <- path_temp(basename(raw_tsv))
for (i in seq_along(tmp_tsv)) {
  message(i)
  read_file(raw_tsv[i]) %>% 
    str_replace_all("\r\r\n", "\r") %>% 
    write_file(tmp_tsv[i])
}
```

Now each fixed file can be read into a single file.

```{r raw-read}
sec <- read_delim(
  file = tmp_tsv,
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  col_names = raw_cols$field_name,
  name_repair = make_clean_names,
  id = "source_file",
  na = c(" ", ""),
  col_types = cols(
    .default = col_character()
  )
)
```

We will rearrange the columns to add the source file and date from the download
page table to the end.

```{r clean-names}
sec <- sec %>% 
  select(-X9) %>% 
  mutate(across(source_file, basename)) %>% 
  relocate(source_file, .after = last_col()) %>% 
  left_join(
    down_tbl %>% 
      mutate(source_file = basename(url)) %>% 
      select(source_file, source_date = File)
  )
```

Many registrants are listed in multiple quarterly files. We are going to keep
only the most recent registration per entry but also create a new column 
indicating the _first_ file they appeared in.

```{r date-add}
sec <- sec %>%
  group_by(cik_number) %>% 
  mutate(
    source_date = parse_date(source_date, "%B %Y"),
    first_date = min(source_date),
    last_date = max(source_date),
  ) %>% 
  arrange(desc(source_date)) %>% 
  slice(1) %>% 
  ungroup() %>% 
  select(-source_date)
```

About half the registrants are listed in the most recent file.

```{r most-recent}
mean(sec$source_date == max(sec$source_date))
```

## Explore

There are `r comma(nrow(sec))` rows of `r ncol(sec)` columns. Each record
represents a single broker-dealer registered with the SEC.

```{r glimpse}
glimpse(sec)
tail(sec)
```

### Missing

Columns vary in their degree of missing values. Only the `address2` and
`state_code` columns are missing data.

```{r na-count}
col_stats(sec, count_na)
```

The rows missing `state_code` have a foreign `city` value, which makes sense.

```{r na-state}
sec %>% 
  filter(is.na(state_code)) %>% 
  count(city, sort = TRUE)
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date-year}
sec <- mutate(sec, last_year = year(last_date))
```

```{r date-range}
min(sec$last_date)
sum(sec$last_year < 2000)
max(sec$last_date)
sum(sec$last_date > today())
```

```{r bar-year, echo=FALSE}
sec %>%
  count(last_year) %>%
  ggplot(aes(x = last_year, y = n)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2007, 2022)) +
  theme(legend.position = "bottom") +
  labs(
    title = "United States Broker-Dealers by Last Listed Year",
    caption = "Source: SEC",
    fill = "Election Year",
    x = "Year Last Registered",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official
USPS suffixes.

```{r address-norm}
sec <- sec %>%
  mutate(
    address1_norm = normal_address(address1, usps_street),
    address2_norm = normal_address(address2, usps_street, abb_end = FALSE)
  ) %>%
  unite(
    col = address_norm,
    matches("^address\\d_norm$"),
    sep = " ",
    remove = TRUE,
    na.rm = TRUE
  )
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip-norm}
sec <- sec %>%
  mutate(
    zip_norm = normal_zip(
      zip = zip_code,
      na_rep = TRUE
    )
  )
```

```{r zip-progress}
progress_table(
  sec$zip_code,
  sec$zip_norm,
  compare = valid_zip
)
```

### State

All of the "invalid" `state_code` values are from overseas.

```{r}
prop_in(sec$state_code, valid_state)
unique(what_out(sec$state_code, valid_state))
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city-norm}
norm_city <- sec %>%
  distinct(city, state_code, zip_norm) %>%
  mutate(
    city_norm = normal_city(
      city = city,
      abbs = usps_city,
      states = c("US", "DC", "UNITED STATES"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city-swap}
norm_city <- norm_city %>%
  rename(city_raw = city) %>%
  left_join(
    y = zipcodes,
    by = c(
      "state_code" = "state",
      "zip_norm" = "zip"
    )
  ) %>%
  rename(city_match = city) %>%
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>%
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

```{r city-rejoin}
sec <- left_join(
  x = sec,
  y = norm_city,
  by = c(
    "city" = "city_raw",
    "state_code",
    "zip_norm"
  )
)
```

#### Progress

Our goal for normalization was to increase the proportion of city values known
to be valid and reduce the total distinct values by correcting misspellings.

```{r city-progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(sec$city),
  sec$city_norm,
  sec$city_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
progress %>%
    mutate(across(stage, md_code)) %>%
    kable(digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar-progress, echo=FALSE}
raw_in <- percent(prop_in(sec$city, valid_city))
progress %>%
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "United States City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar-distinct, echo=FALSE}
progress %>%
  select(
    stage,
    all = n_distinct,
    bad = n_diff
  ) %>%
  mutate(good = all - bad) %>%
  pivot_longer(c("good", "bad")) %>%
  mutate(name = name == "good") %>%
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "United States City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean-select}
sec <- sec %>%
  select(
    -city_norm,
    city_clean = city_swap,
  ) %>%
  rename_all(~str_replace(., "_norm", "_clean")) %>%
  rename_all(~str_remove(., "_raw")) %>%
  relocate(address_clean, city_clean, .before = zip_clean)
```

## Conclude

```{r clean-glimpse}
glimpse(sample_n(sec, 1000))
```

1. There are `r comma(nrow(sec))` records in the database.
1. There are `r comma(sum(sec$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(sec$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server. We
will name the object using a date range of the records included.

```{r clean-timestamp}
min_dt <- str_remove_all(min(sec$first_date), "-")
max_dt <- str_remove_all(max(sec$last_date), "-")
csv_ts <- paste(min_dt, max_dt, sep = "-")
```

```{r clean-dir}
clean_dir <- dir_create(here("us", "sec_brokers", "data", "clean"))
clean_csv <- path(clean_dir, glue("sec_broker-dealers_{csv_ts}.csv"))
clean_rds <- path_ext_set(clean_csv, "rds")
basename(clean_csv)
```

```{r clean-write}
write_csv(sec, clean_csv, na = "")
write_rds(sec, clean_rds, compress = "xz")
(clean_size <- file_size(clean_csv))
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws-upload, eval=FALSE}
aws_key <- path("csv", basename(clean_csv))
if (!object_exists(aws_key, "publicaccountability")) {
  put_object(
    file = clean_csv,
    object = aws_key,
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_key, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
