---
title: "Ohio Contributions"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("oh", "contribs", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  rvest, # read html pages
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The data is obtained from the [Ohio Secretary of State][sos]. The OH SOS offers
a file transfer page (FTP) to download data in bulk rather than via searches.

>  Welcome to the Ohio Secretary of State's Campaign Finance File Transfer Page.
This page was developed to allow users to obtain large sets of data faster than
the normal query process. At this page you can download files of pre-queried
data, such as all candidate contributions for a particular year or a list of all
active political action committees registered with the Secretary of State. In
addition, campaign finance data filed prior to 2000 is available only on this
site. These files contain all relevant and frequently requested information. If
you are looking for smaller or very specific sets of data please use the regular
Campaign Finance queries listed on the tabs above.

[sos]: https://www.ohiosos.gov/

## Import

### Download

> On the FTP page, please decide which information you would like to download.
Click "Download File" on the right hand side. The system will then proceed to
download the file into Microsoft Excel or provide you will an opportunity to
download the file to the location on your computer (the settings on your
computer will dictate this). You may see a series of dialog boxes on your screen
asking you if you want to run or save the zipped `.exe` file. Follow the dialog
boxes for whichever you chose telling the computer where you want the files
saved. The end result will be a `.csv` file that you can open in Microsoft Excel
or some other database application.

We can download all the contribution files by reading the FTP website itself
and scraping each of the "Download" links in the table. This process needs to
be repeated for candidates, PACs, and parties.

```{r raw_urls}
ftp_base <- "https://www6.ohiosos.gov/ords/"
t <- c("CAN", "PAC", "PARTY")
ftp_url <- glue("f?p=CFDISCLOSURE:73:7027737052457:{t}:NO:RP:P73_TYPE:{t}:")
ftp_url <- str_c(ftp_base, ftp_url)
ftp_params <- character()
ftp_table <- rep(list(NA), length(t))
for (i in seq_along(t)) {
  ftp_page <- read_html(ftp_url[i])
  table_id <- paste0("#", str_extract(ftp_page, '(?<=id\\=")report_.*(?="\\s)'))
  ftp_table[[i]] <- ftp_page %>%
    html_node(table_id) %>%
    html_table() %>%
    as_tibble() %>%
    select(-last_col()) %>%
    set_names(c("file", "date", "size")) %>%
    mutate_at(vars(2), parse_date_time, "%m/%d/%Y %H:%M:%S %p")
  con_index <- str_which(ftp_table[[i]]$file, "Contributions\\s-\\s\\d+")
  ftp_params <- ftp_page %>%
    html_node(table_id) %>%
    html_nodes("tr") %>%
    html_nodes("a") %>%
    html_attr("href") %>%
    str_subset("f\\?p") %>%
    extract(con_index) %>%
    append(ftp_params)
}
```

Then each link can be downloaded to the `/data/raw` directory.

```{r wget}
wget <- function(url, dir) {
  system2(
    command = "wget",
    args = c(
      "--no-verbose", 
      "--content-disposition", 
      url, 
      paste("-P", raw_dir)
    )
  )
}

quiet_wget <- quietly(wget)
```

```{r raw_download}
raw_dir <- dir_create(here("oh", "contribs", "data", "raw"))
raw_urls <- paste0(ftp_base, ftp_params)
if (length(dir_ls(raw_dir)) < 82) {
  map(raw_urls, quiet_wget, raw_dir)
}
raw_paths <- dir_ls(raw_dir)
```

### Read

> The data is in a "comma delimited" format that loads easily into Microsoft
Excel or Access as well as many other spreadsheet or database programs. Many of
the available files contain a significant quantity of data records. A
spreadsheet program, such as Microsoft Excel, may not allow all of the data in a
file to be loaded because of a limit on the number of available rows. For this
reason, it is advised that a database application be utilized to load and work
with the data available at this site...

We can read all `r length(raw_paths)` raw CSV files into a single data frame
using `purrr::map_df()` and `readr::read_csv()`.

```{r read_raw}
ohc <- map_df(
  .x = sample(raw_paths, 2),
  .f = read_csv,
  na = c("", "NA", "N/A"),
  col_types = cols(.default = "c")
)

# ohc <- ohc %>% 
#   mutate_at(vars(FILE_DATE), parse_date, "%m/%d/%Y") %>% 
#   mutate_at(vars(AMOUNT), parse_double) %>% 
#   mutate_at(vars(MASTER_KEY, RPT_YEAR, REPORT_KEY, DISTRICT), parse_integer) 
```

```{r raw_rename, echo=FALSE}
ohc <- ohc %>% 
  clean_names() %>% 
  rename(
    rpt_desc = report_description,
    rpt_key = report_key,
    desc = short_description,
    first = first_name,
    middle = middle_name,
    last = last_name,
    suffix = suffix_name,
    non_ind = non_individual,
    pac_no = pac_reg_no,
    date = file_date,
    event = event_date,
    occupation = emp_occupation,
    inkind = inkind_description,
    other_type = other_income_type,
    cand_first = candidate_first_name,
    cand_last = candidate_last_name
  )
```

## Explore

```{r glimpse}
head(ohc)
tail(ohc)
glimpse(sample_n(ohc, 20))
```

### Missing

```{r na_count}
col_stats(ohc, count_na)
```

```{r na_flag}
ohc <- ohc %>% 
  unite(
    first, middle, last, non_ind,
    col = con_name,
    sep = " ",
    na.rm = TRUE
  ) %>% 
  flag_na(date, con_name, amount, com_name)

sum(ohc$na_flag)
percent(mean(ohc$na_flag), 0.01)
```
