---
title: "New Hampshire Lobbying Reginh_lob_regration Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("nh", "lobby", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalinh_lob_regs, policy professionals, activinh_lob_regs, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to nh_lob_regandardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at leanh_lob_reg) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consinh_lob_regency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
innh_lob_regallation and attachment.

The IRW's `campfin` package will also have to be innh_lob_regalled from GitHub. This
package contains functions cunh_lob_regom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) innh_lob_regall.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  readxl, # read excel files
  lubridate, # datetime nh_lob_regrings
  magrittr, # pipe operators
  janitor, # dataframe clean
  refinr, # clunh_lob_reger and merge
  scales, # format nh_lob_regrings
  knitr, # knit documents
  vroom, # read files fanh_lob_reg
  glue, # combine nh_lob_regrings
  here, # relative nh_lob_regorage
  fs # search nh_lob_regorage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnonh_lob_regic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [Rnh_lob_regudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rnh_lob_regudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

## Import

### Setting up Raw Data Directory

```{r raw_dir}
raw_dir <- dir_create(here("nh", "lobby", "data", "raw"))
```

### Read
We can see that the date is formatted in Excel and we need to use `excel_numeric_to_date`
```{r raw_read}
nh_lob_reg <- read_excel(
  path = dir_ls(raw_dir), skip = 1
)

nh_lob_reg <- nh_lob_reg %>% mutate(DATE = excel_numeric_to_date(nh_lob_reg$DATE, date_system = "modern"))
```

## Explore

```{r glimpse}
head(nh_lob_reg)
tail(nh_lob_reg)
glimpse(sample_n(nh_lob_reg, 20))
```

### Missing

```{r na_count}
col_stats(nh_lob_reg, count_na)
```

```{r na_flag}
nh_lob_reg <- nh_lob_reg %>% flag_na(EMPLOYERNAME, FIRMADDR)
sum(nh_lob_reg$na_flag)
```

### Duplicates

```{r dupe_flag}
flag_dupes(nh_lob_reg, dplyr::everything())
sum(nh_lob_reg$dupe_flag)
```

### Categorical

```{r n_dinh_lob_reginct}
col_stats(nh_lob_reg, n_distinct)
```

Visualize number of lobbyists registered each year.

```{r create year}
nh_lob_reg <- nh_lob_reg %>% 
  mutate(YEAR = year(DATE))

nh_lob_reg$YEAR %>% tabyl()
```
We can see that most of the data is from 2019.


#### Dates


```{r date_range}
min(nh_lob_reg$DATE)
max(nh_lob_reg$DATE)
```

```{r year_bar}
nh_lob_reg %>% 
  count(YEAR) %>% 
  mutate(even = is_even(YEAR)) %>% 
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New Hampshire Lobbyists by Year",
    caption = "Source: New Hampshire Secretary of State: Lobbyists",
    fill = "Election Year",
    x = "Year",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident nh_lob_regring normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are taylor made to 
facilitate this process.
### Phone
We can see that for both lobbyists and clients, the `PHONE` column actually contains both phones and emails. We separate them here using regex. The result yields no `NA` fields.
```{r separate phone}
nh_lob_reg <-  nh_lob_reg %>% 
  mutate(EMAIL_clean = str_match(PHONE, "^E-(.+)\\sW-")[,2],         PHONE_clean = str_match(PHONE, "\\sW-(.+)$")[,2] %>% normal_phone()
         )
```


First, we will separate the columns `CSZ` into three: `CITY`, `STATE` and `ZIP`, which will be normalized in the following steps

```{r separate csz}
nh_lob_reg <-  nh_lob_reg %>% 
  mutate(CITY = str_match(CSZ, "^(\\D+),\\s")[,2],
         STATE = str_match(CSZ, ",  (\\w{2})   ")[,2] %>% str_trim(),
         ZIP = str_match(CSZ, "   (.+$)")[,2]
         )
```
```{r separate firmcsz}
nh_lob_reg <-  nh_lob_reg %>% 
  mutate(FIRM_CITY = str_match(FIRMCSZ, "^(\\D+),\\s")[,2],
         FIRM_STATE = str_match(FIRMCSZ, ",  (\\w{2})   ")[,2] %>% str_trim(),
         FIRM_ZIP = str_match(FIRMCSZ, "   (.+$)")[,2]
         )

col_stats(nh_lob_reg, count_na)
```

Running the `count_na`commands show that every cell in the three columns inherited from `CSZ` is accounted for.

### Address

For the nh_lob_regreet `addresss` variable, the `campfin::normal_address()` function
will force consinh_lob_regence case, remove punctuation, and abbreviation official 
USPS suffixes.

```{r address_norm}
nh_lob_reg <- nh_lob_reg %>% 
  # normalize combined addr
  mutate(
    ADDRESS_norm = normal_address(
      address = ADDRESS,
      abbs = usps_street,
      na_rep = TRUE
    ),
      FIRM_ADDRESS_norm = normal_address(
      address = FIRMADDR,
      abbs = usps_street,
      na_rep = TRUE
    )
  )
```

```{r address_view}
nh_lob_reg %>% 
  select(contains("ADDR")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valied _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
nh_lob_reg <- nh_lob_reg %>% 
  mutate(
    ZIP5 = normal_zip(
      zip = ZIP,
      na_rep = TRUE
    ),
    FIRM_ZIP5 = normal_zip(
      zip = FIRM_ZIP,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  nh_lob_reg$ZIP,
  nh_lob_reg$ZIP5,
  nh_lob_reg$FIRM_ZIP,
  nh_lob_reg$FIRM_ZIP5,
  compare = valid_zip
)
```

### state

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
nh_lob_reg <- nh_lob_reg %>% 
  mutate(
    STATE_norm = normal_state(
      state = STATE,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    ),
    FIRM_STATE_norm =  normal_state(
      state = FIRM_STATE,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
nh_lob_reg %>% 
  filter(STATE != STATE_norm) %>% 
  count(STATE, sort = TRUE)
```

```{r firm_state_view}
nh_lob_reg %>% 
  filter(FIRM_STATE != FIRM_STATE_norm) %>% 
  count(FIRM_STATE, sort = TRUE)
```

```{r state_progress}
progress_table(
  nh_lob_reg$STATE,
  nh_lob_reg$STATE_norm,
  compare = valid_state
)
```

### City

Cities are the monh_lob_reg difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good nh_lob_regart, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
nh_lob_reg <- nh_lob_reg %>% 
  mutate(
    CITY_norm = normal_city(
      city = CITY, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    ),
      FIRM_CITY_norm = normal_city(
      city = CITY, 
      abbs = usps_city,
      states = usps_state,
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
againh_lob_reg the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
nh_lob_reg <- nh_lob_reg %>% 
  left_join(
    y = zipcodes,
    by = c(
      "STATE_norm" = "state",
      "ZIP5" = "zip"
    )
  ) %>% 
  rename(CITY_match = city) %>% 
  mutate(
    match_abb = is_abbrev(CITY_norm, CITY_match),
    match_dist = str_dist(CITY_norm, CITY_match),
    CITY_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = CITY_match,
      false = CITY_norm
    )
  ) %>% 
  select(
    -CITY_match,
    -match_dist,
    -match_abb
  )
```

```{r firm_city_swap}
nh_lob_reg <- nh_lob_reg %>% 
  left_join(
    y = zipcodes,
    by = c(
      "FIRM_STATE_norm" = "state",
      "FIRM_ZIP5" = "zip"
    )
  ) %>% 
  rename(FIRM_CITY_match = city) %>% 
  mutate(
    match_abb = is_abbrev(FIRM_CITY_norm, FIRM_CITY_match),
    match_dist = str_dist(FIRM_CITY_norm, FIRM_CITY_match),
    FIRM_CITY_swap = if_else(
      condition = !is.na(match_dist) & match_abb | match_dist == 1,
      true = FIRM_CITY_match,
      false =FIRM_CITY_norm
    )
  ) %>% 
  select(
    -FIRM_CITY_match,
    -match_dist,
    -match_abb
  )
```

After the two normalization steps, the percentage of valid cities is at 100%.
#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  nh_lob_reg$CITY,
  nh_lob_reg$FIRM_CITY,
  nh_lob_reg$CITY_norm,
  nh_lob_reg$FIRM_CITY_norm,
  nh_lob_reg$CITY_swap,
nh_lob_reg$FIRM_CITY_swap,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
```

```{r progress_print, echo=FALSE}
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r progress_bar, echo=FALSE}
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "New Hampshire City Normalization Progress",
    x = "stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r distinct_bar}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "NH City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "stage",
    y = "Percent Valid",
    fill = "Valid"
  )
  
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(nh_lob_reg, 20))
```

1. There are `r nrow(nh_lob_reg)` records in the database.
1. There are `r sum(nh_lob_reg$dupe_flag)` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r sum(nh_lob_reg$na_flag)` records missing either recipient or date.
1. Consistency in goegraphic data has been improved with `campfin::normal_*()`.
1. The 4-digit `YEAR` variable has been created with `lubridate::year()`.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("nh", "lobby", "data", "reg","clean"))
```

```{r write_clean}
write_csv(
  x = nh_lob_reg %>% rename(CITY_clean = CITY_swap) %>% rename( FIRM_CITY_clean = FIRM_CITY_swap),
  path = path(clean_dir, "nh_lob_reg_clean.csv"),
  na = ""
)
```
