---
title: "California Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ca", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the Secretary of State's [Cal-Access database][03].

> Cal-Access provides financial information supplied by state candidates,
donors, lobbyists, and others. Simply start by selecting either Campaign Finance
Activity, Lobbying Activity, or Cal-Access Resources and we will guide you to
the information.

The [Political Reform Division][04] (PRD) collects the data regarding expenditures made.

> The Political Reform Division administers provisions of California's Political
Reform Act, including the law's most fundamental purpose of ensuring that
"receipts and expenditures in election campaigns should be fully and truthfully
disclosed in order that the voters may be fully informed and the activities of
lobbyists should be regulated and their finances disclosed..."

> * Receive campaign disclosure statements (itemizing contributions received and
expenditures made) filed by individuals and committees raising or spending
campaign funds to support or oppose state candidates or ballot measures. (Local
campaign committees file their itemized disclosure statements with local filing
officers).

[03]: https://www.sos.ca.gov/
[04]: https://www.sos.ca.gov/campaign-lobbying/about-political-reform-division/

## Documentation

Using the data key provided by Cal-Access, we can find the expenditure file and understand it's 
contents.

```{r key_size}
raw_dir <- dir_create(here("ca", "expends", "data", "raw"))
key_url <- "https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip"
key_file <- str_c(raw_dir, basename(key_url), sep = "/")
url_file_size(key_url)
```

If they ZIP file containing the documentation files has not yet been downloaded, we can do so now.

```{r key_download}
download.file(key_url, key_file)
```

Before we unzip the file, we can view it's contents.

```{r key_list}
key_content <- unzip(key_file, list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>% 
  mutate(across(length, as_fs_bytes)) %>% 
  mutate(across(name, as_fs_path))
```

```{r key_extract}
key_files <- as_fs_path(unzip(key_file, exdir = raw_dir))
```

## Download

All California campaign finance data can be downloaded in a single file. As described on PRD
website:

> ### Raw Data for Campaign Finance and Lobbying Activity > In addition to
presenting California campaign finance and lobbying activity on the
user-friendly [CAL-ACCESS website][05], the Secretary of State provides the raw
data to allow people with technical expertise to create their own databases. 
>
> The raw data is presented in tab-delimited text files from corresponding
tables in the CAL-ACCESS database. Users can uncompress and extract the data
with standard software such as PKZIP, WinZip, or MacZip. The Secretary of State
does not provide technical support to users who want to download and analyze
CAL-ACCESS data in this advanced way. However, the Secretary of State offers
guides to the CAL-ACCESS data structure and fields.

> ### Download Files  
* [Guides to CAL-ACCESS data structure and fields (ZIP)][06]
* [CAL-ACCESS raw data (ZIP)][07]
> 
> All CAL-ACCESS users should keep in mind that campaign finance and lobbying
activity information changes often. The raw data extracts are updated once a
day. Campaign finance and lobbying activity filings can also be obtained in hard
copy by contacting the Secretary of State's Political Reform Division.

[05]: http://cal-access.sos.ca.gov/
[06]: https://campaignfinance.cdn.sos.ca.gov/calaccess-documentation.zip
[07]: https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip

```{r zip_check}
zip_url <- "https://campaignfinance.cdn.sos.ca.gov/dbwebexport.zip"
zip_file <- path(raw_dir, basename(zip_url))
url_file_size(zip_url)
```

```{r zip_download}
if (!this_file_new(zip_file)) {
  download.file(zip_url, zip_file)
}
```

```{r zip_list}
zip_contents <- unzip(zip_file, list = TRUE) %>% 
  as_tibble(.name_repair = make_clean_names) %>% 
  mutate(across(length, as_fs_bytes)) %>% 
  mutate(across(name, as_fs_path))
```

```{r zip_extract}
all_paths <- as_fs_path(unzip(
  zipfile = zip_file,
  exdir = raw_dir
))
```

## Join

There are `r length(dir_ls(raw_dir, recurse = TRUE))` files in the data and
documentation ZIP archives. 

The `r md_code(exp_name)` file contains records of expenditures made by
campaigns, but it lacks some of the variables needed to fully identify both
parties to each transactions. We will need to join records with 

## Read

```{r raw_read}
expn_path <- str_subset(all_paths, "EXPN")
expn_cd <- read_delim(
  file = expn_path,
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  guess_max = 0,
  locale = locale(tz = "US/Pacific"),
  col_types = cols(
    .default = col_character(),
    FILING_ID = col_integer(),
    AMEND_ID = col_integer(),
    EXPN_DATE = col_date("%m/%d/%Y %H:%M:%S %p"),
    AMOUNT = col_double(),
    CUM_YTD = col_double(),
    CUM_OTH = col_double(),
    LINE_ITEM = col_integer()
  )
)
```

## Explore

```{r glimpse}
glimpse(cae)
tail(cae)
```

### Missing

```{r na_count}
col_stats(cae, count_na)
```

```{r na_flag}
cae <- cae %>% flag_na(date, last, amount, committee)
sum(cae$na_flag)
```

```{r na_view}
cae %>% 
  filter(na_flag) %>% 
  select(date, last, amount, committee)
```

### Duplicates

```{r dupe_flag}
cae <- flag_dupes(cae, -id)
sum(cae$dupe_flag)
```

```{r dupe_view}
cae %>% 
  filter(dupe_flag) %>% 
  select(date, last, amount, committee)
```

### Categorical

```{r distinct_count}
col_stats(cae, n_distinct)
```

```{r distinct_plots}
explore_plot(cae, type)
```

### Amounts

```{r ammount_summary}
summary(cae$amount)
mean(cae$amount <= 0)
```

```{r hist_amount, echo=FALSE}
cae %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "California Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
cae <- mutate(cae, year = year(date))
```

```{r date_range}
min(cae$date)
sum(cae$year < 2000)
max(cae$date)
sum(cae$date > today())
```

```{r bar_year, echo=FALSE}
cae %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "California Expenditures by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
cae <- cae %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
cae %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
cae <- cae %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  cae$zip,
  cae$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
cae <- cae %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
cae %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  cae$state,
  cae$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
cae <- cae %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("CA", "DC", "CALIFORNIA"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
cae <- cae %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- cae %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
cae <- cae %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(cae$city_raw),
  cae$city_norm,
  cae$city_swap,
  cae$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(cae$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "California City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "California City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
cae <- cae %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(cae, 50))
```

1. There are `r comma(nrow(cae))` records in the database.
1. There are `r comma(sum(cae$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(cae$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ca", "expends", "data", "clean"))
clean_path <- path(clean_dir, "ca_expends_clean.csv")
write_csv(cae, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(cae)),
  type = md_code(map_chr(cae, typeof)),
  def = c(
    "Date contribution was made",
    "Contribution amount or correction",
    "Contributor name",
    "Contributor unique ID",
    "Contributor street address",
    "Contributor secondary address",
    "Contributor city name",
    "Contributor state abbreviation",
    "Contributor ZIP+4 code",
    "Contributor occupation",
    "Contributor employer name",
    "Contributor type",
    "Recipient committee name",
    "Recipient unique ID",
    "Contribution method",
    "Recipient election or office sought",
    "Flag for missing date, amount, or name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

```{r dict_write}
write_lines(
  x = c("# California Contracts Data Dictionary\n", dict_md),
  path = here("ca", "expends", "ca_expends_dict.md"),
)
```
