---
title: "Illinois Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("il", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Data is obtained from the [Illinois State Board of Election (SBOE)][sboe].
Using the [SBOE search portal][portal], we can search for all expenditures
between two dates and export them to a tab-delimited file.

Infuriatingly, a maximum of 25,000 records can be downloaded at a time. This is
typically the number of expenditures made in 1-3 months, depending on the year.
To obtain all records between 2000 and 2020, we will have export hundreds of
small files.

> Click the link below for the file type desired. Choose "Save" to copy the file
to your computer. The maximum number of records available for download is
25,000. File Type:
> 
> * Tab-Delimited Text File
> * CSV File
> * XML File

> Chapter 10 ICLS 5/9-17, Illinois Compiled Statutes, governs the dissemination
and use of campaign disclosure data.
>
> * Any person who shall sell or utilize information from statements and reports
filed with the State Board of Elections for the purpose of soliciting
contributions or for the purpose of business solicitation is guilty of a Class B
misdemeanor.
> * Any person who alters or falsifies this information and publishes,
circulates, or distributes such altered or falsified information with the intent
to misrepresent contributions received or expenditures made by a candidate or
political committee is guilty of a Class B misdemeanor.

> Please review the following facts concerning campaign disclosure data files:
>
> * Data is updated continuously.
> * _This data may contain errors that have not yet been identified and
corrected._
> * The Board provides files of campaign disclosure data; the Board does not
provide software to process this data.

[sboe]: https://www.elections.il.gov
[portal]: https://www.elections.il.gov/CampaignDisclosure/ExpenditureSearchByAllExpenditures.aspx

## Download

Ideally, we would automate the download of these hundreds of files using
a combination of `httr::POST()` and `httr::GET()`, but the SBOE portal uses
a confusing back-end HTML structure we can't easily replicate. Instead, we have
downloaded the files manually.

```{r raw_info}
raw_dir <- dir_create(here("il", "expends", "data", "raw"))
raw_info <- as_tibble(dir_info(raw_dir))
nrow(raw_info)
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, path.abbrev))
```

## Fix

To improve the reading of these files, we can make new versions of the data
by reading the lines of each and looking for any double-quote character (`"`)
not next to a tab delimiter. These quotation marks are used _inside_ a cell, 
often around nicknames (e.g., "Committee to Elect JP "Rick" Carney").

Use of double quotations in this way can cause parsing errors, so we can
identify them with a regular expression and replace them with a single quote
(`'`).

```{r fix_write}
fix_dir <- dir_create(path(dirname(raw_dir), "fix"))
pb <- txtProgressBar(max = nrow(raw_info), style = 3)
for (i in seq_along(raw_info$path)) {
  read_file(f <- raw_info$path[i]) %>% 
    # any quote not next to tab or end
    str_replace_all("(?<!\t|^|\r\n)\"(?!\t|$|\r\n)", "\'") %>% 
    write_file(path(fix_dir, basename(f)))
  setTxtProgressBar(pb, i)
}
fix_info <- dir_info(fix_dir)
```

```{r fix_check}
nrow(fix_info) == nrow(raw_info)
sum(fix_info$size) == sum(raw_info$size)
```

## Read

All the `r nrow(raw_info)` fixed files can be read into a single data frame 
using `vroom::vroom()`

```{r raw_read}
ile <- vroom(
  file = fix_info$path,
  id = "file",
  delim = "\t",
  escape_backslash = FALSE,
  escape_double = FALSE,
  na = c("", " ", "       "),
  num_threads = 1,
  col_types = cols(
    .default = col_character(),
    ExpndDate = col_date("%m/%d/%Y"),
    Amount  = col_double(),
    RptPdBegDate = col_date("%m/%d/%Y"),
    RptPdEndDate = col_date("%m/%d/%Y"),
    RptRcvdDate = col_date("%m/%d/%Y")
  )
)
```

If all these files were correctly read into a single data frame, we should be
able count the distinct values fo a discrete variable like `doc_name`.

```{r raw_check}
comma(nrow(ile))
count(ile, DocName)
```

```{r raw_rename}
old_names <- names(ile)
ile <- ile %>% 
  clean_names("snake") %>% 
  remove_empty("rows") %>% 
  remove_empty("cols") %>% 
  remove_constant() %>% 
  rename(
    comm_id = committee_id,
    vendor = received_by,
    date = expnd_date,
    type = expended_by_type,
    support = supporting_opposing,
    candidate = candidate_name,
    report_begin = rpt_pd_beg_date,
    report_end = rpt_pd_end_date,
    committee = expended_by
  )
```

```{r echo=FALSE}
ile <- mutate(ile, across(file, basename))
```

## Explore

There are `r comma(nrow(ile))` records with `r comma(ncol(ile))` variables.

```{r glimpse}
glimpse(ile)
tail(ile)
```

### Missing

Variables differ in the degree of values they are missing.

```{r na_count}
col_stats(ile, count_na)
```

We can flag any record missing a key variable used to identify a transaction.

```{r na_flag}
ile <- ile %>% flag_na(date, vendor, amount, committee)
sum(ile$na_flag)
```

If a record is _any_ of these values, they are missing all of them.

```{r na_view}
ile %>% 
  filter(na_flag) %>% 
  select(date, vendor, amount, committee) %>% 
  distinct()
```

The only variable present for any of these records is the `file` one we added.

```{r na_prop}
ile %>% 
  filter(na_flag) %>% 
  col_stats(count_na)
```

We can simply remove these rows because they hold no data.

```{r na_remove}
file_col <- ile$file[!ile$na_flag]
ile <- ile %>% 
  select(-file, -na_flag) %>% 
  remove_empty("rows") %>% 
  mutate(file = file_col, .before = 1)
```

Now, no records are missing any of these key variables.

```{r na_recheck}
ile %>% 
  select(date, vendor, amount, committee) %>% 
  col_stats(count_na)
```

### Duplicates

```{r dupe_write}
dupe_file <- file_temp(pattern = "dupes")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  n <- 1e5 # rows per chunk
  nr <- nrow(ile)
  # split file into chunks
  ils <- split(ile, rep(1:ceiling(nr/n), each = n, length.out = nr))
  for (i in seq_along(ils)) {
    # check dupes from both ends
    d1 <- duplicated(ils[[i]], fromLast = FALSE)
    d2 <- duplicated(ils[[i]], fromLast = TRUE)
    # append to disk
    write_lines(d1 | d2, dupe_file, append = TRUE)
    rm(d1, d2)
    ils[[i]] <- NA
    message(percent(i/length(ils)))
    Sys.sleep(10)
    flush_memory()
  }
  rm(ils)
}
```

```{r dupe_read}
ile <- mutate(ile, dupe_flag = as.logical(read_lines(dupe_file)))
percent(mean(ile$dupe_flag), 0.1)
```

```{r dupe_view}
ile %>% 
  filter(dupe_flag) %>% 
  select(date, vendor, amount, committee)
```

### Categorical

```{r distinct_count}
col_stats(ile, n_distinct)
```

```{r distinct_plots}
explore_plot(ile, type)
explore_plot(ile, support)
explore_plot(ile, doc_name)
explore_plot(ile, election)
```

### Amounts

The max `amount` value is `r dollar(max(ile$amount))` and 
`r percent(mean(ile$amount <= 0), 0.1)` of values are less than or equal to
$0.

```{r ammount_summary}
noquote(map_chr(summary(ile$amount), dollar))
mean(ile$amount <= 0)
```

Here is the record with the largest `amount` value.

```{r amount_max}
glimpse(ile[which.max(ile$amount), ])
```

```{r hist_amount, echo=FALSE}
ile %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Illinois Expenditures Amount Distribution",
    caption = "Source: IL SBOE",
    x = "Amount",
    y = "Count"
  )
```

```{r amount_violin, echo=FALSE}
ile %>%
  filter(amount > 1) %>% 
  ggplot(
    mapping = aes(
      y = amount,
      x = reorder(
        type, 
        X = amount, 
        FUN = median, 
        na.rm = TRUE
      )
    )
  ) +
  geom_violin(
    mapping = aes(fill = type),
    trim = TRUE,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "area",
    adjust = 2
  ) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Illinois Expenditures Amount Distribution",
    caption = "Source: IL SBOE",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
ile <- mutate(ile, year = year(date))
```

Since we exported the records from the search portal based on year, there are
none outside the expected range.

```{r date_range}
min(ile$date)
sum(ile$year < 2000)
max(ile$date)
sum(ile$date > today())
```

```{r bar_year, echo=FALSE}
ile %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Illinois Expenditures by Year",
    caption = "Source: IL SBOE",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
ile <- ile %>% 
  unite(
    col = address_full,
    starts_with("address"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(
    address_norm = normal_address(
      address = address_full,
      abbs = usps_street,
      na_rep = TRUE
    )
  ) %>% 
  select(-address_full)
```

```{r address_view}
ile %>% 
  select(contains("address")) %>% 
  distinct() %>% 
  sample_n(10)
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
ile <- ile %>% 
  mutate(
    zip_norm = normal_zip(
      zip = zip,
      na_rep = TRUE
    )
  )
```

```{r zip_progress}
progress_table(
  ile$zip,
  ile$zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
ile <- ile %>% 
  mutate(
    state_norm = normal_state(
      state = state,
      abbreviate = TRUE,
      na_rep = TRUE,
      valid = valid_state
    )
  )
```

```{r state_view}
ile %>% 
  filter(state != state_norm) %>% 
  count(state, state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  ile$state,
  ile$state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
ile <- ile %>% 
  mutate(
    city_norm = normal_city(
      city = city, 
      abbs = usps_city,
      states = c("IL", "DC", "ILLINOIS"),
      na = invalid_city,
      na_rep = TRUE
    )
  )
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
ile <- ile %>% 
  rename(city_raw = city) %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(city_norm, city_match),
    match_dist = str_dist(city_norm, city_match),
    city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- ile %>% 
  mutate(
    city_refine = city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(city_refine != city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "city_refine" = "city",
      "state_norm" = "state",
      "zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    state_norm, 
    zip_norm, 
    city_swap, 
    city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
ile <- ile %>% 
  left_join(good_refine) %>% 
  mutate(city_refine = coalesce(city_refine, city_swap))
```

#### Progress

```{r city_progress, echo=FALSE}
many_city <- c(valid_city, extra_city)
progress <- progress_table(
  str_to_upper(ile$city_raw),
  ile$city_norm,
  ile$city_swap,
  ile$city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(ile$city_raw, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Illinois City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Illinois City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
ile <- ile %>% 
  select(
    -city_norm,
    -city_swap,
    city_clean = city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(ile, 100))
```

1. There are `r comma(nrow(ile))` records in the database.
1. `r percent(mean(ile$dupe_flag), 0.01)` of records are duplicates.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ile$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("il", "expends", "data", "clean"))
clean_path <- path(clean_dir, "il_expends_clean.csv")
write_csv(ile, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(ile)),
  old = c(md_code(old_names), rep("", ncol(ile) - length(old_names))),
  type = md_code(map_chr(ile, typeof)),
  def = c(
    "Raw file (period) containing the record",
    "Unique expending committee ID",
    "Receiving vendor name",
    "Date expenditure was made",
    "Expenditure amount or correction",
    "Vendor street address",
    "Vendor secondary address",
    "Vendor city name",
    "Vendor state abbreviation",
    "Vendor ZIP+4 code",
    "Vendor type",
    "Expenditure purpose",
    "For whom the expenditure was made",
    "Expending candidate name",
    "Office and district sought",
    "Whether committee suports or opposes",
    "Document name expenditure reported on",
    "Election during which expenditure was made",
    "Reporting period begining date",
    "Reporting period end date",
    "Report received date",
    "Spending committee name",
    "Flag for completely duplicated record",
    "Calendar year of contribution date",
    "Normalized combined street address",
    "Normalized 5-digit ZIP code",
    "Normalized 2-digit state abbreviation",
    "Normalized city name"
  )
)
```

```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Original Name", "Type", "Definition")
))
```
