---
title: "Arizona Expenditures"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("az", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load_gh("kiernann/mdbr")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  magrittr, # pipe operators
  janitor, # clean data frames
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  rvest, # html scraping
  glue, # combine strings
  here, # relative paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Expenditures data for the state of Arizona was retrieved under open records law 
Arizona Rev. Stat. Ann. Secs. 39-121 to 39-122. After a $25 fee was paid to the
state by the Investigative Reporting Workshop we were given access to the
state's file sharing website from which a Microsoft Access Database was
downloaded.

```{r raw_dir}
raw_dir <- dir_create(here("az", "expends", "data", "raw"))
mdb_file <- path(raw_dir, "CFS_Export.mdb")
```

```{r raw_download}
file_size(mdb_file)
```

## Read

The Microsoft Access database provided by the state cannot easily be read into
R like one might with a text file. We can use the [mdbr] package to access the
open source [MDB Tools][mdbtools] command line utility. This package can list
the tables in a database and read them as data frames.

[mdbtools]: https://github.com/brianb/mdbtools
[mdbr]: https://github.com/kiernann/mdbr

To use the mdbr package, MDB Tools must first be installed. The `mdbtools`
utility can be installed from the command line on Debian systems:

```bash
# https://github.com/brianb/mdbtools
sudo apt install mdbtools
```

We can use the `mdb_tableS()` function to list all the tables in the database.

```{r mdb_tables}
az_tables <- mdb_tables(mdb_file)
```

|Table                  |Description                                     |Rows |
|:----------------------|:-----------------------------------------------|:----|
|`BallotMeasures`       |Information on ballot measures voted on         |50   |
|`Categories`           |Category names with codes                       |71   |
|`Counties`             |County names with codes                         |15   |
|`Cycles`               |Election cycle start and dates with codes       |15   |
|`EntityTypes`          |Filing entity type names with codes             |43   |
|`IncomeExpenseNeutral` |Filing type with codes                          |3    |
|`Names`                |The names and addresses of recipients      |1,197,512 |
|`Offices`              |Office names sought with codes                  |70   |
|`Parties`              |Political party names with codes                |6    |
|`ReportNames`          |Filing report type, usual a time of year        |237  |
|`Reports`              |All of the reports filed by committees       |143,497 |
|`ReportTypes`          |Report types with codes                         |8    |
|`TransactionTypes`     |Transaction type name and `IncomeExpenseNeutral` |159 |
|`Committees`           |Names, IDs, and addresses of filing committees |4,556 |
|`Transactions`         |All expenditures with amounts and dates    |6,241,937 |

These tables can be read with `read_mdb()`, which uses `readr::read_csv()`. We
will read every table into a single list of multiple data frames.

```{r mdb_read}
a <- rep(list(NA), length(az_tables))
names(a) <- make_clean_names(az_tables)
for (i in seq_along(az_tables)) {
  a[[i]] <- read_mdb(mdb_file, az_tables[i])
}
```

## Join

These many data frames can then be joined together to create a single collection
of transactions between parties.

```{r echo=FALSE}
az_rename <- function(string) {
  string %>%
    make_clean_names("snake") %>% 
    str_replace("transaction_", "tx_") %>% 
    str_replace("modifies_", "mod_") %>% 
    str_replace("address", "addr") %>% 
    str_replace("committee_", "com_") %>% 
    str_replace("financial_institution", "fin_inst") %>% 
    str_replace("candidate_", "cand_") %>% 
    str_replace("organization_", "org_") %>% 
    str_replace("termination_", "term_") %>% 
    str_replace("physical_", "com_") %>% 
    str_replace("_relationship", "_rel") %>% 
    str_replace("_relationship", "_rel") %>% 
    str_replace("cand_other_party_name", "cand_other") %>% 
    str_replace("benefits_ballot_measure", "ben_ballot") %>% 
    str_remove("_name")
}
```

```{r}
a$committees <- select(
  .data = a$committees, CommitteeID, NameID, 
  starts_with("Candidate"), -CandidateOtherPartyName
)
```

```{r}
a$names <- select(a$names, -NameGroupID, -CountyID, Zip = ZipCode)
```

```{r raw_join}
aze <- a$transactions %>% # 6,241,937
  # Transaction type, category info -------------------------------------------
  left_join(a$transaction_types, by = "TransactionTypeID") %>% 
  select(-TransactionTypeID) %>% 
  left_join(a$income_expense_neutral, by = "IncomeExpenseNeutralID") %>% 
  select(-IncomeExpenseNeutralID, TransactionMethod = IncomeExpenseNeutral) %>% 
  left_join(a$categories, by = "CategoryID") %>% 
  select(-CategoryID, -SubjectCommitteeID) %>%
  # Spending COMMITTEE info ---------------------------------------------------
  rename(PayeeID = NameID) %>% # expenditure recipient
  relocate(CommitteeID, .after = last_col()) %>% 
  left_join(a$committees, by = "CommitteeID") %>% 
  # COMMITTEE party, county, office, cycle ------------------------------------
  # join by ID, remove ID and rename columns
  left_join(a$parties, by = c("CandidatePartyID" = "PartyID")) %>% 
  select(-CandidatePartyID, CandidateParty = PartyName) %>% 
  left_join(a$offices, by = c("CandidateOfficeID" = "OfficeID")) %>% 
  select(-CandidateOfficeID, CandidateOffice = OfficeName) %>% 
  left_join(a$counties, by = c("CandidateCountyID" = "CountyID")) %>% 
  select(-CandidateCountyID, CandidateCounty = CountyName) %>%
  left_join(select(a$cycles, 1:2), by = c("CandidateCycleID" = "CycleID")) %>% 
  select(-CandidateCycleID, CandidateCycle = CycleName) %>%
  # CANDIDATE name and info ---------------------------------------------------
  left_join(a$names, by = "NameID") %>% 
  rename(CandidateID = NameID) %>% 
  left_join(a$entity_types, by = "EntityTypeID") %>% 
  select(-EntityTypeID, CandidateType = EntityTypeName) %>% 
  rename_with(~paste0("Candidate", .), .cols = 21:last_col(1)) %>% 
  # PAYEE name and info -------------------------------------------------------
  relocate(PayeeID, .after = last_col()) %>% 
  left_join(a$names, by = c("PayeeID" = "NameID")) %>% 
  left_join(a$entity_types, by = "EntityTypeID") %>% 
  select(-EntityTypeID, PayeeType = EntityTypeName) %>% 
  rename_with(~paste0("Payee", .), .cols = 33:last_col(1)) %>% 
  select(-14) %>% 
  rename_with(.fn = az_rename) %>% 
  rename(date = tx_date) %>% 
  mutate(across(ends_with("date"), as_date))
```

```{r}
rm(a); flush_memory()
```

## Explore

```{r glimpse}
glimpse(aze)
tail(aze)
```

### Missing

```{r na_count}
col_stats(aze, count_na)
```

```{r na_flag}
key_vars <- c("date", "payee_last", "amount", "cand_last")
aze <- flag_na(aze, all_of(key_vars))
sum(aze$na_flag)
```

```{r na_view}
aze %>% 
  filter(na_flag) %>% 
  select(all_of(key_vars)) %>% 
  sample_frac()
```

### Duplicates

```{r dupe_write}
dupe_file <- here("az", "expends", "dupes.txt")
if (!file_exists(dupe_file)) {
  file_create(dupe_file)
  # split file into monthly chunks
  azs <- aze %>% 
    mutate(month = month(date), year = year(date)) %>% 
    group_split(month, year, .keep = FALSE)
  pb <- txtProgressBar(max = length(azs), style = 3)
  for (i in seq_along(azs)) {
    # check dupes from both ends
    d1 <- duplicated(azs[[i]], fromLast = FALSE)
    d2 <- duplicated(azs[[i]], fromLast = TRUE)
    # append to disk
    write_lines(d1 | d2, dupe_file, append = TRUE)
    rm(d1, d2)
    azs[[i]] <- NA
    Sys.sleep(1); flush_memory(1)
    setTxtProgressBar(pb, value = i)
  }
  rm(azs)
}
```

```{r dupe_read}
aze <- mutate(aze, dupe_flag = as.logical(read_lines(dupe_file)))
if (sum(aze$dupe_flag) == 0) {
  file_delete(dupe_file)
  aze <- select(aze, -dupe_flag)
}
```

### Categorical

```{r distinct_count}
col_stats(aze, n_distinct)
```

```{r distinct_plots, echo=FALSE}
explore_plot(aze, account_type)
explore_plot(aze, tx_method)
explore_plot(aze, category)
explore_plot(aze, cand_party)
explore_plot(aze, cand_office) + scale_x_truncate()
explore_plot(aze, cand_county)
explore_plot(aze, cand_type) + scale_x_truncate()
explore_plot(aze, payee_type) + scale_x_truncate()
```

### Amounts

```{r ammount_summary}
summary(aze$amount)
mean(aze$amount <= 0)
```

```{r hist_amount, echo=FALSE}
aze %>%
  filter(amount >= 1) %>% 
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Arizona Contracts Amount Distribution",
    caption = "Source:  Arizona Secretary of State",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `date` with `lubridate::year()`

```{r date_year}
aze <- mutate(aze, year = year(date))
```

```{r date_range}
min(aze$date)
sum(aze$year < 2000)
max(aze$date)
sum(aze$date > today())
```

```{r bar_year, echo=FALSE}
aze %>% 
  count(year) %>% 
  mutate(even = is_even(year)) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona Expenditures by Year",
    caption = "Source:  Arizona Secretary of State",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

To improve the searchability of the database, we will perform some consistent,
confident string normalization. For geographic variables like city names and
ZIP codes, the corresponding `campfin::normal_*()` functions are tailor made to 
facilitate this process.

### Address

For the street `addresss` variable, the `campfin::normal_address()` function
will force consistence case, remove punctuation, and abbreviate official 
USPS suffixes.

```{r address_norm}
aze <- aze %>% 
  unite(
    col = payee_addr,
    starts_with("payee_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  unite(
    col = cand_addr,
    starts_with("cand_addr"),
    sep = " ",
    remove = FALSE,
    na.rm = TRUE
  ) %>% 
  mutate(across(
    .cols = ends_with("_addr"),
    .fns = list(norm = normal_address),
    abbs = usps_street,
    na_rep = TRUE
  )) %>% 
  select(-ends_with("_addr"))
```

```{r address_view}
aze %>% 
  sample_n(10) %>% 
  select(contains("_addr"))
```

### ZIP

For ZIP codes, the `campfin::normal_zip()` function will attempt to create
valid _five_ digit codes by removing the ZIP+4 suffix and returning leading
zeroes dropped by other programs like Microsoft Excel.

```{r zip_norm}
aze <- mutate(
  .data = aze, across(
    .cols = ends_with("_zip"),
    .fns = list(norm = normal_zip),
    na_rep = TRUE
  )
)
```

```{r zip_progress}
progress_table(
  aze$payee_zip,
  aze$payee_zip_norm,
  compare = valid_zip
)
```

### State

Valid two digit state abbreviations can be made using the 
`campfin::normal_state()` function.

```{r state_norm}
aze <- mutate(
  .data = aze, across(
    .cols = ends_with("_state"),
    .fns = list(norm = normal_state),
    abbreviate = TRUE,
    na_rep = TRUE,
  )
)
```

```{r state_view}
aze %>% 
  filter(payee_state != payee_state_norm) %>% 
  count(payee_state, payee_state_norm, sort = TRUE)
```

```{r state_progress}
progress_table(
  aze$payee_state,
  aze$payee_state_norm,
  compare = valid_state
)
```

### City

Cities are the most difficult geographic variable to normalize, simply due to
the wide variety of valid cities and formats.

#### Normal

The `campfin::normal_city()` function is a good start, again converting case,
removing punctuation, but _expanding_ USPS abbreviations. We can also remove
`invalid_city` values.

```{r city_norm}
aze <- mutate(
  .data = aze, across(
    .cols = ends_with("_city"),
    .fns = list(norm = normal_city),
    abbs = usps_city,
      states = c("AZ", "DC", "ARIZONA"),
      na = invalid_city,
      na_rep = TRUE
  )
)
```

#### Swap

We can further improve normalization by comparing our normalized value
against the _expected_ value for that record's state abbreviation and ZIP code.
If the normalized value is either an abbreviation for or very similar to the
expected value, we can confidently swap those two.

```{r city_swap}
aze <- aze %>% 
  left_join(
    y = zipcodes,
    by = c(
      "payee_state_norm" = "state",
      "payee_zip_norm" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_abb = is_abbrev(payee_city_norm, city_match),
    match_dist = str_dist(payee_city_norm, city_match),
    payee_city_swap = if_else(
      condition = !is.na(match_dist) & (match_abb | match_dist == 1),
      true = city_match,
      false = payee_city_norm
    )
  ) %>% 
  select(
    -city_match,
    -match_dist,
    -match_abb
  )
```

#### Refine

The [OpenRefine][or] algorithms can be used to group similar strings and replace
the less common versions with their most common counterpart. This can greatly
reduce inconsistency, but with low confidence; we will only keep any refined
strings that have a valid city/state/zip combination.

[or]: https://openrefine.org/

```{r city_refine}
good_refine <- aze %>% 
  mutate(
    payee_city_refine = payee_city_swap %>% 
      key_collision_merge() %>% 
      n_gram_merge(numgram = 1)
  ) %>% 
  filter(payee_city_refine != payee_city_swap) %>% 
  inner_join(
    y = zipcodes,
    by = c(
      "payee_city_refine" = "city",
      "payee_state_norm" = "state",
      "payee_zip_norm" = "zip"
    )
  )
```

```{r city_count, echo=FALSE}
good_refine %>%
  count(
    payee_state_norm, 
    payee_zip_norm, 
    payee_city_swap, 
    payee_city_refine,
    sort = TRUE
  )
```

Then we can join the refined values back to the database.

```{r city_join}
aze <- aze %>% 
  left_join(good_refine, by = names(aze)) %>% 
  mutate(payee_city_refine = coalesce(payee_city_refine, payee_city_swap))
```

#### Progress

```{r city_recheck}
many_city <- c(valid_city, extra_city)
aze %>% 
  filter(payee_city_refine %out% many_city) %>% 
  count(payee_city_refine, payee_state_norm, payee_zip_norm, sort = TRUE) %>% 
  drop_na()
```

```{r city_remove}
aze$payee_city_refine <- na_in(
  x = aze$payee_city_refine, 
  y = c("<MISSING VALUE>", "AZ")
)
```

```{r city_increase}
many_city <- c(
  many_city,
  "FARMINGTON HILLS",
  "SHELBY TOWNSHIP",
  "SADDLEBROOKE",
  "DESERT HILLS"
)
```

```{r city_progress, echo=FALSE}
progress <- progress_table(
  str_to_upper(aze$payee_city),
  aze$payee_city_norm,
  aze$payee_city_swap,
  aze$payee_city_refine,
  compare = many_city
) %>% mutate(stage = as_factor(stage))
kable(progress, digits = 3)
```

You can see how the percentage of valid values increased with each stage.

```{r bar_progress, echo=FALSE}
raw_in <- percent(prop_in(aze$payee_city, valid_city))
progress %>% 
  ggplot(aes(x = stage, y = prop_in)) +
  geom_hline(yintercept = 0.99) +
  geom_col(fill = dark2["purple"]) +
  coord_cartesian(ylim = c(0.75, 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = glue("Raw at {raw_in} before conversion to uppercase"),
    x = "Stage",
    y = "Percent Valid"
  )
```

More importantly, the number of distinct values decreased each stage. We were
able to confidently change many distinct invalid values to their valid
equivalent.

```{r bar_distinct, echo=FALSE}
progress %>% 
  select(
    stage, 
    all = n_distinct,
    bad = n_diff
  ) %>% 
  mutate(good = all - bad) %>% 
  pivot_longer(c("good", "bad")) %>% 
  mutate(name = name == "good") %>% 
  ggplot(aes(x = stage, y = value)) +
  geom_col(aes(fill = name)) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "bottom") +
  labs(
    title = "Arizona City Normalization Progress",
    subtitle = "Distinct values, valid and invalid",
    x = "Stage",
    y = "Distinct Values",
    fill = "Valid"
  )
```

## Conclude

Before exporting, we can remove the intermediary normalization columns and
rename all added variables with the `_clean` suffix.

```{r clean_select}
aze <- aze %>% 
  select(
    -payee_city_norm,
    -payee_city_swap,
    payee_city_clean = payee_city_refine
  ) %>% 
  rename_all(~str_replace(., "_norm", "_clean")) %>% 
  rename_all(~str_remove(., "_raw"))
```

```{r clean_glimpse}
glimpse(sample_n(aze, 50))
```

1. There are `r comma(nrow(aze))` records in the database.
1. There are `r comma(sum(aze$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(aze$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("az", "expends", "data", "clean"))
clean_path <- path(clean_dir, "az_expends_clean.csv")
write_csv(aze, clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Upload

Using the [duckr] R package, we can wrap around the [duck] command line tool to
upload the file to the IRW server.

[duckr]: https://github.com/kiernann/duckr
[duck]: https://duck.sh/

```{r clean_upload, eval=FALSE}
# remotes::install_github("kiernann/duckr")
s3_dir <- "s3:/publicaccountability/csv/"
s3_path <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_path)
}
```
