---
title: "Ohio Contracts"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("oh", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  readxl, # read excel file
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::i_am("oh/contracts/docs/oh_contracts_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data


```{r raw_dir}
raw_dir <- dir_create(here("oh", "contracts", "data", "raw"))
raw_xls <- path(raw_dir, "Contracts_1.1.2015_-_Present.xlsx")
```

## Read

```{r}
raw_sheets <- set_names(excel_sheets(raw_xls))[1:6]
```

```{r}
ohc <- map(
  .x = raw_sheets,
  .f = read_excel,
  path = raw_xls,
  range = "A5:E10000"
)
```

```{r}
ohc <- ohc %>% 
  bind_rows(.id = "src_sheet") %>% 
  filter(!is.na(`Vendor Name`)) %>% 
  clean_names(case = "snake")
```

```{r}
num_dates <- str_which(ohc$po_date, "^\\d+$")
ohc$po_date[num_dates] <- format(
  x = excel_numeric_to_date(as.double(ohc$po_date[num_dates])),
  format = "%b %d, %Y"
)
```

```{r}
ohc$po_date <- parse_date(
  x = ohc$po_date,
  format = "%b %d, %Y"
)
```

## Explore

There are `r comma(nrow(ohc))` rows of `r ncol(ohc)` columns. Each record
represents a single contract between a state agency and an outside vendor.

```{r glimpse}
glimpse(ohc)
tail(ohc)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(ohc, count_na)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
ohc <- flag_dupes(ohc, everything())
sum(ohc$dupe_flag)
```

```{r dupe_view}
ohc %>% 
  filter(dupe_flag) %>% 
  select(vendor_name, po_date, created_amt, po_description)
```

### Amounts

```{r amount_summary}
summary(ohc$created_amt)
mean(ohc$created_amt <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(ohc[c(which.max(ohc$created_amt), which.min(ohc$created_amt)), ])
```

```{r hist_amount, echo=FALSE}
ohc %>%
  filter(created_amt >= 1) %>% 
  ggplot(aes(created_amt)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Ohio Contracts Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

We can add the calendar year from `po_date` with `lubridate::year()`

```{r date_year}
ohc <- mutate(ohc, po_year = year(po_date))
```

```{r date_range}
min(ohc$po_date)
max(ohc$po_date)
```

```{r bar_year, echo=FALSE}
ohc %>% 
  count(po_year) %>% 
  mutate(even = is_even(po_year)) %>% 
  ggplot(aes(x = po_year, y = n)) +
  geom_col(aes(fill = even)) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "Ohio Contracts by Year",
    caption = "Source: {source}",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Wrangle

We can manually add the state and agency spending the money.

```{r agency_add}
ohc <- mutate(ohc, state = "OH", agency = "OHIO TREASURER OF STATE")
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(ohc, 50))
```

1. There are `r comma(nrow(ohc))` records in the database.
1. There are `r comma(sum(ohc$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(ohc$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("oh", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "oh_contracts_clean.csv")
write_csv(ohc, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
