---
title: "Wisconsin Lobbying Data Diary"
author: "Yanqi Xu"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wi", "data", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:

1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction

## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This package contains
functions custom made to help facilitate the processing of campaign finance data.

```{r load_packages, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  patchwork, # position plots
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  readxl, # read data frames from xls
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.

The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[01]: https://github.com/irworkshop/accountability_datacleaning "TAP repo"
[02]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects "Rproj"

## Data
The Accountability Project obtained the Wisconsin lobbying expenditures from the Wisconsin Ethics Commission from 2013 to 2019 through an open records request. The dataset is as current as 
Nov 6, 2019. 

### About

### Variables

For the `reporting_period` column; periods are as follows:

P1: January-June odd numbered year of session

P2: July-December odd numbered year of session

P3: January-June even numbered year of session

P4: July-December even numbered year of session

 

For the `session` column; sessions are as follows:

2013REG: 2013-2014

2015REG: 2015-2016

2017REG: 2017-2018

2019REG: 2019-2020

## Import

```{r create_raw_dir}
reg_dir <- here("wi", "lobbyists", "data", "raw", "registration")
exp_dir <- here("wi", "lobbyists", "data", "raw", "expends")
dir_create(c(reg_dir, exp_dir))
```

### Read
There are two excel spreadsheets for Wisconsin Lobbying Expenditures records obtained by the Accountability Project, namely, payments made directly to lobbyists, and payments made to lobbying firms.
```{r read_raw}
wi_lobbyist <- read_xlsx(glue("{exp_dir}/WILobbyingExpense.xlsx"), 
                         sheet = 1, col_types = "text") %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper)
# Change "NULL" to NAs.

wi_lobbyist_13 <- read_xlsx(glue("{exp_dir}/wi_lobbying_expenses_2013.xlsx"), 
                         sheet = 1, col_types = "text") %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper)

wi_lobbyist <- wi_lobbyist %>% bind_rows(wi_lobbyist_13)

wi_lobbyist[,6:10] <- wi_lobbyist[,6:10] %>% na_if("NULL") %>% map_dfc(as.numeric)

wi_org <- read_xlsx(glue("{exp_dir}/WILobbyingExpense.xlsx"), 
                         sheet = 2, col_types = "text") %>% 
  clean_names() %>% 
  mutate_if(is_character, str_to_upper)

wi_org[,5:7] <- wi_org[,5:7] %>% map_dfc(as.numeric)

reg_dir <- here("wi", "lobbyists", "data", "processed", "registration")
wi_lobby_reg <- read_csv(glue("{reg_dir}/wi_lobbyists_reg.csv"))

```

## Explore

```{r glimpse}
head(wi_org)
tail(wi_org)
glimpse(sample_frac(wi_org))

head(wi_lobbyist)
tail(wi_lobbyist)
glimpse(sample_frac(wi_lobbyist))
```

### Missing

```{r glimpse_na}
col_stats(wi_lobbyist, count_na)
col_stats(wi_org, count_na)
```

```{r flag_na}
wi_org <- wi_org %>% flag_na()
wi_lobbyist <- wi_lobbyist %>% flag_na()
```

### Duplicates

```{r flag_dupes}
wi_lobbyist <- wi_lobbyist %>% flag_dupes(dplyr::everything())
wi_org <- wi_org %>%  flag_dupes(dplyr::everything())
```

### Categorical

```{r glimpse_distinct}
col_stats(wi_lobbyist, n_distinct)
col_stats(wi_org, n_distinct)
```

```{r summary_amount}

wi_lobbyist <- wi_lobbyist %>% 
  mutate(attributable_expenses = payments_attributable_to_lobbying + travel_reimbursement + other_reimbursement,
         inhouse_expenses = in_house_lobbyist_compensation + in_house_lobbyist_reimbursement)

wi_org <- wi_org %>% 
  mutate(org_expenses = payments_attributable_to_lobbying + travel_reimbursement + other_reimbursement)

summary(wi_lobbyist$inhouse_expenses)
summary(wi_lobbyist$attributable_expenses)
```

#### Year
```{r mutate year}
year_calculator <- function(rep_per){
  year_result = ifelse(test = rep_per > 2,yes = 1, no = 0)
  return(year_result)
}

wi_org <- wi_org %>% 
  mutate(year = as.numeric(str_sub(session, start = 1L, end = 4L)) + year_calculator(as.numeric(reporting_period %>% str_remove("^P"))))


wi_lobbyist <- wi_lobbyist %>% 
    mutate(year = as.numeric(str_sub(session, start = 1L, end = 4L)) + year_calculator(as.numeric(reporting_period %>% str_remove("^P"))))

```

```{r plot year, echo=FALSE}
bar_attr <- wi_lobbyist %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(!is.na(attributable_expenses) & attributable_expenses != 0) %>% 
  count(year, sort = T) %>% 
  mutate(on = is_even(year),) %>%
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = on)) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_y_continuous(labels = comma) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Count by Year",
    subtitle = "Payments to lobbyists attributable to lobbying",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Election Year",
    x = "Year",
    y = "Distinct Expenditures"
  ) +
  theme(legend.position = "bottom")

bar_inhouse <- wi_lobbyist %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(!is.na(inhouse_expenses) & inhouse_expenses != 0) %>% 
  count(year, sort = T) %>% 
  mutate(on = is_even(year),) %>%
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = on)) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_y_continuous(labels = comma) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Count by Year",
    subtitle = "Payments to in-house lobbyists",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Election Year",
    x = "Year",
    y = "Distinct Expenditures"
  ) +
  theme(legend.position = "bottom")

bar_org <- wi_org %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(!is.na(org_expenses) & org_expenses != 0) %>% 
  count(year, sort = T) %>% 
  mutate(on = is_even(year),) %>%
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = on)) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_y_continuous(labels = comma) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Count by Year",
    subtitle = "Payments to lobbying organizations",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Election Year",
    x = "Year",
    y = "Distinct Expenditures"
  ) +
  theme(legend.position = "bottom")

bar_attr + bar_org + bar_inhouse + plot_layout(ncol = 2)
```

```{r bar_med_plot, echo=FALSE}
bar_med_attr <- wi_lobbyist %>% 
  mutate(on = is_even(year)) %>%
  group_by(on, year) %>% 
  summarize(median = median(attributable_expenses[attributable_expenses!=0], na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = median)) +
  geom_col(aes(fill = on)) +
  scale_y_continuous(labels = dollar) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures by Year",
    subtitle = "How much is spent on lobbying activities?",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Year",
    x = "Election Year",
    y = "Median Amount"
  ) +
  theme(legend.position = "bottom")

bar_med_org <- wi_org %>% 
  mutate(on = is_even(year)) %>%
  group_by(on, year) %>% 
  summarize(median = median(org_expenses[org_expenses!=0], na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = median)) +
  geom_col(aes(fill = on)) +
  scale_y_continuous(labels = dollar) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures by Year",
    subtitle = "How much is paid to lobbying organizations?",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Year",
    x = "Election Year",
    y = "Median Amount"
  ) +
  theme(legend.position = "bottom")

bar_med_inhouse <- wi_lobbyist %>% 
  mutate(on = is_even(year)) %>%
  group_by(on, year) %>% 
  summarize(median = median(inhouse_expenses[inhouse_expenses!=0], na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = median)) +
  geom_col(aes(fill = on)) +
  scale_y_continuous(labels = dollar) +
  scale_x_continuous(breaks = 2013:2019) +
  scale_fill_brewer(
    type = "qual",
    palette = "Dark2"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures by Year",
    subtitle = "How much is paid to compensate in-house lobbyists?",
    caption = "Source: Wisconsin Ethics Commission",
    fill = "Year",
    x = "Election Year",
    y = "Median Amount"
  ) +
  theme(legend.position = "bottom")

bar_med_attr + bar_med_org + bar_med_inhouse + plot_layout(nrow = 2)
```

### Continuous

#### Amounts
We'll examine expenses attributable to lobbying paid to lobbyists, expenses attributable to lobbying paid to organizations and expenses paid to in-house lobbyists individually.

```{r amount_histogram, echo=FALSE}
hist_attr <- wi_lobbyist %>%
  ggplot(aes(attributable_expenses)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  geom_vline(xintercept = median(wi_lobbyist$attributable_expenses[wi_lobbyist$attributable_expenses!= 0], na.rm = TRUE), linetype = 2) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Distribution",
    subtitle = "attributable expenses to lobbyists from 2013 to 2019",
    caption = "Source: Wisconsin Ethics Commission",
    x = "Amount",
    y = "Count"
  )

hist_org <- wi_org %>%
  ggplot(aes(org_expenses)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[1]) +
  geom_vline(xintercept = median(wi_org$org_expenses[wi_org$org_expenses!= 0], na.rm = TRUE), linetype = 2) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Distribution",
    subtitle = "attributable expenses to lobbying organizations from 2013 to 2019",
    caption = "Source: Wisconsin Ethics Commission",
    x = "Amount",
    y = "Count"
  )

hist_inhouse <- wi_lobbyist %>%
  ggplot(aes(inhouse_expenses)) +
  geom_histogram(fill = RColorBrewer::brewer.pal(3, "Dark2")[3]) +
  geom_vline(xintercept = median(wi_lobbyist$inhouse_expenses[wi_lobbyist$inhouse_expenses!= 0], na.rm = TRUE), linetype = 2) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Wisconsin Lobbying Expenditures Distribution",
    subtitle = "In-house expenses to lobbyists from 2013 to 2019",
    caption = "Source: Wisconsin Ethics Commission",
    x = "Amount",
    y = "Count"
  )


hist_attr + 
  hist_org +
  hist_inhouse + 
  plot_layout(nrow = 3)

```

## Wrangle
### Session ID
```{r session id}
wi_org <- wi_org %>% mutate(session = session %>% str_replace("REG", " REGULAR SESSION"))
wi_lobbyist <- wi_lobbyist %>% mutate(session = session %>% str_replace("REG", " REGULAR SESSION"))
```

## Join
Since neither of the expenditure tables has IDs or addresses to identify lobbyists, principals or lobbying organizations. We join them back to the registration dataset processed before.

```{r join reg}
wi_reg <- read_csv(file.path(here("wi","lobbyists","data","processed","registration"),"wi_lobbyists_reg.csv"))

wi_org <- wi_reg %>% select(principal, principal_mailing_address_norm, principal_mailing_city_clean, principal_mailing_state, principal_mailing_zip_norm, legislative_session, organization) %>% unique() %>% 
  right_join(wi_org, by = c("principal" = "principal_name", "organization" = "organization_name", "legislative_session" = "session")) 

wi_lobby_reg <- wi_reg %>% 
  select(lobbyist_first_name, lobbyist_last_name, lobbyist_address_norm, lobbyist_city_clean, lobbyist_state, lobbyist_zip_norm, principal, principal_mailing_address_norm, principal_mailing_city_clean, principal_mailing_state, principal_mailing_zip_norm, legislative_session, organization) %>% unique() %>% 
  right_join(wi_lobbyist, by = c("principal" = "principal_name", "legislative_session" = "session", "lobbyist_first_name", "lobbyist_last_name")) 
```


## Conclude

1. There are `nrow(wi_lobbyist)` records in the lobbyist database and `nrow(wi_org)` in the organization database.
1. There are `sum(wi_lobbyist$dupe_flag)` duplicate records in the database and `nrow(wi_org)` duplicates in the organization database.
1. The range and distribution of expenses-related fields seem reasonable.
1. There are `sum(wi_lobbyist$na_flag)` records missing either receipient or 
1. The 4-digit `year` variable has been created based on the `session` and `reporting_period` columns.

## Export

```{r create_proc_dir}
proc_dir <- here("wi", "type", "data", "processed","expends")
dir_create(proc_dir)
```

```{r write_clean}
wi_org %>% 
  select(
  ) %>% 
  write_csv(
    path = glue("{proc_dir}/wi_org_exp_clean.csv"),
    na = ""
  )

wi_lobbyist %>% 
  write_csv(
    path = glue("{proc_dir}/wi_lobbyists_exp_clean.csv"),
    na = ""
  )
```
