---
title: "Mississippi PPP Breakdown"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 4
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 12,
  fig.height = 6,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ms", "ppp", "docs"))
```

## Data

This is an analysis of Paycheck Protection Program loans in Mississippi.

On Dec. 1, 2020 the Small Business Administration was ordered by the court to
release more detailed data on loans made through the [Paycheck Protection
Program][ppp], a major facet of the 2020 CARES Act to provide stimulus funding
during the ongoing COVID-19 pandemic. This detailed release came after the SBA
initially refused to publish any database, then released only partial data on
the largest loan recipients. The full database now contains all recipient names,
addresses, and exact loan amounts.

[According to the SBA][facts], PPP loans are forgiven as long as:

1. The loan proceeds are used to cover payroll costs, and most mortgage
interest, rent, and utility costs over the 8 week period after the loan is made.
2. Employee and compensation levels are maintained.

[The SBA clarifies key aspects][aspects] of the PPP, namely how the money is
distributed and what information is actually contained in the data.

> #### PPP Is A Delegated Loan Making Process
> PPP loans are not made by SBA. PPP loans are made by lending institutions and
then guaranteed by SBA. Accordingly, borrowers apply to lenders and self-certify
that they are eligible for PPP loans. The self-certification includes a good
faith certification that the borrower has economic need requiring the loan and
that the borrower has applied the affiliation rules and is a small business. The
lender then reviews the borrower’s application, and if all the paperwork is in
order, approves the loan and submits it to SBA...

> #### Cancelled Loans Do Not Appear In The PPP Loan Data
The public PPP data includes only active loans. Loans that were cancelled for
any reason are not included in the public data release.

[facts]: https://home.treasury.gov/system/files/136/PPP%20Borrower%20Information%20Fact%20Sheet.pdf
[ppp]: https://www.sba.gov/funding-programs/loans/coronavirus-relief-options/paycheck-protection-program
[aspects]: https://www.sba.gov/sites/default/files/2020-12/PPP%20Loan%20Data%20-%20Key%20Aspects%2012-01-20%20-%20Final.pdf

## Preparation

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  patchwork, # combine plots
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  usmap, # plot us maps
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

```{r custom_funs, echo=FALSE}
top_bottom <- function(.data, nnn = 10) {
  nr = nrow(x = .data)
  top = seq(from = 1, to = nnn)
  bot = seq(from = nr - (nnn - 1), to = nr)
  mutate(
    .data = .data[c(top, bot), ], 
    .before = 1,
    sort = rep(c("head", "tail"), each = nnn)
  )
}
head_tail <- function(data, n = 10) {
  list(
    head = head(data, n = n), 
    tail = tail(data, n = n)
  )
}
str_empty_na <- function(x) {
  str_remove(str_replace_na(x), "NA")
}
solo_types <- c(
  "Sole Proprietorship",
  "Independent Contractors",
  "Self-Employed Individuals"
)
```

We read the normalized PPP data from the IRW S3 server.

```{r clean_find}
ppp_dir <- here("us", "covid", "ppp")
ppp_file <- path(ppp_dir, "data", "clean", "sba_ppp_full.csv")
```

```{r clean_down}
if (!file_exists(ppp_file)) {
  save_object(
    object = "csv/sba_ppp_full.csv",
    bucket = "publicaccountability",
    file = ppp_file
  )
}
```

```{r clean_read}
ppp <- vroom(
  file = ppp_file,
  col_types = cols(
    naics_code = col_character(),
    zip = col_character()
  )
)
```

```{r clean_rename, echo=FALSE}
ppp <- ppp %>% 
  select(-address, -city, -state) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip) %>% 
  rename_with(~str_remove(., "_clean")) %>% 
  rename(
    date = date_approved,
    amount = loan_amount,
    jobs = jobs_reported,
    race = race_ethnicity,
    business = business_name,
    type = business_type,
    naics = naics_code
  ) %>% 
  arrange(date)
```

We added additional variables to identify which records are in Mississippi.

```{r clean_state_id}
ppp <- ppp %>% 
  mutate(
    is_ms = !is.na(state) & state == "MS",
    ms_lab = ifelse(is_ms, "Miss.", "Other")
  )
```

And then created a separate table of just Mississippi loans.

```{r clean_filter}
ms_ppp <- filter(ppp, is_ms)
ms2x <- bind_rows(
  "United States" = ppp,
  "Mississippi" = ms_ppp,
  .id = "ms_lab2"
)
```

## Explore

Mississippi has `r comma(nrow(ms_ppp))` loans with `r ncol(ms_ppp)` columns. 
In total, `r dollar(sum(ms_ppp$amount))` was distributed to Mississippi
businesses.

Mississippi accounts for `r percent(nrow(ms_ppp)/nrow(ppp), 0.1)` of the total 
PPP loans, and `r percent(sum(ms_ppp$amount)/sum(ppp$amount), 0.1)` of 
the total amount disbursed. This is close to the 
`r percent(usa::facts$population[25]/sum(usa::facts$population), 0.1)` of the
U.S. population that lives in the state.

## Jobs

We examined the number of _jobs_ reportedly retained. These numbers were self
reported and a significant amount of applications reported zero jobs or omitted
a number altogether, although the percent of loans missing job saved was smaller
in Mississippi than the country at large.

```{r jobs_missing, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2) %>% 
  summarise(
    no_job = mean(is.na(jobs)),
    zero_jobs = mean(jobs == 0, na.rm = TRUE)
  )
```

These loans, with missing or invalid jobs reported, do not appear to be tied
to the time the loan was approved and probably isn't related to confusion.

```{r jobs_missing_time, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2, week = epiweek(date)) %>% 
  summarise(
    `No Jobs` = mean(is.na(jobs)),
    `Zero Jobs` = mean(jobs == 0, na.rm = TRUE)
  ) %>% 
  pivot_longer(
    cols = 3:4
  ) %>% 
  ggplot(aes(x = week, y = value)) +
  geom_col(aes(fill = name)) +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks = 14:32) +
  labs(
    title = "Loans Missing Reported Jobs Over Time",
    fill = "Type",
    x = "Week",
    y = "Proportion"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  facet_wrap(
    facets = ~ms_lab2,
    ncol = 1,
    scales = "free_y"
  )
```

```{r id_vars, echo=FALSE}
id_vars <- c(
  "date", "lender", "amount", "jobs", "business",
  "address", "city", "state", "zip"
)
```

The loans missing job data seem to be tied to the lending bank. In Mississippi,
**100%** of loans from Wells Fargo were missing job information while **0%** of
those from PriorityOne Bank were.

```{r jobs_bank_count, echo=FALSE}
ms_bank_jobs <- ms_ppp %>% 
  group_by(lender) %>% 
  summarise(
    n = n(),
    no_job = round(mean(is.na(jobs) | jobs == 0), 4)
  ) %>% 
  arrange(desc(n))

data_dir <- dir_create(here("ms", "ppp", "data"))
write_csv(
  x = ms_bank_jobs,
  file = path(data_dir, "ms_ppp_banks_nojobs.csv")
)
```

Here are the top ten banks with the greatest and least proportion of their loans
missing job information (sorted by most loans given).

```{r jobs_bank_view, echo=FALSE}
ms_bank_jobs <- ms_bank_jobs %>% 
  filter(n > 100) %>% 
  arrange(desc(no_job)) %>%
  mutate(no_job = percent(no_job, 1)) %>% 
  mutate(n = comma(n, 1)) %>% 
  top_bottom(nnn = 10) %>%
  group_split(sort) %>% 
  map(select, -1)

kable(
  x = ms_bank_jobs[1], 
  col.names = c("Lending Bank", "# Loans", "% w/out Jobs")
)
kable(
  x = ms_bank_jobs[2], 
  col.names = c("Lending Bank", "# Loans", "% w/out Jobs")
)
```

We ignored missing data to continue to breakdown the reported jobs.

Most loans did go to small businesses saving less than **25** jobs. This was
true in both Mississippi and the rest of the country.

```{r jobs_small}
ms2x %>% 
  group_by(ms_lab2) %>% 
  summarise(small_jobs = mean(jobs < 25, na.rm = TRUE))
```

```{r hist_jobs, echo=FALSE, fig.height=10}
ms2x %>%
  filter(jobs > 0, jobs <= 25) %>% 
  ggplot(aes(jobs)) +
  geom_histogram(aes(fill = ms_lab2), binwidth = 1) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 1:25) +
  labs(
    title = "Mississippi PPP Jobs Reportedly Retained",
    subtitle = "Excluding missing and zero jobs",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab2, scales = "free_y", ncol = 1)
```

`r sum(ms_ppp$jobs == 500, na.rm = TRUE)` businesses in Mississippi
reported that they retained the max of `r max(ms_ppp$jobs, na.rm = TRUE)` jobs.

```{r jobs_max, echo=FALSE}
ms_ppp %>% 
  filter(jobs == max(jobs, na.rm = TRUE)) %>% 
  select(all_of(id_vars)) %>%
  arrange(desc(amount)) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, ~dollar(., 1))) %>% 
  select(-date) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

The loan amounts for companies that retained 500 jobs vary widely from \$2,253 
to \$10,000,000.

We looked at the amount of money compared to the number of jobs. Some businesses
in Mississippi received hundreds of thousands to save one or two jobs. Other
received a few thousand to save more than 100 jobs.

```{r jobs_dollar_add, echo=FALSE}
ppp <- mutate(
  .data = ppp,
  .after = jobs,
  dollar_job = amount/jobs
)
ms2x <- mutate(
  .data = ms2x,
  .after = jobs,
  dollar_job = amount/jobs
)
```

These are the businesses that received the most and least money per job 
retained:

```{r jobs_dollar_minmax, echo=FALSE}
ht_jobs <- ppp %>%
  filter(is_ms) %>% 
  filter(!is.na(jobs) & jobs > 0) %>% 
  filter(amount > 10, jobs != 500) %>% 
  arrange(desc(dollar_job)) %>% 
  select(all_of(id_vars), dollar_job) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  top_bottom(10) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  group_split(sort) %>% 
  map(select, -date, -sort)

kable(
  x = ht_jobs[1],
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Address"
    )
  )

kable(
  x = ht_jobs[2],
  col.names = c(
    # "Date",
    "Lending Bank",
    "Amount",
    "Jobs",
    "Cost",
    "Business Name",
    "Business Address"
  )
)
```

Some of these businesses with small loans and high jobs are sole proprietorship
or self employed, which brings into question whether the job numbers are
accurate.

```{r jobs_dollar_types, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(!is.na(jobs) & jobs > 0) %>% 
  filter(amount > 10, jobs != 500) %>% 
  arrange(desc(dollar_job)) %>% 
  select(all_of(id_vars), dollar_job, type) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  filter(dollar_job < 100) %>% 
  select(-address, -city, -state, -zip, -date) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Type"
    )
  )
```

There are 26 "Self-Employed Individual" businesses which took more than a
thousand dollars in PPP money and reported having saved more than 10 jobs.

```{r jobs_dollar_self, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(type == "Self-Employed Individuals") %>% 
  arrange(desc(jobs)) %>% 
  select(type, all_of(id_vars), dollar_job) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  select(-address, -city, -state, -zip, -date) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  head(10) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Type"
    )
  )
```

The distribution of loan amount per job retained is similar for both Mississippi and the United States, both with a spike of loans at $21,000 per job.

```{r jobs_dollar_hist, echo=FALSE}
ms2x %>% 
  filter(dollar_job < 25000, dollar_job > 0) %>% 
  ggplot(aes(dollar_job)) +
  geom_histogram(aes(fill = ms_lab2), bins = 50) +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(labels = dollar, n.breaks = 10) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Dollars Loaned per Job Reportedly Retained",
    x = "Dollars/Job",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~ms_lab2,
    ncol = 1,
    scales = "free_y"
  )
```

According to [PPP loan forgiveness rules][forgive], the maximum loan amount
forgivable for sole proprietors and those self-employed is $20,833:

> In addition to the specific caps described below, the amount of loan
forgiveness requested for owner-employees and self-employed individuals’ payroll
compensation is capped at $20,833 per individual in total across all businesses
in which he or she has an ownership stake.

[forgive]: https://www.sba.gov/sites/default/files/2020-10/PPP%20--%20Loan%20Forgiveness%20FAQs%20%28October%2013%2C%202020%29.pdf

## Amounts

```{r amount_summary}
summary(ms_ppp$amount)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
ms_ppp %>% 
  filter(
    amount == min(amount) | amount == max(amount)
  ) %>% 
  select(all_of(id_vars))
```

Overall, Mississippi has a similar loan amount distribution to the United States
at large.

```{r hist_amount, echo=FALSE, fig.height=10}
ms2x %>%
  filter(amount >= 100, amount <= 100000) %>% 
  ggplot(aes(amount)) +
  geom_histogram(aes(fill = ms_lab2), bins = 50) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Mississippi PPP Loan Amounts",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab2, scales = "free_y", ncol = 1)
```

These are the 20 largest loans in the state:

```{r amount_most, echo=FALSE}
ms_ppp %>% 
  arrange(desc(amount)) %>% 
  select(all_of(id_vars)) %>% 
  head(20) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  select(-date) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Type"
    )
  )
```

## Businesses

There are `r comma(n_distinct(ms_ppp$business))` business names in the
`r comma(nrow(ms_ppp))` Mississippi PPP loans.

### Duplicate Names

There are a number of businesses which apparently relieved multiple loans, which
is not allowed under the PPP rules.

```{r biz_multi, echo=FALSE}
multi_biz <- ms_ppp %>% 
  group_by(business, address, city, state, zip) %>% 
  mutate(n = n(), group = cur_group_id()) %>% 
  filter(n > 1) %>% 
  arrange(desc(n), business) %>% 
  select(all_of(id_vars), type, n, group) %>% 
  relocate(type, .after = business)
```

```{r biz_multi_write, echo=FALSE}
write_csv(
  x = multi_biz,
  file = path(data_dir, "ms_ppp_multi_biz.csv")
)
```

```{r biz_multi_view, echo=FALSE}
multi_biz %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-date, -n, -group) %>% 
  head(11) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Type",
      "Business Address"
    )
  )
```

This could be in line with [earlier reporting][multi] that some small businesses
had received multiple PPP loans:

> The issue stems from the hectic early weeks of the program, when funding ran
out quickly and borrowers were not hearing back from their banks, industry
sources told NBC News. Although businesses must certify they are only applying
for one loan, some small-business owners applied at more than one bank to ensure
they could secure a financial lifeline amid the economic shutdown.

Some of these duplicate loans appear to be similar in value, but none of them
are exactly the same and some are from the same bank. A separate text file with
these loans to duplicate businesses has been created.

[multi]: https://www.nbcnews.com/business/business-news/ppp-loan-program-accidentally-paid-some-small-business-owners-twice-n1224061

### Duplicate Addresses

There are thousands of loans issued to duplicate addresses. In some cases, the
addresses may be for the headquarters of an organization. For example, the
Catholic Diocese of Jackson is headquartered at 237 E Amite ST, where **51**
other church locations are also incorporated (the 10 largest of those loans
are listed below).

Other cases may be worth digging into.

```{r geo_count, echo=FALSE}
ms_geo <- ms_ppp %>% 
  select(all_of(id_vars), type) %>% 
  group_by(address, city, state, zip) %>% 
  mutate(num_loans = n(), group = cur_group_id(), .before = 1) %>% 
  arrange(desc(num_loans)) %>% 
  filter(num_loans > 1)
```

```{r geo_multi_church, echo=FALSE}
ms_geo %>% 
  filter(group == 19589) %>% 
  ungroup() %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  arrange(desc(amount)) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-num_loans, -group, -type, -date) %>% 
  head(10) %>% 
  kable(
    col.names = c(
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

Loans under the Paycheck Protection Program had a maximum value of $10 million,
but business at some of these addresses received more when combined.

```{r geo_max_10, echo=FALSE}
ms_geo %>% 
  group_by(group, num_loans) %>% 
  summarise(mil_loan = sum(amount)/1e6) %>% 
  arrange(desc(mil_loan))
```

This address appears to be home to multiple staffing businesses, which
received a combined $17.6 million ([view on Google Streetview][gsv]).

[gsv]: https://goo.gl/maps/CnLrRaejf9UNKEPF9

```{r geo_multi_staff, echo=FALSE}
ms_geo %>% 
  filter(group == 29332) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  arrange(desc(amount)) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-num_loans, -group, -type) %>% 
  kable(
    col.names = c(
      "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

It seems common in the data for separate franchises or subsidiaries to receive
their own loans. Dozens of BUMPER'S DRIVE-IN restaurants all incorporated at the
same address but operating in different locations received their own loans.

```{r geo_multi_dbi, echo=FALSE}
ms_geo %>% 
    filter(group == 11714) %>% 
    mutate(
        .keep = "unused",
        address = paste(
            address, 
            paste(city, state, zip), 
            sep = ", "
        )
    ) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-num_loans, -group, -type, -date) %>% 
  kable(
    col.names = c(
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

This address in Gulfport is also the location of incorporation for nine Domino's
Pizza franchises, _three_ of which reportedly retained the maximum 500 jobs
and took a combined $10.2 million in PPP loans, which exceeds the maximum loan
amount for any single business entity. All of these loans were issued by the
same bank and all but one were issued on the same day. Weirdly enough, most of
these businesses seem to cover regions outside Mississippi.

It's worth noting that businesses with more than 500 employees not _necessarily_
ineligible for PPP loans, according to [this SBA FAQ sheet][faq]. Businesses
only have to attest eligibility.

> **Question:** Are small business concerns (as defined in section 3 of the
Small Business Act, 15 U.S.C. 632) required to have 500 or fewer employees to be
eligible borrowers in the PPP?
> 
> **Answer:** No. Small business concerns can be eligible borrowers even if they
have more than 500 employees, as long as they satisfy the existing statutory and
regulatory definition of a “small business concern” under section 3 of the Small
Business Act, 15 U.S.C. 632. A business can qualify if it meets the SBA
employee-based or revenue-based size standard corresponding to its primary
industry. Go to <www.sba.gov/size> for the industry size standards...
> 
> A business that qualifies as a small business concern under section 3 of the
Small Business Act, 15 U.S.C. 632, may truthfully attest to its eligibility for
PPP loans on the Borrower Application Form, unless otherwise ineligible.

[faq]: https://www.sba.gov/sites/default/files/2020-12/Final%20PPP%20FAQs%20%28December%209%202020%29-508.pdf

```{r geo_multi_rpm, echo=FALSE}
ms_geo %>% 
  filter(group == 11558) %>% 
  arrange(desc(amount)) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-num_loans, -group, -type, -date) %>% 
  kable(
    col.names = c(
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

## Business Type

There are `r n_distinct(ppp$type)` different kinds of business types.

The PPP was ostensibly designed to help small businesses, but most loans
nationwide were given to Corporations. In Mississippi, a greater percentage of
loans were given to LLC’s and sole proprietorships.

```{r biz_type, echo=FALSE}
ms2x %>% 
  filter(!is.na(type)) %>% 
  count(ms_lab2, type) %>% 
  ggplot(aes(x = reorder(type, n), y = n)) +
  geom_col(aes(fill = type)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 0)) +
  scale_y_continuous(labels = comma) +
  scale_fill_discrete(guide = FALSE) +
  labs(
    title = "PPP Loans Issued by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~ms_lab2,
    ncol = 1,
    scales = "free_y"
  )
```

The distribution of loan amounts by business type is roughly similar between
loans to businesses in Mississippi and the country at large. The only major 
discrepancies come from the business types with very few loans in the state.

```{r biz_amount_violin, echo=FALSE}
ms2x %>% 
  filter(
    !is.na(type), 
    amount > 100,
    type %in% most_common(type, 8)
  ) %>% 
  ggplot(aes(x = reorder(type, amount), y = amount)) +
  geom_violin(
    mapping = aes(fill = ms_lab2, alpha = ms_lab2),
    draw_quantiles = c(0.50),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
    position = "identity"
  ) +
  scale_alpha_manual(values = c(1, 0.5), guide = FALSE) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(
    breaks = c(1 %o% 10^(2:7)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Mississippi PPP: Loan Distribution by Business Type",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r biz_jobs_count, echo=FALSE}
ms2x %>% 
  filter(!is.na(type), 
         type %in% most_common(type, 8)) %>% 
  group_by(type, ms_lab2) %>% 
  summarise(jobs = sum(jobs, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(type, jobs), y = jobs)) +
  geom_col(aes(fill = type)) +
  facet_wrap(~ms_lab2, scales = "free_y", ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
   labs(
    title = "Mississippi PPP: Jobs Retained by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Total Jobs Reported"
  )
```

```{r biz_jobs_dollar, echo=FALSE}
ms2x %>% 
  filter(!is.na(type), 
         type %in% most_common(type, 8)) %>% 
  group_by(type, ms_lab2) %>% 
  summarise(dollar_job = sum(amount, na.rm = TRUE)/sum(jobs, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(type, dollar_job), y = dollar_job)) +
  geom_col(aes(fill = type)) +
  facet_wrap(~ms_lab2, ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
   labs(
    title = "Mississippi PPP: Cost of Retaining each Job by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Jobs/Dollar"
  )
```

## Business Industry

```{r naics_download, echo=FALSE}
nurl <- "https://www.census.gov/eos/www/naics/2017NAICS/6-digit_2017_Codes.xlsx"
download.file(nurl, destfile = naics_file <- file_temp(ext = "xlsx"))
```

```{r naics_read, echo=FALSE}
naics <- read_excel(
  path = naics_file,
  range = "A3:B1059",
  col_names = c("naics", "industry"),
  col_types = "text"
)
```

```{r naics_join, echo=FALSE}
ppp <- left_join(ppp, naics)
ppp <- relocate(ppp, industry, .after = naics)
ms2x <- left_join(ms2x, naics)
ms2x <- relocate(ms2x, industry, .after = naics)
```

```{r industry_count_bar, echo=FALSE}
ms2x %>% 
  filter(industry %in% most_common(industry, 10)) %>% 
  group_by(ms_lab2) %>% 
  count(industry, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(x = reorder(industry, p), y = p)) +
  geom_col(aes(fill = ms_lab2), position = "dodge") +
  scale_y_continuous(labels = percent) +
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Count Proportion by Business Industry and State",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r industry_mean_bar, echo=FALSE}
ms2x %>% 
  filter(industry %in% most_common(industry, 10)) %>% 
  group_by(ms_lab2, industry) %>% 
  summarise(sum = mean(amount, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(industry, sum), y = sum)) +
  geom_col(aes(fill = ms_lab2), position = "dodge") +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Count Proportion by Business Industry and State",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

## Business Ownership

There are four variables in the data describing the ownership of the businesses.

1. The race and ethnicity
2. Gender
3. Veteran status
3. Non-profit status

Most loans did not report this information.

```{r no_owner_info}
ms2x %>% 
  group_by(ms_lab2) %>% 
  summarise(
    no_race = mean(race == "Unanswered", na.rm = TRUE),
    no_gender = mean(gender == "Unanswered", na.rm = TRUE),
    no_vet = mean(veteran == "Unanswered", na.rm = TRUE)
  )
```

The degree of missing race data, for example, does not seem correlated with a
particular type of business.

```{r no_owner_type}
ms_ppp %>% 
  group_by(type) %>% 
  summarise(
    n = n(),
    no_race = mean(race == "Unanswered", na.rm = TRUE)
  ) %>% 
  arrange(desc(n))
```

### Race & Ethnicity

```{r race_table, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2) %>% 
  count(race, sort = TRUE) %>% 
  mutate(
    p = if_else(race != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(ms_lab2) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = ms_lab2, 
    values_from = c(n, p),
    names_glue = "{ms_lab2}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(Mississippi_p, .after = Mississippi_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Race", "MS Count", "", "US Count", ""))
```

### Gender

```{r gender_table, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2) %>% 
  count(gender, sort = TRUE) %>% 
  mutate(
    p = if_else(gender != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(ms_lab2) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = ms_lab2, 
    values_from = c(n, p),
    names_glue = "{ms_lab2}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(Mississippi_p, .after = Mississippi_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Gender", "MS Count", "", "US Count", ""))
```

### Veteran

```{r vet_table, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2) %>% 
  count(veteran, sort = TRUE) %>% 
  mutate(
    p = if_else(veteran != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(ms_lab2) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = ms_lab2, 
    values_from = c(n, p),
    names_glue = "{ms_lab2}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(Mississippi_p, .after = Mississippi_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Veteran", "MS Count", "", "US Count", ""))
```

### Non-Profit

```{r nonprofit_table, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2) %>% 
  count(non_profit, sort = TRUE) %>% 
  mutate(
    p = if_else(non_profit != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(ms_lab2) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = ms_lab2, 
    values_from = c(n, p),
    names_glue = "{ms_lab2}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(Mississippi_p, .after = Mississippi_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Non-Profit", "MS Count", "", "US Count", ""))
```

## Dates

```{r bar_day_n, echo=FALSE}
ms2x %>% 
  count(ms_lab2, date) %>% 
  ggplot(aes(x = date, y = n)) +
  geom_col(aes(fill = ms_lab2), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1, 52, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loans Approved by Day of Year",
    caption = "Source: SBA",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab2, ncol = 1, scales = "free_y")
```

```{r bar_day_amount, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2, date) %>% 
  summarise(sum = sum(amount)/1e9) %>% 
  ggplot(aes(x = date, y = sum)) +
  geom_col(aes(fill = ms_lab2), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = function(x) dollar(x, suffix = "B")) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Amount Approved by Day of Year",
    caption = "Source: SBA",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab2, ncol = 1, scales = "free_y")
```

## Lenders

Kabbage, Inc., an online automated lending platform, is the only one of the top
10 national lending banks to show up in the top 10 banks in Mississippi.

```{r lend_count_bar, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2, lender) %>% 
  summarise(
    n = n(),
    sum = sum(amount),
    mean = mean(amount)
  ) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(lender, n), n)) +
  geom_col(aes(fill = lender)) +
  coord_flip() +
  facet_wrap(~ms_lab2, ncol = 1, scales = "free") +
  scale_y_continuous(labels = comma) +
  scale_fill_discrete(guide = FALSE) +
  labs(
    title = "PPP Loans by Bank",
    caption = "Source: SBA",
    x = "Lending Bank",
    y = "Count"
  )
```

```{r lend_amount_bar, echo=FALSE}
ms2x %>% 
  group_by(ms_lab2, lender) %>% 
  summarise(
    n = n(),
    sum = sum(amount),
    mean = mean(amount)
  ) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(lender, sum), sum)) +
  geom_col(aes(fill = lender)) +
  coord_flip() +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  facet_wrap(~ms_lab2, ncol = 1, scales = "free") +
  labs(
    title = "PPP Loans by Bank",
    caption = "Source: SBA",
    x = "Lending Bank",
    y = "Sum"
  )
```

