---
title: "Mississippi PPP Breakdown"
author: "Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 4
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/ms/",
  fig.width = 12,
  fig.height = 6,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ms", "ppp", "docs"))
```

## Data

This is an analysis of Paycheck Protection Program loans in Mississippi.

On December 1, 2020 the Small Business Administration was ordered by the court
to release more detailed data on loans made through the Paycheck Protection
Program (PPP), a major facet of the 2020 CARES Act to provide stimulus funding
during the ongoing COVID-19 pandemic. This detailed release came after the SBA
initially refused to publish any database, then released only partial data on
the largest loan recipients. The full database now contains all recipient names,
addresses, and exact loan amounts.

> #### PPP Is A Delegated Loan Making Process
> PPP loans are not made by SBA. PPP loans are made by lending institutions and
then guaranteed by SBA. Accordingly, borrowers apply to lenders and self-certify
that they are eligible for PPP loans. The self- certification includes a good
faith certification that the borrower has economic need requiring the loan and
that the borrower has applied the affiliation rules and is a small business. The
lender then reviews the borrower’s application, and if all the paperwork is in
order, approves the loan and submits it to SBA...

> #### Cancelled Loans Do Not Appear In The PPP Loan Data
The public PPP data includes only active loans. Loans that were cancelled for
any reason are not included in the public data release.

## Prep

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  patchwork, # combine plots
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  readxl, # read excel files
  usmap, # plot us maps
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs # local storage 
)
```

```{r custom_funs, echo=FALSE}
top_bottom <- function(.data, nnn = 10) {
  nr = nrow(x = .data)
  top = seq(from = 1, to = nnn)
  bot = seq(from = nr - (nnn - 1), to = nr)
  mutate(
    .data = .data[c(top, bot), ], 
    .before = 1,
    sort = rep(c("head", "tail"), each = nnn)
  )
}
head_tail <- function(data, n = 10) {
  list(
    head = head(data, n = n), 
    tail = tail(data, n = n)
  )
}
str_empty_na <- function(x) {
  str_remove(str_replace_na(x), "NA")
}
solo_types <- c(
  "Sole Proprietorship",
  "Independent Contractors",
  "Self-Employed Individuals"
)
```

We can read the normalized PPP data from the IRW S3 server.

```{r clean_find}
ppp_dir <- here("us", "covid", "ppp")
ppp_file <- path(ppp_dir, "data", "clean", "sba_ppp_full.csv")
```

```{r clean_down}
if (!file_exists(ppp_file)) {
  save_object(
    object = "csv/sba_ppp_full.csv",
    bucket = "publicaccountability",
    file = ppp_file
  )
}
```

```{r clean_read}
ppp <- vroom(
  file = ppp_file,
  col_types = cols(
    naics_code = col_character(),
    zip = col_character()
  )
)
```

```{r clean_rename, echo=FALSE}
ppp <- ppp %>% 
  select(-address, -city, -state) %>% 
  relocate(address_clean, city_clean, state_clean, .before = zip) %>% 
  rename_with(~str_remove(., "_clean")) %>% 
  rename(
    date = date_approved,
    amount = loan_amount,
    jobs = jobs_reported,
    race = race_ethnicity,
    business = business_name,
    type = business_type,
    naics = naics_code
  ) %>% 
  arrange(date)
```

We can add additional variables to identify which records are in Mississippi.

```{r clean_state_id}
ppp <- ppp %>% 
  mutate(
    is_ms = !is.na(state) & state == "MS",
    ms_lab = ifelse(is_ms, "Miss.", "Other")
  )
```

And then create a separate table of Mississippi loans only.

```{r clean_filter}
ms_ppp <- filter(ppp, is_ms)
```

## Explore

There are `r comma(nrow(ms_ppp))` loans with `r ncol(ms_ppp)` columns.

Mississippi accounts for `r percent(nrow(ms_ppp)/nrow(ppp), 0.1)` of the total 
PPP loans, and `r percent(sum(ms_ppp$amount)/sum(ppp$amount), 0.1)` of 
the total amount disbursed. This is close to the 
`r percent(usa::facts$population[25]/sum(usa::facts$population), 0.1)` of the
US population that lives in the state.

## Jobs

We can also explore the number of _jobs_ reportedly retained. These numbers were
self reported and a significant amount of applications reported zero jobs or
omitted a number altogether, although the number of missing jobs reportedly
saved was smaller in Mississippi than the country at large.

```{r jobs_missing}
ppp %>% 
  group_by(ms_lab) %>% 
  summarise(
    no_job = mean(is.na(jobs)),
    zero_jobs = mean(jobs == 0, na.rm = TRUE)
  )
```

These loans, with missing or invalid jobs reported, does not appear to be tied
to the time the loan was approved and probably isn't related to confusion.

```{r jobs_missing_time, echo=FALSE}
ppp %>% 
  group_by(ms_lab, week = epiweek(date)) %>% 
  summarise(
    `No Jobs` = mean(is.na(jobs)),
    `Zero Jobs` = mean(jobs == 0, na.rm = TRUE)
  ) %>% 
  pivot_longer(
    cols = 3:4
  ) %>% 
  ggplot(aes(x = week, y = value)) +
  geom_col(aes(fill = name)) +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks = 14:32) +
  labs(
    title = "Loans Missing Reported Jobs Over Time",
    fill = "Type",
    x = "Week",
    y = "Proportion"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  facet_wrap(
    facets = ~ms_lab,
    ncol = 1,
    scales = "free_y"
  )
```

```{r id_vars, echo=FALSE}
id_vars <- c(
  "date", "lender", "amount", "jobs", "business",
  "address", "city", "state", "zip"
)
```

The loans missing job data seem to be tied to the lending bank. In Mississippi,
**100%** of loans from Wells Fargo were missing job information while **0%** of
those from PriorityOne Bank were.

```{r jobs_bank_count, echo=FALSE}
ms_bank_jobs <- ms_ppp %>% 
  group_by(lender) %>% 
  summarise(
    n = n(),
    no_job = mean(is.na(jobs) | jobs == 0)
  ) %>% 
  arrange(desc(n))

data_dir <- dir_create(here("ms", "ppp", "data"))
write_csv(
  x = ms_bank_jobs,
  file = path(data_dir, "ms_ppp_banks_nojobs.csv")
)
```

Here are the top ten banks with the greatest and least proportion of their loans
missing job information.

```{r jobs_bank_view, echo=FALSE}
ms_bank_jobs %>% 
  filter(n > 100) %>% 
  arrange(desc(no_job)) %>%
  mutate(no_job = percent(no_job, 1)) %>% 
  mutate(n = comma(n, 1)) %>% 
  top_bottom(nnn = 10) %>% 
  kable(
    col.names = c(
      "sort", "Lending Bank", "# Loans", "% w/out Jobs"
    )
  )
```

If we ignore this missing data, we can continue to breakdown the reported jobs.

Most loans did go to small businesses saving less than **25** jobs. This was
true in both Mississippi and the rest of the country.

```{r jobs_small}
ppp %>% 
  group_by(is_ms) %>% 
  summarise(small_jobs = mean(jobs < 25, na.rm = TRUE))
```

```{r hist_jobs, echo=FALSE, fig.height=10}
ppp %>%
  filter(jobs > 0, jobs <= 25) %>% 
  ggplot(aes(jobs)) +
  geom_histogram(aes(fill = ms_lab), binwidth = 1) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 1:25) +
  labs(
    title = "Mississippi PPP Jobs Reportedly Retained",
    subtitle = "Excluding missing and zero jobs",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1)
```

`r sum(ms_ppp$jobs == 500, na.rm = TRUE)` businesses in Mississippi
reportedly retained the max of `r max(ms_ppp$jobs, na.rm = TRUE)` jobs.

```{r jobs_max}
ms_ppp %>% 
  filter(jobs == max(jobs, na.rm = TRUE)) %>% 
  select(all_of(id_vars)) %>%
  arrange(desc(amount)) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, ~dollar(., 1))) %>% 
  select(-date) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

From this list, we can see an incredible range of loan amounts for the same 500
jobs retained, from 
`r dollar(min(ms_ppp$amount[ms_ppp$jobs == 500], na.rm = TRUE))`
to 
`r dollar(max(ms_ppp$amount[ms_ppp$jobs == 500], na.rm = TRUE))`.

This leads us to investigate the dollars spent to save each job. There are
some businesses in Mississippi which took hundreds of thousands to save one
or two jobs and others which took only a couple thousand to save over a hundred
jobs.

```{r jobs_dollar_add}
ppp <- mutate(
  .data = ppp,
  .after = jobs,
  dollar_job = amount/jobs
)
```

```{r jobs_dollar_minmax, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(!is.na(jobs) & jobs > 0) %>% 
  filter(amount > 10, jobs != 500) %>% 
  arrange(desc(dollar_job)) %>% 
  select(all_of(id_vars), dollar_job) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  top_bottom(10) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  select(-date, -sort) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Address"
    )
  )
```

Some of these businesses with small loans and high jobs are sole proprietorship
or self employed, which questions whether the job numbers are accurate.

```{r jobs_dollar_types, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(!is.na(jobs) & jobs > 0) %>% 
  filter(amount > 10, jobs != 500) %>% 
  arrange(desc(dollar_job)) %>% 
  select(all_of(id_vars), dollar_job, type) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  filter(dollar_job < 100) %>% 
  select(-address, -city, -state, -zip, -date) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Type"
    )
  )
```

There are "Self-Employed Individual" businesses which took thousands of dollars
to retain dozens of jobs.

```{r jobs_dollar_self, echo=FALSE}
ppp %>%
  filter(is_ms) %>% 
  filter(type == "Self-Employed Individuals") %>% 
  arrange(desc(jobs)) %>% 
  select(type, all_of(id_vars), dollar_job) %>% 
  relocate(dollar_job, .after = jobs) %>% 
  select(-address, -city, -state, -zip, -date) %>% 
  mutate(across(amount, dollar, 1)) %>% 
  mutate(across(dollar_job, dollar, 1)) %>% 
  mutate(across(business, str_trunc, 35)) %>% 
  head(10) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Cost",
      "Business Name",
      "Business Type"
    )
  )
```

The distribution of loan amount per job retained is similar for both Mississippi
and the United States, both with a spike of loans at $21,000 per job.

```{r jobs_dollar_hist, echo=FALSE}
ppp %>% 
  filter(dollar_job < 25000, dollar_job > 0) %>% 
  ggplot(aes(dollar_job)) +
  geom_histogram(aes(fill = ms_lab), bins = 50) +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(labels = dollar, n.breaks = 10) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Dollars Loaned per Job Reportedly Retained",
    x = "Dollars/Job",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~ms_lab,
    ncol = 1,
    scales = "free_y"
  )
```

The maximum amount forgivable by single people is $20,833:

> In addition to the specific caps described below, the amount of loan
forgiveness requested for owner-employees and self-employedindividuals’ payroll
compensation is capped at $20,833 per individual in total across allbusinesses
in which he or she has an ownership stake.

## Amounts

```{r amount_summary}
summary(ms_ppp$amount)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
ms_ppp %>% 
  filter(
    amount == min(amount) | amount == max(amount)
  ) %>% 
  select(all_of(id_vars))
```

Overall, Mississippi has a similar loan amount distribution to the United States
at large.

```{r hist_amount, echo=FALSE, fig.height=10}
ppp %>%
  filter(amount >= 100, amount <= 100000) %>% 
  ggplot(aes(amount)) +
  geom_histogram(aes(fill = ms_lab), bins = 50) +
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = dollar) +
  labs(
    title = "Mississippi PPP Loan Amounts",
    caption = "Source: SBA",
    x = "Amount",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1)
```

These are the 20 largest loans in the state:

```{r amount_most, echo=FALSE}
ms_ppp %>% 
  arrange(desc(amount)) %>% 
  select(all_of(id_vars)) %>% 
  head(20) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, ~dollar(./1e6))) %>% 
  select(-date) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "$ Mil.",
      "Jobs",
      "Business Name",
      "Business Type"
    )
  )
```

## Businesses

There are a number of businesses which apparently relieved multiple loans. 

```{r biz_multi, echo=FALSE}
multi_biz <- ms_ppp %>% 
  group_by(business, address, city, state, zip) %>% 
  mutate(n = n(), group = cur_group_id()) %>% 
  filter(n > 1) %>% 
  arrange(desc(n), business) %>% 
  select(all_of(id_vars), type, n, group) %>% 
  relocate(type, .after = business)
```

```{r biz_multi_write, echo=FALSE}
write_csv(
  x = multi_biz,
  file = path(data_dir, "ms_ppp_multi_biz.csv")
)
```

```{r biz_multi_view, echo=FALSE}
multi_biz %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-date, -n, -group) %>% 
  head(21) %>% 
  kable(
    col.names = c(
      # "Date",
      "Lending Bank",
      "$ Mil.",
      "Jobs",
      "Business Name",
      "Business Type",
      "Business Address"
    )
  )
```

### Duplicate Addresses

There are thousands of loans issued to duplicate addresses. This isn't unusual
in and of itself but we can explore them nonetheless.

```{r geo_count}
ms_geo <- ms_ppp %>% 
  select(all_of(id_vars), type) %>% 
  group_by(address, city, state, zip) %>% 
  mutate(num_loans = n(), group = cur_group_id(), .before = 1) %>% 
  arrange(desc(num_loans)) %>% 
  filter(num_loans > 1)
```

Loans under the Paycheck Protection Program had a maximum value of $10 million,
but we can see that some of business at some of these addresses received more
than that.

```{r geo_max_10, echo=FALSE}
ms_geo %>% 
  group_by(group, num_loans) %>% 
  summarise(mil_loan = sum(amount)/1e6) %>% 
  arrange(desc(mil_loan))
```

This address appears to be home to multiple staffing businesses, which received
a combined $17.6 million.

```{r geo_multi_staff, echo=FALSE}
ms_geo %>% 
  filter(group == 29323) %>% 
  mutate(
    .keep = "unused",
    address = paste(
      address, 
      paste(city, state, zip), 
      sep = ", "
    )
  ) %>% 
  mutate(across(amount, dollar)) %>% 
  select(-num_loans, -group, -type) %>% 
  kable(
    col.names = c(
      "Date",
      "Lending Bank",
      "Amount",
      "Jobs",
      "Business Name",
      "Business Address"
    )
  )
```

It seems common in the data for separate franchises or subsidiaries to receive
their own loans. Dozens of BUMPER'S DRIVE-IN restaurants all incorporated at the
same address but operating in different locations received their own loans.

```{r geo_multi_dbi, echo=FALSE}
ms_geo %>% 
    filter(group == 11711) %>% 
    mutate(
        .keep = "unused",
        address = paste(
            address, 
            paste(city, state, zip), 
            sep = ", "
        )
    ) %>% 
    mutate(across(amount, dollar)) %>% 
    select(-num_loans, -group, -type) %>% 
    kable(
        col.names = c(
            "Date",
            "Lending Bank",
            "Amount",
            "Jobs",
            "Business Name",
            "Business Address"
        )
    )
```

## Business Type

There are `r n_distinct(ppp$type)` different kinds of business types.

The PPP was ostensibly designed to help small businesses, but we can see that
most loans nationwide were given to Corporations. In Mississippi, a greater
percentage of loans were given to LLC's and sole proprietorship.

```{r biz_type, echo=FALSE}
ppp %>% 
  filter(!is.na(type)) %>% 
  count(ms_lab, type) %>% 
  ggplot(aes(x = reorder(type, n), y = n)) +
  geom_col(aes(fill = type)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 0)) +
  scale_y_continuous(labels = comma) +
  scale_fill_discrete(guide = FALSE) +
  labs(
    title = "PPP Loans Issued by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Loan Count"
  ) +
  facet_wrap(
    facets = ~ms_lab,
    ncol = 1,
    scales = "free_y"
  )
```

The distribution of loan amounts by business type is roughly similar between
loans to businesses in Mississippi and the country at large. The only major 
discrepancies come from the business types with very few loans in the state.

```{r biz_amount_violin, echo=FALSE}
ppp %>% 
  filter(
    !is.na(type), 
    amount > 100,
    type %in% most_common(type, 8)
  ) %>% 
  ggplot(aes(x = reorder(type, amount), y = amount)) +
  geom_violin(
    mapping = aes(fill = ms_lab, alpha = ms_lab),
    draw_quantiles = c(0.50),
    scale = "width",
    trim = TRUE,
    na.rm = TRUE,
    position = "identity"
  ) +
  scale_alpha_manual(values = c(1, 0.5), guide = FALSE) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(
    breaks = c(1 %o% 10^(2:7)),
    labels = dollar,
    trans = "log10"
  ) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Mississippi PPP: Loan Distribution by Business Type",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r biz_jobs_count, echo=FALSE}
ppp %>% 
  filter(!is.na(type), 
         type %in% most_common(type, 8)) %>% 
  group_by(type, ms_lab) %>% 
  summarise(jobs = sum(jobs, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(type, jobs), y = jobs)) +
  geom_col(aes(fill = type)) +
  facet_wrap(~ms_lab, scales = "free_y", ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
   labs(
    title = "Mississippi PPP: Jobs Retained by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Total Jobs Reported"
  )
```

```{r biz_jobs_dollar, echo=FALSE}
ppp %>% 
  filter(!is.na(type), 
         type %in% most_common(type, 8)) %>% 
  group_by(type, ms_lab) %>% 
  summarise(dollar_job = sum(amount, na.rm = TRUE)/sum(jobs, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(type, dollar_job), y = dollar_job)) +
  geom_col(aes(fill = type)) +
  facet_wrap(~ms_lab, ncol = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(
    labels = function(x) {
      stringr::str_wrap(x, width = 10)
    }
  ) +
   labs(
    title = "Mississippi PPP: Cost of Retaining each Job by Business Type",
    caption = "Source: SBA",
    x = "Business Type",
    y = "Jobs/Dollar"
  )
```

## Business Industry

```{r naics_download, echo=FALSE}
nurl <- "https://www.census.gov/eos/www/naics/2017NAICS/6-digit_2017_Codes.xlsx"
download.file(nurl, destfile = naics_file <- file_temp(ext = "xlsx"))
```

```{r naics_read, echo=FALSE}
naics <- read_excel(
  path = naics_file,
  range = "A3:B1059",
  col_names = c("naics", "industry"),
  col_types = "text"
)
```

```{r naics_join, echo=FALSE}
ppp <- left_join(ppp, naics)
ppp <- relocate(ppp, industry, .after = naics)
```

```{r industry_count_bar, echo=FALSE}
ppp %>% 
  filter(industry %in% most_common(industry, 10)) %>% 
  group_by(ms_lab) %>% 
  count(industry, sort = TRUE) %>% 
  mutate(p = n/sum(n)) %>% 
  ggplot(aes(x = reorder(industry, p), y = p)) +
  geom_col(aes(fill = ms_lab), position = "dodge") +
  scale_y_continuous(labels = percent) +
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Count Proportion by Business Industry and State",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

```{r industry_mean_bar, echo=FALSE}
ppp %>% 
  filter(industry %in% most_common(industry, 10)) %>% 
  group_by(ms_lab, industry) %>% 
  summarise(sum = mean(amount, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(industry, sum), y = sum)) +
  geom_col(aes(fill = ms_lab), position = "dodge") +
  scale_y_continuous(labels = dollar) +
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loan Count Proportion by Business Industry and State",
    caption = "Source: SBA",
    fill = "State",
    x = "Amount",
    y = "Count"
  )
```

## Business Ownership

There are four variables in the data describing the ownership of the businesses.

1. The race and ethnicity
2. Gender
3. Veteran status
3. Non-profit status

Most loans did not report this information.

```{r no_owner_info}
ppp %>% 
  group_by(is_ms) %>% 
  summarise(
    no_race = mean(race == "Unanswered", na.rm = TRUE),
    no_gender = mean(gender == "Unanswered", na.rm = TRUE),
    no_vet = mean(veteran == "Unanswered", na.rm = TRUE)
  )
```

The degree of missing race data, for example, does not seem correlated with a
particular type of business.

```{r no_owner_type}
ms_ppp %>% 
  group_by(type) %>% 
  summarise(
    n = n(),
    no_race = mean(race == "Unanswered", na.rm = TRUE)
  ) %>% 
  arrange(desc(n))
```

### Race & Ethnicity

```{r race_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state == "MS" & !is.na(state)) %>% 
  count(race, sort = TRUE) %>% 
  mutate(
    p = if_else(race != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Race", "US Count", "", "MS Count", ""))
```

### Gender

```{r gender_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state == "MS" & !is.na(state)) %>% 
  count(gender, sort = TRUE) %>% 
  mutate(
    p = if_else(gender != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Gender", "US Count", "", "MS Count", ""))
```

### Veteran

```{r vet_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state == "MS" & !is.na(state)) %>% 
  count(veteran, sort = TRUE) %>% 
  mutate(
    p = if_else(veteran != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Veteran", "US Count", "", "MS Count", ""))
```

### Non-Profit

```{r nonprofit_table, echo=FALSE}
ppp %>% 
  group_by(is_ms = state == "MS" & !is.na(state)) %>% 
  count(non_profit, sort = TRUE) %>% 
  mutate(
    p = if_else(non_profit != "Unanswered", n, NA_integer_),
    p = p/sum(p, na.rm = TRUE)
  ) %>% 
  arrange(is_ms) %>% 
  ungroup() %>% 
  mutate(across(n, comma, 1)) %>% 
  mutate(across(p, percent, 0.1)) %>% 
  pivot_wider(
    id_cols = 1:2, 
    names_from = is_ms, 
    values_from = c(n, p),
    names_glue = "{is_ms}_{.value}",
    names_sort = TRUE
  ) %>% 
  relocate(FALSE_p, .after = FALSE_n) %>% 
  mutate(across(3:5, str_empty_na)) %>% 
  kable(col.names = c("Non-Profit", "US Count", "", "MS Count", ""))
```

## Dates

```{r bar_day_n, echo=FALSE}
ppp %>% 
  count(ms_lab, date) %>% 
  ggplot(aes(x = date, y = n)) +
  geom_col(aes(fill = ms_lab), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(1, 52, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Loans Approved by Day of Year",
    caption = "Source: SBA",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, ncol = 1, scales = "free_y")
```

```{r bar_day_amount, echo=FALSE}
ppp %>% 
  group_by(ms_lab, date) %>% 
  summarise(sum = sum(amount)/1e9) %>% 
  ggplot(aes(x = date, y = sum)) +
  geom_col(aes(fill = ms_lab), position = "dodge") + 
  scale_fill_brewer(palette = "Dark2", guide = FALSE) +
  scale_y_continuous(labels = function(x) dollar(x, suffix = "B")) +
  theme(legend.position = "bottom") +
  labs(
    title = "PPP Amount Approved by Day of Year",
    caption = "Source: SBA",
    x = "Day of year",
    y = "Count"
  ) +
  facet_wrap(~ms_lab, ncol = 1, scales = "free_y")
```

## Lenders

```{r lend_count_bar, echo=FALSE}
ppp %>% 
  group_by(ms_lab, lender) %>% 
  summarise(
    n = n(),
    sum = sum(amount),
    mean = mean(amount)
  ) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(lender, n), n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ms_lab, ncol = 1, scales = "free") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "PPP Loans by Bank",
    caption = "Source: SBA",
    x = "Lending Bank",
    y = "Count"
  )
```

```{r lend_amount_bar, echo=FALSE}
ppp %>% 
  group_by(ms_lab, lender) %>% 
  summarise(
    n = n(),
    sum = sum(amount),
    mean = mean(amount)
  ) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(lender, sum), sum)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ms_lab, ncol = 1, scales = "free") +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "PPP Loans by Bank",
    caption = "Source: SBA",
    x = "Lending Bank",
    y = "Sum"
  )
```

