---
title: "Kentucky Expenditures"
author: "Yanqi Xu & Kiernan Nicholls"
date: "`r date()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 120)
  set.seed(5)
}
```

```{r create-docs-dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("state", "ky", "expends", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

```{r load-packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
}
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  refinr, # cluster & merge
  scales, # format strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  cli, # commend line
  fs # local storage 
)
```

This diary was run using `campfin` version `r packageVersion("campfin")`.

```{r campfin-version}
packageVersion("campfin")
```

```{r package-options, echo=FALSE}
options(options(knitr.kable.NA = ""))
```

This document should be run as part of the `R_tap` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_tap` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where-here}
# where does this document knit?
here::i_am("state/ky/expends/docs/ky_expends_diary.Rmd")
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Source

## Data

State expenditures can be obtained from the [Kentucky Registry of Election 
Finance (KREF)][kref]. Data can be exported from the KREF candidate search
[page](https://secure.kentucky.gov/kref/publicsearch/CandidateSearch/).

[kref]: https://kref.ky.gov/Pages/default.aspx

## Download

We can request all expenditures for each month since Jan 1, 2011.

```{r raw_dir}
raw_dir <- dir_create(here("state","ky", "expends", "data", "raw"))
```

```{r raw_info}
raw_info <- dir_info(raw_dir)
sum(raw_info$size)
raw_info %>% 
  select(path, size, modification_time) %>% 
  mutate(across(path, basename))
```

We have downloaded `nrow(raw_info)` files totaling `sum(raw_info$size)` in 
size.

## Read

Given all of these files have the same structure, we can read them all into a
single data frame at once.

```{r raw_read}
kye <- read_csv(
  raw_info$path,
  name_repair = make_clean_names,
  col_types = cols(
    .default = col_character(),
   disbursement_date = col_datetime(format="%m/%d/%Y %H:%M:%S"),
   election_date = col_datetime(format = "%m/%d/%Y %H:%M:%S"),
    disbursement_amount = col_double()
  )
)
```


## Explore

There are `r comma(nrow(kye))` rows of `r ncol(kye)` columns. Each record
represents a single contribution from an organization or individual to a
campaign or committee.

```{r glimpse}
glimpse(kye)
tail(kye)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(kye, count_na)
```

expenditures can be made to an individual (with a `from_candidate_last_name`), an
organization (with a `from_organization_name` name), or an individual _with_ a committee
name as well. We only want to flag records that are truly missing _any_ way to
identify the parties of the transaction.

We can flag any record missing a key variable needed to identify a transaction.

```{r na_flag}
kye <- kye %>% 
   unite(
    col = from_candidate_any_name,
    from_candidate_first_name, from_candidate_last_name,
    sep = " ",
    remove = FALSE
  ) %>% 
  mutate(
    from_any = coalesce(from_organization_name, from_candidate_any_name)
  ) %>% 
   unite(
    col = recipient_any_name,
    recipient_first_name, recipient_last_name,
    sep = " ",
    remove = FALSE
  ) %>% 
  mutate(
    to_any = coalesce(recipient_any_name, organization_name)
  ) %>% 
  flag_na(to_any, from_any, disbursement_amount) %>% 
  select(
    -from_candidate_any_name,
    -recipient_any_name,
    -to_any, -from_any
    )
```

```{r na_sum}
sum(kye$na_flag)
```

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
kye <- flag_dupes(kye, everything())
sum(kye$dupe_flag)
```

`percent(mean(kye$dupe_flag))` of all records are duplicates.

```{r dupe_view}
kye %>% 
  filter(dupe_flag) %>% 
  select(
    from_candidate_last_name, recipient_last_name, 
    disbursement_amount, election_date, office_sought
  ) %>% 
  arrange(from_candidate_last_name, election_date, disbursement_amount)
```

Without a contribution date, it's difficult to identify these duplicates as
stemming from a flaw in the data (rather than "true" duplicate repeat payments).
We will flag them but not remove any.

### Categorical

```{r distinct_count}
col_stats(kye, n_distinct)
```

```{r distinct_plots, echo=FALSE, fig.height=3}
explore_plot(kye, election_type)
explore_plot(kye, office_sought)
```

### Amounts

```{r amount_summary}
summary(kye$disbursement_amount)
mean(kye$disbursement_amount <= 0)
```

These are the records with the minimum and maximum amounts.

```{r amount_minmax}
glimpse(kye[c(which.max(kye$disbursement_amount), which.min(kye$disbursement_amount)), ])
```

```{r hist_amount, echo=FALSE}
kye %>%
  ggplot(aes(disbursement_amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Kentucky expenditures Amount Distribution",
    caption = "Source: {source}",
    x = "Amount",
    y = "Count"
  )
```

### Dates

There is no variable containing the actual date the expenditures were made.

The `election_date` is the only date variable.

```{r election_date_bar, echo=FALSE}
kye %>% 
  mutate(year = year(election_date)) %>% 
  filter(year>=2010) %>% 
  count(year) %>% 
  ggplot(aes(year, n)) +
  geom_col(aes(fill = n)) +
  scale_fill_viridis_c(end = 0.5, guide = FALSE) +
  scale_x_continuous(n.breaks = 20) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "expenditures by election year",
    x = "Year",
    y = "Count"
  )
```

## Conclude

```{r clean_glimpse}
glimpse(sample_n(kye, 50))
```

1. There are `r comma(nrow(kye))` records in the database.
1. There are `r comma(sum(kye$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(kye$na_flag))` records missing key variables.
1. Consistency in geographic data has been improved with `campfin::normal_*()`.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

## Export

```{r}
kye <- kye %>% select(-na_flag)
```

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("state","ky", "expends", "data", "clean"))
clean_path <- path(clean_dir, "ky_expends_clean.csv")
write_csv(kye, clean_path, na = "")
(clean_size <- file_size(clean_path))
non_ascii(clean_path)
```

## Upload

We can use the `aws.s3::put_object()` to upload the text file to the IRW server.

```{r aws_upload, eval=FALSE}
aws_path <- path("csv", basename(clean_path))
if (!object_exists(aws_path, "publicaccountability")) {
  put_object(
    file = clean_path,
    object = aws_path, 
    bucket = "publicaccountability",
    acl = "public-read",
    show_progress = TRUE,
    multipart = TRUE
  )
}
aws_head <- head_object(aws_path, "publicaccountability")
(aws_size <- as_fs_bytes(attr(aws_head, "content-length")))
unname(aws_size == clean_size)
```
