---
title: "Alaska voters"
author: "Jennifer LaFleur"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 3
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
doc_dir <- fs::dir_create(here::here("ak", "voters", "docs"))
```


## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across public data about individuals, organizations and locations.

Our goal is to standardize public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
2. Check for entirely duplicated records.
3. Check ranges of continuous variables.
4. Is there anything blank or missing?
5. Check for consistency issues.
6. Create a five-digit ZIP Code called `zip`.
7. Create a `year` field from the transaction date.
8. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.


```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  gluedown, # printing markdown
  janitor, # clean data frames
  campfin, # custom irw tools
  aws.s3, # aws cloud storage
  readxl, # read excel files
  refinr, # cluster & merge
  scales, # format strings
  digest, # hash strings
  knitr, # knit documents
  vroom, # fast reading
  rvest, # scrape html
  glue, # code strings
  here, # project paths
  httr, # http requests
  fs, # local storage 
  stringr # string functions
 )
```
## Read

```{r read voters}
akv <- read.csv('ak_Public_Voter_List_20230118.csv',strip.white=TRUE)
```


```{r read party}
party <- read.csv('party.csv')
akv <- akv %>%  
    left_join(party, by = c("PARTY"="partycode"))
```

## Data

The Alaska voter data was obtained via open records request from the Alaska Division of Elections. And received in January 2023.


Two files were provided to the Accountability Project:

1. a csv containing the voter data 
2. a pdf containing the record layout 


#### Columns

```{r rec_layout, echo=FALSE}
rec_layout <- tribble(
  ~column, ~description,
"UN",                 "Address undeliverable",
"PARTY",	            "Voter party affiliation",
"DP",	                "Voter house district and precinct",
"LAST_NAME",          "Voter last name",
"FIRST_NAME",	        "Voter first name",
"MIDDLE_NAME",	      "Voter middle init_name",
"SUFFIX_NAME",	      "Voter suffix",
"ASCENSION",	        "Random number assigned to the voter",                    
"REG_DATE",	          "Date of most recent registration (if included)",
"ORG_REG_DATE",	      "Date when voter originally registered (if included)",
"DIST_DATE",	        "Date when voter registered in house district (if included)",
"RESIDENCE_ADDRESS",  "Voter residence address",
"RESIDENCE_CITY",	    "Voter residence city",
"RESIDENCE_ZIP",	    "Voter residence ZIP",
"MAILING_ADDRESS1",	  "Line 1 voter mailing address",
"MAILING_ADDRESS2",   "Line 2 voter mailing address",
"MAILING_ADDRESS3",	  "Line 3 voter mailing address",  
"MAILING_CITY",	      "Voter mailing city",
"MAILING_STATE",	    "Voter mailing state",
"MAILING_ZIP",	      "Voter mailing ZIP",
"MAILING_COUNTRY",	  "If overseas, voter mailing country",
"GENDER",	            "Voter gender",
"VH1",	              "voter history columns", 
"VH2",	              " ",
"VH3",	              " ",	
"VH4",		            " ",
"VH5",		            " ",
"VH6",	              " ",	
"VH7",	              " ",	
"VH8",	              " ",	
"VH9",	              " ",
"VH10",	              " ",	
"VH11",	              " ",	
"VH12",	              " ",
"VH13",	              " ",	
"VH14",	              " ",	
"VH15",	              " ",	
"VH16",	              " "	
)
```

```{r echo=FALSE}
kable(rec_layout)
```

Voter history is represented by an election ID that consists of five to six characters with the first	two digits being the year of the election followed by a three to four lettered election name code. The second set of data is the method that the voter voted:	

P Voted in person at the polls	
E Voted an Early Vote ballot	
A Voted an Absentee ballot	
Q Voted a Questioned ballot	

Statewide primary and general elections are consistently named with the year and	
election codes of PRIM and GENR:	
18PRIM 2018 Primary Election	
18GENR 2018 General Election	

Statewide special elections are reflected as:	
07SPEC 2007 Statewide Special Election	

State conducted Regional Educational Attendance Area elections (school board elections	
in unorganized boroughs of Alaska) are reflected as REAA, RE## or R###:	
18REAA 2018 Regional Educational Attendance Area elections	
18RE17 2018 Regional Educational Attendance Area election in REAA 17	
18R123 2018 Regional Education Attendance Area election in REAA 12	
Section 3	

Local and special city and borough elections from 2016 forward, the election ID will show the two-digit year and three to four alpha characters that represents the	city/borough name or geographic region name:	

18VALD 2018 local election for the City of Valdez	
18MOA 2018 local election for the Municipality of Anchorage	
18WASI 2018 local election for the City of Wasilla	
18FNSB 2018 local election for the Fairbanks North Star Borough	
18NSLB 2018 local election for the North Slope Borough	

At times, city and boroughs will conduct special or run-off elections.  The last characterof the election ID will reflect either a ‘S’ or ‘R’ for special or runoff.	Prior to 2016, local and special city and borough elections were:	
12REGL 2012 Regular Local Election	
12SPEL 2012 Special Local Election	

For a list of city and/or boroughs within the State of Alaska, visit the division	
Research webpage below and locate the Alaska Community List selections:	
http://www.elections.alaska.gov/Core/electionresources.php	




#### Political party

Voter political party affiliation.
Recognized Political Parties: Are those parties that have gained recognized
political party status under Alaska Statute 15.80.010(27) and voters appear on
voter lists as follows:
A:Alaskan Independence Party
D:Alaska Democratic Party
R:Alaska Republican Party

Political Groups: Are those groups that have applied for party status but have
not met the qualifications to be a recognized political party under Alaska
Statute 15.80.010(26) and voters appear on voter lists as follows:
C:Alaska Constitution Party
L:Alaska Libertarian Party
E:Moderate Party of Alaska
O:Progressive Party of Alaska
F:Freedom Reform Party
P:Patriot's Party of Alaska
G:Green Party of Alaska
V:Veterans Party of Alaska
H:OWL Party
W:UCES Clowns Party
K:Alliance Party of Alaska
No Affiliation / Undeclared: Voters who chose not to declare an affiliation or
who are unaffiliated appear on voter lists as follows:
N:Nonpartisan (no affiliation)
U:Undeclared


## Explore

There are `r scales::comma(nrow(akv))` rows of `r ncol(akv)` columns.

```{r glimpse}
glimpse(akv)
tail(akv)
```

### Missing

Columns vary in their degree of missing values.

```{r na_count}
col_stats(akv, count_na)
```

No columns are missing the registration date or last name needed to identify a
voter.

### Duplicates

We can also flag any record completely duplicated across every column.

```{r dupe_flag}
d1 <- duplicated(akv, fromLast = FALSE)
d2 <- duplicated(akv, fromLast = TRUE)
akv <- mutate(akv, dupe_flag = d1 | d2)
sum(akv$dupe_flag)
```

```{r dupe_rm, echo=FALSE}
rm(d1, d2)
flush_memory(1)
```

```{r dupe_view}
akv %>% 
  filter(dupe_flag) %>% 
  select(ORG_REG_DATE, LAST_NAME, PARTY) %>% 
  arrange(ORG_REG_DATE)
```

### Categorical

```{r distinct_count}
col_stats(akv, n_distinct)
```


### Cleaning

Pull the year from ORG_REG_DATE

```{r date_year}
akv <- mutate(akv, year = str_sub(ORG_REG_DATE, start= -4))
```

Add the state

```{r state}
akv <- mutate(akv, state = "Alaska")
```

Create a five-digit zip field

```{r ZIP}
akv <- mutate(akv, zip = str_sub(RESIDENCE_ZIP, 1, 5))
```

Clean column names

```{r names}
akv <- clean_names(akv)
```

## Explore


```{r bar_year, echo=FALSE}
explore_plot(akv, year)
explore_plot(akv, party)
```


## Wrangle

Check geo variables

```{r geo_check, echo=FALSE}
unique(akv$residence_city)
unique(akv$state)
prop_in(akv$zip, valid_zip)
```

## Conclude


```{r clean_glimpse}
glimpse(sample_n(akv, 25))
```

1. There are `r comma(nrow(akv))` records in the database.
2. There are `r comma(sum(akv$dupe_flag))` duplicate records in the database.
3. The range and distribution of `year` seem reasonable.


## Export

Now the file can be saved on disk for upload to the Accountability server.

```{r clean_dir}
clean_dir <- dir_create(here("ak", "voters", "data", "clean"))
clean_path <- path(clean_dir, "ak_voters_clean.csv")
write_csv(akv, clean_path, na = "")
(clean_size <- file_size(clean_path))
file_encoding(clean_path) %>% 
  mutate(across(path, path.abbrev))
```

## Dictionary

The following table describes the variables in our final exported file:

```{r dict_make, echo=FALSE}
dict_raw <- tibble(
  var = md_code(names(akv)),
  type = md_code(map_chr(akv, typeof)),
  def = c(
    "undeliverable",
    "party",
    "district precinct",
    "last name",
    "first name",
    "middle name/init",
    "suffix",
    "random ID",
    "Most recent registration date",
    "Original registration date",
    "date voter registered in district",
    "residence address",
    "residence city",
    "residence zip",
    "mailing address1",
    "mailing address2",
    "mailing address3",
    "mailing city",
    "mailing state",
    "mailing ZIP code",
    "mailing country if outside U.S.",
    "gender",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "voting history",
    "party name",
    "duplicate flag",
    "Original registration year",
    "Residential state",
    "5-digit residential ZIP code"
       )
)
```


```{r dict_md, echo=FALSE}
(dict_md <- kable(
  x = dict_raw,
  format = "markdown",
  col.names = c("Column", "Type", "Definition")
))
```

