---
title: "New York Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
options(width = 99)
set.seed(5)
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("nyc", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Software

This data is processed using the free, open-source statistical computing
language R, which can be [installed from CRAN][cran] for various opperating
systems. For example, R can be installed from the apt package repository on
Ubuntu.

```bash
sudo apt update
sudo apt -y upgrade
sudo apt -y install r-base
```

[cran]: https://cran.r-project.org/

The following additional R packages are needed to collect, manipulate,
visualize, analyze, and communicate these results. The `pacman` package will
facilitate their installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  gluedown, # print markdown
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  readxl, # read excel
  knitr, # knit documents
  vroom, # read files fast
  furrr, # parallel map
  glue, # combine strings
  here, # relative storage
  httr, # http requests
  pryr, # memory usage
  fs # search storage 
)
```

This document should be run as part of the `us_spending` project, which lives as
a sub-directory of the more general, language-agnostic [`irworkshop/tap`][tap]
GitHub repository.

The `us_spending` project uses the [RStudio projects][rproj] feature and should
be run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

Contracts data is obtained from the New York City 
[Office of Citywide Purchasing (OCP)][ocp] via the city's [OpenData portal][od].
In light of the ongoing COVID-19 outbreak in New York City, our typical efforts
to ensure a portal contains all relevant data were not made. While the number of
records is fairly small, the amount of data seems consitent year to year and up
to date.

[od]: https://opendata.cityofnewyork.us/

## Import

We can import [the data][data] directly from the OpenData portal.

[data]: https://data.cityofnewyork.us/City-Government/Recent-Contract-Awards/qyyg-4tf5
[ocp]: https://www1.nyc.gov/site/dcas/business/working-with-citywide-procurement.page

First, we simply download using an `httr::GET()` request to the portal site.

```{r raw_path}
raw_url <- "https://data.cityofnewyork.us/api/views/qyyg-4tf5/rows.csv"
raw_dir <- dir_create(here("nyc", "contracts", "data", "raw"))
raw_path <- path(raw_dir, "Recent_Contract_Awards.csv")
```

```{r raw_get}
if (!file_exists(raw_path)) {
  GET(
    url = raw_url,
    write_disk(raw_path, overwrite = TRUE),
    query = list(accessType = "DOWNLOAD")
  )
}
```

Then the local raw text file can be read into memory with `vroom::vroom()`. 

```{r raw_read}
nyc <- vroom(
  file = raw_path,
  .name_repair = make_clean_names,
  escape_double = FALSE,
  escape_backslash = FALSE,
  na = c("", "NA", "......"),
  col_types = cols(
    .default = col_character(),
    StartDate = col_date_usa(),
    EndDate = col_date_usa(),
    ContractAmount = col_double()
  )
)
```

A number of columns contain no informatio and can be removed.

```{r raw_remove}
nyc <- remove_empty(nyc, "cols")
```

## Wrangle

The `vendor_address` string contains both the street address, the city, state,
and ZIP code. We will have to extract the last two for their own variables.

```{r addr_wrangle}
# copy address in new column
nyc <- mutate(nyc, addr = str_to_upper(vendor_address))
# remove periods in abbrevs
nyc$addr <- str_remove_all(nyc$addr, "\\.(?=\\S)")
# extract and remove zip from end
nyc <- mutate(nyc, zip = normal_zip(str_extract(addr, "\\d+(?:-\\d{4})?$")))
nyc$addr <- str_trim(str_remove(nyc$addr, "[:punct:]?\\s?\\d+(?:-\\d{4})?$"))
# abbrev full states at end of addr
rx_abbs <- valid_state
names(rx_abbs) <- glue("{valid_name}$")
nyc$addr <- str_replace_all(nyc$addr, rx_abbs)
# extract upper abbs from end
nyc <- mutate(nyc, state = str_extract(addr, "(?<=\\W)[:upper:]{2}$"))
# remove abbs and extra from end
nyc$addr <- str_remove(nyc$addr, "[:punct:]?\\s[:upper:]{2}$")
nyc <- relocate(nyc, state, zip, .after = addr)
nyc$addr <- nyc$addr %>% 
  # normalize remaining addr
  str_normal(punct = FALSE) %>% 
  # abbreviate street abbs
  abbrev_full(usps_street)
```

```{r addr_view}
nyc %>% 
  select(vendor_address, addr, state, zip) %>% 
  distinct() %>% 
  sample_n(20)
```

We can also add "NY" as the state for spending agencies.

```{r state_add}
nyc <- mutate(nyc, agency_state = "NY", .after = agency_name)
```

## Explore

```{r glimpse}
head(nyc)
tail(nyc)
glimpse(sample_n(nyc, 20))
```

A number of records are missing one or more key variables.

```{r na_count}
col_stats(nyc, count_na)
```

We can flag these records with `campfin::flag_na()`.

```{r na_flag}
nyc <- flag_na(nyc, agency_name, vendor_name, start_date, contract_amount)
percent(mean(nyc$na_flag), 0.01)
```

We can also use `campfin::dupe_flag()` to flag duplicate records. In this case,
there are none!

```{r dupe_flag, warning=TRUE}
nyc <- flag_dupes(nyc, everything(), .check = TRUE)
sum(nyc$dupe_flag)
```

We can also count and plot some of the categorical values.

```{r n_distinct}
col_stats(nyc, n_distinct)
```

```{r bar-type, fig.keep=FALSE}
explore_plot(nyc, category_description)
explore_plot(nyc, special_case_reason_description)
```

### Continuous

We should also check to ensure the amount and dates are the reasonable.

Only a few records of a `contract_amount` value less than or equal to zero.

```{r ammount_summary}
summary(nyc$contract_amount)
mean(nyc$contract_amount <= 0, na.rm = TRUE) # less than zero
```

The amounts are logarithmically normal, as we'd expect with something like this.

```{r hist_amount, echo=FALSE}
nyc %>%
  filter(contract_amount >= 1e3) %>% 
  ggplot(aes(contract_amount)) +
  geom_histogram(fill = dark2["purple"], bins = 30) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(2:8)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "New York Contracts Amount Distribution",
    caption = "Source: NYC OCP",
    x = "Amount",
    y = "Count"
  )
```

Most of the `start_date` and `end_date` values are the same. We will use the
year of the `start_date` to create a new 4 digit `year` variable.

```{r date_year}
mean(nyc$start_date == nyc$end_date) # mostly same
nyc <- mutate(nyc, year = year(start_date))
```

No records have bad dates in the distant past or future.

```{r date_range}
min(nyc$start_date)
sum(nyc$year < 2002)
max(nyc$end_date)
sum(nyc$end_date > today())
```

There are a fairly constant number of contracts per year.

```{r bar_year, echo=FALSE}
nyc %>% 
  count(year) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = dark2["purple"]) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2000, 2020, by = 2)) +
  theme(legend.position = "bottom") +
  labs(
    title = "New York Contracts by Year",
    caption = "Source: NYC OCP",
    x = "Year Made",
    y = "Count"
  )
```

## Export

```{r clean_glimpse}
glimpse(sample_n(nyc, 20))
```

1. There are `r comma(nrow(nyc))` records in the database.
1. There are `r comma(sum(nyc$dupe_flag))` duplicate records in the database.
1. The range and distribution of amount and dates seem mostly reasonable.
1. Geographic variables were extracted from a single string or added manually.
1. The 4-digit `year` variable has been created with `lubridate::year()`.

```{r clean_write}
clean_dir <- dir_create(here("nyc", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "nyc_contracts_clean.csv")
write_csv(nyc, path = clean_path, na = "")
file_size(clean_path)
file_encoding(clean_path)
```

## Upload

```{r clean_upload}
s3_dir <- "s3:/publicaccountability/csv/"
s3_url <- path(s3_dir, basename(clean_path))
if (require(duckr)) {
  duckr::duck_upload(clean_path, s3_url)
}
```
