---
title: "Washington Contracts"
author: "Kiernan Nicholls"
date: "`r Sys.time()`"
output:
  github_document: 
    df_print: tibble
    toc: true
    toc_dept: 2
editor_options: 
  chunk_output_type: console
---

<!-- Place comments regarding knitting here -->

```{r setup, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "../plots/",
  fig.width = 10,
  dpi = 300
)
if (!interactive()) {
  options(width = 99)
  set.seed(5)
}
```

```{r create_docs_dir, eval=FALSE, echo=FALSE, include=FALSE}
fs::dir_create(here::here("wa", "contracts", "docs"))
```

## Project

The Accountability Project is an effort to cut across data silos and give
journalists, policy professionals, activists, and the public at large a simple
way to search across huge volumes of public data about people and organizations.

Our goal is to standardizing public data on a few key fields by thinking of each
dataset row as a transaction. For each transaction there should be (at least) 3
variables:

1. All **parties** to a transaction.
2. The **date** of the transaction.
3. The **amount** of money involved.

## Objectives

This document describes the process used to complete the following objectives:

1. How many records are in the database?
1. Check for entirely duplicated records.
1. Check ranges of continuous variables.
1. Is there anything blank or missing?
1. Check for consistency issues.
1. Create a five-digit ZIP Code called `zip`.
1. Create a `year` field from the transaction date.
1. Make sure there is data on both parties to a transaction.

## Packages

The following packages are needed to collect, manipulate, visualize, analyze,
and communicate these results. The `pacman` package will facilitate their
installation and attachment.

The IRW's `campfin` package will also have to be installed from GitHub. This
package contains functions custom made to help facilitate the processing of
campaign finance data.

```{r load_packages, message=FALSE, warning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("irworkshop/campfin")
pacman::p_load(
  tidyverse, # data manipulation
  lubridate, # datetime strings
  magrittr, # pipe operators
  jsonlite, # parse json files
  janitor, # dataframe clean
  refinr, # cluster and merge
  scales, # format strings
  knitr, # knit documents
  vroom, # read files fast
  glue, # combine strings
  here, # relative storage
  fs # search storage 
)
```

This document should be run as part of the `R_campfin` project, which lives as a
sub-directory of the more general, language-agnostic
[`irworkshop/accountability_datacleaning`][tap] GitHub repository.

The `R_campfin` project uses the [RStudio projects][rproj] feature and should be
run as such. The project also uses the dynamic `here::here()` tool for file
paths relative to _your_ machine.

```{r where_here}
# where does this document knit?
here::here()
```

[tap]: https://github.com/irworkshop/accountability_datacleaning
[rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

## Data

The contracts [data] is from the Washington state [Department of Enterprise 
Services][des]. The data can be retrieved from the Washington OpenData portal
under the title "Master Contract Sales Data by Customer, Contract, Vendor" in
the "Procurements and Contracts" category. The data file was last updated on
January 21, 2020. As of writing, due to the ongoing Covid-19 pandemic, our
typical efforts to verify this OpenData source as the most updated were not 
made. 

[data]: https://data.wa.gov/Procurements-and-Contracts/Master-Contract-Sales-Data-by-Customer-Contract-Ve/n8q6-4twj
[des]: https://des.wa.gov/

## Import

We can import the text file directly into R from the OpenData portal URL.

```{r raw_read}
wac <- vroom(
  file = "https://data.wa.gov/api/views/n8q6-4twj/rows.csv",
  .name_repair = make_clean_names,
  col_types = cols(
    `Calendar Year` = col_integer(),
    `Q1 Sales Reported` = col_double(),
    `Q2 Sales Reported` = col_double(),
    `Q3 Sales Reported` = col_double(),
    `Q4 Sales Reported` = col_double(),
  )
)
```

Then we can convert some quasi-logical variables to a true logical type.

```{r raw_convert}
count(wac, vet_owned)
wac <- mutate_at(
  .tbl = wac,
  .vars = vars(12:15),
  .funs = equals, "Y"
)
```

## Explore

```{r glimpse}
head(wac)
tail(wac)
glimpse(wac)
```

### Missing

```{r na_count}
col_stats(wac, count_na)
```

About 6% of transactions are missing the customer name.

```{r na_flag}
wac <- wac %>% flag_na(customer_name, vendor_name)
percent(mean(wac$na_flag), 0.01)
```

### Duplicates

There are also a small handful of duplicate records, all also missing a name.

```{r dupe_flag}
wac <- flag_dupes(wac, everything())
sum(wac$dupe_flag)
```

```{r dupe_view}
wac %>% 
  filter(dupe_flag) %>% 
  select(customer_name, vendor_name, calendar_year)
```

### Categorical

```{r distinct_each}
col_stats(wac, n_distinct)
```

```{r distinct_cound}
wac %>% 
  select(12:15) %>% 
  map(~mutate(count(data.frame(x = .x), x), p = n/sum(n)))
```

### Amounts

The amount paid to each vendor for a contract is broken up into fiscal quarters.
We will sum the quarters to find the total annual spending.

```{r amouunt_sum}
wac <- wac %>% 
  rowwise() %>% 
  mutate(amount = sum(c_across(ends_with("sales_reported"))))
```

```{r ammount_summary}
summary(wac$amount)
percent(mean(wac$amount <= 0), 0.01)
```

```{r hist_amount, echo=FALSE}
wac %>%
  ggplot(aes(amount)) +
  geom_histogram(fill = dark2["purple"]) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(
    breaks = c(1 %o% 10^(0:6)),
    labels = dollar,
    trans = "log10"
  ) +
  labs(
    title = "Washington Contracts Amount Distribution",
    subtitle = "from 2000 to 2019",
    caption = "Source: WA DES",
    x = "Amount",
    y = "Count"
  )
```

### Dates

The data goes back to `r min(wac$calendar_year)` and the number of contracts is
fairly consistent year to year.

```{r bar_year, echo=FALSE}
wac %>% 
  count(calendar_year) %>% 
  ggplot(aes(x = calendar_year, y = n)) +
  geom_col(fill = dark2["orange"]) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous() +
  theme(legend.position = "bottom") +
  labs(
    title = "Washington Contracts by Year",
    caption = "Source: WA DES",
    fill = "Election Year",
    x = "Year Made",
    y = "Count"
  )
```

## Conclude

```{r clean_glimpse}
glimpse(wac)
```

1. There are `r comma(nrow(wac))` records in the database.
1. There are `r comma(sum(wac$dupe_flag))` duplicate records in the database.
1. The range and distribution of `amount` and `date` seem reasonable.
1. There are `r comma(sum(wac$na_flag))` records missing a name.
1. There is no geographic data to normalize.
1. The 4-digit `calendar_year` variable already exists.

## Export

```{r clean_dir}
clean_dir <- dir_create(here("wa", "contracts", "data", "clean"))
clean_path <- path(clean_dir, "wa_Contracts_clean.csv")
write_csv(wac, clean_path, na = "")
file_size(clean_path)
guess_encoding(clean_path)
```
