---
title: "Data Diary"
subtitle: "Wisconsin Lobbying"
author: "Yanqi Xu"
date: "`r format(Sys.time())`"
output:
   github_document:    
    df_print: tibble
    # fig_caption: yes
    # highlight: tango
    # keep_md: yes
    # max.print: 32
    # toc: yes
    # toc_float: no
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  warning = FALSE
)
options(width = 99)
```

## Project


The Accountability Project is an effort to cut across data silos and give journalists, policy
professionals, activists, and the public at large a simple way to search across huge volumes of
public data about people and organizations.


Our goal is to standardizing public data on a few key fields by thinking of each dataset row as a
transaction. For each transaction there should be (at least) 3 variables:


1. All **parties** to a transaction
2. The **date** of the transaction
3. The **amount** of money involved


## Objectives


This document describes the process used to complete the following objectives:


1. How many records are in the database?
1. Check for duplicates
1. Check ranges
1. Is there anything blank or missing?
1. Check for consistency issues
1. Create a five-digit ZIP Code called `ZIP5`
1. Create a `YEAR` field from the transaction date
1. Make sure there is data on both parties to a transaction


## Packages

The following packages are needed to collect, manipulate, visualize, analyze, and communicate
these results. The `pacman` package will facilitate their installation and attachment.

```{r p_load, message=FALSE, dfrning=FALSE, error=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("irworkshop/campfin")
pacman::p_load(
  stringdist, # levenshtein value
  tidyverse, # data manipulation
  lubridate, # datetime strings
  tidytext, # string analysis
  magrittr, # pipe opperators
  janitor, # dataframe clean
  refinr, # cluster and merge
  knitr, # knit documents
  glue, # combine strings
  scales, #format strings
  here, # relative storage
  fs, # search storage 
  vroom, #read deliminated files
  readxl #read excel files
)
```

```{r fix_fun, echo=FALSE, collapse = TRUE}
# fix conflict
here <- here::here
print_all <- function(df) df %>% print(n = nrow(.)) 
```
This document should be run as part of the `R_campfin` project, which lives as a sub-directory of
the more general, language-agnostic [`irworkshop/accountability_datacleaning`][01] GitHub
repository.


The `R_campfin` project uses the [RStudio projects][02] feature and should be run as such. The
project also uses the dynamic `here::here()` tool for file paths relative to _your_ machine.


```{r where_here, collapse=TRUE}
# where dfs this document knit?
here::here()
```

## Download
Set the download directory first.
```{r raw_dir}
# create a directory for the raw data
raw_dir <- here("wi", "lobbyists", "data", "raw")
principal <- here(raw_dir,"principal")
lobbyists <- here(raw_dir, "lobbyists")

dir_create(raw_dir,principal)
```
The [Louisiana Board of Ethics] [03] 

> Lobbyists and principals submit three kinds of reports.

> Principals must file a report within 15 days of lobbying on a specific legislative proposal, budget bill subject, administrative rule, or topic. These reports appear on the website under each principal’s “Lobbying Efforts” tab.  
Lobbyists are required to submit a Lobbyist Time Report every six months for each principal that they represent. This report contains only the amount of time that was spent by the lobbyist on lobbying in that six month period. The report must be filed even if the lobbyist has zero hours to report.  
Principals are required to complete a Statement of Lobbying Activities and Expenditures (SLAE) every six months. The SLAE contains the Lobbyist Time Report for each lobbyist the principal has authorized, and the amount of money the principal spent on lobbying. Registered principals are required to file the SLAE even if no activity has occurred.

[03]: https://www.legis.iowa.gov/lobbyist/reports/searchLobby?type=lobbyist
[05]: http://ethics.la.gov/Pub/Laws/Title24LegislativeLobbying.pdf

```{r lobbyists, eval = FALSE}

wi_url <- glue("https://lobbying.wi.gov/Reports/Report.aspx?ReportPath=/GAB/Lobbying/Public&ReportName=DirectoryOfLicensedLobbyistsFlat&SessionID={(2005:2020)[c(TRUE,FALSE)]}REG&SessionStartYear={(2005:2020)[c(TRUE,FALSE)]}&SessionEndYear={(2005:2020)[c(FALSE,TRUE)]}&OutputFormat=Excel")

glue(https://lobbying.wi.gov/Reports/Report.aspx?ReportPath=/GAB/Lobbying/Public&ReportName=DirectoryOfLicensedLobbyistsFlat&SessionID={c()})

if (!all_files_new(raw_dir)) {
    download.file(
      url = wi_url,
      destfile = glue("{raw_dir}/wi_lobbyist_{(2005:2020)[c(TRUE,FALSE)]}.xls"))
}

```

Then, we'll merge each dataset into a master dataset. Note that there is no date or year field in the individual databases, and we will need to create such fields in the master file retaining the legislative period information. 
[La. Stat. Ann. § 24:53.][05] regulates lobbyists registration as such.
> Each lobbyist shall register with the board as soon as possible after employment as a lobbyist or after
the first action requiring his registration as a lobbyist, whichever occurs first, and in any event not later than
five days after employment as a lobbyist or not later than five days after the first action requiring his
registration as a lobbyist, whichever occurs first. 


## Reading
The excel files contain some empty rows or text in some cells that are not part of the dataset itself. There are usually 3 top rows with table title and metadata, but the invalid rows at the bottom vary in lengths. We'll get rid of these columns by skipping the first three rows and requiring at least 3 non-na cells, `name`, `organization` and `licensed date` to be a valid registration entry. 
```{r read csv}
wi_lobby_files <- list.files(raw_dir, pattern = ".xls", recursive = TRUE, full.names = TRUE)
# wi_lobby_colnames <- c("lobbyist_name", "licensed_date", "title", "organization_name", "address", "city_state_zip", "phone_1", "phone_2", "fax", "email", "license_type", "surrendered_date")
wi_lobby <- wi_lobby_files %>% map(read_xls, skip = 3) %>% 
                                   
                                   # col_types = cols(.default = col_character(),
                                   #                  licensed_date = col_date(format = "%m/%d/%Y"),
                                   #                  surrendered_date = col_date(format = "%m/%d/%Y"))) %>% 
                   bind_rows() %>% 
                   clean_names()


wi_lobby <- wi_lobby %>%filter(rowSums(!is.na(wi_lobby))>=3)

wi_lobby$licensed_date <-  wi_lobby$licensed_date %>% as.Date(format = "%m/%d/%Y")
wi_lobby$surrendered_date <- wi_lobby$surrendered_date %>% excel_numeric_to_date(date_system = "modern")
```

## Duplicates

We'll use the `flag_dupes()` function to see if there are records identical to one another and flag the duplicates. A new variable `dupe_flag` will be created.

```{r flag dupe}
wi_lobby <- flag_dupes(wi_lobby, dplyr::everything())
```

## Missing

```{r count_na}
wi_lobby  %>% glimpse_fun(count_na)
```
Few values are missing from the lobbyists database.

## Wrangling
We'll wrangle the three datasets one by one. First, lobbyists registration directory.
Since all the lobbyists were jumbled together, we turn the lobbyists column into a vector, and then use `unnest_longer()` to unnest all the elements in the vector to keep each in individual row. 
```{r name}
wi_lobby <- wi_lobby %>% 
  mutate(lobbyist_name = str_to_upper(lobbyist_name))
```

### Phone
```{r normal phone}
wi_lobby <- wi_lobby %>% mutate(phone_1_norm = normal_phone(phone_1))
wi_lobby <- wi_lobby %>% mutate(phone_2_norm = normal_phone(phone_2))
```


### Year 
```{r separate active start/end year}

```

### CityStateZip
Separate the address, city, state and zip columns.
```{r separate columns}
wi_lobby <- wi_lobby %>% 
    separate(city_state_zip, ", ", into = c("city_raw", "state_zip"), remove = FALSE) %>% 
    add_column(state = NA, .after = "city_raw") %>% 
    add_column(zip = NA, .after = "state") %>% 
    mutate(
         state = str_match(city_state_zip, ",\\s([A-Z]{2})\\s")[,2],
         zip = str_remove(state_zip, state) %>% normal_zip()) %>% 
    select(-state_zip)
```
### Address
```{r normal address, echo=FALSE}
wi_lobby <- wi_lobby %>% 
  mutate(lobbyist_address = normal_address(address = address,
      add_abbs = usps_city,
      na_rep = TRUE))
```

### ZIP 
Running the following commands tells us the zipcode fields are mostly clean.
```{r client normal zip}
prop_in(wi_lobby$zip, valid_zip, na.rm = TRUE) %>% percent()
```

### State
Running the following commands tells us the state fields are mostly clean.
```{r clients clean state}
prop_in(wi_lobby$state, valid_state, na.rm = TRUE) %>% percent()
```
### City
The city field needs some work. We'll go to our 
```{r}
wi_lobby$city_raw <- wi_lobby$city_raw %>% str_to_upper()

```

#### Prep
```{r prep_city, collapse = TRUE}
wi_lobby <- wi_lobby %>% mutate(city_norm = normal_city(city = city_raw,
                                            geo_abbs = usps_city,
                                            st_abbs = c(valid_state),
                                            na = invalid_city,
                                            na_rep = TRUE))
n_distinct(wi_lobby$city_raw)
n_distinct(wi_lobby$city_norm)

prop_in(wi_lobby$city_raw, valid_city, na.rm = TRUE)
prop_in(wi_lobby$city_norm, valid_city, na.rm = TRUE)
```

#### Swap
Then, we will compare these normalized `city_norm` values to the _expected_ city value for that
vendor's ZIP code. If the [levenshtein distance][09] is less than 3, we can confidently swap these
two values.

[09]: https://en.wikipedia.org/wiki/Levenshtein_distance

```{r swap_city}
wi_lobby <- wi_lobby %>% 
  left_join(
    y = zipcodes,
    by = c(
      "state" = "state",
      "zip" = "zip"
    )
  ) %>% 
  rename(city_match = city) %>% 
  mutate(
    match_dist = stringdist(city_norm, city_match),
city_swap = if_else(condition = is.na(city_match) == FALSE,
                        if_else(
      condition = match_dist <= 2,
      true = city_match,
      false = city_norm
    ),
      false = city_norm
  ))

prop_in(wi_lobby$city_swap, valid_city, na.rm = TRUE)
```

This is a very fast way to increase the valid proportion to
`r percent(prop_in(wi_lobby$city_swap, valid_city, na.rm = TRUE))` and reduce the number of distinct
_invalid_ values from `r length(setdiff(wi_lobby$city_norm, valid_city))` to only
`r length(setdiff(wi_lobby$city_swap, valid_city))`

## Export

```{r writeean}
clean_dir <- here("wi", "lobbying", "data", "processed")
dir_create(clean_dir)
wi_lobby %>% 
  write_csv(
    path = glue("{clean_dir}/wi_lobbyists.csv"),
    na = ""
  )
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

